### 2020-03-12 Notes/BoundEffRed.tex
notes

\documentclass[twoside,letterpaper]{article}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{euscript}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{ifpdf}
\usepackage[english]{babel}
\usepackage{caption}
\usepackage{subfig} % or can use subcaption package
\usepackage{color}
\usepackage{graphicx}
\usepackage{mathrsfs}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{dsfont}
\usepackage[shortlabels]{enumitem}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{hyperref}
\usepackage{mathpazo}
\usepackage[left=80pt,right=80pt,top=70pt,bottom=70pt]{geometry}

\title{Tackling boundary effects on kernel-based representations (draft)}
%\author{}
\date{}

\input{mydefs}

\newtheorem{remark}{Remark}

\begin{document}

\maketitle

\section{Introduction}
In any digital acquisition system, the measured signals are sampled and of finite length. Then, when performing any kernel-based transform, such as synchrosqueezing transform, boundaries effect is going to affect the quality of the representation in the vicinity of the borders. Consequently, signal characteristics, like instantaneous frequencies, appear to be imprecisely determined near borders.

We propose here a method to limit the disruptions caused by the boundary effect. To reduce the influence of boundaries on the kernel-based representations, we proceed in two steps:
\begin{enumerate}
\item Extend the signal by forecasting it.
\item Run the kernel-based algorithm on the extended signal.
\end{enumerate}

\section{Algorithm}
\input{algorithm}

\section{Theoretical performance}
\input{theory}

%\section{Numerical results}
%\input{results}

\end{document}

### 2020-03-12 Notes/algorithm.tex
notes

\subsection{Forecasting}

\paragraph{Notations.} Let $x:\RR\to\RR$ denote a continuous-time signal. In this work, we consider a finite-length discretization of that one. Thus, the sampled signal $\bx$, whose length is denoted by $N$, is such that
\[
\bx[n] = x\left(\frac{n}{\fs}\right)\ ,\quad \forall n\in\{0,\ldots,N-1\}\ , 
\]
where $\fs$ denotes the sampling frequency.

Let $M<N$, $K<N$. Then, for all $k\in\{0,\ldots,K-1\}$, we extract from $\bx\in\RR^N$ the sub-signal $\bx_k\in\RR^M$ given by:
\[
\bx_k = 
\begin{pmatrix}
\bx[N-K+(k-1)-(M-1)] \\
\vdots \\
\bx[N-K+(k-1)]
\end{pmatrix}\ .
\]  
These sub-signals are gathered into the matrix $\bX\in\RR^{M\times K}$ such that:
\[
\bX = 
\begin{pmatrix}
\bx_0 & \cdots & \bx_{K-1}
\end{pmatrix}\ .
\]
Notice that these sub-signals are overlapping each other. Indeed, $\bx_{k+1}$ is a shifting of $\bx_k$ from one sample.

We also consider the matrix $\bY\in\RR^{M\times K}$ given by:
\[
\bY = 
\begin{pmatrix}
\bx_1 & \cdots & \bx_{K}
\end{pmatrix}\ .
\]

\paragraph{Dynamic model and forecasting.} Establishing a dynamic model consists in determining the relation linking $\bY$ to $\bX$, that is finding a function $f$ so that
\[
\bY = f(\bX)\ .
\] 
We consider here a naive dynamic model, assuming that we have the following relation:
\begin{equation}
\bY = \bA\bX\ ,
\end{equation}
where $\bA\in\RR^{M\times M}$. This is a linear dynamic model, which can be written equivalently in function of the sub-signals $\bx_k$, as:
\begin{equation}
\bx_{k+1} = \bA\bx_{k},\ ,\forall k\in\{0,\dots,K-1\}\ .
\end{equation}

The forecasting method consists in estimating the unknown matrix $\bA$. Indeed, let $\tilde\bA$ denotes the estimate of $\bA$, we then obtain the forecasting of the signal at time $\frac{N-1+\ell}{\fs}$ by:
\begin{equation}
\label{eq:prediction}
\tilde \bx [N-1+\ell] = \be_M^T\tilde\bA^\ell\bx_{K}\ ,
\end{equation}  
where $\be_M$ is the vector of length $M$ given by $\be_M = \begin{pmatrix} 0 & \cdots & 0 & 1\end{pmatrix}^T$.

\paragraph{Model estimation.} To estimate the matrix $\bA$, we basically implement the least square estimator. Thus, we solve the following problem:
\begin{equation}
\label{eq:ls.pb}
\tilde\bA = \arg\min_{\balpha} \cL(\bA)\ ,
\end{equation}
where the loss function $\cL$ is given by:
\[
\cL(\bA) = \|\bY-\bA\bX\|^2 = \sum_{k=0}^{K-1} \|\bx_{k+1}-\bA\bx_k\|^2.
\]
Therefore, solving the problem~\eqref{eq:ls.pb}, \ie~$\nabla \cL(\tilde\bA)=\bzero$, gives the following estimate $\tilde\bA$ of the dynamic model matrix $\bA$:
\begin{align}
\tilde\bA &= \bY\bX^T(\bX\bX^T)^{-1}\ .
\label{eq:lse}
\end{align}

\begin{remark}
This expression clearly shows that the matrix $\tilde\bA$ takes the following form:
\[
\tilde\bA =
\begin{pmatrix}
0       & 1       & 0      & \cdots & 0      \\
\vdots  & \ddots  & \ddots & \ddots & \vdots  \\
\vdots  &         & \ddots & \ddots & 0  \\
0       & \cdots  & \cdots & 0      & 1  \\
\alpha_1& \cdots  & \cdots & \cdots & \alpha_M  \\
\end{pmatrix}.
\]
Then, except the row vector $\balpha = \left(\alpha_1 \cdots\alpha_M\right)$, the matrix $\bA$ is fully determined by the dynamic model.
\end{remark}

\paragraph{Signal extension.} We can finally construct the extended signal $\tilde\bx\in\RR^{N+2L}$ concatenating the backward prediction $\tilde\bx_{\rm bw}$, the observed signal $\bx$, and the forward prediction $\tilde\bx_{\rm fw}$. We summarize the extension step in Algorithm~\ref{alg:extension}. Notice that we handle the backward estimation using the same strategy than described above, but applying it to the reverse signal $\bx^{\rm r} = \begin{pmatrix} \bx[N-1] & \cdots & \bx[0] \end{pmatrix}^T$.

\begin{algorithm}
\caption{Signal extension. $\tilde\bx = \mathsf{SigExt}(\bx,M,K,L)$}
\label{alg:extension}
\begin{algorithmic}
\STATE {\bf Inputs}: $\bx$, $M$, $K$, $L$
\STATE 
\STATE {\bf Forward forecasting.}
\STATE \quad\textbullet\ LS estimation of the forward matrix $\tilde\bA_{\rm fw}$ via equation~\eqref{eq:lse}.
\STATE \quad\textbullet\ Forward forecasting $\tilde\bx_{\rm bw}$ obtained applying equation~\eqref{eq:prediction} with $\ell\in\{1,\ldots,L\}$.
\STATE 
\STATE {\bf Backward forecasting.}
\STATE \quad\textbullet\ Reverse signal $\bx$ to $\bx^{\rm r}$. 
\STATE \quad\textbullet\ LS estimation of the backward matrix $\tilde\bA_{\rm bw}$ via equation~\eqref{eq:lse} applied to $\bx^{\rm r}$.
\STATE \quad\textbullet\ Reversed backward forecasting $\tilde\bx_{\rm bw}^{\rm r}$ obtained applying equation~\eqref{eq:prediction} to $\bx^r$ with $\ell\in\{1,\ldots,L\}$.
\STATE \quad\textbullet\ Reverse $\tilde\bx_{\rm bw}^{\rm r}$ to obtain the estimate $\tilde\bx_{\rm bw}$.
\STATE 
\STATE {\bf Output}: Extended signal $\tilde\bx = \begin{pmatrix} \tilde\bx_{\rm bw}  & \bx & \tilde\bx_{\rm fw}\end{pmatrix}^T$.
\end{algorithmic}
\end{algorithm}

\subsection{Kernel-based representation}

\paragraph{Transform restriction.} 
Let $\ccF_{N}:\RR^{N}\to\RR^{F\times N}$ generically denotes the kernel-based representation we are interested in. It can be, for instance, such as short-time Fourier transform (STFT), the continuous wavelet transform (CWT), the synchrosqueezing transform (SST), or the reassignment (RS). Here, $F$ typically denotes the size of the representation along the frequency axis. Due to the boundary effects, the representation $\ccF_{N}(\bx)$ shows undesired patterns when approaching its edges.For example, the instantaneous frequencies highlighted by the SST can be blurred near that edges. To limit these phenomena, we apply the representation to the estimated extended signal $\tilde\bx$. This strategy moves the boundary effects out of the time interval $I=[0, \frac{N-1}{\fs}]$. Finally, the boundary-effects insensitive representation $\ccF_N^{\mathrm{ext}}:\RR^{N}\to\RR^{F\times N}$ of $\bx$ is given by:
\begin{equation}
\ccF_N^{\mathrm{ext}}(\bx)[f,n] = \ccF_{N+2L}(\tilde \bx)[f,L+n]\ ,\quad\forall f\in\{0,\cdots,F-1\},\ n\in\{0,\cdots,N-1\}\ .
\label{eq:restriction}
\end{equation}
This amounts to restricting the representation $\ccF_{N+2L}(\tilde \bx)$ to the original measurement interval of $\bx$. For the sake of simplicity, we denotes the restriction operator by $\cR$, where $\cR:\RR^{F\times (N+2L)}\to\RR^{F\times N}$. Consequently, we have:
\begin{equation*}
\ccF_N^{\mathrm{ext}}(\bx) = \cR\left( \ccF_{N+2L}(\tilde \bx) \right) \ .
\end{equation*}

\paragraph{Conclusion.}
Finally, the global procedure we implement to reduce boundary effects on kernel-based representations is summarized by the pseudo-code of Algorithm~\ref{alg:boundary}.

\begin{algorithm}
\caption{Tackling boundary effects. $\bF_\bx = \mathsf{BoundEffRed(}\bx,M,K,L,\ccF)$}
\label{alg:boundary}
\begin{algorithmic}
\STATE {\bf Inputs}: $\bx$, $M$, $K$, $L$, $\ccF$
\STATE 
\STATE {\bf Forecasting step.}
\STATE \quad\textbullet\ Signal extension: $\tilde\bx = \mathrm{SigExt}(\bx)$.
\STATE 
\STATE {\bf Representation step.}
\STATE \quad\textbullet\ Representation evaluation: $\ccF_{N+2L}(\tilde\bx)$.
\STATE \quad\textbullet\ Restriction of $\ccF_{N+2L}(\tilde\bx)$ to the central time interval (see~\eqref{eq:restriction})to obtain $\bF_\bx=\ccF_N^{\mathrm{ext}}(\bx)$.
\STATE 
\STATE {\bf Output}: Signal representation $\bF_\bx$
\end{algorithmic}
\end{algorithm}

### 2020-03-12 Notes/results.tex
notes

\subsection{Evaluation the forecasting performance}
In that section, we first evaluate the quality of the forecasting step. That one depends on at least two conditions:
\begin{itemize}
\item The validity of the dynamic model we use to implement forecasting.
\item An optimal choice of the set of hyperparameters on which our algorithm depends. 
\end{itemize}
We study that two points in the following.

\subsubsection{Validation of the linear dynamic model}
\label{ssse:validation}
We proved that the linear dynamic model is sufficient to describe the dynamical behavior of signals as long as the adaptive harmonic model is relevant to describe those signals. We confirm here this theoretical result on real-life signals. To this end, we apply Algorithm~\ref{alg:extension} to a larger number of pieces of a respiratory signal. The signal is 50 minutes-long and is sampled at $\fs=100$~Hz. A zoom on a small portion of the signal is displayed in Figure~\ref{fig:tho}. From that large signal, we build a dataset of $48$ non-overlapping signals of 60 seconds, \ie~$N=6000$.

\begin{figure}
\includegraphics[width=\textwidth]{THOsig.eps}
\caption{Zoom on the respiratory signal.}
\label{fig:tho}
\end{figure}

The forecasting Algorithm~\ref{alg:extension} is applied to each of the signals in the dataset we built. These ones are extended of 7 second on each border, corresponding to $L =700$. Moreover, we choose the following set of hyperparameters: $M=\lfloor 1.5L\rfloor$ and $K=\lfloor2.5M\rfloor$. We justify such a choice in section~\ref{ssse:hyperparam}. To quantify the quality of our forecasting approach, we evaluate the Mean Square Error $\mathrm{MSE}(\tilde\bx)$ of the estimated extended signal, namely:
\[
\mathrm{MSE}(\tilde\bx) = \dfrac1{2L}\|\tilde\bx_{\rm bw} -\bx_{\rm bw}\|^2 + \dfrac1{2L}\|\tilde\bx_{\rm fw} -\bx_{\rm fw}\|^2\ .
\]
The mean and the standard deviation of the MSE with respect to the simulations are given in Table~\ref{tab:mse}.

The results are compared with the strategy consisting in a zero-padding extension of the signal. The results neatly shows on improvement of the quality of the forecasting. Indeed, in average, the forecasting MSE is improved of almost $20\%$. Nevertheless, the variance forecasting MSE shows a huge variability in the quality of the estimation. This is probably due to the outliers and pulse that can be find in the respiratory signal. That ones make the adaptive harmonic model temporary irrelevant, and break the validity of the linear dynamic model.

\begin{table}
\centering
\caption{Averaged MSE.}
\begin{tabular}{|c||c|c|}
  \hline
   \multirow{2}{*}{Algorithm} & \multicolumn{2}{c|}{MSE} \\
   \cline{2-3}
      & Mean & Standard deviation\\
   \hhline{|=#=|=|}
   {\sf SigExt} & $0.7935$ & $0.2615$ \\
   \hline
   Zero padding & $0.9903$ & $0.0579$ \\
   \hline
\end{tabular}
\label{tab:mse}
\end{table}

\subsubsection{Influence of the hyperparameters}
\label{ssse:hyperparam}
We evaluate the influence of the hyperparameters $K$ and $M$ on the forecasting MSE. This study is performed on a 32 second-long photoplethysmogram (PPG) signal, which is displayed on the top of Figure~\ref{fig:ppg}. This signal is sampled at $\fs=125$~Hz.

\begin{figure}
\includegraphics[width=\textwidth]{PPGsig.eps}
\includegraphics[width=\textwidth]{PPGforecast.eps}
\caption{PPG signal. Top: original measured signal. Extended signal obtained by forecasting superimposed with the ground truth signal.}
\label{fig:ppg}
\end{figure}

In this study, the extension length $L$ is fixed because the second step of the Algorithm~\ref{alg:boundary} determines the length $L$. Indeed, in order to avoid boundary effect, this one should be at least equal to the half of the width of the window used to perform the kernel-based transform. In practice, we choose to extend the signal if 2~seconds on each side, \ie~$L=250$.
\begin{enumerate}[a),leftmargin=*]
\item 
We perform the forecasting Algorithm~\ref{alg:extension} for different values of the sub-signals length $M$. Meanwhile, the value of $K$ is unchanged and set to $K=10L$. The graph on the left in Figure~\ref{fig:influence.M} shows the influence of $M$ on the forecasting MSE. 
%When $M$ is smaller than $L$, the forecasting algorithm cannot provide accurate forecasting because . As soon as $M$ is greater than $L$, the MSE will not be influenced by the sub-signal lengths. Indeed, the information provided by
\item
We perform the forecasting Algorithm~\ref{alg:extension} for different values of the sub-signals dataset size $K$, when the value of $M$ is set to $M=\lfloor 1.5L\rfloor$. The graph on the left in Figure~\ref{fig:influence.M} shows the influence of $K$ on the forecasting MSE.
\end{enumerate}

\begin{figure}
\includegraphics[width=.5\textwidth]{mseM.eps}
\includegraphics[width=.5\textwidth]{mseK.eps}
\caption{Influence of the size parameters on the forecasting MSE. Left: MSE in function of the sub-signals length $M$. Right: MSE in function of the sub-signal dataset size $K$.}
\label{fig:influence.M}
\end{figure}

\paragraph{Influence of noise.}
The same kind of study is done. The results are displayed on Figure~\ref{fig:influence.noise}. Here, $M=\lfloor 1.5L\rfloor$ and $K=2M$.

\begin{figure}
\centering
\includegraphics[width=.7\textwidth]{mseNoise.eps}
\caption{Influence of the noise variance on the forecasting MSE.}
\label{fig:influence.noise}
\end{figure}

\subsection{Evaluation of the quality of the boundary effect reduction}
The quality of the boundary effect reduction step must be evaluated of the synchrosqueezing representation. To that aim, we compare the obtained synchrosqueezing transform to the optimal synchrosqueezing transform $\ccF_N^\mathrm{opt}(\bx)$. The optimal synchrosqueezing transform is defined as the restriction of the synchrosqueezing of the ground truth extended signal $\bx^\mathrm{ext} = \begin{pmatrix}\bx[-L] & \cdots & \bx[N-1+L] \end{pmatrix}$. Therefore:
\begin{equation*}
\ccF_N^\mathrm{opt}(\bx) = \cR\left( \ccF_{N+2L}(\bx^\mathrm{ext}) \right) \ .
\end{equation*}

The different representations we obtained must be compared to the optimal SST. To that aim, we establish a criterion that quantify the distance of these time-frequency representations to the optimal one. To do this, we reason in analogy with the optimal transport distance which allows us to quantify the distance between two probability density functions. Let us generically denote a time frequency representation $\ccQ$. Then, for $t$ fixed, we consider the following pseudo probability density function:
\begin{align}
p_\ccQ^t(\xi) &= \dfrac{|\ccQ(\xi,t)|^2}{\int_\RR |\ccQ(\nu,t)|^2\dd\nu}\ .
\label{eq:ppdf}
\end{align}
Similarly, let $p_{\ccF_0}^t$ denotes the associated pseudo-probability density function (defined as in~\eqref{eq:ppdf}) associated with the reference time-frequency representation $\ccF_0$. At each instant $t$, we can then determine the optimal transport distance $d_{t}$ between the two pseudo densities. It is given by the $L^1$ norm of the difference between the associated distribution functions. In other words, either $P_{\ccQ}^t(\xi)=\int_{-\infty}^\xi p_{\ccQ}^t(\nu)\dd\nu$ and $\tilde P_{\ccF_0}^t(\xi)=\int_{-\infty}^\xi\tilde p_{\ccF_0}^t(\nu)\dd\nu$, we have
\begin{equation*}
d_{t}(\ccQ,\ccF_0) = \int_\RR\left|\tilde P_{\ccQ}^t(\xi)-  P_{\ccF_0}^t(\xi)\right|\dd\xi\ .
\end{equation*}
Finally, the distance between the two time-frequency representations is obtained by averaging all the optimal transport distances with respect to time:
\begin{equation}
D(\ccQ,\ccF_0) = 100\times\frac1{|I|}\int_I d_{t}(\ccQ,\ccF_0)\dd t\ .
\end{equation}
The \textit{Optimal Transport Distance} (OTD) quantifies the proximity between the estimated and actual instantaneous frequencies while favouring the sparsity of the estimated time-frequency representation.

This distance is evaluated on the dataset of respiratory signals introduced in section~\ref{ssse:validation}. The results are summarize in Table~\ref{tab:otd}. The OTD averaged over the simulations clearly shows that $\ccF_\mathrm{FC}$ outperforms $\ccF$. This highlights the ability of our approach to limit the distortion due the boundary effect and provide a more accurate synchrosqueezing transform. Moreover, the OTD decreases by almost $50\%$ from the standard SST $\ccF$ to the improved SST $\ccF_{\mathrm{FC}}$. Thus, the improvement obtained by forecasting in the time domain is amplified in the time-frequency domain.

\begin{table}
\centering
\caption{Performance of the Boundary effect reduction on the SST.}
\begin{tabular}{|c||c|c|}
  \hline
   \multirow{2}{*}{Algorithm} & \multicolumn{2}{c|}{OTD to $\ccF_N^{\mathrm{opt}}(\bx)$} \\
   \cline{2-3}
      & Mean & Standard deviation\\
   \hhline{|=#=|=|}
   $\ccF_N^{\mathrm{ext}}(\bx)$ & $4.705\times 10^{-3}$ & $3.147\times 10^{-3}$ \\
   \hline
   $\ccF_N(\bx)$ & $9.195\times 10^{-3}$ & $3.453\times 10^{-3}$ \\
   \hline
\end{tabular}
\label{tab:otd}
\end{table}

### 2020-03-12 Notes/algorithm.tex
notes

@@
 \]
-Notice that these sub-signals are overlapping each other. Indeed, $\bx_{k+1}$ is a shifting of $\bx_k$ from one sample.
+%Notice that these sub-signals are overlapping each other. Indeed, $\bx_{k+1}$ is a shifting of $\bx_k$ from one sample.
 
@@
 Let $\ccF_{N}:\RR^{N}\to\RR^{F\times N}$ generically denotes the kernel-based representation we are interested in.
-It can be, for instance, such as short-time Fourier transform (STFT), the continuous wavelet transform (CWT), the synchrosqueezing transform (SST), or the reassignment (RS).
+It can be, for instance, the short-time Fourier transform (STFT), the continuous wavelet transform (CWT), the synchrosqueezing transform (SST), or the reassignment (RS).
 Here, $F$ typically denotes the size of the representation along the frequency axis.
-Due to the boundary effects, the representation $\ccF_{N}(\bx)$ shows undesired patterns when approaching its edges.For example, the instantaneous frequencies highlighted by the SST can be blurred near that edges.
+Due to the boundary effects, the representation $\ccF_{N}(\bx)$ shows undesired patterns when approaching its edges.
+For example, the instantaneous frequencies highlighted by the SST can be blurred near that edges.
 To limit these phenomena, we apply the representation to the estimated extended signal $\tilde\bx$.

### 2020-03-19 Notes/algorithm.tex
update notes

--- 2020-03-12 Notes/algorithm.tex
@@
 \STATE \quad\textbullet\ Representation evaluation: $\ccF_{N+2L}(\tilde\bx)$.
-\STATE \quad\textbullet\ Restriction of $\ccF_{N+2L}(\tilde\bx)$ to the central time interval (see~\eqref{eq:restriction})to obtain $\bF_\bx=\ccF_N^{\mathrm{ext}}(\bx)$.
+\STATE \quad\textbullet\ Restriction of $\ccF_{N+2L}(\tilde\bx)$ to the central time interval (see~\eqref{eq:restriction}) to obtain $\bF_\bx=\ccF_N^{\mathrm{ext}}(\bx)$.
 \STATE 

### 2020-04-14 Notes/BoundEffRed.tex
notes
new version notes

@@
 \usepackage{mathpazo}
-\usepackage[left=80pt,right=80pt,top=70pt,bottom=70pt]{geometry}
+\usepackage[left=90pt,right=90pt,top=70pt,bottom=70pt]{geometry}
 
-\title{Tackling boundary effects on kernel-based representations (draft)}
+\title{An approach for the reduction of boundary effects in time-frequency and time-scale representations}
 %\author{}
@@
 \section{Introduction}
-In any digital acquisition system, the measured signals are sampled and of finite length.
-Then, when performing any kernel-based transform, such as synchrosqueezing transform, boundaries effect is going to affect the quality of the representation in the vicinity of the borders.
-Consequently, signal characteristics, like instantaneous frequencies, appear to be imprecisely determined near borders.
+In any digital acquisition system, the study and the interpretation of the measured signals generally require the use an analysis tool, which enable to point out the useful characteristics of the signal.
+For example, any  measured physiological signal, such as photoplethysmogram (PPG), may not be interpreted as it is, from its run chart.
+An analysis tool would make possible the extraction of some useful characteristics such as heart rate or blood pressure.
 
-We propose here a method to limit the disruptions caused by the boundary effect. To reduce the influence of boundaries on the kernel-based representations, we proceed in two steps:
+In general, the observed quantities are produced by transient phenomena, that can vary rapidly and irregularly.
+As a consequence, the measured signals exhibit nonstationary behavior.
+In order to adapt the analysis to nonstationarities, local analysis is generally performed.
+The short time Fourier transform (STFT) is a typical tool which is build that way, and allows to determine locally the frequency content of a nonstationary signal.
+
+Windowing is a common method for performing local analysis.
+Among others, STFT, wavelet transform~\cite{Mallat99wavelet}, synchrosqueezing transform (SST), and reassignment~\cite{Auger13time} are representations that fall back on the use of an analysis window.
+Let $x:I\to\RR$ denote the observed signal, where $I$ denotes the finite interval where the signal is measured.
+Let $g_s:\RR\to\RR$ denote the analysis window, where $s$ is a shape parameter.
+The support of $g_s$ is is localized around the origin, and is small with respect to $|I|$.
+The translation operator is $T_\tau$ defined as:
+\[
+T_\tau f = f(t-\tau)\ ,\quad \forall\ f:\RR\to\RR\ .
+\]
+Then, the local analysis of $x$ around the instant $\tau\in I$ rely on the evaluation of the following dot product:
+\begin{equation}
+V_x(s,\tau) = \langle x, T_\tau g_s \rangle_I \ .
+\label{eq:windowing}
+\end{equation}
+A major shortcoming of this technique occurs when analyzing the signal $x$ near the boundaries of the interval $I$.
+Indeed, at these points, half of the information is missing.
+Consequently the results of the dot product~\ref{eq:windowing} is disturbed by this lack of information.
+This phenomenon is the so-called \emph{boundary effect}.
+We display, on the left of Fig.~\ref{fig:ex.intro} the result of the SST of a PPG (see section~\ref{ssse:ppg} for a comprehensive description).
+The distortion resulting from the boundary effect is clearly visible on both sides of this representation.
+Indeed, while in the major central part of the image, clear lines stand out, they become blurred as they approach the left and right boundaries of the image.
+A zoom on the right boundary of this SST, displayed on the right of Fig.~\ref{fig:ex.intro}, emphasizes the result of boundary effect.
+More generally, signal characteristics, like instantaneous frequencies or amplitudes appear to be imprecisely determinable in the vicinity of the boundaries.
+
+\begin{figure}
+\includegraphics[width=.49\textwidth]{SSTintro.eps}
+\includegraphics[width=.49\textwidth]{zoomSSTintro.eps}
+\caption{Synchrosqueezing transform of a PPG (left) with a zoom on its right boundary (right). }
+\label{fig:ex.intro}
+\end{figure}
+
+The purpose of this paper is to provide a fast algorithm allowing to tackle boundary effect and limit the disruptions it may cause on the signals representations based on windowing.
+In that aim, we proceed in two steps:
 \begin{enumerate}
-\item Extend the signal by forecasting it.
-\item Run the kernel-based algorithm on the extended signal.
+\item \emph{Extend the signal by forecasting it.} The aim is to use a dynamic model to predict the values taken by the measured signal outside the measurement interval.
+Then, once this operation is done, we have access to an extended signal defined on a larger interval $I_\Delta$, where $\Delta$ denotes the size of the extension on both boundaries of $I$.
+\item \emph{Run the local analysis tool on the extended signal.} Assuming that the support of the analysis window is smaller that $2\Delta$, the local analysis near the boundary of $I$ is now possible without lack of information thanks to knowledge brought by the extension on both sides.
 \end{enumerate}
+Thus, assuming that the quality of the extension step is sufficient, the representation obtained that way will be less sensitive to the boundary effect than the result of the analysis tool applied directly to the non-extended signal.
+Notice that, in order to obtain a fast algorithm the extension step we propose must have a low computational cost.
 
 \section{Algorithm}
+\label{se:algo}
 \input{algorithm}
@@
 \section{Theoretical performance}
+\label{se:theoretical}
 \input{theory}
 
-%\section{Numerical results}
-%\input{results}
+\section{Numerical results}
+\input{results}
+
+\bibliographystyle{plain}
+\bibliography{AM041320}
 

### 2020-04-14 Notes/algorithm.tex
new version notes

@@
 \bY = \bA\bX\ ,
+\label{eq:dyn.model}
 \end{equation}
@@
 
-\subsection{Kernel-based representation}
+\subsection{Representation}
 
 \paragraph{Transform restriction.} 
-Let $\ccF_{N}:\RR^{N}\to\RR^{F\times N}$ generically denotes the kernel-based representation we are interested in.
+Let $\ccF_{N}:\RR^{N}\to\RR^{F\times N}$ generically denotes the time-frequency or time-scale representation we are interested in.
 It can be, for instance, such as short-time Fourier transform (STFT), the continuous wavelet transform (CWT), the synchrosqueezing transform (SST), or the reassignment (RS).
@@
 \paragraph{Conclusion.}
-Finally, the global procedure we implement to reduce boundary effects on kernel-based representations is summarized by the pseudo-code of Algorithm~\ref{alg:boundary}.
+Finally, the global procedure we implement to reduce boundary effects on windowing-based representations is summarized by the pseudo-code of Algorithm~\ref{alg:boundary}.
 

### 2020-04-14 Notes/results.tex
new version notes

@@
 \subsection{Evaluation the forecasting performance}
-In that section, we first evaluate the quality of the forecasting step. That one depends on at least two conditions:
+In that section, we first evaluate the quality of the forecasting step. That one depends on at least two parameters:
 \begin{itemize}
-\item The validity of the dynamic model we use to implement forecasting.
-\item An optimal choice of the set of hyperparameters on which our algorithm depends. 
+\item The noise variance $\sigma^2$.
+\item The size of the training dataset $K$. 
 \end{itemize}
-We study that two points in the following.
+In subsections~\ref{ssse:res.sine} and~\ref{ssse:res.ahm}, we study the influence of these parameters. A comparison with the theoretical results of section~\ref{se:theoretical} is also available.
 
-\subsubsection{Validation of the linear dynamic model}
-\label{ssse:validation}
-We proved that the linear dynamic model is sufficient to describe the dynamical behavior of signals as long as the adaptive harmonic model is relevant to describe those signals.
-We confirm here this theoretical result on real-life signals.
-To this end, we apply Algorithm~\ref{alg:extension} to a larger number of pieces of a respiratory signal.
-The signal is 50 minutes-long and is sampled at $\fs=100$~Hz.
-A zoom on a small portion of the signal is displayed in Figure~\ref{fig:tho}.
-From that large signal, we build a dataset of $48$ non-overlapping signals of 60 seconds, \ie~$N=6000$.
+\subsubsection{Sum of sine waves}
+\label{ssse:res.sine}
+We proved that the linear dynamic model is sufficient to catch the dynamical behavior of signals taking the form~\eqref{eq:sum.sine}.
+In order to validate this theoretical result, we apply the forecasting Algorithm~\ref{alg:extension} to a large number of realizations of the random vector $\bz$ such that:
+\[
+\bz = \bx+ \sigma\bw\ ,
+\]
+where
+\[
+\bx[n] =\cos\left(2\pi p_1 \dfrac{n}{M} \right) + R\cos\left(2\pi p_2 \dfrac{n}{M} \right)\ ,\quad\forall\ n\in\{1,\ldots,N\}\ ,
+\]
+and $N=10^4$, $M=750$, $p_1=10$, $p_2=33$ and $R=1.4$. Besides, the additive noise is chosen to be Gaussian: $\bw\sim\cN(\bzero,\bI)$.
 
+\paragraph{Influence of the noise variance $\sigma^2$.} Here, the size of the training dataset is set to $K=1875$.
+Then, the forecasting algorithm is run on 100 realizations on the discrete signal $\bx$ for 100 different values of $\sigma^2$  logarithmically spaced from $5.10^{-3}$ to $10^{-1}$.
+For each of these values, we determine the experimental bias $\mu_{xp}[\ell]$ and variance $\gamma_{xp}[\ell,\ell]$ in function of the forecasting sample index $\ell$ (going from 1 to $L=500$).
+These results are displayed on Figure~\ref{fig:res.noise.sine}.
+Each color corresponds to these experimental results obtained for a given value of $\sigma$.
+On the left of Figure~\ref{fig:res.noise.sine}, the result show that the bias is neither depending on the noise variance $\sigma^2$ nor the forecasting length $\ell$, and always insignificant with respect to the signal values.
+On the right of Figure~\ref{fig:res.noise.sine}, we display the normalized experimental variance, that is $\frac{\gamma[\ell,\ell]}{\sigma^2}$.
+This highligths the fact that the forecasting variance increases linearly with respect to $\sigma^2$.
+More surprisingly, this result shows that the forecasting variance decreases with $\ell$, what is counterintuitive.
 \begin{figure}
-\includegraphics[width=\textwidth]{THOsig.eps}
-\caption{Zoom on the respiratory signal.}
-\label{fig:tho}
+\includegraphics[width=.49\textwidth]{biasNoiseSine.eps}
+\includegraphics[width=.49\textwidth]{VarianceNoiseSine.eps}
+\caption{Evolution of the experimental forecasting bias (left) and the normalized forecasting variance (right) in function of the forecasting sample index.
+Each color characterizes the result for a given value of $\sigma$.}
+\label{fig:res.noise.sine}
 \end{figure}
 
-The forecasting Algorithm~\ref{alg:extension} is applied to each of the signals in the dataset we built.
-These ones are extended of 7 second on each border, corresponding to $L =700$.
-Moreover, we choose the following set of hyperparameters: $M=\lfloor 1.5L\rfloor$ and $K=\lfloor2.5M\rfloor$.
-We justify such a choice in section~\ref{ssse:hyperparam}.
-To quantify the quality of our forecasting approach, we evaluate the Mean Square Error $\mathrm{MSE}(\tilde\bx)$ of the estimated extended signal, namely:
+\paragraph{Influence of the training dataset size $K$.} Here, the noise variance $\sigma$ is set to $\sigma=10^{-2}$.
+Then, the forecasting algorithm is run on 3000 realizations on the discrete signal $\bx$ for 25 different values of $K$, logarithmically spaced from $3.5\times 10^{3}$ to $6\times 10^{3}$.
+For each of these values, we determine the experimental bias $\mu_{xp}[\ell]$ and variance $\gamma_{xp}[\ell,\ell]$ in function of the forecasting sample index $\ell$ (going from $1$ to $500$).
+These results are displayed on Figure~\ref{fig:res.size.sine}.
+Each color corresponds to these experimental results obtained for a given value of $K$.
+On the left of Figure~\ref{fig:res.size.sine}, as expected, the experimental bias vanishes.
+On the right of Figure~\ref{fig:res.size.sine}, we display the product of $K$ with the normalized experimental variance, that is $\frac{\gamma[\ell,\ell]}{\sigma^2}$.
+This result shows the asymptotic behavior of the variance.
+Indeed, as soon as $K$ is sufficiently high, the product $K\gamma[\ell,\ell]$ is approximately independent of $K$.
+\begin{figure}
+\includegraphics[width=.49\textwidth]{BiasKSine.eps}
+\includegraphics[width=.49\textwidth]{VarianceKSine.eps}
+\caption{Evolution of the experimental forecasting bias (left) and the normalized forecasting variance (right) in function of the forecasting sample index.
+Each color characterizes the result for a given value of $K$.}
+\label{fig:res.size.sine}
+\end{figure}
+
+\paragraph{Conclusion.}
+Both previous results allows us to describe the influence of the noise variance and the size of the training dataset on the variance of the forecasting noise, which is empirically summarized as follows:
+\begin{equation}
+\gamma[\ell,\ell] \sim \dfrac{\sigma^2}{K}g[\ell] \ ,\quad\mbox{when}\ \sigma^2\to 0\ \mbox{and}\ K\to\infty \ ,
+\end{equation} 
+where $g$ is a bounded positive function.
+
+\subsubsection{Adaptive harmonic model}
+\label{ssse:res.ahm}
+We now consider a signal whose instantaneous frequencies and amplitudes of its components vary over time. The toy signal we use takes the following form:
 \[
-\mathrm{MSE}(\tilde\bx) = \dfrac1{2L}\|\tilde\bx_{\rm bw} -\bx_{\rm bw}\|^2 + \dfrac1{2L}\|\tilde\bx_{\rm fw} -\bx_{\rm fw}\|^2\ .
+\bx[n] = \cos\left(2\pi \phi_1[n] \right) + R[n]\cos\left(2\pi \phi_2[n] \right)\ ,\quad\forall\ n\in\{1,\ldots,N\} \ ,
 \] 
-The mean and the standard deviation of the MSE with respect to the simulations are given in Table~\ref{tab:mse}.
+where the instantaneous amplitude $R$ is given by:
+\[
+R[n] = 1.4 + 0.2\cos\left(4\pi\frac{n}{N}\right)\ ,
+\]
+and the instantaneous phases are such that:
+\begin{align*}
+\phi_1[n] &= \frac{p_1}{M}\left( n + \frac{0.01}{2\pi}\cos\left(2\pi\frac{n}{N}\right) \right) \\
+\phi_2[n] & = p_2\frac{n}{M} + \frac{20}{2N\fs}n^2
+\end{align*}
+Besides, the additive noise is chosen to be Gaussian: $\bw\sim\cN(\bzero,\bI)$. Numerically, we take: $N=10^4$, $M=750$, $p_1=10$, $p_2=23$ and $R=1.4$.
 
-The results are compared with the strategy consisting in a zero-padding extension of the signal.
-The results neatly shows on improvement of the quality of the forecasting.
-Indeed, in average, the forecasting MSE is improved of almost $20\%$.
-Nevertheless, the variance forecasting MSE shows a huge variability in the quality of the estimation.
-This is probably due to the outliers and pulse that can be find in the respiratory signal.
-That ones make the adaptive harmonic model temporary irrelevant, and break the validity of the linear dynamic model.
+To highlight the fact that the linear dynamical model is sufficient to catch most of the dynamical behavior of signals following the AHM, we compare the performance of the Algorithm~\ref{alg:boundary} with reference forecasting algorithm that could be used for extending such signals.
+These methods are:
+\begin{itemize}
+\item The \emph{Extended Dynamic Mode Decomposition (EDMD)} has been developped by Williams \etal~\cite{Williams15data}.
+The proposed algorithm is a way to obtain an approximation of the so-called Koopman operator of the observed system, which theoretically allows to catch dynamic of nonlinear systems~\cite{Korda18linear}.
+\item The \emph{Gaussian Process Regression (GPR)}~\cite{Rasmussen06gaussian} is a method relying on a probabilistic dynamical model.
+That one is based on the Gaussian process structure, and therefore offer more flexibility in the type of dynamic that could be modeled than the linear model~\eqref{eq:dyn.model}.
+\end{itemize}
+
+To quantify the global quality (\ie~not depending on $\ell$) of the forecasting approaches, we evaluate the Experimental Mean Square Error $\mathrm{MSE_{xp}}(\tilde\bx)$ of the forward forecast extended signals, namely:
+\begin{align}
+\label{eq:mse}
+\mathrm{MSE_{\xp}}(\tilde\bx) &= \dfrac1{L}\|\tilde\bx -\bx^\mathrm{ext}\|^2 \\
+\nonumber
+&= \dfrac1{L}\sum_{\ell=1}^L \bmu_{\xp}[\ell]^2 + \bgamma_{\xp}[\ell,\ell]\ .
+\end{align}
+where $\bx^\mathrm{ext}$ is the ground-truth extended signal, that is: $\bx^\mathrm{ext} = \begin{pmatrix}\bx[-L] & \cdots & \bx[N-1+L] \end{pmatrix}$.
+Then, as long as the bias $\bmu[\ell]$ and the variance $\bgamma[\ell,\ell]$ of the forecasting estimator remain small for all $\ell$, the MSE takes small values either.
+Corresponding results are given in Table~\ref{tab:mse.sine}.
+They show that the naive extension we propose gives satisfying results, even though the oher methods, more sophisticated gives MSE values that are somewhat smaller.
+Netherveless, a major limit of those methods is the computing time they require, which prevent them from being used to exploit real-time data.
+Thus, {\sf SigExt} is the method that optimize the trade-off between the forecasting quality and the computing time as extension method.
+That is why, it is implemented in our algorithm for the reduction of boundary effect.
 
@@
 \centering
-\caption{Averaged MSE.}
-\begin{tabular}{|c||c|c|}
+\caption{Performance of the extension methods.}
+\begin{tabular}{|c||c|c|c|}
   \hline
-   \multirow{2}{*}{Algorithm} & \multicolumn{2}{c|}{MSE} \\
-   \cline{2-3}
-      & Mean & Standard deviation\\
-   \hhline{|=#=|=|}
-   {\sf SigExt} & $0.7935$ & $0.2615$ \\
+   \multirow{2}{*}{Algorithm} & \multicolumn{2}{c|}{MSE}  & \multirow{2}{*}{Computing time (sec.)} \\
+   \cline{2-3} & Mean & Standard deviation & \\
+   \hhline{|=#=|=|=|}
+   {\sf SigExt} & $1.433\times 10^{-3}$ & $4.361\times 10^{-4}$ & $0.15$ \\
    \hline
-   Zero padding & $0.9903$ & $0.0579$ \\
+   EDMD & $3.076\times 10^{-2}$ & $8.095\times 10^{-2}$ & $2.53$\\
+   \hline
+   GPR & $1.436\times 10^{-3}$ & $4.346\times 10^{-4}$ & $146.33$ \\
    \hline
 \end{tabular}
-\label{tab:mse}
+\label{tab:mse.sine}
 \end{table}
 
-\subsubsection{Influence of the hyperparameters}
-\label{ssse:hyperparam}
-We evaluate the influence of the hyperparameters $K$ and $M$ on the forecasting MSE.
-This study is performed on a 32 second-long photoplethysmogram (PPG) signal, which is displayed on the top of Figure~\ref{fig:ppg}.
-This signal is sampled at $\fs=125$~Hz.
+\subsection{Evaluation of the quality of the boundary effect reduction}
 
-\begin{figure}
-\includegraphics[width=\textwidth]{PPGsig.eps}
-\includegraphics[width=\textwidth]{PPGforecast.eps}
-\caption{PPG signal. Top: original measured signal. Extended signal obtained by forecasting superimposed with the ground truth signal.}
-\label{fig:ppg}
-\end{figure}
-
-In this study, the extension length $L$ is fixed because the second step of the Algorithm~\ref{alg:boundary} determines the length $L$.
-Indeed, in order to avoid boundary effect, this one should be at least equal to the half of the width of the window used to perform the kernel-based transform.
-In practice, we choose to extend the signal if 2~seconds on each side, \ie~$L=250$.
-\begin{enumerate}[a),leftmargin=*]
-\item 
-We perform the forecasting Algorithm~\ref{alg:extension} for different values of the sub-signals length $M$.
-Meanwhile, the value of $K$ is unchanged and set to $K=10L$.
-The graph on the left in Figure~\ref{fig:influence.M} shows the influence of $M$ on the forecasting MSE.
-%When $M$ is smaller than $L$, the forecasting algorithm cannot provide accurate forecasting because .
-As soon as $M$ is greater than $L$, the MSE will not be influenced by the sub-signal lengths.
-Indeed, the information provided by
-\item
-We perform the forecasting Algorithm~\ref{alg:extension} for different values of the sub-signals dataset size $K$, when the value of $M$ is set to $M=\lfloor 1.5L\rfloor$.
-The graph on the left in Figure~\ref{fig:influence.M} shows the influence of $K$ on the forecasting MSE.
-\end{enumerate}
-
-\begin{figure}
-\includegraphics[width=.5\textwidth]{mseM.eps}
-\includegraphics[width=.5\textwidth]{mseK.eps}
-\caption{Influence of the size parameters on the forecasting MSE. Left: MSE in function of the sub-signals length $M$. Right: MSE in function of the sub-signal dataset size $K$.}
-\label{fig:influence.M}
-\end{figure}
-
-\paragraph{Influence of noise.}
-The same kind of study is done. The results are displayed on Figure~\ref{fig:influence.noise}. Here, $M=\lfloor 1.5L\rfloor$ and $K=2M$.
-
-\begin{figure}
-\centering
-\includegraphics[width=.7\textwidth]{mseNoise.eps}
-\caption{Influence of the noise variance on the forecasting MSE.}
-\label{fig:influence.noise}
-\end{figure}
-
-\subsection{Evaluation of the quality of the boundary effect reduction}
-The quality of the boundary effect reduction step must be evaluated of the synchrosqueezing representation.
+\subsubsection{Metrics}
+The quality of the boundary effect reduction must be evaluated on the synchrosqueezing representation.
 To that aim, we compare the obtained synchrosqueezing transform to the optimal synchrosqueezing transform $\ccF_N^\mathrm{opt}(\bx)$.
-The optimal synchrosqueezing transform is defined as the restriction of the synchrosqueezing of the ground truth extended signal $\bx^\mathrm{ext} = \begin{pmatrix}\bx[-L] & \cdots & \bx[N-1+L] \end{pmatrix}$.
-Therefore:
+The optimal synchrosqueezing transform is defined as the restriction of the synchrosqueezing of the ground-truth extended signal $\bx^\mathrm{ext}$.
+Therefore, we have:
 \begin{equation*}
@@
 
-This distance is evaluated on the dataset of respiratory signals introduced in section~\ref{ssse:validation}.
-The results are summarize in Table~\ref{tab:otd}.
-The OTD averaged over the simulations clearly shows that $\ccF_\mathrm{FC}$ outperforms $\ccF$.
-This highlights the ability of our approach to limit the distortion due the boundary effect and provide a more accurate synchrosqueezing transform.
-Moreover, the OTD decreases by almost $50\%$ from the standard SST $\ccF$ to the improved SST $\ccF_{\mathrm{FC}}$.
-Thus, the improvement obtained by forecasting in the time domain is amplified in the time-frequency domain.
+\subsubsection{Respiratory signal}
+
+The signal is 50 minutes-long and is sampled at $\fs=100$~Hz. A zoom on a small portion of the signal is displayed in Figure~\ref{fig:tho}.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{THOsig.eps}
+\caption{Zoom on the respiratory signal.}
+\label{fig:tho}
+\end{figure}
+
+From that large signal, we build a dataset of $48$ non-overlapping signals of 60 seconds, \ie~$N=6000$.
+On each of these pieces of signal, we implement the Algorithm~\ref{alg:boundary} on diverse time-frequency and time-scale representations.
+The corresponding results are given in Table~
+
+The forecasting Algorithm~\ref{alg:extension} is applied to each of the signals in the dataset we built.
+These ones are extended of 7 second on each border, corresponding to $L =700$.
+Thus, in order to catch slowly varying dynamical behaviors, the size of the training signal $M$ is chosen so that $M=\lfloor 1.5L\rfloor$.
+As a result of section~\ref{ssse:res.sine}, we take: $K=\lfloor2.5M\rfloor$.
+The average and the standard deviation of the MSE~\ref{eq:mse} with respect to the simulations are given in Table~\ref{tab:mse}.
 
@@
 \centering
-\caption{Performance of the Boundary effect reduction on the SST.}
+\caption{Averaged MSE.}
 \begin{tabular}{|c||c|c|}
   \hline
-   \multirow{2}{*}{Algorithm} & \multicolumn{2}{c|}{OTD to $\ccF_N^{\mathrm{opt}}(\bx)$} \\
+   \multirow{2}{*}{Algorithm} & \multicolumn{2}{c|}{MSE} \\
    \cline{2-3}
@@
    \hhline{|=#=|=|}
-   $\ccF_N^{\mathrm{ext}}(\bx)$ & $4.705\times 10^{-3}$ & $3.147\times 10^{-3}$ \\
+   {\sf SigExt} & $0.7935$ & $0.2615$ \\
    \hline
-   $\ccF_N(\bx)$ & $9.195\times 10^{-3}$ & $3.453\times 10^{-3}$ \\
+   Zero padding & $0.9903$ & $0.0579$ \\
    \hline
 \end{tabular}
-\label{tab:otd}
+\label{tab:mse}
 \end{table}
+
+The results are compared with the strategy consisting in a zero-padding extension of the signal.
+These results  show a significant improvement of the quality of the forecasting.
+Indeed, in average, the forecasting MSE is improved of almost $20\%$.
+Nevertheless, the variance forecasting MSE shows a huge variability in the quality of the estimation.
+This is probably due to the outliers and pulse that can be find in the respiratory signal.
+That ones make the adaptive harmonic model temporary irrelevant, and break the validity of the linear dynamic model.
+
+Even though the performance of the forecasting algorithm is moderate, the boundary effect can be reduced dramatically on the time-frequency and time-scale representations.
+We evaluate the OTD to the optimal representation for different time-frequency and time-scale representation.
+Notice that the extension length $L$ is set accordingly to the window length used by the time-frequency analysis tool.
+For instance, here the window length we use to evaluate the STFT is of 1500 samples.
+To prevent the STFT from being sensitive to the boundary effect, we set $L=750$.
+In this way, when evaluating the spectral content of the signal near its boundaries, the analysis is not limited by a lack of information all along the window support.
+From now on, all results are given for $L$ at equal to the half of the width of the window used in the time-frequency transform.
+
+\begin{table}
+\centering
+\caption{Performance of the Boundary effect reduction on TFR.}
+\begin{tabular}{|c||c|c|c|}
+  \hline
+   \multirow{2}{*}{Extension method} & \multicolumn{3}{c|}{Time-Frequency Representation} \\
+   \cline{2-4}
+      & STFT & SST & RS\\
+   \hhline{|=#=|=|=|}
+   Without extension & $2.16\times 10^{-2}$ & $5.26\times 10^{-3}$ & $3.07\times 10^{-2}$ \\
+   \hline
+   {\sf SigExt} & $1.72\times 10^{-2}$ & $4.00\times 10^{-3}$ & $2.43\times 10^{-2}$ \\
+   \hline
+   EDMD & $1.75\times 10^{-2}$ & $4.23\times 10^{-3}$ & $2.45\times 10^{-2}$ \\
+   \hline
+   GPR & $1.84\times 10^{-2}$ & $4.10\times 10^{-3}$ & $2.48\times 10^{-2}$ \\
+   \hline
+\end{tabular}
+\label{tab:otd.tho}
+\end{table}
+
+\subsubsection{Photoplethysmogram}
+\label{ssse:ppg}
+We perform a study similar to the previous one on a 640 second-long photoplethysmogram (PPG) signal, sampled at $\fs=125$~Hz.
+A 32 second-long piece of this signal is displayed on the top of Figure~\ref{fig:ppg}.
+The estimated 2-second extension on both sides of this signal is superimposed to the ground-truth signal is provided on the bottom of Figure~\ref{fig:ppg}.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{PPGsig.eps}
+\includegraphics[width=\textwidth]{PPGforecast.eps}
+\caption{PPG signal. Top: original measured signal. Extended signal obtained by forecasting superimposed with the ground truth signal.}
+\label{fig:ppg}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=.49\textwidth]{SSTBoundEffRed.eps}
+\includegraphics[width=.49\textwidth]{zoomSSTBoundEffRed.eps}
+\caption{Result of {\sf BoundEffRed} on the synchrosqueezing transform of a PPG (left) with a zoom on its right boundary (right). }
+\label{fig:ppg.boundeffred}
+\end{figure}
+
+%\begin{enumerate}[a),leftmargin=*]
+%\item 
+%We perform the forecasting Algorithm~\ref{alg:extension} for different values of the sub-signals length $M$.
+Meanwhile, the value of $K$ is unchanged and set to $K=10L$.
+The graph on the left in Figure~\ref{fig:influence.M} shows the influence of $M$ on the forecasting MSE.
+%When $M$ is smaller than $L$, the forecasting algorithm cannot provide accurate forecasting because .
+As soon as $M$ is greater than $L$, the MSE will not be influenced by the sub-signal lengths.
+Indeed, the information provided by
+%\item
+%We perform the forecasting Algorithm~\ref{alg:extension} for different values of the sub-signals dataset size $K$, when the value of $M$ is set to $M=\lfloor 1.5L\rfloor$.
+The graph on the left in Figure~\ref{fig:influence.M} shows the influence of $K$ on the forecasting MSE.
+%\end{enumerate} 
+%\begin{figure}
+%\includegraphics[width=.5\textwidth]{mseM.eps}
+%\includegraphics[width=.5\textwidth]{mseK.eps}
+%\caption{Influence of the size parameters on the forecasting MSE. Left: MSE in function of the sub-signals length $M$. Right: MSE in function of the sub-signal dataset size $K$.}
+%\label{fig:influence.M}
+%\end{figure}
+
+We divide the signal into 32-second long pieces, and apply Algorithm~\ref{alg:boundary} on each piece.
+We provide in Table~\ref{tab:otd.ppg} the OTD to the optimal time-frequency representation averaged over the signals.
+For all the considered time-frequency representations, the results clearly shows that our algorithm reduce the influence of the boundary effect.
+This highlights the ability of our approach to limit the distortion due the boundary effect and provide a more accurate representations.
+
+On Fig.~\ref{fig:ppg.boundeffred}, we display the SST resulting from {\sf BoundEffRed} strategy, applied to the same portion of PPG than what is used to display Fig.~\ref{fig:ppg}.
+We clearly observe an improvement of the quality of the STT near boundaries.
+Indeed, the blurring visible on the zoom on the right boundary of the SST has almost vanished (see right of Fig.~\ref{fig:ppg.boundeffred}).
+The real-time tracking of the instantaneous frequencies contained the measured signal is therefore largely facilitated.
+
+\begin{table}
+\centering
+\caption{Performance of the Boundary effect reduction on TFR.}
+\begin{tabular}{|c||c|c|c|}
+  \hline
+   \multirow{2}{*}{Extension method} & \multicolumn{3}{c|}{Time-Frequency Representation} \\
+   \cline{2-4}
+      & STFT & SST & RS\\
+   \hhline{|=#=|=|=|}
+   Without extension & $2.52\times 10^{-2}$ & $9.41\times 10^{-2}$ & $1.03\times 10^{-1}$ \\
+   \hline
+   {\sf SigExt} & $1.22\times 10^{-3}$ & $7.31\times 10^{-2}$ & $1.11\times 10^{-1}$ \\
+   \hline
+   EDMD & $9.83\times 10^{-4}$ & $5.59\times 10^{-2}$ & $9.80\times 10^{-2}$ \\
+   \hline
+   GPR & $1.14\times 10^{-3}$ & $7.02\times 10^{-2}$ & $1.07\times 10^{-1}$ \\
+   \hline
+\end{tabular}
+\label{tab:otd.ppg}
+\end{table}

### 2020-05-14 Notes/Supplementary material/SupMat.tex
update notes

\documentclass[journal, onecolumn]{IEEEtran}

\usepackage{graphicx}
\usepackage{bbm}
\usepackage{amsfonts,amssymb,amsmath,amsthm}
\usepackage{type1cm,eso-pic,color}
\usepackage{mathrsfs} 
\usepackage{mathpazo}
\usepackage[scaled=.95]{helvet}
\usepackage{courier}
\usepackage{xr}
\usepackage{xspace}
\usepackage{etoolbox}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{algorithm,algorithmic}

\input{../mydefs}

\title{An Approach for the Reduction of Boundary Effects in Time-Frequency Representations:\\ Part II}

\author{Adrien~Meynard, %~\IEEEmembership{Member,~IEEE,}
        Hau-Tieng~Wu
\thanks{A. Meynard and H.-T. Wu are with the Department
of Mathematics, Duke University, Durham,
NC, 27708 USA e-mail: adrien.meynard@duke.edu}}

\newtheorem{remark}{Remark}

\begin{document}
\maketitle

\setcounter{equation}{8}
\setcounter{figure}{3}

\section{Proof of Lemma~1}
\label{ap:lm.error}
We have:
\begin{align}
\dfrac1K\bX\bX^T &= \underbrace{\dfrac1K\bZ\bZ'+\sigma^2\bI}_{\defeq S^{(0)}} + \bE^{(0)} \\
\dfrac1K\bY\bX^T &= \underbrace{\dfrac1K\bZ\bZ^T +\sigma^2\bD}_{\defeq S^{(1)}}+ \bE^{(1)} \ ,
\end{align}
where $\bE^{(a)} = \sigma\bE_1^{(a)} + \sigma^2\bE_2^{(a)}$ with:
\[
\bE_1^{(a)}[m,m'] = \dfrac1K\sum_{k=0}^{K-1} \bz[N_0+m+a+k]\bw[N_0+m'+k] + \bw[N_0+m+a+k]\bz[N_0+m'+k]\ ,
\]
and
\[
\bE_2^{(a)}[m,m'] =  \dfrac1K\sum_{k=0}^{K-1} \bw[N_0+m+a+k]\bw[N_0+m'+k] - \delta_{(m+a)m'}\ ,
\]
with $a\in\{0,1\}$.
\begin{remark}
The matrices $\bE^{(0)}$ and $\bE^{(1)}$ are said to be error matrices because:
\begin{align*}
\EE\{\bE^{(0)}\} &= \EE\{\bE_1^{(0)}\} = \EE\{\bE_2^{(0)}\} = \bzero \\
\EE\{\bE^{(1)}\} &= \EE\{\bE_1^{(1)}\} = \EE\{\bE_2^{(1)}\} = \bzero\ .
\end{align*}
\end{remark}
Thus:
\begin{align*}
\tilde\bA &= (\bS^{(1)}+\bE^{(1)})(\bS^{(0)}+\bE^{(0)})^\inv \\
\bA_0&=\bS^{(1)}{\bS^{(0)}}^\inv
\end{align*}
Then:
\begin{align}
\nonumber
\bh^{(\ell)} &= \balpha^{(\ell)} - \balpha^{(\ell)}_0 \\
\nonumber
&= \be_M^T\left(\tilde\bA^\ell-\tilde\bA_0^\ell\right) \\
&= \be_M^T\left(\left((\bS^{(1)}+\bE^{(1)})(\bS^{(0)}+\bE^{(0)})^\inv\right)^\ell- \tilde\bA_0^\ell\right)\ .
\label{eq:error.vec}
\end{align}
Thus, the randomness of $\bh^{(\ell)}$ is completely originating from the random vector $\bg\in\RR^{M(M+1)}$ defined as:
\[
\bg = \vec\left(
\begin{pmatrix}
\bE^{(0)} \\
\be_M^T\bE^{(1)}
\end{pmatrix}
\right)\ .
\]
{\color{blue}
Thus, one can write $\bh^{\ell}$ as $\bh^{\ell}=f^{(\ell)}(g)$ where $f^{(\ell)}$ is a deterministic function:
\begin{align*}
f^{(\ell)} : \RR^{M(M+1)} &\to \RR^{M} \\
 \bg &\mapsto \bh^{(\ell)}\ . \\
\end{align*}
Besides, by application of the central limit theorem, the random vector $\bg$ converges in law to a Gaussian random vector:
\[
\sqrt{K}\ \bg \xrightarrow[K\to\infty]{\cD} \cN(\bzero,\bGamma_0)\ .
\]
Then, using Delta method, we have:
\begin{align*}
\sqrt{K}\ \bh^{(\ell)} \xrightarrow[K\to\infty]{\cD} \cN(\bzero,{\bF^{(\ell)}}^T\bGamma_0\bF^{(\ell)})\ ,
\end{align*}
where:
\[
\bF^{(\ell)}[m,m'] = \left.\dfrac{\partial f^{(\ell)}_m}{\partial\bg[m']}\right\vert_{\bg=\bzero}\ .
\]
}

\section{Proof of Theorem~1}
\label{ap:th.error}

\paragraph{Expression of the bias $\bmu$.}
Clearly, $\bmu[n]=0$ when $n\in I$. When $n=N-1+\ell$, we have:
\begin{align*}
\bmu[n] &=  \EE\{\balpha^{(\ell)}\}\bz_{K} + \sigma\EE\{\balpha^{(\ell)}\bw_{K}\} - \bz[n]\\
& = \balpha_0^{(\ell)}\bz_{K} + \EE\{\bh^{(\ell)}\}\bz_{K} + \sigma\EE\{\bh^{(\ell)}\bw_{K}\} - \bz[N-1+\ell]
\end{align*}
Let us first evaluate the expression of $\balpha_0^{(\ell)}\bz_K$. We have:
\begin{align*}
\bS^{(a)}[m,m'] &= \sigma^2\delta_{(m+a)m'}+\sum_{j,j'=1}^J\dfrac{\Omega_j\Omega_{j'}}{K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(N_0+m+a+k)\right)\cos\left(2\pi \frac{f_{j'}}{\fs}(N_0+m'+k)\right) \\
&= \sigma^2\delta_{(m+a)m'}+\sum_{j=1}^J\dfrac{\Omega_j^2}{2K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)\\
&= \sigma^2\delta_{(m+a)m'}+\sum_{j=1}^J\left( \dfrac{\Omega_j^2}2\cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \dfrac{\Omega_j^2}{2K}\underbrace{\sum_{k=0}^{K-1}\cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)}_{=0\ \mathrm{because}\ \frac{f_j}\fs=\frac{p'_j}K } \right)\\
&= \sigma^2\delta_{(m+a)m'}+\sum_{j=1}^J\dfrac{\Omega_j^2}2\cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right)\ .
\end{align*}
Thus, $\bS^{(0)}$ is a circulant matrix and is therefore diagonalizable in the Fourier basis:
\[
\bS^{(0)} = \bU\bLambda^{(0)}\bU^*\ ,
\]
where $\bU[m,m']=\frac1{\sqrt{M}}e^{-2\ii\pi mm'/M}$ and $\bLambda^{(0)} = \mathrm{diag}(\lambda_0^{(0)},\dots,\lambda_{M-1}^{(0)})$ with:
\begin{align*}
\lambda_m^{(0)} &= \sigma^2 + \sum_{j=1}^J\dfrac{\Omega_j^2}2\sum_{q=0}^{M-1} \cos\left(2\pi\frac{f_j}{\fs} q\right) e^{-2\ii\pi qm/M} \\
&= \sigma^2 + \dfrac{M}{4}\sum_{j=1}^J\Omega_j^2(\delta_{m,p_j} + \delta_{m,M-p_j})\ .
\end{align*}
Therefore:
\begin{align*}
{\bS^{(0)}}^\inv  &= \bU{\bLambda^{(0)}}^\inv\bU^*
\end{align*}
which leads to:
\begin{align*}
{\bS^{(0)}}^\inv[m,m']  &= \dfrac1{\sigma^2}\delta_{m,m'}-\sum_{j=1}^J\dfrac{\Omega_j^2}{2\sigma^2(\sigma^2+\Omega_j^2M/4)}\cos\left(2\pi p_j \dfrac{m-m'}{M}\right)\ ,
\end{align*}
and, consequently:
\begin{align}
\nonumber
\tilde\bA_0[m,m']  &= \sum_{q=0}^{M-1} {\bS^{(1)}}[m,q]{\bS^{(0)}}^\inv[q,m'] \\
&= \delta_{m+1,m'} + \sum_{j=1}^J\dfrac{2\Omega_j^2}{\Omega_j^2M+4\sigma^2}\cos\left(2\pi p_j\frac{m'}{M}\right)\delta_{m+1,M}
\label{eq:A0.sine}
\end{align}
Thus:
\begin{align*}
\tilde\balpha_0^{(1)}[m]  &=\sum_{j=1}^J\dfrac{2\Omega_j^2}{\Omega_j^2M+4\sigma^2}\cos\left(2\pi p_j\frac{m}{M}\right) \\
&= \dfrac{2}{M}\sum_{j=1}^J\cos\left(2\pi p_j\frac{m}{M}\right) + o(\sigma)\ .
\end{align*}
Besides, from equation~\eqref{eq:A0.sine}, we have
\begin{align*}
\tilde\bA_0\bz_{K} &= 
\begin{pmatrix}
\bz[N-M+1] \\
\vdots \\
\bz[N-1] \\
\balpha_0^{(1)}\bz_K
\end{pmatrix}
\end{align*}
By induction, we have:
\begin{align*}
\tilde\bA_0^{\ell}\bz_{K} &= 
\begin{pmatrix}
\bz[N-M+\ell] \\
\vdots \\
\bz[N-1] \\
\balpha_0^{(1)}\bz_K \\
\vdots \\
\balpha_0^{(\ell)}\bz_K
\end{pmatrix}
\ .
\end{align*}
Then:
\begin{align}
\nonumber
\balpha_0^{(\ell)}\bz_{K} &= \tilde\balpha_0^{(1)}\tilde\bA_0^{\ell-1}\bz_{K} \\
&=\sum_{m=0}^{M-\ell}\balpha_0^{(1)}[m]\bz[N-M+\ell+m-1]+\sum_{m=M-\ell+1}^{M-1}\balpha_0^{(1)}[m]\balpha_0^{(m-M+\ell)}\bz_K
\label{eq:seq}
\end{align}
But:
\begin{align*}
\balpha_0^{(1)}\bz_{K} &=\sum_{m=0}^{M-1}\balpha_0^{(1)}[m]\bz[N-M+m] \\
&=\sum_{j,j'=1}^J\Omega_{j'}\dfrac{2}{M}\underbrace{\sum_{m=0}^{M-1}\cos\left(2\pi p_j\frac{m}{M}\right)\cos\left(2\pi p_{j'}\dfrac{N+m}{M}\right)}_{=\delta_{j,j'}\frac{M}2\cos\left(2\pi p_j\frac{N}M\right)} + o(\sigma) \\
&= \sum_{j=1}^J\Omega_j\cos\left(2\pi p_j\dfrac{N}{M}\right) + o(\sigma) \\
&= \bz[N] + o(\sigma)
\end{align*}
and, by induction from~\eqref{eq:seq}:
\begin{equation}
\label{eq:alpha0z.sine}
\balpha_0^{(\ell)}\bz_{K} = \bz[N-1+\ell] + o(\sigma)
\end{equation}
Then:
\begin{equation*}
\bmu[N-1+\ell] = \EE\{\bh^{(\ell)}\}\bz_{K} + \sigma\EE\{\bh^{(\ell)}\bw_{K}\} + o(\sigma)\ .
\end{equation*}
{\color{blue}
Besides, from Lemma~1, we have the following results:
\begin{align*}
\EE\{\bh^{(\ell)}\} &\underset{K\to\infty}{=} o\left(\dfrac1{\sqrt{K}}\right)\\
\EE\{\bh^{(\ell)}\bw_{K}\} &\underset{K\to\infty}{=} o\left(\dfrac1{\sqrt{K}}\right)
\end{align*}
Consequently:
\begin{equation*}
\bmu[N-1+\ell] \underset{K\to\infty}{\sim} o(\sigma)\ .
\end{equation*}

\paragraph{Expression of the covariance $\bgamma$.}
Let us segregate the cases.

First, when $(n,n')\in I^2$, we clearly have $\bgamma[n,n']=\sigma^2\delta_{n,m}$.

Second, we focus on the case where $n=N-1+\ell\geq N$ and $n'=N-1+\lambda\geq N$. Thus:
\begin{align*}  
\bgamma[n,n'] &= \bz_K^T\EE\left\{{\balpha^{(\ell)}}^T\balpha^{(\lambda)}\right\}\bz_K + \sigma\EE\{\balpha^{(\ell)}\bw_K\balpha^{(\lambda)}\}\bz_K+ \sigma\EE\{\balpha^{(\lambda)}\bw_K\balpha^{(\ell)}\}\bz_K \\
& +\sigma^2\EE\{\balpha^{(\ell)}\bw_K\balpha^{(\lambda)}\bw_K\}-\bz[n]\bz[n']-\bz[n]\bmu[n']-\bz[n]\bmu[n'] - \bmu[n]\bmu[n'] \\
&= \bz_K^T\EE\left\{{\bh^{(\ell)}}^T\bh^{(\lambda)}\right\}\bz_K + \balpha_0^{(\ell)}\sigma\EE\{\bw_K\bh^{(\lambda)}\}\bz_K + \sigma\EE\{\bh^{(\ell)}\bw_K\}\bz(n']  \\
&+ \sigma\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\}\bz_K+ \sigma\EE\{\bh^{(\lambda)}\bw_K\}\bz[n] + \sigma\balpha_0^{(\lambda)}\EE\{\bw_K\bh^{(\ell)}\}\bz_K \\
& + \sigma\EE\{\bh^{(\lambda)}\bw_K\bh^{(\ell)}\}\bz_K+\sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\lambda)}\bw_K\}+\sigma^2\balpha_0^{(\lambda)}\EE\{\bw_K\bh^{(\ell)}\bw_K\} \\
&+\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\ell)}\right\rangle +\sigma^2\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} + o(\sigma^2)
\end{align*}
From lemma~1, we have:
To be continued...

Third, we focus on the case where $n=N-1+\ell\geq N$ and $n'\in I$. The case $n\in I$ and $n'=N-1+\lambda$ is directly derived from the current one.
To be continued...
%\begin{align*}
%\bgamma[\ell,\ell] &\leq \dfrac1K\left(\sqrt{\bz_K^T\EE\{{\bu^{(\ell)}}^T\bu^{(\ell)}\}\bz_K} + \sigma\sqrt{\EE\{(\bu^{(\ell)}\bw_K)^2\}}\right)^2 \\
%&\ + \dfrac{2\sigma}{\sqrt{K}}\|\balpha_0^{(\ell)}\|\left(\sqrt{\bz_K^T\EE\{{\bu^{(\ell)}}^T\bu^{(\ell)}\}\bz_K} +\sigma\sqrt{\EE\{(\bu^{(\ell)}\bw_K)^2\}} \right)+ \sigma^2\|\balpha_0^{(\ell)}\|^2 - \bmu[\ell]^2 + o(\sigma^2) \\
%\end{align*}
%
%Then, Isserlis formula gives us:
%\begin{align*}
%\EE\{(\bu^{(\ell)}\bw_K)^2\} &= \sum_{m,m'=0}^{M-1}\EE\{\bu^{(\ell)}[m]\bw_K[m]\bu^{(\ell)}[m']\bw_K[m']\}\\
%&\underset{K\to\infty}{\sim} \sum_{m,m'=0}^{M-1}\EE\{\bu^{(\ell)}[m]\bw_K[m]\}\EE\{\bu^{(\ell)}[m']\bw_K[m']\}+\EE\{\bu^{(\ell)}[m]\bu^{(\ell)}[m']\}\EE\{\bw_K[m]\bw_K[m']\} \\
%&\hspace{30pt}+\EE\{\bu^{(\ell)}[m]\bw_K[m']\}\EE\{\bu^{(\ell)}[m']\bw_K[m]\}\\
%&\leq \Tr(\bGamma^{(\ell)}) + 2\sum_{m,m'=0}^{M-1}\sqrt{\bGamma^{(\ell)}[m,m]\bGamma^{(\ell)}[m',m']}
%\end{align*}
%So:
%\begin{align*}
%\bgamma[\ell,\ell] &\leq \dfrac1K\left(\sqrt{\bz_K^T\bGamma^{(\ell)}\bz_K} + \sigma\sqrt{\Tr(\bGamma^{(\ell)}) + 2\sum_{m,m'=0}^{M-1}\sqrt{\bGamma^{(\ell)}[m,m]\bGamma^{(\ell)}[m',m']}}\right)^2 \\
%&\ + \dfrac{2\sigma}{\sqrt{K}}\|\balpha_0^{(\ell)}\|\left(\sqrt{\bz_K^T\bGamma^{(\ell)}\bz_K} +\sigma\sqrt{\Tr(\bGamma^{(\ell)}) + 2\sum_{m,m'=0}^{M-1}\sqrt{\bGamma^{(\ell)}[m,m]\bGamma^{(\ell)}[m',m']}} \right)+ \sigma^2\|\balpha_0^{(\ell)}\|^2 + o(\sigma^2) \\
%\end{align*}
}

\end{document}

### 2020-05-14 Notes/algorithm.tex
update notes

@@
+As explained above, the algorithm for the reduction of boundary effects on time-frequency representations relies on two steps. These ones are detailed in the current section.
+
 \subsection{Forecasting}
 
-\paragraph{Notations.} Let $x:\RR\to\RR$ denote a continuous-time signal.
-In this work, we consider a finite-length discretization of that one.
-Thus, the sampled signal $\bx$, whose length is denoted by $N$, is such that
+Let $x:\RR\to\RR$ denote a continuous-time signal. In this work, we consider a finite-length discretization of that one. Thus, the sampled signal $\bx$, whose length is denoted by $N$, is such that
 \[
@@
 
-Let $M<N$, $K<N$. Then, for all $k\in\{0,\ldots,K-1\}$, we extract from $\bx\in\RR^N$ the sub-signal $\bx_k\in\RR^M$ given by:
-\[
+\paragraph{Notations} 
+Let $M$ and $K$ be two integers such that $M<N$ and $K+M<N$. Then, for all $k\in\{0,\ldots,K-1\}$, we extract from $\bx\in\RR^N$ the sub-signal $\bx_k\in\RR^M$ given by:
+\begin{equation}
 \bx_k = 
@@
 \end{pmatrix}\ .
-\]  
+\label{eq:xk}
+\end{equation} 
 These sub-signals are gathered into the matrix $\bX\in\RR^{M\times K}$ such that:
-\[
+\begin{equation*}
 \bX = 
@@
 \end{pmatrix}\ .
-\]
-Notice that these sub-signals are overlapping each other. Indeed, $\bx_{k+1}$ is a shifting of $\bx_k$ from one sample.
-
-We also consider the matrix $\bY\in\RR^{M\times K}$ given by:
-\[
+\end{equation*}
+Notice that these sub-signals are overlapping each other. Indeed, $\bx_{k+1}$ is a shifting of $\bx_k$ from one sample. We also consider the matrix $\bY\in\RR^{M\times K}$ given by:
+\begin{equation*}
 \bY = 
@@
 \end{pmatrix}\ .
-\]
+\end{equation*}
 
-\paragraph{Dynamic model and forecasting.} Establishing a dynamic model consists in determining the relation linking $\bY$ to $\bX$, that is finding a function $f$ so that
+\paragraph{Dynamical model and forecasting} 
+Establishing a dynamical model consists in determining the relation linking $\bY$ to $\bX$, that is finding a function $f$ so that
 \[
@@
 \] 
-We consider here a naive dynamic model, assuming that we have the following relation:
+We consider here a naive dynamical model, assuming that we have the following relation:
 \begin{equation}
@@
 \end{equation}
-where $\bA\in\RR^{M\times M}$. This is a linear dynamic model, which can be written equivalently in function of the sub-signals $\bx_k$, as:
+where $\bA\in\RR^{M\times M}$.
+We adopt a classical strategy in the study of dynamical systems, that is the linearization of a nonlinear phenomenon.
+Notice that this linear dynamical model can be written equivalently in function of the sub-signals $\bx_k$, as:
 \begin{equation}
@@
 \label{eq:prediction}
-\tilde \bx [N-1+\ell] = \be_M^T\tilde\bA^\ell\bx_{K}\ ,
+\tilde \bx [N-1+\ell] = \balpha^{(\ell)}\bx_{K}\ ,
+\end{equation}  
+where $\balpha^{(\ell)}$ denotes the last row of $\tilde\bA^\ell$, that is to say:
+\begin{equation}
+\balpha^{(\ell)}=\be_M^T\tilde\bA^\ell\ ,
+\label{eq:alpha.l}
 \end{equation}
@@
 
-\paragraph{Model estimation.} To estimate the matrix $\bA$, we basically implement the least square estimator. Thus, we solve the following problem:
+\paragraph{Model estimation} To estimate the matrix $\bA$, we basically implement the least square estimator. Thus, we solve the following problem:
 \begin{equation}
@@
 \]
-Therefore, solving the problem~\eqref{eq:ls.pb}, \ie~$\nabla \cL(\tilde\bA)=\bzero$, gives the following estimate $\tilde\bA$ of the dynamic model matrix $\bA$:
+Therefore, solving the problem~\eqref{eq:ls.pb}, \ie~$\nabla \cL(\tilde\bA)=\bzero$, gives the following estimate $\tilde\bA$ of the dynamical model matrix $\bA$:
 \begin{align}
@@
 \]
-Then, except the row vector $\balpha = \left(\alpha_1 \cdots\alpha_M\right)$, the matrix $\bA$ is fully determined by the dynamic model.
+Then, except the row vector $\balpha = \left(\alpha_1 \cdots\alpha_M\right)$, the matrix $\bA$ is fully determined by the dynamical model.
 \end{remark}
 
-\paragraph{Signal extension.} We can finally construct the extended signal $\tilde\bx\in\RR^{N+2L}$ concatenating the backward prediction $\tilde\bx_{\rm bw}$, the observed signal $\bx$, and the forward prediction $\tilde\bx_{\rm fw}$.
+\paragraph{Signal extension} In order to reduce the boundary effects on both "sides" of the time-frequency (or time-scale) representation, we finally construct the extended signal $\tilde\bx\in\RR^{N+2L}$ concatenating the backward prediction $\tilde\bx_{\rm bw}$, the observed signal $\bx$, and the forward prediction $\tilde\bx_{\rm fw}$.
 We summarize the extension step in Algorithm~\ref{alg:extension}.
@@
 
-\paragraph{Transform restriction.} 
 Let $\ccF_{N}:\RR^{N}\to\RR^{F\times N}$ generically denotes the time-frequency or time-scale representation we are interested in.
@@
 This strategy moves the boundary effects out of the time interval $I=[0, \frac{N-1}{\fs}]$.
-Finally, the boundary-effects insensitive representation $\ccF_N^{\mathrm{ext}}:\RR^{N}\to\RR^{F\times N}$ of $\bx$ is given by:
+Finally, the boundary-effects insensitive representation $\ccF_N^{\mathrm{ext}}:\RR^{N}\to\RR^{F\times N}$ of $\bx$ is given for all $f\in\{0,\cdots,F-1\}$, $n\in\{0,\cdots,N-1\}$ by:
 \begin{equation}
-\ccF_N^{\mathrm{ext}}(\bx)[f,n] = \ccF_{N+2L}(\tilde \bx)[f,L+n]\ ,\quad\forall f\in\{0,\cdots,F-1\},\ n\in\{0,\cdots,N-1\}\ .
+\ccF_N^{\mathrm{ext}}(\bx)[f,n] = \ccF_{N+2L}(\tilde \bx)[f,L+n]\ .
 \label{eq:restriction}
@@
 
-\paragraph{Conclusion.}
+\subsection{Global algorithm}
 Finally, the global procedure we implement to reduce boundary effects on windowing-based representations is summarized by the pseudo-code of Algorithm~\ref{alg:boundary}.

### 2020-05-14 Notes/results.tex
update notes

@@
 \subsection{Evaluation the forecasting performance}
-In that section, we first evaluate the quality of the forecasting step. That one depends on at least two parameters:
+In that section, we first evaluate the quality of the forecasting step and compare it the the theoretical results provided by Theorem~\ref{th:error}.
+The level of the forecasting error one depends on at least two parameters:
 \begin{itemize}
@@
 We proved that the linear dynamic model is sufficient to catch the dynamical behavior of signals taking the form~\eqref{eq:sum.sine}.
-In order to validate this theoretical result, we apply the forecasting Algorithm~\ref{alg:extension} to a large number of realizations of the random vector $\bz$ such that:
+In order to validate this theoretical result, we apply the forecasting Algorithm~\ref{alg:extension} to a large number of realizations of the random vector $\bx$ following the model~\eqref{eq:model.noise}, and such that the deterministic component $\bz$ takes the form:
 \[
-\bz = \bx+ \sigma\bw\ ,
+\bz[n]\! =\!\cos\!\left(2\pi p_1 \dfrac{n}{M} \right)\! +\! R\cos\!\left(2\pi p_2 \dfrac{n}{M} \right),\quad \!\forall n\!\in\!\{1,\ldots,N\},
 \]
-where
-\[
-\bx[n] =\cos\left(2\pi p_1 \dfrac{n}{M} \right) + R\cos\left(2\pi p_2 \dfrac{n}{M} \right)\ ,\quad\forall\ n\in\{1,\ldots,N\}\ ,
-\]
-and $N=10^4$, $M=750$, $p_1=10$, $p_2=33$ and $R=1.4$. Besides, the additive noise is chosen to be Gaussian: $\bw\sim\cN(\bzero,\bI)$.
+with $N=10^4$, $M=150$, $p_1=10$, $p_2=33$ and $R=1.4$. Besides, the additive noise is chosen to be Gaussian: $\bw\sim\cN(\bzero,\bI)$.
 
-\paragraph{Influence of the noise variance $\sigma^2$.} Here, the size of the training dataset is set to $K=1875$.
-Then, the forecasting algorithm is run on 100 realizations on the discrete signal $\bx$ for 100 different values of $\sigma^2$  logarithmically spaced from $5.10^{-3}$ to $10^{-1}$.
-For each of these values, we determine the experimental bias $\mu_{xp}[\ell]$ and variance $\gamma_{xp}[\ell,\ell]$ in function of the forecasting sample index $\ell$ (going from 1 to $L=500$).
-These results are displayed on Figure~\ref{fig:res.noise.sine}.
-Each color corresponds to these experimental results obtained for a given value of $\sigma$.
-On the left of Figure~\ref{fig:res.noise.sine}, the result show that the bias is neither depending on the noise variance $\sigma^2$ nor the forecasting length $\ell$, and always insignificant with respect to the signal values.
-On the right of Figure~\ref{fig:res.noise.sine}, we display the normalized experimental variance, that is $\frac{\gamma[\ell,\ell]}{\sigma^2}$.
-This highligths the fact that the forecasting variance increases linearly with respect to $\sigma^2$.
-More surprisingly, this result shows that the forecasting variance decreases with $\ell$, what is counterintuitive.
+\paragraph{Influence of the noise variance $\sigma^2$} Here, the size of the training dataset is set to $K=450$.
+Then, the forecasting algorithm is run on $1000$ realizations on the discrete signal $\bx$ for three different values of $\sigma$, logarithmically spaced from $10^{-3}$ to $10^{-1}$.
+For each of these values, we determine the experimental bias $\mu_{xp}[N-1\ell]$ and variance $\gamma_{xp}[N-1+\ell,N-1+\ell]$ in function of the forecasting sample index $\ell$ (going from $1$ to $L=100$).
+
+The experimental results show that the bias is neither depending on the noise variance $\sigma^2$ nor the forecasting length $\ell$.
+Indeed, independently of $sigma$, we always have $\mu_{xp}[N-1+\ell]\in\left[-0.03\sigma,0.03\sigma\right]$, which is negligible with respect to the magnitude of $\bz$.
+This result confirms the theoretical result~\eqref{eq:mean.error}.
+
+On Figure~\ref{fig:res.noise.sine}, we display the experimental variance $\gamma_{xp}[N-1+\ell,N-1+\ell]$ for each value of $\sigma$ ($+$ markers).
+The associated theoretical asymptotic forecasting variance~\eqref{eq:cov.error.2} is also displayed in solid line.
+As expected, this result highlights the fact that the forecasting variance increases linearly with respect to $\sigma^2$.
+More surprisingly, this result shows that the forecasting variance slightly decreases with $\ell$, what is counterintuitive.
+It should be noted that, contrary to what expression~\eqref{eq:cov.error.2} suggests, smaller values of $\sigma$ do not  cause a decrease of the experimental variance, but an increase.
+This comes from the fact that when $\sigma$ is small, the matrix $\bX\bX^T$ becomes ill-conditioned.
+The calculation of the forecasting matrix $\tilde\bA$ in~\eqref{eq:lse} is then strongly disturbed.
 \begin{figure}
-\includegraphics[width=.49\textwidth]{biasNoiseSine.eps}
-\includegraphics[width=.49\textwidth]{VarianceNoiseSine.eps}
-\caption{Evolution of the experimental forecasting bias (left) and the normalized forecasting variance (right) in function of the forecasting sample index.
-Each color characterizes the result for a given value of $\sigma$.}
+%\includegraphics[width=.24\textwidth]{biasNoiseSine.eps}
+\includegraphics[width=.48\textwidth]{VarianceNoiseSine.eps}
+\caption{Evolution of the experimental and theoretical forecasting variance in function of the forecasting sample index for different values of $\sigma$.}
 \label{fig:res.noise.sine}
@@
 
-\paragraph{Influence of the training dataset size $K$.} Here, the noise variance $\sigma$ is set to $\sigma=10^{-2}$.
-Then, the forecasting algorithm is run on 3000 realizations on the discrete signal $\bx$ for 25 different values of $K$, logarithmically spaced from $3.5\times 10^{3}$ to $6\times 10^{3}$.
-For each of these values, we determine the experimental bias $\mu_{xp}[\ell]$ and variance $\gamma_{xp}[\ell,\ell]$ in function of the forecasting sample index $\ell$ (going from $1$ to $500$).
-These results are displayed on Figure~\ref{fig:res.size.sine}.
+\paragraph{Influence of the training dataset size $K$} Here, the noise variance $\sigma$ is set to $\sigma=10^{-2}$.
+Then, the forecasting algorithm is run on 3000 realizations on the discrete signal $\bx$ for three different values of $K$, logarithmically spaced from $4.5\times 10^{2}$ to $2\times 10^{3}$.
+For each of these values, we determine the experimental bias $\mu_{xp}[N-1+\ell]$ and variance $\gamma_{xp}[N-1+\ell,N-1+\ell]$ in function of the forecasting sample index $\ell$ (going from $1$ to $500$).
+
+As in the previous study, the experimental bias vanishes when $K$ increases, what confirms the approximation result~\eqref{eq:mean.error}.
+Besides, the experimental variance  is displayed on Figure~\ref{fig:res.size.sine} ($+$ markers), and compared with the associated theoretical variance~\eqref{eq:cov.error.2}.
 Each color corresponds to these experimental results obtained for a given value of $K$.
-On the left of Figure~\ref{fig:res.size.sine}, as expected, the experimental bias vanishes.
-On the right of Figure~\ref{fig:res.size.sine}, we display the product of $K$ with the normalized experimental variance, that is $\frac{\gamma[\ell,\ell]}{\sigma^2}$.
-This result shows the asymptotic behavior of the variance.
-Indeed, as soon as $K$ is sufficiently high, the product $K\gamma[\ell,\ell]$ is approximately independent of $K$.
+This result validates the asymptotic behavior provided by~\eqref{eq:cov.error.2}, and we can show that, as soon as $K$ is sufficiently high, the product $K\gamma[\ell,\ell]$ is approximately independent of $K$.
 \begin{figure}
-\includegraphics[width=.49\textwidth]{BiasKSine.eps}
-\includegraphics[width=.49\textwidth]{VarianceKSine.eps}
-\caption{Evolution of the experimental forecasting bias (left) and the normalized forecasting variance (right) in function of the forecasting sample index.
-Each color characterizes the result for a given value of $K$.}
+%\includegraphics[width=.24\textwidth]{BiasKSine.eps}
+\includegraphics[width=.48\textwidth]{VarianceKSine.eps}
+\caption{Evolution of the experimental and theoretical forecasting variance in function of the forecasting sample index for different values of $K$.}
 \label{fig:res.size.sine}
@@
 
-\paragraph{Conclusion.}
-Both previous results allows us to describe the influence of the noise variance and the size of the training dataset on the variance of the forecasting noise, which is empirically summarized as follows:
+\paragraph{Conclusion}
+Both previous experimental results combined with the theoretical asymptotic equation~\eqref{eq:cov.error.2} allow us to describe the influence of the noise variance and the size of the training dataset on the variance of the forecasting noise, which is empirically summarized as follows:
 \begin{equation}
-\gamma[\ell,\ell] \sim \dfrac{\sigma^2}{K}g[\ell] \ ,\quad\mbox{when}\ \sigma^2\to 0\ \mbox{and}\ K\to\infty \ ,
+\gamma[N-1+\ell,N-1+\ell] \underset{K\to\infty}{\approx} \dfrac{\sigma^2}{K}g[\ell] .
 \end{equation} 
-where $g$ is a bounded positive function.
+where $g$ is a bounded positive function. The empirical result is coherent with the theoretical result provided by Theorem~\ref{th:error}.
+
+This study neglects the analysis of the influence of the parameter $M$, whose influence on the value of the experimental variance is numerically not significant as long as $M\ll 2K$.
+The choice of this parameter is especially crucial when the deterministic component of the signal is no longer stationary.
+The adaptive harmonic model, discussed below, is an example.
 
@@
 \label{ssse:res.ahm}
-We now consider a signal whose instantaneous frequencies and amplitudes of its components vary over time. The toy signal we use takes the following form:
+We now consider a signal whose instantaneous frequencies and amplitudes of its components vary over time.
+The deterministic component $\bz$ of the random vector $\bx$ ( constructed following the model~\eqref{eq:model.noise}) takes the following form, for all $n\in\{1,\ldots,N\}$:
 \[
-\bx[n] = \cos\left(2\pi \phi_1[n] \right) + R[n]\cos\left(2\pi \phi_2[n] \right)\ ,\quad\forall\ n\in\{1,\ldots,N\} \ ,
+\bx[n] = \cos\left(2\pi \phi_1[n] \right) + R[n]\cos\left(2\pi \phi_2[n] \right) \ ,
 \] 
@@
 \end{align*}
-Besides, the additive noise is chosen to be Gaussian: $\bw\sim\cN(\bzero,\bI)$. Numerically, we take: $N=10^4$, $M=750$, $p_1=10$, $p_2=23$ and $R=1.4$.
+Besides, the additive noise is chosen to be Gaussian: $\bw\sim\cN(\bzero,\bI)$. Numerically, we take: $N=10^4$, $M=750$, $p_1=10$, $p_2=23$.
 
@@
 That one is based on the Gaussian process structure, and therefore offer more flexibility in the type of dynamic that could be modeled than the linear model~\eqref{eq:dyn.model}.
+\item The \emph{Trigonometric , Box-Cox transformation, ARMA errors, Trend and Seasonal components (TBATS)} method~\cite{Rasmussen06gaussian} is based on a classical decomposition of times series into a trend, a seasonal and an ARMA components, with a specific dynamic for the seasonal component.
+This model demands the estimation of numerous parameters and, by implication, may be slow.
 \end{itemize}
@@
 \nonumber
-&= \dfrac1{L}\sum_{\ell=1}^L \bmu_{\xp}[\ell]^2 + \bgamma_{\xp}[\ell,\ell]\ .
+&=\! \dfrac1{L}\sum_{\ell=1}^L \bmu_{\xp}[N\!-\!1\!+\!\ell]^2 \!+\! \bgamma_{\xp}[N\!-\!1\!+\!\ell,N\!-\!1\!+\!\ell] .
 \end{align}
 where $\bx^\mathrm{ext}$ is the ground-truth extended signal, that is: $\bx^\mathrm{ext} = \begin{pmatrix}\bx[-L] & \cdots & \bx[N-1+L] \end{pmatrix}$.
-Then, as long as the bias $\bmu[\ell]$ and the variance $\bgamma[\ell,\ell]$ of the forecasting estimator remain small for all $\ell$, the MSE takes small values either.
+Then, as long as the bias $\bmu[N-1+\ell]$ and the variance $\bgamma[N-1+\ell,N-1+\ell]$ of the forecasting estimator remain small for all $\ell$, the MSE takes small values either.
 Corresponding results are given in Table~\ref{tab:mse.sine}.
-They show that the naive extension we propose gives satisfying results, even though the oher methods, more sophisticated gives MSE values that are somewhat smaller.
-Netherveless, a major limit of those methods is the computing time they require, which prevent them from being used to exploit real-time data.
-Thus, {\sf SigExt} is the method that optimize the trade-off between the forecasting quality and the computing time as extension method.
-That is why, it is implemented in our algorithm for the reduction of boundary effect.
+They show that the naive extension we propose gives satisfying results, even though the other methods, more sophisticated, give MSE values that are somewhat smaller.
+Nevertheless, a major limit of those methods is the computing time they require, which prevent them from being used to exploit real-time data.
+Thus, {\sf SigExt} is the extension method that optimize the trade-off between the forecasting quality and the computing time.
+That is why, it is implemented in our algorithm for the reduction of boundary effects.
 
@@
   \hline
-   \multirow{2}{*}{Algorithm} & \multicolumn{2}{c|}{MSE}  & \multirow{2}{*}{Computing time (sec.)} \\
+   \multirow{2}{*}{Algorithm} & \multicolumn{2}{c|}{MSE}  & \multirow{2}{41pt}{Computing time (sec.)} \\
    \cline{2-3} & Mean & Standard deviation & \\
@@
    \hline
+   TBATS & ${\color{red}X}$ & ${\color{red}X}$ & ${\color{red}X}$ \\
+   \hline
 \end{tabular}
@@
 \subsubsection{Metrics}
-The quality of the boundary effect reduction must be evaluated on the synchrosqueezing representation.
-To that aim, we compare the obtained synchrosqueezing transform to the optimal synchrosqueezing transform $\ccF_N^\mathrm{opt}(\bx)$.
-The optimal synchrosqueezing transform is defined as the restriction of the synchrosqueezing of the ground-truth extended signal $\bx^\mathrm{ext}$.
+The quality of the boundary effect reduction must be evaluated directly on the time-frequency representation.
+To that aim, we compare the obtained representation to the optimal representation $\ccF_N^\mathrm{opt}(\bx)$.
+The optimal representation is defined as the restriction of the representation of the ground-truth extended signal $\bx^\mathrm{ext}$.
 Therefore, we have:
@@
 
-The different representations we obtained must be compared to the optimal SST.
-To that aim, we establish a criterion that quantify the distance of these time-frequency representations to the optimal one.
-To do this, we reason in analogy with the optimal transport distance which allows us to quantify the distance between two probability density functions.
+In the aim of comparing the different techniques, we uses a criterion, proposed in~\cite{Daubechies16conceft}, that quantify the distance between a given time-frequency representation and the optimal one.
+It is built in analogy with the optimal transport distance, which enables quantifying the distance between two probability density functions.
 Let us generically denote a time frequency representation $\ccQ$.
-Then, for $t$ fixed, we consider the following pseudo probability density function:
+Then, for $t$ fixed, we consider the following pseudo-probability density function:
 \begin{align}
@@
 \end{align}
-Similarly, let $p_{\ccF_0}^t$ denotes the associated pseudo-probability density function (defined as in~\eqref{eq:ppdf}) associated with the reference time-frequency representation $\ccF_0$.
-At each instant $t$, we can then determine the optimal transport distance $d_{t}$ between the two pseudo densities.
+Then, at each instant $t$, we can then determine the optimal transport distance $d_{t}$ between the two pseudo densities.
 It is given by the $L^1$ norm of the difference between the associated distribution functions.
-In other words, either $P_{\ccQ}^t(\xi)=\int_{-\infty}^\xi p_{\ccQ}^t(\nu)\dd\nu$ and $\tilde P_{\ccF_0}^t(\xi)=\int_{-\infty}^\xi\tilde p_{\ccF_0}^t(\nu)\dd\nu$, we have
+In other words, we have:
 \begin{equation*}
-d_{t}(\ccQ,\ccF_0) = \int_\RR\left|\tilde P_{\ccQ}^t(\xi)-  P_{\ccF_0}^t(\xi)\right|\dd\xi\ .
+d_{t}(\ccQ,\ccF_0) = \int_\RR\left|\tilde P_{\ccQ}^t(\xi)-  P_{\ccF_0}^t(\xi)\right|\dd\xi\ ,
 \end{equation*}
+where $P_{\ccQ}^t(\xi)=\int_{-\infty}^\xi p_{\ccQ}^t(\nu)\dd\nu$ and $\tilde P_{\ccF_0}^t(\xi)=\int_{-\infty}^\xi\tilde p_{\ccF_0}^t(\nu)\dd\nu$.
 Finally, the distance between the two time-frequency representations is obtained by averaging all the optimal transport distances with respect to time:
 \begin{equation}
-D(\ccQ,\ccF_0) = 100\times\frac1{|I|}\int_I d_{t}(\ccQ,\ccF_0)\dd t\ .
+D(\ccQ,\ccF_0) = \frac1{|I|}\int_I d_{t}(\ccQ,\ccF_0)\dd t\ .
 \end{equation}
@@
 
+Let us evaluate the quality of the boundary effects reduction on biomedical signals.
+
 \subsubsection{Respiratory signal}
-
-The signal is 50 minutes-long and is sampled at $\fs=100$~Hz. A zoom on a small portion of the signal is displayed in Figure~\ref{fig:tho}.
+We first consider a 50 minutes-long respiratory signal. This signal is sampled at $\fs=100$~Hz. A zoom on a small portion of the signal is displayed in Figure~\ref{fig:tho}.
 
 \begin{figure}
-\includegraphics[width=\textwidth]{THOsig.eps}
-\caption{Zoom on the respiratory signal.}
+\includegraphics[width=.48\textwidth]{THOsig.eps}
+\includegraphics[width=.48\textwidth]{THOforecast.eps}
+\caption{Zoom on the respiratory signal. Top: original measured signal. Extended signal obtained by {\sf SigExt} forecasting superimposed with the ground truth signal.}
 \label{fig:tho}
@@
 From that large signal, we build a dataset of $48$ non-overlapping signals of 60 seconds, \ie~$N=6000$.
-On each of these pieces of signal, we implement the Algorithm~\ref{alg:boundary} on diverse time-frequency and time-scale representations.
-The corresponding results are given in Table~
-
-The forecasting Algorithm~\ref{alg:extension} is applied to each of the signals in the dataset we built.
-These ones are extended of 7 second on each border, corresponding to $L =700$.
+On each of these pieces of signal, we implement the forecasting method introduced in section~\ref{ssse:res.ahm}, including the {\sf SigExt} method detailed in Algorithm~\ref{alg:extension}.
+However, the TBATS extension method in not implemented here because of its excessive computing time.
+The extensions of $7$ seconds-long on each boundary, corresponding to $L =700$.
 Thus, in order to catch slowly varying dynamical behaviors, the size of the training signal $M$ is chosen so that $M=\lfloor 1.5L\rfloor$.
 As a result of section~\ref{ssse:res.sine}, we take: $K=\lfloor2.5M\rfloor$.
-The average and the standard deviation of the MSE~\ref{eq:mse} with respect to the simulations are given in Table~\ref{tab:mse}.
+The average and the median of the MSE~\eqref{eq:mse} with respect to the simulations are given in Table~\ref{tab:mse}.
+The averaged MSE of {\sf SigExt} method is overwhelmingly higher than the averaged MSE of the other methods.
+This result is caused by the presence of outliers in some extensions by {\sf SigExt}.
+Indeed, the value median of the MSE is of the same order than those of the other methods.
+These outliers are probably due to the presence of pulse in the respiratory signal.
+That ones make the adaptive harmonic model temporary irrelevant, and break the validity of the linear dynamical model~\eqref{eq:dyn.model} used to extend the signal.
+That is why {\sf SigExt} is less robust to the vanishing of oscillations that the GPR or EDMD extensions.
 
@@
 \centering
-\caption{Averaged MSE.}
+\caption{MSE of the respiratory signal extensions.}
 \begin{tabular}{|c||c|c|}
@@
    \cline{2-3}
-      & Mean & Standard deviation\\
+      & Mean & Median\\
    \hhline{|=#=|=|}
-   {\sf SigExt} & $0.7935$ & $0.2615$ \\
+   {\sf SigExt} & $1.704 \times 10^{1}$ & $7.346 \times 10^{-1}$ \\
    \hline
-   Zero padding & $0.9903$ & $0.0579$ \\
+   EDMD & $2.172\times 10^{-2}$ & $7.349\times 10^{-1}$ \\
+   \hline
+   GPR & $4.476\times 10^{-2}$ & $1.998\times 10^{-2}$ \\
    \hline
@@
 
-The results are compared with the strategy consisting in a zero-padding extension of the signal.
-These results  show a significant improvement of the quality of the forecasting.
-Indeed, in average, the forecasting MSE is improved of almost $20\%$.
-Nevertheless, the variance forecasting MSE shows a huge variability in the quality of the estimation.
-This is probably due to the outliers and pulse that can be find in the respiratory signal.
-That ones make the adaptive harmonic model temporary irrelevant, and break the validity of the linear dynamic model.
-
-Even though the performance of the forecasting algorithm is moderate, the boundary effect can be reduced dramatically on the time-frequency and time-scale representations.
-We evaluate the OTD to the optimal representation for different time-frequency and time-scale representation.
-Notice that the extension length $L$ is set accordingly to the window length used by the time-frequency analysis tool.
+Then, we apply the boundary effects reduction method {\sf BoundEffRed} (Algorithm~\ref{alg:boundary}) on these extensions for diverse time-frequency representations (STFT, SST and RS).
+In Table~\ref{tab:otd.tho}, the results are compared in terms of OTD.
+They are also compared with the strategy consisting in a zero-padding extension of the signal.
+Even though the performance of the forecasting Algorithm~\ref{alg:extension} is somehow moderate, the boundary effects can be reduced dramatically on the time-frequency representations.
+Notice that the extension length $L$ has been set accordingly to the window length used by the time-frequency analysis tool.
 For instance, here the window length we use to evaluate the STFT is of 1500 samples.
@@
 \centering
-\caption{Performance of the Boundary effect reduction on TFR.}
+\caption{Respiratory signal: OTD of the boundary effects reduction methods on diverse TFR.}
 \begin{tabular}{|c||c|c|c|}
@@
 A 32 second-long piece of this signal is displayed on the top of Figure~\ref{fig:ppg}.
-The estimated 2-second extension on both sides of this signal is superimposed to the ground-truth signal is provided on the bottom of Figure~\ref{fig:ppg}.
+The estimated 2-second extension obtained by {\sf SigExt} on both boundary of this signal is superimposed to the ground-truth signal in the bottom of Figure~\ref{fig:ppg}.
 
 \begin{figure}
-\includegraphics[width=\textwidth]{PPGsig.eps}
-\includegraphics[width=\textwidth]{PPGforecast.eps}
+\includegraphics[width=.48\textwidth]{PPGsig.eps}
+\includegraphics[width=.48\textwidth]{PPGforecast.eps}
 \caption{PPG signal. Top: original measured signal. Extended signal obtained by forecasting superimposed with the ground truth signal.}
@@
 \end{figure}
-
-\begin{figure}
-\includegraphics[width=.49\textwidth]{SSTBoundEffRed.eps}
-\includegraphics[width=.49\textwidth]{zoomSSTBoundEffRed.eps}
-\caption{Result of {\sf BoundEffRed} on the synchrosqueezing transform of a PPG (left) with a zoom on its right boundary (right). }
-\label{fig:ppg.boundeffred}
-\end{figure}
-
-%\begin{enumerate}[a),leftmargin=*]
-%\item 
-%We perform the forecasting Algorithm~\ref{alg:extension} for different values of the sub-signals length $M$.
-Meanwhile, the value of $K$ is unchanged and set to $K=10L$.
-The graph on the left in Figure~\ref{fig:influence.M} shows the influence of $M$ on the forecasting MSE.
-%When $M$ is smaller than $L$, the forecasting algorithm cannot provide accurate forecasting because .
-As soon as $M$ is greater than $L$, the MSE will not be influenced by the sub-signal lengths.
-Indeed, the information provided by
-%\item
-%We perform the forecasting Algorithm~\ref{alg:extension} for different values of the sub-signals dataset size $K$, when the value of $M$ is set to $M=\lfloor 1.5L\rfloor$.
-The graph on the left in Figure~\ref{fig:influence.M} shows the influence of $K$ on the forecasting MSE.
-%\end{enumerate} 
-%\begin{figure}
-%\includegraphics[width=.5\textwidth]{mseM.eps}
-%\includegraphics[width=.5\textwidth]{mseK.eps}
-%\caption{Influence of the size parameters on the forecasting MSE. Left: MSE in function of the sub-signals length $M$. Right: MSE in function of the sub-signal dataset size $K$.}
-%\label{fig:influence.M}
-%\end{figure}
 
@@
 This highlights the ability of our approach to limit the distortion due the boundary effect and provide a more accurate representations.
-
-On Fig.~\ref{fig:ppg.boundeffred}, we display the SST resulting from {\sf BoundEffRed} strategy, applied to the same portion of PPG than what is used to display Fig.~\ref{fig:ppg}.
-We clearly observe an improvement of the quality of the STT near boundaries.
-Indeed, the blurring visible on the zoom on the right boundary of the SST has almost vanished (see right of Fig.~\ref{fig:ppg.boundeffred}).
-The real-time tracking of the instantaneous frequencies contained the measured signal is therefore largely facilitated.
+Even though the {\sc SigExt} extension yields time-frequency  representations slightly more sensitive to boundary effects that the extensions given by EDMD or GPR, it is the only technique that allows a real-time implementation.
 
@@
 \end{table}
+
+On Fig.~\ref{fig:ppg.boundeffred}, we display the SST resulting from the {\sf BoundEffRed} strategy, applied to the same portion of PPG than what is used to display Fig.~\ref{fig:ppg}.
+We clearly observe an improvement of the quality of the SST near boundaries.
+Indeed, the blurring visible when zooming on the right boundary of the SST has almost vanished (see bottom of Fig.~\ref{fig:ppg.boundeffred}).
+The real-time tracking of the instantaneous frequencies contained in the measured signal is therefore largely facilitated.
+
+\begin{figure}
+\centering
+\includegraphics[width=.48\textwidth]{SSTBoundEffRed.eps}
+\includegraphics[width=.48\textwidth]{zoomSSTBoundEffRed.eps}
+\caption{Result of {\sf BoundEffRed} on the synchrosqueezing transform of a PPG (top) with a zoom on its right boundary (bottom). }
+\label{fig:ppg.boundeffred}
+\end{figure}

### 2020-05-14 TimeFrequencyScaleRep/cwt_JEFAS.m

@@ -1,0 +1,65 @@
+function W = cwt_JEFAS(sig,scales,wav_typ,wav_par)
+% CWT_JEFAS:	Continuous wavelet transform with analytic derivative of Gaussian
+%       wavelet or the sharp wavelet
+% usage:	W = cwt_JEFAS(sig,scales,wav_typ,wav_par)
+%
+% Input:
+%   sig: vector containing the input signal
+%   scales : vector of scales
+%   wav_typ: wavelet type:
+%     wav_typ='sharp': The sharp wavelet
+%     wav_typ='dgauss': analytic derivative of gaussian
+%   wav_par: parameter depending on the wavelet type
+%     if wav_typ='sharp': wav_par = -ln(epsilon)>0 where epsilon=value of \hat{\psi}(Fs/2)
+%     if wav_typ='dgauss': wav_par = number of vanishing moments
+%
+% Output:
+%   W: wavelet transform coefficients matrix (complex valued)
+
+% Copyright (C) 2017 Adrien MEYNARD
+% 
+% This program is free software; you can redistribute it and/or modify it
+% under the terms of the GNU General Public License as published by
+% the Free Software Foundation; either version 3 of the License, or
+% (at your option) any later version.
+% 
+% This program is distributed in the hope that it will be useful,
+% but WITHOUT ANY WARRANTY; without even the implied warranty of
+% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+% GNU General Public License for more details.
+% 
+% You should have received a copy of the GNU General Public License
+% along with this program.  If not, see <http://www.gnu.org/licenses/>.
+% 
+% Author: Adrien MEYNARD
+% Created: 2017-12-19
+
+siglength = length(sig);
+fff = (0:(siglength-1))*2*pi/siglength;
+
+fsig = fft(sig(:)).';
+
+scales = scales(:);
+
+tmp = scales * fff;
+
+% Generate wavelets in the Fourier domain
+switch wav_typ
+    case 'sharp'
+        fpsi = exp(-2*wav_par*((pi./(2*tmp)) + (2*tmp/pi) - 2));
+
+    case 'dgauss'
+        Cst = 4*wav_par/(pi^2); % such that argmax fpsi = pi/2
+        K = (2/pi)^wav_par*exp(wav_par/2); % such that fpsi(pi/2) = 1
[…]

### 2020-05-14 Notes/BoundEffRed.tex
tex

@@
-\documentclass[twoside,letterpaper]{article}
+\documentclass[journal]{IEEEtran}
+
 \usepackage{amsmath,amssymb,amsfonts,amsthm}
@@
 \usepackage[T1]{fontenc}
-\usepackage{ifpdf}
+%\usepackage{ifpdf}
 \usepackage[english]{babel}
-\usepackage{caption}
-\usepackage{subfig} % or can use subcaption package
+%\usepackage{caption}
+\usepackage{subfig}
 \usepackage{color}
@@
 \usepackage{mathpazo}
-\usepackage[left=90pt,right=90pt,top=70pt,bottom=70pt]{geometry}
 
-\title{An approach for the reduction of boundary effects in time-frequency and time-scale representations}
-%\author{}
-\date{}
-
+\graphicspath{{figures/}}
 \input{mydefs}
 
+\title{An Approach for the Reduction of Boundary Effects in Time-Frequency Representations}
+
+\author{Adrien~Meynard, %~\IEEEmembership{Member,~IEEE,}
+        Hau-Tieng~Wu
+\thanks{A. Meynard and H.-T. Wu are with the Department
+of Mathematics, Duke University, Durham,
+NC, 27708 USA e-mail: adrien.meynard@duke.edu}}
+
 \newtheorem{remark}{Remark}
+\newtheorem{lemma}{Lemma}
+\newtheorem{theorem}{Theorem}
 
@@
 
+\begin{abstract}
+Time-frequency representations are intrinsically subject to the boundary effects.
+As a consequence, the structures of signals that are highlighted by the representations are garbled when approaching the boundaries of the time-frequency domain.
+In this paper, we propose an approach for the reduction of such boundary effects, whose objective is to be able to be used in real time.
+That one relies on an extension of the analyzed signal obtained by a forecasting technique.
+In the case of the study of a class of locally oscillating signals, we provide a theoretical study of the performance of our approach.
+Following a numerical verification of the performance of the algorithm, we implement it on real-life biomedical signals.
+\end{abstract}
+
+\begin{IEEEkeywords}
+Boundary effects, time-frequency, forecasting
+\end{IEEEkeywords}
+
 \section{Introduction}
-In any digital acquisition system, the study and the interpretation of the measured signals generally require the use an analysis tool, which enable to point out the useful characteristics of the signal.
-For example, any  measured physiological signal, such as photoplethysmogram (PPG), may not be interpreted as it is, from its run chart.
-An analysis tool would make possible the extraction of some useful characteristics such as heart rate or blood pressure.
+\IEEEPARstart{I}{n} any digital acquisition system, the study and the interpretation of the measured signals generally require the use an analysis tool, which enable to point out the useful characteristics of the signal.
+For example, any  measured biomedical signal, such as photoplethysmogram (PPG), may not be interpreted as it is, from its run chart.
+An analysis tool would make possible the extraction of some useful characteristics such as heart rate or blood pressure~\cite{Akay96detection}.
 
@@
 In order to adapt the analysis to nonstationarities, local analysis is generally performed.
-The short time Fourier transform (STFT) is a typical tool which is build that way, and allows to determine locally the frequency content of a nonstationary signal.
+The short time Fourier transform (STFT) is a typical tool which is build that way, and enables the determination of the local frequency content of a nonstationary signal.
 
 Windowing is a common method for performing local analysis.
-Among others, STFT, wavelet transform~\cite{Mallat99wavelet}, synchrosqueezing transform (SST), and reassignment~\cite{Auger13time} are representations that fall back on the use of an analysis window.
+Among others, STFT, wavelet transform~\cite{Mallat99wavelet}, synchrosqueezing transform (SST), and reassignment~\cite{Auger13time} (RS) are representations that fall back on the use of an analysis window.
 Let $x:I\to\RR$ denote the observed signal, where $I$ denotes the finite interval where the signal is measured.
@@
 A zoom on the right boundary of this SST, displayed on the right of Fig.~\ref{fig:ex.intro}, emphasizes the result of boundary effect.
-More generally, signal characteristics, like instantaneous frequencies or amplitudes appear to be imprecisely determinable in the vicinity of the boundaries.
+More generally, signal characteristics, like instantaneous frequencies or amplitudes appear to be imprecisely determinable from the SST in the vicinity of the boundaries.
 
 \begin{figure}
-\includegraphics[width=.49\textwidth]{SSTintro.eps}
-\includegraphics[width=.49\textwidth]{zoomSSTintro.eps}
-\caption{Synchrosqueezing transform of a PPG (left) with a zoom on its right boundary (right). }
+\centering
+%\includegraphics[width=.48\textwidth]{PPGsig.eps}
+\includegraphics[width=.48\textwidth]{SSTintro.eps}
+\includegraphics[width=.48\textwidth]{zoomSSTintro.eps}
+\caption{Synchrosqueezing transform of a PPG (top) with a zoom on its right boundary (bottom). }
 \label{fig:ex.intro}
@@
 The purpose of this paper is to provide a fast algorithm allowing to tackle boundary effect and limit the disruptions it may cause on the signals representations based on windowing.
-In that aim, we proceed in two steps:
+Attempts to minimize the boundary effects generally consists in softening the discontinuity on signals edges.
+In that goal, a natural idea is to extend to signal beyond its edges.
+Hence, to implement our boundary effects reduction procedure, we proceed in two steps.
 \begin{enumerate}
@@
 Thus, assuming that the quality of the extension step is sufficient, the representation obtained that way will be less sensitive to the boundary effect than the result of the analysis tool applied directly to the non-extended signal.
-Notice that, in order to obtain a fast algorithm the extension step we propose must have a low computational cost.
+
+Various extension schemes have been proposed.
+On one hand, there exists simple extension schemes such as zero-padding, periodic extension, symmetric extension~\cite{Kharitonenko02wavelet}, that do not take into account the dynamical behavior of the signal.
+On the other hand, there exists extension schemes based on a dynamical model, such as the Empirical Dynamic Mode Decomposition~\cite{Williams15data} (EDMD), the Gaussian process regression~\cite{Rasmussen06gaussian} (GPR), the TBATS algorithm~\cite{DeLivera11forecasting}, or interpolation algorithm~\cite{Su13adaptive}.
+While the second class of extension schemes give better extended signals than the first class, they generally have a great computational cost.
+We propose here a fast extension algorithm based on a simple dynamical model, in order to optimize the trade-off between the extension quality and the computational cost.
+Furthermore, we prove that the dynamical model we consider is sufficient to forecast locally oscillating signals.
+
+Therefore, in section~\ref{se:algo}, we provide an extension method based on a linear dynamic model.
+We derive the corresponding algorithm for boundary effect reduction.
+In section~\ref{se:theoretical}, we show that, even though we consider a simplistic dynamic model, it is sufficient to extend signals takings the form of sums of sine waves.
+An evaluation of the theoretical performance of our algorithm on a class of signals, the sums of sine waves, is given in section~\ref{se:theoretical}.
+In section~\ref{se:results}, we compare our extension method with more sophisticated methods such as EDMD, GPR, or TBATS.
+We show that our algorithm gives fast results of reasonable quality.
+Finally, we evaluate the performance of our boundary effect reduction algorithm on biomedical signals such as respiratory signals, and compare it to the theoretical results.
 
@@
 \section{Numerical results}
+\label{se:results}
 \input{results}
 
-\bibliographystyle{plain}
+{\color{red}
+\section{Conclusion}
+\label{se:conclusion}
+The MATLAB code and datasets used to produce the numerical results of this paper are available online\footnote{\url{https://github.com/AdMeynard/BoundaryEffectsReduction}}.
+}
+
+%\appendices
+%\input{appendices}
+
+%\section*{Acknowledgment}
+
+\bibliographystyle{IEEEtran}
 \bibliography{AM041320}
 
+%\begin{IEEEbiography}{Michael Shell}
+%Biography text here.
+%\end{IEEEbiography}
+%
+%\begin{IEEEbiographynophoto}{John Doe}
+%Biography text here.
+%\end{IEEEbiographynophoto}
+
 \end{document}

### 2020-05-27 Notes/Supplementary material/SupMat.tex
update code

@@
 \usepackage{etoolbox}
+\usepackage[shortlabels]{enumitem}
 \usepackage{multirow}
+\usepackage{dsfont}
 \usepackage{hhline}
 \usepackage{algorithm,algorithmic}
+\usepackage{xr}
+\externaldocument[I-]{../BoundEffRed}
 
@@
 \label{ap:lm.error}
+
+\subsection{Notations}
 We have:
 \begin{align}
-\dfrac1K\bX\bX^T &= \underbrace{\dfrac1K\bZ\bZ'+\sigma^2\bI}_{\defeq S^{(0)}} + \bE^{(0)} \\
-\dfrac1K\bY\bX^T &= \underbrace{\dfrac1K\bZ\bZ^T +\sigma^2\bD}_{\defeq S^{(1)}}+ \bE^{(1)} \ ,
+\dfrac1K\bX\bX^T &= \underbrace{\dfrac1K\bZ\bZ^T+\sigma^2\bI}_{\defeq S^{(0)}} + \bE^{(0)} \\
+\dfrac1K\bY\bX^T &= \underbrace{\dfrac1K\bZ'\bZ^T +\sigma^2\bD}_{\defeq S^{(1)}}+ \bE^{(1)} \ ,
 \end{align}
@@
 with $a\in\{0,1\}$.
+
 \begin{remark}
@@
 \end{remark}
-Thus:
+
+It follows from this:
 \begin{align*}
 \tilde\bA &= (\bS^{(1)}+\bE^{(1)})(\bS^{(0)}+\bE^{(0)})^\inv \\
-\bA_0&=\bS^{(1)}{\bS^{(0)}}^\inv
+\bA_0 &=\bS^{(1)}{\bS^{(0)}}^\inv\ .
 \end{align*}
@@
 \end{align}
-Thus, the randomness of $\bh^{(\ell)}$ is completely originating from the random vector $\bg\in\RR^{M(M+1)}$ defined as:
+
+\subsection{Study of $\bh^{(\ell)}$}
+The randomness of $\bh^{(\ell)}$ is completely originating from the error matrices.
+Besides, notice that the first $M-1$ rows in $\bE^{(1)}$ are the same as the last $M-1$ rows of $\bE^{(0)}$.
+Thus, we gather all the sources of randomness into an vector $\bg\in\RR^{M(M+1)}$ defined as:
 \[
@@
 \]
-{\color{blue}
-Thus, one can write $\bh^{\ell}$ as $\bh^{\ell}=f^{(\ell)}(g)$ where $f^{(\ell)}$ is a deterministic function:
+Here, "$\vec$" denotes the vectorization operator, concatenating the columns of a given matrix on top of one another. Then, by definition, we have:
+\[
+\bg = \sigma\bg_1 + \sigma^2\bg_2\ ,
+\]
+where:
+\begin{align*}
+\bg_1 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left( \tilde\bz_k\bw_k^T + \tilde\bw_k\bz_k^T \right) \\
+\bg_2 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left(\tilde\bw_k\bw_k^T-\tilde\bI \right) \ .
+\end{align*}
+with $\tilde\bz_k^T= \begin{pmatrix} \bz_k^T & \bz_{k+1}[0] \end{pmatrix}$ and $\tilde\bw_k^T= \begin{pmatrix} \bw_k^T & \bw_{k+1}[0] \end{pmatrix}$.
+Then, $\bg_1$ is a Gaussian random vector as a linear combination of Gaussian random vectors.
+Moreover, using the central limit theorem under weak dependence, we can show that $\bg_2$ also converges towards a Gaussian random vector as $K\to\infty$.
+Combining these two results gives the following result:
+\[
+\sqrt{K}\ \bg \xrightarrow[K\to\infty]{\cD} \cN(\bzero,\bGamma_0)\ ,
+\]
+where $\bGamma_0=\EE\left\{\bg\bg^T\right\}$ is a covariance matrix.
+
+Furthermore, one can write $\bh^{(\ell)}$ as $\bh^{(\ell)}=f^{(\ell)}(\bg)$ where $f^{(\ell)}$ is a deterministic function such that:
 \begin{align*}
@@
 \end{align*}
-Besides, by application of the central limit theorem, the random vector $\bg$ converges in law to a Gaussian random vector:
-\[
-\sqrt{K}\ \bg \xrightarrow[K\to\infty]{\cD} \cN(\bzero,\bGamma_0)\ .
-\]
-Then, using Delta method, we have:
+Then, as $f^{(\ell)}$ is a differentiable function, using the Delta method gives:
 \begin{align*}
@@
 \end{align*}
-where:
+where $\bF^{(\ell)}$ is the Jacobian matrix such that:
 \[
@@
 \]
-}
 
@@
 
-\paragraph{Expression of the bias $\bmu$.}
+\subsection{Expression of the bias $\bmu$.}
 Clearly, $\bmu[n]=0$ when $n\in I$. When $n=N-1+\ell$, we have:
@@
 \end{equation*}
-{\color{blue}
 Besides, from Lemma~1, we have the following results:
 \begin{align*}
-\EE\{\bh^{(\ell)}\} &\underset{K\to\infty}{=} o\left(\dfrac1{\sqrt{K}}\right)\\
-\EE\{\bh^{(\ell)}\bw_{K}\} &\underset{K\to\infty}{=} o\left(\dfrac1{\sqrt{K}}\right)
+\EE\{\bh^{(\ell)}\} &\underset{K\to\infty}{\longrightarrow} 0\\
+\EE\{\bh^{(\ell)}\bw_{K}\} &\underset{K\to\infty}{\longrightarrow} 0
 \end{align*}
@@
 \bmu[N-1+\ell] \underset{K\to\infty}{\sim} o(\sigma)\ .
+\label{eq:bias.ap}
 \end{equation*}
 
-\paragraph{Expression of the covariance $\bgamma$.}
-Let us segregate the cases.
+\subsection{Expression of the covariance $\bgamma$.}
+We recall that $n>N$, we consequently denote $n=N-1+\ell$.  To derive the covariance, let us segregate the cases.
 
-First, when $(n,n')\in I^2$, we clearly have $\bgamma[n,n']=\sigma^2\delta_{n,m}$.
+\paragraph{If $n'\in I^2$}
+From Lemma~1, we have that $h^{(\ell)}$ is asymptotically Gaussian.
+Then, as a direct consequence of the Isserlis' theorem, odd-order moments are vanishing.
+Then, combining result~\eqref{eq:bias.ap} and equation~\eqref{eq:bias0}, we obtain:
+\begin{align*}
+\bgamma[n,n'] &= \sigma\EE\{\bw[n']\bh^{(\ell)}\}\bz_K + \sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bw[n']\} + o(\sigma^2) \ .
+\end{align*}
+We remark that
+\[
+\balpha_0^{(\ell)}\EE\{\bw_K\bw[n']\}=\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)}\ .
+\]
+Moreover, using Delta method one can show that we have the asymptotic result:
+\begin{equation}
+\EE\{\bw[n']\bh^{(\ell)}\} \underset{K\to\infty}{\longrightarrow} 0
+\label{eq:cov.ap}
+\end{equation}
+Finally:
+\begin{align*}
+\bgamma[n,n'] &\underset{K\to\infty}{\sim} \sigma^2\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)} + o(\sigma^2) \ .
+\end{align*}
 
-Second, we focus on the case where $n=N-1+\ell\geq N$ and $n'=N-1+\lambda\geq N$. Thus:
+\paragraph{If $n'\geq N$}
+From Lemma~1, we have that $h^{(\ell)}$ is asymptotically Gaussian.
+Then, as a direct consequence of the Isserlis' theorem, odd-order moments are vanishing.
+Then, combining result~\eqref{eq:bias.ap} and equations~\eqref{eq:cov0} and~\eqref{eq:cov.ap}, we obtain:
 \begin{align*}  
 \bgamma[n,n'] &= \bz_K^T\EE\left\{{\balpha^{(\ell)}}^T\balpha^{(\lambda)}\right\}\bz_K + \sigma\EE\{\balpha^{(\ell)}\bw_K\balpha^{(\lambda)}\}\bz_K+ \sigma\EE\{\balpha^{(\lambda)}\bw_K\balpha^{(\ell)}\}\bz_K \\
-& +\sigma^2\EE\{\balpha^{(\ell)}\bw_K\balpha^{(\lambda)}\bw_K\}-\bz[n]\bz[n']-\bz[n]\bmu[n']-\bz[n]\bmu[n'] - \bmu[n]\bmu[n'] \\
-&= \bz_K^T\EE\left\{{\bh^{(\ell)}}^T\bh^{(\lambda)}\right\}\bz_K + \balpha_0^{(\ell)}\sigma\EE\{\bw_K\bh^{(\lambda)}\}\bz_K + \sigma\EE\{\bh^{(\ell)}\bw_K\}\bz(n']  \\
+& +\sigma^2\EE\{\balpha^{(\ell)}\bw_K\balpha^{(\lambda)}\bw_K\}-\bz[n]\bz[n']+o(\sigma^2) \\
+&= \bz_K^T\EE\left\{{\bh^{(\ell)}}^T\bh^{(\lambda)}\right\}\bz_K + \balpha_0^{(\ell)}\sigma\EE\{\bw_K\bh^{(\lambda)}\}\bz_K + \sigma\EE\{\bh^{(\ell)}\bw_K\}\bz[n']  \\
 &+ \sigma\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\}\bz_K+ \sigma\EE\{\bh^{(\lambda)}\bw_K\}\bz[n] + \sigma\balpha_0^{(\lambda)}\EE\{\bw_K\bh^{(\ell)}\}\bz_K \\
 & + \sigma\EE\{\bh^{(\lambda)}\bw_K\bh^{(\ell)}\}\bz_K+\sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\lambda)}\bw_K\}+\sigma^2\balpha_0^{(\lambda)}\EE\{\bw_K\bh^{(\ell)}\bw_K\} \\
-&+\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\ell)}\right\rangle +\sigma^2\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} + o(\sigma^2)
+&+\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\sigma^2\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} + o(\sigma^2) \\
+&\underset{K\to\infty}{\sim} \dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K +\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\sigma^2\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} + o(\sigma^2)
 \end{align*}
-From lemma~1, we have:
-To be continued...
+The Isserlis' theorem also gives the following result:
+\begin{align*}
+\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} &= \sum_{m,m'=0}^{M-1}\EE\{\bh^{(\ell)}[m]\bw_K[m]\bh^{(\lambda)}[m']\bw_K[m']\} \\
+&= \sum_{m,m'=0}^{M-1}\EE\{\bh^{(\ell)}[m]\bw_K[m]\}\EE\{\bh^{(\lambda)}[m']\bw_K[m']\} + \EE\{\bh^{(\ell)}[m]\bw_K[m']\}\EE\{\bh^{(\lambda)}[m']\bw_K[m]\} \\
+&\hspace{35pt} + \EE\{\bh^{(\ell)}[m]\bh^{(\lambda)}[m']\}\underbrace{\EE\{\bw_K[m]\bw_K[m']\}}_{=\delta_{m,m'}} \\
+&\underset{K\to\infty}{\sim} \sum_{m=0}^{M-1} \EE\{\bh^{(\ell)}[m]\bh^{(\lambda)}[m]\} = \dfrac1K\Tr\left(\bGamma^{(\ell,\lambda)}\right)\ .
+\end{align*}
+We conclude that:
+\begin{align*}  
+\bgamma[n,n'] &\underset{K\to\infty}{\sim} \dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K +\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\dfrac{\sigma^2}{K}\Tr\left(\bGamma^{(\ell,\lambda)}\right) + o(\sigma^2)\ .
+\end{align*}
 
-Third, we focus on the case where $n=N-1+\ell\geq N$ and $n'\in I$. The case $n\in I$ and $n'=N-1+\lambda$ is directly derived from the current one.
-To be continued...
 %\begin{align*}
@@
 %\end{align*}
-}
 

### 2020-05-27 Notes/results.tex
update code

@@
 
-\paragraph{Conclusion}
+\paragraph{Summary}
 Both previous experimental results combined with the theoretical asymptotic equation~\eqref{eq:cov.error.2} allow us to describe the influence of the noise variance and the size of the training dataset on the variance of the forecasting noise, which is empirically summarized as follows:
@@
 \centering
-\caption{Performance of the extension methods.}
+\caption{AHM signal. Performance of the extension methods.}
 \begin{tabular}{|c||c|c|c|}
@@
 
-In the aim of comparing the different techniques, we uses a criterion, proposed in~\cite{Daubechies16conceft}, that quantify the distance between a given time-frequency representation and the optimal one.
+In the aim of comparing the different techniques, we use a criterion, proposed in~\cite{Daubechies16conceft}, that quantify the distance between a given time-frequency representation and the optimal one.
 It is built in analogy with the optimal transport distance, which enables quantifying the distance between two probability density functions.
@@
 \centering
-\caption{MSE of the respiratory signal extensions.}
+\caption{Respiratory signal. MSE of the respiratory signal extensions.}
 \begin{tabular}{|c||c|c|}
@@
 
-Then, we apply the boundary effects reduction method {\sf BoundEffRed} (Algorithm~\ref{alg:boundary}) on these extensions for diverse time-frequency representations (STFT, SST and RS).
+Then, we apply the boundary effects reduction method {\sf BoundEffRed} (Algorithm~\ref{alg:boundary}) on these extensions for diverse time-frequency representations: STFT, SST, RS, as well as ConceFT, an SST-based representation introduced in~\cite{Daubechies16conceft}.
 In Table~\ref{tab:otd.tho}, the results are compared in terms of OTD.
@@
 \centering
-\caption{Respiratory signal: OTD of the boundary effects reduction methods on diverse TFR.}
-\begin{tabular}{|c||c|c|c|}
+\caption{Respiratory signal. Averaged OTD of the boundary effects reduction methods on diverse representations.}
+\begin{tabular}{|>{\centering}m{35pt}||c|c|c|c|}
   \hline
-   \multirow{2}{*}{Extension method} & \multicolumn{3}{c|}{Time-Frequency Representation} \\
-   \cline{2-4}
-      & STFT & SST & RS\\
-   \hhline{|=#=|=|=|}
-   Without extension & $2.16\times 10^{-2}$ & $5.26\times 10^{-3}$ & $3.07\times 10^{-2}$ \\
+   \multirow{2}{35pt}{\centering Extension method} & \multicolumn{4}{c|}{Time-Frequency Representation} \\
+   \cline{2-5}
+      & STFT & SST & RS & ConceFT \\
+   \hhline{|=#=|=|=|=|}
+   Without extension & $2.16\times 10^{-2}$ & $5.26\times 10^{-3}$ & $3.07\times 10^{-2}$ & $1.41\times 10^{-2}$ \\
    \hline
-   {\sf SigExt} & $1.72\times 10^{-2}$ & $4.00\times 10^{-3}$ & $2.43\times 10^{-2}$ \\
+   {\sf SigExt} & $1.72\times 10^{-2}$ & $4.00\times 10^{-3}$ & $2.43\times 10^{-2}$ & $1.12\times 10^{-2}$ \\
    \hline
-   EDMD & $1.75\times 10^{-2}$ & $4.23\times 10^{-3}$ & $2.45\times 10^{-2}$ \\
+   EDMD & $1.75\times 10^{-2}$ & $4.23\times 10^{-3}$ & $2.45\times 10^{-2}$ & {\color{red}X} \\
    \hline
-   GPR & $1.84\times 10^{-2}$ & $4.10\times 10^{-3}$ & $2.48\times 10^{-2}$ \\
+   GPR & $1.84\times 10^{-2}$ & $4.10\times 10^{-3}$ & $2.48\times 10^{-2}$ & {\color{red}X} \\
    \hline
@@
 \centering
-\caption{Performance of the Boundary effect reduction on TFR.}
+\caption{PPG signal. Averaged OTD of the boundary effects reduction methods on diverse representations.}
 \begin{tabular}{|c||c|c|c|}
@@
 \end{figure}
+
+To evaluate the influence of the noise level on the performance of {\sf BoundEffRed}, we artificially add a Gaussian noise to the measured PPG signal, whose variance is denoted $\sigma_{\mathrm{add}}^2$.
+It is thus an additional noise to the measurement noise actually contained in the signal.
+Figure~\ref{fig:otd.noise} shows the averaged OTD of {\sf BoundEffRed} for different values of $\sigma_{\mathrm{add}}^2$.
+At first glance, we notice that STFT is slightly more sensitive to noise than SST.
+On one hand, the robustness to noise of the SST is the direct consequence of the approximation result of~\cite{Daubechies16conceft}, discussed in section~\ref{sse:perf.BoundEffRed}.
+On the other hand, the sentivity to noise of the STFT may be interpreted as the inability to forecast the values of the time-frequency coefficients where only noise is active, while sharp time-frequency representations, such as SST, ideally vanish at these points.
+\begin{figure}
+\centering
+\includegraphics[width=.48\textwidth]{OTDBoundEffRed.eps}
+\caption{PPG signal. Averaged OTD of {\sf BoundEffRed} in function of the noise variance.}
+\label{fig:otd.noise}
+\end{figure}

### 2020-05-27 Scripts/EvalPerfBoundEffRed/EvalECG.m

@@ -12,20 +12,18 @@ fs = 200 ; % sampling frequency
 N = 15e3; % subsignals length
 
 % Forecasting parameters
+methods = {'lse','edmd','gpr'} ;
 HOP = 1 ;
 extSEC = 6 ; % the extension is of extSEC second
 L = round(extSEC*fs) ;
 extM = round( 1.5*L ) ; % dimension of embedding / signals length
 extK = round( 2.5*extM ) ;  % number of points to estimate A / size of datasets
 
-basicTF.meth = 'sstSTFT' ;
+basicTF.meth = {'sstSTFT','RS'} ;
 basicTF.hop = 10 ;
 basicTF.win = 2301 ;
 basicTF.fmin = 0 ;
 basicTF.fmax = 0.01 ;
-[forecastErr1, CompTime1, OTD1] = SucForecast(xtot,fs,HOP,N,extM,extK,extSEC,basicTF) ;
-basicTF.meth = 'RS' ;
-[forecastErr2, CompTime2, OTD2] = SucForecast(xtot,fs,HOP,N,extM,extK,extSEC,basicTF) ;
+[forecastErr, CompTime, OTD] = SucForecast(xtot,fs,methods,HOP,N,extM,extK,extSEC,TFR,basicTF) ;
 
-
-save('../../Results/resultSucForECGcompMeth','forecastErr1','CompTime1','OTD1','forecastErr2','CompTime2','OTD2');
+save('../../Results/resultSucForECGcompMeth','forecastErr','CompTime','OTD');

### 2020-05-27 Scripts/EvalPerfBoundEffRed/ResultsBoundEffRed.m

@@ -27,28 +27,29 @@ fprintf('|__________________|___________|___________|___________|\n\n')
 
 %% THO
 load ../../Results/resultSucForTHOcompMeth ;
-fprintf('================================================================\n')
-fprintf('                                THO                             \n')
-fprintf('================================================================\n')
+fprintf('====================================================================\n')
+fprintf('                              THO                                  \n')
+fprintf('====================================================================\n')
 
-fprintf('===================Forecasting Performance======================\n')
-fprintf(' ______________________________________________________________\n')
-fprintf('| Extension Method | Computing time (sec.) | Forecasting Error |\n')
-fprintf('|------------------|-----------------------|-------------------|\n')
-fprintf('|      LSE         |           %.2f        |     %.3e     |\n', CompTime1.LSE , mean(forecastErr1.LSE))
-fprintf('|      EDMD        |           %.2f        |     %.3e     |\n', CompTime1.EDMD , mean(forecastErr1.EDMD))
-fprintf('|      GPR         |         %.2f        |     %.3e     |\n', CompTime1.GPR , mean(forecastErr1.GPR))
-fprintf('|__________________|_______________________|___________________|\n\n')
+fprintf('===========================Forecasting Performance==================\n')
+fprintf(' __________________________________________________________________\n')
+fprintf('| Extension Method | Computing time (sec.) |   Forecasting Error   |\n')
+fprintf('|                  |                       |    Mean   |   Median  |\n')
+fprintf('|------------------|-----------------------|-----------|-----------|\n')
+fprintf('|      LSE         |           %.2f        | %.3e | %.3e |\n', CompTime1.LSE , mean(forecastErr1.LSE(~isnan(OTD1.sst.LSE))),median(forecastErr1.LSE(~isnan(OTD1.sst.LSE))) )
[…]

### 2020-05-27 Scripts/EvalPerfForecasting/CompPerfAHM.m

@@ -9,7 +9,7 @@ t = linspace(0,1,N);
 
 % forecasting parameters
 HOP = 1 ;
-extSEC = 0.05 ; % the extension is of extSEC second
+extSEC = 0.1 ; % the extension is of extSEC second
 L = round( extSEC*fs ) ;
 extM = round( 1.5*L ) ; % dimension of embedding / signals length
 extK = round( 2.5*extM );  % number of points to estimate A / size of datasets
@@ -18,19 +18,33 @@ tt = linspace(-L/fs, 1+L/fs, N+2*L) ;
 
 %% Synthesize signal
 
-p0 = 25 ;
-phi00 = p0*fs/extM * ( tt + (0.3/(2*pi))*cos(2*pi*tt) ) ;
+tmp = cumsum(randn(N+2*L,1)) ;
+tmp = smooth(tmp,0.1,'rloess') ;
+tmp = tmp - mean(tmp) ;
+tmp = tmp/max(abs(tmp));
+
+p0 = 20 ;
+freq00 = p0*fs/extM * (1 + 0.0*tmp) ;
+% phi00 = p0*fs/extM * ( tt + (0.01/(2*pi))*cos(2*pi*tt) ) ;
+phi00 = (1/fs) * cumsum(freq00).' ;
 xx00 = cos(2*pi*phi00) ;
 
-p1 = 60 ;
-R = 1.4 + 0.2*cos(4*pi*tt) ;
-phi01 = (p1*fs/extM) * tt + (250/2)*tt.^2 ;
+
+% phi01 = (p1*fs/extM) * tt + (20/2)*tt.^2 ;
 
 nComp= 2;
 switch nComp
     case 1
         xx01 = 0 ;
     case 2
+        R = 1.4 + 0.2*cos(4*pi*tt) ;
+        p1 = 35 ;
+        tmp = cumsum(randn(N+2*L,1)) ;
+        tmp = smooth(tmp,0.25,'rloess') ;
+        tmp = tmp - mean(tmp) ;
+        tmp = tmp/max(abs(tmp));
+        freq01 = p1*fs/extM * (1 + 0.01*tmp) ;
+        phi01 = (1/fs) * cumsum(freq01).' ;
         xx01 = R.*cos(2*pi*phi01) ;
 end
 xx0 = xx00 + xx01 ; % extended signal
@@ -81,5 +95,17 @@ BiasXP.GPR = mean(MeanGPR) ;
[…]

### 2020-05-27 Scripts/EvalPerfForecasting/InfluenceSizeDatasetSine.m

@@ -9,10 +9,9 @@ t = linspace(0,1,N);
 
 % forecasting parameters
 HOP = 1 ;
-extSEC = 0.05 ; % the extension is of extSEC second
+extSEC = 0.01 ; % the extension is of extSEC second
 L = round( extSEC*fs ) ;
 extM = round( 1.5*L ) ; % dimension of embedding / signals length
-extK = round( 2.5*extM );  % number of points to estimate A / size of datasets
 
 tt = linspace(-L/fs, 1+L/fs, N+2*L) ;
 
@@ -36,26 +35,68 @@ end
 xx0 = xx00 + xx01 ; % extended signal
 x0 = xx0( (L+1) : (L+N) ) ; % restriction to the measurement interval
 
-sigman = 1e-2 ;
+zK = x0(end-extM+1 : HOP: end).' ;
 
 %% Forecasting
-method.name = 'lseV' ;
-nbXP = 50 ;
+sigman = 1e-2 ;
+
+method.name = 'lse' ;
+nbXP = 3 ;
 nbXPP = 3000 ;
-KK = round(linspace(1e3,6e3,nbXP)) ;
+KK = round(logspace(log10(4.5e2),log10(2e3),nbXP)) ;
+Covh = zeros(extM,extM,L) ;
+Ehw2 = zeros(L,1);
 k = 1 ;
 for k = 1:nbXP
     extK = KK(k) ;
+    
+    X0 = [] ; Y0 = [] ;
+    for kk = 1: extK
+        X0(:,kk) = x0(end-extK-extM+kk: HOP: end-extK+kk-1) ;
+        Y0(:,kk) = x0(end-extK-extM+kk+1: HOP: end-extK+kk) ;
+    end
+    S0 = (1/extK)*(X0*X0') + sigman^2*eye(extM) ;
+    r = zeros(1,extM) ; r(2) = 1 ;
+    S1 = (1/extK)*(Y0*X0') + sigman^2*toeplitz(zeros(extM,1),r) ;
+    A0 = S1 / S0 ; % ideal least square estimation
+    for l = 1:L
+        tmp0 = A0^l ;
+        alpha0l(l,:) = tmp0(end,:) ;
+    end
+    a2 = sum(alpha0l.^2, 2) ;
+    
     for nb2 = 1:nbXPP
         noise = sigman*randn(N+2*L,1) ;
+        wK = noise((N+L-extM+1):(N+L))/sigman ;
         x = x0.' + noise((L+1):(N+L)) ; % signal to be extended
[…]

### 2020-05-27 Scripts/EvalPerfForecasting/ResultsInfluenceNoiseSine.m

@@ -5,21 +5,22 @@ L = size(VarXPm,2) ;
 nbXP = length (Sigma) ;
 
 %% Plot
-tmp = 2/extM ;
 ll = 1:L ;
-
 
 for k = 1:nbXP
     sigma = Sigma(k) ;
+    txtXP = ['Experimental \sigma = ',num2str(sigma,'%.0e')];
+    txtTH = ['Theoretical    \sigma = ',num2str(sigma,'%.0e')];
     figure(1);
-    plot(ll,MeanXPm(k,:),'linewidth',2); grid on; hold on;
+    plot(ll,MeanXPm(k,:)/sigma,'linewidth',2,'DisplayName',txtXP); grid on; hold on;
     figure(2);
-    semilogy(ll,VarXPm(k,:)/sigma^2,'linewidth',2); grid on; hold on; %,ll,BndVarTH,'--',
+    fig = semilogy(ll,VarXPm(k,:),'+','linewidth',2,'MarkerSize',8,'DisplayName',txtXP); grid on; hold on;
+    semilogy(ll,VarTH(k,:),'-','color',fig.Color,'linewidth',2,'DisplayName',txtTH); hold on;
 end
 %legend('Experimental MSE[10]','Experimental MSE[100]','Experimemtal MSE[200]','Experimental MSE[500]','\sigma^2');
-figure(1); xlabel('Forecasting sample index ($\ell$)','interpreter', 'latex','fontsize',15); ylabel('Experimental forecasting bias ($\mu_{xp}[\ell]$)','interpreter','latex','fontsize',15);
-ylim([-0.05 0.05]);
-figure(2); xlabel('Forecasting sample index ($\ell$)','interpreter', 'latex','fontsize',15); ylabel('Normalized experimental forecasting variance ($\gamma_{xp}[\ell]/\sigma^2$)','interpreter','latex','fontsize',15);
-ylim([0.2 1]);
-% BndVarTH = 2*tmp*(1+tmp).^(2*ll-2) ;
-% semilogy(ll,BndVarTH,'--','linewidth',2);
+figure(1); xlabel('Forecasting sample index ($\ell$)','interpreter', 'latex','fontsize',28); ylabel('Normalized experimental forecasting bias ($\mu_{xp}[N-1+\ell]/sigma$)','interpreter','latex','fontsize',28);
+legend show ; %ylim([-0.05 0.05]);
[…]

### 2020-05-28 Notes/BoundEffRed.tex
update code
new files

@@
 \usepackage[T1]{fontenc}
-%\usepackage{ifpdf}
 \usepackage[english]{babel}
-%\usepackage{caption}
 \usepackage{subfig}
@@
 \usepackage{mathpazo}
+\usepackage{array}
 
@@
 
-{\color{red}
 \section{Conclusion}
 \label{se:conclusion}
-The MATLAB code and datasets used to produce the numerical results of this paper are available online\footnote{\url{https://github.com/AdMeynard/BoundaryEffectsReduction}}.
-}
+In this paper, we propose an algorithm, named {\sf BoundEffRed}, for the boundary effects reduction on time-frequency representations.
+This method is based on an extension of the signal obtained by forecasting.
+We have shown theoretically that the chosen dynamic model is sufficient for the extension of signals formed by a sum of sine waves whose instantaneous frequencies and amplitudes vary slowly.
+Moreover, the numerical results show that this strategy allows us to switch to a real-time implementation of {\sf BoundEffRed} is possible, due to the low computation time, unlike other pre-existing forecasting methods.
+The numerical results also confirmed the robustness to noise of {\sf BoundEffRed}, as well as its ability to be applied to many time-frequency representations.
+The MATLAB code and datasets used to produce the numerical results of this paper are available online at the following address:
+\begin{center}
+    {\small\url{https://github.com/AdMeynard/BoundaryEffectsReduction}}
+\end{center}
+Other numerical examples are also included, such as the application of {\sf BoundEffRed} to other types of biomedical signals such as ECGs.
+
+Various improvements can be considered to make the algorithm more robust to certain situations.
+In particular, we have noticed that when the observed signal does not oscillate, the forecasting is no longer relevant, and outlier time-frequency representations are obtained.
+A preliminary step should then be added to the algorithm to select the portions of the signal that exhibit oscillations, in order to reject the portions for which accurate time-frequency analysis is practically useless.
+A more fundamental improvement would be to perform the prediction step directly in the time-frequency plane, with the aim of specifically predicting the evolution of each instantaneous frequency.
+This would lead to seeing a time-frequency representation as a set of interdependent time series (corresponding to each frequency channel).
 

### 2020-05-28 Scripts/EvalPerfForecasting/ResultsInfluenceSizeDataSetSine.m

@@ -1,0 +1,24 @@
+clear all; close all;
+
+load('../../Results/PerfSizeDataset.mat')
+L = size(VarXPm,2) ;
+nbXP = length (KK) ;
+
+%% Plot
+ll = 1:L ;
+
+for k = 1:nbXP % k must be sufficiently large to observe the asymptotic behavior
+    txtXP = ['Experimental K = ',num2str(KK(k),'%.0f')];
+    txtTH = ['Theoretical    K = ',num2str(KK(k),'%.0f')];
+    figure(1);
+    plot(ll,MeanXPm(k,:),'linewidth',2,'DisplayName',txtXP); grid on; hold on;
+    figure(2);
+    fig = semilogy(ll,VarXPm(k,:),'+','linewidth',2,'DisplayName',txtXP); grid on; hold on;
+    semilogy(ll,VarTH(k,:),'-','color',fig.Color,'linewidth',2,'DisplayName',txtTH); hold on;
+end
+figure(1); xlabel('Forecasting sample index ($\ell$)','interpreter', 'latex','fontsize',28); ylabel('Experimental forecasting bias ($\mu_{xp}[\ell]$)','interpreter','latex','fontsize',28);
+ylim([-0.01 0.01]); legend show ;
+figure(2); xlabel('Forecasting sample index ($\ell$)','interpreter', 'latex','fontsize',28); ylabel('Forecasting variance','interpreter','latex','fontsize',28);
+lgd = legend('show') ;
+lgd.FontSize = 18 ;
+set(gca,'fontsize',24) ;

### 2020-05-28 Notes/BoundEffRed.tex
paper

@@
 \usepackage[english]{babel}
-\usepackage{subfig}
+%\usepackage{subfig}
 \usepackage{color}
@@
 \section{Introduction}
+\label{se:introduction}
 \IEEEPARstart{I}{n} any digital acquisition system, the study and the interpretation of the measured signals generally require the use an analysis tool, which enable to point out the useful characteristics of the signal.
-For example, any  measured biomedical signal, such as photoplethysmogram (PPG), may not be interpreted as it is, from its run chart.
-An analysis tool would make possible the extraction of some useful characteristics such as heart rate or blood pressure~\cite{Akay96detection}.
+This problem arises for the analysis of signals as varied as audio~\cite{Stowell18computational,Muller11signal}, mechanical~\cite{Peng02vibration} or biomedical signals~\cite{Akay96detection}.
+For instance, any  measured biomedical signal, such as photoplethysmogram (PPG), may not be interpreted as it is, from its run chart.
+An analysis tool would make possible the extraction of some useful characteristics such as heart rate or blood pressure.
 
@@
 As a consequence, the measured signals exhibit nonstationary behavior.
-In order to adapt the analysis to nonstationarities, local analysis is generally performed.
-The short time Fourier transform (STFT) is a typical tool which is build that way, and enables the determination of the local frequency content of a nonstationary signal.
+In order to adapt the analysis to nonstationarities, local spectral analysis is generally performed~\cite{Stoica05spectral,Matz97generalized}.
+The short time Fourier transform~\cite{Grochenig01foundations} (STFT) is a typical tool which is build that way, and enables the determination of the local frequency content of a nonstationary signal.
 
@@
 We display, on the left of Fig.~\ref{fig:ex.intro} the result of the SST of a PPG (see section~\ref{ssse:ppg} for a comprehensive description).
-The distortion resulting from the boundary effect is clearly visible on both sides of this representation.
+The distortion resulting from boundary effects is clearly visible on both sides of this representation.
 Indeed, while in the major central part of the image, clear lines stand out, they become blurred as they approach the left and right boundaries of the image.
-A zoom on the right boundary of this SST, displayed on the right of Fig.~\ref{fig:ex.intro}, emphasizes the result of boundary effect.
-More generally, signal characteristics, like instantaneous frequencies or amplitudes appear to be imprecisely determinable from the SST in the vicinity of the boundaries.
+A zoom on the right boundary of this SST, displayed on the right of Fig.~\ref{fig:ex.intro}, emphasizes the result of boundary effects.
+More generally, estimations of signal characteristics, like instantaneous frequencies~\cite{Delprat92asymptotic} or amplitudes, from its time-frequency representation appear to be imprecisely determinable (or even likely to fail) in the vicinity of the boundaries.
 
@@
 
-The purpose of this paper is to provide a fast algorithm allowing to tackle boundary effect and limit the disruptions it may cause on the signals representations based on windowing.
+The purpose of this paper is to provide a fast algorithm allowing to tackle boundary effects and limit the disruptions it may cause on the signals representations based on windowing.
 Attempts to minimize the boundary effects generally consists in softening the discontinuity on signals edges.
-In that goal, a natural idea is to extend to signal beyond its edges.
+For instance, judicious choices of analysis windows whose support does not interfere with the boundary  points can minimize the occurrence of aberrant patterns near the boundaries of the time-frequency plane~\cite{Chui92wavelets,Depczynski99fast}.
+These techniques therefore do not make it possible to reduce the boundary effects of any time-frequency representations, and for any analysis window.
+For this purpose, a natural idea consists in carrying out a preliminary step to the analysis allowing to extend the signal beyond its boundaries.
 Hence, to implement our boundary effects reduction procedure, we proceed in two steps.
@@
 \end{enumerate}
-Thus, assuming that the quality of the extension step is sufficient, the representation obtained that way will be less sensitive to the boundary effect than the result of the analysis tool applied directly to the non-extended signal.
+Thus, assuming that the quality of the extension step is sufficient, the representation obtained that way will be less sensitive to the boundary effects than the result of the analysis tool applied directly to the non-extended signal.
 
 Various extension schemes have been proposed.
-On one hand, there exists simple extension schemes such as zero-padding, periodic extension, symmetric extension~\cite{Kharitonenko02wavelet}, that do not take into account the dynamical behavior of the signal.
-On the other hand, there exists extension schemes based on a dynamical model, such as the Empirical Dynamic Mode Decomposition~\cite{Williams15data} (EDMD), the Gaussian process regression~\cite{Rasmussen06gaussian} (GPR), the TBATS algorithm~\cite{DeLivera11forecasting}, or interpolation algorithm~\cite{Su13adaptive}.
+On one hand, there exists simple extension schemes that do not take into account the dynamical behavior of the signal, such as zero-padding, periodic extension, symmetric extension~\cite{Kharitonenko02wavelet,Chen95symmetric}, or polynomial extrapolation~\cite{Williams97discrete}.
+On the other hand, there exists extension schemes based on physically relevant dynamical models, such as the Extented Dynamic Mode Decomposition~\cite{Williams15data} (EDMD), the Gaussian process regression~\cite{Rasmussen06gaussian,Roberts13Gaussian} (GPR), the TBATS algorithm~\cite{DeLivera11forecasting}.
 While the second class of extension schemes give better extended signals than the first class, they generally have a great computational cost.
@@
 Therefore, in section~\ref{se:algo}, we provide an extension method based on a linear dynamic model.
-We derive the corresponding algorithm for boundary effect reduction.
+We derive the corresponding algorithm for boundary effects reduction.
 In section~\ref{se:theoretical}, we show that, even though we consider a simplistic dynamic model, it is sufficient to extend signals takings the form of sums of sine waves.
@@
 We show that our algorithm gives fast results of reasonable quality.
-Finally, we evaluate the performance of our boundary effect reduction algorithm on biomedical signals such as respiratory signals, and compare it to the theoretical results.
+Finally, we evaluate the performance of our boundary effects reduction algorithm on biomedical signals such as respiratory signals, and compare it to the theoretical results.
 

### 2020-05-28 Notes/algorithm.tex
paper

@@
 \]
-We consider here a naive dynamical model, assuming that we have the following relation:
+In a general framework, forecasting means estimating the function $f$ from the observed values taken by the signal, in order to predict its future values.
+For instance, the dynamic mode decomposition~\cite{Schmid10dynamic,Williams15data} allows this by setting very few additional constraints on the regularity of $f$.
+We will see later, in section~\ref{se:theoretical}, that it is not necessary to consider such a complex dynamic model for the study of the signals of interest to us.
+That is why we consider here a naive dynamical model, assuming that we have the following relation:
 \begin{equation}

### 2020-05-28 Notes/results.tex
paper

@@
 The experimental results show that the bias is neither depending on the noise variance $\sigma^2$ nor the forecasting length $\ell$.
-Indeed, independently of $sigma$, we always have $\mu_{xp}[N-1+\ell]\in\left[-0.03\sigma,0.03\sigma\right]$, which is negligible with respect to the magnitude of $\bz$.
+Indeed, independently of $\sigma$, we always have $\mu_{xp}[N-1+\ell]\in\left[-0.03\sigma,0.03\sigma\right]$, which is negligible with respect to the magnitude of $\bz$.
 This result confirms the theoretical result~\eqref{eq:mean.error}.
 
-On Figure~\ref{fig:res.noise.sine}, we display the experimental variance $\gamma_{xp}[N-1+\ell,N-1+\ell]$ for each value of $\sigma$ ($+$ markers).
+On Fig.~\ref{fig:res.noise.sine}, we display the experimental variance $\gamma_{xp}[N-1+\ell,N-1+\ell]$ for each value of $\sigma$ ($+$ markers).
 The associated theoretical asymptotic forecasting variance~\eqref{eq:cov.error.2} is also displayed in solid line.
@@
 As in the previous study, the experimental bias vanishes when $K$ increases, what confirms the approximation result~\eqref{eq:mean.error}.
-Besides, the experimental variance  is displayed on Figure~\ref{fig:res.size.sine} ($+$ markers), and compared with the associated theoretical variance~\eqref{eq:cov.error.2}.
+Besides, the experimental variance  is displayed on Fig.~\ref{fig:res.size.sine} ($+$ markers), and compared with the associated theoretical variance~\eqref{eq:cov.error.2}.
 Each color corresponds to these experimental results obtained for a given value of $K$.
@@
 That one is based on the Gaussian process structure, and therefore offer more flexibility in the type of dynamic that could be modeled than the linear model~\eqref{eq:dyn.model}.
-\item The \emph{Trigonometric , Box-Cox transformation, ARMA errors, Trend and Seasonal components (TBATS)} method~\cite{Rasmussen06gaussian} is based on a classical decomposition of times series into a trend, a seasonal and an ARMA components, with a specific dynamic for the seasonal component.
+\item The \emph{Trigonometric , Box-Cox transformation, ARMA errors, Trend and Seasonal components (TBATS)} method~\cite{DeLivera11forecasting} is based on a classical decomposition of times series into a trend, a seasonal and an ARMA components, with a specific dynamic for the seasonal component.
 This model demands the estimation of numerous parameters and, by implication, may be slow.
@@
    \hline
-   TBATS & ${\color{red}X}$ & ${\color{red}X}$ & ${\color{red}X}$ \\
+   TBATS & $1.732\times 10^{-3}$ & $4.924\times 10^{-4}$ & $1837.12$ \\
    \hline
@@
 
-\subsection{Evaluation of the quality of the boundary effect reduction}
+\subsection{Evaluation of the quality of the boundary effects reduction}
 
 \subsubsection{Metrics}
-The quality of the boundary effect reduction must be evaluated directly on the time-frequency representation.
+The quality of the boundary effects reduction must be evaluated directly on the time-frequency representation.
 To that aim, we compare the obtained representation to the optimal representation $\ccF_N^\mathrm{opt}(\bx)$.
@@
 \subsubsection{Respiratory signal}
-We first consider a 50 minutes-long respiratory signal. This signal is sampled at $\fs=100$~Hz. A zoom on a small portion of the signal is displayed in Figure~\ref{fig:tho}.
+We first consider a 50 minutes-long respiratory signal. This signal is sampled at $\fs=100$~Hz. A zoom on a small portion of the signal is displayed in Fig.~\ref{fig:tho}.
 
@@
 These outliers are probably due to the presence of pulse in the respiratory signal.
-That ones make the adaptive harmonic model temporary irrelevant, and break the validity of the linear dynamical model~\eqref{eq:dyn.model} used to extend the signal.
+That ones make the AM-FM model temporarily irrelevant, and break the validity of the linear dynamical model~\eqref{eq:dyn.model} used to extend the signal.
 That is why {\sf SigExt} is less robust to the vanishing of oscillations that the GPR or EDMD extensions.
@@
 For instance, here the window length we use to evaluate the STFT is of 1500 samples.
-To prevent the STFT from being sensitive to the boundary effect, we set $L=750$.
+To prevent the STFT from being sensitive to the boundary effects, we set $L=750$.
 In this way, when evaluating the spectral content of the signal near its boundaries, the analysis is not limited by a lack of information all along the window support.
@@
    \hhline{|=#=|=|=|=|}
-   Without extension & $2.16\times 10^{-2}$ & $5.26\times 10^{-3}$ & $3.07\times 10^{-2}$ & $1.41\times 10^{-2}$ \\
+   Without extension & $2.16\!\times\! 10^{-2}$ & $5.26\!\times\! 10^{-3}$ & $3.07\!\times\! 10^{-2}$ & $1.41\!\times\! 10^{-2}$ \\
    \hline
-   {\sf SigExt} & $1.72\times 10^{-2}$ & $4.00\times 10^{-3}$ & $2.43\times 10^{-2}$ & $1.12\times 10^{-2}$ \\
+   {\sf SigExt} & $1.72\!\times\! 10^{-2}$ & $4.00\!\times\! 10^{-3}$ & $2.43\!\times\! 10^{-2}$ & $1.12\!\times\! 10^{-2}$ \\
    \hline
-   EDMD & $1.75\times 10^{-2}$ & $4.23\times 10^{-3}$ & $2.45\times 10^{-2}$ & {\color{red}X} \\
+   EDMD & $1.75\!\times\! 10^{-2}$ & $4.23\!\times\! 10^{-3}$ & $2.45\!\times\! 10^{-2}$ & $1.14\!\times\! 10^{-2}$ \\
    \hline
-   GPR & $1.84\times 10^{-2}$ & $4.10\times 10^{-3}$ & $2.48\times 10^{-2}$ & {\color{red}X} \\
+   GPR & $1.84\!\times\! 10^{-2}$ & $4.10\!\times\! 10^{-3}$ & $2.48\!\times\! 10^{-2}$ & $1.10\!\times\! 10^{-2}$ \\
    \hline
@@
 We perform a study similar to the previous one on a 640 second-long photoplethysmogram (PPG) signal, sampled at $\fs=125$~Hz.
-A 32 second-long piece of this signal is displayed on the top of Figure~\ref{fig:ppg}.
-The estimated 2-second extension obtained by {\sf SigExt} on both boundary of this signal is superimposed to the ground-truth signal in the bottom of Figure~\ref{fig:ppg}.
+A 32 second-long piece of this signal is displayed on the top of Fig.~\ref{fig:ppg}.
+The estimated 2-second extension obtained by {\sf SigExt} on both boundary of this signal is superimposed to the ground-truth signal in the bottom of Fig.~\ref{fig:ppg}.
 
@@
 We provide in Table~\ref{tab:otd.ppg} the OTD to the optimal time-frequency representation averaged over the signals.
-For all the considered time-frequency representations, the results clearly shows that our algorithm reduce the influence of the boundary effect.
-This highlights the ability of our approach to limit the distortion due the boundary effect and provide a more accurate representations.
+For all the considered time-frequency representations, the results clearly shows that our algorithm reduce the influence of the boundary effects.
+This highlights the ability of our approach to limit the distortion due the boundary effects and provide a more accurate representations.
 Even though the {\sc SigExt} extension yields time-frequency  representations slightly more sensitive to boundary effects that the extensions given by EDMD or GPR, it is the only technique that allows a real-time implementation.
@@
 It is thus an additional noise to the measurement noise actually contained in the signal.
-Figure~\ref{fig:otd.noise} shows the averaged OTD of {\sf BoundEffRed} for different values of $\sigma_{\mathrm{add}}^2$.
-At first glance, we notice that STFT is slightly more sensitive to noise than SST.
-On one hand, the robustness to noise of the SST is the direct consequence of the approximation result of~\cite{Daubechies16conceft}, discussed in section~\ref{sse:perf.BoundEffRed}.
-On the other hand, the sentivity to noise of the STFT may be interpreted as the inability to forecast the values of the time-frequency coefficients where only noise is active, while sharp time-frequency representations, such as SST, ideally vanish at these points.
+Fig.~\ref{fig:otd.noise} shows the averaged OTD of {\sf BoundEffRed} for different values of $\sigma_{\mathrm{add}}^2$.
+At first glance, we notice that STFT is slightly more sensitive to noise than SST or RS.
+On one hand, the robustness to noise of the SST and the RS is the direct consequence of the approximation result of~\cite{Daubechies16conceft}, discussed in section~\ref{sse:perf.BoundEffRed}.
+On the other hand, the sentivity to noise of the STFT may be interpreted as the inability to forecast the values of the time-frequency coefficients where only noise is active, while sharp time-frequency representations, such as SST or RS, ideally vanish at these points.
 \begin{figure}

### 2020-05-28 Scripts/EvalPerfBoundEffRed/ResultsBoundEffRed.m

@@ -48,7 +48,7 @@ fprintf('|------------------|-----------|-----------|-----------|---------|\n')
 fprintf('|  No extension    | %.3e | %.3e | %.3e | %.3e |\n', mean(OTD.sst.S),mean(OTD.stft.S),mean(OTD.rs.S),mean(OTD.conceft.S))
 fprintf('|  LSE extension   | %.3e | %.3e | %.3e | %.3e |\n', mean(OTD.sst.LSE(~isnan(OTD.sst.LSE))),mean(OTD.stft.LSE(~isnan(OTD.sst.LSE))),mean(OTD.rs.LSE(~isnan(OTD.sst.LSE))),mean(OTD.conceft.LSE(~isnan(OTD.sst.LSE))))
 fprintf('|  EDMD extension  | %.3e | %.3e | %.3e | %.3e |\n', mean(OTD.sst.EDMD),mean(OTD.stft.EDMD),mean(OTD.rs.EDMD),mean(OTD.conceft.EDMD))
-fprintf('|  GPR extension   | %.3e | %.3e | %.3e | %.3e |\n', mean(OTD.sst.GPR(~isnan(OTD.sst.GPR))),mean(OTD.stft.GPR(~isnan(OTD.sst.GPR))),mean(OTD.rs.GPR(~isnan(OTD.sst.GPR)))mean(OTD.conceft.GPR(~isnan(OTD.sst.GPR))))
+fprintf('|  GPR extension   | %.3e | %.3e | %.3e | %.3e |\n', mean(OTD.sst.GPR(~isnan(OTD.sst.GPR))),mean(OTD.stft.GPR(~isnan(OTD.sst.GPR))),mean(OTD.rs.GPR(~isnan(OTD.sst.GPR))),mean(OTD.conceft.GPR(~isnan(OTD.sst.GPR))))
 fprintf('|__________________|___________|___________|___________|_________|\n\n')
 
 %% EEG

### 2020-05-28 Notes/algorithm.tex
paper

@@
 This strategy moves the boundary effects out of the time interval $I=[0, \frac{N-1}{\fs}]$.
-Finally, the boundary-effects insensitive representation $\ccF_N^{\mathrm{ext}}:\RR^{N}\to\RR^{F\times N}$ of $\bx$ is given for all $f\in\{0,\cdots,F-1\}$, $n\in\{0,\cdots,N-1\}$ by:
+Finally, the boundary-effects insensitive representation $\ccF_N^{\mathrm{ext}}:\RR^{N}\to\RR^{F\times N}$ of $\bx$ is given for all $\nu\in\{0,\cdots,F-1\}$, $n\in\{0,\cdots,N-1\}$ by:
 \begin{equation}
-\ccF_N^{\mathrm{ext}}(\bx)[f,n] = \ccF_{N+2L}(\tilde \bx)[f,L+n]\ .
+\ccF_N^{\mathrm{ext}}(\bx)[\nu,n] = \ccF_{N+2L}(\tilde \bx)[\nu,L+n]\ .
 \label{eq:restriction}

### 2020-06-01 Notes/BoundEffRed.tex
paper
updates

@@
 \centering
-%\includegraphics[width=.48\textwidth]{PPGsig.eps}
+\includegraphics[width=.48\textwidth]{zoomPPGsig.eps}
 \includegraphics[width=.48\textwidth]{SSTintro.eps}
 \includegraphics[width=.48\textwidth]{zoomSSTintro.eps}
-\caption{Synchrosqueezing transform of a PPG (top) with a zoom on its right boundary (bottom). }
+\caption{Zoom on a PPG signal (top), with its synchrosqueezing transform (middle) and a zoom on the right boundary of the SST (bottom). The window length is 12~sec. }
 \label{fig:ex.intro}
@@
 A more fundamental improvement would be to perform the prediction step directly in the time-frequency plane, with the aim of specifically predicting the evolution of each instantaneous frequency.
-This would lead to seeing a time-frequency representation as a set of interdependent time series (corresponding to each frequency channel).
+This would lead to seeing a time-frequency representation as a set of interdependent time series (corresponding to each frequency band).
 

### 2020-06-01 Notes/Supplementary material/SupMat.tex
updates

@@
 \usepackage{dsfont}
-\usepackage{hhline}
-\usepackage{algorithm,algorithmic}
+%\usepackage{hhline}
+%\usepackage{algorithm,algorithmic}
+
 \usepackage{xr}
@@
 
-\setcounter{equation}{8}
-\setcounter{figure}{3}
-
 \section{Proof of Lemma~1}
@@
 \subsection{Notations}
-We have:
+By definition the matrix $\bX$ and $\bY$ and because of the noisy signal model~\eqref{eq:model.noise}, we have:
 \begin{align}
@@
 \end{align}
-where $\bE^{(a)} = \sigma\bE_1^{(a)} + \sigma^2\bE_2^{(a)}$ with:
+where $\bE^{(a)} = \sigma\bE_1^{(a)} + \sigma^2\bE_2^{(a)}$, with:
 \[
@@
 The randomness of $\bh^{(\ell)}$ is completely originating from the error matrices.
-Besides, notice that the first $M-1$ rows in $\bE^{(1)}$ are the same as the last $M-1$ rows of $\bE^{(0)}$.
-Thus, we gather all the sources of randomness into an vector $\bg\in\RR^{M(M+1)}$ defined as:
+Besides, notice that the first $M-1$ rows in $\bE^{(1)}$ are equals to the last $M-1$ rows of $\bE^{(0)}$.
+Thus, we gather all the sources of randomness into an vector $\bg\in\RR^{M(M+1)}$, containing $M$ rows and defined as:
 \[
@@
 with $\tilde\bz_k^T= \begin{pmatrix} \bz_k^T & \bz_{k+1}[0] \end{pmatrix}$ and $\tilde\bw_k^T= \begin{pmatrix} \bw_k^T & \bw_{k+1}[0] \end{pmatrix}$.
-Then, $\bg_1$ is a Gaussian random vector as a linear combination of Gaussian random vectors.
+Then, $\bg_1$ is a Gaussian random vector because it is a linear combination of Gaussian random vectors.
 Moreover, using the central limit theorem under weak dependence, we can show that $\bg_2$ also converges towards a Gaussian random vector as $K\to\infty$.

### 2020-06-01 Notes/results.tex
updates

@@
 
-To highlight the fact that the linear dynamical model is sufficient to catch most of the dynamical behavior of signals following the AHM, we compare the performance of the Algorithm~\ref{alg:boundary} with reference forecasting algorithm that could be used for extending such signals.
+To highlight the fact that the linear dynamical model is sufficient to catch most of the dynamical behavior of signals following the AM-FM, we compare the performance of the Algorithm~\ref{alg:boundary} with reference forecasting algorithm that could be used for extending such signals.
 These methods are:
@@
 \centering
-\caption{AHM signal. Performance of the extension methods.}
+\caption{AM-FM signal. Performance of the extension methods.}
 \begin{tabular}{|c||c|c|c|}
@@
    \hline
-   EDMD & $1.75\!\times\! 10^{-2}$ & $4.23\!\times\! 10^{-3}$ & $2.45\!\times\! 10^{-2}$ & $1.14\!\times\! 10^{-2}$ \\
+   EDMD & $1.75\!\times\! 10^{-2}$ & $4.23\!\times\! 10^{-3}$ & $2.45\!\times\! 10^{-2}$ & $1.09\!\times\! 10^{-2}$ \\
    \hline
-   GPR & $1.84\!\times\! 10^{-2}$ & $4.10\!\times\! 10^{-3}$ & $2.48\!\times\! 10^{-2}$ & $1.10\!\times\! 10^{-2}$ \\
+   GPR & $1.84\!\times\! 10^{-2}$ & $4.10\!\times\! 10^{-3}$ & $2.48\!\times\! 10^{-2}$ & $1.24\!\times\! 10^{-2}$ \\
    \hline

### 2020-06-01 Scripts/EvalPerfBoundEffRed/figPPG.m

@@ -30,11 +30,23 @@ tsst = t(1:basicTF.hop:end);
 ff = linspace(0,fs*basicTF.fmax,400) ;
 
 figure;
+subplot(3,1,1);
+plot(t,x,'linewidth',2); axis tight ; 
+ylabel('Signal'); 
+set(gca,'fontsize',18); grid on; xticklabels([]);
+
+subplot(3,1,[2 3]);
 imagesc(tsst,ff,log1p(abs(sstS)/1e1)); axis xy; 
 xlabel('Time (sec.)'); ylabel('Frequency (Hz)'); colormap(1-gray);
 set(gca,'fontsize',18);
 
 figure;
+subplot(3,1,1);
+plot(t,x,'linewidth',2); axis tight ; xlim([27 t(end)]) ;
+ylabel('Signal'); 
+set(gca,'fontsize',18); grid on; xticklabels([]);
+
+subplot(3,1,[2 3]);
 imagesc(tsst,ff,log1p(abs(sstS)/1e1)); axis xy; xlim([27 t(end)])
 xlabel('Time (sec.)'); ylabel('Frequency (Hz)'); colormap(1-gray)
 set(gca,'fontsize',18);

### 2020-06-02 Notes/Supplementary material/SupMat.tex
update tex

@@
-\documentclass[journal, onecolumn]{IEEEtran}
+\documentclass[journal,onecolumn]{IEEEtran}
 
@@
 \usepackage{courier}
-\usepackage{xr}
 \usepackage{xspace}
@@
 \usepackage{dsfont}
-%\usepackage{hhline}
-%\usepackage{algorithm,algorithmic}
 
 \usepackage{xr}
-\externaldocument[I-]{../BoundEffRed}
+\externaldocument[I-]{../BoundEffRed2}
 
@@
 
+\setcounter{equation}{26}
+
 \begin{document}
@@
 \subsection{Notations}
-By definition the matrix $\bX$ and $\bY$ and because of the noisy signal model~\eqref{eq:model.noise}, we have:
+By definition the matrix $\bX$ and $\bY$ and because of the noisy signal model~\eqref{I-eq:model.noise}, we have:
 \begin{align}
@@
 \begin{align*}
-\bS^{(a)}[m,m'] &= \sigma^2\delta_{(m+a)m'}+\sum_{j,j'=1}^J\dfrac{\Omega_j\Omega_{j'}}{K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(N_0+m+a+k)\right)\cos\left(2\pi \frac{f_{j'}}{\fs}(N_0+m'+k)\right) \\
+\bS^{(a)}[m,m'] &= \sigma^2\delta_{(m+a)m'}+\sum_{j,j'=1}^J\dfrac{\Omega_j\Omega_{j'}}{K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(N_0+m+a+k)+\varphi_j\right)\cos\left(2\pi \frac{f_{j'}}{\fs}(N_0+m'+k)+\varphi_{j'}\right) \\
 &= \sigma^2\delta_{(m+a)m'}+\sum_{j=1}^J\dfrac{\Omega_j^2}{2K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)\\
@@
 \balpha_0^{(1)}\bz_{K} &=\sum_{m=0}^{M-1}\balpha_0^{(1)}[m]\bz[N-M+m] \\
-&=\sum_{j,j'=1}^J\Omega_{j'}\dfrac{2}{M}\underbrace{\sum_{m=0}^{M-1}\cos\left(2\pi p_j\frac{m}{M}\right)\cos\left(2\pi p_{j'}\dfrac{N+m}{M}\right)}_{=\delta_{j,j'}\frac{M}2\cos\left(2\pi p_j\frac{N}M\right)} + o(\sigma) \\
-&= \sum_{j=1}^J\Omega_j\cos\left(2\pi p_j\dfrac{N}{M}\right) + o(\sigma) \\
+&=\sum_{j,j'=1}^J\Omega_{j'}\dfrac{2}{M}\underbrace{\sum_{m=0}^{M-1}\cos\left(2\pi p_j\frac{m}{M}\right)\cos\left(2\pi p_{j'}\dfrac{N+m}{M}+\varphi_{j'}\right)}_{=\delta_{j,j'}\frac{M}2\cos\left(2\pi p_j\frac{N}M +\varphi_{j}\right)} + o(\sigma) \\
+&= \sum_{j=1}^J\Omega_j\cos\left(2\pi p_j\dfrac{N}{M}+\varphi_j\right) + o(\sigma) \\
 &= \bz[N] + o(\sigma)
@@
 Consequently:
-\begin{equation*}
+\begin{equation}
 \bmu[N-1+\ell] \underset{K\to\infty}{\sim} o(\sigma)\ .
 \label{eq:bias.ap}
-\end{equation*}
+\end{equation}
 
@@
 Then, as a direct consequence of the Isserlis' theorem, odd-order moments are vanishing.
-Then, combining result~\eqref{eq:bias.ap} and equation~\eqref{eq:bias0}, we obtain:
+Then, combining result~\eqref{eq:bias.ap} and equation~\eqref{I-eq:bias0}, we obtain:
 \begin{align*}
@@
 Then, as a direct consequence of the Isserlis' theorem, odd-order moments are vanishing.
-Then, combining result~\eqref{eq:bias.ap} and equations~\eqref{eq:cov0} and~\eqref{eq:cov.ap}, we obtain:
+Then, combining result~\eqref{eq:bias.ap} and equations~\eqref{I-eq:cov01} and~\eqref{eq:cov.ap}, we obtain:
 \begin{align*}  

### 2020-06-02 Notes/results.tex
update tex

@@
 Then, the forecasting algorithm is run on $1000$ realizations on the discrete signal $\bx$ for three different values of $\sigma$, logarithmically spaced from $10^{-3}$ to $10^{-1}$.
-For each of these values, we determine the experimental bias $\mu_{xp}[N-1\ell]$ and variance $\gamma_{xp}[N-1+\ell,N-1+\ell]$ in function of the forecasting sample index $\ell$ (going from $1$ to $L=100$).
+For each of these values, we determine the experimental bias $\mu_{\xp}[N-1\ell]$ and variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$ in function of the forecasting sample index $\ell$ (going from $1$ to $L=100$).
 
 The experimental results show that the bias is neither depending on the noise variance $\sigma^2$ nor the forecasting length $\ell$.
-Indeed, independently of $\sigma$, we always have $\mu_{xp}[N-1+\ell]\in\left[-0.03\sigma,0.03\sigma\right]$, which is negligible with respect to the magnitude of $\bz$.
+Indeed, independently of $\sigma$, we always have $\mu_{\xp}[N-1+\ell]\in\left[-0.03\sigma,0.03\sigma\right]$, which is negligible with respect to the magnitude of $\bz$.
 This result confirms the theoretical result~\eqref{eq:mean.error}.
 
-On Fig.~\ref{fig:res.noise.sine}, we display the experimental variance $\gamma_{xp}[N-1+\ell,N-1+\ell]$ for each value of $\sigma$ ($+$ markers).
+On Fig.~\ref{fig:res.noise.sine}, we display the experimental variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$ for each value of $\sigma$ ($+$ markers).
 The associated theoretical asymptotic forecasting variance~\eqref{eq:cov.error.2} is also displayed in solid line.
@@
 Then, the forecasting algorithm is run on 3000 realizations on the discrete signal $\bx$ for three different values of $K$, logarithmically spaced from $4.5\times 10^{2}$ to $2\times 10^{3}$.
-For each of these values, we determine the experimental bias $\mu_{xp}[N-1+\ell]$ and variance $\gamma_{xp}[N-1+\ell,N-1+\ell]$ in function of the forecasting sample index $\ell$ (going from $1$ to $500$).
+For each of these values, we determine the experimental bias $\mu_{\xp}[N-1+\ell]$ and variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$ in function of the forecasting sample index $\ell$ (going from $1$ to $500$).
 

### 2020-07-02 Notes/BoundEffRed.tex
paper and code

--- 2020-05-28 Notes/BoundEffRed.tex
@@
 
-\title{An Approach for the Reduction of Boundary Effects in Time-Frequency Representations}
+\title{An Efficient Forecasting Approach for the Reduction of Boundary Effects in Time-Frequency Representations}
 
@@
 \begin{abstract}
-Time-frequency representations are intrinsically subject to the boundary effects.
-As a consequence, the structures of signals that are highlighted by the representations are garbled when approaching the boundaries of the time-frequency domain.
-In this paper, we propose an approach for the reduction of such boundary effects, whose objective is to be able to be used in real time.
-That one relies on an extension of the analyzed signal obtained by a forecasting technique.
-In the case of the study of a class of locally oscillating signals, we provide a theoretical study of the performance of our approach.
-Following a numerical verification of the performance of the algorithm, we implement it on real-life biomedical signals.
+Time-frequency (TF) representations of time series are intrinsically subject to the boundary effects.
+As a consequence, the structures of signals that are highlighted by the representations are garbled when approaching the boundaries of the TF domain.
+In this paper, for the purpose of real-time TF information acquisition for nonstationary oscillatory time series, we propose a numerically efficient approach for the reduction of such boundary effects.
+The solution relies on an extension of the analyzed signal obtained by a forecasting technique.
+In the case of the study of a class of locally oscillating signals, we provide a theoretical guarantee of the performance of our approach.
+Following a numerical verification of the performance of the algorithm, we validated its performance on various real-life biomedical signals.
 \end{abstract}
@@
 \label{se:introduction}
-\IEEEPARstart{I}{n} any digital acquisition system, the study and the interpretation of the measured signals generally require the use an analysis tool, which enable to point out the useful characteristics of the signal.
-This problem arises for the analysis of signals as varied as audio~\cite{Stowell18computational,Muller11signal}, mechanical~\cite{Peng02vibration} or biomedical signals~\cite{Akay96detection}.
-For instance, any  measured biomedical signal, such as photoplethysmogram (PPG), may not be interpreted as it is, from its run chart.
-An analysis tool would make possible the extraction of some useful characteristics such as heart rate or blood pressure.
+\IEEEPARstart{I}{n} any digital acquisition system, the study and the interpretation of the measured signals generally require an analysis tool, which enables researchers to point out the useful characteristics of the signal.
+The need for signal analysis arises from various signals, ranging from audio~\cite{Stowell18computational,Muller11signal}, mechanical~\cite{Peng02vibration} or biomedical signals~\cite{Akay96detection}.
+In this paper, we focus on oscillatory time series.
+For instance, biomedical signal, such as photoplethysmogram (PPG), contain several characteristics, such as respiratory rate or blood pressure, that cannot be interpreted from its running chart in the time domain.
+An analysis tool would make possible the extraction of these useful characteristics.
 
-In general, the observed quantities are produced by transient phenomena, that can vary rapidly and irregularly.
-As a consequence, the measured signals exhibit nonstationary behavior.
+Usually, the measured signals exhibit nonstationary behavior, and the observed quantities might be interfered by transient phenomena that can vary rapidly and irregularly.
+For example, the signal might oscillate fast with large amplitude at one moment, and then oscillate slow with small amplitude at the next moment.
 In order to adapt the analysis to nonstationarities, local spectral analysis is generally performed~\cite{Stoica05spectral,Matz97generalized}.
-The short time Fourier transform~\cite{Grochenig01foundations} (STFT) is a typical tool which is build that way, and enables the determination of the local frequency content of a nonstationary signal.
+The short time Fourier transform~\cite{Grochenig01foundations} (STFT) is a typical tool which is build for this purpose, and enables the determination of the local frequency content of a nonstationary signal.
 
 Windowing is a common method for performing local analysis.
-Among others, STFT, wavelet transform~\cite{Mallat99wavelet}, synchrosqueezing transform (SST), and reassignment~\cite{Auger13time} (RS) are representations that fall back on the use of an analysis window.
+Among many others, STFT~\cite{Flandrin:1999}, continuous wavelet transform (CWT)~\cite{Da1992}, synchrosqueezing transform (SST) \cite{Daubechies11synchrosqueezed}, and reassignment~\cite{Auger13time} (RS) are representations that fall back on the use of an analysis window.
 Let $x:I\to\RR$ denote the observed signal, where $I$ denotes the finite interval where the signal is measured.
@@
 \]
-Then, the local analysis of $x$ around the instant $\tau\in I$ rely on the evaluation of the following dot product:
+Then, the local analysis of $x$ around the instant $\tau\in I$ rely on the evaluation of the following inner product:
 \begin{equation}
@@
 \end{equation}
-A major shortcoming of this technique occurs when analyzing the signal $x$ near the boundaries of the interval $I$.
-Indeed, at these points, half of the information is missing.
-Consequently the results of the dot product~\ref{eq:windowing} is disturbed by this lack of information.
-This phenomenon is the so-called \emph{boundary effect}.
-We display, on the left of Fig.~\ref{fig:ex.intro} the result of the SST of a PPG (see section~\ref{ssse:ppg} for a comprehensive description).
+A major shortcoming of this technique occurs when analyzing the signal $x$ {\em near the boundaries} of the interval $I$.
+Clearly, at these points, half of the information is missing.
+Consequently the results of the inner product~\eqref{eq:windowing} is distorted.
+This phenomenon is usually understood as the \emph{boundary effect}.
+We display on Fig.~\ref{fig:ex.intro} the result of the SST of a PPG (see section~\ref{ssse:ppg} for a comprehensive description).
 The distortion resulting from boundary effects is clearly visible on both sides of this representation.
 Indeed, while in the major central part of the image, clear lines stand out, they become blurred as they approach the left and right boundaries of the image.
-A zoom on the right boundary of this SST, displayed on the right of Fig.~\ref{fig:ex.intro}, emphasizes the result of boundary effects.
-More generally, estimations of signal characteristics, like instantaneous frequencies~\cite{Delprat92asymptotic} or amplitudes, from its time-frequency representation appear to be imprecisely determinable (or even likely to fail) in the vicinity of the boundaries.
+The right boundary is zoomed in and displayed on the right of Fig.~\ref{fig:ex.intro}, which emphasizes the result of boundary effects.
+Clearly, estimations of signal characteristics, like instantaneous frequencies~\cite{Delprat92asymptotic} or amplitudes, from this TF representation appear to be imprecisely determinable (or even likely to fail) in the vicinity of the boundaries.
+Moreover, this boundary effect would unavoidably limit applying these window based TF analysis tools for the real-time analysis purpose.
+It is thus desirable to have a solution to eliminate this boundary effect.
 
@@
 \centering
-%\includegraphics[width=.48\textwidth]{PPGsig.eps}
-\includegraphics[width=.48\textwidth]{SSTintro.eps}
-\includegraphics[width=.48\textwidth]{zoomSSTintro.eps}
-\caption{Synchrosqueezing transform of a PPG (top) with a zoom on its right boundary (bottom). }
+\includegraphics[width=.48\textwidth]{PPGsig.eps}
+\includegraphics[width=.24\textwidth]{zoomSSTintro.eps}
+\includegraphics[width=.24\textwidth]{zoomSSTBoundEffRed.eps}
+\caption{A segment of PPG signal (top) and the right boundary of a TF representation determined by the SST without extension (bottom left), and the right boundary of a TF representation determined by the SST with the proposed boundary effect reduction algorithm by forecasting (bottom right).
+The window length for the SST is 12~sec.}
 \label{fig:ex.intro}
@@
 
-The purpose of this paper is to provide a fast algorithm allowing to tackle boundary effects and limit the disruptions it may cause on the signals representations based on windowing.
+%The purpose of this paper is providing a fast algorithm allowing to tackle boundary effects and limit the disruptions it may cause on the signals representations based on windowing. 
 Attempts to minimize the boundary effects generally consists in softening the discontinuity on signals edges.
-For instance, judicious choices of analysis windows whose support does not interfere with the boundary  points can minimize the occurrence of aberrant patterns near the boundaries of the time-frequency plane~\cite{Chui92wavelets,Depczynski99fast}.
-These techniques therefore do not make it possible to reduce the boundary effects of any time-frequency representations, and for any analysis window.
-For this purpose, a natural idea consists in carrying out a preliminary step to the analysis allowing to extend the signal beyond its boundaries.
-Hence, to implement our boundary effects reduction procedure, we proceed in two steps.
+For instance, judicious choices of analysis windows whose support does not interfere with the boundary points can minimize the occurrence of aberrant patterns near the boundaries of the TF plane~\cite{Chui92wavelets,Depczynski99fast}.
+Due to the specific relationship of chosen analysis windows and TF analysis tool, these techniques do not make it possible to reduce the boundary effects of any TF representations, and for any analysis window.
+Another natural idea consists of carrying out a preliminary step of extending the signal beyond its boundaries,
+%
+and due to its flexibility, various extension schemes have been proposed.
+For example, there exists simple extension schemes that do not take into account the dynamical behavior of the signal, such as zero-padding, periodic extension, symmetric extension~\cite{Kharitonenko02wavelet,Chen95symmetric}, or polynomial extrapolation~\cite{Williams97discrete}.
+There also exist extension schemes based on physically relevant dynamical models, such as the Extended Dynamic Mode Decomposition~\cite{Williams15data} (EDMD), the Gaussian process regression~\cite{Rasmussen06gaussian,Roberts13Gaussian} (GPR), the Trigonometric, Box-Cox transformation, ARMA errors, Trend and Seasonal components (TBATS) algorithm~\cite{DeLivera11forecasting}, the dynamic linear models \cite{west2006bayesian}, and long short-term memory model \cite{vlachas2018data}.
+It is a far from exhaustive list, and we refer readers with interest to \cite{hyndman2018forecasting} for a friendly monograph on the general forecasting topic.
+While the second class of extension schemes give better extended signals than the first class, they generally have a great computational cost.
+%
+
+In this paper, we propose a fast extension algorithm based on a simple dynamical model, in order to optimize the trade-off between the extension quality and the computational cost.
+The proposed algorithm is composed of two steps.
 \begin{enumerate}
@@
 \end{enumerate}
-Thus, assuming that the quality of the extension step is sufficient, the representation obtained that way will be less sensitive to the boundary effects than the result of the analysis tool applied directly to the non-extended signal.
+Thus, assuming that the quality of the extension step is sufficient, the analysis results obtained that way will be less sensitive to the boundary effects than the result of the analysis tool applied directly to the non-extended signal.
+%
+We claim and prove that forecasting oscillatory signals based on the simple dynamical model combined with the simple least square approach is sufficient for reducing the boundary effects for the TF analysis, or other kernel based analysis.
+See Figure \ref{fig:ex.intro} for a snapshot of the result.
+The main benefit of this simple approach is a numerically efficient solution with theoretical guarantee for the real-time analysis purpose.
 
-Various extension schemes have been proposed.
-On one hand, there exists simple extension schemes that do not take into account the dynamical behavior of the signal, such as zero-padding, periodic extension, symmetric extension~\cite{Kharitonenko02wavelet,Chen95symmetric}, or polynomial extrapolation~\cite{Williams97discrete}.
-On the other hand, there exists extension schemes based on physically relevant dynamical models, such as the Extented Dynamic Mode Decomposition~\cite{Williams15data} (EDMD), the Gaussian process regression~\cite{Rasmussen06gaussian,Roberts13Gaussian} (GPR), the TBATS algorithm~\cite{DeLivera11forecasting}.
-While the second class of extension schemes give better extended signals than the first class, they generally have a great computational cost.
-We propose here a fast extension algorithm based on a simple dynamical model, in order to optimize the trade-off between the extension quality and the computational cost.
-Furthermore, we prove that the dynamical model we consider is sufficient to forecast locally oscillating signals.
-
-Therefore, in section~\ref{se:algo}, we provide an extension method based on a linear dynamic model.
+The paper is organized in the following.
+In section~\ref{se:algo}, we provide an extension method based on a linear dynamic model.
 We derive the corresponding algorithm for boundary effects reduction.
@@
 \label{se:conclusion}
-In this paper, we propose an algorithm, named {\sf BoundEffRed}, for the boundary effects reduction on time-frequency representations.
-This method is based on an extension of the signal obtained by forecasting.
-We have shown theoretically that the chosen dynamic model is sufficient for the extension of signals formed by a sum of sine waves whose instantaneous frequencies and amplitudes vary slowly.
-Moreover, the numerical results show that this strategy allows us to switch to a real-time implementation of {\sf BoundEffRed} is possible, due to the low computation time, unlike other pre-existing forecasting methods.
+In this paper, we propose an algorithm, named {\sf BoundEffRed}, for the boundary effects reduction on TF representations to achieve a real-time signal analysis.
+This method is based on an extension of the signal obtained by a simple-minded and numerically efficient forecasting.
+We have shown theoretically that the chosen dynamic model is sufficient for the extension of signals formed by a sum of sine waves.
+Moreover, the numerical results show that this strategy allows us to switch to a real-time implementation of {\sf BoundEffRed} is possible, due to the low computation time, unlike other existing forecasting methods.
 The numerical results also confirmed the robustness to noise of {\sf BoundEffRed}, as well as its ability to be applied to many time-frequency representations.
-The MATLAB code and datasets used to produce the numerical results of this paper are available online at the following address:
-\begin{center}
-    {\small\url{https://github.com/AdMeynard/BoundaryEffectsReduction}}
-\end{center}
-Other numerical examples are also included, such as the application of {\sf BoundEffRed} to other types of biomedical signals such as ECGs.
+%Other numerical examples are also included, such as the application of {\sf BoundEffRed} to other types of biomedical signals such as ECGs.
 
 Various improvements can be considered to make the algorithm more robust to certain situations.
-In particular, we have noticed that when the observed signal does not oscillate, the forecasting is no longer relevant, and outlier time-frequency representations are obtained.
-A preliminary step should then be added to the algorithm to select the portions of the signal that exhibit oscillations, in order to reject the portions for which accurate time-frequency analysis is practically useless.
-A more fundamental improvement would be to perform the prediction step directly in the time-frequency plane, with the aim of specifically predicting the evolution of each instantaneous frequency.
-This would lead to seeing a time-frequency representation as a set of interdependent time series (corresponding to each frequency channel).
+In particular, we have noticed that when the observed signal does not oscillate {\color{red}[HT: we need to confirm this statement]}, the forecasting is no longer relevant, and the TF representations are not ideal.
+A preliminary step should then be added to the algorithm to select the portions without sufficient quality, in order to reject the portions for which a better forecasting scheme is needed.
+A more fundamental improvement would be to perform the prediction step directly in the TF plane, or even with the aim of specifically predicting the evolution of each instantaneous frequency.
+This would lead to seeing a TF representation as a multivariate time series (corresponding to each frequency band).
+We will explore these possibilities in our future work.
 

### 2020-07-02 Notes/algorithm.tex
paper and code

@@
-As explained above, the algorithm for the reduction of boundary effects on time-frequency representations relies on two steps. These ones are detailed in the current section.
+As explained above, the algorithm for the reduction of boundary effects on TF representations relies on extending the signal by forecasting it before applying the TF analysis. 
+%
 
-\subsection{Forecasting}
-
-Let $x:\RR\to\RR$ denote a continuous-time signal. In this work, we consider a finite-length discretization of that one. Thus, the sampled signal $\bx$, whose length is denoted by $N$, is such that
+%\paragraph{Notations} 
+We start with the notation.
+Let $x:\RR\to\RR$ denote a continuous-time signal.
+In this work, we consider a finite-length discretization of that one.
+Thus, the sampled signal $\bx$, whose length is denoted by $N$, is such that
 \[
@@
 where $\fs$ denotes the sampling frequency. 
-
-\paragraph{Notations} 
 Let $M$ and $K$ be two integers such that $M<N$ and $K+M<N$. Then, for all $k\in\{0,\ldots,K-1\}$, we extract from $\bx\in\RR^N$ the sub-signal $\bx_k\in\RR^M$ given by:
@@
 \end{equation*}
+The boundary effect reduction algorithm is based on manipulating $\bX$ and $\bY$.
+
+%
+The pseudo-code of the proposed algorithm to reduce boundary effects on windowing-based TF representations is shown in the Algorithm~\ref{alg:boundary}.
+We coined the algorithm {\sf BoundEffRed}.
+Below, we detail the algorithm, particularly the Signal extension algorithm {\sf SigExt} in Algorithm~\ref{alg:boundary}.
+
+\begin{algorithm}
+\caption{Tackling boundary effects. $\bF_\bx = \mbox{\sf BoundEffRed}(\bx,M,K,L,\ccF)$}
+\label{alg:boundary}
+\begin{algorithmic}
+\STATE {\bf Inputs}: $\bx$, $M$, $K$, $L$, $\ccF$
+\STATE 
+\STATE {\bf Forecasting step.}
+\STATE \quad\textbullet\ Signal extension: $\tilde\bx = \mbox{\sf SigExt}(\bx)$.
+\STATE 
+\STATE {\bf Representation step.}
+\STATE \quad\textbullet\ Representation evaluation: $\ccF_{N+2L}(\tilde\bx)$.
+\STATE \quad\textbullet\ Restriction of $\ccF_{N+2L}(\tilde\bx)$ to the central time interval (see~\eqref{eq:restriction}) to obtain $\bF_\bx=\ccF_N^{\mathrm{ext}}(\bx)$.
+\STATE 
+\STATE {\bf Output}: Signal representation $\bF_\bx$
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Step 1: Extension by forecasting}
 
@@
 In a general framework, forecasting means estimating the function $f$ from the observed values taken by the signal, in order to predict its future values.
-For instance, the dynamic mode decomposition~\cite{Schmid10dynamic,Williams15data} allows this by setting very few additional constraints on the regularity of $f$.
-We will see later, in section~\ref{se:theoretical}, that it is not necessary to consider such a complex dynamic model for the study of the signals of interest to us.
+For instance, the dynamic mode decomposition~\cite{Schmid10dynamic,Williams15data} or other more complicated models \cite{Roberts13Gaussian,DeLivera11forecasting,west2006bayesian,vlachas2018data}, allow this by setting additional constraints on the behavior of $f$.
+We will see later, in section~\ref{se:theoretical}, that it is not necessary to consider such a complex dynamic model for the study of the oscillatory signals of interest to us.
 That is why we consider here a naive dynamical model, assuming that we have the following relation:
@@
 where $\bA\in\RR^{M\times M}$.
-We adopt a classical strategy in the study of dynamical systems, that is the linearization of a nonlinear phenomenon.
-Notice that this linear dynamical model can be written equivalently in function of the sub-signals $\bx_k$, as:
+In other words, we adopt a classical strategy in the study of dynamical systems, the {\em linearization} of a nonlinear phenomenon, when the system is sufficient regular.
+Notice that this linearized dynamical model can be written equivalently in function of the sub-signals $\bx_k$, as:
 \begin{equation}
@@
 The forecasting method consists in estimating the unknown matrix $\bA$.
-Indeed, let $\tilde\bA$ denotes the estimate of $\bA$, we then obtain the forecasting of the signal at time $\frac{N-1+\ell}{\fs}$ by:
+Indeed, let $\tilde\bA$ denotes the estimate of $\bA$.
+We then obtain the forecasting of the signal at time $\frac{N-1+\ell}{\fs}$ by:
 \begin{equation}
@@
 \end{equation}  
-where $\balpha^{(\ell)}$ denotes the last row of $\tilde\bA^\ell$, that is to say:
+where $\balpha^{(\ell)}$ denotes the last row of $\tilde\bA^\ell$; that is to say,
 \begin{equation}
@@
 \end{equation}
-where $\be_M$ is the vector of length $M$ given by $\be_M = \begin{pmatrix} 0 & \cdots & 0 & 1\end{pmatrix}^T$.
+where $\be_M$ is the unit vector of length $M$ given by $\be_M = \begin{pmatrix} 0 & \cdots & 0 & 1\end{pmatrix}^T$.
 
-\paragraph{Model estimation} To estimate the matrix $\bA$, we basically implement the least square estimator. Thus, we solve the following problem:
+\paragraph{Model estimation} To estimate the matrix $\bA$, we consider the simple but numerically efficient least square estimator. That is, we solve the following problem:
 \begin{equation}
@@
 
-\paragraph{Signal extension} In order to reduce the boundary effects on both "sides" of the time-frequency (or time-scale) representation, we finally construct the extended signal $\tilde\bx\in\RR^{N+2L}$ concatenating the backward prediction $\tilde\bx_{\rm bw}$, the observed signal $\bx$, and the forward prediction $\tilde\bx_{\rm fw}$.
+\paragraph{Signal extension} In order to reduce the boundary effects on both ``sides'' of the TF (or time-scale) representation, we finally construct the extended signal $\tilde\bx\in\RR^{N+2L}$ concatenating the backward prediction $\tilde\bx_{\rm bw}$, the observed signal $\bx$, and the forward prediction $\tilde\bx_{\rm fw}$.
 We summarize the extension step in Algorithm~\ref{alg:extension}.
@@
 \begin{algorithm}
-\caption{Signal extension. $\tilde\bx = \mathsf{SigExt}(\bx,M,K,L)$}
+\caption{Signal extension. $\tilde\bx = \mbox{\sf SigExt}(\bx,M,K,L)$}
 \label{alg:extension}
@@
 
-\subsection{Representation}
+\subsection{Step 2: Extended Time-Frequency Representation}
 
-Let $\ccF_{N}:\RR^{N}\to\RR^{F\times N}$ generically denotes the time-frequency or time-scale representation we are interested in.
-It can be, for instance, such as short-time Fourier transform (STFT), the continuous wavelet transform (CWT), the synchrosqueezing transform (SST), or the reassignment (RS).
-Here, $F$ typically denotes the size of the representation along the frequency axis.
-Due to the boundary effects, the representation $\ccF_{N}(\bx)$ shows undesired patterns when approaching its edges.For example, the instantaneous frequencies highlighted by the SST can be blurred near that edges.
-To limit these phenomena, we apply the representation to the estimated extended signal $\tilde\bx$.
+Let $\ccF_{N}:\RR^{N}\to\RR^{F\times N}$ generically denotes the TF or time-scale representation we are interested in, which could be, for instance, STFT, CWT, SST, or RS.
+Here, $F$ typically denotes the size of the discretization along the frequency axis.
+Due to the boundary effects, the representation $\ccF_{N}(\bx)$ shows undesired patterns when approaching its edges.
+%For example, the instantaneous frequencies highlighted by the SST can be blurred near that edges.
+To alleviate the boundary effect, we apply the representation to the estimated extended signal $\tilde\bx$.
 This strategy moves the boundary effects out of the time interval $I=[0, \frac{N-1}{\fs}]$.
@@
 \end{equation*}
-
-\subsection{Global algorithm}
-Finally, the global procedure we implement to reduce boundary effects on windowing-based representations is summarized by the pseudo-code of Algorithm~\ref{alg:boundary}.
-
-\begin{algorithm}
-\caption{Tackling boundary effects. $\bF_\bx = \mathsf{BoundEffRed(}\bx,M,K,L,\ccF)$}
-\label{alg:boundary}
-\begin{algorithmic}
-\STATE {\bf Inputs}: $\bx$, $M$, $K$, $L$, $\ccF$
-\STATE 
-\STATE {\bf Forecasting step.}
-\STATE \quad\textbullet\ Signal extension: $\tilde\bx = \mathrm{SigExt}(\bx)$.
-\STATE 
-\STATE {\bf Representation step.}
-\STATE \quad\textbullet\ Representation evaluation: $\ccF_{N+2L}(\tilde\bx)$.
-\STATE \quad\textbullet\ Restriction of $\ccF_{N+2L}(\tilde\bx)$ to the central time interval (see~\eqref{eq:restriction}) to obtain $\bF_\bx=\ccF_N^{\mathrm{ext}}(\bx)$.
-\STATE 
-\STATE {\bf Output}: Signal representation $\bF_\bx$
-\end{algorithmic}
-\end{algorithm}
+We call $\ccF_N^{\mathrm{ext}}$ the {\em boundary-free TF representation}.

### 2020-07-02 Notes/results.tex
paper and code

@@
+For the reproducibility purpose, the MATLAB code and datasets used to produce the numerical results in this section are available online at \url{https://github.com/AdMeynard/BoundaryEffectsReduction}.
+
 \subsection{Evaluation the forecasting performance}
 In that section, we first evaluate the quality of the forecasting step and compare it the the theoretical results provided by Theorem~\ref{th:error}.
-The level of the forecasting error one depends on at least two parameters:
+The level of the forecasting error depends on at least two parameters:
 \begin{itemize}
@@
 \paragraph{Influence of the noise variance $\sigma^2$} Here, the size of the training dataset is set to $K=450$.
-Then, the forecasting algorithm is run on $1000$ realizations on the discrete signal $\bx$ for three different values of $\sigma$, logarithmically spaced from $10^{-3}$ to $10^{-1}$.
-For each of these values, we determine the experimental bias $\mu_{\xp}[N-1\ell]$ and variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$ in function of the forecasting sample index $\ell$ (going from $1$ to $L=100$).
+Then, the forecasting algorithm is run on $1000$ realizations on the discrete signal $\bx$ for three different values of $\sigma$, logarithmically equi-spaced from $10^{-3}$ to $10^{-1}$.
+For each of these values, we determine the experimental bias, denoted as $\mu_{\xp}[N-1+\ell]$, and experimental variance, denoted as $\gamma_{\xp}[N-1+\ell,N-1+\ell]$, in function of the forecasting sample index $\ell$ (going from $1$ to $L=100$).
 
@@
 
-On Fig.~\ref{fig:res.noise.sine}, we display the experimental variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$ for each value of $\sigma$ ($+$ markers).
+On Fig.~\ref{fig:res.noise.sine}, we display the experimental variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$ for each value of $\sigma$.
 The associated theoretical asymptotic forecasting variance~\eqref{eq:cov.error.2} is also displayed in solid line.
 As expected, this result highlights the fact that the forecasting variance increases linearly with respect to $\sigma^2$.
-More surprisingly, this result shows that the forecasting variance slightly decreases with $\ell$, what is counterintuitive.
-It should be noted that, contrary to what expression~\eqref{eq:cov.error.2} suggests, smaller values of $\sigma$ do not  cause a decrease of the experimental variance, but an increase.
-This comes from the fact that when $\sigma$ is small, the matrix $\bX\bX^T$ becomes ill-conditioned.
+Surprisingly, this result shows that the forecasting variance {\em slightly} decreases with $\ell$, what is counterintuitive.
+%
+{\color{red} [HT: need a more convincing explanation.] It should be noted that, contrary to what expression~\eqref{eq:cov.error.2} suggests, smaller values of $\sigma$ do not cause a decrease of the experimental variance, but an increase.
+[HT: strange.]} This comes from the fact that when $\sigma$ is small, the matrix $\bX\bX^T$ becomes ill-conditioned.
 The calculation of the forecasting matrix $\tilde\bA$ in~\eqref{eq:lse} is then strongly disturbed.
@@
 \paragraph{Influence of the training dataset size $K$} Here, the noise variance $\sigma$ is set to $\sigma=10^{-2}$.
-Then, the forecasting algorithm is run on 3000 realizations on the discrete signal $\bx$ for three different values of $K$, logarithmically spaced from $4.5\times 10^{2}$ to $2\times 10^{3}$.
+Then, the forecasting algorithm is run on 3000 realizations on the discrete signal $\bx$ for three different values of $K$, logarithmically equi-spaced from $4.5\times 10^{2}$ to $2\times 10^{3}$.
 For each of these values, we determine the experimental bias $\mu_{\xp}[N-1+\ell]$ and variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$ in function of the forecasting sample index $\ell$ (going from $1$ to $500$).
@@
 As in the previous study, the experimental bias vanishes when $K$ increases, what confirms the approximation result~\eqref{eq:mean.error}.
-Besides, the experimental variance  is displayed on Fig.~\ref{fig:res.size.sine} ($+$ markers), and compared with the associated theoretical variance~\eqref{eq:cov.error.2}.
+Besides, the experimental variance  is displayed on Fig.~\ref{fig:res.size.sine}, and compared with the associated theoretical variance~\eqref{eq:cov.error.2}.
 Each color corresponds to these experimental results obtained for a given value of $K$.
-This result validates the asymptotic behavior provided by~\eqref{eq:cov.error.2}, and we can show that, as soon as $K$ is sufficiently high, the product $K\gamma[\ell,\ell]$ is approximately independent of $K$.
+This result validates the asymptotic behavior provided by~\eqref{eq:cov.error.2}, and we can show that the second order moment $\gamma[\ell,\ell]$ is less dependent on $K$ when $K$ is large.
+
 \begin{figure}
@@
 The choice of this parameter is especially crucial when the deterministic component of the signal is no longer stationary.
-The adaptive harmonic model, discussed below, is an example.
+The AHM, discussed below, is an example.
 
@@
 \label{ssse:res.ahm}
-We now consider a signal whose instantaneous frequencies and amplitudes of its components vary over time.
-The deterministic component $\bz$ of the random vector $\bx$ ( constructed following the model~\eqref{eq:model.noise}) takes the following form, for all $n\in\{1,\ldots,N\}$:
+We now consider a signal satisfying the AHM so that the instantaneous frequencies and amplitudes of its components vary over time.
+The deterministic component $\bz$ of the random vector $\bx$ (constructed following the model~\eqref{eq:model.noise}) takes the following form, for all $n\in\{1,\ldots,N\}$:
 \[
@@
 
-To highlight the fact that the linear dynamical model is sufficient to catch most of the dynamical behavior of signals following the AM-FM, we compare the performance of the Algorithm~\ref{alg:boundary} with reference forecasting algorithm that could be used for extending such signals.
+To highlight the fact that the linear dynamical model is sufficient to catch most of the dynamical behavior of signals following the AHM, we compare the performance of the Algorithm~\ref{alg:boundary} with a simple extension obtained by pointwise symmetrization~\cite{Kharitonenko02wavelet}.
+We also evaluate the performance of reference forecasting algorithm that could be used for extending such signals.
 These methods are:
 \begin{itemize}
-\item The \emph{Extended Dynamic Mode Decomposition (EDMD)} has been developped by Williams \etal~\cite{Williams15data}.
+\item The EDMD has been developed by Williams \etal~\cite{Williams15data}.
 The proposed algorithm is a way to obtain an approximation of the so-called Koopman operator of the observed system, which theoretically allows to catch dynamic of nonlinear systems~\cite{Korda18linear}.
-\item The \emph{Gaussian Process Regression (GPR)}~\cite{Rasmussen06gaussian} is a method relying on a probabilistic dynamical model.
+\item The GPR~\cite{Rasmussen06gaussian} is a method relying on a probabilistic dynamical model.
 That one is based on the Gaussian process structure, and therefore offer more flexibility in the type of dynamic that could be modeled than the linear model~\eqref{eq:dyn.model}.
-\item The \emph{Trigonometric , Box-Cox transformation, ARMA errors, Trend and Seasonal components (TBATS)} method~\cite{DeLivera11forecasting} is based on a classical decomposition of times series into a trend, a seasonal and an ARMA components, with a specific dynamic for the seasonal component.
+\item The TBATS method~\cite{DeLivera11forecasting} is based on a classical decomposition of times series into a trend, a seasonal and an ARMA components, with a specific dynamic for the seasonal component.
 This model demands the estimation of numerous parameters and, by implication, may be slow.
@@
 Corresponding results are given in Table~\ref{tab:mse.sine}.
-They show that the naive extension we propose gives satisfying results, even though the other methods, more sophisticated, give MSE values that are somewhat smaller.
-Nevertheless, a major limit of those methods is the computing time they require, which prevent them from being used to exploit real-time data.
+They show that the naive extension we propose gives satisfying results, in particular in comparison with the point-symmetric extension.
+Besides, even though the other more sophisticated methods, like GPR, give MSE values that have a slightly smaller standard deviation, these methods are substantially limited by the computing time they require, which prevent them from being used to exploit real-time data.
 Thus, {\sf SigExt} is the extension method that optimize the trade-off between the forecasting quality and the computing time.
-That is why, it is implemented in our algorithm for the reduction of boundary effects.
+%That is why, it is implemented in our algorithm for the reduction of boundary effects.
 
@@
 \centering
-\caption{AM-FM signal. Performance of the extension methods.}
+\caption{AHM signal. Performance of the extension methods.}
 \begin{tabular}{|c||c|c|c|}
@@
    \hhline{|=#=|=|=|}
-   {\sf SigExt} & $1.433\times 10^{-3}$ & $4.361\times 10^{-4}$ & $0.15$ \\
+   {\sf SigExt} & $1.433\times 10^{-3}$ & $4.361\times 10^{-4}$ & $0.152$ \\
    \hline
-   EDMD & $3.076\times 10^{-2}$ & $8.095\times 10^{-2}$ & $2.53$\\
+   Symmetric & $1.019\times 10^{1}$ & $1.192\times 10^{2}$ & $0.002$ \\
    \hline
-   GPR & $1.436\times 10^{-3}$ & $4.346\times 10^{-4}$ & $146.33$ \\
+   EDMD & $3.076\times 10^{-2}$ & $8.095\times 10^{-2}$ & $2.537$\\
    \hline
-   TBATS & $1.732\times 10^{-3}$ & $4.924\times 10^{-4}$ & $1837.12$ \\
+   GPR & $1.436\times 10^{-3}$ & $4.346\times 10^{-4}$ & $146.331$ \\
+   \hline
+   TBATS & $1.732\times 10^{-3}$ & $4.924\times 10^{-4}$ & $1837.120$ \\
    \hline
@@
 \subsubsection{Metrics}
-The quality of the boundary effects reduction must be evaluated directly on the time-frequency representation.
+The quality of the boundary effects reduction must be evaluated directly on the TF representation.
 To that aim, we compare the obtained representation to the optimal representation $\ccF_N^\mathrm{opt}(\bx)$.
@@
 
-In the aim of comparing the different techniques, we use a criterion, proposed in~\cite{Daubechies16conceft}, that quantify the distance between a given time-frequency representation and the optimal one.
+In the aim of comparing the different techniques, we use a criterion, proposed in~\cite{Daubechies16conceft}, that quantify the distance between a given TF representation and the optimal one.
 It is built in analogy with the optimal transport distance, which enables quantifying the distance between two probability density functions.
 Let us generically denote a time frequency representation $\ccQ$.
-Then, for $t$ fixed, we consider the following pseudo-probability density function:
+Then, for $t$ fixed, we consider the following probability density function:
 \begin{align}
@@
 \end{align}
-Then, at each instant $t$, we can then determine the optimal transport distance $d_{t}$ between the two pseudo densities.
+Then, at each instant $t$, we can then determine the optimal transport distance $d_{t}$ between the two densities.
 It is given by the $L^1$ norm of the difference between the associated distribution functions.
@@
 where $P_{\ccQ}^t(\xi)=\int_{-\infty}^\xi p_{\ccQ}^t(\nu)\dd\nu$ and $\tilde P_{\ccF_0}^t(\xi)=\int_{-\infty}^\xi\tilde p_{\ccF_0}^t(\nu)\dd\nu$.
-Finally, the distance between the two time-frequency representations is obtained by averaging all the optimal transport distances with respect to time:
+Finally, the distance between the two TF representations is obtained by averaging all the optimal transport distances with respect to time:
 \begin{equation}
@@
 \end{equation}
-The \textit{Optimal Transport Distance} (OTD) quantifies the proximity between the estimated and actual instantaneous frequencies while favouring the sparsity of the estimated time-frequency representation.
+The \textit{Optimal Transport Distance} (OTD) quantifies the proximity between the estimated and actual instantaneous frequencies while favouring the sparsity of the estimated TF representation.
 
@@
 \begin{figure}
-\includegraphics[width=.48\textwidth]{THOsig.eps}
 \includegraphics[width=.48\textwidth]{THOforecast.eps}
-\caption{Zoom on the respiratory signal. Top: original measured signal. Extended signal obtained by {\sf SigExt} forecasting superimposed with the ground truth signal.}
+\caption{Extended THO signal (blue) obtained by the {\sf SigExt} forecasting (top), the EDMD forecasting (middle), and the GPR forecasting (bottom), superimposed with the ground truth signal (red dash).}
 \label{fig:tho}
@@
 These outliers are probably due to the presence of pulse in the respiratory signal.
-That ones make the AM-FM model temporarily irrelevant, and break the validity of the linear dynamical model~\eqref{eq:dyn.model} used to extend the signal.
-That is why {\sf SigExt} is less robust to the vanishing of oscillations that the GPR or EDMD extensions.
+{\color{red}[HT: plot histogram.
+Show a failure case.]} Those cases make the AHM model temporarily irrelevant, and break the validity of the linear dynamical model~\eqref{eq:dyn.model} used to extend the signal.
+We see that {\sf SigExt} is less robust to the vanishing of oscillations that the GPR or EDMD extensions.
 
@@
    \hhline{|=#=|=|}
-   {\sf SigExt} & $1.704 \times 10^{1}$ & $7.346 \times 10^{-1}$ \\
+   {\sf SigExt} & $1.704 \times 10^{0}$ & $7.346 \times 10^{-1}$ \\
+   \hline
+   Symmetric & ${\color{red}X}$ & ${\color{red}X}$ \\
    \hline
@@
 
-Then, we apply the boundary effects reduction method {\sf BoundEffRed} (Algorithm~\ref{alg:boundary}) on these extensions for diverse time-frequency representations: STFT, SST, RS, as well as ConceFT, an SST-based representation introduced in~\cite{Daubechies16conceft}.
+Then, we apply the boundary effects reduction method {\sf BoundEffRed} (Algorithm~\ref{alg:boundary}) on these extensions for diverse TF representations: STFT, SST, RS, as well as concentration of frequency and time (ConceFT), an generalized multitaper SST-based representation introduced in~\cite{Daubechies16conceft}.
 In Table~\ref{tab:otd.tho}, the results are compared in terms of OTD.
-They are also compared with the strategy consisting in a zero-padding extension of the signal.
-Even though the performance of the forecasting Algorithm~\ref{alg:extension} is somehow moderate, the boundary effects can be reduced dramatically on the time-frequency representations.
-Notice that the extension length $L$ has been set accordingly to the window length used by the time-frequency analysis tool.
-For instance, here the window length we use to evaluate the STFT is of 1500 samples.
+{\color{red}[HT: make clear where you do the comparison.
+On the whole signal or only near the boundary.]} They are also compared with the strategy consisting in a zero-padding extension of the signal.
+Even though the performance of the forecasting Algorithm~\ref{alg:extension} is somehow moderate, the boundary effects can be reduced dramatically on the TF representations.
+Notice that the extension length $L$ has been set accordingly to the window length used by the TF analysis tool.
+For instance, the window length we use to evaluate the STFT is of 1500 samples.
 To prevent the STFT from being sensitive to the boundary effects, we set $L=750$.
 In this way, when evaluating the spectral content of the signal near its boundaries, the analysis is not limited by a lack of information all along the window support.
-From now on, all results are given for $L$ at equal to the half of the width of the window used in the time-frequency transform.
+From now on, all results are given for $L$ at equal to the half of the width of the window used in the TF transform.
 
@@
   \hline
-   \multirow{2}{35pt}{\centering Extension method} & \multicolumn{4}{c|}{Time-Frequency Representation} \\
+   \multirow{2}{35pt}{\centering Extension method} & \multicolumn{4}{c|}{TF Representation} \\
    \cline{2-5}
@@
    {\sf SigExt} & $1.72\!\times\! 10^{-2}$ & $4.00\!\times\! 10^{-3}$ & $2.43\!\times\! 10^{-2}$ & $1.12\!\times\! 10^{-2}$ \\
+   \hline
+   Symmetric & ${\color{red}X}$ & ${\color{red}X}$ & ${\color{red}X}$ & ${\color{red}X}$\\
    \hline
@@
 \label{ssse:ppg}
-We perform a study similar to the previous one on a 640 second-long photoplethysmogram (PPG) signal, sampled at $\fs=125$~Hz.
+We perform a study similar to the previous one on a 640 second-long photoplethysmogram (PPG) signal extracted from the Physionet dataset~\cite{Pimentel17toward, Goldberger00physiobank}, sampled at $\fs=125$~Hz.
 A 32 second-long piece of this signal is displayed on the top of Fig.~\ref{fig:ppg}.
@@
 \begin{figure}
-\includegraphics[width=.48\textwidth]{PPGsig.eps}
+%\includegraphics[width=.48\textwidth]{PPGsig.eps}
 \includegraphics[width=.48\textwidth]{PPGforecast.eps}
-\caption{PPG signal. Top: original measured signal. Extended signal obtained by forecasting superimposed with the ground truth signal.}
+\caption{Extended PPG signal (blue) obtained by the {\sf SigExt} forecasting (top), the EDMD forecasting (middle), and the GPR forecasting (bottom), superimposed with the ground truth signal (red dash).}
 \label{fig:ppg}
@@
 We divide the signal into 32-second long pieces, and apply Algorithm~\ref{alg:boundary} on each piece.
-We provide in Table~\ref{tab:otd.ppg} the OTD to the optimal time-frequency representation averaged over the signals.
-For all the considered time-frequency representations, the results clearly shows that our algorithm reduce the influence of the boundary effects.
+We provide in Table~\ref{tab:otd.ppg} the OTD to the optimal TF representation averaged over the signals.
+For all the considered TF representations, the results clearly shows that our algorithm reduce the influence of the boundary effects.
 This highlights the ability of our approach to limit the distortion due the boundary effects and provide a more accurate representations.
-Even though the {\sc SigExt} extension yields time-frequency  representations slightly more sensitive to boundary effects that the extensions given by EDMD or GPR, it is the only technique that allows a real-time implementation.
+Even though the {\sf SigExt} extension yields TF representations slightly more sensitive to boundary effects than the extensions given by EDMD or GPR, it is the only technique that allows a real-time implementation.
 
@@
   \hline
-   \multirow{2}{*}{Extension method} & \multicolumn{3}{c|}{Time-Frequency Representation} \\
+   \multirow{2}{*}{Extension method} & \multicolumn{3}{c|}{TF Representation} \\
    \cline{2-4}
@@
    {\sf SigExt} & $1.22\times 10^{-3}$ & $7.31\times 10^{-2}$ & $1.11\times 10^{-1}$ \\
+   \hline
+   Symmetric & ${\color{red}X}$ & ${\color{red}X}$ & ${\color{red}X}$ \\
    \hline
@@
 
-On Fig.~\ref{fig:ppg.boundeffred}, we display the SST resulting from the {\sf BoundEffRed} strategy, applied to the same portion of PPG than what is used to display Fig.~\ref{fig:ppg}.
+On the bottom panel of Fig.~\ref{fig:ex.intro}, we display the SST resulting from the {\sf BoundEffRed} strategy, applied to the same portion of PPG than what is used to display Fig.~\ref{fig:ppg}.
 We clearly observe an improvement of the quality of the SST near boundaries.
-Indeed, the blurring visible when zooming on the right boundary of the SST has almost vanished (see bottom of Fig.~\ref{fig:ppg.boundeffred}).
+Indeed, the blurring visible when zooming on the right boundary of the SST has almost vanished.
 The real-time tracking of the instantaneous frequencies contained in the measured signal is therefore largely facilitated.
+
+%\begin{figure}
+%\centering
+%\includegraphics[width=.48\textwidth]{SSTBoundEffRed.eps}
+%\includegraphics[width=.48\textwidth]{zoomSSTBoundEffRed.eps}
+%\caption{Result of {\sf BoundEffRed} on the synchrosqueezing transform of a PPG (top) with a zoom on its right boundary (bottom). }
+%\label{fig:ppg.boundeffred}
+%\end{figure}
+
+To evaluate the influence of the noise level on the performance of {\sf BoundEffRed}, we artificially add a Gaussian noise to the measured PPG signal.
+It is thus an additional noise to the measurement noise actually contained in the signal.
+Fig.~\ref{fig:otd.noise} shows the averaged OTD of {\sf BoundEffRed} for different values of the Signal to Noise Ratio (SNR).
+We notice that STFT is slightly more sensitive to noise than SST or RS, and STFT performs the best when the added noise is small, \ie~when the SNR is great.
+%On one hand, the robustness to noise of the SST and the RS is the direct consequence of the approximation result of~\cite{Daubechies16conceft}, discussed in section~\ref{sse:perf.BoundEffRed}.
+On the other hand, the sensitivity to noise of the STFT may be interpreted as the inability to forecast the values of the TF coefficients where only noise is active, while sharp TF representations, such as SST or RS, ideally vanish at these points.
 
@@
 \centering
-\includegraphics[width=.48\textwidth]{SSTBoundEffRed.eps}
-\includegraphics[width=.48\textwidth]{zoomSSTBoundEffRed.eps}
-\caption{Result of {\sf BoundEffRed} on the synchrosqueezing transform of a PPG (top) with a zoom on its right boundary (bottom). }
-\label{fig:ppg.boundeffred}
-\end{figure}
-
-To evaluate the influence of the noise level on the performance of {\sf BoundEffRed}, we artificially add a Gaussian noise to the measured PPG signal, whose variance is denoted $\sigma_{\mathrm{add}}^2$.
-It is thus an additional noise to the measurement noise actually contained in the signal.
-Fig.~\ref{fig:otd.noise} shows the averaged OTD of {\sf BoundEffRed} for different values of $\sigma_{\mathrm{add}}^2$.
-At first glance, we notice that STFT is slightly more sensitive to noise than SST or RS.
-On one hand, the robustness to noise of the SST and the RS is the direct consequence of the approximation result of~\cite{Daubechies16conceft}, discussed in section~\ref{sse:perf.BoundEffRed}.
-On the other hand, the sentivity to noise of the STFT may be interpreted as the inability to forecast the values of the time-frequency coefficients where only noise is active, while sharp time-frequency representations, such as SST or RS, ideally vanish at these points.
-\begin{figure}
-\centering
 \includegraphics[width=.48\textwidth]{OTDBoundEffRed.eps}
-\caption{PPG signal. Averaged OTD of {\sf BoundEffRed} in function of the noise variance.}
+\caption{PPG signal. Averaged OTD of {\sf BoundEffRed} in function of the SNR.}
 \label{fig:otd.noise}

### 2020-07-02 Scripts/BoundEffRedExamples/PPGexample.m

@@ -151,20 +151,20 @@ figure;
 subplot(2,2,1);
 [WxxZP,~] = display_cwt_JEFAS(xxZP,fs,fmin,fmax,Ms,wav_typ,wav_par) ;
 [Wx,~] = display_cwt_JEFAS(x,fs,fmin,fmax,Ms,wav_typ,wav_par) ;
-xlabel('Time (s)'); ylabel('Frequency (Hz)'); title('SST on original short signal');
+xlabel('Time (s)'); ylabel('Frequency (Hz)'); title('CWT on original short signal');
 % On the true extended signal
 subplot(2,2,2);
 [WxxTRUE,~] = display_cwt_JEFAS(xxTRUE,fs,fmin,fmax,Ms,wav_typ,wav_par) ;
-xlabel('Time (s)'); ylabel('Frequency (Hz)'); title('SST on original long signal');
+xlabel('Time (s)'); ylabel('Frequency (Hz)'); title('CWT on original long signal');
 % On the estimated extended signal
 subplot(2,2,3);
 [WxxLSE,~] = display_cwt_JEFAS(xxLSE,fs,fmin,fmax,Ms,wav_typ,wav_par) ;
-xlabel('Time (s)'); ylabel('Frequency (Hz)'); title('SST on estimated long signal (short signal extended)');
+xlabel('Time (s)'); ylabel('Frequency (Hz)'); title('CWT on estimated long signal (short signal extended)');
 % On the estimated extended signal
 subplot(2,2,4);
 [WxxGPR,~] = display_cwt_JEFAS(xxGPR,fs,fmin,fmax,Ms,wav_typ,wav_par) ;
 [WxxEDMD,~] = display_cwt_JEFAS(xxEDMD,fs,fmin,fmax,Ms,wav_typ,wav_par) ;
-xlabel('Time (s)'); ylabel('Frequency (Hz)'); title('SST on estimated long signal (short signal extended)');
+xlabel('Time (s)'); ylabel('Frequency (Hz)'); title('CWT on estimated long signal (short signal extended)');
 
 %save('results','tfrsq3','tfrsq3EXT');
 

### 2020-07-02 Scripts/EvalPerfBoundEffRed/EvalPPGnoise.m

@@ -38,4 +38,6 @@ for n = 1:nXP
     NoiseTFR.RS(n) = mean(OTD.rs.LSE) ;
 end
 
-save('../../Results/resultSucForPPGnoise','StdNoise','NoiseForecastErr','NoiseTFR');
+SNR = 20*log10(std(x0)./StdNoise) ;
+
+save('../../Results/resultSucForPPGnoise','SNR','NoiseForecastErr','NoiseTFR');

### 2020-07-02 Scripts/EvalPerfBoundEffRed/EvalTHO.m

@@ -12,7 +12,7 @@ fs = 100 ; % sampling frequency
 N = 6e3; % subsignals length
 
 % Forecasting parameters
-methods = {'lse','edmd','gpr'} ;
+methods = {'lse','symmetrization'} ;%,'edmd','gpr'} ;
 HOP = 1 ;
 extSEC = 7 ; % the extension is of extSEC second
 L = round(extSEC*fs) ;
@@ -26,4 +26,4 @@ basicTF.fmin = 0 ;
 basicTF.fmax = 0.01 ;
 [forecastErr, CompTime, OTD] = SucForecast(xtot,fs,methods,HOP,N,extM,extK,extSEC,TFR,basicTF) ;
 
-save('../../Results/resultSucForTHOcompMeth','forecastErr','CompTime','OTD');
+% save('../../Results/resultSucForTHOcompMeth','forecastErr','CompTime','OTD');

### 2020-07-02 Scripts/EvalPerfBoundEffRed/ResultsBoundEffRed.m

@@ -6,14 +6,15 @@ fprintf('================================================================\n')
 fprintf('                              PPG                               \n')
 fprintf('================================================================\n')
 
-fprintf('===================Forecasting Performance======================\n')
-fprintf(' ______________________________________________________________\n')
-fprintf('| Extension Method | Computing time (sec.) | Forecasting Error |\n')
-fprintf('|------------------|-----------------------|-------------------|\n')
-fprintf('|      LSE         |           %.2f        |     %.3e     |\n', CompTime.LSE , mean(forecastErr.LSE))
-fprintf('|      EDMD        |           %.2f        |     %.3e     |\n', CompTime.EDMD , mean(forecastErr.EDMD))
-fprintf('|      GPR         |         %.2f        |     %.3e     |\n', CompTime.GPR , mean(forecastErr.GPR))
-fprintf('|__________________|_______________________|___________________|\n\n')
+fprintf('====================Forecasting Performance======================\n')
+fprintf(' _______________________________________________________________\n')
+fprintf('| Extension Method | Computing time (sec.)  | Forecasting Error |\n')
+fprintf('|------------------|------------------------|-------------------|\n')
+fprintf('|      LSE         |           %.3f        |     %.3e     |\n', CompTime.LSE , mean(forecastErr.LSE) )
+fprintf('|  Symmetrization  |           %.3f        | %.3e | %.3e |\n', CompTime.SYM , mean(forecastErr.SYM) )
+fprintf('|      EDMD        |           %.3f        |     %.3e     |\n', CompTime.EDMD , mean(forecastErr.EDMD) )
+fprintf('|      GPR         |         %.3f        |     %.3e     |\n', CompTime.GPR , mean(forecastErr.GPR) )
+fprintf('|__________________|________________________|___________________|\n\n')
 
 fprintf('==============Optimal Transport Distance===============\n')
 fprintf(' ______________________________________________________\n')
@@ -21,6 +22,7 @@ fprintf('| Extension Method |    SST    |    STFT   |    RS     |\n')
[…]

### 2020-07-02 Scripts/EvalPerfBoundEffRed/SucForecast.m

@@ -33,7 +33,15 @@ while nend <= Ntot
         tic;
         xxLSE = SigExtension(x,fs,HOP,extK,extM,extSEC,method) ; % Forecasted signal via LSE
         LSEtime(ind) = toc;
-        forecastErr.LSE(ind) = sqrt( (1/(2*L)) * sum( abs(xxLSE - xxTRUE).^2 ) ) ;
+        forecastErr.LSE(ind) =  (1/(2*L)) * sqrt( sum( abs(xxLSE - xxTRUE).^2 ) ) ;
+    end
+    
+    if any(strcmp(methods,'symmetrization'))
+        method.name = 'symmetrization' ;
+        tic;
+        xxSYM = SigExtension(x,fs,HOP,extK,extM,extSEC,method) ; % Forecasted signal via LSE
+        SYMtime(ind) = toc;
+        forecastErr.SYM(ind) =  (1/(2*L)) * sqrt( sum( abs(xxSYM - xxTRUE).^2 ) ) ;
     end
     
     if any(strcmp(methods,'edmd'))
@@ -42,7 +50,7 @@ while nend <= Ntot
         tic; 
         xxEDMD = SigExtension(x,fs,HOP,extK,extM,extSEC,method); 
         EDMDtime(ind) = toc;
-        forecastErr.EDMD(ind) = (1/(2*extSEC*fs)) * sqrt(sum(abs(xxEDMD - xxTRUE).^2)) ;
+        forecastErr.EDMD(ind) = (1/(2*L)) * sqrt(sum(abs(xxEDMD - xxTRUE).^2)) ;
     end
     
     if any(strcmp(methods,'gpr'))
@@ -50,11 +58,11 @@ while nend <= Ntot
         tic; 
         xxGPR = SigExtension(x,fs,HOP,extK,extM,extSEC,method); 
         GPRtime(ind) = toc;
-        forecastErr.GPR(ind) = (1/(2*extSEC*fs)) * sqrt(sum(abs(xxGPR - xxTRUE).^2)) ;
+        forecastErr.GPR(ind) = (1/(2*L)) * sqrt(sum(abs(xxGPR - xxTRUE).^2)) ;
     end
     
     xxZP = [zeros(L,1); x; zeros(L,1)];
-    forecastErr.ZP(ind) = sqrt( (1/(2*L)) * sum( abs(xxZP - xxTRUE).^2 ) ) ;
[…]

### 2020-07-02 Scripts/EvalPerfBoundEffRed/figPPGnoise.m

@@ -2,13 +2,13 @@ clear all; close all; clc;
 load ../../Results/resultSucForPPGnoise ;
 
 figure;
-loglog(StdNoise.^2,NoiseForecastErr,'+','linewidth',2);
-xlabel('Noise Variance \sigma_{add}^2'); ylabel('Forecasting Variance');
+semilogy(SNR,NoiseForecastErr,'+','linewidth',2);
+xlabel('SNR (dB)'); ylabel('Forecasting Variance');
 grid on;
 set(gca,'fontsize',24) ;
 
 figure;
-loglog(StdNoise.^2,NoiseTFR.SST,'+',StdNoise.^2,NoiseTFR.STFT,'^',StdNoise.^2,NoiseTFR.RS,'o','linewidth',2,'MarkerSize',8);
-xlabel('Noise Variance \sigma_{add}^2'); ylabel('Averaged OTD'); grid on;
-legend({'SST','RS','STFT'},'location','southeast');
+semilogy(SNR,NoiseTFR.STFT,'^',SNR,NoiseTFR.SST,'+',SNR,NoiseTFR.RS,'o','linewidth',2,'MarkerSize',8);
+xlabel('SNR (dB)'); ylabel('Averaged OTD'); grid on;
+legend({'STFT','SST','RS'},'location','southwest');
 set(gca,'fontsize',24) ;

### 2020-07-02 Scripts/EvalPerfForecasting/ResultsCompPerfAHM.m

@@ -6,17 +6,19 @@ VarianceXP.TBATS = dataTBATS(:,2) ;
 CPUtimeXP.TBATS = dataTBATS(~isnan(dataTBATS(:,3)),3) ;
 
 MSE.LSE = BiasXP.LSE.^2 + VarianceXP.LSE ;
+MSE.SYM = BiasXP.SYM.^2 + VarianceXP.SYM ;
 MSE.EDMD = BiasXP.EDMD.^2 + VarianceXP.EDMD ;
 MSE.GPR = BiasXP.GPR.^2 + VarianceXP.GPR ;
 MSE.TBATS = BiasXP.TBATS.^2 + VarianceXP.TBATS ;
 
 fprintf('===================Forecasting Performance=========\n')
-fprintf(' _________________________________________________\n')
-fprintf('| Extension |  Computing  |   Forecasting Error   |\n')
-fprintf('|  Method   | time (sec.) |    Mean   |  Variance |\n')
-fprintf('|-----------|-------------|-----------|-----------|\n')
-fprintf('|   LSE     |     %.2f    | %.3e | %.3e |\n', CPUtimeXP.LSE , median(MSE.LSE), var(MSE.LSE))
-fprintf('|   EDMD    |     %.2f    | %.3e | %.3e |\n', CPUtimeXP.EDMD , median(MSE.EDMD), var(MSE.EDMD))
-fprintf('|   GPR     |   %.2f    | %.3e | %.3e |\n', CPUtimeXP.GPR , median(MSE.GPR), var(MSE.GPR))
-fprintf('|   TBATS   |   %.2f    | %.3e | %.3e |\n', CPUtimeXP.TBATS , median(MSE.TBATS), var(MSE.TBATS))
-fprintf('|___________|_____________|___________|___________|\n\n')
+fprintf(' ______________________________________________________\n')
+fprintf('|    Extension   |  Computing  |   Forecasting Error   |\n')
+fprintf('|     Method     | time (sec.) |    Mean   |  Variance |\n')
+fprintf('|----------------|-------------|-----------|-----------|\n')
+fprintf('|      LSE       |     %.3f   | %.3e | %.3e |\n', CPUtimeXP.LSE , mean(MSE.LSE), var(MSE.LSE))
+fprintf('| Symmetrization |     %.3f   | %.3e | %.3e |\n', CPUtimeXP.SYM , mean(MSE.SYM), var(MSE.SYM))
[…]

### 2020-07-02 TimeFrequencyScaleRep/ConceFT_rsSTFT.m

@@ -22,7 +22,7 @@ N = length(x) ;
 %=======================================
 
 fprintf(['Run ordinary RS\n']) ;
-[tfr, tfrtic, tfrrs, tfrrstic] = rsSTFTbase(x, lowFreq, highFreq, alpha, 1, h(1,:)', Dh(1,:)', 1);
+[tfr, tfrtic, tfrrs, tfrrstic] = rsSTFTbase(x, lowFreq, highFreq, alpha, hop, h(1,:)', Dh(1,:)', 1);
 
 
 %=======================================
@@ -41,7 +41,7 @@ if MT > 1
 		rh = rv * h ; 
 		rDh = rv * Dh ;
 
-		[~, ~, tfrrs, tfrrstic] = rsSTFTbase(x, lowFreq, highFreq, alpha, 1, rh', rDh', 1);
+		[~, ~, tfrrs, tfrrstic] = rsSTFTbase(x, lowFreq, highFreq, alpha, hop, rh', rDh', 1);
 
 	 	ConceFT = ConceFT + tfrrs ;
     end

### 2020-07-10 Notes/BoundEffRed.tex
code and tex

@@
 \usepackage{array}
+\usepackage{bigints}
 
@@
 
-\title{An Efficient Forecasting Approach for the Reduction of Boundary Effects in Time-Frequency Representations}
+\title{An Efficient Forecasting Approach for the Real-Time Reduction of Boundary Effects in Time-Frequency Representations}
 
@@
 Various improvements can be considered to make the algorithm more robust to certain situations.
-In particular, we have noticed that when the observed signal does not oscillate {\color{red}[HT: we need to confirm this statement]}, the forecasting is no longer relevant, and the TF representations are not ideal.
+In particular, we have noticed (see Fig.~\ref{fig:THO.failure}) that when the regular oscillations of the observed signal break, the forecasting is no longer relevant, and the TF representations are not ideal.
 A preliminary step should then be added to the algorithm to select the portions without sufficient quality, in order to reject the portions for which a better forecasting scheme is needed.

### 2020-07-10 Notes/algorithm.tex
code and tex

@@
 We coined the algorithm {\sf BoundEffRed}.
-Below, we detail the algorithm, particularly the Signal extension algorithm {\sf SigExt} in Algorithm~\ref{alg:boundary}.
+Below, we detail the algorithm, particularly the Signal extension algorithm {\sf SigExt} in Algorithm~\ref{alg:extension}.
 
 \begin{algorithm}
-\caption{Tackling boundary effects. $\bF_\bx = \mbox{\sf BoundEffRed}(\bx,M,K,L,\ccF)$}
+\caption{Tackling boundary effects of a TF representation in real-time. $\bF_\bx = \mbox{\sf BoundEffRed}(\bx,M,K,L,\ccF)$}
 \label{alg:boundary}
 \begin{algorithmic}
-\STATE {\bf Inputs}: $\bx$, $M$, $K$, $L$, $\ccF$
+\STATE {\bf Inputs}: $\bx$, $M$, $K$, $N_0$, $L$, $\ccF$
 \STATE
+\WHILE {$N$ increases}
 \STATE {\bf Forecasting step.}
-\STATE \quad\textbullet\ Signal extension: $\tilde\bx = \mbox{\sf SigExt}(\bx)$.
+\STATE \quad\textbullet\ Signal extension: $\tilde\bx = \mbox{\sf SigExt}(\bx_N)$.
 \STATE 
@@
 \STATE \quad\textbullet\ Representation evaluation: $\ccF_{N+2L}(\tilde\bx)$.
-\STATE \quad\textbullet\ Restriction of $\ccF_{N+2L}(\tilde\bx)$ to the central time interval (see~\eqref{eq:restriction}) to obtain $\bF_\bx=\ccF_N^{\mathrm{ext}}(\bx)$.
+\STATE \quad\textbullet\ Restriction of $\ccF_{N+2L}(\tilde\bx)$ to the central time interval (see~\eqref{eq:restriction}) to obtain $\bF_\bx^{(N)}=\ccF_N^{\mathrm{ext}}(\bx_N)$.
 \STATE 
-\STATE {\bf Output}: Signal representation $\bF_\bx$
+\STATE {\bf Real-time output}: Signal representation $\bF_\bx^{(N)}$
+\STATE $N \leftarrow N+1$
+\ENDWHILE
 \end{algorithmic}
@@
 Consequently, we have:
-\begin{equation*}
+\begin{equation}
 \ccF_N^{\mathrm{ext}}(\bx) = \cR\left( \ccF_{N+2L}(\tilde \bx) \right) \ .
-\end{equation*}
+\label{eq:bound.free.TFR}
+\end{equation}
 We call $\ccF_N^{\mathrm{ext}}$ the {\em boundary-free TF representation}.

### 2020-07-10 Notes/results.tex
code and tex

@@
 %
-{\color{red} [HT: need a more convincing explanation.] It should be noted that, contrary to what expression~\eqref{eq:cov.error.2} suggests, smaller values of $\sigma$ do not cause a decrease of the experimental variance, but an increase.
-[HT: strange.]} This comes from the fact that when $\sigma$ is small, the matrix $\bX\bX^T$ becomes ill-conditioned.
+{\color{red} [HT: need a more convincing explanation.] } 
+%It should be noted that, contrary to what expression~\eqref{eq:cov.error.2} suggests, smaller values of $\sigma$ do not cause a decrease of the experimental variance, but an increase.
+This comes from the fact that when $\sigma$ is small, the matrix $\bX\bX^T$ becomes ill-conditioned.
 The calculation of the forecasting matrix $\tilde\bA$ in~\eqref{eq:lse} is then strongly disturbed.
+
 \begin{figure}
@@
 \begin{align*}
-\phi_1[n] &= \frac{p_1}{M}\left( n + \frac{0.01}{2\pi}\cos\left(2\pi\frac{n}{N}\right) \right) \\
+\phi_1[n] &= \frac{p_1}{M}\left( n + \frac{0.01}{2\pi}\cos\left(2\pi\frac{n}{N}\right) \right) \\[-1mm]
 \phi_2[n] & = p_2\frac{n}{M} + \frac{20}{2N\fs}n^2
@@
 \label{eq:mse}
-\mathrm{MSE_{\xp}}(\tilde\bx) &= \dfrac1{L}\|\tilde\bx -\bx^\mathrm{ext}\|^2 \\
+\mathrm{MSE_{\xp}}(\tilde\bx) &= \dfrac1{L}\|\tilde\bx -\bx^\mathrm{ext}\|^2 \\[-1mm]
 \nonumber
@@
 \subsubsection{Metrics}
-The quality of the boundary effects reduction must be evaluated directly on the TF representation.
-To that aim, we compare the obtained representation to the optimal representation $\ccF_N^\mathrm{opt}(\bx)$.
-The optimal representation is defined as the restriction of the representation of the ground-truth extended signal $\bx^\mathrm{ext}$.
+The quality of the boundary effects reduction is evaluated directly on the TF representation.
+To that aim, we compare the obtained representation to the optimal representation $\ccF_N^\mathrm{opt}(\bx)$, defined as the restriction of the representation of the ground-truth extended signal $\bx^\mathrm{ext}$.
 Therefore, we have:
-\begin{equation*}
+\begin{equation}
 \ccF_N^\mathrm{opt}(\bx) = \cR\left( \ccF_{N+2L}(\bx^\mathrm{ext}) \right) \ .
-\end{equation*}
+\label{eq:opt.TFR}
+\end{equation}
 
@@
 Let us generically denote a time frequency representation $\ccQ$.
-Then, for $t$ fixed, we consider the following probability density function:
-\begin{align}
-p_\ccQ^t(\xi) &= \dfrac{|\ccQ(\xi,t)|^2}{\int_\RR |\ccQ(\nu,t)|^2\dd\nu}\ .
-\label{eq:ppdf}
-\end{align}
-Then, at each instant $t$, we can then determine the optimal transport distance $d_{t}$ between the two densities.
+Then, for $t$ fixed, we consider the following probability density function: $p_\ccQ^t(\xi) = \left.
+|\ccQ(\xi,t)|^2 \middle/ \int_\RR |\ccQ(\nu,t)|^2\dd\nu \right.$.
+At each instant $t$, we can then determine the optimal transport distance $d_{t}$ between the two densities.
 It is given by the $L^1$ norm of the difference between the associated distribution functions.
@@
 where $P_{\ccQ}^t(\xi)=\int_{-\infty}^\xi p_{\ccQ}^t(\nu)\dd\nu$ and $\tilde P_{\ccF_0}^t(\xi)=\int_{-\infty}^\xi\tilde p_{\ccF_0}^t(\nu)\dd\nu$.
-Finally, the distance between the two TF representations is obtained by averaging all the optimal transport distances with respect to time:
+The \textit{optimal transport distance} quantifies the proximity between the estimated and actual instantaneous frequencies while favoring the sparsity of the estimated TF representation.
+That is why, the performance index $D(\ccQ)$ for the reduction of boundary effects of a given TF representation $\ccQ$ is given by the averaged ratio between the optimal transport distance to the optimal TF representation~\eqref{eq:opt.TFR} and the optimal transport distance of to the optimal TF representation, that is:
 \begin{equation}
-D(\ccQ,\ccF_0) = \frac1{|I|}\int_I d_{t}(\ccQ,\ccF_0)\dd t\ .
+D(\ccQ) = \dfrac1{|I|}\bigintss_I\ \frac{d_{t}\left(\ccQ,\ccF_N^\mathrm{opt}\right)}{d_{t}\left(\ccF_N,\ccF_N^\mathrm{opt}\right)}\ \dd t\ .
+\label{eq:index.perf}
 \end{equation}
-The \textit{Optimal Transport Distance} (OTD) quantifies the proximity between the estimated and actual instantaneous frequencies while favouring the sparsity of the estimated TF representation.
-
-Let us evaluate the quality of the boundary effects reduction on biomedical signals.
+Thus, $D(\ccQ)<1$ means a reduction of the boundary effects. Let us evaluate the quality of the boundary effects reduction on biomedical signals.
 
 \subsubsection{Respiratory signal}
-We first consider a 50 minutes-long respiratory signal. This signal is sampled at $\fs=100$~Hz. A zoom on a small portion of the signal is displayed in Fig.~\ref{fig:tho}.
+We first consider a 6 hours 20 minutes-long respiratory signal. This signal is sampled at $\fs=100$~Hz. A zoom on a small portion of the signal is displayed in Fig.~\ref{fig:tho}.
 
@@
 
-From that large signal, we build a dataset of $48$ non-overlapping signals of 60 seconds, \ie~$N=6000$.
+From that large signal, we build a dataset of $378$ non-overlapping signals of 60 seconds, \ie~$N=6000$.
 On each of these pieces of signal, we implement the forecasting method introduced in section~\ref{ssse:res.ahm}, including the {\sf SigExt} method detailed in Algorithm~\ref{alg:extension}.
@@
 As a result of section~\ref{ssse:res.sine}, we take: $K=\lfloor2.5M\rfloor$.
-The average and the median of the MSE~\eqref{eq:mse} with respect to the simulations are given in Table~\ref{tab:mse}.
-The averaged MSE of {\sf SigExt} method is overwhelmingly higher than the averaged MSE of the other methods.
-This result is caused by the presence of outliers in some extensions by {\sf SigExt}.
-Indeed, the value median of the MSE is of the same order than those of the other methods.
-These outliers are probably due to the presence of pulse in the respiratory signal.
-{\color{red}[HT: plot histogram.
-Show a failure case.]} Those cases make the AHM model temporarily irrelevant, and break the validity of the linear dynamical model~\eqref{eq:dyn.model} used to extend the signal.
-We see that {\sf SigExt} is less robust to the vanishing of oscillations that the GPR or EDMD extensions.
+The resulting MSE~\eqref{eq:mse} averaged with respect to the simulations are given in Table~\ref{tab:THO}.
+The MSE of {\sf SigExt} is higher than the MSE of the other methods.
+This is mainly caused by the presence of a few segments of the whole signal involving complicated behaviors, that are unpredictable via a too simple dynamical model like~\eqref{eq:dyn.model}.
+The left of Fig.~\ref{fig:THO.failure} illustrates one of those cases, where {\sf SigExt} fails to catch the fast varying dynamic of the instantaneous amplitude to satisfactorily forecast the signal.
+The EDMD an Symmetric extensions are more robust to those situations, as shown on this example and in Table~\ref{tab:THO}.
+Nevertheless, {\sf SigExt} provides a sufficiently relevant extension to give TF representations sparingly affected by boundary effects.
+On the right of Fig.~\ref{fig:THO.failure}, we display comparison between the right boundary of SST of the same segment of signal (top-right), and its boundary-free SST obtained after the {\sf SigExt} forecasting (bottom-right).
+The extension of the instantaneous frequency visible on the right side of the image, illustrates the reduction of boundary effects produced despite an inaccurate signal forecasting.
+
+We then apply {\sf BoundEffRed} (\ie~Algorithm~\ref{alg:boundary}) for diverse TF representations: STFT, SST, RS, as well as concentration of frequency and time (ConceFT), a generalized multitaper SST-based representation introduced in~\cite{Daubechies16conceft}.
+In Table~\ref{tab:THO}, we give the averaged performance index~\eqref{eq:index.perf}, evaluated the whole TF representations (including  boundaries).
+Even though {\sf SigExt} performs somehow moderately, the boundary effects are dramatically reduced on the TF representations, in the same order of magnitude than with the forecastings given by EDMD or GPR.
+Notice that the extension length $L$ has been set accordingly to the window length used by the TF analysis tool.
+For instance, the window length used to evaluate the STFT is of $1500$ samples.
+To prevent the STFT from being sensitive to the boundaries, we set $L=750$.
+In this way, the evaluation of the spectral content of the signal near its boundaries is not limited by a lack of information all along the window support.
+From now on, all results are given for $L$ equal to the half of the width of the window used in the TF transform.
 
@@
 \centering
-\caption{Respiratory signal. MSE of the respiratory signal extensions.}
-\begin{tabular}{|c||c|c|}
+\caption{Performance of the extension methods on a respiratory signal.}
+\begin{tabular}{|c||c||c|c|c|c|}
   \hline
-   \multirow{2}{*}{Algorithm} & \multicolumn{2}{c|}{MSE} \\
-   \cline{2-3}
-      & Mean & Median\\
-   \hhline{|=#=|=|}
-   {\sf SigExt} & $1.704 \times 10^{0}$ & $7.346 \times 10^{-1}$ \\
+   \multirow{2}{*}{Algorithm} & \multirow{2}{35pt}{\centering Averaged MSE} & \multicolumn{4}{c|}{Averaged performance index $D$} \\
+   \cline{3-6}
+      & & STFT & SST & RS & ConceFT \\
+   \hhline{|=#=#=|=|=|=|}
+   {\sf SigExt} & $0.046$ & $0.705$ & $0.698$ & $0.694$ & $0.776$ \\
    \hline
-   Symmetric & ${\color{red}X}$ & ${\color{red}X}$ \\
+   Symmetric & $0.039$ & $1.280$ & $0.984$ & $9.063$ & $1.295$ \\
    \hline
-   EDMD & $2.172\times 10^{-2}$ & $7.349\times 10^{-1}$ \\
+   EDMD & $0.022$ & $0.757$ & $0.635$ & $0.701$ & $0.793$ \\
    \hline
-   GPR & $4.476\times 10^{-2}$ & $1.998\times 10^{-2}$ \\
+   GPR & $0.045$ & $0.817$ & $0.793$ & $0.727$ & $0.876$ \\ 
    \hline
 \end{tabular}
-\label{tab:mse}
+\label{tab:THO}
 \end{table}
 
-Then, we apply the boundary effects reduction method {\sf BoundEffRed} (Algorithm~\ref{alg:boundary}) on these extensions for diverse TF representations: STFT, SST, RS, as well as concentration of frequency and time (ConceFT), an generalized multitaper SST-based representation introduced in~\cite{Daubechies16conceft}.
-In Table~\ref{tab:otd.tho}, the results are compared in terms of OTD.
-{\color{red}[HT: make clear where you do the comparison.
-On the whole signal or only near the boundary.]} They are also compared with the strategy consisting in a zero-padding extension of the signal.
-Even though the performance of the forecasting Algorithm~\ref{alg:extension} is somehow moderate, the boundary effects can be reduced dramatically on the TF representations.
-Notice that the extension length $L$ has been set accordingly to the window length used by the TF analysis tool.
-For instance, the window length we use to evaluate the STFT is of 1500 samples.
-To prevent the STFT from being sensitive to the boundary effects, we set $L=750$.
-In this way, when evaluating the spectral content of the signal near its boundaries, the analysis is not limited by a lack of information all along the window support.
-From now on, all results are given for $L$ at equal to the half of the width of the window used in the TF transform.
-
-\begin{table}
+\begin{figure}
 \centering
-\caption{Respiratory signal. Averaged OTD of the boundary effects reduction methods on diverse representations.}
-\begin{tabular}{|>{\centering}m{35pt}||c|c|c|c|}
-  \hline
-   \multirow{2}{35pt}{\centering Extension method} & \multicolumn{4}{c|}{TF Representation} \\
-   \cline{2-5}
-      & STFT & SST & RS & ConceFT \\
-   \hhline{|=#=|=|=|=|}
-   Without extension & $2.16\!\times\! 10^{-2}$ & $5.26\!\times\! 10^{-3}$ & $3.07\!\times\! 10^{-2}$ & $1.41\!\times\! 10^{-2}$ \\
-   \hline
-   {\sf SigExt} & $1.72\!\times\! 10^{-2}$ & $4.00\!\times\! 10^{-3}$ & $2.43\!\times\! 10^{-2}$ & $1.12\!\times\! 10^{-2}$ \\
-   \hline
-   Symmetric & ${\color{red}X}$ & ${\color{red}X}$ & ${\color{red}X}$ & ${\color{red}X}$\\
-   \hline
-   EDMD & $1.75\!\times\! 10^{-2}$ & $4.23\!\times\! 10^{-3}$ & $2.45\!\times\! 10^{-2}$ & $1.09\!\times\! 10^{-2}$ \\
-   \hline
-   GPR & $1.84\!\times\! 10^{-2}$ & $4.10\!\times\! 10^{-3}$ & $2.48\!\times\! 10^{-2}$ & $1.24\!\times\! 10^{-2}$ \\
-   \hline
-\end{tabular}
-\label{tab:otd.tho}
-\end{table}
+\includegraphics[width=.48\textwidth]{THOfailure.eps}
+\caption{Extensions of a segment of the respiratory signal (left) where {\sf SigExt} is outperformed by the EDMD and Symmetric extensions.
+Corresponding SST (top-right) and boundary-free SST obtained with {\sf SigExt} (bottom-right).}
+\label{fig:THO.failure}
+\end{figure}
 
@@
 
-We divide the signal into 32-second long pieces, and apply Algorithm~\ref{alg:boundary} on each piece.
-We provide in Table~\ref{tab:otd.ppg} the OTD to the optimal TF representation averaged over the signals.
-For all the considered TF representations, the results clearly shows that our algorithm reduce the influence of the boundary effects.
-This highlights the ability of our approach to limit the distortion due the boundary effects and provide a more accurate representations.
-Even though the {\sf SigExt} extension yields TF representations slightly more sensitive to boundary effects than the extensions given by EDMD or GPR, it is the only technique that allows a real-time implementation.
+We divide the signal into 32-second long segments, and apply Algorithm~\ref{alg:boundary} on each piece.
+We provide in Table~\ref{tab:otd.ppg} the performance index $D$ of the boundary-free TF representations averaged over the signals.
+For all the considered TF representations, the results clearly shows that our algorithm reduces the boundary effects.
+%This highlights the ability of our approach to limit the distortion due the boundary effects and provide a more accurate representations.
+Even though, on this signal, the {\sf SigExt} extension yields TF representations slightly more sensitive to boundary effects than the extensions given by EDMD or GPR, it is the only technique that allows a real-time implementation.
 
@@
 \centering
-\caption{PPG signal. Averaged OTD of the boundary effects reduction methods on diverse representations.}
+\caption{PPG signal: averaged performance index of the boundary-free TF representations for different extensions.}
 \begin{tabular}{|c||c|c|c|}
   \hline
-   \multirow{2}{*}{Extension method} & \multicolumn{3}{c|}{TF Representation} \\
+   \multirow{2}{*}{Extension method} & \multicolumn{3}{c|}{Performance index $D$} \\
    \cline{2-4}
-      & STFT & SST & RS\\
+      & STFT & SST & ConceFT\\
    \hhline{|=#=|=|=|}
-   Without extension & $2.52\times 10^{-2}$ & $9.41\times 10^{-2}$ & $1.03\times 10^{-1}$ \\
+%   Without extension & $2.52\times 10^{-2}$ & $9.41\times 10^{-2}$ & $1.03\times 10^{-1}$ \\
+%   \hline
+   {\sf SigExt} & $0.774$ & $0.048$ & $X$ \\
    \hline
-   {\sf SigExt} & $1.22\times 10^{-3}$ & $7.31\times 10^{-2}$ & $1.11\times 10^{-1}$ \\
+   Symmetric & $0.818$ & $0.059$ & $X$ \\
    \hline
-   Symmetric & ${\color{red}X}$ & ${\color{red}X}$ & ${\color{red}X}$ \\
+   EDMD & $0.566$ & $0.038$ & $X$ \\
    \hline
-   EDMD & $9.83\times 10^{-4}$ & $5.59\times 10^{-2}$ & $9.80\times 10^{-2}$ \\
-   \hline
-   GPR & $1.14\times 10^{-3}$ & $7.02\times 10^{-2}$ & $1.07\times 10^{-1}$ \\
+   GPR & $0.735$ & $0.045$ & $X$ \\
    \hline
@@
 
-On the bottom panel of Fig.~\ref{fig:ex.intro}, we display the SST resulting from the {\sf BoundEffRed} strategy, applied to the same portion of PPG than what is used to display Fig.~\ref{fig:ppg}.
+On the bottom-right panel of Fig.~\ref{fig:ex.intro}, we display the SST resulting from the {\sf BoundEffRed} strategy, applied to the portion of PPG displayed on Fig.~\ref{fig:ppg}.
 We clearly observe an improvement of the quality of the SST near boundaries.

### 2020-07-10 Scripts/EvalPerfBoundEffRed/ResultsBoundEffRed.m

@@ -11,21 +11,21 @@ fprintf(' _______________________________________________________________\n')
 fprintf('| Extension Method | Computing time (sec.)  | Forecasting Error |\n')
 fprintf('|------------------|------------------------|-------------------|\n')
 fprintf('|      LSE         |           %.3f        |     %.3e     |\n', CompTime.LSE , mean(forecastErr.LSE) )
-fprintf('|  Symmetrization  |           %.3f        | %.3e | %.3e |\n', CompTime.SYM , mean(forecastErr.SYM) )
+fprintf('|  Symmetrization  |           %.3f        |     %.3e     |\n', CompTime.SYM , mean(forecastErr.SYM) )
 fprintf('|      EDMD        |           %.3f        |     %.3e     |\n', CompTime.EDMD , mean(forecastErr.EDMD) )
 fprintf('|      GPR         |         %.3f        |     %.3e     |\n', CompTime.GPR , mean(forecastErr.GPR) )
 fprintf('|__________________|________________________|___________________|\n\n')
 
-fprintf('==============Optimal Transport Distance===============\n')
-fprintf(' ______________________________________________________\n')
-fprintf('| Extension Method |    SST    |    STFT   |    RS     |\n')
-fprintf('|------------------|-----------|-----------|-----------|\n')
-fprintf('|  No extension    | %.3e | %.3e | %.3e |\n', mean(OTD.sst.S),mean(OTD.stft.S),mean(OTD.rs.S))
-fprintf('|  LSE extension   | %.3e | %.3e | %.3e |\n', mean(OTD.sst.LSE),mean(OTD.stft.LSE),mean(OTD.rs.LSE))
-fprintf('|  Symmetrization  | %.3e | %.3e | %.3e |\n', mean(OTD.sst.SYM),mean(OTD.stft.SYM),mean(OTD.rs.SYM))
-fprintf('|  EDMD extension  | %.3e | %.3e | %.3e |\n', mean(OTD.sst.EDMD),mean(OTD.stft.EDMD),mean(OTD.rs.EDMD))
[…]

### 2020-07-10 TimeFrequencyScaleRep/ConceFT_sqSTFT_C.m

@@ -15,10 +15,10 @@ N = length(x) ;
 
 
 if ~Second
-	fprintf(['Run ordinary STFT-SST x (Smoothing = ',num2str(Smooth),')\n']) ;
+% 	fprintf(['Run ordinary STFT-SST x (Smoothing = ',num2str(Smooth),')\n']) ;
     [tfr, tfrtic, tfrsq, tfrsqtic] = sqSTFTbase(x, lowFreq, highFreq, alpha, hop, h(1,:)', Dh(1,:)', Smooth);
 else
-	fprintf(['Run 2nd-order STFT-SST (no Smoothing)\n']) ;
+% 	fprintf(['Run 2nd-order STFT-SST (no Smoothing)\n']) ;
     [tfr, tfrtic, ~, tfrsq, tfrsqtic] = sqSTFTbase2nd(x, lowFreq, highFreq, alpha, hop, h(1,:)', Dh(1,:)', dwindow(Dh(1,:)'));
 end
 
@@ -27,9 +27,9 @@ end
 
 ConceFT = abs(tfrsq) ;
 if strcmp(method, 'MEAN')
-	fprintf('\tUse mean\n') ;
+% 	fprintf('\tUse mean\n') ;
 else
-	fprintf('\tUse quantile\n') ;
+% 	fprintf('\tUse quantile\n') ;
 	ConceFTcum = zeros([size(tfrsq) MT+1]) ;
 	ConceFTcum(:,:,1) = tfrsq ;
 end

### 2020-07-10 Notes/algorithm.tex
code

@@
 \begin{algorithmic}
-\STATE {\bf Inputs}: $\bx$, $M$, $K$, $N_0$, $L$, $\ccF$
+\STATE {\bf Inputs}: $\bx_{N}$, $M$, $K$, $N_0$, $L$, $\ccF$
+%\STATE {\bf Initialization}: $\bF_\bx^{(N_0)}=\ccF(\bx_{N_0})$
 \STATE
 \WHILE {$N$ increases}
+\STATE {\bf Real-time input}: $\bx_N$
+\STATE
 \STATE {\bf Forecasting step.}
-\STATE \quad\textbullet\ Signal extension: $\tilde\bx = \mbox{\sf SigExt}(\bx_N)$.
-\STATE 
-\STATE {\bf Representation step.}
-\STATE \quad\textbullet\ Representation evaluation: $\ccF_{N+2L}(\tilde\bx)$.
-\STATE \quad\textbullet\ Restriction of $\ccF_{N+2L}(\tilde\bx)$ to the central time interval (see~\eqref{eq:restriction}) to obtain $\bF_\bx^{(N)}=\ccF_N^{\mathrm{ext}}(\bx_N)$.
+\STATE \quad\textbullet\ Signal extension: $\tilde\bx_{N+L} = \mbox{\sf SigExt}(\bx_N)$. 
+\STATE {\bf Representation estimation step.}
+\STATE \quad\textbullet\ Extended representation evaluation: $\ccF(\tilde\bx_{N+L})$.
+\STATE \quad\textbullet\ Restriction of $\ccF(\tilde\bx_{N+L})$ to the current time interval (see~\eqref{eq:restriction}) to obtain $\bF_\bx^{(N)}=\ccF^{\mathrm{ext}}(\bx_N)$.
 \STATE
 \STATE {\bf Real-time output}: Signal representation $\bF_\bx^{(N)}$
-\STATE $N \leftarrow N+1$
+%\STATE $N \leftarrow N+1$
 \ENDWHILE
@@
 
-\paragraph{Signal extension} In order to reduce the boundary effects on both ``sides'' of the TF (or time-scale) representation, we finally construct the extended signal $\tilde\bx\in\RR^{N+2L}$ concatenating the backward prediction $\tilde\bx_{\rm bw}$, the observed signal $\bx$, and the forward prediction $\tilde\bx_{\rm fw}$.
+\paragraph{Signal extension} Since we are building a real-time algorithm, we consider that only the right ``side'' of the TF representation  is affected by boundary effects, the left "side" being fixed since it only concerns the past values of the signal.
+We therefore construct the extended signal $\tilde\bx\in\RR^{N+L}$ concatenating the observed signal $\bx$, and the forward forecasting $\tilde\bx_{\rm fw}$.
 We summarize the extension step in Algorithm~\ref{alg:extension}.
-Notice that we handle the backward estimation using the same strategy than described above, but applying it to the reverse signal $\bx^{\rm r} = \begin{pmatrix} \bx[N-1] & \cdots & \bx[0] \end{pmatrix}^T$.
+Notice that we we could, if necessary, handle the backward estimation using the same strategy than described above, but applying it to the reverse signal $\bx^{\rm r} = \begin{pmatrix} \bx[N-1] & \cdots & \bx[0] \end{pmatrix}^T$.
 
@@
 \begin{algorithmic}
-\STATE {\bf Inputs}: $\bx$, $M$, $K$, $L$
+\STATE {\bf Inputs}: $\bx_N$, $M$, $K$, $L$
 \STATE 
@@
 \STATE \quad\textbullet\ Forward forecasting $\tilde\bx_{\rm bw}$ obtained applying equation~\eqref{eq:prediction} with $\ell\in\{1,\ldots,L\}$.
+%\STATE 
+%\STATE {\bf Backward forecasting.}
+%\STATE \quad\textbullet\ Reverse signal $\bx$ to $\bx^{\rm r}$. 
+%\STATE \quad\textbullet\ LS estimation of the backward matrix $\tilde\bA_{\rm bw}$ via equation~\eqref{eq:lse} applied to $\bx^{\rm r}$.
+%\STATE \quad\textbullet\ Reversed backward forecasting $\tilde\bx_{\rm bw}^{\rm r}$ obtained applying equation~\eqref{eq:prediction} to $\bx^r$ with $\ell\in\{1,\ldots,L\}$.
+%\STATE \quad\textbullet\ Reverse $\tilde\bx_{\rm bw}^{\rm r}$ to obtain the estimate $\tilde\bx_{\rm bw}$.
 \STATE 
-\STATE {\bf Backward forecasting.}
-\STATE \quad\textbullet\ Reverse signal $\bx$ to $\bx^{\rm r}$. 
-\STATE \quad\textbullet\ LS estimation of the backward matrix $\tilde\bA_{\rm bw}$ via equation~\eqref{eq:lse} applied to $\bx^{\rm r}$.
-\STATE \quad\textbullet\ Reversed backward forecasting $\tilde\bx_{\rm bw}^{\rm r}$ obtained applying equation~\eqref{eq:prediction} to $\bx^r$ with $\ell\in\{1,\ldots,L\}$.
-\STATE \quad\textbullet\ Reverse $\tilde\bx_{\rm bw}^{\rm r}$ to obtain the estimate $\tilde\bx_{\rm bw}$.
-\STATE 
-\STATE {\bf Output}: Extended signal $\tilde\bx = \begin{pmatrix} \tilde\bx_{\rm bw}  & \bx & \tilde\bx_{\rm fw}\end{pmatrix}^T$.
+\STATE {\bf Output}: Extended signal $\tilde\bx_{N+L} = \begin{pmatrix} \bx_N & \tilde\bx_{\rm fw}\end{pmatrix}^T$. %\tilde\bx_{\rm bw}  &
 \end{algorithmic}
@@
 
-Let $\ccF_{N}:\RR^{N}\to\RR^{F\times N}$ generically denotes the TF or time-scale representation we are interested in, which could be, for instance, STFT, CWT, SST, or RS.
+Let $\ccF:\RR^{N}\to\CC^{F\times N}$ generically denotes the TF representation of interest to us, which could be, for instance, STFT, CWT, SST, or RS.
 Here, $F$ typically denotes the size of the discretization along the frequency axis.
-Due to the boundary effects, the representation $\ccF_{N}(\bx)$ shows undesired patterns when approaching its edges.
+Due to the boundary effects, the representation $\ccF(\bx_N)$ shows undesired patterns when approaching its edges.
 %For example, the instantaneous frequencies highlighted by the SST can be blurred near that edges.
@@
 This strategy moves the boundary effects out of the time interval $I=[0, \frac{N-1}{\fs}]$.
-Finally, the boundary-effects insensitive representation $\ccF_N^{\mathrm{ext}}:\RR^{N}\to\RR^{F\times N}$ of $\bx$ is given for all $\nu\in\{0,\cdots,F-1\}$, $n\in\{0,\cdots,N-1\}$ by:
+Finally, the boundary-effects insensitive representation $\ccF^{\mathrm{ext}}:\RR^{N}\to\CC^{F\times N}$ of $\bx_N$ is given for all $\nu\in\{0,\cdots,F-1\}$, $n\in\{0,\cdots,N-1\}$ by:
 \begin{equation}
-\ccF_N^{\mathrm{ext}}(\bx)[\nu,n] = \ccF_{N+2L}(\tilde \bx)[\nu,L+n]\ .
+\ccF^{\mathrm{ext}}(\bx_N)[\nu,n] = \ccF(\tilde \bx_{N+L})[\nu,n]\ .
 \label{eq:restriction}
 \end{equation}
-This amounts to restricting the representation $\ccF_{N+2L}(\tilde \bx)$ to the original measurement interval of $\bx$.
-For the sake of simplicity, we denotes the restriction operator by $\cR$, where $\cR:\RR^{F\times (N+2L)}\to\RR^{F\times N}$.
+This amounts to restricting the representation $\ccF(\tilde \bx_{N+L})$ to the current measurement interval of $\bx_N$.
+For the sake of simplicity, we denotes the restriction operator by $\cR$, where $\cR:\CC^{F\times (N+L)}\to\CC^{F\times N}$.
 Consequently, we have:
 \begin{equation}
-\ccF_N^{\mathrm{ext}}(\bx) = \cR\left( \ccF_{N+2L}(\tilde \bx) \right) \ .
+\ccF^{\mathrm{ext}}(\bx_N) = \cR\left( \ccF(\tilde \bx_{N+L}) \right) \ .
 \label{eq:bound.free.TFR}
 \end{equation}
-We call $\ccF_N^{\mathrm{ext}}$ the {\em boundary-free TF representation}.
+We call $\ccF^{\mathrm{ext}}$ the {\em boundary-free TF representation}.
+
+$\bF_\bx^{(N)}\in\CC^{F\times N}$ is the estimation of the boundary-free TF representation at iteration $N$ in Algorithm~\ref{alg:boundary}.
+For numerical purpose, and to make the real-time implementation achievable, we do not perform a full re-estimation of $\ccF^{\mathrm{ext}}$.
+Instead, the additional knowledge provided by $\bx[N+1]$ only influences the values of the last $L_{\mathrm{win}}$ columns of $\ccF^{\mathrm{ext}}(\bx_{N+1})$, where $L_{\mathrm{win}}$ denotes the window half-length used by the TF representation.
+Thus, $\bF_\bx^{(N+1)}$ is obtained by the concatenation of the first $N-L_{\mathrm{win}}+ 1$ columns of $\bF_\bx^{(N)}$ with the last $L_{\mathrm{win}}$ columns of $\ccF^{\mathrm{ext}}(\bx_{N+1})$.

### 2020-07-10 Scripts/BoundEffRedExamples/RealTime.m

@@ -1,0 +1,65 @@
+clear all; close all; clc;
+addpath(genpath('../../TimeFrequencyScaleRep'));
+addpath('../../Algorithm/');
+addpath('../../Signals/');
+
+%% load THO signal
+data = load('PPG.mat');
+xtot = data.PPG;
+fs = 100 ; % sampling frequency
+Ntot = length(xtot) ;
+
+method.name = 'lse' ;
+HOP = 1 ;
+extSEC = 5 ; % the extension is of extSEC second
+L = round(extSEC*fs) ;
+extM = round(1.5*L) ; % dimension of embedding / signals length
+extK = round( 2.5*extM ) ;  % number of points to estimate A / size of datasets
+
+basicTF.hop = 10 ;
+basicTF.win = 2*L+1 ; % window length (in samples)
+fmin = 0/fs ;
+fmax = 4/fs ;
+df = 2e-5 ;
+
+Nt = extK+extM+1 ; % segments length
+Lo = round(1.5*L/basicTF.hop); % overlap in TF domain
+LL = round(L/basicTF.hop); % extension in TF domain
+%% Initialization
+n0 = 1 ;
+n1 = 4000 ;
+x = xtot(n0:n1) ;
+[~, ~, SSTtot, ~, tfrsqticEXT] = ConceFT_sqSTFT_C(x, fmin, fmax,...
+            df, basicTF.hop, basicTF.win, 1, 10, 1, 0, 0) ;
+
+n0 = n1-L;
+n1 = n0+Nt-1 ;
+        
+% figure;
+% imagesc(log1p(abs(SSTtot)/5e1)); colormap(1-gray);drawnow;
+
+%% Real-time update
+k = 1 ;
+while n1<Ntot
+    tic;
+    x = xtot(n0:n1) ;
+
+    %% Forecasting
+    xext = forecasting(x,fs,HOP,extK,extM,extSEC,method);
+    xxLSE = [x; xext]; % size Nt + L
+%     xxLSE = SigExtension(x,fs,HOP,extK,extM,extSEC,method); 
+    
+    %% SST
+    [~, ~, SSTxx, ~, tfrsqticEXT] = ConceFT_sqSTFT_C(xxLSE((end-L-round(1.5*Lo*basicTF.hop)):end), fmin, fmax,...
+            df, basicTF.hop, basicTF.win, 1, 10, 1, 0, 0) ;
+    
[…]

### 2020-07-10 Scripts/EvalPerfBoundEffRed/ResultsBoundEffRed.m

@@ -21,10 +21,10 @@ fprintf(' __________________________________________\n')
 fprintf('| Extension Method |  SST  |  STFT |ConceFT|\n')
 fprintf('|------------------|-------|-------|-------|\n')
 %fprintf('|  No extension    | %.3e | %.3e | %.3e |\n', mean(OTD.sst.S),mean(OTD.stft.S),mean(OTD.rs.S))
-fprintf('|  LSE extension   | %.3f | %.3f | %.3f |\n', mean(OTD.sst.LSE./OTD.sst.S),mean(OTD.stft.LSE./OTD.stft.S),mean(OTD.conceft.LSE./OTD.rs.S))
-fprintf('|  Symmetrization  | %.3f | %.3f | %.3f |\n', mean(OTD.sst.SYM./OTD.sst.S),mean(OTD.stft.SYM./OTD.stft.S),mean(OTD.conceft.SYM./OTD.rs.S))
-fprintf('|  EDMD extension  | %.3f | %.3f | %.3f |\n', mean(OTD.sst.EDMD./OTD.sst.S),mean(OTD.stft.EDMD./OTD.stft.S),mean(OTD.conceft.EDMD./OTD.rs.S))
-fprintf('|  GPR extension   | %.3f | %.3f | %.3f |\n', mean(OTD.sst.GPR./OTD.sst.S),mean(OTD.stft.GPR./OTD.stft.S),mean(OTD.conceft.GPR./OTD.rs.S))
+fprintf('|  LSE extension   | %.3f | %.3f | %.3f |\n', mean(OTD.sst.LSE./OTD.sst.S),mean(OTD.stft.LSE./OTD.stft.S),mean(OTD.conceft.LSE./OTD.conceft.S))
+fprintf('|  Symmetrization  | %.3f | %.3f | %.3f |\n', mean(OTD.sst.SYM./OTD.sst.S),mean(OTD.stft.SYM./OTD.stft.S),mean(OTD.conceft.SYM./OTD.conceft.S))
+fprintf('|  EDMD extension  | %.3f | %.3f | %.3f |\n', mean(OTD.sst.EDMD./OTD.sst.S),mean(OTD.stft.EDMD./OTD.stft.S),mean(OTD.conceft.EDMD./OTD.conceft.S))
[…]

### 2020-07-10 Notes/results.tex
code

@@
 %
-{\color{red} [HT: need a more convincing explanation.] } 
+%{\color{red} [HT: need a more convincing explanation.] } 
 %It should be noted that, contrary to what expression~\eqref{eq:cov.error.2} suggests, smaller values of $\sigma$ do not cause a decrease of the experimental variance, but an increase.
@@
 
+Finally, in order to verify the ability of {\sf BoundEffRed} to be implemented in real time, we evaluate the computation time required for each iteration of Algorithm~\ref{alg:boundary}, corresponding to the update of the boundary-free TF representation.
+This check is performed on the real-time SST of a 10-minute sample of the respiratory signal.
+As the SST is temporally sub-sampled by a ratio of 10, the update should not exceed $10/\fs=0.1$~sec.
+This is indeed the case, as no iteration lasted longer than xx sec on a 6-Core Xeon CPU running at 3.5~GHz and 64~GB of RAM.
+
 \subsubsection{Photoplethysmogram}

### 2020-07-10 Scripts/BoundEffRedExamples/RealTime.m

@@ -27,20 +27,22 @@ Lo = round(1.5*L/basicTF.hop); % overlap in TF domain
 LL = round(L/basicTF.hop); % extension in TF domain
 %% Initialization
 n0 = 1 ;
-n1 = 4000 ;
+n1 = n0+Nt-1 ;
 x = xtot(n0:n1) ;
 [~, ~, SSTtot, ~, tfrsqticEXT] = ConceFT_sqSTFT_C(x, fmin, fmax,...
             df, basicTF.hop, basicTF.win, 1, 10, 1, 0, 0) ;
 
-n0 = n1-L;
-n1 = n0+Nt-1 ;
+%n0 = n1-L;
+%n1 = n0+Nt-1 ;
         
 % figure;
 % imagesc(log1p(abs(SSTtot)/5e1)); colormap(1-gray);drawnow;
 
 %% Real-time update
+Nmax = 10*60*fs ;
+
 k = 1 ;
-while n1<Ntot
+while n1<Nmax
     tic;
     x = xtot(n0:n1) ;
 
@@ -63,3 +65,6 @@ while n1<Ntot
     dt(k) = toc;
     k = k+1 ;
 end
+
+dtmax = basicTF.hop/fs ;
+save('RealTime','dt','dtmax')

### 2020-07-10 Notes/BoundEffRed.tex
code

@@
 \label{se:conclusion}
-In this paper, we propose an algorithm, named {\sf BoundEffRed}, for the boundary effects reduction on TF representations to achieve a real-time signal analysis.
+In this paper, we propose an algorithm, named {\sf BoundEffRed}, for the reduction of boundary effects on TF representations in real-time.
 This method is based on an extension of the signal obtained by a simple-minded and numerically efficient forecasting.
@@
 
-Various improvements can be considered to make the algorithm more robust to certain situations.
+Various improvements can be considered to make the algorithm more robust.
 In particular, we have noticed (see Fig.~\ref{fig:THO.failure}) that when the regular oscillations of the observed signal break, the forecasting is no longer relevant, and the TF representations are not ideal.
-A preliminary step should then be added to the algorithm to select the portions without sufficient quality, in order to reject the portions for which a better forecasting scheme is needed.
+A preliminary step should then be added to the algorithm to detect the portions to be rejected, for which a better forecasting scheme is needed.
 A more fundamental improvement would be to perform the prediction step directly in the TF plane, or even with the aim of specifically predicting the evolution of each instantaneous frequency.

### 2020-07-10 Notes/results.tex
code

@@
 \paragraph{Summary}
-Both previous experimental results combined with the theoretical asymptotic equation~\eqref{eq:cov.error.2} allow us to describe the influence of the noise variance and the size of the training dataset on the variance of the forecasting noise, which is empirically summarized as follows:
+Both previous experimental results combined with the theoretical asymptotic equation~\eqref{eq:cov.error.2} allow us to describe the influence of the noise variance and the size of the training dataset on the variance of the forecasting noise, empirically summarized as follows:
 \begin{equation}
@@
 \end{align*}
-Besides, the additive noise is chosen to be Gaussian: $\bw\sim\cN(\bzero,\bI)$. Numerically, we take: $N=10^4$, $M=750$, $p_1=10$, $p_2=23$.
+Besides, the noise is chosen to be Gaussian: $\bw\sim\cN(\bzero,\bI)$, and we take: $N=10^4$, $M=750$, $p_1=10$, $p_2=23$.
 
@@
 \begin{equation}
-\ccF_N^\mathrm{opt}(\bx) = \cR\left( \ccF_{N+2L}(\bx^\mathrm{ext}) \right) \ .
+\ccF^\mathrm{opt}(\bx_N) = \cR\left( \ccF(\bx^\mathrm{ext}) \right) \ .
 \label{eq:opt.TFR}
@@
 \begin{equation}
-D(\ccQ) = \dfrac1{|I|}\bigintss_I\ \frac{d_{t}\left(\ccQ,\ccF_N^\mathrm{opt}\right)}{d_{t}\left(\ccF_N,\ccF_N^\mathrm{opt}\right)}\ \dd t\ .
+D(\ccQ) = \dfrac1{|I|}\bigintss_I\ \frac{d_{t}\left(\ccQ,\ccF^\mathrm{opt}\right)}{d_{t}\left(\ccF,\ccF^\mathrm{opt}\right)}\ \dd t\ .
 \label{eq:index.perf}
@@
 \subsubsection{Respiratory signal}
-We first consider a 6 hours 20 minutes-long respiratory signal. This signal is sampled at $\fs=100$~Hz. A zoom on a small portion of the signal is displayed in Fig.~\ref{fig:tho}.
+We first consider a respiratory signal of 6 hours 20 minutes. This signal, a small portion of which is displayed in Fig.~\ref{fig:tho}, is sampled at $\fs=100$~Hz.
 

### 2020-07-10 Scripts/BoundEffRedExamples/RealTime.m

@@ -20,7 +20,7 @@ basicTF.hop = 10 ;
 basicTF.win = 2*L+1 ; % window length (in samples)
 fmin = 0/fs ;
 fmax = 4/fs ;
-df = 2e-5 ;
+df = 8e-5 ;
 
 Nt = extK+extM+1 ; % segments length
 Lo = round(1.5*L/basicTF.hop); % overlap in TF domain
@@ -39,7 +39,8 @@ x = xtot(n0:n1) ;
 % imagesc(log1p(abs(SSTtot)/5e1)); colormap(1-gray);drawnow;
 
 %% Real-time update
-Nmax = 10*60*fs ;
+minutes = 5; 
+Nmax = minutes*60*fs ;
 
 k = 1 ;
 while n1<Nmax

### 2020-07-11 Scripts/BoundEffRedExamples/RealTime.m
code

@@ -16,11 +16,11 @@ L = round(extSEC*fs) ;
 extM = round(1.5*L) ; % dimension of embedding / signals length
 extK = round( 2.5*extM ) ;  % number of points to estimate A / size of datasets
 
-basicTF.hop = 10 ;
+basicTF.hop = 20 ;
 basicTF.win = 2*L+1 ; % window length (in samples)
 fmin = 0/fs ;
 fmax = 4/fs ;
-df = 8e-5 ;
+df = 2e-4 ;
 
 Nt = extK+extM+1 ; % segments length
 Lo = round(1.5*L/basicTF.hop); % overlap in TF domain
@@ -29,8 +29,9 @@ LL = round(L/basicTF.hop); % extension in TF domain
 n0 = 1 ;
 n1 = n0+Nt-1 ;
 x = xtot(n0:n1) ;
-[~, ~, SSTtot, ~, tfrsqticEXT] = ConceFT_sqSTFT_C(x, fmin, fmax,...
+[~, ~, SSTcurrent, ~, tfrsqticEXT] = ConceFT_sqSTFT_C(x, fmin, fmax,...
             df, basicTF.hop, basicTF.win, 1, 10, 1, 0, 0) ;
+SSTtot = SSTcurrent ;
 
 %n0 = n1-L;
 %n1 = n0+Nt-1 ;
@@ -56,7 +57,7 @@ while n1<Nmax
     [~, ~, SSTxx, ~, tfrsqticEXT] = ConceFT_sqSTFT_C(xxLSE((end-L-round(1.5*Lo*basicTF.hop)):end), fmin, fmax,...
             df, basicTF.hop, basicTF.win, 1, 10, 1, 0, 0) ;
     
-    SSTtot = [SSTtot(:,1:(end-Lo)) SSTxx(:,(end-LL-Lo):(end-LL))] ;
+    SSTcurrent = [SSTcurrent(:,2:(end-Lo)) SSTxx(:,(end-LL-Lo):(end-LL))] ; % sliding sst
 
 %     imagesc(log1p(abs(SSTtot)/5e1)); colormap(1-gray);drawnow;
     
@@ -64,8 +65,9 @@ while n1<Nmax
     n1 = n0 + Nt - 1 ;
     
     dt(k) = toc;
+    SSTtot = [SSTtot(:,1:(end-Lo)) SSTxx(:,(end-LL-Lo):(end-LL))] ;
     k = k+1 ;
 end
 
[…]

### 2020-07-11 Notes/results.tex
code
notes

@@
 This check is performed on the real-time SST of a 10-minute sample of the respiratory signal.
-As the SST is temporally sub-sampled by a ratio of 10, the update should not exceed $10/\fs=0.1$~sec.
-This is indeed the case, as no iteration lasted longer than xx sec on a 6-Core Xeon CPU running at 3.5~GHz and 64~GB of RAM.
+As the SST is temporally sub-sampled by a ratio of 20, the update should not exceed $20/\fs=0.2$~sec.
+This is indeed the case, as no iteration lasted longer than $0.17$ sec on a 6-Core Xeon CPU running at 3.5~GHz and 64~GB of RAM.
 
@@
 %   \hline
-   {\sf SigExt} & $0.774$ & $0.048$ & $X$ \\
+   {\sf SigExt} & $0.774$ & $0.048$ & $0.826$ \\
    \hline
-   Symmetric & $0.818$ & $0.059$ & $X$ \\
+   Symmetric & $0.818$ & $0.059$ & $1.393$ \\
    \hline
-   EDMD & $0.566$ & $0.038$ & $X$ \\
+   EDMD & $0.566$ & $0.038$ & $0.850$ \\
    \hline
-   GPR & $0.735$ & $0.045$ & $X$ \\
+   GPR & $0.735$ & $0.045$ & $0.816$ \\
    \hline

### 2020-07-11 Notes/results.tex
typo

@@
 
-Finally, in order to verify the ability of {\sf BoundEffRed} to be implemented in real time, we evaluate the computation time required for each iteration of Algorithm~\ref{alg:boundary}, corresponding to the update of the boundary-free TF representation.
+Finally, in order to verify the ability of {\sf BoundEffRed} to be implemented in real time, we evaluate the computation time required for each iteration of Algorithm~\ref{alg:boundary}, updating of the boundary-free TF representation.
 This check is performed on the real-time SST of a 10-minute sample of the respiratory signal.

### 2020-07-14 Notes/results.tex
code

@@
 \centering
-\caption{AHM signal. Performance of the extension methods.}
+\caption{AHM signal: performance of the extension methods.}
 \begin{tabular}{|c||c|c|c|}
   \hline
-   \multirow{2}{*}{Algorithm} & \multicolumn{2}{c|}{MSE}  & \multirow{2}{41pt}{Computing time (sec.)} \\
+   \multirow{2}{40pt}{\centering Extension method} & \multicolumn{2}{c|}{MSE}  & \multirow{2}{41pt}{Computing time (sec.)} \\
    \cline{2-3} & Mean & Standard deviation & \\
@@
 where $P_{\ccQ}^t(\xi)=\int_{-\infty}^\xi p_{\ccQ}^t(\nu)\dd\nu$ and $\tilde P_{\ccF_0}^t(\xi)=\int_{-\infty}^\xi\tilde p_{\ccF_0}^t(\nu)\dd\nu$.
-The \textit{optimal transport distance} quantifies the proximity between the estimated and actual instantaneous frequencies while favoring the sparsity of the estimated TF representation.
-That is why, the performance index $D(\ccQ)$ for the reduction of boundary effects of a given TF representation $\ccQ$ is given by the averaged ratio between the optimal transport distance to the optimal TF representation~\eqref{eq:opt.TFR} and the optimal transport distance of to the optimal TF representation, that is:
+The \textit{Optimal Transport Distance (OTD)} quantifies the proximity between the estimated and actual instantaneous frequencies while favoring the sparsity of the estimated TF representation.
+That is why, the performance index $D(\ccQ)$ for the reduction of boundary effects of a given TF representation $\ccQ$ is given by the ratio between its averaged OTD to the optimal TF representation $\ccF^\mathrm{opt}$ (see~\eqref{eq:opt.TFR}) and the averaged OTD of the original TF representation $\ccF$ to $\ccF^\mathrm{opt}$, that is:
+%\begin{equation}
+%D(\ccQ) = \dfrac1{|I|}\bigintssss_I\ \frac{d_{t}\left(\ccQ,\ccF^\mathrm{opt}\right)}{d_{t}\left(\ccF,\ccF^\mathrm{opt}\right)}\ \dd t\ .
+%\label{eq:index.perf}
+%\end{equation}
 \begin{equation}
-D(\ccQ) = \dfrac1{|I|}\bigintss_I\ \frac{d_{t}\left(\ccQ,\ccF^\mathrm{opt}\right)}{d_{t}\left(\ccF,\ccF^\mathrm{opt}\right)}\ \dd t\ .
+D(\ccQ) = \dfrac{\displaystyle\int_I d_{t}\left(\ccQ,\ccF^\mathrm{opt}\right)\dd t}{\displaystyle\int_I d_{t}\left(\ccF,\ccF^\mathrm{opt}\right)\dd t}\ .
 \label{eq:index.perf}
@@
 \centering
-\caption{Performance of the extension methods on a respiratory signal.}
+\caption{Respiratory signal: Performance of extension methods and associated boundary-free TF representations.}
 \begin{tabular}{|c||c||c|c|c|c|}
   \hline
-   \multirow{2}{*}{Algorithm} & \multirow{2}{35pt}{\centering Averaged MSE} & \multicolumn{4}{c|}{Averaged performance index $D$} \\
+   \multirow{2}{40pt}{\centering Extension method} & \multirow{2}{35pt}{\centering Averaged MSE} & \multicolumn{4}{c|}{Averaged performance index $D$} \\
    \cline{3-6}
@@
 \centering
-\caption{PPG signal: averaged performance index of the boundary-free TF representations for different extensions.}
+\caption{PPG signal: Performance of the boundary-free TF representations according to the extension method.}
 \begin{tabular}{|c||c|c|c|}
   \hline
-   \multirow{2}{*}{Extension method} & \multicolumn{3}{c|}{Performance index $D$} \\
+   \multirow{2}{40pt}{\centering Extension method} & \multicolumn{3}{c|}{Averaged performance index $D$} \\
    \cline{2-4}

### 2020-07-15 Notes/BoundEffRed.tex
code

@@
 Time-frequency (TF) representations of time series are intrinsically subject to the boundary effects.
-As a consequence, the structures of signals that are highlighted by the representations are garbled when approaching the boundaries of the TF domain.
-In this paper, for the purpose of real-time TF information acquisition for nonstationary oscillatory time series, we propose a numerically efficient approach for the reduction of such boundary effects.
+As a result, the structures of signals that are highlighted by the representations are garbled when approaching the boundaries of the TF domain.
+In this paper, for the purpose of real-time TF information acquisition of nonstationary oscillatory time series, we propose a numerically efficient approach for the reduction of such boundary effects.
 The solution relies on an extension of the analyzed signal obtained by a forecasting technique.
 In the case of the study of a class of locally oscillating signals, we provide a theoretical guarantee of the performance of our approach.
-Following a numerical verification of the performance of the algorithm, we validated its performance on various real-life biomedical signals.
+Following a numerical verification of the algorithmic performance of our approach, we validate it by implementing it on biomedical signals.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Boundary effects, time-frequency, forecasting
+Boundary effects, time-frequency, forecasting, nonstationarity
 \end{IEEEkeywords}
@@
 In this paper, we focus on oscillatory time series.
-For instance, biomedical signal, such as photoplethysmogram (PPG), contain several characteristics, such as respiratory rate or blood pressure, that cannot be interpreted from its running chart in the time domain.
+For instance, biomedical signal, such as photoplethysmogram (PPG), contain several characteristics, such as respiratory rate or blood pressure, that cannot be interpreted from its run-sequence plot in the time domain.
 An analysis tool would make possible the extraction of these useful characteristics.

### 2020-07-15 Notes/results.tex
code

@@
 However, the TBATS extension method in not implemented here because of its excessive computing time.
-The extensions of $7$ seconds-long on each boundary, corresponding to $L =700$.
+We forecast $7$-second-long extensions on each segment of signal, corresponding to $L =700$.
 Thus, in order to catch slowly varying dynamical behaviors, the size of the training signal $M$ is chosen so that $M=\lfloor 1.5L\rfloor$.
@@
    \hhline{|=#=#=|=|=|=|}
-   {\sf SigExt} & $0.046$ & $0.705$ & $0.698$ & $0.694$ & $0.776$ \\
+   {\sf SigExt} & $0.046$ & $0.698$ &  $0.705$ & $0.694$ & $0.776$ \\
    \hline
-   Symmetric & $0.039$ & $1.280$ & $0.984$ & $9.063$ & $1.295$ \\
+   Symmetric & $0.039$ & $0.984$ & $1.280$ & $9.063$ & $1.295$ \\
    \hline
-   EDMD & $0.022$ & $0.757$ & $0.635$ & $0.701$ & $0.793$ \\
+   EDMD & $0.022$ & $0.635$ &  $0.757$ & $0.701$ & $0.793$ \\
    \hline
-   GPR & $0.045$ & $0.817$ & $0.793$ & $0.727$ & $0.876$ \\ 
+   GPR & $0.045$ & $0.793$ &  $0.817$ & $0.727$ & $0.876$ \\ 
    \hline
@@
 We perform a study similar to the previous one on a 640 second-long photoplethysmogram (PPG) signal extracted from the Physionet dataset~\cite{Pimentel17toward, Goldberger00physiobank}, sampled at $\fs=125$~Hz.
-A 32 second-long piece of this signal is displayed on the top of Fig.~\ref{fig:ppg}.
-The estimated 2-second extension obtained by {\sf SigExt} on both boundary of this signal is superimposed to the ground-truth signal in the bottom of Fig.~\ref{fig:ppg}.
+A portion of this signal is displayed in Fig.~\ref{fig:ppg}.
+The estimated 5 seconds extensions of this segment obtained by {\sf SigExt}, EDMD and GPR forecastings  are superimposed to the ground-truth extension in Fig.~\ref{fig:ppg}.
 
@@
 
-We divide the signal into 32-second long segments, and apply Algorithm~\ref{alg:boundary} on each piece.
+We divide the signal into 32-second-long segments, and apply Algorithm~\ref{alg:boundary} on each piece.
 We provide in Table~\ref{tab:otd.ppg} the performance index $D$ of the boundary-free TF representations averaged over the signals.
@@
 %   \hline
-   {\sf SigExt} & $0.774$ & $0.048$ & $0.826$ \\
+   {\sf SigExt} & $0.425$ & $0.713$ & $0.687$ \\
    \hline
-   Symmetric & $0.818$ & $0.059$ & $1.393$ \\
+   Symmetric & $1.335$ & $1.340$ & $1.258$ \\
    \hline
-   EDMD & $0.566$ & $0.038$ & $0.850$ \\
+   EDMD & $0.406$ & $0.642$ & $0.615$ \\
    \hline
-   GPR & $0.735$ & $0.045$ & $0.816$ \\
+   GPR & $0.448$ & $0.766$ & $0.730$ \\
    \hline

### 2020-07-15 Algorithm/SigExtension.m

@@ -1,6 +1,8 @@
 function xx = SigExtension(x,fs,HOP,extK,extM,extSEC,method)
 
-xext = forecasting(x,fs,HOP,extK,extM,extSEC,method);
-xexti = forecasting(x,fs,HOP,extK,extM,extSEC,method,'backward');
+L = round(fs*extSEC);
+
+xext = forecasting(x,L,HOP,extK,extM,method);
+xexti = forecasting(x,L,HOP,extK,extM,method,'backward');
 
 xx = [xexti; x; xext];

### 2020-07-15 Algorithm/forecasting.m

@@ -1,4 +1,4 @@
-function xext = forecasting(x,fs,HOP,extK,extM,extSEC,method,varargin)
+function xext = forecasting(x,L,HOP,extK,extM,method,varargin)
 
 if (~isempty(varargin))&&(isequal(varargin{1},'backward'))
     x = flipud(x);
@@ -8,8 +8,9 @@ if ~strcmpi(method.name,'symmetrization')
     X = [] ; Y = [] ;
     for kk = 1: extK
         X(:,kk) = x(end-extK-extM+kk: HOP: end-extK+kk-1) ;
-        Y(:,kk) = x(end-extK-extM+kk+1: HOP: end-extK+kk) ;
+%         Y(:,kk) = x(end-extK-extM+kk+1: HOP: end-extK+kk) ;
     end
+    Y = [X(:,2:end) x(end-extM+1: HOP: end)] ;
 end
 
 
@@ -43,7 +44,6 @@ switch method.name
 end
 
 %% Extension
-L = round(fs*extSEC);
 Z = zeros(extM,L) ;
 
 switch method.name

### 2020-07-15 Scripts/EvalPerfForecasting/InfluenceNoiseSine.m

@@ -4,17 +4,16 @@ addpath('../../Algorithm/');
 
 %% Parameters
 
-N = 10000 ; fs = N-1 ;
-t = linspace(0,1,N);
-
 % forecasting parameters
 HOP = 1 ;
-extSEC = 0.01 ; % the extension is of extSEC second
-L = round( extSEC*fs ) ;
+L = 350 ;
 extM = round( 1.5*L ) ; % dimension of embedding / signals length
 extK = round( 3*extM );  % number of points to estimate A / size of datasets
 
-tt = linspace(-L/fs, 1+L/fs, N+2*L) ;
+N = extK + extM + 1 ; 
+fs = N-1 ;
+t = linspace(0,1,N);
+tt = linspace(0, 1+L/fs, N+L) ;
 
 %% Synthesize signal
 
@@ -34,7 +33,7 @@ switch nComp
         xx01 = R*cos(2*pi*f1*tt) ;
 end
 xx0 = xx00 + xx01 ; % extended signal
-x0 = xx0( (L+1) : (L+N) ) ; % restriction to the measurement interval
+x0 = xx0(1:N) ; % restriction to the measurement interval
 
 %% Ideal Forecasting vector
 
@@ -47,28 +46,35 @@ zK = x0(end-extM+1 : HOP: end).' ;
 
 %% Forecasting
 method.name = 'lseV' ;
-nbXP = 3 ;
-nbXPP = 3000 ;
+nbXP = 3 ; % number of noise levels
+nbXPP = 3000 ; % number of XP / noise level
 Sigma = logspace(-3,-1,nbXP) ;
-Covh = zeros(extM,extM,L) ;
-Ehw2 = zeros(L,1);
+
+hopL = 10 ; % hop for forecasting
+lenForcast = length(1:hopL:L) ;
+Covh = zeros(extM,extM, lenForcast) ;
+Ehw2 = zeros(lenForcast,1);
 k = 1 ;
 for k = 1:nbXP
     sigman = Sigma(k) ;
+    
+    % Theoretical foreacstion Matrix
     S0 = (1/extK)*(X0*X0') + sigman^2*eye(extM) ;
     r = zeros(1,extM) ; r(2) = 1 ;
     S1 = (1/extK)*(Y0*X0') + sigman^2*toeplitz(zeros(extM,1),r) ;
     A0 = S1 / S0 ; % ideal least square estimation
-    for l = 1:L
-        tmp0 = A0^l ;
[…]

### 2020-07-16 Notes/BoundEffRed.tex
paper

@@
 \IEEEPARstart{I}{n} any digital acquisition system, the study and the interpretation of the measured signals generally require an analysis tool, which enables researchers to point out the useful characteristics of the signal.
-The need for signal analysis arises from various signals, ranging from audio~\cite{Stowell18computational,Muller11signal}, mechanical~\cite{Peng02vibration} or biomedical signals~\cite{Akay96detection}.
-In this paper, we focus on oscillatory time series.
-For instance, biomedical signal, such as photoplethysmogram (PPG), contain several characteristics, such as respiratory rate or blood pressure, that cannot be interpreted from its run-sequence plot in the time domain.
+The need for signal analysis arises from various signals, ranging from audio~\cite{Stowell18computational,Muller11signal}, mechanical~\cite{Peng02vibration}, or biomedical signals~\cite{Akay96detection}.
+For instance, biomedical signals, such as photoplethysmogram (PPG), contain several characteristics, including respiratory rate or blood pressure, that cannot be interpreted from its run-sequence plot in the time domain.
 An analysis tool would make possible the extraction of these useful characteristics.
@@
 Usually, the measured signals exhibit nonstationary behavior, and the observed quantities might be interfered by transient phenomena that can vary rapidly and irregularly.
-For example, the signal might oscillate fast with large amplitude at one moment, and then oscillate slow with small amplitude at the next moment.
+In this paper, we focus on oscillatory time series.
+These signals might, for example, oscillate fast with large amplitude at one moment, and then oscillate slowly with small amplitude at the next moment.
 In order to adapt the analysis to nonstationarities, local spectral analysis is generally performed~\cite{Stoica05spectral,Matz97generalized}.
-The short time Fourier transform~\cite{Grochenig01foundations} (STFT) is a typical tool which is build for this purpose, and enables the determination of the local frequency content of a nonstationary signal.
+The short-time Fourier transform~\cite{Grochenig01foundations} (STFT), a typical tool built for this purpose, enables the determination of the local frequency content of a nonstationary signal.
 
@@
 Let $g_s:\RR\to\RR$ denote the analysis window, where $s$ is a shape parameter.
-The support of $g_s$ is is localized around the origin, and is small with respect to $|I|$.
+The support of $g_s$ is localized around the origin and is small with respect to $|I|$.
 The translation operator is $T_\tau$ defined as:
@@
 Clearly, at these points, half of the information is missing.
-Consequently the results of the inner product~\eqref{eq:windowing} is distorted.
+Consequently, the results of the inner product~\eqref{eq:windowing} are distorted.
 This phenomenon is usually understood as the \emph{boundary effect}.
-We display on Fig.~\ref{fig:ex.intro} the result of the SST of a PPG (see section~\ref{ssse:ppg} for a comprehensive description).
-The distortion resulting from boundary effects is clearly visible on both sides of this representation.
-Indeed, while in the major central part of the image, clear lines stand out, they become blurred as they approach the left and right boundaries of the image.
-The right boundary is zoomed in and displayed on the right of Fig.~\ref{fig:ex.intro}, which emphasizes the result of boundary effects.
-Clearly, estimations of signal characteristics, like instantaneous frequencies~\cite{Delprat92asymptotic} or amplitudes, from this TF representation appear to be imprecisely determinable (or even likely to fail) in the vicinity of the boundaries.
-Moreover, this boundary effect would unavoidably limit applying these window based TF analysis tools for the real-time analysis purpose.
-It is thus desirable to have a solution to eliminate this boundary effect.
+We display on the bottom left of Fig.~\ref{fig:ex.intro} the result of the SST of a PPG (see section~\ref{ssse:ppg} for a comprehensive description).
+The distortion resulting from boundary effects is clearly visible on the right side of this representation.
+Indeed, while in the major left part of the image, clear lines stand out, they become blurred as they approach the right boundary of the image.
+Therefore, estimations of signal characteristics, like instantaneous frequencies~\cite{Delprat92asymptotic} or amplitudes, from this TF representation appear to be imprecisely determinable (or even likely to fail) in the vicinity of the boundaries.
+Moreover, this boundary effect would unavoidably limit applying these window-based TF analysis tools for real-time analysis purposes.
+It is thus desirable to have a solution to eliminating the boundary effects.
 
@@
 
-%The purpose of this paper is providing a fast algorithm allowing to tackle boundary effects and limit the disruptions it may cause on the signals representations based on windowing. 
-Attempts to minimize the boundary effects generally consists in softening the discontinuity on signals edges.
+Attempts to minimize the boundary effects generally consists in softening the discontinuity on signal edges.
 For instance, judicious choices of analysis windows whose support does not interfere with the boundary points can minimize the occurrence of aberrant patterns near the boundaries of the TF plane~\cite{Chui92wavelets,Depczynski99fast}.
 Due to the specific relationship of chosen analysis windows and TF analysis tool, these techniques do not make it possible to reduce the boundary effects of any TF representations, and for any analysis window.
-Another natural idea consists of carrying out a preliminary step of extending the signal beyond its boundaries,
-%
-and due to its flexibility, various extension schemes have been proposed.
-For example, there exists simple extension schemes that do not take into account the dynamical behavior of the signal, such as zero-padding, periodic extension, symmetric extension~\cite{Kharitonenko02wavelet,Chen95symmetric}, or polynomial extrapolation~\cite{Williams97discrete}.
+Another natural idea consists of carrying out a preliminary step of extending the signal beyond its boundaries, and due to its flexibility, various extension schemes have been proposed.
+For example, there exist simple extension schemes that do not take into account the dynamical behavior of the signal, such as zero-padding, periodic extension, symmetric extension~\cite{Kharitonenko02wavelet,Chen95symmetric}, or polynomial extrapolation~\cite{Williams97discrete}.
 There also exist extension schemes based on physically relevant dynamical models, such as the Extended Dynamic Mode Decomposition~\cite{Williams15data} (EDMD), the Gaussian process regression~\cite{Rasmussen06gaussian,Roberts13Gaussian} (GPR), the Trigonometric, Box-Cox transformation, ARMA errors, Trend and Seasonal components (TBATS) algorithm~\cite{DeLivera11forecasting}, the dynamic linear models \cite{west2006bayesian}, and long short-term memory model \cite{vlachas2018data}.
-It is a far from exhaustive list, and we refer readers with interest to \cite{hyndman2018forecasting} for a friendly monograph on the general forecasting topic.
-While the second class of extension schemes give better extended signals than the first class, they generally have a great computational cost.
-%
+It is a far from an exhaustive list, and we refer readers with interest to \cite{hyndman2018forecasting} for a friendly monograph on the general forecasting topic.
+While the second class of extension schemes gives better-extended signals than the first class, they generally have a great computational cost.
 
 In this paper, we propose a fast extension algorithm based on a simple dynamical model, in order to optimize the trade-off between the extension quality and the computational cost.
-The proposed algorithm is composed of two steps.
+The proposed algorithm is then composed of two steps.
 \begin{enumerate}
@@
 Then, once this operation is done, we have access to an extended signal defined on a larger interval $I_\Delta$, where $\Delta$ denotes the size of the extension on both boundaries of $I$.
-\item \emph{Run the local analysis tool on the extended signal.} Assuming that the support of the analysis window is smaller that $2\Delta$, the local analysis near the boundary of $I$ is now possible without lack of information thanks to knowledge brought by the extension on both sides.
+\item \emph{Run the local analysis tool on the extended signal.} Assuming that the support of the analysis window is smaller than $2\Delta$, the local analysis near the boundary of $I$ is now possible without lack of information thanks to knowledge brought by the extension.
 \end{enumerate}
 Thus, assuming that the quality of the extension step is sufficient, the analysis results obtained that way will be less sensitive to the boundary effects than the result of the analysis tool applied directly to the non-extended signal.
-%
-We claim and prove that forecasting oscillatory signals based on the simple dynamical model combined with the simple least square approach is sufficient for reducing the boundary effects for the TF analysis, or other kernel based analysis.
-See Figure \ref{fig:ex.intro} for a snapshot of the result.
-The main benefit of this simple approach is a numerically efficient solution with theoretical guarantee for the real-time analysis purpose.
+We claim and prove that forecasting oscillatory signals based on the simple dynamical model combined with the simple least square approach is sufficient for reducing the boundary effects for the TF analysis, or other kernel-based analysis.
+See the bottom right of Fig.~\ref{fig:ex.intro} for a snapshot of the result.
+The main benefit of this simple approach is a numerically efficient solution with a theoretical guarantee for real-time analysis purposes.
 
@@
 We derive the corresponding algorithm for boundary effects reduction.
-In section~\ref{se:theoretical}, we show that, even though we consider a simplistic dynamic model, it is sufficient to extend signals takings the form of sums of sine waves.
+In section~\ref{se:theoretical}, we show that the dynamic model we consider is sufficient to extend signals takings the form of sums of sine waves.
 An evaluation of the theoretical performance of our algorithm on a class of signals, the sums of sine waves, is given in section~\ref{se:theoretical}.
@@
 We show that our algorithm gives fast results of reasonable quality.
-Finally, we evaluate the performance of our boundary effects reduction algorithm on biomedical signals such as respiratory signals, and compare it to the theoretical results.
+Finally, we evaluate the performance of our boundary effects reduction algorithm on biomedical signals, such as respiratory signals, and compare it to the theoretical results.
 
@@
 \label{se:conclusion}
-In this paper, we propose an algorithm, named {\sf BoundEffRed}, for the reduction of boundary effects on TF representations in real-time.
+In this paper, we propose an algorithm, named {\sf BoundEffRed}, for the real-time reduction of boundary effects in TF representations.
 This method is based on an extension of the signal obtained by a simple-minded and numerically efficient forecasting.
-We have shown theoretically that the chosen dynamic model is sufficient for the extension of signals formed by a sum of sine waves.
-Moreover, the numerical results show that this strategy allows us to switch to a real-time implementation of {\sf BoundEffRed} is possible, due to the low computation time, unlike other existing forecasting methods.
+We have shown theoretically that the chosen dynamic model is sufficient to extend signals formed by a sum of sine waves.
+Moreover, thanks to the low computational time, the numerical results show that this strategy allows us to switch to a real-time implementation of {\sf BoundEffRed}, unlike other existing forecasting methods.
 The numerical results also confirmed the robustness to noise of {\sf BoundEffRed}, as well as its ability to be applied to many time-frequency representations.
@@
 Various improvements can be considered to make the algorithm more robust.
-In particular, we have noticed (see Fig.~\ref{fig:THO.failure}) that when the regular oscillations of the observed signal break, the forecasting is no longer relevant, and the TF representations are not ideal.
-A preliminary step should then be added to the algorithm to detect the portions to be rejected, for which a better forecasting scheme is needed.
-A more fundamental improvement would be to perform the prediction step directly in the TF plane, or even with the aim of specifically predicting the evolution of each instantaneous frequency.
+In particular, we have noticed (see Fig.~\ref{fig:THO.failure}) that when the regular oscillations of the observed signal break, the forecasting step is no longer relevant, and only slows down the calculation of the TF representation.
+A preliminary step should then be added to the algorithm to detect signal activity, and disable the forecasting step when possible.
+A more fundamental improvement would be to perform the forecasting step directly in the TF plane, to specifically predict the evolution of each instantaneous frequency.
 This would lead to seeing a TF representation as a multivariate time series (corresponding to each frequency band).

### 2020-07-16 Notes/algorithm.tex
paper

@@
 As explained above, the algorithm for the reduction of boundary effects on TF representations relies on extending the signal by forecasting it before applying the TF analysis.
-%
 
-%\paragraph{Notations} 
 We start with the notation.
@@
 %
-The pseudo-code of the proposed algorithm to reduce boundary effects on windowing-based TF representations is shown in the Algorithm~\ref{alg:boundary}.
+The pseudo-code of the proposed real-time algorithm to reduce boundary effects on windowing-based TF representations is shown in Algorithm~\ref{alg:boundary}.
 We coined the algorithm {\sf BoundEffRed}.
-Below, we detail the algorithm, particularly the Signal extension algorithm {\sf SigExt} in Algorithm~\ref{alg:extension}.
+Below, we detail the algorithm, particularly the signal extension {\sf SigExt} in Algorithm~\ref{alg:extension}.
 
@@
 %\STATE {\bf Initialization}: $\bF_\bx^{(N_0)}=\ccF(\bx_{N_0})$
-\STATE
+\STATE \vspace{-2mm}
 \WHILE {$N$ increases}
 \STATE {\bf Real-time input}: $\bx_N$
-\STATE
+\STATE \vspace{-2mm}
 \STATE {\bf Forecasting step.}
@@
 \STATE \quad\textbullet\ Restriction of $\ccF(\tilde\bx_{N+L})$ to the current time interval (see~\eqref{eq:restriction}) to obtain $\bF_\bx^{(N)}=\ccF^{\mathrm{ext}}(\bx_N)$.
-\STATE
+\STATE \vspace{-2mm}
 \STATE {\bf Real-time output}: Signal representation $\bF_\bx^{(N)}$
@@
 For instance, the dynamic mode decomposition~\cite{Schmid10dynamic,Williams15data} or other more complicated models \cite{Roberts13Gaussian,DeLivera11forecasting,west2006bayesian,vlachas2018data}, allow this by setting additional constraints on the behavior of $f$.
-We will see later, in section~\ref{se:theoretical}, that it is not necessary to consider such a complex dynamic model for the study of the oscillatory signals of interest to us.
+We will see, in section~\ref{se:theoretical}, that considering such a complex dynamic model is not necessary for the study of the oscillatory signals of interest to us.
 That is why we consider here a naive dynamical model, assuming that we have the following relation:
@@
 where $\bA\in\RR^{M\times M}$.
-In other words, we adopt a classical strategy in the study of dynamical systems, the {\em linearization} of a nonlinear phenomenon, when the system is sufficient regular.
+In other words, we adopt a classical strategy in the study of dynamical systems, the {\em linearization} of a nonlinear phenomenon, when the system is sufficiently regular.
 Notice that this linearized dynamical model can be written equivalently in function of the sub-signals $\bx_k$, as:
@@
 \]
-Then, except the row vector $\balpha = \left(\alpha_1 \cdots\alpha_M\right)$, the matrix $\bA$ is fully determined by the dynamical model.
+Then, except for the row vector $\balpha = \left(\alpha_1 \cdots\alpha_M\right)$, the matrix $\bA$ is fully determined by the dynamical model.
 \end{remark}
 
-\paragraph{Signal extension} Since we are building a real-time algorithm, we consider that only the right ``side'' of the TF representation  is affected by boundary effects, the left "side" being fixed since it only concerns the past values of the signal.
-We therefore construct the extended signal $\tilde\bx\in\RR^{N+L}$ concatenating the observed signal $\bx$, and the forward forecasting $\tilde\bx_{\rm fw}$.
+\paragraph{Signal extension} Since we are building a real-time algorithm, we consider that only the right ``side'' of the signal has to be extended, the left ``side'' being fixed since it only concerns the past values of the signal.
+We therefore construct the extended signal $\tilde\bx\in\RR^{N+L}$ concatenating the observed signal $\bx$, and the forward forecasting $\tilde\bx_{\mathrm{fw}}\in\RR^L$.
 We summarize the extension step in Algorithm~\ref{alg:extension}.
-Notice that we we could, if necessary, handle the backward estimation using the same strategy than described above, but applying it to the reverse signal $\bx^{\rm r} = \begin{pmatrix} \bx[N-1] & \cdots & \bx[0] \end{pmatrix}^T$.
+Notice that we could, if necessary, handle the backward estimation using the same strategy as described above, but applying it to the reverse signal $\bx^{\rm r} = \begin{pmatrix} \bx[N-1] & \cdots & \bx[0] \end{pmatrix}^T$.
 
@@
 \STATE {\bf Inputs}: $\bx_N$, $M$, $K$, $L$
-\STATE 
+\STATE \vspace{-2mm}
 \STATE {\bf Forward forecasting.}
-\STATE \quad\textbullet\ LS estimation of the forward matrix $\tilde\bA_{\rm fw}$ via equation~\eqref{eq:lse}.
-\STATE \quad\textbullet\ Forward forecasting $\tilde\bx_{\rm bw}$ obtained applying equation~\eqref{eq:prediction} with $\ell\in\{1,\ldots,L\}$.
+\STATE \quad\textbullet\ Estimation of the matrix $\tilde\bA$ via equation~\eqref{eq:lse}.
+\STATE \quad\textbullet\ Forecasting $\tilde\bx_{\mathrm{fw}}\in\RR^L$ obtained applying equation~\eqref{eq:prediction} with $\ell\in\{1,\ldots,L\}$.
 %\STATE 
@@
 %\STATE \quad\textbullet\ Reverse $\tilde\bx_{\rm bw}^{\rm r}$ to obtain the estimate $\tilde\bx_{\rm bw}$.
-\STATE 
-\STATE {\bf Output}: Extended signal $\tilde\bx_{N+L} = \begin{pmatrix} \bx_N & \tilde\bx_{\rm fw}\end{pmatrix}^T$. %\tilde\bx_{\rm bw}  &
+\STATE \vspace{-2mm}
+\STATE {\bf Output}: Extended signal $\tilde\bx_{N+L} = \begin{pmatrix} \bx_N & \tilde\bx_{\mathrm{fw}}\end{pmatrix}^T$. %\tilde\bx_{\rm bw}  &
 \end{algorithmic}
@@
 Due to the boundary effects, the representation $\ccF(\bx_N)$ shows undesired patterns when approaching its edges.
-%For example, the instantaneous frequencies highlighted by the SST can be blurred near that edges.
-To alleviate the boundary effect, we apply the representation to the estimated extended signal $\tilde\bx$.
+To alleviate the boundary effects, we apply the representation to the estimated extended signal $\tilde\bx$.
 This strategy moves the boundary effects out of the time interval $I=[0, \frac{N-1}{\fs}]$.
@@
 This amounts to restricting the representation $\ccF(\tilde \bx_{N+L})$ to the current measurement interval of $\bx_N$.
-For the sake of simplicity, we denotes the restriction operator by $\cR$, where $\cR:\CC^{F\times (N+L)}\to\CC^{F\times N}$.
+For the sake of simplicity, we denote the restriction operator by $\cR$, where $\cR:\CC^{F\times (N+L)}\to\CC^{F\times N}$.
 Consequently, we have:
@@
 $\bF_\bx^{(N)}\in\CC^{F\times N}$ is the estimation of the boundary-free TF representation at iteration $N$ in Algorithm~\ref{alg:boundary}.
-For numerical purpose, and to make the real-time implementation achievable, we do not perform a full re-estimation of $\ccF^{\mathrm{ext}}$.
+For numerical purposes, and to make the real-time implementation achievable, we do not perform a full re-estimation of $\ccF^{\mathrm{ext}}$ at iteration $N+1$.
 Instead, the additional knowledge provided by $\bx[N+1]$ only influences the values of the last $L_{\mathrm{win}}$ columns of $\ccF^{\mathrm{ext}}(\bx_{N+1})$, where $L_{\mathrm{win}}$ denotes the window half-length used by the TF representation.
 Thus, $\bF_\bx^{(N+1)}$ is obtained by the concatenation of the first $N-L_{\mathrm{win}}+ 1$ columns of $\bF_\bx^{(N)}$ with the last $L_{\mathrm{win}}$ columns of $\ccF^{\mathrm{ext}}(\bx_{N+1})$.
+
+\begin{remark}
+Unlike other representations, RS is not causal and rigorously requires knowing the STFT on the whole time-frequency plane before making any reassignment.
+Nevertheless, in~\cite{Lin17conceft}, the authors propose a causal RS that only requires the future values of the STFT within a limited time horizon (typically less than $L$).
+We use this version in the paper to enable real-time implementation of RS.
+\end{remark}

### 2020-07-16 Notes/results.tex
paper

@@
 
-\subsection{Evaluation the forecasting performance}
-In that section, we first evaluate the quality of the forecasting step and compare it the the theoretical results provided by Theorem~\ref{th:error}.
+\subsection{Evaluation of the forecasting performance}
+In that section, we first evaluate the quality of the forecasting step and compare it to the theoretical results provided by Theorem~\ref{th:error}.
 The level of the forecasting error depends on at least two parameters:
@@
 We proved that the linear dynamic model is sufficient to catch the dynamical behavior of signals taking the form~\eqref{eq:sum.sine}.
-In order to validate this theoretical result, we apply the forecasting Algorithm~\ref{alg:extension} to a large number of realizations of the random vector $\bx$ following the model~\eqref{eq:model.noise}, and such that the deterministic component $\bz$ takes the form:
+In order to validate this theoretical result, we apply the forecasting Algorithm~\ref{alg:extension} to a large number of realizations of the random vector $\bx$ of size $N=10^4$, following model~\eqref{eq:model.noise}, and such that the deterministic component $\bz$ takes the form:
 \[
@@
 \]
-with $N=10^4$, $M=150$, $p_1=10$, $p_2=33$ and $R=1.4$. Besides, the additive noise is chosen to be Gaussian: $\bw\sim\cN(\bzero,\bI)$.
+with $M=150$, $p_1=10$, $p_2=33$ and $R=1.4$. Besides, the additive noise is chosen to be Gaussian: $\bw\sim\cN(\bzero,\bI)$.
 
 \paragraph{Influence of the noise variance $\sigma^2$} Here, the size of the training dataset is set to $K=450$.
-Then, the forecasting algorithm is run on $1000$ realizations on the discrete signal $\bx$ for three different values of $\sigma$, logarithmically equi-spaced from $10^{-3}$ to $10^{-1}$.
+Then, the forecasting algorithm is run on $1000$ realizations on the discrete signal $\bx$ for three different values of $\sigma$, logarithmically equispaced from $10^{-3}$ to $10^{-1}$.
 For each of these values, we determine the experimental bias, denoted as $\mu_{\xp}[N-1+\ell]$, and experimental variance, denoted as $\gamma_{\xp}[N-1+\ell,N-1+\ell]$, in function of the forecasting sample index $\ell$ (going from $1$ to $L=100$).
@@
 As expected, this result highlights the fact that the forecasting variance increases linearly with respect to $\sigma^2$.
-Surprisingly, this result shows that the forecasting variance {\em slightly} decreases with $\ell$, what is counterintuitive.
-%
-%{\color{red} [HT: need a more convincing explanation.] } 
-%It should be noted that, contrary to what expression~\eqref{eq:cov.error.2} suggests, smaller values of $\sigma$ do not cause a decrease of the experimental variance, but an increase.
-This comes from the fact that when $\sigma$ is small, the matrix $\bX\bX^T$ becomes ill-conditioned.
-The calculation of the forecasting matrix $\tilde\bA$ in~\eqref{eq:lse} is then strongly disturbed.
+The {\em slight} decrease of the forecasting variance with $\ell$ is characteristic of the stationarity of the forecast signal.
 
@@
 \paragraph{Influence of the training dataset size $K$} Here, the noise variance $\sigma$ is set to $\sigma=10^{-2}$.
-Then, the forecasting algorithm is run on 3000 realizations on the discrete signal $\bx$ for three different values of $K$, logarithmically equi-spaced from $4.5\times 10^{2}$ to $2\times 10^{3}$.
+Then, the forecasting algorithm is run on 3000 realizations on the discrete signal $\bx$ for three different values of $K$, logarithmically equispaced from $4.5\times 10^{2}$ to $2\times 10^{3}$.
 For each of these values, we determine the experimental bias $\mu_{\xp}[N-1+\ell]$ and variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$ in function of the forecasting sample index $\ell$ (going from $1$ to $500$).
 
-As in the previous study, the experimental bias vanishes when $K$ increases, what confirms the approximation result~\eqref{eq:mean.error}.
+As in the previous study, the experimental bias vanishes when $K$ increases, confirming the approximation result~\eqref{eq:mean.error}.
 Besides, the experimental variance  is displayed on Fig.~\ref{fig:res.size.sine}, and compared with the associated theoretical variance~\eqref{eq:cov.error.2}.
 Each color corresponds to these experimental results obtained for a given value of $K$.
-This result validates the asymptotic behavior provided by~\eqref{eq:cov.error.2}, and we can show that the second order moment $\gamma[\ell,\ell]$ is less dependent on $K$ when $K$ is large.
+This result validates the asymptotic behavior provided by~\eqref{eq:cov.error.2}, and we can show that the second-order moment $\gamma[\ell,\ell]$ is less dependent on $K$ when $K$ is large.
 
@@
 
-To highlight the fact that the linear dynamical model is sufficient to catch most of the dynamical behavior of signals following the AHM, we compare the performance of the Algorithm~\ref{alg:boundary} with a simple extension obtained by pointwise symmetrization~\cite{Kharitonenko02wavelet}.
-We also evaluate the performance of reference forecasting algorithm that could be used for extending such signals.
+To highlight the fact that the linear dynamical model is sufficient to catch most of the dynamical behavior of signals following the AHM, we compare the performance of Algorithm~\ref{alg:boundary} with a simple extension obtained by pointwise symmetrization~\cite{Kharitonenko02wavelet}.
+We also evaluate the performance of reference forecasting algorithms that could be used for extending such signals.
 These methods are:
@@
 \item The EDMD has been developed by Williams \etal~\cite{Williams15data}.
-The proposed algorithm is a way to obtain an approximation of the so-called Koopman operator of the observed system, which theoretically allows to catch dynamic of nonlinear systems~\cite{Korda18linear}.
+The proposed algorithm is a way to obtain an approximation of the so-called Koopman operator of the observed system, which theoretically allows catching dynamic of nonlinear systems~\cite{Korda18linear}.
 \item The GPR~\cite{Rasmussen06gaussian} is a method relying on a probabilistic dynamical model.
-That one is based on the Gaussian process structure, and therefore offer more flexibility in the type of dynamic that could be modeled than the linear model~\eqref{eq:dyn.model}.
-\item The TBATS method~\cite{DeLivera11forecasting} is based on a classical decomposition of times series into a trend, a seasonal and an ARMA components, with a specific dynamic for the seasonal component.
+That one is based on the Gaussian process structure, and therefore offers more flexibility in the type of dynamic that could be modeled than the linear model~\eqref{eq:dyn.model}.
+\item The TBATS method~\cite{DeLivera11forecasting} is based on a classical decomposition of times series into a trend, seasonal and ARMA components, with a specific dynamic for the seasonal component.
 This model demands the estimation of numerous parameters and, by implication, may be slow.
@@
 They show that the naive extension we propose gives satisfying results, in particular in comparison with the point-symmetric extension.
-Besides, even though the other more sophisticated methods, like GPR, give MSE values that have a slightly smaller standard deviation, these methods are substantially limited by the computing time they require, which prevent them from being used to exploit real-time data.
-Thus, {\sf SigExt} is the extension method that optimize the trade-off between the forecasting quality and the computing time.
-%That is why, it is implemented in our algorithm for the reduction of boundary effects.
+Besides, even though the other more sophisticated methods, like GPR, give MSE values that have a slightly smaller standard deviation, these methods are substantially limited by the computing time they require, which prevents them from being used to exploit real-time data.
+Thus, {\sf SigExt} is the extension method that optimizes the trade-off between forecasting quality and computing time.
 
@@
 
-In the aim of comparing the different techniques, we use a criterion, proposed in~\cite{Daubechies16conceft}, that quantify the distance between a given TF representation and the optimal one.
+In the aim of comparing the different techniques, we use a criterion, proposed in~\cite{Daubechies16conceft}, that quantifies the distance between a given TF representation and the optimal one.
 It is built in analogy with the optimal transport distance, which enables quantifying the distance between two probability density functions.
-Let us generically denote a time frequency representation $\ccQ$.
+Let us generically denote a time-frequency representation $\ccQ$.
 Then, for $t$ fixed, we consider the following probability density function: $p_\ccQ^t(\xi) = \left.
 |\ccQ(\xi,t)|^2 \middle/ \int_\RR |\ccQ(\nu,t)|^2\dd\nu \right.$.
-At each instant $t$, we can then determine the optimal transport distance $d_{t}$ between the two densities.
+At each instant $t$, we then determine the optimal transport distance $d_{t}$ between the two densities.
 It is given by the $L^1$ norm of the difference between the associated distribution functions.
@@
 
-From that large signal, we build a dataset of $378$ non-overlapping signals of 60 seconds, \ie~$N=6000$.
+From that long signal, we build a dataset of $378$ non-overlapping signals of 60 seconds, \ie~$N=6000$.
 On each of these pieces of signal, we implement the forecasting method introduced in section~\ref{ssse:res.ahm}, including the {\sf SigExt} method detailed in Algorithm~\ref{alg:extension}.
-However, the TBATS extension method in not implemented here because of its excessive computing time.
-We forecast $7$-second-long extensions on each segment of signal, corresponding to $L =700$.
+However, the TBATS extension method is not implemented here because of its excessive computing time.
+We forecast $7$-second-long extensions on each segment of the signal, corresponding to $L =700$.
 Thus, in order to catch slowly varying dynamical behaviors, the size of the training signal $M$ is chosen so that $M=\lfloor 1.5L\rfloor$.
@@
 The left of Fig.~\ref{fig:THO.failure} illustrates one of those cases, where {\sf SigExt} fails to catch the fast varying dynamic of the instantaneous amplitude to satisfactorily forecast the signal.
-The EDMD an Symmetric extensions are more robust to those situations, as shown on this example and in Table~\ref{tab:THO}.
+The EDMD and Symmetric extensions are more robust to those situations, as shown in this example, and in Table~\ref{tab:THO}.
 Nevertheless, {\sf SigExt} provides a sufficiently relevant extension to give TF representations sparingly affected by boundary effects.
-On the right of Fig.~\ref{fig:THO.failure}, we display comparison between the right boundary of SST of the same segment of signal (top-right), and its boundary-free SST obtained after the {\sf SigExt} forecasting (bottom-right).
-The extension of the instantaneous frequency visible on the right side of the image, illustrates the reduction of boundary effects produced despite an inaccurate signal forecasting.
+On the right of Fig.~\ref{fig:THO.failure}, we display a comparison between the right boundary of SST of the same segment of signal (top-right), and its boundary-free SST obtained after the {\sf SigExt} forecasting (bottom-right).
+The extension of the instantaneous frequency visible on the right side of the image illustrates the reduction of boundary effects produced despite an inaccurate signal forecasting.
 
-We then apply {\sf BoundEffRed} (\ie~Algorithm~\ref{alg:boundary}) for diverse TF representations: STFT, SST, RS, as well as concentration of frequency and time (ConceFT), a generalized multitaper SST-based representation introduced in~\cite{Daubechies16conceft}.
+We then apply {\sf BoundEffRed} (\ie~Algorithm~\ref{alg:boundary}) for diverse TF representations: STFT, SST, RS, as well as \textit{Concentration of Frequency and Time} (ConceFT), a generalized multitaper SST-based representation introduced in~\cite{Daubechies16conceft}.
 In Table~\ref{tab:THO}, we give the averaged performance index~\eqref{eq:index.perf}, evaluated the whole TF representations (including boundaries).
@@
 Finally, in order to verify the ability of {\sf BoundEffRed} to be implemented in real time, we evaluate the computation time required for each iteration of Algorithm~\ref{alg:boundary}, updating of the boundary-free TF representation.
-This check is performed on the real-time SST of a 10-minute sample of the respiratory signal.
+This check is first performed on the real-time SST of a 10-minute sample of the respiratory signal.
 As the SST is temporally sub-sampled by a ratio of 20, the update should not exceed $20/\fs=0.2$~sec.
@@
 A portion of this signal is displayed in Fig.~\ref{fig:ppg}.
-The estimated 5 seconds extensions of this segment obtained by {\sf SigExt}, EDMD and GPR forecastings  are superimposed to the ground-truth extension in Fig.~\ref{fig:ppg}.
+The estimated 5 seconds extensions of this segment obtained by {\sf SigExt}, EDMD, and GPR forecastings  are superimposed to the ground-truth extension in Fig.~\ref{fig:ppg}.
 
@@
 
-We divide the signal into 32-second-long segments, and apply Algorithm~\ref{alg:boundary} on each piece.
+We divide the signal into 32-second-long segments and apply Algorithm~\ref{alg:boundary} on each piece.
 We provide in Table~\ref{tab:otd.ppg} the performance index $D$ of the boundary-free TF representations averaged over the signals.
-For all the considered TF representations, the results clearly shows that our algorithm reduces the boundary effects.
-%This highlights the ability of our approach to limit the distortion due the boundary effects and provide a more accurate representations.
-Even though, on this signal, the {\sf SigExt} extension yields TF representations slightly more sensitive to boundary effects than the extensions given by EDMD or GPR, it is the only technique that allows a real-time implementation.
+For all the considered TF representations, the results clearly show that our algorithm reduces the boundary effects.
+Even though on this signal, the {\sf SigExt} extension yields TF representations slightly more sensitive to boundary effects than the extensions given by EDMD or GPR, it is the only technique that allows a real-time implementation.
 
@@
 
-On the bottom-right panel of Fig.~\ref{fig:ex.intro}, we display the SST resulting from the {\sf BoundEffRed} strategy, applied to the portion of PPG displayed on Fig.~\ref{fig:ppg}.
-We clearly observe an improvement of the quality of the SST near boundaries.
+On the bottom-right panel of Fig.~\ref{fig:ex.intro}, we display the SST resulting from the {\sf BoundEffRed} strategy, applied to the portion of PPG displayed in Fig.~\ref{fig:ppg}.
+We distinctly observe an improvement in the quality of the SST near boundaries.
 Indeed, the blurring visible when zooming on the right boundary of the SST has almost vanished.
@@
 
-To evaluate the influence of the noise level on the performance of {\sf BoundEffRed}, we artificially add a Gaussian noise to the measured PPG signal.
+To evaluate the influence of the noise level on the performance of {\sf BoundEffRed}, we artificially add Gaussian noise to the measured PPG signal.
 It is thus an additional noise to the measurement noise actually contained in the signal.
@@
 We notice that STFT is slightly more sensitive to noise than SST or RS, and STFT performs the best when the added noise is small, \ie~when the SNR is great.
-%On one hand, the robustness to noise of the SST and the RS is the direct consequence of the approximation result of~\cite{Daubechies16conceft}, discussed in section~\ref{sse:perf.BoundEffRed}.
-On the other hand, the sensitivity to noise of the STFT may be interpreted as the inability to forecast the values of the TF coefficients where only noise is active, while sharp TF representations, such as SST or RS, ideally vanish at these points.
 

### 2020-07-16 Notes/results.tex
paper

@@
 This result confirms the theoretical result~\eqref{eq:mean.error}.
-
-On Fig.~\ref{fig:res.noise.sine}, we display the experimental variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$ for each value of $\sigma$.
+In Fig.~\ref{fig:res.noise.sine}, we display the experimental variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$ for each value of $\sigma$.
 The associated theoretical asymptotic forecasting variance~\eqref{eq:cov.error.2} is also displayed in solid line.
@@
 The {\em slight} decrease of the forecasting variance with $\ell$ is characteristic of the stationarity of the forecast signal.
+This shows that as $\ell$ increases, the prediction becomes less dependent on the measurement noise.
 
 \begin{figure}
+\centering
 %\includegraphics[width=.24\textwidth]{biasNoiseSine.eps}

### 2020-07-17 Algorithm/BoundEffRed_RT.m
polish code

@@ -1,0 +1,113 @@
+function [TFRtot, dt] = BoundEffRed_RT(xtot,fs,forecastMethod,basicTF,VideoWriting)
+% BOUNDEFFRED_RT Real-time implementation of BoundEffRed
+% Usage:	[TFRtot, dt] = BoundEffRed_RT(xtot,fs,forecastMethod,basicTF,VideoWriting)
+%
+% Input:
+%   xtot: signal to analyze
+%   fs: sampling frequency
+%   forecastMethod: forecasting method
+%   basicTF: time-frequency representation and associated parameters
+%   VideoWriting: When set to 1, the real-time TF represention image is recorded 
+%
+% Output:
+%   TFRtot: Whole time-frequency representation
+%   dt: iteration time
+
+
+%% Parameters
+Ntot = length(xtot) ;
+
+HOP = 1 ;
+L = round((basicTF.win-1)/2) ; % extension length
+extM = round(1.5*L) ; % dimension of embedding / signals length
+extK = round( 2.5*extM ) ;  % number of points to estimate A / size of datasets
+
+Nt = extK + extM + 1 ; % segments length
+Lo = round(1.5*L/basicTF.hop); % overlap in TF domain
+LL = round(L/basicTF.hop); % extension in TF domain
+
+freqs = fs * (basicTF.fmin:basicTF.df:basicTF.fmax) ;
+
+%% Initialization
+n0 = 1 ;
+n1 = n0+Nt-1 ;
+
+t = (n0:n1)/fs ;
+x = xtot(n0:n1) ;
+
+switch basicTF.representation
+    case 'SST'
+        [~, ~, TFRcurrent, ~, ~] = ConceFT_sqSTFT_C(x, basicTF.fmin, basicTF.fmax, basicTF.df, basicTF.hop, basicTF.win, 1, 10, 1, 0, 0) ;
+    case 'conceFT'
+        [~, ~, ~, TFRcurrent, ~] = ConceFT_sqSTFT_C(x, basicTF.fmin, basicTF.fmax, basicTF.df, basicTF.hop, basicTF.win, 1, 10, 1, 0, 0) ;
+    case 'STFT'
[…]

### 2020-07-17 Algorithm/SigExtension.m

@@ -1,4 +1,22 @@
 function xx = SigExtension(x,fs,HOP,extK,extM,extSEC,method)
+% SIGEXTENSION Extend signal on both sides by forecasting
+% Usage:	xext = forecasting(x,L,HOP,extK,extM,method,side)
+%
+% Input:
+%   x: signal to be forecasted
+%   fs: sampling frequency
+%   HOP: subsampling rate for forecasting
+%   extK: size of the dataset for forecasting
+%   extM: lengths of the segments used for forecasting
+%   extSEC: extension length in seconds
+%   method: forcasting method. To be chosen between
+%       'SigExt': Least square estimation
+%       'edmd': Empirical dynamical Mode decompostion
+%       'gpr': Gaussian process regression
+%       'symmetrization': Symmetric extension
+%   side (optionnal): left blank for forward forecasting, set to 'backward' otherwise
+% Output:
+%   xx: extended signal
 
 L = round(fs*extSEC);
 

### 2020-07-17 Algorithm/forecasting.m

@@ -1,4 +1,21 @@
 function xext = forecasting(x,L,HOP,extK,extM,method,varargin)
+% FORECASTING Predict future values of a signal
+% Usage:	xext = forecasting(x,L,HOP,extK,extM,method,side)
+%
+% Input:
+%   x: signal to be forecasted
+%   L: number of samples to be forecasted
+%   HOP: subsampling rate for forecasting
+%   extK: size of the dataset for forecasting
+%   extM: lengths of the segments used for forecasting
+%   method: forcasting method. To be chosen between
+%       'SigExt': Least square estimation
+%       'edmd': Empirical dynamical Mode decompostion
+%       'gpr': Gaussian process regression
+%       'symmetrization': Symmetric extension
+%   side (optionnal): left blank for forward forecasting, set to 'backward' otherwise
+% Output:
+%   xext: forecasted extension
 
 if (~isempty(varargin))&&(isequal(varargin{1},'backward'))
     x = flipud(x);
@@ -17,7 +34,7 @@ end
 %% Estimate the parameters of the forecasting model
 switch method.name
     
-    case 'lse'
+    case {'SigExt','lse'}
         A = (Y*X') / (X*X') ; % least square estimation
         
     case 'lseV'
@@ -66,7 +83,7 @@ switch method.name
             xext(kk) = zpred ;
         end
         
-    case {'lse','lseV'}
+    case {'SigExt','lse','lseV'}
         Z = zeros(extM,L) ;
         Z(:,1) = A*Y(:,end) ;
         for kk = 2:L

### 2020-07-17 Scripts/BoundEffRedExamples/RealTime.m

@@ -5,94 +5,31 @@ addpath('../../Signals/');
 
 %% load THO signal
 data = load('PPG.mat');
+
 subT = 2 ; % Sub sampling (pre-processing)
 xtot = data.PPG(1:subT:end);
 fs = 100/subT ; % sampling frequency
-Ntot = length(xtot) ;
-ttot = linspace(0,(Ntot-1)/fs,Ntot) ;
+
+minutes = 3 ; 
+Nmax = minutes*60*fs ;
+xtot = xtot(1:Nmax) ; % shorten signal
+
+%% Parameters
+forecastMethod.name = 'lse' ;
+
+basicTF.representation = 'SST' ;
+
+basicTF.hop = round(20/subT) ;
+basicTF.win = 501 ; % window length (in samples)
+basicTF.fmin = 0.5/fs ;
+basicTF.fmax = 4/fs ;
+basicTF.df = 2e-4 ;
 
 VideoWriting = 0 ; % Change to 1 when recording
 
-%% Forecasting parameters
-method.name = 'lse' ;
-HOP = 1 ;
-extSEC = 5 ; % the extension is of extSEC second
-L = round(extSEC*fs) ;
-extM = round(1.5*L) ; % dimension of embedding / signals length
-extK = round( 2.5*extM ) ;  % number of points to estimate A / size of datasets
-
-basicTF.hop = round(20/subT) ;
-basicTF.win = 2*L+1 ; % window length (in samples)
-fmin = 0.5/fs ;
-fmax = 4/fs ;
-df = 2e-4 ;
-
-freqs = fs * (fmin:df:fmax) ;
-
-Nt = extK+extM+1 ; % segments length
-Lo = round(1.5*L/basicTF.hop); % overlap in TF domain
-LL = round(L/basicTF.hop); % extension in TF domain
-%% Initialization
-n0 = 1 ;
-n1 = n0+Nt-1 ;
-
-t = ttot(n0:n1) ;
-x = xtot(n0:n1) ;
-[~, ~, SSTcurrent, ~, tfrsqticEXT] = ConceFT_sqSTFT_C(x, fmin, fmax,...
-            df, basicTF.hop, basicTF.win, 1, 10, 1, 0, 0) ;
-SSTtot = SSTcurrent ;
-        
-figure;
-imagesc(t,freqs,log1p(abs(SSTtot)/5e0)); colormap(1-gray);
[…]

### 2020-07-17 TimeFrequencyScaleRep/ConceFT_rsSTFT.m

@@ -21,7 +21,7 @@ N = length(x) ;
 
 %=======================================
 
-fprintf(['Run ordinary RS\n']) ;
+%fprintf(['Run ordinary RS\n']) ;
 [tfr, tfrtic, tfrrs, tfrrstic] = rsSTFTbase(x, lowFreq, highFreq, alpha, hop, h(1,:)', Dh(1,:)', 1);
 
 

### 2020-07-17 TimeFrequencyScaleRep/sqSTFTbase.m

@@ -69,9 +69,9 @@ end
 
 %====================================================================
 	%% run STFT and reassignment rule
-tfr = zeros(N/2, tLen); 	% for h
+tfr = zeros(fLen, tLen) ; %zeros(N/2, tLen); 	% for h
 tfrtic = linspace(0, 0.5, N/2)' ;
-tfrsq = zeros(fLen, tLen); 
+tfrsq = zeros(fLen, tLen) ; 
 tfrsqtic = linspace(lowFreq, highFreq, fLen)' ;
 
 
@@ -139,7 +139,7 @@ for tidx = 1:tLen
         end
     end
 
-	tfr(:, tidx) = tf0(1:N/2) ;
+	tfr(:, tidx) = tf0(Lidx:Hidx); % tf0(1:N/2) ;
 	tfrsq(:, tidx) = sst ;
 
 end

### 2020-07-17 Notes/results.tex
polish code
code and typos

@@
 The {\em slight} decrease of the forecasting variance with $\ell$ is characteristic of the stationarity of the forecast signal.
-This shows that as $\ell$ increases, the prediction becomes less dependent on the measurement noise.
+This shows that as $\ell$ increases, the prediction becomes less correlated with the measurement noise.
 
@@
 \label{ssse:res.ahm}
-We now consider a signal satisfying the AHM so that the instantaneous frequencies and amplitudes of its components vary over time.
+We now consider a signal satisfying the AHM so that the instantaneous frequencies and amplitudes of its components vary from time to time.
 The deterministic component $\bz$ of the random vector $\bx$ (constructed following the model~\eqref{eq:model.noise}) takes the following form, for all $n\in\{1,\ldots,N\}$:

### 2020-07-17 Algorithm/BoundEffRed_RT.m

@@ -37,11 +37,11 @@ x = xtot(n0:n1) ;
 
 switch basicTF.representation
     case 'SST'
-        [~, ~, TFRcurrent, ~, ~] = ConceFT_sqSTFT_C(x, basicTF.fmin, basicTF.fmax, basicTF.df, basicTF.hop, basicTF.win, 1, 10, 1, 0, 0) ;
+        [~, ~, TFRcurrent, ~] = sqSTFT_C(x, basicTF.fmin, basicTF.fmax, basicTF.df, basicTF.hop, basicTF.win, 1, 10, 0, 0) ;
     case 'conceFT'
-        [~, ~, ~, TFRcurrent, ~] = ConceFT_sqSTFT_C(x, basicTF.fmin, basicTF.fmax, basicTF.df, basicTF.hop, basicTF.win, 1, 10, 1, 0, 0) ;
+        [~, ~, ~, TFRcurrent, ~] = ConceFT_sqSTFT_C(x, basicTF.fmin, basicTF.fmax, basicTF.df, basicTF.hop, basicTF.win, 1, 10, basicTF.MT, 0, 0) ;
     case 'STFT'
-        [TFRcurrent, ~, ~, ~, ~] = ConceFT_sqSTFT_C(x, basicTF.fmin, basicTF.fmax, basicTF.df, basicTF.hop, basicTF.win, 1, 10, 1, 0, 0) ;
+        [TFRcurrent, ~] = STFT_C(x, basicTF.fmin, basicTF.fmax, basicTF.df, basicTF.hop, basicTF.win, 1, 10) ;
     case 'RS'
         [~, ~, TFRcurrent, ~, ~] = ConceFT_rsSTFT(x, basicTF.fmin, basicTF.fmax, basicTF.df, basicTF.hop, basicTF.win, 1, 10, 1) ;
 end
@@ -74,11 +74,11 @@ while n1<Ntot
     
     switch basicTF.representation
         case 'SST'
-            [~, ~, TFRxx, ~, ~] = ConceFT_sqSTFT_C(xxPREC, basicTF.fmin, basicTF.fmax, basicTF.df, basicTF.hop, basicTF.win, 1, 10, 1, 0, 0) ;
+            [~, ~, TFRxx, ~] = sqSTFT_C(xxPREC, basicTF.fmin, basicTF.fmax, basicTF.df, basicTF.hop, basicTF.win, 1, 10, 0, 0) ;
         case 'conceFT'
[…]

### 2020-07-17 Algorithm/approxKoopman.m

@@ -1,6 +1,15 @@
 function [Xi,mu,phi_end] = approxKoopman(X,Y,sigma2)
-% evaluate Koopman modes V, eigen values Mu, and eigen function Psi*Xi
-% EDMD (from paper of Hua et al.)
+% APPROXKOOPMAN Evaluate Koopman modes, eigen values, and eigen function (from paper of Hua et al.)
+% Usage:	[Xi,mu,phi_end] = approxKoopman(X,Y,sigma2)
+%
+% Input:
+%   X: input dataset
+%   Y: output dataset
+%   sigma2: shape parameter
+% Output:
+%   Xi: Koopman modes
+%   Mu: Koopman eigen values
+%   phi_end: Koopman eigen functions
 
 M = size(X,1) ;
 

### 2020-07-17 Scripts/BoundEffRedExamples/RealTime.m

@@ -17,13 +17,17 @@ xtot = xtot(1:Nmax) ; % shorten signal
 %% Parameters
 forecastMethod.name = 'lse' ;
 
-basicTF.representation = 'SST' ;
+basicTF.representation = 'conceFT' ;
 
 basicTF.hop = round(20/subT) ;
 basicTF.win = 501 ; % window length (in samples)
 basicTF.fmin = 0.5/fs ;
 basicTF.fmax = 4/fs ;
 basicTF.df = 2e-4 ;
+
+if strcmp(basicTF.representation,'conceFT')
+    basicTF.MT = 10 ;
+end
 
 VideoWriting = 0 ; % Change to 1 when recording
 

### 2020-07-17 TimeFrequencyScaleRep/STFT_C.m

@@ -1,0 +1,7 @@
+function [tfr, tfrtic] = STFT_C(x, lowFreq, highFreq, alpha, hop, WinLen, dim, supp)
+
+% generate the window for short time Fourier transform (STFT)
+[h, ~, ~] = hermf(WinLen, dim, supp) ;
+[tfr, tfrtic] = STFTbase(x, lowFreq, highFreq, alpha, hop, h(1,:)');
+
+end

### 2020-07-17 TimeFrequencyScaleRep/STFTbase.m

@@ -1,0 +1,50 @@
+function [tfr, tfrtic] = STFTbase(x, lowFreq, highFreq, alpha, tDS, h)
+
+[xrow,xcol] = size(x) ;
+t = 1:length(x) ;
+tLen = length(t(1:tDS:length(x))) ;
+
+	% for tfr
+N = length(-0.5+alpha:alpha:0.5) ;
+
+Lidx = round( (N/2)*(lowFreq/0.5) ) + 1 ; 
+Hidx = round( (N/2)*(highFreq/0.5) ) ; 
+fLen = Hidx - Lidx + 1 ;
+
+
+
+%====================================================================
+	%% check input signals
+if (xcol~=1)
+    error('X must have only one column');
+elseif highFreq > 0.5
+    error('TopFreq must be a value in [0, 0.5]');
+elseif (tDS < 1) || (rem(tDS,1)) 
+    error('tDS must be an integer value >= 1');
+end
+
+[hrow,hcol] = size(h); Lh = (hrow-1)/2; 
+if (hcol~=1)||(rem(hrow,2)==0)
+    error('H must be a smoothing window with odd length');
+end
+
+
+%====================================================================
+	%% run STFT
+tfr = zeros(fLen, tLen) ;
+tfrtic = linspace(lowFreq, highFreq, fLen)' ;
+
+for tidx = 1:tLen
+
+    ti = t((tidx-1)*tDS+1); 
+    tau = -min([round(N/2)-1,Lh,ti-1]):min([round(N/2)-1,Lh,xrow-ti]);
+    indices = rem(N+tau,N)+1;
+    norm_h=norm(h(Lh+1+tau));
+
+	tf0 = zeros(N, 1) ;
+    tf0(indices) = x(ti+tau).*conj( h(Lh+1+tau)) /norm_h;
+    tf0 = fft(tf0) ;
+
+	tfr(:, tidx) = tf0(Lidx:Hidx);
+
+end

### 2020-07-18 Notes/BoundEffRed.tex
code and typos
typos

@@
 This phenomenon is usually understood as the \emph{boundary effect}.
-We display on the bottom left of Fig.~\ref{fig:ex.intro} the result of the SST of a PPG (see section~\ref{ssse:ppg} for a comprehensive description).
+The result of the SST of a PPG is shown in the lower left corner of Fig.~\ref{fig:ex.intro} (see section~\ref{ssse:ppg} for a comprehensive description).
 The distortion resulting from boundary effects is clearly visible on the right side of this representation.
@@
 
-The paper is organized in the following.
+The paper is organized in the following way.
 In section~\ref{se:algo}, we provide an extension method based on a linear dynamic model.
@@
 A more fundamental improvement would be to perform the forecasting step directly in the TF plane, to specifically predict the evolution of each instantaneous frequency.
-This would lead to seeing a TF representation as a multivariate time series (corresponding to each frequency band).
+This would lead to seeing a TF representation as a multivariate time series, with each frequency band containing a univariate time series.
 We will explore these possibilities in our future work.

### 2020-07-18 Notes/results.tex
precision
typos

@@
 Then, the forecasting algorithm is run on 3000 realizations on the discrete signal $\bx$ for three different values of $K$, logarithmically equispaced from $4.5\times 10^{2}$ to $2\times 10^{3}$.
-For each of these values, we determine the experimental bias $\mu_{\xp}[N-1+\ell]$ and variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$ in function of the forecasting sample index $\ell$ (going from $1$ to $500$).
+For each of these values, we determine the experimental bias $\mu_{\xp}[N-1+\ell]$ and variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$ in function of the forecasting sample index $\ell$, going from $1$ to $500$.
 
 As in the previous study, the experimental bias vanishes when $K$ increases, confirming the approximation result~\eqref{eq:mean.error}.
-Besides, the experimental variance  is displayed on Fig.~\ref{fig:res.size.sine}, and compared with the associated theoretical variance~\eqref{eq:cov.error.2}.
+Besides, the experimental variance  is displayed in Fig.~\ref{fig:res.size.sine}, and compared with the associated theoretical variance~\eqref{eq:cov.error.2}.
 Each color corresponds to these experimental results obtained for a given value of $K$.
@@
 We now consider a signal satisfying the AHM so that the instantaneous frequencies and amplitudes of its components vary from time to time.
-The deterministic component $\bz$ of the random vector $\bx$ (constructed following the model~\eqref{eq:model.noise}) takes the following form, for all $n\in\{1,\ldots,N\}$:
+The deterministic component $\bz$ of the random vector $\bx$, constructed following the model~\eqref{eq:model.noise}, takes the following form, for all $n\in\{1,\ldots,N\}$:
 \[
@@
 We then apply {\sf BoundEffRed} (\ie~Algorithm~\ref{alg:boundary}) for diverse TF representations: STFT, SST, RS, as well as \textit{Concentration of Frequency and Time} (ConceFT), a generalized multitaper SST-based representation introduced in~\cite{Daubechies16conceft}.
+In this application and throughout the document, we use a Gaussian window as the analysis window, regardless of the representation involved.
 In Table~\ref{tab:THO}, we give the averaged performance index~\eqref{eq:index.perf}, evaluated the whole TF representations (including boundaries).

### 2020-07-21 Notes/BoundEffRed.tex
simulations

@@
 This phenomenon is usually understood as the \emph{boundary effect}.
-The result of the SST of a PPG is shown in the lower left corner of Fig.~\ref{fig:ex.intro} (see section~\ref{ssse:ppg} for a comprehensive description).
+The result of the SST of a PPG is shown in the lower-left corner of Fig.~\ref{fig:ex.intro} (see section~\ref{ssse:ppg} for a comprehensive description).
 The distortion resulting from boundary effects is clearly visible on the right side of this representation.
@@
 There also exist extension schemes based on physically relevant dynamical models, such as the Extended Dynamic Mode Decomposition~\cite{Williams15data} (EDMD), the Gaussian process regression~\cite{Rasmussen06gaussian,Roberts13Gaussian} (GPR), the Trigonometric, Box-Cox transformation, ARMA errors, Trend and Seasonal components (TBATS) algorithm~\cite{DeLivera11forecasting}, the dynamic linear models \cite{west2006bayesian}, and long short-term memory model \cite{vlachas2018data}.
-It is a far from an exhaustive list, and we refer readers with interest to \cite{hyndman2018forecasting} for a friendly monograph on the general forecasting topic.
+It is far from an exhaustive list, and we refer readers with interest to \cite{hyndman2018forecasting} for a friendly monograph on the general forecasting topic.
 While the second class of extension schemes gives better-extended signals than the first class, they generally have a great computational cost.

### 2020-07-22 Notes/algorithm.tex
new fig and notes

@@
 where $\fs$ denotes the sampling frequency. 
-Let $M$ and $K$ be two integers such that $M<N$ and $K+M<N$. Then, for all $k\in\{0,\ldots,K-1\}$, we extract from $\bx\in\RR^N$ the sub-signal $\bx_k\in\RR^M$ given by:
+Let $M$ and $K$ be two integers such that $M<N$ and $K+M<N$. Then, for all $k\in\{0,\ldots,K-1\}$, we extract from $\bx\in\RR^N$ the subsignal $\bx_k\in\RR^M$ given by:
 \begin{equation}
@@
 \end{equation} 
-These sub-signals are gathered into the matrix $\bX\in\RR^{M\times K}$ such that:
+These subsignals are gathered into the matrix $\bX\in\RR^{M\times K}$ such that:
 \begin{equation*}
@@
 \end{equation*}
-Notice that these sub-signals are overlapping each other. Indeed, $\bx_{k+1}$ is a shifting of $\bx_k$ from one sample. We also consider the matrix $\bY\in\RR^{M\times K}$ given by:
+Notice that these subsignals are overlapping each other. Indeed, $\bx_{k+1}$ is a shifting of $\bx_k$ from one sample. We also consider the matrix $\bY\in\RR^{M\times K}$ given by:
 \begin{equation*}
@@
 
-\subsection{Step 1: Extension by forecasting}
+\subsection{Step 1: Extension by Forecasting}
 
-\paragraph{Dynamical model and forecasting} 
+\paragraph{Dynamical Model and Forecasting} 
 Establishing a dynamical model consists in determining the relation linking $\bY$ to $\bX$, that is finding a function $f$ so that
@@
 In other words, we adopt a classical strategy in the study of dynamical systems, the {\em linearization} of a nonlinear phenomenon, when the system is sufficiently regular.
-Notice that this linearized dynamical model can be written equivalently in function of the sub-signals $\bx_k$, as:
+Notice that this linearized dynamical model can be written equivalently in function of the subsignals $\bx_k$, as:
 \begin{equation}
@@
 
-\paragraph{Model estimation} To estimate the matrix $\bA$, we consider the simple but numerically efficient least square estimator. That is, we solve the following problem:
+\paragraph{Model Estimation} To estimate the matrix $\bA$, we consider the simple but numerically efficient least square estimator. That is, we solve the following problem:
 \begin{equation}
@@
 
-\paragraph{Signal extension} Since we are building a real-time algorithm, we consider that only the right ``side'' of the signal has to be extended, the left ``side'' being fixed since it only concerns the past values of the signal.
+\paragraph{Signal Extension} Since we are building a real-time algorithm, we consider that only the right ``side'' of the signal has to be extended, the left ``side'' being fixed since it only concerns the past values of the signal.
 We therefore construct the extended signal $\tilde\bx\in\RR^{N+L}$ concatenating the observed signal $\bx$, and the forward forecasting $\tilde\bx_{\mathrm{fw}}\in\RR^L$.

### 2020-07-22 Notes/results.tex
new fig and notes

@@
 
-\subsection{Evaluation of the forecasting performance}
+\subsection{Evaluation of the Forecasting Performance}
 In that section, we first evaluate the quality of the forecasting step and compare it to the theoretical results provided by Theorem~\ref{th:error}.
@@
 
-\paragraph{Influence of the noise variance $\sigma^2$} Here, the size of the training dataset is set to $K=450$.
+\paragraph{Influence of the Noise Variance $\sigma^2$} Here, the size of the training dataset is set to $K=450$.
 Then, the forecasting algorithm is run on $1000$ realizations on the discrete signal $\bx$ for three different values of $\sigma$, logarithmically equispaced from $10^{-3}$ to $10^{-1}$.
@@
 
-\paragraph{Influence of the training dataset size $K$} Here, the noise variance $\sigma$ is set to $\sigma=10^{-2}$.
-Then, the forecasting algorithm is run on 3000 realizations on the discrete signal $\bx$ for three different values of $K$, logarithmically equispaced from $4.5\times 10^{2}$ to $2\times 10^{3}$.
+\paragraph{Influence of the Training Dataset Size $K$} Here, the noise variance $\sigma$ is set to $\sigma=10^{-2}$.
+Then, {\sf SigExt} is run on 3000 realizations of the discrete signal $\bx$ for three different values of $K$, logarithmically equispaced from $4.5\times 10^{2}$ to $2\times 10^{3}$.
 For each of these values, we determine the experimental bias $\mu_{\xp}[N-1+\ell]$ and variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$ in function of the forecasting sample index $\ell$, going from $1$ to $500$.
@@
 
-\subsubsection{Adaptive harmonic model}
+\subsubsection{Adaptive Harmonic Model}
 \label{ssse:res.ahm}
@@
 \centering
-\caption{AHM signal: performance of the extension methods.}
+\caption{AHM Signal: Performance of the Extension Methods}
 \begin{tabular}{|c||c|c|c|}
@@
 
-\subsection{Evaluation of the quality of the boundary effects reduction}
+\subsection{Quality of the Reduction of Boundary Effects}
 
@@
 
-\subsubsection{Respiratory signal}
+\subsubsection{Respiratory Signal}
 We first consider a respiratory signal of 6 hours 20 minutes. This signal, a small portion of which is displayed in Fig.~\ref{fig:tho}, is sampled at $\fs=100$~Hz.
@@
 
-From that long signal, we build a dataset of $378$ non-overlapping signals of 60 seconds, \ie~$N=6000$.
+From that long signal, we build a dataset of $378$ non\-overlapping signals of 60 seconds, \ie~$N=6000$.
 On each of these pieces of signal, we implement the forecasting method introduced in section~\ref{ssse:res.ahm}, including the {\sf SigExt} method detailed in Algorithm~\ref{alg:extension}.
@@
 \centering
-\caption{Respiratory signal: Performance of extension methods and associated boundary-free TF representations.}
+\caption{Respiratory Signal: Performance of the Extension Methods and the Associated Boundary-Free TF Representations}
 \begin{tabular}{|c||c||c|c|c|c|}
@@
 This check is first performed on the real-time SST of a 10-minute sample of the respiratory signal.
-As the SST is temporally sub-sampled by a ratio of 20, the update should not exceed $20/\fs=0.2$~sec.
+As the SST is temporally subsampled by a ratio of 20, the update should not exceed $20/\fs=0.2$~sec.
 This is indeed the case, as no iteration lasted longer than $0.17$ sec on a 6-Core Xeon CPU running at 3.5~GHz and 64~GB of RAM.
@@
 \centering
-\caption{PPG signal: Performance of the boundary-free TF representations according to the extension method.}
+\caption{PPG Signal: Performance of the Boundary-Free TF Representations According to the Extension Method}
 \begin{tabular}{|c||c|c|c|}
@@
    \hhline{|=#=|=|=|}
-%   Without extension & $2.52\times 10^{-2}$ & $9.41\times 10^{-2}$ & $1.03\times 10^{-1}$ \\
-%   \hline
-   {\sf SigExt} & $0.425$ & $0.713$ & $0.687$ \\
+   {\sf SigExt} & $0.280$ & $0.309$ & $0.367$ \\
    \hline
-   Symmetric & $1.335$ & $1.340$ & $1.258$ \\
+   Symmetric & $1.168$ & $1.209$ & $1.310$ \\
    \hline
-   EDMD & $0.406$ & $0.642$ & $0.615$ \\
+   EDMD & $0.289$ & $0.319$ & $0.375$ \\
    \hline
-   GPR & $0.448$ & $0.766$ & $0.730$ \\
+   GPR & $0.276$ & $0.303$ & $0.361$ \\
    \hline
@@
 
-On the bottom-right panel of Fig.~\ref{fig:ex.intro}, we display the SST resulting from the {\sf BoundEffRed} strategy, applied to the portion of PPG displayed in Fig.~\ref{fig:ppg}.
+On the bottom-right panel of Fig.~\ref{fig:ex.intro}, we display the SST resulting from {\sf BoundEffRed} applied to the portion of PPG displayed in Fig.~\ref{fig:ppg}.
 We distinctly observe an improvement in the quality of the SST near boundaries.
@@
 It is thus an additional noise to the measurement noise actually contained in the signal.
-Fig.~\ref{fig:otd.noise} shows the averaged OTD of {\sf BoundEffRed} for different values of the Signal to Noise Ratio (SNR).
-We notice that STFT is slightly more sensitive to noise than SST or RS, and STFT performs the best when the added noise is small, \ie~when the SNR is great.
+Fig.~\ref{fig:otd.noise} shows the averaged performance index of {\sf BoundEffRed} for different values of the Signal to Noise Ratio (SNR).
+This graph illustrates the robustness to noise of {\sf BoundEffRed}.
+Indeed, whatever the representation considered, the reduction of boundary effects is relatively independent of the noise magnitude.
 

### 2020-07-22 Scripts/EvalPerfBoundEffRed/EvalPPGnoise.m

@@ -34,13 +34,13 @@ nXP = 12 ;
 StdNoise = logspace(-3.5,-1,nXP) ;
 for n = 1:nXP
     x = x0 + StdNoise(n)*randn(N0,1) ;
-    TFR = {'sstSTFT','RS'} ;
+    TFR = {'sst','RS'} ;
     [forecastErr, CompTime, OTD] = SucForecast(x,fs,methods,HOP,N,extM,extK,extSEC,TFR,basicTF) ;
     
     NoiseForecastErr(n) = mean(forecastErr.LSE) ;
-    NoiseTFR.SST(n) = mean(OTD.sst.LSE) ;
-    NoiseTFR.STFT(n) = mean(OTD.stft.LSE) ;
-    NoiseTFR.RS(n) = mean(OTD.rs.LSE) ;
+    NoiseTFR.SST(n) = mean(OTD.sst.LSE./OTD.sst.S) ;
+    NoiseTFR.STFT(n) = mean(OTD.stft.LSE./OTD.stft.S) ;
+    NoiseTFR.RS(n) = mean(OTD.rs.LSE./OTD.rs.S) ;
 end
 
 SNR = 20*log10(std(x0)./StdNoise) ;

### 2020-07-22 Notes/BoundEffRed.tex

@@
 \usepackage[english]{babel}
-%\usepackage{subfig}
 \usepackage{color}
@@
 
-\title{An Efficient Forecasting Approach for the Real-Time Reduction of Boundary Effects in Time-Frequency Representations}
+\title{An Efficient Forecasting Approach for the Real-Time Reduction of Boundary Effects\\ in Time-Frequency Representations}
 
@@
 \caption{A segment of PPG signal (top) and the right boundary of a TF representation determined by the SST without extension (bottom left), and the right boundary of a TF representation determined by the SST with the proposed boundary effect reduction algorithm by forecasting (bottom right).
-The window length for the SST is 12~sec.}
+The window length for the SST is 12~seconds.}
 \label{fig:ex.intro}
@@
 It is far from an exhaustive list, and we refer readers with interest to \cite{hyndman2018forecasting} for a friendly monograph on the general forecasting topic.
-While the second class of extension schemes gives better-extended signals than the first class, they generally have a great computational cost.
+Whereas the second class of extension schemes gives better-extended signals than the first class, they generally have a great computational cost.
 
@@
 \end{enumerate}
-Thus, assuming that the quality of the extension step is sufficient, the analysis results obtained that way will be less sensitive to the boundary effects than the result of the analysis tool applied directly to the non-extended signal.
+Thus, assuming that the quality of the extension step is sufficient, the analysis results obtained that way will be less sensitive to the boundary effects than the result of the analysis tool applied directly to the nonextended signal.
 We claim and prove that forecasting oscillatory signals based on the simple dynamical model combined with the simple least square approach is sufficient for reducing the boundary effects for the TF analysis, or other kernel-based analysis.
@@
 
-\section{Theoretical performance}
+\section{Theoretical Performance}
 \label{se:theoretical}
@@
 
-\section{Numerical results}
+\section{Numerical Results}
 \label{se:results}

### 2020-07-22 Scripts/EvalPerfBoundEffRed/EvalPPG.m
new XP

@@ -23,7 +23,7 @@ L = round(extSEC*fs) ;
 extM = round( 1.5*L ) ; % dimension of embedding / signals length
 extK = round( 2.5*extM ) ;  % number of points to estimate A / size of datasets
 
-TFR = {'conceFT'} ;
+TFR = {'conceFT','RS'} ;
 basicTF.hop = 10 ;
 basicTF.win = 1501 ;
 basicTF.fmin = 0 ;

### 2020-07-22 Notes/BoundEffRed.tex
result

@@
 \usepackage[T1]{fontenc}
-\usepackage[english]{babel}
 \usepackage{color}
@@
 Moreover, thanks to the low computational time, the numerical results show that this strategy allows us to switch to a real-time implementation of {\sf BoundEffRed}, unlike other existing forecasting methods.
-The numerical results also confirmed the robustness to noise of {\sf BoundEffRed}, as well as its ability to be applied to many time-frequency representations.
+The numerical results also confirmed the robustness to noise of {\sf BoundEffRed}, as well as its ability to be applied to many TF representations.
 %Other numerical examples are also included, such as the application of {\sf BoundEffRed} to other types of biomedical signals such as ECGs.

### 2020-07-22 Notes/results.tex
new XP
result

@@
 It is built in analogy with the optimal transport distance, which enables quantifying the distance between two probability density functions.
-Let us generically denote a time-frequency representation $\ccQ$.
+Let us generically denote a TF representation $\ccQ$.
 Then, for $t$ fixed, we consider the following probability density function: $p_\ccQ^t(\xi) = \left.
@@
 \label{ssse:ppg}
-We perform a study similar to the previous one on a 640 second-long photoplethysmogram (PPG) signal extracted from the Physionet dataset~\cite{Pimentel17toward, Goldberger00physiobank}, sampled at $\fs=125$~Hz.
+We perform a study similar to the previous one on a $640$-second-long photoplethysmogram (PPG) signal extracted from the Physionet dataset~\cite{Pimentel17toward, Goldberger00physiobank}, sampled at $\fs=125$~Hz.
 A portion of this signal is displayed in Fig.~\ref{fig:ppg}.
@@
 \caption{PPG Signal: Performance of the Boundary-Free TF Representations According to the Extension Method}
-\begin{tabular}{|c||c|c|c|}
+\begin{tabular}{|c||c|c|c|c|}
   \hline
-   \multirow{2}{40pt}{\centering Extension method} & \multicolumn{3}{c|}{Averaged performance index $D$} \\
-   \cline{2-4}
-      & STFT & SST & ConceFT\\
-   \hhline{|=#=|=|=|}
-   {\sf SigExt} & $0.280$ & $0.309$ & $0.367$ \\
+   \multirow{2}{40pt}{\centering Extension method} & \multicolumn{4}{c|}{Averaged performance index $D$} \\
+   \cline{2-5}
+      & STFT & SST & ConceFT & RS\\
+   \hhline{|=#=|=|=|=|}
+   {\sf SigExt} & $0.280$ & $0.309$ & $0.367$ & $0.534$ \\
    \hline
-   Symmetric & $1.168$ & $1.209$ & $1.310$ \\
+   Symmetric & $1.168$ & $1.209$ & $1.310$ & $0.983$ \\
    \hline
-   EDMD & $0.289$ & $0.319$ & $0.375$ \\
+   EDMD & $0.289$ & $0.319$ & $0.375$ & $0.503$ \\
    \hline
-   GPR & $0.276$ & $0.303$ & $0.361$ \\
+   GPR & $0.276$ & $0.303$ & $0.361$ & $0.544$ \\
    \hline
@@
 \includegraphics[width=.48\textwidth]{OTDBoundEffRed.eps}
-\caption{PPG signal. Averaged OTD of {\sf BoundEffRed} in function of the SNR.}
+\caption{Performance index of {\sf BoundEffRed} applied on a PPG, in function of the SNR.}
 \label{fig:otd.noise}

### 2020-07-23 Notes/results.tex
new fig.

@@
 It is thus an additional noise to the measurement noise actually contained in the signal.
-Fig.~\ref{fig:otd.noise} shows the averaged performance index of {\sf BoundEffRed} for different values of the Signal to Noise Ratio (SNR).
-This graph illustrates the robustness to noise of {\sf BoundEffRed}.
-Indeed, whatever the representation considered, the reduction of boundary effects is relatively independent of the noise magnitude.
+Fig.~\ref{fig:otd.noise} shows the averaged performance index of {\sf BoundEffRed} for different values of the Signal-to-Noise Ratio (SNR).
+Whatever the representation considered, {\sf BoundEffRed} is relatively robust to noise, as long as the SNR remains above $10$~dB.
+Below this level, the reduction of boundary effects gradually deteriorates.
 

### 2020-07-23 Scripts/EvalPerfBoundEffRed/EvalPPGnoise.m

@@ -24,6 +24,8 @@ L = round(extSEC*fs) ;
 extM = round( 1.5*L ) ; % dimension of embedding / signals length
 extK = round( 2.5*extM ) ;  % number of points to estimate A / size of datasets
 
+% TFR parameters
+TFR = {'conceFT','RS'} ;
 basicTF.hop = 10 ;
 basicTF.win = 1501 ;
 basicTF.fmin = 0 ;
@@ -36,7 +38,6 @@ NoiseForecastErr = zeros(nXP,1) ;
 StdNoise = logspace(-3.5,-1,nXP) ;
 for n = 1:nXP
     x = x0 + StdNoise(n)*randn(N0,1) ;
-    TFR = {'conceFT','RS'} ;
     [forecastErr, CompTime, OTD] = SucForecast(x,fs,methods,HOP,N,extM,extK,extSEC,TFR,basicTF) ;
     
     NoiseForecastErr(n) = mean(forecastErr.LSE) ;

### 2020-07-23 Scripts/EvalPerfBoundEffRed/figPPGnoise.m

@@ -14,7 +14,7 @@ set(gca,'fontsize',24) ;
 
 figure;
 plot(SNR,NoiseTFR.STFT,'^',SNR,NoiseTFR.SST,'+',SNR,NoiseTFR.conceFT,'s',SNR,NoiseTFR.RS,'o','linewidth',2,'MarkerSize',11);
-ylim([0 0.6]) ;
+%ylim([0 0.6]) ;
 xlabel('SNR (dB)','interpreter','latex'); ylabel('Averaged performance index $D$','interpreter','latex'); grid on;
 legend('STFT','SST','ConceFT','RS');
 set(gca,'fontsize',24) ;

### 2020-07-24 Notes/BoundEffRed.tex
notes

@@
 We have shown theoretically that the chosen dynamic model is sufficient to extend signals formed by a sum of sine waves.
-Moreover, thanks to the low computational time, the numerical results show that this strategy allows us to switch to a real-time implementation of {\sf BoundEffRed}, unlike other existing forecasting methods.
+Moreover, the low computational time allows us to switch to a real-time implementation of {\sf BoundEffRed}, unlike other existing forecasting methods.
 The numerical results also confirmed the robustness to noise of {\sf BoundEffRed}, as well as its ability to be applied to many TF representations.
-%Other numerical examples are also included, such as the application of {\sf BoundEffRed} to other types of biomedical signals such as ECGs.
+An additional application to ECG is included into the supplementary material.
 
@@
 A more fundamental improvement would be to perform the forecasting step directly in the TF plane, to specifically predict the evolution of each instantaneous frequency.
-This would lead to seeing a TF representation as a multivariate time series, with each frequency band containing a univariate time series.
-We will explore these possibilities in our future work.
+This leads to seeing a TF representation as a multivariate time series, with each frequency band containing a univariate time series.
+These avenues will be explored in our future work.
 

### 2020-07-24 Notes/Supplementary material/SupMat.tex
small updates
paper
notes

@@
 \usepackage{dsfont}
+\usepackage{hhline}
+\usepackage{graphicx}
 
@@
 
-\title{An Approach for the Reduction of Boundary Effects in Time-Frequency Representations:\\ Part II}
+\title{An Efficient Forecasting Approach for the Real-Time Reduction of Boundary Effects\\ in Time-Frequency Representations:\\ Part II}
 
@@
 
-\setcounter{equation}{26}
+\setcounter{equation}{27}
+\setcounter{figure}{7}
+\setcounter{table}{3}
 
@@
 \nonumber
-&= \be_M^T\left(\tilde\bA^\ell-\tilde\bA_0^\ell\right) \\
-&= \be_M^T\left(\left((\bS^{(1)}+\bE^{(1)})(\bS^{(0)}+\bE^{(0)})^\inv\right)^\ell- \tilde\bA_0^\ell\right)\ .
+&= \be_M^T\left(\tilde\bA^\ell-\bA_0^\ell\right) \\
+&= \be_M^T\left(\left((\bS^{(1)}+\bE^{(1)})(\bS^{(0)}+\bE^{(0)})^\inv\right)^\ell- \bA_0^\ell\right)\ .
 \label{eq:error.vec}
@@
 \end{align*}
-with $\tilde\bz_k^T= \begin{pmatrix} \bz_k^T & \bz_{k+1}[0] \end{pmatrix}$ and $\tilde\bw_k^T= \begin{pmatrix} \bw_k^T & \bw_{k+1}[0] \end{pmatrix}$.
+with $\tilde\bz_k^T= \begin{pmatrix} \bz_k^T & \bz_{k+1}[M-1] \end{pmatrix}$ and $\tilde\bw_k^T= \begin{pmatrix} \bw_k^T & \bw_{k+1}[M-1] \end{pmatrix}$.
 Then, $\bg_1$ is a Gaussian random vector because it is a linear combination of Gaussian random vectors.
@@
 
-\subsection{Expression of the bias $\bmu$.}
+\subsection{Expression of the Bias $\bmu$.}
 Clearly, $\bmu[n]=0$ when $n\in I$. When $n=N-1+\ell$, we have:
@@
 
-\subsection{Expression of the covariance $\bgamma$.}
+\subsection{Expression of the Covariance $\bgamma$.}
 We recall that $n>N$, we consequently denote $n=N-1+\ell$.  To derive the covariance, let us segregate the cases.
@@
 
-%\begin{align*}
-%\bgamma[\ell,\ell] &\leq \dfrac1K\left(\sqrt{\bz_K^T\EE\{{\bu^{(\ell)}}^T\bu^{(\ell)}\}\bz_K} + \sigma\sqrt{\EE\{(\bu^{(\ell)}\bw_K)^2\}}\right)^2 \\
-%&\ + \dfrac{2\sigma}{\sqrt{K}}\|\balpha_0^{(\ell)}\|\left(\sqrt{\bz_K^T\EE\{{\bu^{(\ell)}}^T\bu^{(\ell)}\}\bz_K} +\sigma\sqrt{\EE\{(\bu^{(\ell)}\bw_K)^2\}} \right)+ \sigma^2\|\balpha_0^{(\ell)}\|^2 - \bmu[\ell]^2 + o(\sigma^2) \\
-%\end{align*}
-%
-%Then, Isserlis formula gives us:
-%\begin{align*}
-%\EE\{(\bu^{(\ell)}\bw_K)^2\} &= \sum_{m,m'=0}^{M-1}\EE\{\bu^{(\ell)}[m]\bw_K[m]\bu^{(\ell)}[m']\bw_K[m']\}\\
-%&\underset{K\to\infty}{\sim} \sum_{m,m'=0}^{M-1}\EE\{\bu^{(\ell)}[m]\bw_K[m]\}\EE\{\bu^{(\ell)}[m']\bw_K[m']\}+\EE\{\bu^{(\ell)}[m]\bu^{(\ell)}[m']\}\EE\{\bw_K[m]\bw_K[m']\} \\
-%&\hspace{30pt}+\EE\{\bu^{(\ell)}[m]\bw_K[m']\}\EE\{\bu^{(\ell)}[m']\bw_K[m]\}\\
-%&\leq \Tr(\bGamma^{(\ell)}) + 2\sum_{m,m'=0}^{M-1}\sqrt{\bGamma^{(\ell)}[m,m]\bGamma^{(\ell)}[m',m']}
-%\end{align*}
-%So:
-%\begin{align*}
-%\bgamma[\ell,\ell] &\leq \dfrac1K\left(\sqrt{\bz_K^T\bGamma^{(\ell)}\bz_K} + \sigma\sqrt{\Tr(\bGamma^{(\ell)}) + 2\sum_{m,m'=0}^{M-1}\sqrt{\bGamma^{(\ell)}[m,m]\bGamma^{(\ell)}[m',m']}}\right)^2 \\
-%&\ + \dfrac{2\sigma}{\sqrt{K}}\|\balpha_0^{(\ell)}\|\left(\sqrt{\bz_K^T\bGamma^{(\ell)}\bz_K} +\sigma\sqrt{\Tr(\bGamma^{(\ell)}) + 2\sum_{m,m'=0}^{M-1}\sqrt{\bGamma^{(\ell)}[m,m]\bGamma^{(\ell)}[m',m']}} \right)+ \sigma^2\|\balpha_0^{(\ell)}\|^2 + o(\sigma^2) \\
-%\end{align*}
+\section{Application to an Electrocardiogram}
+We provide here an additional implementation of {\sf BoundEffRed}, applied to an electrocardiogram (ECG) dataset.
+The dataset is constructed from a $500$-second-long ECG, sampled at $\fs=200$~Hz, cut into $14$ segments of $35$ seconds each.
+Fig.~\ref{fig:ecg} depicts one of these subsignals, together with the estimated 5 seconds extensions obtained by {\sf SigExt}, EDMD, and GPR forecastings  are superimposed to the ground-truth extension.
+The sharp and spiky ECG patterns make the AHM model too simplistic to describe this type of signal.
+Consequently, the forecast produced by {\sf SigExt} is moderately satisfactory.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{ECGforecast.eps}
+\caption{Extended ECG (blue) obtained by the {\sf SigExt} forecasting (top), the EDMD forecasting (middle), and the GPR forecasting (bottom), superimposed with the ground truth signal (red).}
+\label{fig:ecg}
+\end{figure}
+
+Table~\ref{tab:otd.ecg} contains the performance index $D$ of the boundary-free TF representations, averaged over the $N$ subsignals, according to the extension method.
+As a result of the fair quality of the forecasts, the reduction of boundary effects is less significant than for PPG signal.
+Nevertheles, the results show that {\sf BoundEffRed} has the same efficiency when the {\sf SigExt} extension, the EDMD extension or the GPR extension is chosen.
+This justifies the choice of {\sf SigExt} for real-time implementation.
+
+\begin{table}
+\centering
+\caption{ECG: Performance of the Boundary-Free TF Representations According to the Extension Method}
+\begin{tabular}{|c||c|c|c|c|}
+  \hline
+   \multirow{2}{40pt}{\centering Extension method} & \multicolumn{4}{c|}{Averaged performance index $D$} \\
+   \cline{2-5}
+      & STFT & SST & ConceFT & RS\\
+   \hhline{|=#=|=|=|=|}
+   {\sf SigExt} & $0.713$ & $0.700$ & $0.753$ & $0.761$ \\
+   \hline
+   Symmetric & $1.395$ & $1.402$ & $1.416$ & $0.976$ \\
+   \hline
+   EDMD & $0.856$ & $0.792$ & $0.559$ & $0.688$ \\
+   \hline
+   GPR & $0.723$ & $0.734$ & $0.734$ & $0.761$ \\
+   \hline
+\end{tabular}
+\label{tab:otd.ecg}
+\end{table}
 

### 2020-07-24 Notes/results.tex
notes

@@
 
-We then apply {\sf BoundEffRed} (\ie~Algorithm~\ref{alg:boundary}) for diverse TF representations: STFT, SST, RS, as well as \textit{Concentration of Frequency and Time} (ConceFT), a generalized multitaper SST-based representation introduced in~\cite{Daubechies16conceft}.
+We then apply {\sf BoundEffRed} for diverse TF representations: STFT, SST, RS, as well as \textit{Concentration of Frequency and Time} (ConceFT), a generalized multitaper SST-based representation introduced in~\cite{Daubechies16conceft}.
 In this application and throughout the document, we use a Gaussian window as the analysis window, regardless of the representation involved.
@@
 
-We divide the signal into 32-second-long segments and apply Algorithm~\ref{alg:boundary} on each piece.
+We divide the signal into 32-second-long segments and apply {\sf BoundEffRed} on each piece.
 We provide in Table~\ref{tab:otd.ppg} the performance index $D$ of the boundary-free TF representations averaged over the signals.
 For all the considered TF representations, the results clearly show that our algorithm reduces the boundary effects.
-Even though on this signal, the {\sf SigExt} extension yields TF representations slightly more sensitive to boundary effects than the extensions given by EDMD or GPR, it is the only technique that allows a real-time implementation.
+In addition to being the only technique that allows real-time implementation, {\sf SigExt} produces TF representations that erase boundary effects about as efficiently as extensions given by EDMD or GPR.
 

### 2020-07-24 Scripts/BoundEffRedExamples/ECGexample.m

@@ -9,7 +9,7 @@ data = load('ECG');
 xtot = data.ECG ;
 fs = 200 ; % sampling frequency
 
-n0 = 42e3;
+n0 = 51e3;
 nf = 58e3;
 x = xtot(n0:nf);
 mu = mean(x) ;
@@ -21,16 +21,12 @@ N = length(x);
 %% Forecasting
 
 HOP = 1 ;
-extSEC = 11 ; % the extension is of extSEC second
+extSEC = 6 ; % the extension is of extSEC second
 L = round( extSEC*fs ) ;
 extM = round( 1.5*L ) ; % dimension of embedding / signals length
 extK = round( 2.5*extM );  % number of points to estimate A / size of datasets
-if extK + extM >length(x) - 10 
-    extK = extK/2 ; extM = extM/2 ;
-end
 
-
-method.name = 'lseV' ;
+method.name = 'SigExt' ;
 tic; 
 xxLSE = SigExtension(x,fs,HOP,extK,extM,extSEC,method); 
 LSEtime = toc;
@@ -48,49 +44,68 @@ GPRtime = toc;
 
 fprintf(' __________________________________________\n')
 fprintf('| Extension Method | Computing time (sec.) |\n')
-fprintf('|------------------------------------------|\n')
-fprintf('|  LSE extension   |        %.2f          |\n', LSEtime)
-fprintf('|  EDMD extension  |       %.2f          |\n', EDMDtime)
-fprintf('|  GPR extension   |      %.2f          |\n', GPRtime)
+fprintf('|------------------|-----------------------|\n')
+fprintf('|     SigExt       |        %.2f          |\n', LSEtime)
+fprintf('|  EDMD extension  |        %.2f          |\n', EDMDtime)
+fprintf('|  GPR extension   |       %.2f          |\n', GPRtime)
 fprintf('|__________________|_______________________|\n')
 
 
 %% Plot
 t = linspace(0, (N-1)/fs, N) ;
-tt = linspace(-extSEC, (N-1)/fs+extSEC, N+2*round(extSEC*fs)) ;
+tt = linspace(-L/fs, (N-1+L)/fs, N+2*L) ;
 
[…]

### 2020-07-24 Scripts/BoundEffRedExamples/PPGexample.m

@@ -28,7 +28,7 @@ if extK + extM >length(x) - 10
 end
 
 
-method.name = 'lseV' ;
+method.name = 'SigExt' ;
 tic; 
 xxLSE = SigExtension(x,fs,HOP,extK,extM,extSEC,method); 
 LSEtime = toc;
@@ -47,7 +47,7 @@ GPRtime = toc;
 fprintf(' __________________________________________\n')
 fprintf('| Extension Method | Computing time (sec.) |\n')
 fprintf('|------------------------------------------|\n')
-fprintf('|  LSE extension   |         %.2f          |\n', LSEtime)
+fprintf('|     SigExt       |         %.2f          |\n', LSEtime)
 fprintf('|  EDMD extension  |         %.2f          |\n', EDMDtime)
 fprintf('|  GPR extension   |       %.2f          |\n', GPRtime)
 fprintf('|__________________|_______________________|\n')
@@ -130,13 +130,12 @@ sstOTDGPR = slicedOT(tfrsqGPRw, tfrsqTRUEw) ;
 
 fprintf('==============SST===============\n')
 fprintf(' _______________________________\n')
-fprintf('|  Extension Method |    OTD    | \n')
+fprintf('|  Extension Method |  Index D  | \n')
 fprintf('|-------------------|-----------|\n')
-%fprintf('|   Short signal    | %.3e |\n', sstOTDshort)
-fprintf('| Without extension | %.3e |\n', sstOTDZP)
-fprintf('|   LSE extension   | %.3e |\n', sstOTDLSE)
-fprintf('|   EDMD extension  | %.3e |\n', sstOTDEDMD)
-fprintf('|   GPR extension   | %.3e |\n', sstOTDGPR)
+fprintf('|   Zero-padding    | %.3f |\n', sstOTDZP/sstOTDshort)
+fprintf('|      SigExt       | %.3f |\n', sstOTDLSE/sstOTDshort)
+fprintf('|   EDMD extension  | %.3f |\n', sstOTDEDMD/sstOTDshort)
+fprintf('|   GPR extension   | %.3f |\n', sstOTDGPR/sstOTDshort)
 fprintf('|___________________|___________|\n')
 
 %% CWT
@@ -188,13 +187,12 @@ cwtOTDGPR = slicedOT(scaloGPRw, scaloTRUEw) ;
 
 fprintf('================CWT==============\n')
[…]

### 2020-07-24 Scripts/EvalPerfBoundEffRed/ResultsBoundEffRed.m

@@ -20,7 +20,7 @@ fprintf('|      EDMD        |           %.3f        |     %.3e     |\n', CompTim
 fprintf('|      GPR         |         %.3f        |     %.3e     |\n', CompTime.GPR , mean(forecastErr.GPR) )
 fprintf('|__________________|________________________|___________________|\n\n')
 
-fprintf('=============Optimal Transport Distance============\n')
+fprintf('===============Performance Index D=================\n')
 fprintf(' __________________________________________________\n')
 fprintf('| Extension Method |  SST  |  STFT |ConceFT|   RS  |\n')
 fprintf('|------------------|-------|-------|-------|-------|\n')
@@ -73,7 +73,7 @@ xlabel('Time (s)'); ylabel('Signals'); xlim([40 tt(end)]); ylim([-20 20]);
 set(gca,'fontsize',20)
 
 
-fprintf('============Optimal Transport Distance=============\n')
+fprintf('===============Performance Index D=================\n')
 fprintf(' __________________________________________________\n')
 fprintf('| Extension Method |  SST  |  STFT |  RS   |ConceFT|\n')
 fprintf('|------------------|-------|-------|-------|-------|\n')
@@ -98,15 +98,15 @@ fprintf('|      EDMD        |           %.2f        |     %.3e     |\n', CompTim
 fprintf('|      GPR         |          %.2f        |     %.3e     |\n', CompTime.GPR , mean(forecastErr.GPR))
 fprintf('|__________________|_______________________|___________________|\n\n')
 
-fprintf('=========Optimal Transport Distance========\n')
-fprintf(' __________________________________________\n')
-fprintf('| Extension Method |  SST  |  STFT |  RS   |\n')
-fprintf('|------------------|-------|-------|-------|\n')
[…]

### 2020-07-24 Notes/Supplementary material/SupMat.tex
sup mat

@@
 The dataset is constructed from a $500$-second-long ECG, sampled at $\fs=200$~Hz, cut into $14$ segments of $35$ seconds each.
-Fig.~\ref{fig:ecg} depicts one of these subsignals, together with the estimated 5 seconds extensions obtained by {\sf SigExt}, EDMD, and GPR forecastings  are superimposed to the ground-truth extension.
+Fig.~\ref{fig:ecg} depicts the right boundary of one of these subsignals, together with the $6$-second extensions estimated by {\sf SigExt} (top panel), or EDMD (middle panel), or GPR (bottom panel).
+These extensions are superimposed to the ground-truth extension, plotted in red.
 The sharp and spiky ECG patterns make the AHM model too simplistic to describe this type of signal.

### 2020-07-24 Scripts/BoundEffRedExamples/FailureCase.m
matlab

@@ -55,7 +55,7 @@ SYMtime = toc;
 
 %% Plot
 t = linspace(0, (N-1)/fs, N) ;
-tt = linspace(-extSEC, (N-1)/fs+extSEC, N+2*round(extSEC*fs)) ;
+tt = linspace(-L/fs, (N-1+L)/fs, N+2*L) ;
 
 xxTRUE = ( xtot((n0-L):(n1+L)) - mu ) / s ;
 xxZP = [ zeros(L,1); x; zeros(L,1) ] ; % zero padding
@@ -84,8 +84,8 @@ set(gca,'FontSize',20); %xticklabels([]) ;
 basicTF.fmin = 0/fs ;
 basicTF.fmax = 1/fs ;
 basicTF.df = 1e-5 ;
-basicTF.hop = 10;
-basicTF.win = 2*L+1;
+basicTF.hop = 10 ;
+basicTF.win = 2*L+1 ;
 
 
 [~, SSTx, ~, tfrsqticX] = ConceFT_sqSTFT_C(x, basicTF.fmin, basicTF.fmax,...

### 2020-07-24 Scripts/BoundEffRedExamples/THOexample.m

@@ -24,9 +24,6 @@ extSEC = 7 ; % the extension is of extSEC second
 L = round(extSEC*fs) ;
 extM = round(1.5*L) ; % dimension of embedding / signals length
 extK = round( 2.5*extM ) ;  % number of points to estimate A / size of datasets
-if extK + extM >length(x) - 10 
-    extK = extK/2 ; extM = extM/2 ;
-end
 
 method.name = 'SigExt' ;
 tic; 
@@ -46,50 +43,59 @@ GPRtime = toc;
 
 %% Plot
 t = linspace(0, (N-1)/fs, N) ;
-tt = linspace(-extSEC, (N-1)/fs+extSEC, N+2*round(extSEC*fs)) ;
+tt = linspace(-L/fs, (N-1+L)/fs, N+2*L) ;
 
 xxTRUE = ( xtot((n0-L):(n1+L)) - mu ) / s ;
 xxZP = [ zeros(L,1); x; zeros(L,1) ] ; % zero padding
 
 figure;
-plot(tt,xxLSE,tt,xxEDMD,tt,xxGPR,tt,xxTRUE,'--',t,x,'linewidth',2); grid on;
-legend({'{\sf SigExt} Extended signal','EDMD Estimated Extended signal','GPR Estimated Extended signal','Ground truth Extended signal','Original signal'},'interpreter','latex'); 
-xlabel('Time (s)'); ylabel('Signals'); title('Time series');
 
-figure;
+subplot(3,1,1)
 plot(tt,xxLSE,tt,xxTRUE,'--',t,x,'linewidth',2); grid on;
-legend({'{\sf SigExt} Extended signal','Ground truth Extended signal','Original signal'},'interpreter','latex'); 
-xlabel('Time (s)'); ylabel('Signals'); axis tight; ylim([-2 2]);
-set(gca,'FontSize',20);
+ylabel('Signals'); axis tight; xlim([t(end)-1.5*extSEC tt(end)]);
+xticks([]);
+legend({'{\sf SigExt} extension','Ground truth extension','Original signal'},'location','northwest','interpreter','latex');
+set(gca,'fontsize',18);
 
-figure;
-plot(tt,xxTRUE,'linewidth',2); grid on;
[…]

### 2020-07-24 Scripts/EvalPerfBoundEffRed/EvalECG.m

@@ -13,7 +13,7 @@ xtot = data.ECG;
 
 fs = 200 ; % sampling frequency
 
-N = 7e3; % subsignals length
+N = 10e3; % subsignals length
 
 % Forecasting parameters
 methods = {'SigExt','symmetrization','edmd','gpr'} ;

### 2020-07-24 Scripts/EvalPerfBoundEffRed/EvalEEG.m

@@ -13,7 +13,7 @@ xtot = data.EEG;
 
 fs = 200 ; % sampling frequency
 
-N = 6e3; % subsignals length
+N = 3e3; % subsignals length
 
 % Forecasting parameters
 methods = {'SigExt','symmetrization','edmd','gpr'} ;

### 2020-07-24 Scripts/EvalPerfBoundEffRed/EvalPPG.m

@@ -25,7 +25,7 @@ extK = round( 2.5*extM ) ;  % number of points to estimate A / size of datasets
 
 TFR = {'conceFT','RS'} ;
 basicTF.hop = 10 ;
-basicTF.win = 1501 ;
+basicTF.win = 2*L+1 ;
 basicTF.fmin = 0 ;
 basicTF.fmax = 0.04 ;
 basicTF.df = 1e-5 ;

### 2020-07-24 Scripts/EvalPerfBoundEffRed/EvalPPGnoise.m

@@ -27,7 +27,7 @@ extK = round( 2.5*extM ) ;  % number of points to estimate A / size of datasets
 % TFR parameters
 TFR = {'conceFT','RS'} ;
 basicTF.hop = 10 ;
-basicTF.win = 1501 ;
+basicTF.win = 2*L+1 ;
 basicTF.fmin = 0 ;
 basicTF.fmax = 0.04 ;
 basicTF.df = 1e-5 ;

### 2020-07-24 Scripts/EvalPerfBoundEffRed/ResultsBoundEffRed.m

@@ -43,8 +43,8 @@ fprintf('|                  |                        |    Mean   |   Median  |\n
 fprintf('|------------------|------------------------|-----------|-----------|\n')
 fprintf('|     SigExt       |           %.3f        | %.3e | %.3e |\n', CompTime.LSE , mean(forecastErr.LSE(~isnan(OTD.sst.LSE))),median(forecastErr.LSE(~isnan(OTD.sst.LSE))) )
 fprintf('|  Symmetrization  |           %.3f        | %.3e | %.3e |\n', CompTime.SYM , mean(forecastErr.SYM), median(forecastErr.SYM) )
-fprintf('|      EDMD        |          %.3f        | %.3e | %.3e |\n', CompTime.EDMD , mean(forecastErr.EDMD), median(forecastErr.EDMD) )
-fprintf('|      GPR         |         %.3f        | %.3e | %.3e |\n', CompTime.GPR , mean(forecastErr.GPR(~isnan(OTD.sst.GPR))), median(forecastErr.GPR(~isnan(OTD.sst.GPR))) )
+% fprintf('|      EDMD        |          %.3f        | %.3e | %.3e |\n', CompTime.EDMD , mean(forecastErr.EDMD), median(forecastErr.EDMD) )
+% fprintf('|      GPR         |         %.3f        | %.3e | %.3e |\n', CompTime.GPR , mean(forecastErr.GPR(~isnan(OTD.sst.GPR))), median(forecastErr.GPR(~isnan(OTD.sst.GPR))) )
 fprintf('|__________________|________________________|___________|___________|\n\n')
 
 figure;
@@ -79,8 +79,8 @@ fprintf('| Extension Method |  SST  |  STFT |  RS   |ConceFT|\n')
 fprintf('|------------------|-------|-------|-------|-------|\n')
[…]

### 2020-07-24 Notes/results.tex
matlab
xp

@@
 Then, as long as the bias $\bmu[N-1+\ell]$ and the variance $\bgamma[N-1+\ell,N-1+\ell]$ of the forecasting estimator remain small for all $\ell$, the MSE takes small values either.
-Corresponding results are given in Table~\ref{tab:mse.sine}.
+The numerical results when $L=1000$ are given in Table~\ref{tab:mse.sine}.
 They show that the naive extension we propose gives satisfying results, in particular in comparison with the point-symmetric extension.
-Besides, even though the other more sophisticated methods, like GPR, give MSE values that have a slightly smaller standard deviation, these methods are substantially limited by the computing time they require, which prevents them from being used to exploit real-time data.
+Besides, even though the other more sophisticated methods, like GPR, are able to give slightly better forecasts, these methods are substantially limited by the computing time they require, which prevents them from being used to exploit real-time data.
 Thus, {\sf SigExt} is the extension method that optimizes the trade-off between forecasting quality and computing time.
@@
 \caption{AHM Signal: Performance of the Extension Methods}
-\begin{tabular}{|c||c|c|c|}
+\begin{tabular}{|c||c|c|}
   \hline
-   \multirow{2}{40pt}{\centering Extension method} & \multicolumn{2}{c|}{MSE}  & \multirow{2}{41pt}{Computing time (sec.)} \\
-   \cline{2-3} & Mean & Standard deviation & \\
-   \hhline{|=#=|=|=|}
-   {\sf SigExt} & $1.433\times 10^{-3}$ & $4.361\times 10^{-4}$ & $0.152$ \\
+   \multirow{2}{40pt}{\centering Extension method} & \multirow{2}{70pt}{\centering MSE averaged on $50$ realizations}  & \multirow{2}{80pt}{\centering Averaged computing time (sec.)} \\
+    &  & \\
+   \hhline{|=#=|=|}
+   {\sf SigExt} & $1.458$ & $0.152$ \\
    \hline
-   Symmetric & $1.019\times 10^{1}$ & $1.192\times 10^{2}$ & $0.002$ \\
+   Symmetric & $8.735$ & $0.002$ \\
    \hline
-   EDMD & $3.076\times 10^{-2}$ & $8.095\times 10^{-2}$ & $2.537$\\
+   EDMD & $2.313$ &$2.537$\\
    \hline
-   GPR & $1.436\times 10^{-3}$ & $4.346\times 10^{-4}$ & $146.331$ \\
+   GPR & $1.441$ &$146.331$ \\
    \hline
-   TBATS & $1.732\times 10^{-3}$ & $4.924\times 10^{-4}$ & $1837.120$ \\
+   TBATS & $1.501$ & $1267.120$ \\
    \hline
@@
 
-From that long signal, we build a dataset of $378$ non\-overlapping signals of 60 seconds, \ie~$N=6000$.
+From that long signal, we build a dataset of $379$ non\-overlapping signals of 60 seconds, \ie~$N=6000$.
 On each of these pieces of signal, we implement the forecasting method introduced in section~\ref{ssse:res.ahm}, including the {\sf SigExt} method detailed in Algorithm~\ref{alg:extension}.

### 2020-07-24 Scripts/BoundEffRedExamples/THOexample.m

@@ -99,16 +99,16 @@ basicTF.df = 1e-5 ;
 
 figure;
 subplot(2,2,1);
-imagesc(t(1:basicTF.hop:end),tfrsqtic*fs,log1p(abs(SSTx)/1e2));
+imagesc(t(1:basicTF.hop:end),tfrsqtic*fs,log1p(abs(SSTx)/1e2)); axis xy ;
 xlabel('Time (s)'); ylabel('Frequency (Hz)'); title('SST on original short signal');
 subplot(2,2,2);
-imagesc(tt(1:basicTF.hop:end),tfrsqticTRUE*fs,log1p(abs(SSTxxTRUE)/1e2)); xlim([0 t(end)]);
+imagesc(tt(1:basicTF.hop:end),tfrsqticTRUE*fs,log1p(abs(SSTxxTRUE)/1e2)); xlim([0 t(end)]); axis xy ;
 xlabel('Time (s)'); ylabel('Frequency (Hz)'); title('SST on original long signal');
 subplot(2,2,3);
-imagesc(tt(1:basicTF.hop:end),tfrsqticEXT*fs,log1p(abs(SSTxxLSE)/1e2)); xlim([0 t(end)]);
+imagesc(tt(1:basicTF.hop:end),tfrsqticEXT*fs,log1p(abs(SSTxxLSE)/1e2)); xlim([0 t(end)]); axis xy ;
 xlabel('Time (s)'); ylabel('Frequency (Hz)'); title('SST on LSE estimated long signal (short signal extended)');
 subplot(2,2,4);
-imagesc(tt(1:basicTF.hop:end),tfrsqticEXT*fs,log1p(abs(SSTxxEDMD)/1e2)); xlim([0 t(end)]);
+imagesc(tt(1:basicTF.hop:end),tfrsqticEXT*fs,log1p(abs(SSTxxEDMD)/1e2)); xlim([0 t(end)]); axis xy ;
 xlabel('Time (s)'); ylabel('Frequency (Hz)'); title('SST on EDMD estimated long signal (short signal extended)');
 
 %save('results','tfrsq3','tfrsq3EXT');

### 2020-07-24 Scripts/EvalPerfSigExt/CompPerfAHM.m

@@ -42,7 +42,7 @@ x0 = xx0( (L+1) : (L+N) ) ; % restriction to the measurement interval
 sigman = 1e-2 ;
 
 %% Forecasting
-nbXP = 1000 ;
+nbXP = 50 ;
 
 for ind = 1:nbXP
     noise = sigman*randn(N+2*L,1) ;
@@ -79,20 +79,20 @@ for ind = 1:nbXP
     
 end
 
-BiasXP.LSE = mean(MeanLSE) ;
-VarianceXP.LSE = mean(VarLSE) ;
+BiasXP.LSE = mean(MeanLSE,2) ;
+VarianceXP.LSE = mean(VarLSE,2) ;
 CPUtimeXP.LSE = mean(LSEtime) ;
 
-BiasXP.SYM = mean(MeanSYM) ;
-VarianceXP.SYM = mean(VarSYM) ;
+BiasXP.SYM = mean(MeanSYM,2) ;
+VarianceXP.SYM = mean(VarSYM,2) ;
 CPUtimeXP.SYM = mean(SYMtime) ;
 
-BiasXP.EDMD = mean(MeanEDMD) ;
-VarianceXP.EDMD = mean(VarEDMD) ;
+BiasXP.EDMD = mean(MeanEDMD,2) ;
+VarianceXP.EDMD = mean(VarEDMD,2) ;
 CPUtimeXP.EDMD = mean(EDMDtime) ;
 
-BiasXP.GPR = mean(MeanGPR) ;
-VarianceXP.GPR = mean(VarGPR) ;
+BiasXP.GPR = mean(MeanGPR,2) ;
+VarianceXP.GPR = mean(VarGPR,2) ;
 CPUtimeXP.GPR = mean(GPRtime) ;
 
 save('../../Results/PerfAHM','BiasXP','VarianceXP','CPUtimeXP');

### 2020-07-24 Scripts/EvalPerfSigExt/ResultsCompPerfAHM.m

@@ -15,14 +15,14 @@ MSE.EDMD = BiasXP.EDMD.^2 + VarianceXP.EDMD ;
 MSE.GPR = BiasXP.GPR.^2 + VarianceXP.GPR ;
 MSE.TBATS = BiasXP.TBATS.^2 + VarianceXP.TBATS ;
 
-fprintf('===================Forecasting Performance=========\n')
-fprintf(' ______________________________________________________\n')
-fprintf('|    Extension   |  Computing  |   Forecasting Error   |\n')
-fprintf('|     Method     | time (sec.) |    Mean   |  Variance |\n')
-fprintf('|----------------|-------------|-----------|-----------|\n')
-fprintf('|      LSE       |     %.3f   | %.3e | %.3e |\n', CPUtimeXP.LSE , mean(MSE.LSE), var(MSE.LSE))
-fprintf('| Symmetrization |     %.3f   | %.3e | %.3e |\n', CPUtimeXP.SYM , mean(MSE.SYM), var(MSE.SYM))
-fprintf('|      EDMD      |    %.3f   | %.3e | %.3e |\n', CPUtimeXP.EDMD , mean(MSE.EDMD), var(MSE.EDMD))
-fprintf('|      GPR       |   %.3f   | %.3e | %.3e |\n', CPUtimeXP.GPR , mean(MSE.GPR), var(MSE.GPR))
-fprintf('|      TBATS     |  %.3f   | %.3e | %.3e |\n', CPUtimeXP.TBATS , mean(MSE.TBATS), var(MSE.TBATS))
-fprintf('|________________|_____________|___________|___________|\n\n')
+fprintf('======== Forecasting Performance ========\n')
+fprintf(' ________________________________________\n')
+fprintf('|    Extension   |  Computing  | Averaged|\n')
+fprintf('|     Method     | time (sec.) |   MSE   |\n')
+fprintf('|----------------|-------------|---------|\n')
+fprintf('|      LSE       |     %.3f   |  %.3f  |\n', CPUtimeXP.LSE , mean(MSE.LSE))
+fprintf('| Symmetrization |     %.3f   |  %.3f  |\n', CPUtimeXP.SYM , mean(MSE.SYM))
+fprintf('|      EDMD      |    %.3f   |  %.3f  |\n', CPUtimeXP.EDMD , mean(MSE.EDMD))
[…]

### 2020-07-25 Notes/Supplementary material/SupMat.tex
matlab
real time improved

@@
 We provide here an additional implementation of {\sf BoundEffRed}, applied to an electrocardiogram (ECG) dataset.
-The dataset is constructed from a $500$-second-long ECG, sampled at $\fs=200$~Hz, cut into $14$ segments of $35$ seconds each.
+The dataset is constructed from a $500$-second-long ECG, sampled at $\fs=200$~Hz, cut into $10$ segments of $50$ seconds each.
 Fig.~\ref{fig:ecg} depicts the right boundary of one of these subsignals, together with the $6$-second extensions estimated by {\sf SigExt} (top panel), or EDMD (middle panel), or GPR (bottom panel).
@@
 
-Table~\ref{tab:otd.ecg} contains the performance index $D$ of the boundary-free TF representations, averaged over the $N$ subsignals, according to the extension method.
+Table~\ref{tab:otd.ecg} contains the median performance index $D$ of the boundary-free TF representations, over the $N$ subsignals evaluated, according to the extension method.
 As a result of the fair quality of the forecasts, the reduction of boundary effects is less significant than for PPG signal.
@@
   \hline
-   \multirow{2}{40pt}{\centering Extension method} & \multicolumn{4}{c|}{Averaged performance index $D$} \\
+   \multirow{2}{40pt}{\centering Extension method} & \multicolumn{4}{c|}{Median performance index $D$} \\
    \cline{2-5}
@@
    \hhline{|=#=|=|=|=|}
-   {\sf SigExt} & $0.713$ & $0.700$ & $0.753$ & $0.761$ \\
+   {\sf SigExt} & $0.584$ & $0.630$ & $0.462$ & $0.642$ \\
    \hline
-   Symmetric & $1.395$ & $1.402$ & $1.416$ & $0.976$ \\
+   Symmetric & $1.199$ & $1.354$ & $1.427$ & $0.943$ \\
    \hline
-   EDMD & $0.856$ & $0.792$ & $0.559$ & $0.688$ \\
+   EDMD & $0.538$ & $0.558$ & $0.496$ & $0.714$ \\
    \hline
-   GPR & $0.723$ & $0.734$ & $0.734$ & $0.761$ \\
+   GPR & $0.639$ & $0.588$ & $0.485$ & $0.616$ \\
    \hline

### 2020-07-25 Notes/results.tex
real time improved

@@
   \hline
-   \multirow{2}{40pt}{\centering Extension method} & \multirow{2}{70pt}{\centering MSE averaged on $50$ realizations}  & \multirow{2}{80pt}{\centering Averaged computing time (sec.)} \\
+   \multirow{2}{40pt}{\centering Extension method} & \multirow{2}{70pt}{\centering MSE averaged on $50$ realizations}  & \multirow{2}{80pt}{\centering Averaged computing time (seconds)} \\
     &  & \\
    \hhline{|=#=|=|}
-   {\sf SigExt} & $1.458$ & $0.152$ \\
+   {\sf SigExt} & $1.458$ & $1.069$ \\
    \hline
-   Symmetric & $8.735$ & $0.002$ \\
+   Symmetric & $8.735$ & $0.001$ \\
    \hline
-   EDMD & $2.313$ &$2.537$\\
+   EDMD & $2.313$ &$11.10$\\
    \hline
-   GPR & $1.441$ &$146.331$ \\
+   GPR & $1.441$ &$244.7$ \\
    \hline
-   TBATS & $1.501$ & $1267.120$ \\
+   TBATS & $1.501$ & $633.6$ \\
    \hline
@@
 In the aim of comparing the different techniques, we use a criterion, proposed in~\cite{Daubechies16conceft}, that quantifies the distance between a given TF representation and the optimal one.
-It is built in analogy with the optimal transport distance, which enables quantifying the distance between two probability density functions.
+It is built in analogy with the optimal transport distance, which quantify the distance between two probability density functions.
 Let us generically denote a TF representation $\ccQ$.
@@
 
-Finally, in order to verify the ability of {\sf BoundEffRed} to be implemented in real time, we evaluate the computation time required for each iteration of Algorithm~\ref{alg:boundary}, updating of the boundary-free TF representation.
-This check is first performed on the real-time SST of a 10-minute sample of the respiratory signal.
-As the SST is temporally subsampled by a ratio of 20, the update should not exceed $20/\fs=0.2$~sec.
-This is indeed the case, as no iteration lasted longer than $0.17$ sec on a 6-Core Xeon CPU running at 3.5~GHz and 64~GB of RAM.
-
 \subsubsection{Photoplethysmogram}
@@
 
-We divide the signal into 32-second-long segments and apply {\sf BoundEffRed} on each piece.
-We provide in Table~\ref{tab:otd.ppg} the performance index $D$ of the boundary-free TF representations averaged over the signals.
-For all the considered TF representations, the results clearly show that our algorithm reduces the boundary effects.
-In addition to being the only technique that allows real-time implementation, {\sf SigExt} produces TF representations that erase boundary effects about as efficiently as extensions given by EDMD or GPR.
+We divide the signal into 32-second-long segments and apply {\sf BoundEffRed} to each of them.
+Table~\ref{tab:otd.ppg} shows the performance index $D$ of the boundary-free TF representations, averaged over the signals.
+For all the TF representations considered, the results show that {\sf SigExt} reduces boundary effects about as efficiently as the extensions given by EDMD or GPR.
+Incidentally, the bottom-right panel of Fig.~\ref{fig:ex.intro} shows that the SST resulting from {\sf BoundEffRed} applied to the portion of PPG displayed in Fig.~\ref{fig:ppg} produces a significant improvement in the quality of the SST near boundaries.
+Indeed, the blurring visible when zooming on the right boundary of the SST has almost vanished.
+Real-time tracking of the instantaneous frequencies contained in the measured signal is therefore greatly facilitated.
 
@@
 
-On the bottom-right panel of Fig.~\ref{fig:ex.intro}, we display the SST resulting from {\sf BoundEffRed} applied to the portion of PPG displayed in Fig.~\ref{fig:ppg}.
-We distinctly observe an improvement in the quality of the SST near boundaries.
-Indeed, the blurring visible when zooming on the right boundary of the SST has almost vanished.
-The real-time tracking of the instantaneous frequencies contained in the measured signal is therefore largely facilitated.
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=.48\textwidth]{SSTBoundEffRed.eps}
-%\includegraphics[width=.48\textwidth]{zoomSSTBoundEffRed.eps}
-%\caption{Result of {\sf BoundEffRed} on the synchrosqueezing transform of a PPG (top) with a zoom on its right boundary (bottom). }
-%\label{fig:ppg.boundeffred}
-%\end{figure}
-
 To evaluate the influence of the noise level on the performance of {\sf BoundEffRed}, we artificially add Gaussian noise to the measured PPG signal.
@@
 \end{figure}
+
+\subsection{Real-Time Implementation}
+We simulate the real-time implementation of the SST from {\sf BoundEffRed} applied to the PPG described in section~\ref{ssse:ppg}, subsampled by a factor of $2$.
+The test is performed on a 2-Core Intel Core i5 CPU running at $1.7$~GHz and $7.7$~GB of RAM.
+
+For this signal, sampled at $\fs=65.5$~Hz, the suitable window is $8$ seconds long.
+Therefore, the extension must be over $4$ seconds, \ie~$L=250$.
+The forecasting step in each iteration of {\sf BoundEffRed} then takes no more than $t_{\mathrm{forecast}}=46$~ms.
+As indicated in section~\ref{sse:extension.TFR}, the update of the SST requires the calculation of $\lceil L/H \rceil$ new columns of the SST, where $H$ denotes the hop size.
+In this example, the frequency dimension of the SST being $512$, the computational time of one column is $t_{\mathrm{SST}}=2.08$~ms, on average.
+Thus, acceptable values of $H$ for real-time implementation verify:
+\[
+t_{\mathrm{forecast}} + \left\lceil \frac{L}H \right\rceil t_{\mathrm{SST}} < \frac{H}\fs\ . 
+\]
+In our example, taking $H\geq 4$~samples is sufficient, which is a reasonable latency. A video clip illustrating this is available in the GitHub repository.

### 2020-07-25 Algorithm/BoundEffRed_RT.m

@@ -1,4 +1,4 @@
-function [TFRtot, dt] = BoundEffRed_RT(xtot,fs,forecastMethod,basicTF,VideoWriting)
+function [TFRtot, ForecastTime, TFRtime] = BoundEffRed_RT(xtot,fs,forecastMethod,basicTF,VideoWriting)
 % BOUNDEFFRED_RT Real-time implementation of BoundEffRed
 % Usage:	[TFRtot, dt] = BoundEffRed_RT(xtot,fs,forecastMethod,basicTF,VideoWriting)
 %
@@ -11,7 +11,8 @@ function [TFRtot, dt] = BoundEffRed_RT(xtot,fs,forecastMethod,basicTF,VideoWriti
 %
 % Output:
 %   TFRtot: Whole time-frequency representation
-%   dt: iteration time
+%   ForecastTime: forecastion iteration time
+%   TFRtime: TF representation update time
 
 
 %% Parameters
@@ -23,8 +24,7 @@ extM = round(1.5*L) ; % dimension of embedding / signals length
 extK = round( 2.5*extM ) ;  % number of points to estimate A / size of datasets
 
 Nt = extK + extM + 1 ; % segments length
-Lo = round(1.5*L/basicTF.hop); % overlap in TF domain
-LL = round(L/basicTF.hop); % extension in TF domain
+LL = ceil(L/basicTF.hop); % extension in TF domain
 
 freqs = fs * (basicTF.fmin:basicTF.df:basicTF.fmax) ;
 
@@ -46,7 +46,7 @@ switch basicTF.representation
         [~, TFRcurrent, ~] = rsSTFT(x, basicTF.fmin, basicTF.fmax, basicTF.df, basicTF.hop, basicTF.win, 1, 10) ;
 end
 
-TFRtot = TFRcurrent ;
+TFRtot = TFRcurrent(:,end-LL) ;
         
 figure;
 imagesc(t,freqs,log1p(abs(TFRcurrent)/5e0)); colormap(1-gray);
@@ -62,30 +62,32 @@ end
 k = 1 ;
 while n1<Ntot
     t = (n0:n1)/fs ;
-    tic;
     x = xtot(n0:n1) ;
 
     %% Forecasting
+    tic;
     xext = forecasting(x,L,HOP,extK,extM,forecastMethod);
[…]

### 2020-07-25 Scripts/BoundEffRedExamples/RealTime.m

@@ -12,22 +12,22 @@ data = load('PPG.mat');
 
 subT = 2 ; % Sub sampling (pre-processing)
 xtot = data.PPG(1:subT:end);
-fs = 100/subT ; % sampling frequency
+fs = data.fs/subT ; % sampling frequency
 
 minutes = 3 ; 
 Nmax = minutes*60*fs ;
 xtot = xtot(1:Nmax) ; % shorten signal
 
 %% Parameters
-forecastMethod.name = 'lse' ;
+forecastMethod.name = 'SigExt' ;
 
-basicTF.representation = 'conceFT' ;
+basicTF.representation = 'SST' ;
 
-basicTF.hop = round(20/subT) ;
-basicTF.win = 501 ; % window length (in samples)
+basicTF.hop = round(10/subT) ;
+basicTF.win = 2*round(1000/2/subT)+1 ; % window length (in samples) /!\ must be odd
 basicTF.fmin = 0.5/fs ;
-basicTF.fmax = 4/fs ;
-basicTF.df = 2e-4 ;
+basicTF.fmax = 5.5/fs ;
+basicTF.df = (basicTF.fmax-basicTF.fmin)/512 ;
 
 if strcmp(basicTF.representation,'conceFT')
     basicTF.MT = 10 ;
@@ -37,7 +37,9 @@ VideoWriting = 0 ; % Change to 1 when recording
 
 %% Real-time SST
 
-[SSTtot, dt] = BoundEffRed_RT(xtot,fs,forecastMethod,basicTF,VideoWriting) ;
+[SSTtot, ForecastTime, TFRtime] = BoundEffRed_RT(xtot,fs,forecastMethod,basicTF,VideoWriting) ;
 
 dtmax = basicTF.hop/fs ;
-save('../../Results/RealTime','dt','dtmax')
+Hmin = ceil( fs/2 * ( mean(ForecastTime) + sqrt(mean(ForecastTime)^2 + 4*mean(TFRtime)/fs) ) );
+
+%save('../../Results/RealTime','ForecastTime','TFRtime','dtmax')

### 2020-07-25 Scripts/EvalPerfSigExt/CompPerfAHM.m

@@ -48,16 +48,16 @@ for ind = 1:nbXP
     noise = sigman*randn(N+2*L,1) ;
     x = x0.' + noise((L+1):(N+L)) ; % signal to be extended
 
-    method.name = 'lse' ;
+    method.name = 'SigExt' ;
     tic;
-    xxLSE = SigExtension(x,fs,HOP,extK,extM,extSEC,method).' ; % Forecasted signal via LSE
+    xxLSE = SigExtension(x,fs,HOP,extK,extM,extSEC,method).' ; % Forecasted signal via SigExt
     LSEtime(ind) = toc;
     MeanLSE(ind,:) = xxLSE((N+L+1):end) - xx0((N+L+1):end) ;
     VarLSE(ind,:) = ( xxLSE((N+L+1):end) - xx0((N+L+1):end) ).^2 ;
     
     method.name = 'symmetrization' ;
     tic;
-    xxSYM = SigExtension(x,fs,HOP,extK,extM,extSEC,method).' ; % Forecasted signal via LSE
+    xxSYM = SigExtension(x,fs,HOP,extK,extM,extSEC,method).' ;
     SYMtime(ind) = toc;
     MeanSYM(ind,:) = xxSYM((N+L+1):end) - xx0((N+L+1):end) ;
     VarSYM(ind,:) = ( xxSYM((N+L+1):end) - xx0((N+L+1):end) ).^2 ;

### 2020-07-25 TimeFrequencyScaleRep/ConceFT_rsSTFT_RT.m

@@ -1,0 +1,54 @@
+function [tfr, tfrrs, ConceFT, tfrtic] = ConceFT_rsSTFT_RT(x, lowFreq, highFreq, alpha, hop, WinLen, LL, dim, supp, MT)
+% CONCEFT_RSSTFT ConceFT Reassigment of the STFT
+% Usage: 
+% 	[tfr, tfrrs, ConceFT, tfrtic] = ConceFT_rsSTFT(x, lowFreq, highFreq, alpha, hop, WinLen, LL, dim, supp, MT)
+%
+% Input:
+%   x: signal to be analized
+%   lowFreq: relative lower frequency (0<lowFreq<Highfreq)
+%   highFreq: relative higher frequency (lowFreq<Highfreq<0.5)
+%   alpha: resolution in the frequency axis
+%   hop: hop size (in samples)
+%   WinLen: Window length (in samples)
+%   LL: Wextension length
+%   dim: order of the Hermite function used as window
+%   supp: half time support of the Hermite function used as window 
+%   MT = 1: ordinary RS; MT > 1: ConceFT
+% 
+% Output:
+%   tfr: STFT
+%   tfrrs: Reassigment of STFT
+%   ConceFT: ConceFT RS of STFT
+%   tfrtic: frequencies for which TFRs are evaluated
+
+
+%% Ordinary Reassigment
+[h, Dh, ~] = hermf(WinLen, dim, supp) ; % generate the window for short time Fourier transform (STFT)
+
+[tfr, tfrrs, tfrtic] = rsSTFTbase_RT(x, lowFreq, highFreq, alpha, hop, LL, h(1,:)', Dh(1,:)', 1);
+
+
+%% Multitapering
+ConceFT = [] ;
+
+if MT > 1
+
+	% Conceft
+    ConceFT = zeros(size(tfrrs)) ;
+
+    for ii = 1: MT
+		rv = randn(1, dim) ; 
+        rv = rv ./ norm(rv) ;
+		rh = rv * h ; 
+		rDh = rv * Dh ;
+
+		[~, tfrrs, ~] = rsSTFTbase_RT(x, lowFreq, highFreq, alpha, hop, LL, rh', rDh', 1);
+
+	 	ConceFT = ConceFT + tfrrs ;
+    end
+
+    ConceFT = ConceFT ./ MT ;
+
+end
+
+end

### 2020-07-25 TimeFrequencyScaleRep/ConceFT_sqSTFT_RT.m

@@ -1,0 +1,59 @@
+function [tfr, tfrsq, ConceFT, tfrtic] = ConceFT_sqSTFT_RT(x, lowFreq, highFreq, alpha, hop, WinLen, LL, dim, supp, MT, Second, Smooth)
+% CONCEFT_SQSTFT_RT ConceFT Synchrosqueezing of the STFT
+% Usage: 
+% 	[tfr, tfrsq, ConceFT, tfrtic] = ConceFT_sqSTFT_RT(x, lowFreq, highFreq, alpha, hop, WinLen, LL, dim, supp, MT, Second, Smooth)
+%
+% Input:
+%   x: signal to be analized
+%   lowFreq: relative lower frequency (0<lowFreq<Highfreq)
+%   highFreq: relative higher frequency (lowFreq<Highfreq<0.5)
+%   alpha: resolution in the frequency axis
+%   hop: hop size (in samples)
+%   WinLen: Window length (in samples)
+%   LL: extension length (in samples)
+%   dim: order of the Hermite function used as window
+%   supp: half time support of the Hermite function used as window
+%   MT = 1: ordinary RS; MT > 1: ConceFT
+%   Second: if 1, compute the second-order synchrosqueezing
+%   Smooth: if 1, compute the smoothed version of the synchrosqueezing
+% 
+% Output:
+%   tfr: STFT
+%   tfrsq: Synchrosqueezing of STFT
+%   ConceFT: ConceFT synchrosqueezing of STFT
+%   tfrtic: frequencies for which TFRs are evaluated
+
+%% Ordinary SST
+[h, Dh, ~] = hermf(WinLen, dim, supp) ; % generate the window for short time Fourier transform (STFT)
+
+if ~Second
+    [tfr, tfrsq, tfrtic] = sqSTFTbase_RT(x, lowFreq, highFreq, alpha, hop, LL, h(1,:)', Dh(1,:)', Smooth);
+else
+    [tfr, ~, tfrsq, tfrtic] = sqSTFTbase2nd_RT(x, lowFreq, highFreq, alpha, hop, LL, h(1,:)', Dh(1,:)', dwindow(Dh(1,:)'));
+end
+
+%% Multitapering
+
+ConceFT = abs(tfrsq) ;
+
+if MT > 1
+
+	% Conceft
+    for ii = 1: MT
[…]

### 2020-07-25 TimeFrequencyScaleRep/STFT_RT.m

@@ -1,0 +1,23 @@
+function [tfr, tfrtic] = STFT_RT(x, lowFreq, highFreq, alpha, hop, WinLen, LL, dim, supp)
+% STFT_RT  STFT using Hermite window
+% Usage: 
+% 	[tfr, tfrtic] = STFT_RT(x, lowFreq, highFreq, alpha, hop, LL, WinLen, dim, supp)
+%
+% Input:
+%   x: signal to be analized
+%   lowFreq: relative lower frequency (0<lowFreq<Highfreq)
+%   highFreq: relative higher frequency (lowFreq<Highfreq<0.5)
+%   alpha: resolution in the frequency axis
+%   hop: hop size (in samples)
+%   LL: extension length
+%   WinLen: Window length (in samples)
+%   dim: order of the Hermite function used as window
+%   supp: half time support of the Hermite function used as window
+% 
+% Output:
+%   tfr: STFT
+%   tfrtic: frequencies for which STFT is evaluated
+
+[h, ~, ~] = hermf(WinLen, dim, supp) ; % generate the window for short time Fourier transform (STFT)
+
+[tfr, tfrtic] = STFTbase_RT(x, lowFreq, highFreq, alpha, hop, LL, h(1,:)');

### 2020-07-25 TimeFrequencyScaleRep/rsSTFTbase_RT.m

@@ -1,7 +1,7 @@
-function [tfr, tfrrs, tfrtic] = rsSTFTbase(x, lowFreq, highFreq, alpha, tDS, h, Dh, Causality)
-% RSSTFTBASE Reassigment of the STFT
+function [tfr, tfrrs, tfrtic] = rsSTFTbase_RT(x, lowFreq, highFreq, alpha, tDS, LL, h, Dh, Causality)
+% RSSTFTBASE_RT Reassigment of the STFT
 % Usage: 
-% 	[tfr, tfrrs, tfrtic] = rsSTFTbase(x, lowFreq, highFreq, alpha, tDS, h, Dh, Causality)
+% 	[tfr, tfrrs, tfrtic] = rsSTFTbase_RT(x, lowFreq, highFreq, alpha, tDS, LL, h, Dh, Causality)
 %
 % Input:
 %	x     : analysed signal.
@@ -56,15 +56,15 @@ Dt = 1 ;
 
 
 %% Run STFT and reassignment rule
-tfr = zeros(fLen, tLen) ;
-tfrrs = zeros(fLen, tLen); 
+tfr = zeros(fLen, LL+1) ;
+tfrrs = zeros(fLen, LL+1); 
 tfrtic = linspace(lowFreq, highFreq, fLen)' ;
 
 Ex = mean(abs(x(min(t):max(t))).^2);
 Threshold = 1.0e-8*Ex;  % originally it was 1e-6*Ex
 
-
-for tidx = 1:tLen
+j = 1 ;
+for tidx = (tLen-2*LL):(tLen-LL)
 
     ti = t((tidx-1)*tDS+1); 
     tau = -min([round(N/2)-1,Lh,ti-1]):min([round(N/2)-1,Lh,xrow-ti]);
@@ -91,11 +91,11 @@ for tidx = 1:tLen
         if abs(tfr(jcol)) > Threshold
 
 			% to keep the causality, use max(tidxhat,tidx-Lh) instead of max(tidxhat,1)
-            tidxhat = tidx + omegaT(jcol);
+            tidxhat = j + omegaT(jcol);
 			if Causality
-            	tidxhat = min(max(max(tidxhat, tidx-Lh), 1), tLen);
+            	tidxhat = min(max(max(tidxhat, j-Lh), 1), LL);
 			else
-            	tidxhat = min(max(tidxhat, 1), tLen);
[…]

### 2020-07-28 Notes/BoundEffRed.tex
new results + paper

@@
 \includegraphics[width=.48\textwidth]{PPGsig.eps}
-\includegraphics[width=.24\textwidth]{zoomSSTintro.eps}
-\includegraphics[width=.24\textwidth]{zoomSSTBoundEffRed.eps}
+%\includegraphics[width=.24\textwidth]{zoomSSTintro.eps}
+%\includegraphics[width=.24\textwidth]{zoomSSTBoundEffRed.eps}
 \caption{A segment of PPG signal (top) and the right boundary of a TF representation determined by the SST without extension (bottom left), and the right boundary of a TF representation determined by the SST with the proposed boundary effect reduction algorithm by forecasting (bottom right).
@@
 Attempts to minimize the boundary effects generally consists in softening the discontinuity on signal edges.
-For instance, judicious choices of analysis windows whose support does not interfere with the boundary points can minimize the occurrence of aberrant patterns near the boundaries of the TF plane~\cite{Chui92wavelets,Depczynski99fast}.
-Due to the specific relationship of chosen analysis windows and TF analysis tool, these techniques do not make it possible to reduce the boundary effects of any TF representations, and for any analysis window.
-Another natural idea consists of carrying out a preliminary step of extending the signal beyond its boundaries, and due to its flexibility, various extension schemes have been proposed.
+Usually the approaches can be classified into two main classes, choosing a proper analysis window and extending/forecasting the signal.
+In the first class, judicious choice of analysis window whose support does not interfere with the boundary points can minimize the occurrence of aberrant patterns near the boundaries of the TF plane, for instance, \cite{Chui92wavelets,Depczynski99fast}.
+Due to the specific relationship of chosen analysis windows and TF analysis tool, these techniques do not make it possible to reduce the boundary effects of any TF representations.
+%
+Another natural idea, the second class, consists of carrying out a preliminary step of extending, or forecasting, the signal beyond its boundaries.
+Due to its flexibility, various forecasting schemes have been proposed.
 For example, there exist simple extension schemes that do not take into account the dynamical behavior of the signal, such as zero-padding, periodic extension, symmetric extension~\cite{Kharitonenko02wavelet,Chen95symmetric}, or polynomial extrapolation~\cite{Williams97discrete}.
-There also exist extension schemes based on physically relevant dynamical models, such as the Extended Dynamic Mode Decomposition~\cite{Williams15data} (EDMD), the Gaussian process regression~\cite{Rasmussen06gaussian,Roberts13Gaussian} (GPR), the Trigonometric, Box-Cox transformation, ARMA errors, Trend and Seasonal components (TBATS) algorithm~\cite{DeLivera11forecasting}, the dynamic linear models \cite{west2006bayesian}, and long short-term memory model \cite{vlachas2018data}.
-It is far from an exhaustive list, and we refer readers with interest to \cite{hyndman2018forecasting} for a friendly monograph on the general forecasting topic.
-Whereas the second class of extension schemes gives better-extended signals than the first class, they generally have a great computational cost.
+%
+There exist extension schemes based on physically relevant dynamical models, such as the Extended Dynamic Mode Decomposition~\cite{Williams15data} (EDMD) and the Gaussian process regression~\cite{Rasmussen06gaussian,Roberts13Gaussian} (GPR).
+There also exist extension schemes based on stochastic models, such as the Trigonometric, Box-Cox transformation, ARMA errors, Trend and Seasonal components (TBATS) algorithm~\cite{DeLivera11forecasting}, the dynamic linear models \cite{west2006bayesian}.
+%
+In the physically relevant dynamical models, the oscillation or trend, are usually modeled as the mean of a stochastic process, while in the stochastic models, the oscillation or trend are modeled by the covariance structure of the stochastic process.
+It is also possible to consider polynomial regression \cite{fan1996local} or modeling the mean by splines \cite{hall2005theory}, kernel functions \cite{chang2010training}, and wavelets \cite{marron1998exact}, or nonparametric regression \cite{fan1996local}, to estimate the mean function and hence forecasting the signal.
+Neural network models, such as the long short-term memory model \cite{vlachas2018data}, is another approach.
+%
+The above is a far from exhaustive list, and we refer readers with interest to \cite{hyndman2018forecasting} for a friendly monograph on the general forecasting topic. 
+%
+While the model-based extension scheme gives better-extended signals than those simple extension without dynamical models, they generally have a great computational cost.
 
-In this paper, we propose a fast extension algorithm based on a simple dynamical model, in order to optimize the trade-off between the extension quality and the computational cost.
-The proposed algorithm is then composed of two steps.
+In this paper, we propose a fast extension algorithm based on a simple dynamical model that optimizes the trade-off between the extension quality and the computational cost, so that the real-time analysis can be achieved.
+The algorithm is composed of two steps.
 \begin{enumerate}
-\item \emph{Extend the signal by forecasting it.} The aim is to use a dynamic model to predict the values taken by the measured signal outside the measurement interval.
+\item \emph{Extend the signal by forecasting it.} The aim is to use a simple dynamic model to predict the values taken by the measured signal outside the measurement interval.
 Then, once this operation is done, we have access to an extended signal defined on a larger interval $I_\Delta$, where $\Delta$ denotes the size of the extension on both boundaries of $I$.
@@
 \end{enumerate}
+
 Thus, assuming that the quality of the extension step is sufficient, the analysis results obtained that way will be less sensitive to the boundary effects than the result of the analysis tool applied directly to the nonextended signal.

### 2020-07-28 Notes/Supplementary material/SupMat.tex
new results + paper

@@
 
-\title{An Efficient Forecasting Approach for the Real-Time Reduction of Boundary Effects\\ in Time-Frequency Representations:\\ Part II}
+\title{Supplementary material for ``An Efficient Forecasting Approach for the Real-Time Reduction of Boundary Effects in Time-Frequency Representations''}
 
@@
 
-\subsection{Notations}
-By definition the matrix $\bX$ and $\bY$ and because of the noisy signal model~\eqref{I-eq:model.noise}, we have:
+%\subsection{Notations}
+Recall the model~\eqref{I-eq:model.noise}. Based on the definition of matrices $\bX$ and $\bY$, we have:
 \begin{align}
@@
 \end{align}
-where $\bE^{(a)} = \sigma\bE_1^{(a)} + \sigma^2\bE_2^{(a)}$, with:
+where $\bE^{(a)} := \sigma\bE_1^{(a)} + \sigma^2\bE_2^{(a)}$,
 \[
@@
 with $a\in\{0,1\}$.
-
-\begin{remark}
-The matrices $\bE^{(0)}$ and $\bE^{(1)}$ are said to be error matrices because:
+We call $\bE^{(0)}$ and $\bE^{(1)}$ {\em error matrices} because:
 \begin{align*}
@@
 \end{align*}
-\end{remark}
-
-It follows from this:
+Thus, 
 \begin{align*}
-\tilde\bA &= (\bS^{(1)}+\bE^{(1)})(\bS^{(0)}+\bE^{(0)})^\inv \\
-\bA_0 &=\bS^{(1)}{\bS^{(0)}}^\inv\ .
+\bA_0 :=\bS^{(1)}{\bS^{(0)}}^\inv\,,\quad\tilde\bA := (\bS^{(1)}+\bE^{(1)})(\bS^{(0)}+\bE^{(0)})^\inv\,.
 \end{align*}
-Then:
+As a result, for $\ell\in \mathbb{N}$, 
 \begin{align}
@@
 
-\subsection{Study of $\bh^{(\ell)}$}
-The randomness of $\bh^{(\ell)}$ is completely originating from the error matrices.
-Besides, notice that the first $M-1$ rows in $\bE^{(1)}$ are equals to the last $M-1$ rows of $\bE^{(0)}$.
-Thus, we gather all the sources of randomness into an vector $\bg\in\RR^{M(M+1)}$, containing $M$ rows and defined as:
+%\subsection{Study of $\bh^{(\ell)}$}
+The randomness of $\bh^{(\ell)}$ completely comes from the error matrices.
+Besides, notice that the first $M-1$ rows in $\bE^{(1)}$ equal to the last $M-1$ rows of $\bE^{(0)}$.
+We gather all sources of randomness into an vector $\bg\in\RR^{M(M+1)}$, containing $M$ rows defined as
 \[
 \bg = \vec\left(
-\begin{pmatrix}
+\begin{bmatrix}
 \bE^{(0)} \\
 \be_M^T\bE^{(1)}
-\end{pmatrix}
-\right)\ .
+\end{bmatrix}
+\right)\,,
 \]
-Here, "$\vec$" denotes the vectorization operator, concatenating the columns of a given matrix on top of one another. Then, by definition, we have:
+where "$\vec$" denotes the vectorization operator, concatenating the columns of a given matrix on top of one another. Then, by definition, we have:
 \[
@@
 \bg_1 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left( \tilde\bz_k\bw_k^T + \tilde\bw_k\bz_k^T \right) \\
-\bg_2 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left(\tilde\bw_k\bw_k^T-\tilde\bI \right) \ .
+\bg_2 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left(\tilde\bw_k\bw_k^T-\tilde\bI \right) \,,
 \end{align*}
-with $\tilde\bz_k^T= \begin{pmatrix} \bz_k^T & \bz_{k+1}[M-1] \end{pmatrix}$ and $\tilde\bw_k^T= \begin{pmatrix} \bw_k^T & \bw_{k+1}[M-1] \end{pmatrix}$.
+where $\tilde\bz_k^T= \begin{pmatrix} \bz_k^T & \bz_{k+1}[M-1] \end{pmatrix}$ and $\tilde\bw_k^T= \begin{pmatrix} \bw_k^T & \bw_{k+1}[M-1] \end{pmatrix}$.
 Then, $\bg_1$ is a Gaussian random vector because it is a linear combination of Gaussian random vectors.
 Moreover, using the central limit theorem under weak dependence, we can show that $\bg_2$ also converges towards a Gaussian random vector as $K\to\infty$.
-Combining these two results gives the following result:
+Combining these two results leads to
 \[
@@
 \end{align*}
-Then, as $f^{(\ell)}$ is a differentiable function, using the Delta method gives:
+Then, as $f^{(\ell)}$ is a differentiable function with non-zero differentiation at $0$ {\color{red}[I assume that you have checked all conditions for the delta method.
+You may consider adding its formula to save the readers' time]}, using the Delta method gives:
 \begin{align*}
@@
 
-\subsection{Expression of the Bias $\bmu$.}
-Clearly, $\bmu[n]=0$ when $n\in I$. When $n=N-1+\ell$, we have:
+\subsection{Expression of the bias $\bmu$.}
+Clearly, $\bmu[n]=0$ when $n\in I$. When $n=N-1+\ell$, we have
 \begin{align*}
@@
 \end{align*}
-Let us first evaluate the expression of $\balpha_0^{(\ell)}\bz_K$. We have:
+Let us first evaluate the expression of $\balpha_0^{(\ell)}\bz_K$. We have
 \begin{align*}
@@
 \]
-where $\bU[m,m']=\frac1{\sqrt{M}}e^{-2\ii\pi mm'/M}$ and $\bLambda^{(0)} = \mathrm{diag}(\lambda_0^{(0)},\dots,\lambda_{M-1}^{(0)})$ with:
+where $\bU[m,m']=\frac1{\sqrt{M}}e^{-2\ii\pi mm'/M}$ and $\bLambda^{(0)} = \mathrm{diag}(\lambda_0^{(0)},\dots,\lambda_{M-1}^{(0)})$ with
 \begin{align*}
@@
 \end{align*}
-Therefore:
+Therefore,
 \begin{align*}
-{\bS^{(0)}}^\inv  &= \bU{\bLambda^{(0)}}^\inv\bU^*
+{\bS^{(0)}}^\inv  &= \bU{\bLambda^{(0)}}^\inv\bU^*\,,
 \end{align*}
-which leads to:
+which leads to
 \begin{align*}
-{\bS^{(0)}}^\inv[m,m']  &= \dfrac1{\sigma^2}\delta_{m,m'}-\sum_{j=1}^J\dfrac{\Omega_j^2}{2\sigma^2(\sigma^2+\Omega_j^2M/4)}\cos\left(2\pi p_j \dfrac{m-m'}{M}\right)\ ,
+{\bS^{(0)}}^\inv[m,m']  &= \dfrac1{\sigma^2}\delta_{m,m'}-\sum_{j=1}^J\dfrac{\Omega_j^2}{2\sigma^2(\sigma^2+\Omega_j^2M/4)}\cos\left(2\pi p_j \dfrac{m-m'}{M}\right)\,.
 \end{align*}
-and, consequently:
+Consequently, we have
 \begin{align}
@@
 \tilde\bA_0[m,m']  &= \sum_{q=0}^{M-1} {\bS^{(1)}}[m,q]{\bS^{(0)}}^\inv[q,m'] \\
-&= \delta_{m+1,m'} + \sum_{j=1}^J\dfrac{2\Omega_j^2}{\Omega_j^2M+4\sigma^2}\cos\left(2\pi p_j\frac{m'}{M}\right)\delta_{m+1,M}
+&= \delta_{m+1,m'} + \sum_{j=1}^J\dfrac{2\Omega_j^2}{\Omega_j^2M+4\sigma^2}\cos\left(2\pi p_j\frac{m'}{M}\right)\delta_{m+1,M}\,.
 \label{eq:A0.sine}
 \end{align}
-Thus:
+Thus,
 \begin{align*}
@@
 \balpha_0^{(1)}\bz_K
-\end{pmatrix}
+\end{pmatrix}\,.
 \end{align*}
-By induction, we have:
+By induction, we have
 \begin{align*}
@@
 \end{pmatrix}
-\ .
+\, .
 \end{align*}
-Then:
+Then,
 \begin{align}
@@
 \balpha_0^{(\ell)}\bz_{K} &= \tilde\balpha_0^{(1)}\tilde\bA_0^{\ell-1}\bz_{K} \\
-&=\sum_{m=0}^{M-\ell}\balpha_0^{(1)}[m]\bz[N-M+\ell+m-1]+\sum_{m=M-\ell+1}^{M-1}\balpha_0^{(1)}[m]\balpha_0^{(m-M+\ell)}\bz_K
+&=\sum_{m=0}^{M-\ell}\balpha_0^{(1)}[m]\bz[N-M+\ell+m-1]+\sum_{m=M-\ell+1}^{M-1}\balpha_0^{(1)}[m]\balpha_0^{(m-M+\ell)}\bz_K\,.
 \label{eq:seq}
 \end{align}
-But:
+But,
 \begin{align*}
@@
 \end{align*}
-and, by induction from~\eqref{eq:seq}:
+By induction from~\eqref{eq:seq}, we have
 \begin{equation}
@@
 \end{equation}
-Then:
+Then,
 \begin{equation*}
@@
 \end{align*}
-Consequently:
+{\color{red}[if it is possible to get the convergence rate, like $\EE\{\bh^{(\ell)}\}=O(1/K)$, that would be great!]} Consequently,
 \begin{equation}
-\bmu[N-1+\ell] \underset{K\to\infty}{\sim} o(\sigma)\ .
+\frac{1}{\sigma}\bmu[N-1+\ell] = o(1)
 \label{eq:bias.ap}
 \end{equation}
+when $K\to\infty$.
 
-\subsection{Expression of the Covariance $\bgamma$.}
-We recall that $n>N$, we consequently denote $n=N-1+\ell$.  To derive the covariance, let us segregate the cases.
+\subsection{Expression of the covariance $\bgamma$.}
+Take $n>N$, and denote $n=N-1+\ell$.  To derive the covariance, let us segregate the cases.
 
-\paragraph{If $n'\in I^2$}
-From Lemma~1, we have that $h^{(\ell)}$ is asymptotically Gaussian.
-Then, as a direct consequence of the Isserlis' theorem, odd-order moments are vanishing.
+\paragraph{When $n'\in I$}
+From Lemma~1, we have that $h^{(\ell)}$ is asymptotically Gaussian when $K\to \infty$.
+Then, as a direct consequence of the Isserlis' theorem, odd-order moments are vanishing {\color{red}[there is a gap here.
+Note that when $K$ is finite, it is approximated by Gaussian but not Gaussian.
+The discrepancy should be made clear.]}.
 Then, combining result~\eqref{eq:bias.ap} and equation~\eqref{I-eq:bias0}, we obtain:
 \begin{align*}
-\bgamma[n,n'] &= \sigma\EE\{\bw[n']\bh^{(\ell)}\}\bz_K + \sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bw[n']\} + o(\sigma^2) \ .
+\bgamma[n,n'] &= \sigma\EE\{\bw[n']\bh^{(\ell)}\}\bz_K + \sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bw[n']\} + o(\sigma^2) 
 \end{align*}
+when $K\to \infty$.
 We remark that
@@
 \]
-Moreover, using Delta method one can show that we have the asymptotic result:
+Moreover, using Delta method one can show that we have the asymptotic result
 \begin{equation}
-\EE\{\bw[n']\bh^{(\ell)}\} \underset{K\to\infty}{\longrightarrow} 0
+\EE\{\bw[n']\bh^{(\ell)}\} \underset{K\to\infty}{\longrightarrow} 0\,.
 \label{eq:cov.ap}
 \end{equation}
-Finally:
+As a result, we have
 \begin{align*}
-\bgamma[n,n'] &\underset{K\to\infty}{\sim} \sigma^2\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)} + o(\sigma^2) \ .
+\bgamma[n,n'] &= \sigma^2\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)}+ o(\sigma^2)
 \end{align*}
+when $K\to\infty$.
+{\color{red}[Note that $o(\sigma^2)$ becomes $0$ when $K\to \infty$.
+So use either ``$\bgamma[n,n']= \sigma^2\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)}+ o(\sigma^2)$ when $K\to \infty$'' or ``$\bgamma[n,n'] \to \sigma^2\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)}$ when $K\to \infty$''.]}
 
-\paragraph{If $n'\geq N$}
+\paragraph{When $n'\geq N$}
 From Lemma~1, we have that $h^{(\ell)}$ is asymptotically Gaussian.
 Then, as a direct consequence of the Isserlis' theorem, odd-order moments are vanishing.
-Then, combining result~\eqref{eq:bias.ap} and equations~\eqref{I-eq:cov01} and~\eqref{eq:cov.ap}, we obtain:
+{\color{red}[The same gap here.]} Then, combining equations~\eqref{I-eq:cov01}, \eqref{eq:bias.ap} and~\eqref{eq:cov.ap}, we obtain
 \begin{align*}  
-\bgamma[n,n'] &= \bz_K^T\EE\left\{{\balpha^{(\ell)}}^T\balpha^{(\lambda)}\right\}\bz_K + \sigma\EE\{\balpha^{(\ell)}\bw_K\balpha^{(\lambda)}\}\bz_K+ \sigma\EE\{\balpha^{(\lambda)}\bw_K\balpha^{(\ell)}\}\bz_K \\
+\bgamma[n,n'] = &\,\bz_K^T\EE\left\{{\balpha^{(\ell)}}^T\balpha^{(\lambda)}\right\}\bz_K + \sigma\EE\{\balpha^{(\ell)}\bw_K\balpha^{(\lambda)}\}\bz_K+ \sigma\EE\{\balpha^{(\lambda)}\bw_K\balpha^{(\ell)}\}\bz_K \\
 & +\sigma^2\EE\{\balpha^{(\ell)}\bw_K\balpha^{(\lambda)}\bw_K\}-\bz[n]\bz[n']+o(\sigma^2) \\
-&= \bz_K^T\EE\left\{{\bh^{(\ell)}}^T\bh^{(\lambda)}\right\}\bz_K + \balpha_0^{(\ell)}\sigma\EE\{\bw_K\bh^{(\lambda)}\}\bz_K + \sigma\EE\{\bh^{(\ell)}\bw_K\}\bz[n']  \\
+= &\,\bz_K^T\EE\left\{{\bh^{(\ell)}}^T\bh^{(\lambda)}\right\}\bz_K + \balpha_0^{(\ell)}\sigma\EE\{\bw_K\bh^{(\lambda)}\}\bz_K + \sigma\EE\{\bh^{(\ell)}\bw_K\}\bz[n']  \\
 &+ \sigma\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\}\bz_K+ \sigma\EE\{\bh^{(\lambda)}\bw_K\}\bz[n] + \sigma\balpha_0^{(\lambda)}\EE\{\bw_K\bh^{(\ell)}\}\bz_K \\
@@
 &+\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\sigma^2\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} + o(\sigma^2) \\
-&\underset{K\to\infty}{\sim} \dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K +\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\sigma^2\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} + o(\sigma^2)
+= &\,{\color{blue}\dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K} +\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\sigma^2\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} + o(\sigma^2)
 \end{align*}
-The Isserlis' theorem also gives the following result:
+when $K\to\infty$ {\color{red}[in the last equation, make clear which term depends on which equation \eqref{I-eq:cov01}, \eqref{eq:bias.ap} and~\eqref{eq:cov.ap}.] [The previous or current asymptotic expressions with $\dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K$ are both not ideal.
+Either using your previous notation or following what I prefer here, the behavior of $\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K$ should be made clear.
+Is it bounded?
+If it is bounded, then use $=O(1/K)$ to replace $\to \dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K$.
+If not, how does it depend on $K$ should be made clear.] [Moreover, if you prefer to use your previous notation, you may clarify what it means by $\sim$.]}.
+The Isserlis' theorem also gives the following result {\color{red}[the same gap here.]}
 \begin{align*}
@@
 &\hspace{35pt} + \EE\{\bh^{(\ell)}[m]\bh^{(\lambda)}[m']\}\underbrace{\EE\{\bw_K[m]\bw_K[m']\}}_{=\delta_{m,m'}} \\
-&\underset{K\to\infty}{\sim} \sum_{m=0}^{M-1} \EE\{\bh^{(\ell)}[m]\bh^{(\lambda)}[m]\} = \dfrac1K\Tr\left(\bGamma^{(\ell,\lambda)}\right)\ .
+&\to \sum_{m=0}^{M-1} \EE\{\bh^{(\ell)}[m]\bh^{(\lambda)}[m]\} = \dfrac1K\Tr\left(\bGamma^{(\ell,\lambda)}\right)
 \end{align*}
-We conclude that:
+when $K\to \infty$.
+We thus conclude that
 \begin{align*}  
-\bgamma[n,n'] &\underset{K\to\infty}{\sim} \dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K +\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\dfrac{\sigma^2}{K}\Tr\left(\bGamma^{(\ell,\lambda)}\right) + o(\sigma^2)\ .
+\bgamma[n,n'] &\to {\color{blue}\dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K} +\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\dfrac{\sigma^2}{K}\Tr\left(\bGamma^{(\ell,\lambda)}\right) + o(\sigma^2)
 \end{align*}
+when $K\to \infty$.
 

### 2020-07-28 Notes/algorithm.tex
real time improved
new results + paper

@@
 where $\bA\in\RR^{M\times M}$.
-In other words, we adopt a classical strategy in the study of dynamical systems, the {\em linearization} of a nonlinear phenomenon, when the system is sufficiently regular.
+In other words, we adopt a classical strategy in the study of dynamical systems, the {\em linearization of a nonlinear system}, when the system is sufficiently regular.
 Notice that this linearized dynamical model can be written equivalently in function of the subsignals $\bx_k$, as:
@@
 We summarize the extension step in Algorithm~\ref{alg:extension}.
-Notice that we could, if necessary, handle the backward estimation using the same strategy as described above, but applying it to the reverse signal $\bx^{\rm r} = \begin{pmatrix} \bx[N-1] & \cdots & \bx[0] \end{pmatrix}^T$.
+%We mention that we could, if necessary, handle the backward estimation using the same strategy as described above, but applying it to the reverse signal $\bx^{\rm r} = \begin{pmatrix} \bx[N-1] & \cdots & \bx[0] \end{pmatrix}^T$.
 
@@
 \subsection{Step 2: Extended Time-Frequency Representation}
+\label{sse:extension.TFR}
 
@@
 Here, $F$ typically denotes the size of the discretization along the frequency axis.
-Due to the boundary effects, the representation $\ccF(\bx_N)$ shows undesired patterns when approaching its edges.
+Due to the boundary effects, the representation $\ccF(\bx_N)$ shows undesired patterns near its edges.
 To alleviate the boundary effects, we apply the representation to the estimated extended signal $\tilde\bx$.
@@
 This amounts to restricting the representation $\ccF(\tilde \bx_{N+L})$ to the current measurement interval of $\bx_N$.
-For the sake of simplicity, we denote the restriction operator by $\cR$, where $\cR:\CC^{F\times (N+L)}\to\CC^{F\times N}$.
+For the sake of simplicity, we denote the {\em restriction operator} by $\cR$, where $\cR:\CC^{F\times (N+L)}\to\CC^{F\times N}$.
 Consequently, we have:
@@
 Thus, $\bF_\bx^{(N+1)}$ is obtained by the concatenation of the first $N-L_{\mathrm{win}}+ 1$ columns of $\bF_\bx^{(N)}$ with the last $L_{\mathrm{win}}$ columns of $\ccF^{\mathrm{ext}}(\bx_{N+1})$.
-
-\begin{remark}
-Unlike other representations, RS is not causal and rigorously requires knowing the STFT on the whole time-frequency plane before making any reassignment.
-Nevertheless, in~\cite{Lin17conceft}, the authors propose a causal RS that only requires the future values of the STFT within a limited time horizon (typically less than $L$).
-We use this version in the paper to enable real-time implementation of RS.
-\end{remark}

### 2020-07-28 Notes/results.tex
new results + paper

@@
 
-\subsection{Evaluation of the Forecasting Performance}
-In that section, we first evaluate the quality of the forecasting step and compare it to the theoretical results provided by Theorem~\ref{th:error}.
+\subsection{Extension methods for comparison}
+To highlight the fact that the linear dynamical model is sufficient to catch most of the dynamical behavior of signals following the AHM, we compare the performance of Algorithm~\ref{alg:boundary} with a simple extension obtained by pointwise symmetrization~\cite{Kharitonenko02wavelet}.
+We also evaluate the performance of reference forecasting algorithms that could be used for extending such signals.
+These methods are:
+\begin{itemize}
+\item The EDMD has been developed by Williams \etal~\cite{Williams15data}.
+The proposed algorithm is a way to obtain an approximation of the so-called Koopman operator of the observed system, which theoretically allows catching dynamic of nonlinear systems~\cite{Korda18linear}.
+\item The GPR~\cite{Rasmussen06gaussian} is a method relying on a probabilistic dynamical model.
+That one is based on the Gaussian process structure, and therefore offers more flexibility in the type of dynamic that could be modeled than the linear model~\eqref{eq:dyn.model}.
+\item The TBATS method~\cite{DeLivera11forecasting} is based on a classical decomposition of times series into a trend, seasonal and ARMA components, with a specific dynamic for the seasonal component.
+This model demands the estimation of numerous parameters and, by implication, may be slow.
+\end{itemize}
+
+\subsection{Evaluation metrics}
+
+The first evaluation metric is quantifying the forecasting quality (\ie~not depending on $\ell$) in the time domain.
+We evaluate the Experimental Mean Square Error $\mathrm{MSE_{xp}}(\tilde\bx)$ of the forward forecast extended signals, namely,
+\begin{align}
+\label{eq:mse}
+\mathrm{MSE_{\xp}}(\tilde\bx) &= \dfrac1{L}\|\tilde\bx -\bx^\mathrm{ext}\|^2 \\[-1mm]
+\nonumber
+&=\! \dfrac1{L}\sum_{\ell=1}^L \bmu_{\xp}[N\!-\!1\!+\!\ell]^2 \!+\! \bgamma_{\xp}[N\!-\!1\!+\!\ell,N\!-\!1\!+\!\ell] .
+\end{align}
+where $\bx^\mathrm{ext}$ is the {\em ground-truth extended} signal; that is, $\bx^\mathrm{ext} := \begin{pmatrix}\bx[-L] & \cdots & \bx[N-1+L] \end{pmatrix}$.
+Then, as long as the bias $\bmu[N-1+\ell]$ and the variance $\bgamma[N-1+\ell,N-1+\ell]$ of the forecasting estimator remain small for all $\ell$, the MSE takes small values either.
+
+The second evaluation metric is evaluating the quality of the boundary effect reduction directly on the TF representation.
+We compare the obtained TF representation to the optimal TF representation $\ccF_N^\mathrm{opt}(\bx)$, defined as the restriction of the representation of the ground-truth extended signal $\bx^\mathrm{ext}$; that is, we have
+\begin{equation}
+\ccF^\mathrm{opt}(\bx_N) = \cR\left( \ccF(\bx^\mathrm{ext}) \right) \, ,
+\label{eq:opt.TFR}
+\end{equation} 
+where $\mathcal{R}$ is defined in \eqref{eq:bound.free.TFR}.
+To compare different techniques, we use a criterion proposed in~\cite{Daubechies16conceft} that quantifies the optimal transport (OT) distance between a given TF representation and the optimal one.
+In general, the OT distance is a quantity that measures how different two probability density functions are.
+Denote a TF representation as $\ccQ$.
+For $t$ fixed, we consider the probability density function defined by $p_\ccQ^t(\xi) = \left.
+|\ccQ(\xi,t)|^2 \middle/ \int_\RR |\ccQ(\nu,t)|^2\dd\nu \right.$; that is, normalizing the spectral information at time $t$.
+At each time $t$, the OT distance $d_{t}$ between two TF representations $\ccQ$ and $\ccF^\mathrm{opt}$ at time $t$ is defined by
+\begin{equation*}
+d_{t}(\ccQ,\ccF^\mathrm{opt}) = \int_\RR\left|\tilde P_{\ccQ}^t(\xi)-  P_{\ccF^\mathrm{opt}}^t(\xi)\right|\dd\xi\ ,
+\end{equation*}
+where $P_{\ccQ}^t(\xi)=\int_{-\infty}^\xi p_{\ccQ}^t(\nu)\dd\nu$ and $\tilde P_{\ccF^\mathrm{opt}}^t(\xi)=\int_{-\infty}^\xi\tilde p_{\ccF^\mathrm{opt}}^t(\nu)\dd\nu$.
+In other words, the OT distance between $\ccQ$ and $\ccF^\mathrm{opt}$ at time $t$ is given by the $L^1$ norm of the associated cumulative density function.
+%
+Note that $d_{t}(\ccQ,\ccF^\mathrm{opt})$ quantifies the proximity between the estimated and actual TF representations at time $t$.
+With the OT distance, we define the {\em performance index} $D(\ccQ,\ccF)$ for the reduction of boundary effects of a given TF representation $\ccQ$ over another TF representation $\ccF$ by
+%\begin{equation}
+%D(\ccQ) = \dfrac1{|I|}\bigintssss_I\ \frac{d_{t}\left(\ccQ,\ccF^\mathrm{opt}\right)}{d_{t}\left(\ccF,\ccF^\mathrm{opt}\right)}\ \dd t\ .
+%\label{eq:index.perf}
+%\end{equation}
+\begin{equation}
+D(\ccQ,\ccF) = \dfrac{\displaystyle\int_I d_{t}\left(\ccQ,\ccF^\mathrm{opt}\right)\dd t}{\displaystyle\int_I d_{t}\left(\ccF,\ccF^\mathrm{opt}\right)\dd t}\ .
+\label{eq:index.perf}
+\end{equation}
+In other words, $D(\ccQ,\ccF)$ is the ratio between its averaged OTD to the optimal TF representation $\ccF^\mathrm{opt}$ (see~\eqref{eq:opt.TFR}) and the averaged OTD of the original TF representation $\ccF$ to $\ccF^\mathrm{opt}$.
+%
+Thus, $D(\ccQ)<1$ means a reduction of the boundary effects. Let us evaluate the quality of the boundary effects reduction on biomedical signals.
+
+\subsection{Simulated signals}
+We first evaluate the quality of the forecasting step and compare it to the theoretical results provided by Theorem~\ref{th:error}.
 The level of the forecasting error depends on at least two parameters:
@@
 \[
-\bz[n]\! =\!\cos\!\left(2\pi p_1 \dfrac{n}{M} \right)\! +\! R\cos\!\left(2\pi p_2 \dfrac{n}{M} \right),\quad \!\forall n\!\in\!\{1,\ldots,N\},
+\bz[n]\! =\!\cos\!\left(2\pi p_1 \dfrac{n}{M} \right)\! +\! A\cos\!\left(2\pi p_2 \dfrac{n}{M} \right),\quad \!\!\forall n\!\in\!\{1,\ldots,N\},
 \]
-with $M=150$, $p_1=10$, $p_2=33$ and $R=1.4$. Besides, the additive noise is chosen to be Gaussian: $\bw\sim\cN(\bzero,\bI)$.
+with $M=150$, $p_1=10$, $p_2=33$ and $A=1.4$. Besides, the additive noise is chosen to be Gaussian: $\bw\sim\cN(\bzero,\bI)$.
 
@@
 \label{ssse:res.ahm}
-We now consider a signal satisfying the AHM so that the instantaneous frequencies and amplitudes of its components vary from time to time.
+Consider a signal satisfying the AHM so that the instantaneous frequencies and amplitudes of its components vary over time.
 The deterministic component $\bz$ of the random vector $\bx$, constructed following the model~\eqref{eq:model.noise}, takes the following form, for all $n\in\{1,\ldots,N\}$:
@@
 \end{align*}
-Besides, the noise is chosen to be Gaussian: $\bw\sim\cN(\bzero,\bI)$, and we take: $N=10^4$, $M=750$, $p_1=10$, $p_2=23$.
+Besides, the noise is chosen to be Gaussian, $\bw\sim\cN(\bzero,\bI)$, and we take $N=10^4$, $M=750$, $p_1=10$, $p_2=23$.
 
-To highlight the fact that the linear dynamical model is sufficient to catch most of the dynamical behavior of signals following the AHM, we compare the performance of Algorithm~\ref{alg:boundary} with a simple extension obtained by pointwise symmetrization~\cite{Kharitonenko02wavelet}.
-We also evaluate the performance of reference forecasting algorithms that could be used for extending such signals.
-These methods are:
-\begin{itemize}
-\item The EDMD has been developed by Williams \etal~\cite{Williams15data}.
-The proposed algorithm is a way to obtain an approximation of the so-called Koopman operator of the observed system, which theoretically allows catching dynamic of nonlinear systems~\cite{Korda18linear}.
-\item The GPR~\cite{Rasmussen06gaussian} is a method relying on a probabilistic dynamical model.
-That one is based on the Gaussian process structure, and therefore offers more flexibility in the type of dynamic that could be modeled than the linear model~\eqref{eq:dyn.model}.
-\item The TBATS method~\cite{DeLivera11forecasting} is based on a classical decomposition of times series into a trend, seasonal and ARMA components, with a specific dynamic for the seasonal component.
-This model demands the estimation of numerous parameters and, by implication, may be slow.
-\end{itemize}
-
-To quantify the global quality (\ie~not depending on $\ell$) of the forecasting approaches, we evaluate the Experimental Mean Square Error $\mathrm{MSE_{xp}}(\tilde\bx)$ of the forward forecast extended signals, namely:
-\begin{align}
-\label{eq:mse}
-\mathrm{MSE_{\xp}}(\tilde\bx) &= \dfrac1{L}\|\tilde\bx -\bx^\mathrm{ext}\|^2 \\[-1mm]
-\nonumber
-&=\! \dfrac1{L}\sum_{\ell=1}^L \bmu_{\xp}[N\!-\!1\!+\!\ell]^2 \!+\! \bgamma_{\xp}[N\!-\!1\!+\!\ell,N\!-\!1\!+\!\ell] .
-\end{align}
-where $\bx^\mathrm{ext}$ is the ground-truth extended signal, that is: $\bx^\mathrm{ext} = \begin{pmatrix}\bx[-L] & \cdots & \bx[N-1+L] \end{pmatrix}$.
-Then, as long as the bias $\bmu[N-1+\ell]$ and the variance $\bgamma[N-1+\ell,N-1+\ell]$ of the forecasting estimator remain small for all $\ell$, the MSE takes small values either.
-The numerical results when $L=1000$ are given in Table~\ref{tab:mse.sine}.
+Corresponding results are given in Table~\ref{tab:mse.sine}.
 They show that the naive extension we propose gives satisfying results, in particular in comparison with the point-symmetric extension.
-Besides, even though the other more sophisticated methods, like GPR, are able to give slightly better forecasts, these methods are substantially limited by the computing time they require, which prevents them from being used to exploit real-time data.
+Besides, even though the other more sophisticated methods, like GPR, give MSE values that have a slightly smaller standard deviation, these methods are substantially limited by the computing time they require, which prevents them from being used to exploit real-time data.
 Thus, {\sf SigExt} is the extension method that optimizes the trade-off between forecasting quality and computing time.
@@
 
-\subsection{Quality of the Reduction of Boundary Effects}
+\subsection{Real physiological signals}
 
-\subsubsection{Metrics}
-The quality of the boundary effects reduction is evaluated directly on the TF representation.
-To that aim, we compare the obtained representation to the optimal representation $\ccF_N^\mathrm{opt}(\bx)$, defined as the restriction of the representation of the ground-truth extended signal $\bx^\mathrm{ext}$.
-Therefore, we have:
-\begin{equation}
-\ccF^\mathrm{opt}(\bx_N) = \cR\left( \ccF(\bx^\mathrm{ext}) \right) \ .
-\label{eq:opt.TFR}
-\end{equation}
-
-In the aim of comparing the different techniques, we use a criterion, proposed in~\cite{Daubechies16conceft}, that quantifies the distance between a given TF representation and the optimal one.
-It is built in analogy with the optimal transport distance, which quantify the distance between two probability density functions.
-Let us generically denote a TF representation $\ccQ$.
-Then, for $t$ fixed, we consider the following probability density function: $p_\ccQ^t(\xi) = \left.
-|\ccQ(\xi,t)|^2 \middle/ \int_\RR |\ccQ(\nu,t)|^2\dd\nu \right.$.
-At each instant $t$, we then determine the optimal transport distance $d_{t}$ between the two densities.
-It is given by the $L^1$ norm of the difference between the associated distribution functions.
-In other words, we have:
-\begin{equation*}
-d_{t}(\ccQ,\ccF_0) = \int_\RR\left|\tilde P_{\ccQ}^t(\xi)-  P_{\ccF_0}^t(\xi)\right|\dd\xi\ ,
-\end{equation*}
-where $P_{\ccQ}^t(\xi)=\int_{-\infty}^\xi p_{\ccQ}^t(\nu)\dd\nu$ and $\tilde P_{\ccF_0}^t(\xi)=\int_{-\infty}^\xi\tilde p_{\ccF_0}^t(\nu)\dd\nu$.
-The \textit{Optimal Transport Distance (OTD)} quantifies the proximity between the estimated and actual instantaneous frequencies while favoring the sparsity of the estimated TF representation.
-That is why, the performance index $D(\ccQ)$ for the reduction of boundary effects of a given TF representation $\ccQ$ is given by the ratio between its averaged OTD to the optimal TF representation $\ccF^\mathrm{opt}$ (see~\eqref{eq:opt.TFR}) and the averaged OTD of the original TF representation $\ccF$ to $\ccF^\mathrm{opt}$, that is:
-%\begin{equation}
-%D(\ccQ) = \dfrac1{|I|}\bigintssss_I\ \frac{d_{t}\left(\ccQ,\ccF^\mathrm{opt}\right)}{d_{t}\left(\ccF,\ccF^\mathrm{opt}\right)}\ \dd t\ .
-%\label{eq:index.perf}
-%\end{equation}
-\begin{equation}
-D(\ccQ) = \dfrac{\displaystyle\int_I d_{t}\left(\ccQ,\ccF^\mathrm{opt}\right)\dd t}{\displaystyle\int_I d_{t}\left(\ccF,\ccF^\mathrm{opt}\right)\dd t}\ .
-\label{eq:index.perf}
-\end{equation}
-Thus, $D(\ccQ)<1$ means a reduction of the boundary effects. Let us evaluate the quality of the boundary effects reduction on biomedical signals.
-
-\subsubsection{Respiratory Signal}
-We first consider a respiratory signal of 6 hours 20 minutes. This signal, a small portion of which is displayed in Fig.~\ref{fig:tho}, is sampled at $\fs=100$~Hz.
+\subsubsection{Respiratory signal}
+We first consider a thoracic movement signal recorded by the piezo-electrical sensor of length 6 hours and 20 minutes.
+The signal is sampled at $\fs=100$~Hz.
+A small portion of this signal is displayed in Fig.~\ref{fig:tho}.
 
@@
 \includegraphics[width=.48\textwidth]{THOforecast.eps}
-\caption{Extended THO signal (blue) obtained by the {\sf SigExt} forecasting (top), the EDMD forecasting (middle), and the GPR forecasting (bottom), superimposed with the ground truth signal (red dash).}
+\caption{Extended respiratory signal (blue) obtained by the {\sf SigExt} forecasting (first panel), the EDMD forecasting (second panel), the GPR forecasting (third panel), and the TBATS forecasting (fourth panel), superimposed with the ground truth signal (red-dashed line).}
 \label{fig:tho}
@@
 From that long signal, we build a dataset of $379$ non\-overlapping signals of 60 seconds, \ie~$N=6000$.
-On each of these pieces of signal, we implement the forecasting method introduced in section~\ref{ssse:res.ahm}, including the {\sf SigExt} method detailed in Algorithm~\ref{alg:extension}.
-However, the TBATS extension method is not implemented here because of its excessive computing time.
+For each segment, we apply the forecasting methods introduced in section~\ref{ssse:res.ahm}, including the {\sf SigExt} method detailed in Algorithm~\ref{alg:extension}.
 We forecast $7$-second-long extensions on each segment of the signal, corresponding to $L =700$.
-Thus, in order to catch slowly varying dynamical behaviors, the size of the training signal $M$ is chosen so that $M=\lfloor 1.5L\rfloor$.
+Thus, in order to catch the dynamical behavior, the size of the training signal $M$ is chosen so that $M=\lfloor 1.5L\rfloor$.
 As a result of section~\ref{ssse:res.sine}, we take: $K=\lfloor2.5M\rfloor$.
-The resulting MSE~\eqref{eq:mse} averaged with respect to the simulations are given in Table~\ref{tab:THO}.
-The MSE of {\sf SigExt} is higher than the MSE of the other methods.
-This is mainly caused by the presence of a few segments of the whole signal involving complicated behaviors, that are unpredictable via a too simple dynamical model like~\eqref{eq:dyn.model}.
+The extensions obtained on one of these subsignals are shown in Fig.~\ref{fig:tho}.
+The resulting MSEs of different methods are given in Table~\ref{tab:THO}.
+The TBATS extension method is not implemented, as its excessive computing time makes it impossible to implement it on $6000$ segments within a reasonable period of time.
+%
+Note that the MSE of {\sf SigExt} is higher than the MSE of the other methods.
+This is caused by the presence of few segments contaminated by artifacts so that it is unpredictable via a too simple dynamical model like~\eqref{eq:dyn.model}.
 The left of Fig.~\ref{fig:THO.failure} illustrates one of those cases, where {\sf SigExt} fails to catch the fast varying dynamic of the instantaneous amplitude to satisfactorily forecast the signal.
@@
 We then apply {\sf BoundEffRed} for diverse TF representations: STFT, SST, RS, as well as \textit{Concentration of Frequency and Time} (ConceFT), a generalized multitaper SST-based representation introduced in~\cite{Daubechies16conceft}.
-In this application and throughout the document, we use a Gaussian window as the analysis window, regardless of the representation involved.
-In Table~\ref{tab:THO}, we give the averaged performance index~\eqref{eq:index.perf}, evaluated the whole TF representations (including boundaries).
-Even though {\sf SigExt} performs somehow moderately, the boundary effects are dramatically reduced on the TF representations, in the same order of magnitude than with the forecastings given by EDMD or GPR.
-Notice that the extension length $L$ has been set accordingly to the window length used by the TF analysis tool.
+We mention that unlike other representations, the standard RS is not causal and rigorously requires knowing the STFT on the whole time-frequency plane before making any reassignment.
+To enable real-time implementation of RS, we use here a real-time version of RS proposed in~\cite{Lin17conceft}.
+
+We set the extension length $L$ accordingly to the window length used by the TF analysis tool.
 For instance, the window length used to evaluate the STFT is of $1500$ samples.
 To prevent the STFT from being sensitive to the boundaries, we set $L=750$.
-In this way, the evaluation of the spectral content of the signal near its boundaries is not limited by a lack of information all along the window support.
+In this way, the evaluation of the spectral content of the signal near its boundary is not limited by a lack of information all along the window support.
 From now on, all results are given for $L$ equal to the half of the width of the window used in the TF transform.
+%
+In Table~\ref{tab:THO}, we give the averaged performance index~\eqref{eq:index.perf} by evaluating the whole TF representations (including boundaries).
+Even though {\sf SigExt} performs moderately well in the sense of MSE, the boundary effects are dramatically reduced on the TF representations, and the performance is in the same order of that given by EDMD or GPR.
 
@@
 For all the TF representations considered, the results show that {\sf SigExt} reduces boundary effects about as efficiently as the extensions given by EDMD or GPR.
-Incidentally, the bottom-right panel of Fig.~\ref{fig:ex.intro} shows that the SST resulting from {\sf BoundEffRed} applied to the portion of PPG displayed in Fig.~\ref{fig:ppg} produces a significant improvement in the quality of the SST near boundaries.
+For a visual inspection, the TF representation of SST resulting from the {\sf BoundEffRed} strategy is shown in the bottom-right panel of Fig.~\ref{fig:ex.intro}, where SST is applied to the portion of PPG displayed in Fig.~\ref{fig:ppg}.
+It produces a significant improvement in the quality of the SST near boundaries.
 Indeed, the blurring visible when zooming on the right boundary of the SST has almost vanished.
@@
 
-To evaluate the influence of the noise level on the performance of {\sf BoundEffRed}, we artificially add Gaussian noise to the measured PPG signal.
+To further evaluate the influence of the noise level on the performance of {\sf BoundEffRed}, we artificially add Gaussian noise to the measured PPG signal.
 It is thus an additional noise to the measurement noise actually contained in the signal.
@@
 In this example, the frequency dimension of the SST being $512$, the computational time of one column is $t_{\mathrm{SST}}=2.08$~ms, on average.
-Thus, acceptable values of $H$ for real-time implementation verify:
+A general rule to determine the acceptable values of $H$ for real-time implementation is verifying that the computational time to update the boundary-free TF representation is smaller than the lag between $H$ samples; that is,
 \[
@@
 \]
-In our example, taking $H\geq 4$~samples is sufficient, which is a reasonable latency. A video clip illustrating this is available in the GitHub repository.
+In our example, taking $H\geq 4$~samples is sufficient, which is a sufficiently reasonable latency to allow smooth monitoring of the real-time SST.
+A video clip illustrating this is available on GitHub, at \url{https://github.com/AdMeynard/BoundaryEffectsReduction/tree/master/Animations}.

### 2020-07-28 Scripts/BoundEffRedExamples/ForecastTHO_TBATS.r

@@ -1,0 +1,46 @@
+library('forecast')
+
+LoadConvertTS <- function(TabName) {
+  x = read.table(TabName)
+  y = ts(x)
+  return(y)
+}			
+
+setwd(getSrcDirectory(LoadConvertTS)[1])
+
+x = LoadConvertTS("../../Signals/THO4R")
+xext0 = LoadConvertTS("../../Signals/THOext4R")
+
+fs = 100
+extSEC = 7 # the extension is of extSEC second
+L = round( extSEC*fs )
+
+N = length(x)
+t = seq(from = 0, to = (N-1)*fs, by = fs)
+tt = seq(from = 0, to = (N+L-1)*fs, by = fs)
+
+xn = x+rnorm(N,sd=0.3) # regularization via awgn
+
+ss = 366 #seq(from=363, to=369, by=0.5)
+matExt = matrix(0,nrow=length(ss),ncol=L)
+compVal = rep(0,length(ss))
+k = 1
+for (s in ss) {
+  fit <- tbats(xn,use.box.cox=FALSE,use.trend=FALSE,use.damped.trend=FALSE,seasonal.periods=s)
+  TBATSest = forecast( fit, h=L )
+  xext = as.numeric( TBATSest[["mean"]] )
+  matExt[k,] = xext
+  compVal[k] = sum((xext - xext0)^2)
+  k = k+1
+}
+
+kopt = which.min(compVal)
+xextOpt = matExt[kopt,]
+
+xx0 = c( x, xext0)
+xxTBATS = c( x, xextOpt)
+plot(tt,xx0,type="l", col="blue", lty=2)
+lines(tt,xxTBATS,type="l", col="red", lty=1)
+lines(t,x,type="l", col="blue", lty=1)
+
+write.csv(xextOpt,'../../Results/THO_TBATS.csv')

### 2020-07-28 Scripts/BoundEffRedExamples/PPGexample.m

@@ -82,8 +82,8 @@ set(gca,'fontsize',18);
 %% SST and STFT
 basicTF.hop = 10;
 basicTF.win = 1501;
-fmin = 0 ;
-fmax = 0.01 ;
+fmin = 0/fs ;
+fmax = 4/fs ;
 df = 1e-5 ;
 
 % On the original signal

### 2020-07-28 Scripts/BoundEffRedExamples/THOexample.m

@@ -41,6 +41,10 @@ tic;
 xxGPR = SigExtension(x,fs,HOP,extK,extM,extSEC,method); 
 GPRtime = toc;
 
+dataTBATS = table2array( readtable('../../Results/THO_TBATS.csv') ) ;
+xxTBATS = [zeros(L,1); x; dataTBATS(:,2)] ;
+
+
 %% Plot
 t = linspace(0, (N-1)/fs, N) ;
 tt = linspace(-L/fs, (N-1+L)/fs, N+2*L) ;
@@ -50,24 +54,31 @@ xxZP = [ zeros(L,1); x; zeros(L,1) ] ; % zero padding
 
 figure;
 
-subplot(3,1,1)
+subplot(4,1,1)
 plot(tt,xxLSE,tt,xxTRUE,'--',t,x,'linewidth',2); grid on;
 ylabel('Signals'); axis tight; xlim([t(end)-1.5*extSEC tt(end)]);
 xticks([]);
 legend({'{\sf SigExt} extension','Ground truth extension','Original signal'},'location','northwest','interpreter','latex');
 set(gca,'fontsize',18);
 
-subplot(3,1,2)
+subplot(4,1,2)
 plot(tt,xxEDMD,tt,xxTRUE,'--',t,x,'linewidth',2); grid on;
 ylabel('Signals'); axis tight; xlim([t(end)-1.5*extSEC tt(end)]);
 xticks([]);
 legend({'EDMD extension','Ground truth extension','Original signal'},'location','northwest','interpreter','latex');
 set(gca,'fontsize',18);
 
-subplot(3,1,3)
+subplot(4,1,3)
 plot(tt,xxGPR,tt,xxTRUE,'--',t,x,'linewidth',2); grid on;
-xlabel('Time (s)'); ylabel('Signals'); axis tight; xlim([t(end)-1.5*extSEC tt(end)]);
+ylabel('Signals'); axis tight; xlim([t(end)-1.5*extSEC tt(end)]);
+xticks([]);
 legend({'GPR extension','Ground truth extension','Original signal'},'location','northwest','interpreter','latex');
+set(gca,'fontsize',18);
+
+subplot(4,1,4)
+plot(tt,xxTBATS,tt,xxTRUE,'--',t,x,'linewidth',2); grid on;
+xlabel('Time (s)'); ylabel('Signals'); axis tight; xlim([t(end)-1.5*extSEC tt(end)]);
[…]

### 2020-07-28 Scripts/BoundEffRedExamples/figPPG.m

@@ -9,47 +9,26 @@ data = load('PPG.mat');
 xtot = data.PPG;
 
 fs = data.fs ; % sampling frequency
+
+n0 = 6e3 ;
 N = 4e3; % subsignals length
-% Subsignal
-k = 5 ;
-x = xtot((k*N+1):((k+1)*N)) ;
-mu = mean(x);
-sigma = std(x);
-x = (x - mu) / sigma ;
+x0 = xtot((n0):(n0+N-1)) ; % Subsignal
+mu = mean(x0);
+sigma = std(x0);
+x0 = (x0 - mu) / sigma ;
 
-%% SST
-basicTF.hop = 10 ;
-basicTF.win = 1501 ;
-basicTF.fmin = 0 ;
-basicTF.fmax = 0.06 ;
+dgamma = cumsum(randn(N,1)) ;
+windowSize = 50; 
+b = (1/windowSize)*ones(1,windowSize);
+a = 1;
+dgamma = filter(b,a,dgamma);
+dgamma = dgamma - mean(dgamma) ;
+dgamma = 0.2*dgamma/max(abs(dgamma)) ;
+dgamma = dgamma + 1 ;
 
-t = linspace(0, (N-1)/fs, N) ;
-tsst = t(1:basicTF.hop:end);
-
-[~, sstS, ~] = ConceFT_sqSTFT_C(x, basicTF.fmin, basicTF.fmax,1e-5, basicTF.hop, basicTF.win, 1, 10, 1, 0, 0) ;
-ff = linspace(0,fs*basicTF.fmax,400) ;
-
-figure;
-subplot(3,1,1);
-plot(t,x,'linewidth',2); axis tight ; 
-ylabel('Signal'); 
-set(gca,'fontsize',18); grid on; xticklabels([]);
-
-subplot(3,1,[2 3]);
-imagesc(tsst,ff,log1p(abs(sstS)/1e1)); axis xy; 
-xlabel('Time (sec.)'); ylabel('Frequency (Hz)'); colormap(1-gray);
-set(gca,'fontsize',18);
-
-figure;
-subplot(3,1,1);
-plot(t,x,'linewidth',2); axis tight ; xlim([27 t(end)]) ;
-ylabel('Signal'); 
-set(gca,'fontsize',18); grid on; xticklabels([]);
-
-subplot(3,1,[2 3]);
[…]

### 2020-07-28 Notes/results.tex
paper

@@
 %
-Note that the MSE of {\sf SigExt} is higher than the MSE of the other methods.
+Note that the MSE of {\sf SigExt} is higher than the MSE of EDMD or the symmetric extension.
 This is caused by the presence of few segments contaminated by artifacts so that it is unpredictable via a too simple dynamical model like~\eqref{eq:dyn.model}.
 The left of Fig.~\ref{fig:THO.failure} illustrates one of those cases, where {\sf SigExt} fails to catch the fast varying dynamic of the instantaneous amplitude to satisfactorily forecast the signal.
-The EDMD and Symmetric extensions are more robust to those situations, as shown in this example, and in Table~\ref{tab:THO}.
+The EDMD and symmetric extensions are more robust to those situations, as shown in this example, and in Table~\ref{tab:THO}.
 Nevertheless, {\sf SigExt} provides a sufficiently relevant extension to give TF representations sparingly affected by boundary effects.
@@
 We set the extension length $L$ accordingly to the window length used by the TF analysis tool.
-For instance, the window length used to evaluate the STFT is of $1500$ samples.
-To prevent the STFT from being sensitive to the boundaries, we set $L=750$.
+For instance, the window length used to evaluate the STFT is of $1400$ samples.
+To prevent the STFT from being sensitive to the boundaries, we set $L=700$.
 In this way, the evaluation of the spectral content of the signal near its boundary is not limited by a lack of information all along the window support.
@@
    \hhline{|=#=#=|=|=|=|}
-   {\sf SigExt} & $0.046$ & $0.698$ &  $0.705$ & $0.694$ & $0.776$ \\
+   {\sf SigExt} & $0.292$ & $0.370$ &  $0.408$ & $0.866$ & $0.423$ \\
    \hline
-   Symmetric & $0.039$ & $0.984$ & $1.280$ & $9.063$ & $1.295$ \\
+   Symmetric & $0.044$ & $1.162$ & $1.173$ & $1.022$ & $1.144$ \\
    \hline
-   EDMD & $0.022$ & $0.635$ &  $0.757$ & $0.701$ & $0.793$ \\
+   EDMD & $0.026$ & $0.359$ &  $0.422$ & $0.828$ & $0.449$ \\
    \hline
-   GPR & $0.045$ & $0.793$ &  $0.817$ & $0.727$ & $0.876$ \\ 
+   GPR & $0.331$ & $0.391$ &  $0.411$ & $0.897$ & $0.430$ \\ 
    \hline

### 2020-07-28 Scripts/EvalPerfBoundEffRed/ResultsBoundEffRed.m

@@ -41,10 +41,10 @@ fprintf(' ___________________________________________________________________\n'
 fprintf('| Extension Method | Computing time (sec.)  |   Forecasting Error   |\n')
 fprintf('|                  |                        |    Mean   |   Median  |\n')
 fprintf('|------------------|------------------------|-----------|-----------|\n')
-fprintf('|     SigExt       |           %.3f        | %.3e | %.3e |\n', CompTime.LSE , mean(forecastErr.LSE(~isnan(OTD.sst.LSE))),median(forecastErr.LSE(~isnan(OTD.sst.LSE))) )
-fprintf('|  Symmetrization  |           %.3f        | %.3e | %.3e |\n', CompTime.SYM , mean(forecastErr.SYM), median(forecastErr.SYM) )
-% fprintf('|      EDMD        |          %.3f        | %.3e | %.3e |\n', CompTime.EDMD , mean(forecastErr.EDMD), median(forecastErr.EDMD) )
-% fprintf('|      GPR         |         %.3f        | %.3e | %.3e |\n', CompTime.GPR , mean(forecastErr.GPR(~isnan(OTD.sst.GPR))), median(forecastErr.GPR(~isnan(OTD.sst.GPR))) )
+fprintf('|     SigExt       |           %.3f        | %.3f | %.3e |\n', CompTime.LSE , mean(forecastErr.LSE(~isnan(OTD.sst.LSE))),median(forecastErr.LSE(~isnan(OTD.sst.LSE))) )
+fprintf('|  Symmetrization  |           %.3f        | %.3f | %.3e |\n', CompTime.SYM , mean(forecastErr.SYM), median(forecastErr.SYM) )
+fprintf('|      EDMD        |           %.3f        | %.3f | %.3e |\n', CompTime.EDMD , mean(forecastErr.EDMD), median(forecastErr.EDMD) )
[…]

### 2020-08-03 Notes/Supplementary material/SupMat.tex
paper

@@
 \]
-where "$\vec$" denotes the vectorization operator, concatenating the columns of a given matrix on top of one another. Then, by definition, we have:
+where "$\vec$" denotes the vectorization operator, concatenating the columns of a given matrix on top of one another.
+Then, we can write $\bh^{(\ell)}$ as $\bh^{(\ell)}=f^{(\ell)}(\bg)$ where $f^{(\ell)}$ is a deterministic function such that:
+\begin{align*}
+f^{(\ell)} : \RR^{M(M+1)} &\to \RR^{M} \\
+ \bg &\mapsto \bh^{(\ell)}\ . \\
+\end{align*}
+
+The multivariate version of the delta method is applicable to $\bh^{(\ell)}$ under the following conditions:
+\begin{enumerate}[(i)]
+\item\label{delta.1} $f^{(\ell)}$ is differentiable at the origin;
+\item\label{delta.2} $\sqrt{K}\ \bg \xrightarrow[K\to\infty]{\cD} \cN(\bzero,\bGamma_0)$ where $\bGamma_0$ is a covariance matrix.
+\end{enumerate}
+Condition~\ref{delta.1} is satisfied.
+Indeed, the differentiation of $f^{(\ell)}$ can be checked as a composition of standard matrix derivation rules.
+The detailed calculation of the derivation is not given here for the sake of conciseness.
+Concerning condition~\ref{delta.2}, by definition of $\bg$, we have:
 \[
@@
 \]
-where:
+where $\bg_1$ and $\bg_2$ are defined by
 \begin{align*}
-\bg_1 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left( \tilde\bz_k\bw_k^T + \tilde\bw_k\bz_k^T \right) \\
-\bg_2 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left(\tilde\bw_k\bw_k^T-\tilde\bI \right) \,,
+\bg_1 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left( \tilde\bz_k\bw_k^T + \tilde\bw_k\bz_k^T \right)\quad\mbox{where}\quad \tilde\bz_k^T= \begin{pmatrix} \bz_k^T & \bz_{k+1}[M-1] \end{pmatrix}\,, \\
+\bg_2 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left(\tilde\bw_k\bw_k^T-\tilde\bI \right) \quad\mbox{where}\quad \tilde\bw_k^T= \begin{pmatrix} \bw_k^T & \bw_{k+1}[M-1] \end{pmatrix} \,.
 \end{align*}
-where $\tilde\bz_k^T= \begin{pmatrix} \bz_k^T & \bz_{k+1}[M-1] \end{pmatrix}$ and $\tilde\bw_k^T= \begin{pmatrix} \bw_k^T & \bw_{k+1}[M-1] \end{pmatrix}$.
-Then, $\bg_1$ is a Gaussian random vector because it is a linear combination of Gaussian random vectors.
-Moreover, using the central limit theorem under weak dependence, we can show that $\bg_2$ also converges towards a Gaussian random vector as $K\to\infty$.
+First, $\bg_1$ is intrinsically a Gaussian random vector since it is a linear combination of Gaussian random vectors.
+Second, using the central limit theorem under weak dependence, we can show that $\bg_2$ also converges towards a Gaussian random vector as $K\to\infty$.
 Combining these two results leads to
@@
 \]
-where $\bGamma_0=\EE\left\{\bg\bg^T\right\}$ is a covariance matrix.
-
-Furthermore, one can write $\bh^{(\ell)}$ as $\bh^{(\ell)}=f^{(\ell)}(\bg)$ where $f^{(\ell)}$ is a deterministic function such that:
-\begin{align*}
-f^{(\ell)} : \RR^{M(M+1)} &\to \RR^{M} \\
- \bg &\mapsto \bh^{(\ell)}\ . \\
-\end{align*}
-Then, as $f^{(\ell)}$ is a differentiable function with non-zero differentiation at $0$ {\color{red}[I assume that you have checked all conditions for the delta method.
-You may consider adding its formula to save the readers' time]}, using the Delta method gives:
+where $\bGamma_0$ is the limit covariance matrix so that $\bGamma_0=\lim\limits_{K\to\infty} K\,\EE\{\bg\bg^T\}$.
+Hence, conditions~\ref{delta.1} and~\ref{delta.1} are satisfied. The delta method can therefore be applied; the result is as follows:
 \begin{align*}
@@
 \end{align*}
-where $\bF^{(\ell)}$ is the Jacobian matrix such that:
+where $\bF^{(\ell)}$ is the Jacobian matrix of $f^{(\ell)}$ at the origin, that is
 \[
@@
 
-\subsection{Expression of the bias $\bmu$.}
-Clearly, $\bmu[n]=0$ when $n\in I$. When $n=N-1+\ell$, we have
-\begin{align*}
+\subsection{Expression of the Bias $\bmu$.}
+By definition of the measurement noise, $\bmu[n]=0$ when $n\in I$.
+Outside the measurement interval $I$, denote by $\ell$ the index such that $n=N-1+\ell$.
+Then, given that $\bh^{(\ell)} = \balpha^{(\ell)} - \balpha_0^{(\ell)}$, we have
+\begin{align}
+\nonumber
 \bmu[n] &=  \EE\{\balpha^{(\ell)}\}\bz_{K} + \sigma\EE\{\balpha^{(\ell)}\bw_{K}\} - \bz[n]\\
-& = \balpha_0^{(\ell)}\bz_{K} + \EE\{\bh^{(\ell)}\}\bz_{K} + \sigma\EE\{\bh^{(\ell)}\bw_{K}\} - \bz[N-1+\ell]
-\end{align*}
+& = \balpha_0^{(\ell)}\bz_{K} + \EE\{\bh^{(\ell)}\}\bz_{K} + \sigma\EE\{\bh^{(\ell)}\bw_{K}\} - \bz[N-1+\ell]\ .
+\label{eq:bias.dec}
+\end{align}
 Let us first evaluate the expression of $\balpha_0^{(\ell)}\bz_K$. We have
-\begin{align*}
+\begin{align}
+\nonumber
 \bS^{(a)}[m,m'] &= \sigma^2\delta_{(m+a)m'}+\sum_{j,j'=1}^J\dfrac{\Omega_j\Omega_{j'}}{K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(N_0+m+a+k)+\varphi_j\right)\cos\left(2\pi \frac{f_{j'}}{\fs}(N_0+m'+k)+\varphi_{j'}\right) \\
+\nonumber
 &= \sigma^2\delta_{(m+a)m'}+\sum_{j=1}^J\dfrac{\Omega_j^2}{2K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)\\
-&= \sigma^2\delta_{(m+a)m'}+\sum_{j=1}^J\left( \dfrac{\Omega_j^2}2\cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \dfrac{\Omega_j^2}{2K}\underbrace{\sum_{k=0}^{K-1}\cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)}_{=0\ \mathrm{because}\ \frac{f_j}\fs=\frac{p'_j}K } \right)\\
+\nonumber
+&= \sigma^2\delta_{(m+a)m'}+\sum_{j=1}^J\Bigg( \dfrac{\Omega_j^2}2\cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \dfrac{\Omega_j^2}{2K}\underbrace{\sum_{k=0}^{K-1}\cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)}_{=0\ \mathrm{because}\ \frac{f_j}\fs=\frac{p'_j}K } \Bigg)\\
 &= \sigma^2\delta_{(m+a)m'}+\sum_{j=1}^J\dfrac{\Omega_j^2}2\cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right)\ .
-\end{align*}
+\label{eq:S1}
+\end{align}
 Thus, $\bS^{(0)}$ is a circulant matrix and is therefore diagonalizable in the Fourier basis:
@@
 which leads to
-\begin{align*}
-{\bS^{(0)}}^\inv[m,m']  &= \dfrac1{\sigma^2}\delta_{m,m'}-\sum_{j=1}^J\dfrac{\Omega_j^2}{2\sigma^2(\sigma^2+\Omega_j^2M/4)}\cos\left(2\pi p_j \dfrac{m-m'}{M}\right)\,.
-\end{align*}
-Consequently, we have
+\begin{equation}
+{\bS^{(0)}}^\inv[m,m']  = \dfrac1{\sigma^2}\delta_{m,m'}-\sum_{j=1}^J\dfrac{\Omega_j^2}{2\sigma^2(\sigma^2+\Omega_j^2M/4)}\cos\left(2\pi p_j \dfrac{m-m'}{M}\right)\,.
+\label{eq:S0.inv}
+\end{equation}
+Consequently, combining equations~\eqref{eq:S1} and~\eqref{eq:S0.inv}, we have
 \begin{align}
 \nonumber
-\tilde\bA_0[m,m']  &= \sum_{q=0}^{M-1} {\bS^{(1)}}[m,q]{\bS^{(0)}}^\inv[q,m'] \\
+\bA_0[m,m']  &= \sum_{q=0}^{M-1} {\bS^{(1)}}[m,q]{\bS^{(0)}}^\inv[q,m'] \\
 &= \delta_{m+1,m'} + \sum_{j=1}^J\dfrac{2\Omega_j^2}{\Omega_j^2M+4\sigma^2}\cos\left(2\pi p_j\frac{m'}{M}\right)\delta_{m+1,M}\,.
@@
 \end{align}
-Thus,
+Thus $\balpha_0^{(1)}$, the last row of $\bA_0$, is written as
 \begin{align*}
-\tilde\balpha_0^{(1)}[m]  &=\sum_{j=1}^J\dfrac{2\Omega_j^2}{\Omega_j^2M+4\sigma^2}\cos\left(2\pi p_j\frac{m}{M}\right) \\
+\balpha_0^{(1)}[m]  &=\sum_{j=1}^J\dfrac{2\Omega_j^2}{\Omega_j^2M+4\sigma^2}\cos\left(2\pi p_j\frac{m}{M}\right) \\
 &= \dfrac{2}{M}\sum_{j=1}^J\cos\left(2\pi p_j\frac{m}{M}\right) + o(\sigma)\ .
@@
 \begin{align*}
-\tilde\bA_0\bz_{K} &= 
+\bA_0\bz_{K} &= 
 \begin{pmatrix}
@@
 \nonumber
-\balpha_0^{(\ell)}\bz_{K} &= \tilde\balpha_0^{(1)}\tilde\bA_0^{\ell-1}\bz_{K} \\
+\balpha_0^{(\ell)}\bz_{K} &= \balpha_0^{(1)}\tilde\bA_0^{\ell-1}\bz_{K} \\
 &=\sum_{m=0}^{M-\ell}\balpha_0^{(1)}[m]\bz[N-M+\ell+m-1]+\sum_{m=M-\ell+1}^{M-1}\balpha_0^{(1)}[m]\balpha_0^{(m-M+\ell)}\bz_K\,.
@@
 \label{eq:alpha0z.sine}
-\balpha_0^{(\ell)}\bz_{K} = \bz[N-1+\ell] + o(\sigma)
+\balpha_0^{(\ell)}\bz_{K} = \bz[N-1+\ell] + o(\sigma)\ .
 \end{equation}
-Then,
+Then,  inserting result~\eqref{eq:alpha0z.sine} into equation~\eqref{eq:bias.dec} gives
 \begin{equation*}
@@
 \end{equation*}
-Besides, from Lemma~1, we have the following results:
-\begin{align*}
-\EE\{\bh^{(\ell)}\} &\underset{K\to\infty}{\longrightarrow} 0\\
-\EE\{\bh^{(\ell)}\bw_{K}\} &\underset{K\to\infty}{\longrightarrow} 0
-\end{align*}
-{\color{red}[if it is possible to get the convergence rate, like $\EE\{\bh^{(\ell)}\}=O(1/K)$, that would be great!]} Consequently,
+Besides, from Lemma~1, we have the following result:
+\begin{equation*}
+\EE\{\bh^{(\ell)}\} \underset{K\to\infty}{\longrightarrow} 0\ .
+\end{equation*}
+{\color{red}[if it is possible to get the convergence rate, like $\EE\{\bh^{(\ell)}\}=O(1/K)$, that would be great!]}
+Moreover, applying delta method to $\left(\bg\ \bw[n']\right)^T$, $\forall n'\in I$, one can show that we have the asymptotic result
+\begin{equation}
+\EE\{\bw[n']\bh^{(\ell)}\} \underset{K\to\infty}{\longrightarrow} 0\,.
+\label{eq:cov.ap}
+\end{equation}
+Consequently, $\EE\{\bh^{(\ell)}\bw_{K}\} \underset{K\to\infty}{\longrightarrow} 0$, and
 \begin{equation}
@@
 
-\subsection{Expression of the covariance $\bgamma$.}
+\subsection{Expression of the Covariance $\bgamma$.}
 Take $n>N$, and denote $n=N-1+\ell$.  To derive the covariance, let us segregate the cases.
@@
 The discrepancy should be made clear.]}.
-Then, combining result~\eqref{eq:bias.ap} and equation~\eqref{I-eq:bias0}, we obtain:
-\begin{align*}
-\bgamma[n,n'] &= \sigma\EE\{\bw[n']\bh^{(\ell)}\}\bz_K + \sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bw[n']\} + o(\sigma^2) 
-\end{align*}
+Then, inserting result~\eqref{eq:bias.ap} into equation~\eqref{I-eq:bias0}, we obtain:
+\begin{equation}
+\bgamma[n,n'] = \sigma\EE\{\bw[n']\bh^{(\ell)}\}\bz_K + \sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bw[n']\} + o(\sigma^2) 
+\label{eq:gamma.nn}
+\end{equation}
 when $K\to \infty$.
-We remark that
+In addition, we remark that
 \[
@@
 \]
-Moreover, using Delta method one can show that we have the asymptotic result
-\begin{equation}
-\EE\{\bw[n']\bh^{(\ell)}\} \underset{K\to\infty}{\longrightarrow} 0\,.
-\label{eq:cov.ap}
-\end{equation}
-As a result, we have
+Besides, according to result~\eqref{eq:cov.ap} the first term in~\eqref{eq:gamma.nn}, we thus have
 \begin{align*}
@@
 when $K\to\infty$.
-{\color{red}[Note that $o(\sigma^2)$ becomes $0$ when $K\to \infty$.
-So use either ``$\bgamma[n,n']= \sigma^2\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)}+ o(\sigma^2)$ when $K\to \infty$'' or ``$\bgamma[n,n'] \to \sigma^2\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)}$ when $K\to \infty$''.]}
 
 \paragraph{When $n'\geq N$}
-From Lemma~1, we have that $h^{(\ell)}$ is asymptotically Gaussian.
-Then, as a direct consequence of the Isserlis' theorem, odd-order moments are vanishing.
-{\color{red}[The same gap here.]} Then, combining equations~\eqref{I-eq:cov01}, \eqref{eq:bias.ap} and~\eqref{eq:cov.ap}, we obtain
+Inserting equation~\eqref{eq:bias.ap} into result~\eqref{I-eq:cov01}, we obtain
 \begin{align*}  
@@
 & + \sigma\EE\{\bh^{(\lambda)}\bw_K\bh^{(\ell)}\}\bz_K+\sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\lambda)}\bw_K\}+\sigma^2\balpha_0^{(\lambda)}\EE\{\bw_K\bh^{(\ell)}\bw_K\} \\
-&+\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\sigma^2\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} + o(\sigma^2) \\
-= &\,{\color{blue}\dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K} +\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\sigma^2\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} + o(\sigma^2)
+&+\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\sigma^2\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} + o(\sigma^2)\ .
 \end{align*}
-when $K\to\infty$ {\color{red}[in the last equation, make clear which term depends on which equation \eqref{I-eq:cov01}, \eqref{eq:bias.ap} and~\eqref{eq:cov.ap}.] [The previous or current asymptotic expressions with $\dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K$ are both not ideal.
-Either using your previous notation or following what I prefer here, the behavior of $\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K$ should be made clear.
-Is it bounded?
-If it is bounded, then use $=O(1/K)$ to replace $\to \dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K$.
-If not, how does it depend on $K$ should be made clear.] [Moreover, if you prefer to use your previous notation, you may clarify what it means by $\sim$.]}.
-The Isserlis' theorem also gives the following result {\color{red}[the same gap here.]}
+From Lemma~1, we have that $\bh^{(\ell)}$ is asymptotically Gaussian.
+Then, as a direct consequence of the Isserlis' theorem, odd-order moments are vanishing.
+{\color{red}[The same gap here.]} {\color{blue}Considering this property and equation~\eqref{eq:cov.ap} gives
+\begin{align*}
+\bgamma[n,n'] = &\,\dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K + \sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\sigma^2\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} + o(\sigma^2)
+\end{align*}
+when $K\to\infty$. The Isserlis' theorem also gives the following result {\color{red}[the same gap here.]}
 \begin{align*}
@@
 &\hspace{35pt} + \EE\{\bh^{(\ell)}[m]\bh^{(\lambda)}[m']\}\underbrace{\EE\{\bw_K[m]\bw_K[m']\}}_{=\delta_{m,m'}} \\
-&\to \sum_{m=0}^{M-1} \EE\{\bh^{(\ell)}[m]\bh^{(\lambda)}[m]\} = \dfrac1K\Tr\left(\bGamma^{(\ell,\lambda)}\right)
+&= \sum_{m=0}^{M-1} \EE\{\bh^{(\ell)}[m]\bh^{(\lambda)}[m]\} = \dfrac1K\Tr\left(\bGamma^{(\ell,\lambda)}\right)
 \end{align*}
@@
 \begin{align*}  
-\bgamma[n,n'] &\to {\color{blue}\dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K} +\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\dfrac{\sigma^2}{K}\Tr\left(\bGamma^{(\ell,\lambda)}\right) + o(\sigma^2)
+\bgamma[n,n'] &= \dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K+\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\dfrac{\sigma^2}{K}\Tr\left(\bGamma^{(\ell,\lambda)}\right) + o(\sigma^2)
 \end{align*}
-when $K\to \infty$.
+when $K\to \infty$.}
 

### 2020-08-03 Notes/results.tex
paper

@@
 
-\subsection{Extension methods for comparison}
+\subsection{Extension Methods for Comparison}
 To highlight the fact that the linear dynamical model is sufficient to catch most of the dynamical behavior of signals following the AHM, we compare the performance of Algorithm~\ref{alg:boundary} with a simple extension obtained by pointwise symmetrization~\cite{Kharitonenko02wavelet}.
@@
 
-\subsection{Evaluation metrics}
+\subsection{Evaluation Metrics}
 
@@
 
-\subsection{Simulated signals}
+\subsection{Simulated Signals}
 We first evaluate the quality of the forecasting step and compare it to the theoretical results provided by Theorem~\ref{th:error}.
@@
 
-\subsection{Real physiological signals}
+\subsection{Real Physiological Signals}
 
-\subsubsection{Respiratory signal}
+\subsubsection{Respiratory Signal}
 We first consider a thoracic movement signal recorded by the piezo-electrical sensor of length 6 hours and 20 minutes.
@@
 The forecasting step in each iteration of {\sf BoundEffRed} then takes no more than $t_{\mathrm{forecast}}=46$~ms.
-As indicated in section~\ref{sse:extension.TFR}, the update of the SST requires the calculation of $\lceil L/H \rceil$ new columns of the SST, where $H$ denotes the hop size.
+Denote by $H\geq 1$ the hop size in samples, that is the number of samples between successive columns of the SST.
+Besides, since the support of the analysis window covers $2L$ samples, the update of the SST requires the calculation of $\lceil 2L/H \rceil$ new columns of the SST.
 In this example, the frequency dimension of the SST being $512$, the computational time of one column is $t_{\mathrm{SST}}=2.08$~ms, on average.
@@
 \]
-In our example, taking $H\geq 4$~samples is sufficient, which is a sufficiently reasonable latency to allow smooth monitoring of the real-time SST.
+In our example, taking $H\geq 8$~samples is sufficient to ensure the feasibility of real-time implementation.
+This value is reasonable because it allows a maximum overlap of $98.4\%$ of the window length.
+SST variations are therefore perceived gradually.
+Combined with the low latency of the algorithm, this allows for smooth monitoring of the real-time SST.
 A video clip illustrating this is available on GitHub, at \url{https://github.com/AdMeynard/BoundaryEffectsReduction/tree/master/Animations}.

### 2020-08-03 Scripts/EvalPerfSigExt/InfluenceNoiseSine.m

@@ -49,7 +49,7 @@ end
 zK = x0(end-extM+1 : HOP: end).' ;
 
 %% Forecasting
-method.name = 'lseV' ;
+method.name = 'SigExt' ;
 nbXP = 3 ; % number of noise levels
 nbXPP = 3000 ; % number of XP / noise level
 Sigma = logspace(-3,-1,nbXP) ;

### 2020-08-04 Notes/Supplementary material/SupMat.tex
code

@@
 
-\title{Supplementary material for ``An Efficient Forecasting Approach for the Real-Time Reduction of Boundary Effects in Time-Frequency Representations''}
+\title{Supplementary material for ``An Efficient Forecasting Approach to Reduce Boundary Effects in Real-Time Time-Frequency Analysis''}
 
@@
 \end{enumerate}
-Condition~\ref{delta.1} is satisfied.
-Indeed, the differentiation of $f^{(\ell)}$ can be checked as a composition of standard matrix derivation rules.
-The detailed calculation of the derivation is not given here for the sake of conciseness.
+When these conditions are satisfied, the delta method gives the asymptotic behavior of $\bh^{(\ell)}=f^{(\ell)}(\bg)$, that is
+\begin{equation}
+\sqrt{K}\ \bh^{(\ell)} \xrightarrow[K\to\infty]{\cD} \cN(\bzero,{\bF^{(\ell)}}^T\bGamma_0\bF^{(\ell)})\ ,
+\label{eq:delta}
+\end{equation}
+where $\bF^{(\ell)}$ is the Jacobian matrix of $f^{(\ell)}$ at the origin, that is
+\[
+\bF^{(\ell)}[m,m'] = \left.\dfrac{\partial f^{(\ell)}_m}{\partial\bg[m']}\right\vert_{\bg=\bzero}\ .
+\]
+
+Condition~\ref{delta.1} is satisfied. Indeed, the differentiation of $f^{(\ell)}$ can be checked as a composition of standard matrix derivation rules. {\color{red}Detail it}
+
 Concerning condition~\ref{delta.2}, by definition of $\bg$, we have:
@@
 \bg_1 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left( \tilde\bz_k\bw_k^T + \tilde\bw_k\bz_k^T \right)\quad\mbox{where}\quad \tilde\bz_k^T= \begin{pmatrix} \bz_k^T & \bz_{k+1}[M-1] \end{pmatrix}\,, \\
-\bg_2 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left(\tilde\bw_k\bw_k^T-\tilde\bI \right) \quad\mbox{where}\quad \tilde\bw_k^T= \begin{pmatrix} \bw_k^T & \bw_{k+1}[M-1] \end{pmatrix} \,.
+\bg_2 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left(\tilde\bw_k\bw_k^T-\bI \right) \quad\mbox{where}\quad \tilde\bw_k^T= \begin{pmatrix} \bw_k^T & \bw_{k+1}[M-1] \end{pmatrix} \,.
 \end{align*}
@@
 Second, using the central limit theorem under weak dependence, we can show that $\bg_2$ also converges towards a Gaussian random vector as $K\to\infty$.
-Combining these two results leads to
+Hence, we have
 \[
@@
 where $\bGamma_0$ is the limit covariance matrix so that $\bGamma_0=\lim\limits_{K\to\infty} K\,\EE\{\bg\bg^T\}$.
-Hence, conditions~\ref{delta.1} and~\ref{delta.1} are satisfied. The delta method can therefore be applied; the result is as follows:
-\begin{align*}
-\sqrt{K}\ \bh^{(\ell)} \xrightarrow[K\to\infty]{\cD} \cN(\bzero,{\bF^{(\ell)}}^T\bGamma_0\bF^{(\ell)})\ ,
-\end{align*}
-where $\bF^{(\ell)}$ is the Jacobian matrix of $f^{(\ell)}$ at the origin, that is
-\[
-\bF^{(\ell)}[m,m'] = \left.\dfrac{\partial f^{(\ell)}_m}{\partial\bg[m']}\right\vert_{\bg=\bzero}\ .
-\]
+Hence, conditions~\ref{delta.1} and~\ref{delta.1} are satisfied.
+The delta method can therefore be applied, and result~\eqref{eq:delta} is valid.
 
@@
 \]
-Besides, according to result~\eqref{eq:cov.ap} the first term in~\eqref{eq:gamma.nn}, we thus have
+Besides, according to result~\eqref{eq:cov.ap} the first term in~\eqref{eq:gamma.nn} vanishes when $K\to\infty$. We thus have
 \begin{align*}
@@
 The dataset is constructed from a $500$-second-long ECG, sampled at $\fs=200$~Hz, cut into $10$ segments of $50$ seconds each.
-Fig.~\ref{fig:ecg} depicts the right boundary of one of these subsignals, together with the $6$-second extensions estimated by {\sf SigExt} (top panel), or EDMD (middle panel), or GPR (bottom panel).
+Fig.~\ref{fig:ecg} depicts the right boundary of one of these subsignals, together with the $6$-second extensions estimated by {\sf SigExt} (first panel), or EDMD (second panel), GPR (third panel), or TBATS (fourth panel).
 These extensions are superimposed to the ground-truth extension, plotted in red.
@@
 \includegraphics[width=\textwidth]{ECGforecast.eps}
-\caption{Extended ECG (blue) obtained by the {\sf SigExt} forecasting (top), the EDMD forecasting (middle), and the GPR forecasting (bottom), superimposed with the ground truth signal (red).}
+\caption{Extended ECG (blue) obtained by the {\sf SigExt} forecasting (first panel), the EDMD forecasting (second panel), the GPR forecasting (third panel), and the TBATS forecasting (fourth panel), superimposed with the ground truth signal (dashed red).}
 \label{fig:ecg}

### 2020-08-04 Notes/results.tex
code

@@
 
-Corresponding results are given in Table~\ref{tab:mse.sine}.
+The forecasting methods are applied to $1000$ realizations of the above-described signal.
+The mean and the standard deviation (SD) of the resulting MSEs are given in Table~\ref{tab:mse.sine}.
 They show that the naive extension we propose gives satisfying results, in particular in comparison with the point-symmetric extension.
@@
   \hline
-   \multirow{2}{40pt}{\centering Extension method} & \multirow{2}{70pt}{\centering MSE averaged on $50$ realizations}  & \multirow{2}{80pt}{\centering Averaged computing time (seconds)} \\
+   \multirow{2}{40pt}{\centering Extension method} & \multirow{2}{70pt}{\centering MSE (mean $\pm$ SD)}  & \multirow{2}{80pt}{\centering Averaged computing time (seconds)} \\
     &  & \\
    \hhline{|=#=|=|}
-   {\sf SigExt} & $1.458$ & $1.069$ \\
+   {\sf SigExt} & $1.028 \pm  0.404$ & $1.069$ \\
    \hline
-   Symmetric & $8.735$ & $0.001$ \\
+   Symmetric & $4.368 \pm  0.001$ & $0.001$ \\
    \hline
-   EDMD & $2.313$ &$11.10$\\
+   EDMD & $1.156  \pm  0.003$ &$11.10$\\
    \hline
-   GPR & $1.441$ &$244.7$ \\
+   GPR & $1.025  \pm  0.403$ &$244.7$ \\
    \hline
-   TBATS & $1.501$ & $633.6$ \\
+   TBATS & $1.501 \pm  1.354$ & $633.6$ \\
    \hline

### 2020-08-04 Scripts/BoundEffRedExamples/ECGexample.m

@@ -2,7 +2,7 @@ clear all; close all; clc;
 addpath(genpath('../../TimeFrequencyScaleRep'));
 addpath('../../Algorithm/');
 
-%% load PPG signal
+%% load ECG signal
 addpath ../../Signals/
 data = load('ECG');
 
@@ -42,107 +42,4 @@ tic;
 xxGPR = SigExtension(x,fs,HOP,extK,extM,extSEC,method); 
 GPRtime = toc;
 
-fprintf(' __________________________________________\n')
-fprintf('| Extension Method | Computing time (sec.) |\n')
-fprintf('|------------------|-----------------------|\n')
-fprintf('|     SigExt       |        %.2f          |\n', LSEtime)
-fprintf('|  EDMD extension  |        %.2f          |\n', EDMDtime)
-fprintf('|  GPR extension   |       %.2f          |\n', GPRtime)
-fprintf('|__________________|_______________________|\n')
-
-
-%% Plot
-t = linspace(0, (N-1)/fs, N) ;
-tt = linspace(-L/fs, (N-1+L)/fs, N+2*L) ;
-
-xxTRUE = ( xtot((n0-L):(nf+L)) - mu ) / s ;
-xxZP = [ zeros(L,1); x; zeros(L,1) ] ; % zero padding
-
-
-figure;
-
-subplot(3,1,1)
-plot(tt,xxLSE,tt,xxTRUE,'--',t,x,'linewidth',2); grid on;
-ylabel('Signals'); axis tight; xlim([t(end)-1.5*extSEC tt(end)]);
-xticks([]);
-legend({'{\sf SigExt} extension','Ground truth extension','Original signal'},'location','northwest','interpreter','latex');
-set(gca,'fontsize',18);
-
-subplot(3,1,2)
-plot(tt,xxEDMD,tt,xxTRUE,'--',t,x,'linewidth',2); grid on;
-ylabel('Signals'); axis tight; xlim([t(end)-1.5*extSEC tt(end)]);
-xticks([]);
-legend({'EDMD extension','Ground truth extension','Original signal'},'location','northwest','interpreter','latex');
-set(gca,'fontsize',18);
-
-subplot(3,1,3)
[…]

### 2020-08-04 Scripts/BoundEffRedExamples/ForecastECG_TBATS.r

@@ -43,4 +43,4 @@ plot(tt,xx0,type="l", col="blue", lty=2)
 lines(tt,xxTBATS,type="l", col="red", lty=1)
 lines(t,x,type="l", col="blue", lty=1)
 
-write.csv(xextOpt,'../../Results/THO_TBATS.csv')
+write.csv(xextOpt,'../../Results/ECG_TBATS.csv')

### 2020-08-04 Scripts/EvalPerfSigExt/CompPerfAHM.m

@@ -42,7 +42,7 @@ x0 = xx0( (L+1) : (L+N) ) ; % restriction to the measurement interval
 sigman = 1e-2 ;
 
 %% Forecasting
-nbXP = 50 ;
+nbXP = 1000 ;
 
 for ind = 1:nbXP
     noise = sigman*randn(N+2*L,1) ;

### 2020-08-06 Notes/results.tex
code + paper

@@
 %
-Note that the MSE of {\sf SigExt} is higher than the MSE of EDMD or the symmetric extension.
-This is caused by the presence of few segments contaminated by artifacts so that it is unpredictable via a too simple dynamical model like~\eqref{eq:dyn.model}.
-The left of Fig.~\ref{fig:THO.failure} illustrates one of those cases, where {\sf SigExt} fails to catch the fast varying dynamic of the instantaneous amplitude to satisfactorily forecast the signal.
+Note that the MSE of {\sf SigExt} is in average higher than the MSE of EDMD or the symmetric extension, with a huge standard deviation.
+This variability is caused by the presence of few segments contaminated by artifacts so that it is unpredictable via a too simple dynamical model like~\eqref{eq:dyn.model}.
+The left of Fig.~\ref{fig:THO.failure} illustrates one of these outliers, where {\sf SigExt} fails to catch the fast varying dynamic of the instantaneous amplitude to satisfactorily forecast the signal.
 The EDMD and symmetric extensions are more robust to those situations, as shown in this example, and in Table~\ref{tab:THO}.
@@
 In Table~\ref{tab:THO}, we give the averaged performance index~\eqref{eq:index.perf} by evaluating the whole TF representations (including boundaries).
-Even though {\sf SigExt} performs moderately well in the sense of MSE, the boundary effects are dramatically reduced on the TF representations, and the performance is in the same order of that given by EDMD or GPR.
+Even though {\sf SigExt} performs moderately well in the sense of MSE, the boundary effects are dramatically reduced on the TF representations, and the averaged performance is in the same order of that given by EDMD or GPR.
+A t-test is performed to compare the performance index of {\sf SigExt} with those of other methods.
+Under the null hypothesis that the means are equal, with a $5\%$ significance level, the t-tests show no statistical significant difference between {\sf SigExt} and EDMD or GPR, regardless of the representation considered.
 
@@
   \hline
-   \multirow{2}{40pt}{\centering Extension method} & \multirow{2}{35pt}{\centering Averaged MSE} & \multicolumn{4}{c|}{Averaged performance index $D$} \\
+   \multirow{2}{38pt}{\centering Extension method} & \multirow{2}{43pt}{\centering MSE\\ (mean $\!\!\pm\!$ SD)} & \multicolumn{4}{c|}{Performance index $D$ (mean $\pm$ SD)} \\
    \cline{3-6}
@@
    \hhline{|=#=#=|=|=|=|}
-   {\sf SigExt} & $0.292$ & $0.370$ &  $0.408$ & $0.866$ & $0.423$ \\
+   \multirow{2}{38pt}{\centering {\sf SigExt}} & $0.292$ & $0.370$ &  $0.408$ & $0.866$ & $0.423$ \\
+   & $\pm 4.438$ & $\pm 0.623$ &  $\pm 0.436$ & $\pm 0.879$ & $\pm 0.344$ \\
    \hline
-   Symmetric & $0.044$ & $1.162$ & $1.173$ & $1.022$ & $1.144$ \\
+   \multirow{2}{38pt}{\centering Symmetric} & $0.044$ & $1.162$ & $1.173$ & $1.022$ & $1.144$ \\
+   & $\pm 0.111$ & $\pm 0.893$ &  $\pm 0.886$ & $\pm 0.281$ & $\pm 0.579$ \\
    \hline
-   EDMD & $0.026$ & $0.359$ &  $0.422$ & $0.828$ & $0.449$ \\
+   \multirow{2}{38pt}{\centering EDMD} & $0.026$ & $0.359$ &  $0.422$ & $0.828$ & $0.449$ \\
+   & $\pm 0.112$ & $\pm 0.266$ &  $\pm 0.282$ & $\pm 0.248$ & $\pm 0.296$ \\
    \hline
-   GPR & $0.331$ & $0.391$ &  $0.411$ & $0.897$ & $0.430$ \\ 
+   \multirow{2}{38pt}{\centering GPR} & $0.331$ & $0.391$ &  $0.411$ & $0.897$ & $0.430$ \\ 
+   & $\pm 4.858$ & $\pm 0.853$ &  $\pm 0.406$ & $\pm 1.140$ & $\pm 0.364$ \\
    \hline
@@
 For all the TF representations considered, the results show that {\sf SigExt} reduces boundary effects about as efficiently as the extensions given by EDMD or GPR.
+As for the respiratory signal, t-tests show no statistical significant difference between {\sf SigExt} and EDMD or GPR, regardless of the representation considered.
 For a visual inspection, the TF representation of SST resulting from the {\sf BoundEffRed} strategy is shown in the bottom-right panel of Fig.~\ref{fig:ex.intro}, where SST is applied to the portion of PPG displayed in Fig.~\ref{fig:ppg}.
@@
   \hline
-   \multirow{2}{40pt}{\centering Extension method} & \multicolumn{4}{c|}{Averaged performance index $D$} \\
+   \multirow{2}{40pt}{\centering Extension method} & \multicolumn{4}{c|}{Performance index $D$ (mean $\pm$ SD)} \\
    \cline{2-5}
@@
    \hhline{|=#=|=|=|=|}
-   {\sf SigExt} & $0.280$ & $0.309$ & $0.367$ & $0.534$ \\
+   \multirow{2}{40pt}{\centering {\sf SigExt}} & $0.280$ & $0.309$ & $0.367$ & $0.534$ \\
+    & $\pm 0.107$ & $\pm 0.112$ & $\pm 0.183$ & $\pm 0.160$ \\
    \hline
-   Symmetric & $1.168$ & $1.209$ & $1.310$ & $0.983$ \\
+   \multirow{2}{40pt}{\centering Symmetric} & $1.168$ & $1.209$ & $1.310$ & $0.983$ \\
+   & $\pm 0.390$ & $\pm 0.340$ & $\pm 0.140$ & $\pm 0.304$ \\
    \hline
-   EDMD & $0.289$ & $0.319$ & $0.375$ & $0.503$ \\
+   \multirow{2}{40pt}{\centering EDMD} & $0.289$ & $0.319$ & $0.375$ & $0.503$ \\
+   & $\pm 0.126$ & $\pm 0.134$ & $\pm 0.163$ & $\pm 0.163$ \\
    \hline
-   GPR & $0.276$ & $0.303$ & $0.361$ & $0.544$ \\
+   \multirow{2}{40pt}{\centering GPR} & $0.276$ & $0.303$ & $0.361$ & $0.544$ \\
+   & $\pm 0.106$ & $\pm 0.110$ & $\pm 0.165$ & $\pm 0.157$ \\
    \hline

### 2020-08-26 Notes/Supplementary material/SupMat.tex
new theorem

@@
 
-\section{Proof of Lemma~1}
-\label{ap:lm.error}
+\section{Proof of Theorem~1}
+\label{ap:th.error}
 
-%\subsection{Notations}
+\subsection{Preliminaries}
 Recall the model~\eqref{I-eq:model.noise}. Based on the definition of matrices $\bX$ and $\bY$, we have:
@@
 We gather all sources of randomness into an vector $\bg\in\RR^{M(M+1)}$, containing $M$ rows defined as
-\[
+\begin{equation}
 \bg = \vec\left(
@@
 \right)\,,
-\]
+\label{eq:vec}
+\end{equation}
 where "$\vec$" denotes the vectorization operator, concatenating the columns of a given matrix on top of one another.
@@
 f^{(\ell)} : \RR^{M(M+1)} &\to \RR^{M} \\
- \bg &\mapsto \bh^{(\ell)}\ . \\
+ \bg &\mapsto \bh^{(\ell)}\ .
 \end{align*}
 
-The multivariate version of the delta method is applicable to $\bh^{(\ell)}$ under the following conditions:
-\begin{enumerate}[(i)]
-\item\label{delta.1} $f^{(\ell)}$ is differentiable at the origin;
-\item\label{delta.2} $\sqrt{K}\ \bg \xrightarrow[K\to\infty]{\cD} \cN(\bzero,\bGamma_0)$ where $\bGamma_0$ is a covariance matrix.
-\end{enumerate}
-When these conditions are satisfied, the delta method gives the asymptotic behavior of $\bh^{(\ell)}=f^{(\ell)}(\bg)$, that is
+\subsubsection{Study of $\bg$}
+In the following, for all $r\in\{0,\ldots,M(M+1)-1\}$ we note $\bg[r] = \sigma\bg_1[r] + \sigma^2\bg_2[r]$, where 
+\begin{align*}
+\bg_1[r] &= \bE_1^{(a_r)}[m_r,m'_r] \\
+\bg_2[r] &= \bE_2^{(a_r)}[m_r,m'_r] \ ,
+\end{align*}
+and $m_r,m'_r, a_r$ are the corresponding coordinates of the matrices associated with $r$ through the vectorization operation~\eqref{eq:vec}.
+That is why, order-two moments of this random vector is split as follows:
 \begin{equation}
-\sqrt{K}\ \bh^{(\ell)} \xrightarrow[K\to\infty]{\cD} \cN(\bzero,{\bF^{(\ell)}}^T\bGamma_0\bF^{(\ell)})\ ,
-\label{eq:delta}
+\bE\{\bg[r]\bg[r']\} = \sigma^2\bE\{\bg_1[r]\bg_1[r']\} + \sigma^3\bE\{\bg_1[r]\bg_2[r']\} + \sigma^3\bE\{\bg_2[r]\bg_1[r']\} + \sigma^4\bE\{\bg_1[r]\bg_1[r']\}\ .
+\label{eq:gg}
 \end{equation}
-where $\bF^{(\ell)}$ is the Jacobian matrix of $f^{(\ell)}$ at the origin, that is
-\[
-\bF^{(\ell)}[m,m'] = \left.\dfrac{\partial f^{(\ell)}_m}{\partial\bg[m']}\right\vert_{\bg=\bzero}\ .
-\]
+Then,
+\begin{align}
+\nonumber
+\left|\bE\{\bg_1[r]\bg_1[r']\}\right| & = \bigg|\dfrac1{K^2}\EE\bigg\{\sum_{k,k'=0}^{K-1} \bz_k[m_r+a_r]\bz_{k'}[m_{r'}+a_{r'}]\bw_k[m'_r]\bw_{k'}[m'_{r'}] + \bz_k[m'_r]\bz_{k'}[m_{r'}+a_{r'}]\bw_k[m_r+a_r]\bw_{k'}[m'_{r'}] \\
+\nonumber
+&+ \bz_k[m_r+a_r]\bz_{k'}[m'_{r'}]\bw_k[m'_r]\bw_{k'}[m_{r'}+a_{r'}] + \bz_k[m'_r]\bz_{k'}[m'_{r'}]\bw_k[m_r+a_r]\bw_{k'}[m_{r'}+a_{r'}]\bigg\} \bigg| \\
+\nonumber
+& \leq \dfrac1{K^2}\left(\sum_{j=1}^J \Omega_j\right)^2 \sum_{k,k'=0}^{K-1} \EE\left\{\bw_k[m'_r]\bw_{k'}[m'_{r'}]\right\} + \EE\left\{\bw_k[m_r+a_r]\bw_{k'}[m'_{r'}]\right\} + \EE\left\{\bw_k[m'_r]\bw_{k'}[m_{r'}+a_{r'}]\right\}  \\
+\nonumber
+&\hspace{110pt}  + \EE\left\{\bw_k[m_r+a_r]\bw_{k'}[m_{r'}+a_{r'}]\right\} \\
+\nonumber
+&\leq \dfrac1{K^2}\left(\sum_{j=1}^J \Omega_j\right)^2 \sum_{k=0}^{K-1} 4 \\
+&\leq \dfrac{C_\bz^2}{K}\ ,
+\label{eq:g1g1}
+\end{align}
+where $C_\bz = 2\left(\sum_{j=1}^J \Omega_j\right)$. Besides, since odd-order moments of a zero-mean multivariate Gaussian random vector are zero, we have:
+\begin{align}
+\nonumber
+\bE\{\bg_1[r]\bg_2[r']\} & = \dfrac1{K^2}\sum_{k,k'=0}^{K-1} \bz_k[m_r+a_r]\EE\left\{\bw_k[m'_r]\bw_{k'}[m_{r'}+a_{r'}]\bw_{k'}[m'_{r'}]\right\} + \bz_k[m'_r]\EE\left\{\bw_k[m_r+a_r]\bw_{k'}[m_{r'}+a_{r'}]\bw_{k'}[m'_{r'}]\right\}\\
+\nonumber
+&\hspace{30pt} - \delta_{m_{r'}+a_{r'},m'_{r'}}\bE\{\bg_1[r]\} \\
+& = 0\ ,
+\label{eq:g1g2}
+\end{align}
+Similarly,
+\begin{equation}
+\bE\{\bg_2[r]\bg_1[r']\} = 0\ .
+\label{eq:g2g1}
+\end{equation}
+Besides,
+\begin{align}
+\nonumber
+\left|\bE\{\bg_2[r]\bg_2[r']\}\right| &= \dfrac1{K^2}\bigg|\sum_{k,k'=0}^{K-1} \EE\left\{\bw_{k}[m_{r}+a_{r}]\bw_{k}[m'_{r}]\bw_{k'}[m_{r'}+a_{r'}]\bw_{k'}[m'_{r'}]\right\} - \delta_{m_{r}+a_{r},m'_{r}}\bE\{\bg_2[r']\} - \delta_{m_{r'}+a_{r'},m'_{r'}}\bE\{\bg_2[r]\}\\
+\nonumber
+&\hspace{30pt} + \delta_{m_{r}+a_{r},m'_{r}}\delta_{m_{r'}+a_{r'},m'_{r'}} \bigg| \\
+\nonumber
+&= \dfrac1{K^2} \bigg| \sum_{k,k'=0}^{K-1} \EE\left\{\bw_{k}[m_{r}+a_{r}]\bw_{k}[m'_{r}]\right\}\EE\left\{\bw_{k'}[m_{r'}+a_{r'}]\bw_{k'}[m'_{r'}]\right\} - \delta_{m_{r}+a_{r},m'_{r}}\delta_{m_{r'}+a_{r'},m'_{r'}} \\
+\nonumber
+&\hspace{30pt}+ \dfrac1{K^2}\sum_{k,k'=0}^{K-1}\EE\left\{\bw_{k}[m_{r}+a_{r}]\bw_{k'}[m_{r'}+a_{r'}]\right\}\EE\left\{\bw_{k}[m'_{r}]\bw_{k'}[m'_{r'}]\right\} \\
+\nonumber
+&\hspace{30pt}+ \dfrac1{K^2}\sum_{k,k'=0}^{K-1} \EE\left\{\bw_{k}[m_{r}+a_{r}]\bw_{k'}[m'_{r'}]\right\}\EE\left\{\bw_{k}[m'_{r}]\bw_{k'}[m_{r'}+a_{r'}]\right\} \bigg| \\
+\nonumber
+& \leq \dfrac1{K^2} \sum_{k=0}^{K-1} 2 \\
+& \leq \dfrac2{K}\ .
+\label{eq:g2g2}
+\end{align}
+Thus, combining results~\eqref{eq:g1g1},~\eqref{eq:g1g2},~\eqref{eq:g2g1} and~\eqref{eq:g2g2} into expression~\eqref{eq:gg} gives the following bound:
+\begin{equation}
+\left|\bE\{\bg[r]\bg[r']\}\right| \leq \dfrac1K \left( C_\bz^2\sigma^2 + 2 \sigma^4\right)\ .
+\label{eq:gg.bound}
+\end{equation}
+In the same way, we show that higher-order moments behave as $o\left( \dfrac1K \right)$.
 
-Condition~\ref{delta.1} is satisfied. Indeed, the differentiation of $f^{(\ell)}$ can be checked as a composition of standard matrix derivation rules. {\color{red}Detail it}
+\subsubsection{Study of $f^{(\ell)}$}
+In this section, we provide bounds of the first and second-order derivatives of $f^{(\ell)}$ at the origin.
 
-Concerning condition~\ref{delta.2}, by definition of $\bg$, we have:
-\[
-\bg = \sigma\bg_1 + \sigma^2\bg_2\ ,
-\]
-where $\bg_1$ and $\bg_2$ are defined by
+Concerning the first-order derivative, from~\eqref{eq:error.vec}, we have:
 \begin{align*}
-\bg_1 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left( \tilde\bz_k\bw_k^T + \tilde\bw_k\bz_k^T \right)\quad\mbox{where}\quad \tilde\bz_k^T= \begin{pmatrix} \bz_k^T & \bz_{k+1}[M-1] \end{pmatrix}\,, \\
-\bg_2 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left(\tilde\bw_k\bw_k^T-\bI \right) \quad\mbox{where}\quad \tilde\bw_k^T= \begin{pmatrix} \bw_k^T & \bw_{k+1}[M-1] \end{pmatrix} \,.
+\dfrac{\partial f^{(\ell)}}{\partial \bg[r]} &= \be_M^T\dfrac{\partial \tilde\bA^{\ell}}{\partial \bg[r]} \\
+&= \sum_{\lambda=0}^{\ell-1} \be_M^T\tilde\bA^\lambda\dfrac{\partial \tilde\bA}{\partial \bg[r]}\tilde\bA^{\ell-1-\lambda}\ .
 \end{align*}
-First, $\bg_1$ is intrinsically a Gaussian random vector since it is a linear combination of Gaussian random vectors.
-Second, using the central limit theorem under weak dependence, we can show that $\bg_2$ also converges towards a Gaussian random vector as $K\to\infty$.
-Hence, we have
-\[
-\sqrt{K}\ \bg \xrightarrow[K\to\infty]{\cD} \cN(\bzero,\bGamma_0)\ ,
-\]
-where $\bGamma_0$ is the limit covariance matrix so that $\bGamma_0=\lim\limits_{K\to\infty} K\,\EE\{\bg\bg^T\}$.
-Hence, conditions~\ref{delta.1} and~\ref{delta.1} are satisfied.
-The delta method can therefore be applied, and result~\eqref{eq:delta} is valid.
+Thus,
+\begin{equation}
+\nonumber
+\left.\dfrac{\partial f^{(\ell)}}{\partial \bg[r]}\right|_{\bg=\bzero} = \sum_{\lambda=0}^{\ell-1} \be_M^T\bA_0^\lambda\left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero}\bA_0^{\ell-1-\lambda}\ .
+\label{eq:df}
+\end{equation}
+Furthermore,
+\begin{align*}
+\dfrac{\partial \tilde\bA}{\partial \bg[r]} &= \dfrac{\partial \bE^{(1)}}{\partial \bg[r]}\left( \bS^{(0)}+\bE^{(0)}\right)^{-1} +  \left(\bS^{(1)}+\bE^{(1)}\right)\dfrac{\partial \left( \bS^{(0)}+\bE^{(0)}\right)^{-1}}{\partial \bg[r]} \\
+&= \left\{
+\begin{array}{ll}
+\bJ_{m_r,m'_r}\left( \bS^{(0)}+\bE^{(0)}\right)^{-1} & \mbox{if}\quad a_r=1, \\
+\left( (1-\delta_{m_r,0})\bJ_{m_r-1,m'_r} +\left(\bS^{(1)}+\bE^{(1)}\right)\left( \bS^{(0)}+\bE^{(0)}\right)^{-1}\bJ_{m_r,m'_r} \right) \left( \bS^{(0)}+\bE^{(0)}\right)^{-1} & \mbox{else,}
+\end{array}
+\right.
+\end{align*}
+where $\bJ_{m_r,m'_r}\in\RR^{M\times M}$ is the matrix such that $\bJ_{m_r,m'_r}[m,m'] = \delta_{m,m_r}\delta_{m',m'_r}$.
+Thus,
+\begin{align*}
+\left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=0} &= \left\{
+\begin{array}{ll}
+\bJ_{m_r,m'_r}{\bS^{(0)}}^{-1} & \mbox{if}\quad a_r=1, \\
+\left((1-\delta_{m_r,0})\bJ_{m_r-1,m'_r} +\bA_0\bJ_{m_r,m'_r}\right){\bS^{(0)}}^{-1} & \mbox{else.}
+\end{array}
+\right.
+\end{align*}
+Let $\|\cdot\|_\infty$ denotes the uniform norm of a vector or a matrix, \ie $\|\bM\|_\infty = \max_{n,n'} \bM[n,m']$. Then,
+\begin{equation*}
+\left\| \left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=0}\right\|_\infty \leq \left(1+\|\bA_0\|_\infty\right)\|{\bS^{(0)}}^{-1}\|_\infty\ .
+\end{equation*}
+Given equation~\eqref{eq:A0.sine}, we have 
+\begin{equation}
+\|\bA_0\|_\infty\leq \max\left(1,\frac{2J}{M}\right)\ ;
+\label{eq:A0}
+\end{equation}
+and from~\eqref{eq:S0.inv}, we have that $\|{\bS^{(0)}}^{-1}\|_\infty\leq \dfrac1{\sigma^2}\left(1+\frac{2J}{M}\right)$. Then,
+\begin{equation}
+\left\| \left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=0}\right\|_\infty \leq \dfrac{C_{\bz,M}}{\sigma^2}\ ,
+\label{eq:dA}
+\end{equation}
+where $C_{\bz,M}=\left(1+\max\left(1,\frac{2J}{M}\right)\right)\left(1+\frac{2J}{M}\right)$. Besides given expression~\eqref{eq:df}, for all $r\in\{0,\ldots,M(M+1)-1\}$, we have:
+\begin{align*}
+\left|\left.\dfrac{\partial f_m^{(\ell)}}{\partial \bg[r]}\right\vert_{\bg=\bzero}\right| &\leq d_{1,M,\ell} \|\bA_0\|_\infty^{\ell-1}\left\|\dfrac{\partial\tilde\bA}{\partial \bg[r]}\right\|_\infty\ .
+\end{align*}
+where $d_{1,M,\ell}$ is only depending on $M$ and $\ell$.
+Therefore, given bounds~\eqref{eq:A0} and~\eqref{eq:dA}, we have:
+\begin{equation}
+\left|\left.\dfrac{\partial f_m^{(\ell)}}{\partial \bg[r]}\right\vert_{\bg=\bzero}\right| \leq \dfrac{d_{1,\bz,M,\ell}}{\sigma^2}\ ,
+\label{eq:df.bound}
+\end{equation}
+where $d_{1,\bz,M,\ell}=d_{1,M,\ell}\left(1+\frac{2J}{M}\right)\left(\max\left(1,\frac{2J}{M}\right)\right)^{\ell}$.
 
-\section{Proof of Theorem~1}
-\label{ap:th.error}
+Concerning the second-order derivative, we have:
+\begin{align*}
+\dfrac{\partial^2 f^{(\ell)}}{\partial \bg[r]\partial \bg[r']} &= \sum_{\lambda=0}^{\ell-1} \be_M^T\dfrac{\partial\tilde\bA^\lambda}{\partial \bg[r']}\dfrac{\partial \tilde\bA}{\partial \bg[r]}\tilde\bA^{\ell-1-\lambda} + \be_M^T\tilde\bA^\lambda\dfrac{\partial^2 \tilde\bA}{\partial \bg[r]\partial \bg[r']}\tilde\bA^{\ell-1-\lambda} + \be_M^T\tilde\bA^\lambda\dfrac{\partial \tilde\bA}{\partial \bg[r]}\dfrac{\partial\tilde\bA^{\ell-1-\lambda}}{\partial \bg[r']} \\
+&= \sum_{\lambda=1}^{\ell-1}\sum_{p=0}^{\lambda-1} \be_M^T\tilde\bA^p\dfrac{\partial\tilde\bA}{\partial \bg[r']}\tilde\bA^{\lambda-1-p}\dfrac{\partial \tilde\bA}{\partial \bg[r]}\tilde\bA^{\ell-1-\lambda} + \sum_{\lambda=0}^{\ell-1}\be_M^T\tilde\bA^l\dfrac{\partial^2 \tilde\bA}{\partial \bg[r]\partial \bg[r']}\tilde\bA^{\ell-1-\lambda} \\
+&+ \sum_{\lambda=0}^{\ell-2}\sum_{p=0}^{\ell-\lambda-2}\be_M^T\tilde\bA^\lambda\dfrac{\partial \tilde\bA}{\partial \bg[r]}\tilde\bA^p\dfrac{\partial\tilde\bA}{\partial \bg[r']}\tilde\bA^{\ell-\lambda-2-p}\ .
+\\
+\end{align*}
+Thus,
+\begin{align}
+\nonumber
+\left.\dfrac{\partial^2 f^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero} &= \sum_{\lambda=1}^{\ell-1}\sum_{p=0}^{\lambda-1} \be_M^T\bA_0^{p}\left.\dfrac{\partial\tilde\bA}{\partial \bg[r']}\right\vert_{\bg=\bzero}\bA_0^{\lambda-1-p}\left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right\vert_{\bg=\bzero}\bA_0^{\ell-1-\lambda} \\
+&+ \sum_{\lambda=0}^{\ell-1}\be_M^T\bA_0^{\lambda}\left.\dfrac{\partial^2 \tilde\bA}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero}\bA_0^{\ell-1-\lambda} + \sum_{\lambda=0}^{\ell-2}\sum_{p=0}^{\ell-2-\lambda}\be_M^T\bA_0^{\lambda}\left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right\vert_{\bg=\bzero}\bA_0^p\left.\dfrac{\partial\tilde\bA}{\partial \bg[r']}\right\vert_{\bg=\bzero}\bA_0^{\ell-\lambda-2-p}\ .
+\label{eq:d2f}
+\end{align}
+Besides,
+\begin{align*}
+\dfrac{\partial^2 \tilde\bA}{\partial \bg[r]\partial \bg[r']} &= \left\{
+\begin{array}{lc}
+0 &\mbox{if}\quad a_r=1\quad\mbox{and}\quad {a_r'}=1, \\
+\bJ_{m_{r},m'_{r}}\left( \bS^{(0)}+\bE^{(0)}\right)^{-1} \bJ_{m_{r'},m'_{r'}} \left( \bS^{(0)}+\bE^{(0)}\right)^{-1} & \mbox{if}\quad a_r=1\quad\mbox{and}\quad {a_r'}=0, \\
+\bJ_{m'_r,m'_{r'}}\left( \bS^{(0)}+\bE^{(0)}\right)^{-1}\bJ_{m_r,m'_r}\left( \bS^{(0)}+\bE^{(0)}\right)^{-1} & \mbox{if}\quad a_r=0\quad\mbox{and}\quad {a_r'}=1, \\
+\left( (1-\delta_{m_r,0})\bJ_{m_r-1,m'_r} +\left(\bS^{(1)}+\bE^{(1)}\right)\left( \bS^{(0)}+\bE^{(0)}\right)^{-1}\bJ_{m_r,m'_r} \right) & \\
+\hspace{20pt}\times \left( \bS^{(0)}+\bE^{(0)}\right)^{-1}\bJ_{m_{r'},m'_{r'}}\left( \bS^{(0)}+\bE^{(0)}\right)^{-1} & \\
++ \left( (1-\delta_{m_{r'},0})\bJ_{m_{r'}-1,m'_{r'}} + \left(\bS^{(1)}+\bE^{(1)}\right)\left( \bS^{(0)}+\bE^{(0)}\right)^{-1}\bJ_{m_{r'},m'_{r'}} \right) & \\
+\hspace{20pt}\times \left( \bS^{(0)}+\bE^{(0)}\right)^{-1}\bJ_{m_{r},m'_{r}}\left( \bS^{(0)}+\bE^{(0)}\right)^{-1}  & \mbox{else.}
+\end{array} 
+\right.
+\end{align*}
+Thus,
+\begin{align*}
+\left.\dfrac{\partial^2 \tilde\bA}{\partial \bg[r]\partial \bg[r']}\right|_{\bg=\bzero} &= \left\{
+\begin{array}{ll}
+0 &\mbox{if}\quad a_r=1\quad\mbox{and}\quad a_{r'}=1, \\
+\bJ_{m_r,m'_r}{\bS^{(0)}}^{-1} \bJ_{m_{r'},m'_{r'}} { \bS^{(0)}}^{-1} & \mbox{if}\quad a_r=1\quad\mbox{and}\quad a_{r'}=0, \\
+\bJ_{m_{r'},m'_{r'}}{\bS^{(0)}}^{-1} \bJ_{m_{r},m'_{r}} { \bS^{(0)}}^{-1} & \mbox{if}\quad a_r=0\quad\mbox{and}\quad a_{r'}=1, \\
+\left( (1-\delta_{m_r,0})\bJ_{m_r-1,m'_r} +\bA_0\bJ_{m_r,m'_r} \right){\bS^{(0)}}^{-1}\bJ_{m_{r'},m'_{r'}}{ \bS^{(0)}}^{-1} & \\
++ \left( (1-\delta_{m_{r'},0})\bJ_{m_{r'}-1,m'_{r'}} +\bA_0\bJ_{m_{r'},m'_{r'}} \right){\bS^{(0)}}^{-1}\bJ_{m_{r},m'_{r}}{ \bS^{(0)}}^{-1} & \mbox{else.}
+\end{array} 
+\right.
+\end{align*}
+Then,
+\begin{align}
+\nonumber
+\left\| \left.\dfrac{\partial^2 \tilde\bA}{\partial \bg[r]\partial \bg[r']}\right|_{\bg=0}\right\|_\infty &\leq \left\{
+\begin{array}{ll}
+\|{\bS^{(0)}}^{-1}\|_\infty^2 &\mbox{if}\quad a_r=1\quad\mbox{or}\quad a_{r'}=1 \\
+2\left(1+\|\bA_0\|_\infty\right)\|{\bS^{(0)}}^{-1}\|_\infty^2 &\mbox{else} \\
+\end{array}
+\right.\\
+&\leq 2\left(1+\max\left(1,\frac{2J}{M}\right)\right)\left(1+\frac{2J}{M}\right)^2\dfrac1{\sigma^4}\ .
+\label{eq:d2A}
+\end{align}
+Returning to equation~\eqref{eq:d2f}, for all $r,r'\in\{0,\ldots,M(M+1)-1\}$ and $m\in\{0,\ldots, M-1\}$, we have:
+\begin{align*}
+\left|\left.\dfrac{\partial^2 f_m^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero}\right| &\leq d_{2,M,\ell} \|\bA_0\|_\infty^{\ell-2}\left\|\dfrac{\partial\tilde\bA}{\partial \bg[r]}\right\|_\infty\left\|\dfrac{\partial\tilde\bA}{\partial \bg[r']}\right\|_\infty + d'_{2,M,\ell}\|\bA_0\|_\infty^{\ell-1}\left\|\dfrac{\partial^2\tilde\bA}{\partial \bg[r]\bg[r']}\right\|_\infty \ ,
+\end{align*}
+where $d_{2,M,\ell}$ and $d'_{2,M,\ell}$ are only depending on $M$ and $\ell$.
+Besides, given results~\eqref{eq:dA} and~\eqref{eq:d2A}, we have:
+\begin{equation}
+\left|\left.\dfrac{\partial^2 f_m^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero}\right| \leq \dfrac{d_{2,\bz,M,\ell}}{\sigma^4}\ ,
+\label{eq:d2f.bound}
+\end{equation}
+where 
+\begin{align*}
+d_{2,\bz,M,\ell} &= d_{2,M,\ell}C_{\bz,M}^2\max\left(1,\frac{2J}{M}\right)^{\ell-2} + d'_{2,M,\ell}\max\left(1,\frac{2J}{M}\right)^{\ell-1}2\left(1+\max\left(1,\frac{2J}{M}\right)\right)\left(1+\frac{2J}{M}\right)^2\ .
+\end{align*}
 
 \subsection{Expression of the Bias $\bmu$.}
+
 By definition of the measurement noise, $\bmu[n]=0$ when $n\in I$.
@@
 \bmu[n] &=  \EE\{\balpha^{(\ell)}\}\bz_{K} + \sigma\EE\{\balpha^{(\ell)}\bw_{K}\} - \bz[n]\\
-& = \balpha_0^{(\ell)}\bz_{K} + \EE\{\bh^{(\ell)}\}\bz_{K} + \sigma\EE\{\bh^{(\ell)}\bw_{K}\} - \bz[N-1+\ell]\ .
+\nonumber
+& = \balpha_0^{(\ell)}\bz_{K} + \EE\{\bh^{(\ell)}\}\bz_{K} + \sigma\EE\{\bh^{(\ell)}\bw_{K}\} - \bz[N-1+\ell] \\
+&\defeq \bepsilon_1[\ell] + \bepsilon_2[\ell] + \bepsilon_3[\ell]\ ,
 \label{eq:bias.dec}
 \end{align}
-Let us first evaluate the expression of $\balpha_0^{(\ell)}\bz_K$. We have
+
+where
+\begin{align}
+\label{eq:eps1.def}
+\bepsilon_1[\ell]&=  \balpha_0^{(\ell)}\bz_{K}  - \bz[N-1+\ell] \\
+\label{eq:eps2.def}
+\bepsilon_2[\ell] &=  \EE\{\bh^{(\ell)}\}\bz_{K}  \\
+\label{eq:eps3.def}
+\bepsilon_3[\ell] &=  \sigma\EE\{\bh^{(\ell)}\bw_{K}\}  \ .
+\end{align}
+
+Let us first determine an upper bound on $\left|\bepsilon_1[n]\right|$. We have
 \begin{align}
@@
 \end{align}
-Thus, $\bS^{(0)}$ is a circulant matrix and is therefore diagonalizable in the Fourier basis:
+Thus, $\bS^{(0)}$ is a circulant matrix, and is therefore diagonalizable in the Fourier basis:
 \[
@@
 \begin{equation}
-{\bS^{(0)}}^\inv[m,m']  = \dfrac1{\sigma^2}\delta_{m,m'}-\sum_{j=1}^J\dfrac{\Omega_j^2}{2\sigma^2(\sigma^2+\Omega_j^2M/4)}\cos\left(2\pi p_j \dfrac{m-m'}{M}\right)\,.
+{\bS^{(0)}}^\inv[m,m']  = \dfrac1{\sigma^2}\delta_{m,m'}-\dfrac{2}{M\sigma^2}\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j \dfrac{m-m'}{M}\right)\,.
 \label{eq:S0.inv}
@@
 \bA_0[m,m']  &= \sum_{q=0}^{M-1} {\bS^{(1)}}[m,q]{\bS^{(0)}}^\inv[q,m'] \\
-&= \delta_{m+1,m'} + \sum_{j=1}^J\dfrac{2\Omega_j^2}{\Omega_j^2M+4\sigma^2}\cos\left(2\pi p_j\frac{m'}{M}\right)\delta_{m+1,M}\,.
+&= \delta_{m+1,m'} + \dfrac2M\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j\frac{m'}{M}\right)\delta_{m+1,M}\,.
 \label{eq:A0.sine}
 \end{align}
-Thus $\balpha_0^{(1)}$, the last row of $\bA_0$, is written as
-\begin{align*}
-\balpha_0^{(1)}[m]  &=\sum_{j=1}^J\dfrac{2\Omega_j^2}{\Omega_j^2M+4\sigma^2}\cos\left(2\pi p_j\frac{m}{M}\right) \\
-&= \dfrac{2}{M}\sum_{j=1}^J\cos\left(2\pi p_j\frac{m}{M}\right) + o(\sigma)\ .
-\end{align*}
-Besides, from equation~\eqref{eq:A0.sine}, we have
+Since $\balpha_0^{(1)}$ is the last row of $\bA_0$, we have
+\begin{equation}
+\balpha_0^{(1)}[m]  =\dfrac2M \sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j\frac{m}{M}\right) \ . 
+\label{eq:alpha01}
+\end{equation}
+Besides, from equation~\eqref{eq:A0.sine}, we also have
 \begin{align*}
@@
 \end{align*}
-By induction, we have
+The upward-shift property is thus successively inducted when $\ell$ increases; that is,
 \begin{align*}
-\tilde\bA_0^{\ell}\bz_{K} &= 
+\bA_0^{\ell}\bz_{K} &= 
 \begin{pmatrix}
@@
 \end{align*}
+Then, $\balpha_0^{(\ell)}$, the last row of $\bA_0^\ell$ follows the following recurrence relation:
+\begin{align*}
+\balpha_0^{(\ell)}\bz_{K} &= \balpha_0^{(1)}\tilde\bA_0^{\ell-1}\bz_{K} \\
+&=\sum_{m=0}^{M-\ell}\balpha_0^{(1)}[m]\bz[N-M+\ell+m-1]+\sum_{m=M-\ell+1}^{M-1}\balpha_0^{(1)}[m]\balpha_0^{(m-M+\ell)}\bz_K\,.
+\end{align*}
+Hence,
+\begin{align*}
+\bepsilon_1[\ell] &= \balpha_0^{(\ell)}\bz_{K} - \bz[N-1+\ell] \\
+&=\sum_{m=0}^{M}\balpha_0^{(1)}[m]\bz[N-M+\ell+m-1]  - \bz[N-1+\ell] +\sum_{m=M-\ell+1}^{M-1}\balpha_0^{(1)}[m]\left(\balpha_0^{(m-M+\ell)}\bz_K - \bz[N-M+\ell+m-1]\right) \\
+&=\sum_{m=0}^{M}\balpha_0^{(1)}[m]\bz[N-M+\ell+m-1]  - \bz[N-1+\ell] +\sum_{m=M-\ell+1}^{M-1}\balpha_0^{(1)}[m]\bepsilon_1[m-M+\ell]\,.
+\end{align*}
+Besides, equation~\eqref{eq:alpha01} gives
+\begin{align*}
+\sum_{m=0}^{M}\balpha_0^{(1)}[m]\bz[N-M+\ell+m-1] &=\sum_{j,j'=1}^J\Omega_{j'}\dfrac{2}{M}\dfrac1{1+\frac{4\sigma^2}{M\Omega_j^2}}\underbrace{\sum_{m=0}^{M-1}\cos\left(2\pi p_j\frac{m}{M}\right)\cos\left(2\pi p_{j'}\dfrac{N+m}{M}+\varphi_{j'}\right)}_{=\delta_{j,j'}\frac{M}2\cos\left(2\pi p_j\frac{N}M +\varphi_{j}\right)} \\
+&= \sum_{j=1}^J\Omega_j\dfrac1{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j\dfrac{N}{M}+\varphi_j\right)\ .
+\end{align*}
+Thus:
+\begin{align}
+\nonumber
+\left|\bepsilon_1[\ell]\right| &= \left|\sum_{j=1}^J\Omega_j\left(\dfrac1{1+\frac{4\sigma^2}{M\Omega_j^2}}-1\right)\cos\left(2\pi p_j\dfrac{N}{M}+\varphi_j\right) +\sum_{m=M-\ell+1}^{M-1}\balpha_0^{(1)}[m]\bepsilon_1[m-M+\ell]\right| \\
+\nonumber
+&\leq \sum_{j=1}^J\Omega_j\left|\dfrac1{1+\frac{4\sigma^2}{M\Omega_j^2}}-1\right| +\dfrac{2J}{M}\sum_{\lambda=1}^{\ell-1}\left|\bepsilon_1[\lambda]\right| \\
+&\leq \dfrac{4\sigma^2}{M}\sum_{j=1}^J\dfrac1{\Omega_j} +\dfrac{2J}{M}\sum_{\lambda=1}^{\ell-1}\left|\bepsilon_1[\lambda]\right|\ .
+\label{eq:seq}
+\end{align}
+Then, by induction from the inequality~\eqref{eq:seq}, we have
+\begin{equation}
+\left|\bepsilon_1[\ell]\right| \leq \dfrac{4\sigma^2}{M}\left(1+\frac{2J}{M}\right)^{\ell-1}\sum_{j=1}^J\dfrac1{\Omega_j} \defeq c^{(1)}\,\sigma^2\ ,
+\label{eq:eps1}
+\end{equation}
+where $c^{(1)}=\dfrac{4}{M}\left(1+\frac{2J}{M}\right)^{\ell-1}\sum_{j=1}^J\dfrac1{\Omega_j}$. Note that $c^{(1)}$ is not depending on $\sigma$ or $K$.
+
+Let us now determine an upper bound on $\left|\bepsilon_2[\ell]\right|$. Since $\EE\{\bg[r]\}=0$, a second-order Taylor expansion of $\bh^{(\ell)}$ gives
+\begin{equation}
+\EE\{\bh^{(\ell)}[m]\} = \dfrac12\sum_{r,r'=0}^{M(M+1)-1}\left.\dfrac{\partial^2 f_m^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero} \EE\{\bg[r]\bg[r']\} + o\left(\dfrac1K\right)\ .
+\label{eq:Eh}
+\end{equation}
+Thus, given the bounds~\eqref{eq:gg.bound} on $\EE\{\bg[r]\bg[r']\}$ and~\eqref{eq:d2f.bound} on the second derivative of $f_m^{(\ell)}$, we have:
+\begin{align}
+\nonumber
+\left|\bepsilon_2[\ell]\right| &\leq C_\bz\dfrac{M^3(M+1)^2}{4}\left\|\left.\dfrac{\partial^2 f^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero}\right\|_\infty \dfrac1K \left( C_\bz^2\sigma^2 + 2 \sigma^4\right) + o\left(\dfrac1K\right)  \\
+&\leq \dfrac1K \left(c^{(2)}_1 + \dfrac{c^{(2)}_2}{\sigma^2} \right) + o\left(\dfrac1K\right)\ ,
+\label{eq:eps2}
+\end{align}
+where
+\begin{align*}
+c^{(2)}_1 &=\dfrac{M^3(M+1)^2}{2}C_\bz d_{2,\bz,M,\ell}\ ,\\
+c^{(2)}_2 &=\dfrac{M^3(M+1)^2}{4}C_\bz^3 d_{2,\bz,M,\ell} \ .
+\end{align*}
+
+Let us now determine an upper bound on $\left|\bepsilon_3[\ell]\right|$. 
+A second-order Taylor expansion of $\bh^{(\ell)}$ gives
+\begin{equation}
+\EE\{\bh^{(\ell)}\bw_K\}  = \sum_{m=0}^{M-1} \sum_{r=0}^{M(M+1)-1}\left.\dfrac{\partial f_m^{(\ell)}}{\partial \bg[r]}\right\vert_{\bg=\bzero} \EE\{\bg[r]\bw_K[m]\} + o\left(\dfrac1K\right)
+\label{eq:Ehw}
+\end{equation}
+Indeed, the third-order moments $\EE\{\bg[r]\bg[r']\bw_K[m]\}$, behave as $o\left(\dfrac1K\right)$. Besides,
+\begin{equation}
+\EE\{\bg[r]\bw_K[m]\}= \sigma\EE\{\bg_1[r]\bw_K[m]\} + \sigma^2\EE\{\bg_2[r]\bw_K[m]\}\ .
+\label{eq:gw}
+\end{equation}
 Then,
@@
 \nonumber
-\balpha_0^{(\ell)}\bz_{K} &= \balpha_0^{(1)}\tilde\bA_0^{\ell-1}\bz_{K} \\
-&=\sum_{m=0}^{M-\ell}\balpha_0^{(1)}[m]\bz[N-M+\ell+m-1]+\sum_{m=M-\ell+1}^{M-1}\balpha_0^{(1)}[m]\balpha_0^{(m-M+\ell)}\bz_K\,.
-\label{eq:seq}
+\left|\bE\{\bg_1[r]\bw_K[m]\}\right| & = \dfrac1{K}\left|\sum_{k=0}^{K-1} \bz_k[m_r+a_r]\EE\left\{\bw_k[m'_r]\bw_{K}[m]\right\} + \bz_k[m'_r]\EE\left\{\bw_k[m_r+a_r]\bw_{K}[m]\right\}\right| \\
+& \leq \dfrac2{K}\left(\sum_{j=1}^J \Omega_j\right) = \dfrac{C_\bz}{K}\ .
+\label{eq:g1w}
 \end{align}
+\begin{align}
+\nonumber
+\bE\{\bg_2[r]\bw_K[m]\} & = \dfrac1{K}\sum_{k,=0}^{K-1} \EE\left\{\bw_{k}[m_{r}+a_{r}]\bw_{k}[m'_{r}]\bw_K[m]\right\} -\delta_{m_r+a_r,m'_r}\EE\left\{\bw_K[m]\right\}\\
+& = 0\ .
+\label{eq:g2w}
+\end{align}
+Thus, combining results~\eqref{eq:g1w} and~\eqref{eq:g2w} into expression~\eqref{eq:gw} gives the following bound:
+\begin{equation}
+\left|\bE\{\bg[r]\bw_K[m]\}\right| \leq C_\bz\dfrac{\sigma}K\ .
+\label{eq:gw.bound}
+\end{equation}
+Thus, given the bound~\eqref{eq:df.bound} on the first derivative of $f_m^{(\ell)}$, and from~\eqref{eq:Ehw} we have:
+\begin{equation}
+\left|\bepsilon_3[\ell]\right| \leq \sigma M^2(M+1) \dfrac{d_{1,\bz,M,\ell}}{\sigma^2} C_\bz \dfrac{\sigma}{K} \defeq \dfrac{c^{(3)}}{{K}}\ ,
+\label{eq:eps3}
+\end{equation}
+where $ c^{(3)} = M^2(M+1)d_{1,\bz,M,\ell}C_\bz$.
+
+Thus, combining bounds~\eqref{eq:eps1} on $\bepsilon_1$, ~\eqref{eq:eps2} on $\bepsilon_2$ and ~\eqref{eq:eps3} on $\bepsilon_3$ gives the following bound of the bias:
+\begin{equation}
+\left\vert\bmu[n]\right\vert \leq  c^{(1)}\sigma^2 + \dfrac1K \left( \dfrac{c_2^{(2)}}{\sigma^2}+c_1^{(2)}+c^{(3)} \right)  + o\left(\dfrac1K\right)\ .
+\end{equation}
+
+%\begin{align*}
+%\bh^{(\ell)}[m] &= f_m^{(\ell)}(\bg) \\
+%& = f_m^{(\ell)}(\bzero) + \sum_{r=0}^{M(M-1)}\left.\dfrac{\partial f_m^{(\ell)}}{\partial \bg[r]}\right\vert_{\bg=\bzero} \bg[r] + \dfrac12\sum_{r,r'=0}^{M(M-1)}\left.\dfrac{\partial^2 f_m^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero} \bg[r]\bg[r'] + o\left(\dfrac1K\right)\ .
+%\end{align*}
+
+\subsection{Expression of the Covariance $\bgamma$.}
+
+Outside the measurement interval $I$, denote by $\ell$ the index such that $n=N-1+\ell$. Then, given that $\bh^{(\ell)} = \balpha^{(\ell)} - \balpha_0^{(\ell)}$, we have
+\begin{align*}
+\bgamma[n,n] &= \left( \balpha_0^{(\ell)}\bz_K\right)^2 + 2\left(\balpha_0^{(\ell)}\bz_K\right)\EE\left\{\bh^{(\ell)}\right\}\bz_K + \EE\left\{\left(\bh^{(\ell)}\bz_K\right)^2\right\} + 2\sigma\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\ell)}\}\bz_K + 2\sigma\balpha_0^{(\ell)}\bz_K \EE\{\bh^{(\ell)}\bw_K\} \\
+& + 2\sigma\EE\{\bh^{(\ell)}\bw_K\bh^{(\ell)}\}\bz_K +\sigma^2\left\|\balpha_0^{(\ell)}\right\|^2 +\sigma^2\EE\left\{\left(\bh^{(\ell)}\bw_K\right)^2\right\}  + 2\sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\ell)}\bw_K\} -\bz[n]^2 - 2\bz[n]\bmu[n]- \bmu[n]^2 \\
+&= \sigma^2\left\|\balpha_0^{(\ell)}\right\|^2-\left( \bz[n] - \balpha_0^{(\ell)}\bz_K+\bz[n]\right)^2 + \EE\left\{\left(\bh^{(\ell)}\bz_K\right)^2\right\} + 2\sigma\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\ell)}\}\bz_K+ 2\sigma\EE\{\bh^{(\ell)}\bw_K\bh^{(\ell)}\}\bz_K \\
+&+\sigma^2\EE\left\{\left(\bh^{(\ell)}\bw_K\right)^2\right\}  + 2\sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\ell)}\bw_K\}\\
+&\defeq \eta_1[n] + \eta_2[n] + \eta_3[n] + \eta_4[n] + \eta_5[n] + \eta_6[n] + \eta_7[n]\ , 
+\end{align*}
+where
+\begin{align*}
+\eta_1[n] &= \sigma^2\left\|\balpha_0^{(\ell)}\right\|^2 \,, & \eta_2[n] &= -\left( \bz[n] - \balpha_0^{(\ell)}\bz_K+\bz[n]\right)^2 \,,
+& \eta_3[n] &= \EE\left\{\left(\bh^{(\ell)}\bz_K\right)^2\right\}\,, \\
+\eta_4[n] &= 2\sigma\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\ell)}\}\bz_K\,,
+& \eta_5[n] &= \sigma\EE\{\bh^{(\ell)}\bw_K\bh^{(\ell)}\}\bz_K\,,
+& \eta_6[n] &= \sigma^2\EE\left\{\left(\bh^{(\ell)}\bw_K\right)^2\right\}\,, \\
+\eta_7[n] &= 2\sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\ell)}\bw_K\} \ .\\ 
+\end{align*}
+Let us now determine an upper bound on each of these terms.
+
+First, since $\|\balpha_0^{(\ell}\|^2\leq M\|\balpha_0^{(\ell)}\|_\infty^2\leq M\|\bA_0^{\ell}\|_\infty^2$, we have:
+\begin{align}
+\nonumber
+\eta_1[n] &\leq \sigma^2M\left\|\bA_0^{\ell}\right\|_\infty^2 \\
+\nonumber
+&\leq\sigma^2M^{\ell}\left\|\bA_0\right\|^{2\ell}_\infty \\
+&\leq \sigma^2 M^{\ell}\max\left(1,\left(\frac{2J}{M}\right)^{2\ell}\right)\ .
+\label{eq:sum1}
+\end{align}
+
+Second, by definition of $\epsilon_2$ and $\epsilon_3$ (see expressions~\eqref{eq:eps2.def} and~\eqref{eq:eps3.def}), $\eta_2$ takes the following form:
+\begin{equation}
+\eta_2[n] = \left(\bepsilon_2[n]+\bepsilon_3[n]\right)^2 = o\left(\dfrac1K\right)\ .
+\label{eq:sum2}
+\end{equation}
+
+Third, second-order Taylor expansions of $\bh^{(\ell)}$ give:
+\begin{align}
+\nonumber
+\eta_3[n] &\leq
+\dfrac{C_\bz^2}{4}\sum_{m,m'=0}^{M-1}\left|\EE\left\{\bh^{(\ell)}[m]\bh^{(\ell)}[m']\right\}\right| \\
+\nonumber
+&\leq \dfrac{C_\bz^2}{4} \sum_{r,r'=0}^{M(M+1)-1}\left|\left.\dfrac{\partial f_m^{(\ell)}}{\partial \bg[r]}\right\vert_{\bg=\bzero}\left.\dfrac{\partial f_{m'}^{(\ell)}}{\partial \bg[r']}\right\vert_{\bg=\bzero} \EE\{\bg[r]\bg[r']\}\right| + o\left(\dfrac1K\right) \\
+\nonumber
+&\leq \dfrac{C_\bz^2 M^2(M+1)^2}{4}  \dfrac{d_{1,\bz,M,\ell}^2}{\sigma^4}\dfrac1K \left( C_\bz^2\sigma^2 + 2 \sigma^4\right) + o\left(\dfrac1K\right) \\
+&\leq \dfrac{C_\bz^2 M^2(M+1)^2d_{1,\bz,M,\ell}^2}{4K} \left( \dfrac{C_\bz^2}{\sigma^2} + 2\right)  + o\left(\dfrac1K\right)\ .
+\label{eq:sum3}
+\end{align}
+
+Fourth,
+\begin{align*}
+\left|\eta_4[n]\right| & \leq \sigma C_\bz\|\balpha_0^{(\ell)}\|_\infty \sum_{m,m'=0}^{M-1}\left|\EE\{\bh^{(\ell)}[m]\bw_K[m']\} \right|
+\end{align*}
+But,
+\begin{align}
+\nonumber
+\left|\EE\{\bh^{(\ell)}[m]\bw_K[m']\} \right|  &\leq \sum_{r=0}^{M(M+1)-1}\left|\left.\dfrac{\partial f_m^{(\ell)}}{\partial \bg[r]}\right\vert_{\bg=\bzero}\right| \left|\EE\{\bg[r]\bw_K[m']\}\right| + o\left(\dfrac1K\right) \\
+&\leq M(M+1)\dfrac{d_{1,\bz,M,\ell}}{\sigma^2} C_\bz\dfrac{\sigma}{K} + o\left(\dfrac1K\right)
+\label{eq:Ehwm}
+\end{align}
+Thus,
+\begin{equation}
+\left|\eta_4[n]\right| \leq M^\ell(M+1)C_\bz^2 d_{1,\bz,M,\ell}\left(\max\left(1,\frac{2J}{M}\right)\right)^\ell\dfrac{1}{K} + o\left(\dfrac1K\right)\ .
+\label{eq:sum4}
+\end{equation}
+
+Fifth and sixth, second-order Taylor expansions of $\bh^{(\ell)}$ give
+\begin{align}
+\label{eq:sum5}
+\eta_5[n] &= o\left(\dfrac1K\right) \\
+\label{eq:sum6}
+\eta_6[n] &= o\left(\dfrac1K\right)\ .
+\end{align}
+
+Seventh,
+\begin{align*}
+\left|\eta_7[n] \right| &\leq 2\sigma^2\|\balpha_0^{(\ell)}\|_\infty \sum_{m,m'=0}^{M-1} \left| \EE\{\bh^{(\ell)}[m]\bw_K[m]\bw_K[m']\} \right| \\
+&\leq 2\sigma^2\|\balpha_0^{(\ell)}\|_\infty \sum_{m,m'=0}^{M-1} \sum_{r=0}^{M(M+1)-1}\left|\left.\dfrac{\partial f_m^{(\ell)}}{\partial \bg[r]}\right\vert_{\bg=\bzero}\right| \left| \EE\{\bg^{(\ell)}[r]\bw_K[m]\bw_K[m']\} \right| + o\left( \dfrac1K\right)\ .
+\end{align*}
 But,
 \begin{align*}
-\balpha_0^{(1)}\bz_{K} &=\sum_{m=0}^{M-1}\balpha_0^{(1)}[m]\bz[N-M+m] \\
-&=\sum_{j,j'=1}^J\Omega_{j'}\dfrac{2}{M}\underbrace{\sum_{m=0}^{M-1}\cos\left(2\pi p_j\frac{m}{M}\right)\cos\left(2\pi p_{j'}\dfrac{N+m}{M}+\varphi_{j'}\right)}_{=\delta_{j,j'}\frac{M}2\cos\left(2\pi p_j\frac{N}M +\varphi_{j}\right)} + o(\sigma) \\
-&= \sum_{j=1}^J\Omega_j\cos\left(2\pi p_j\dfrac{N}{M}+\varphi_j\right) + o(\sigma) \\
-&= \bz[N] + o(\sigma)
+\EE\{\bg^{(\ell)}[r]\bw_K[m]\bw_K[m']\} = \sigma\EE\{\bg_1^{(\ell)}[r]\bw_K[m]\bw_K[m']\} + \sigma^2\EE\{\bg_2^{(\ell)}[r]\bw_K[m]\bw_K[m']\}\ .
 \end{align*}
-By induction from~\eqref{eq:seq}, we have
-\begin{equation}
-\label{eq:alpha0z.sine}
-\balpha_0^{(\ell)}\bz_{K} = \bz[N-1+\ell] + o(\sigma)\ .
-\end{equation}
-Then,  inserting result~\eqref{eq:alpha0z.sine} into equation~\eqref{eq:bias.dec} gives
-\begin{equation*}
-\bmu[N-1+\ell] = \EE\{\bh^{(\ell)}\}\bz_{K} + \sigma\EE\{\bh^{(\ell)}\bw_{K}\} + o(\sigma)\ .
-\end{equation*}
-Besides, from Lemma~1, we have the following result:
-\begin{equation*}
-\EE\{\bh^{(\ell)}\} \underset{K\to\infty}{\longrightarrow} 0\ .
-\end{equation*}
-{\color{red}[if it is possible to get the convergence rate, like $\EE\{\bh^{(\ell)}\}=O(1/K)$, that would be great!]}
-Moreover, applying delta method to $\left(\bg\ \bw[n']\right)^T$, $\forall n'\in I$, one can show that we have the asymptotic result
-\begin{equation}
-\EE\{\bw[n']\bh^{(\ell)}\} \underset{K\to\infty}{\longrightarrow} 0\,.
-\label{eq:cov.ap}
-\end{equation}
-Consequently, $\EE\{\bh^{(\ell)}\bw_{K}\} \underset{K\to\infty}{\longrightarrow} 0$, and
-\begin{equation}
-\frac{1}{\sigma}\bmu[N-1+\ell] = o(1)
-\label{eq:bias.ap}
-\end{equation}
-when $K\to\infty$.
+As before, since $\EE\{\bg_1^{(\ell)}[r]\bw_K[m]\bw_K[m']\}$ is a third-order moment of a multivariate zero-mean Gaussian vector, it vanishes. And,
+\begin{align*}
+\left|\EE\{\bg_2^{(\ell)}[r]\bw_K[m]\bw_K[m']\}\right| &= \dfrac1K\sum_{k,=0}^{K-1} \EE\left\{\bw_{k}[m_{r}+a_{r}]\bw_{k}[m'_{r}]\bw_K[m]\bw_K[m']\right\} - \delta_{m_r+a_r,m'_r} \EE\left\{\bw_K[m]\bw_K[m']\right\} \\
+&\leq \dfrac4K\ .
+\end{align*}
+Thus,
+\begin{align}
+\nonumber
+\left| \eta_7[n] \right| &\leq 2\sigma^2\|\balpha_0^{(\ell)}\|_\infty M^3(M+1)\dfrac{d_{1,\bz,M,\ell}}{\sigma^2} \dfrac{4\sigma^2}K + o\left( \dfrac1K\right)\\
+&\leq 8 M^{\ell+2}(M+1)d_{1,\bz,M,\ell}\left(\max\left(1,\frac{2J}{M}\right)\right)^\ell\dfrac{\sigma^2}{K} + o\left( \dfrac1K\right)\ .
+\label{eq:sum7}
+\end{align}
 
-\subsection{Expression of the Covariance $\bgamma$.}
-Take $n>N$, and denote $n=N-1+\ell$.  To derive the covariance, let us segregate the cases.
+To conclude, we combine the expressions~\eqref{eq:sum1},~\eqref{eq:sum2},~\eqref{eq:sum3},~\eqref{eq:sum4},~\eqref{eq:sum5},~\eqref{eq:sum6}, and~\eqref{eq:sum7} to determine an upper bound on the variance $\gamma[n,n]$.
+It follows:
+\begin{align*}
+\gamma[n,n] &\leq c_0^{(n)}\sigma^2 + \dfrac{1}{K} \left( \dfrac{c_1^{(n)}}{\sigma^2} + c_2^{(n)} + c_3^{(n)}\sigma^2\right) + o\left(\dfrac1K\right)\ ,
+\end{align*}
+where
+\begin{align*}
+c_0^{(n)} & =  M^{\ell}\max\left(1,\left(\frac{2J}{M}\right)^{2\ell}\right) \\
+c_1^{(n)} & = \dfrac14C_\bz^4 M^2(M+1)^2d_{1,\bz,M,\ell}^2  \\
+c_2^{(n)} & = \dfrac12C_\bz^2 M^2(M+1)^2d_{1,\bz,M,\ell}^2 + M^\ell(M+1)C_\bz^2 d_{1,\bz,M,\ell}\left(\max\left(1,\frac{2J}{M}\right)\right)^\ell\\
+c_3^{(n)} & =  8 M^{\ell+2}(M+1)d_{1,\bz,M,\ell}\left(\max\left(1,\frac{2J}{M}\right)\right)^\ell\ .
+\end{align*}
 
-\paragraph{When $n'\in I$}
-From Lemma~1, we have that $h^{(\ell)}$ is asymptotically Gaussian when $K\to \infty$.
-Then, as a direct consequence of the Isserlis' theorem, odd-order moments are vanishing {\color{red}[there is a gap here.
-Note that when $K$ is finite, it is approximated by Gaussian but not Gaussian.
-The discrepancy should be made clear.]}.
-Then, inserting result~\eqref{eq:bias.ap} into equation~\eqref{I-eq:bias0}, we obtain:
-\begin{equation}
-\bgamma[n,n'] = \sigma\EE\{\bw[n']\bh^{(\ell)}\}\bz_K + \sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bw[n']\} + o(\sigma^2) 
-\label{eq:gamma.nn}
-\end{equation}
-when $K\to \infty$.
-In addition, we remark that
-\[
-\balpha_0^{(\ell)}\EE\{\bw_K\bw[n']\}=\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)}\ .
-\]
-Besides, according to result~\eqref{eq:cov.ap} the first term in~\eqref{eq:gamma.nn} vanishes when $K\to\infty$. We thus have
+\paragraph{If $n'\geq N$.}
+When $n\geq N$ and $n'\geq N$, applying the Cauchy-Schwarz inequality, we obtain the following bound on the covariance $\gamma[n,n]$:
 \begin{align*}
-\bgamma[n,n'] &= \sigma^2\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)}+ o(\sigma^2)
+\left|\gamma[n,n]\right| & \leq  \sqrt{\gamma[n,n]\gamma[n',n']} \\
+& \leq  \sqrt{\gamma[n,n]\gamma[n',n']} \\
+& \leq   \sqrt{c_0^{(n)}c_0^{(n')}}\sigma^2\sqrt{1 + \dfrac{1}{K\sigma^2} \left( \dfrac1{c_0^{(n)}}\left( \dfrac{c_1^{(n)}}{\sigma^2} + c_2^{(n)} + c_3^{(n)}\sigma^2\right) + \dfrac1{c_0^{(n')}}\left( \dfrac{c_1^{(n')}}{\sigma^2} + c_2^{(n')} + c_3^{(n')}\sigma^2\right)\right) + o\left(\dfrac1K\right)} \\
+& \leq   \sqrt{c_0^{(n)}c_0^{(n')}}\sigma^2 + \dfrac{1}{2K} \left( \sqrt{\dfrac{c_0^{(n')}}{c_0^{(n)}}}\left( \dfrac{c_1^{(n)}}{\sigma^2} + c_2^{(n)} + c_3^{(n)}\sigma^2\right) + \sqrt{\dfrac{c_0^{(n)}}{c_0^{(n')}}}\left( \dfrac{c_1^{(n')}}{\sigma^2} + c_2^{(n')} + c_3^{(n')}\sigma^2\right)\right) + o\left(\dfrac1K\right) \\
+& \leq  c_0^{(n,n')}\sigma^2 + \dfrac{1}{K} \left( \dfrac{c_1^{(n,n')}}{\sigma^2} + c_2^{(n,n')} + c_3^{(n,n')}\sigma^2\right) + o\left(\dfrac1K\right)\ , \\
 \end{align*}
-when $K\to\infty$.
+where
+\begin{align*}
+c_0^{(n,n')} & =  \sqrt{c_0^{(n)}c_0^{(n')}}\,, \\
+c_p^{(n,n')} & =  \dfrac12\left(c_p^{(n)}\sqrt{\dfrac{c_0^{(n')}}{c_0^{(n)}}} + c_p^{(n')}\sqrt{\dfrac{c_0^{(n)}}{c_0^{(n')}}}\right)\,,\quad\forall p\in\{1,2,3\}\ .\\
+\end{align*}
 
-\paragraph{When $n'\geq N$}
-Inserting equation~\eqref{eq:bias.ap} into result~\eqref{I-eq:cov01}, we obtain
+\paragraph{If $n'\in I$.}
+When $n\geq N$ and $n'\in I$, equation~\eqref{I-eq:cov00} shows us that:
 \begin{align*}
-\bgamma[n,n'] = &\,\bz_K^T\EE\left\{{\balpha^{(\ell)}}^T\balpha^{(\lambda)}\right\}\bz_K + \sigma\EE\{\balpha^{(\ell)}\bw_K\balpha^{(\lambda)}\}\bz_K+ \sigma\EE\{\balpha^{(\lambda)}\bw_K\balpha^{(\ell)}\}\bz_K \\
-& +\sigma^2\EE\{\balpha^{(\ell)}\bw_K\balpha^{(\lambda)}\bw_K\}-\bz[n]\bz[n']+o(\sigma^2) \\
-= &\,\bz_K^T\EE\left\{{\bh^{(\ell)}}^T\bh^{(\lambda)}\right\}\bz_K + \balpha_0^{(\ell)}\sigma\EE\{\bw_K\bh^{(\lambda)}\}\bz_K + \sigma\EE\{\bh^{(\ell)}\bw_K\}\bz[n']  \\
-&+ \sigma\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\}\bz_K+ \sigma\EE\{\bh^{(\lambda)}\bw_K\}\bz[n] + \sigma\balpha_0^{(\lambda)}\EE\{\bw_K\bh^{(\ell)}\}\bz_K \\
-& + \sigma\EE\{\bh^{(\lambda)}\bw_K\bh^{(\ell)}\}\bz_K+\sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\lambda)}\bw_K\}+\sigma^2\balpha_0^{(\lambda)}\EE\{\bw_K\bh^{(\ell)}\bw_K\} \\
-&+\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\sigma^2\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} + o(\sigma^2)\ .
+\bgamma[n,n']&=\sigma\EE\{\bw[n']\bh^{(\ell)}\}\bz_K + \sigma^2\EE\{\bw[n']\balpha_0^{(\ell)}\bw_K\} + \sigma^2\EE\{\bw[n']\bh^{(\ell)}\bw_K\} \\
+& \defeq \beta_1[n,n'] + \beta_2[n,n'] + \beta_3[n,n']\ ,
 \end{align*}
-From Lemma~1, we have that $\bh^{(\ell)}$ is asymptotically Gaussian.
-Then, as a direct consequence of the Isserlis' theorem, odd-order moments are vanishing.
-{\color{red}[The same gap here.]} {\color{blue}Considering this property and equation~\eqref{eq:cov.ap} gives
+where
 \begin{align*}
-\bgamma[n,n'] = &\,\dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K + \sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\sigma^2\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} + o(\sigma^2)
+\beta_1[n,n'] &= \sigma\EE\{\bw[n']\bh^{(\ell)}\}\bz_K \\
+\beta_2[n,n'] &= \sigma^2\EE\{\bw[n']\balpha_0^{(\ell)}\bw_K\} \\
+\beta_3[n,n'] &= \sigma^2\EE\{\bw[n']\bh^{(\ell)}\bw_K\}\ .
 \end{align*}
-when $K\to\infty$. The Isserlis' theorem also gives the following result {\color{red}[the same gap here.]}
+Besides, thanks to the bound~\eqref{eq:Ehwm} we have:
+\begin{align}
+\nonumber
+\left|\beta_1[n,n']\right| &\leq \sigma\dfrac{C_\bz}{2}\sum_{m=0}^{M-1}\left|\EE\{\bw[n']\bh^{(\ell)}[m]\}\right|\\
+\nonumber
+&\leq \sigma\dfrac{C_\bz}{2}M^2(M+1)\dfrac{d_{1,\bz,M,\ell}}{\sigma^2} C_\bz\dfrac{\sigma}{K} + o\left(\dfrac1K\right)\\
+&\leq  M^2(M+1)\dfrac{C_\bz^2 d_{1,\bz,M,\ell}}{2}\dfrac{1}{K} + o\left(\dfrac1K\right)\ .
+\label{eq:beta1}
+\end{align}
+Besides,
+\begin{align}
+\nonumber
+\beta_2[n,n'] &\leq \sigma^2\left\|\balpha_0^{(\ell)}\right\|_\infty\sum_{m=0}^{M-1}\left|\EE\{\bw[n']\bw_K[m]\}\right| \\
+\nonumber
+&\leq \sigma^2\left\|\bA_0^{\ell}\right\|_\infty \\
+&\leq \sigma^2M^{\ell-1}\left(\max\left(1,\frac{2J}{M}\right)\right)^{\ell}\ .
+\label{eq:beta2}
+\end{align}
+Besides, identically to the bound~\eqref{eq:sum7} on $\eta_7$, we obtain 
+\begin{align}
+\nonumber
+\beta_3[n,n'] &\leq \sigma^2\sum_{m=0}^{M-1}\left|\EE\{\bw[n']\bh^{(\ell)}[m]\bw_K[m]\}\right|\\
+\nonumber
+&\leq \sigma^2 M^3(M+1)\dfrac{d_{1,\bz,M,\ell}}{\sigma^2} \dfrac{4\sigma^2}K + o\left(\dfrac1K\right) \\
+&\leq M^3(M+1)d_{1,\bz,M,\ell} \dfrac{4\sigma^2}K + o\left(\dfrac1K\right) \ .
+\label{eq:beta3}
+\end{align}
+Finally, we combine expressions~\eqref{eq:beta1},~\eqref{eq:beta2}, and~\eqref{eq:beta3} to determine an upper bound on the variance $\gamma[n,n']$. It follows:
 \begin{align*}
-\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} &= \sum_{m,m'=0}^{M-1}\EE\{\bh^{(\ell)}[m]\bw_K[m]\bh^{(\lambda)}[m']\bw_K[m']\} \\
-&= \sum_{m,m'=0}^{M-1}\EE\{\bh^{(\ell)}[m]\bw_K[m]\}\EE\{\bh^{(\lambda)}[m']\bw_K[m']\} + \EE\{\bh^{(\ell)}[m]\bw_K[m']\}\EE\{\bh^{(\lambda)}[m']\bw_K[m]\} \\
-&\hspace{35pt} + \EE\{\bh^{(\ell)}[m]\bh^{(\lambda)}[m']\}\underbrace{\EE\{\bw_K[m]\bw_K[m']\}}_{=\delta_{m,m'}} \\
-&= \sum_{m=0}^{M-1} \EE\{\bh^{(\ell)}[m]\bh^{(\lambda)}[m]\} = \dfrac1K\Tr\left(\bGamma^{(\ell,\lambda)}\right)
+\left|\gamma[n,n']\right| &\leq b_0^{(n,n')}\sigma^2 + \dfrac{1}{K} \left( b_1^{(n,n')} + b_2^{(n)}\sigma^2\right) + o\left(\dfrac1K\right)\ ,
 \end{align*}
-when $K\to \infty$.
-We thus conclude that
+where
 \begin{align*}
-\bgamma[n,n'] &= \dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K+\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\dfrac{\sigma^2}{K}\Tr\left(\bGamma^{(\ell,\lambda)}\right) + o(\sigma^2)
+b_0^{(n,n')} & =  M^{\ell-1}\left(\max\left(1,\frac{2J}{M}\right)\right)^{\ell} \\
+b_1^{(n,n')} & = M^2(M+1)\dfrac{C_\bz^2 d_{1,\bz,M,\ell}}{2} \\
+b_2^{(n,n')} & = 4M^3(M+1)d_{1,\bz,M,\ell}  \ .
 \end{align*}
-when $K\to \infty$.}
 
@@
 Nevertheles, the results show that {\sf BoundEffRed} has the same efficiency when the {\sf SigExt} extension, the EDMD extension or the GPR extension is chosen.
+Indeed, t-tests performed under the null hypothesis that the mean are equals, with a $5\%$ significance level, show no statistical significant difference between {\sf SigExt} and EDMD or GPR, regardless of the representation considered.
 This justifies the choice of {\sf SigExt} for real-time implementation.
@@
 
+%The multivariate version of the delta method is applicable to $\bh^{(\ell)}$ under the following conditions:
+%\begin{enumerate}[(i)]
+%\item\label{delta.1} $f^{(\ell)}$ is differentiable at the origin;
+%\item\label{delta.2} $\sqrt{K}\ \bg \xrightarrow[K\to\infty]{\cD} \cN(\bzero,\bGamma_0)$ where $\bGamma_0$ is a covariance matrix.
+%\end{enumerate}
+%When these conditions are satisfied, the delta method gives the asymptotic behavior of $\bh^{(\ell)}=f^{(\ell)}(\bg)$, that is
+%\begin{equation}
+%\sqrt{K}\ \bh^{(\ell)} \xrightarrow[K\to\infty]{\cD} \cN(\bzero,{\bF^{(\ell)}}^T\bGamma_0\bF^{(\ell)})\ ,
+%\label{eq:delta}
+%\end{equation}
+%where $\bF^{(\ell)}$ is the Jacobian matrix of $f^{(\ell)}$ at the origin, that is
+%\[
+%\bF^{(\ell)}[m,m'] = \left.\dfrac{\partial f^{(\ell)}_m}{\partial\bg[m']}\right\vert_{\bg=\bzero}\ .
+%\]
+%
+%Condition~\ref{delta.1} is satisfied. Indeed, the differentiation of $f^{(\ell)}$ can be checked as a composition of standard matrix derivation rules. {\color{red}Detail it}
+%
+%Concerning condition~\ref{delta.2}, by definition of $\bg$, we have:
+%\[
+%\bg = \sigma\bg_1 + \sigma^2\bg_2\ ,
+%\]
+%where $\bg_1$ and $\bg_2$ are defined by
+%\begin{align*}
+%\bg_1 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left( \tilde\bz_k\bw_k^T + \tilde\bw_k\bz_k^T \right)\quad\mbox{where}\quad \tilde\bz_k^T= \begin{pmatrix} \bz_k^T & \bz_{k+1}[M-1] \end{pmatrix}\,, \\
+%\bg_2 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left(\tilde\bw_k\bw_k^T-\bI \right) \quad\mbox{where}\quad \tilde\bw_k^T= \begin{pmatrix} \bw_k^T & \bw_{k+1}[M-1] \end{pmatrix} \,.
+%\end{align*}
+%First, $\bg_1$ is intrinsically a Gaussian random vector since it is a linear combination of Gaussian random vectors.
+Second, using the central limit theorem under weak dependence, we can show that $\bg_2$ also converges towards a Gaussian random vector as $K\to\infty$.
+Hence, we have
+%\[
+%\sqrt{K}\ \bg \xrightarrow[K\to\infty]{\cD} \cN(\bzero,\bGamma_0)\ ,
+%\]
+%where $\bGamma_0$ is the limit covariance matrix so that $\bGamma_0=\lim\limits_{K\to\infty} K\,\EE\{\bg\bg^T\}$.
+Hence, conditions~\ref{delta.1} and~\ref{delta.1} are satisfied.
+The delta method can therefore be applied, and result~\eqref{eq:delta} is valid.
+
+%Therefore,
+%\begin{align*}
+%\bgamma[n,n] &\leq \sigma^2\left\|\balpha_0^{(\ell)}\right\|^2-\left(\bepsilon_2[n]+\bepsilon_3[n]\right)^2 + \EE\left\{\left(\bh^{(\ell)}\bz_K\right)^2\right\} + 2\sigma\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\ell)}\}\bz_K+ 2\sigma\EE\{\bh^{(\ell)}\bw_K\bh^{(\ell)}\}\bz_K \\
+%&+\sigma^2\EE\left\{\left(\bh^{(\ell)}\bw_K\right)^2\right\}  + 2\sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\ell)}\bw_K\}
+%\end{align*}
+
+%Take $n>N$, and denote $n=N-1+\ell$.  To derive the covariance, let us segregate the cases. 
+% 
+%\paragraph{When $n'\in I$}
+%From Lemma~1, we have that $h^{(\ell)}$ is asymptotically Gaussian when $K\to \infty$.
+Then, as a direct consequence of the Isserlis' theorem, odd-order moments are vanishing {\color{red}[there is a gap here.
+Note that when $K$ is finite, it is approximated by Gaussian but not Gaussian.
+The discrepancy should be made clear.]}.
+Then, inserting result~\eqref{eq:bias.ap} into equation~\eqref{I-eq:bias0}, we obtain:
+%\begin{equation}
+%\bgamma[n,n'] = \sigma\EE\{\bw[n']\bh^{(\ell)}\}\bz_K + \sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bw[n']\} + o(\sigma^2) 
+%\label{eq:gamma.nn}
+%\end{equation}
+%when $K\to \infty$.
+%In addition, we remark that
+%\[
+%\balpha_0^{(\ell)}\EE\{\bw_K\bw[n']\}=\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)}\ .
+%\]
+%Besides, according to result~\eqref{eq:cov.ap} the first term in~\eqref{eq:gamma.nn} vanishes when $K\to\infty$. We thus have
+%\begin{align*}
+%\bgamma[n,n'] &= \sigma^2\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)}+ o(\sigma^2)
+%\end{align*}
+%when $K\to\infty$.
+%
+%
+%\paragraph{When $n'\geq N$}
+%Inserting equation~\eqref{eq:bias.ap} into result~\eqref{I-eq:cov01}, we obtain
+%\begin{align*}  
+%\bgamma[n,n'] = &\,\bz_K^T\EE\left\{{\balpha^{(\ell)}}^T\balpha^{(\lambda)}\right\}\bz_K + \sigma\EE\{\balpha^{(\ell)}\bw_K\balpha^{(\lambda)}\}\bz_K+ \sigma\EE\{\balpha^{(\lambda)}\bw_K\balpha^{(\ell)}\}\bz_K \\
+%& +\sigma^2\EE\{\balpha^{(\ell)}\bw_K\balpha^{(\lambda)}\bw_K\}-\bz[n]\bz[n']+o(\sigma^2) \\
+%= &\,\bz_K^T\EE\left\{{\bh^{(\ell)}}^T\bh^{(\lambda)}\right\}\bz_K + \balpha_0^{(\ell)}\sigma\EE\{\bw_K\bh^{(\lambda)}\}\bz_K + \sigma\EE\{\bh^{(\ell)}\bw_K\}\bz[n']  \\
+%&+ \sigma\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\}\bz_K+ \sigma\EE\{\bh^{(\lambda)}\bw_K\}\bz[n] + \sigma\balpha_0^{(\lambda)}\EE\{\bw_K\bh^{(\ell)}\}\bz_K \\
+%& + \sigma\EE\{\bh^{(\lambda)}\bw_K\bh^{(\ell)}\}\bz_K+\sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\lambda)}\bw_K\}+\sigma^2\balpha_0^{(\lambda)}\EE\{\bw_K\bh^{(\ell)}\bw_K\} \\
+%&+\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\sigma^2\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} + o(\sigma^2)\ .
+%\end{align*}
+%From Lemma~1, we have that $\bh^{(\ell)}$ is asymptotically Gaussian.
+Then, as a direct consequence of the Isserlis' theorem, odd-order moments are vanishing.
+{\color{red}[The same gap here.]} {\color{blue}Considering this property and equation~\eqref{eq:cov.ap} gives
+%\begin{align*}
+%\bgamma[n,n'] = &\,\dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K + \sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\sigma^2\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} + o(\sigma^2)
+%\end{align*}
+%when $K\to\infty$. The Isserlis' theorem also gives the following result {\color{red}[the same gap here.]}
+%\begin{align*}
+%\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} &= \sum_{m,m'=0}^{M-1}\EE\{\bh^{(\ell)}[m]\bw_K[m]\bh^{(\lambda)}[m']\bw_K[m']\} \\
+%&= \sum_{m,m'=0}^{M-1}\EE\{\bh^{(\ell)}[m]\bw_K[m]\}\EE\{\bh^{(\lambda)}[m']\bw_K[m']\} + \EE\{\bh^{(\ell)}[m]\bw_K[m']\}\EE\{\bh^{(\lambda)}[m']\bw_K[m]\} \\
+%&\hspace{35pt} + \EE\{\bh^{(\ell)}[m]\bh^{(\lambda)}[m']\}\underbrace{\EE\{\bw_K[m]\bw_K[m']\}}_{=\delta_{m,m'}} \\
+%&= \sum_{m=0}^{M-1} \EE\{\bh^{(\ell)}[m]\bh^{(\lambda)}[m]\} = \dfrac1K\Tr\left(\bGamma^{(\ell,\lambda)}\right)
+%\end{align*}
+%when $K\to \infty$.
+%We thus conclude that
+%\begin{align*}  
+%\bgamma[n,n'] &= \dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K+\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\dfrac{\sigma^2}{K}\Tr\left(\bGamma^{(\ell,\lambda)}\right) + o(\sigma^2)
+%\end{align*}
+%when $K\to \infty$.}
+
 \end{document}

### 2020-08-26 Notes/results.tex
new theorem

@@
 Indeed, independently of $\sigma$, we always have $\mu_{\xp}[N-1+\ell]\in\left[-0.03\sigma,0.03\sigma\right]$, which is negligible with respect to the magnitude of $\bz$.
-This result confirms the theoretical result~\eqref{eq:mean.error}.
+This result does not contradict the theoretical result~\eqref{eq:mean.error} of Theorem~\ref{th:error} insofar as the bound is not reached.
 In Fig.~\ref{fig:res.noise.sine}, we display the experimental variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$ for each value of $\sigma$.
@@
 
-As in the previous study, the experimental bias vanishes when $K$ increases, confirming the approximation result~\eqref{eq:mean.error}.
+As in the previous study, the experimental bias becomes negligible when $K$ increases, confirming the approximation result~\eqref{eq:mean.error}.
 Besides, the experimental variance  is displayed in Fig.~\ref{fig:res.size.sine}, and compared with the associated theoretical variance~\eqref{eq:cov.error.2}.
@@
 The resulting MSEs of different methods are given in Table~\ref{tab:THO}.
+Note that the results are given in ``mean $\pm$ standard deviation'' format.
 The TBATS extension method is not implemented, as its excessive computing time makes it impossible to implement it on $6000$ segments within a reasonable period of time.
@@
   \hline
-   \multirow{2}{38pt}{\centering Extension method} & \multirow{2}{43pt}{\centering MSE\\ (mean $\!\!\pm\!$ SD)} & \multicolumn{4}{c|}{Performance index $D$ (mean $\pm$ SD)} \\
+   \multirow{2}{38pt}{\centering Extension method} & \multirow{2}{29pt}{\centering MSE} & \multicolumn{4}{c|}{Performance index $D$ (mean $\pm$ SD)} \\
    \cline{3-6}

### 2020-08-26 Notes/Supplementary material/SupMat.tex
proof

@@
 \end{align}
-Besides,
+Besides, the second-order derivative of the matrix $\tilde\bA$ is given by
 \begin{align*}
@@
 \end{align*}
-Thus,
+Thus, the second-order derivative of the matrix $\tilde\bA$ at the origin is such that
 \begin{align*}
@@
 \begin{array}{ll}
-\|{\bS^{(0)}}^{-1}\|_\infty^2 &\mbox{if}\quad a_r=1\quad\mbox{or}\quad a_{r'}=1 \\
+\|{\bS^{(0)}}^{-1}\|_\infty^2 &\mbox{if}\quad a_r=1\quad\mbox{or}\quad a_{r'}=1\,, \\
 2\left(1+\|\bA_0\|_\infty\right)\|{\bS^{(0)}}^{-1}\|_\infty^2 &\mbox{else} \\
@@
 
-Let us now determine an upper bound on $\left|\bepsilon_2[\ell]\right|$. Since $\EE\{\bg[r]\}=0$, a second-order Taylor expansion of $\bh^{(\ell)}$ gives
+Let us now determine an upper bound on $\left|\bepsilon_2[\ell]\right|$.
+Since $\EE\{\bg[r]\}=0$ and moments of order $3$ and higher behave as $o\left(\dfrac1K\right)$, a second-order Taylor expansion of $\bh^{(\ell)}$ gives
 \begin{equation}

### 2020-08-27 Notes/results.tex
new figs

@@
 \paragraph{Influence of the Noise Variance $\sigma^2$} Here, the size of the training dataset is set to $K=450$.
-Then, the forecasting algorithm is run on $1000$ realizations on the discrete signal $\bx$ for three different values of $\sigma$, logarithmically equispaced from $10^{-3}$ to $10^{-1}$.
-For each of these values, we determine the experimental bias, denoted as $\mu_{\xp}[N-1+\ell]$, and experimental variance, denoted as $\gamma_{\xp}[N-1+\ell,N-1+\ell]$, in function of the forecasting sample index $\ell$ (going from $1$ to $L=100$).
-
-The experimental results show that the bias is neither depending on the noise variance $\sigma^2$ nor the forecasting length $\ell$.
-Indeed, independently of $\sigma$, we always have $\mu_{\xp}[N-1+\ell]\in\left[-0.03\sigma,0.03\sigma\right]$, which is negligible with respect to the magnitude of $\bz$.
-This result does not contradict the theoretical result~\eqref{eq:mean.error} of Theorem~\ref{th:error} insofar as the bound is not reached.
-In Fig.~\ref{fig:res.noise.sine}, we display the experimental variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$ for each value of $\sigma$.
-The associated theoretical asymptotic forecasting variance~\eqref{eq:cov.error.2} is also displayed in solid line.
-As expected, this result highlights the fact that the forecasting variance increases linearly with respect to $\sigma^2$.
-The {\em slight} decrease of the forecasting variance with $\ell$ is characteristic of the stationarity of the forecast signal.
-This shows that as $\ell$ increases, the prediction becomes less correlated with the measurement noise.
+Then, the forecasting algorithm is run on $500$ realizations on the discrete signal $\bx$ for $200$ different values of $\sigma$, logarithmically equispaced from $10^{-7}$ to $10^{-1}$.
+For each of these values, we determine the experimental bias, denoted as $\mu_{\xp}[N-1+\ell]$, and experimental variance, denoted as $\gamma_{\xp}[N-1+\ell,N-1+\ell]$.
+Fig.~\ref{fig:res.noise.sine} shows the the experimental forecasting variance in function of the noise standard deviation for three different values the forecasting sample index: $\ell=1$, $\ell=10$, and $\ell=100$.
 
@@
 \includegraphics[width=.48\textwidth]{VarianceNoiseSine.eps}
-\caption{Evolution of the experimental and theoretical forecasting variance in function of the forecasting sample index for different values of $\sigma$.}
+\caption{Evolution of the experimental forecasting variance in function of the noise standard deviation for three different values the forecasting sample index $\ell$.}
 \label{fig:res.noise.sine}
@@
 
+{\color{blue}
+Comment Fig. + Comparison with Theorem 1
+}
+
 \paragraph{Influence of the Training Dataset Size $K$} Here, the noise variance $\sigma$ is set to $\sigma=10^{-2}$.
-Then, {\sf SigExt} is run on 3000 realizations of the discrete signal $\bx$ for three different values of $K$, logarithmically equispaced from $4.5\times 10^{2}$ to $2\times 10^{3}$.
-For each of these values, we determine the experimental bias $\mu_{\xp}[N-1+\ell]$ and variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$ in function of the forecasting sample index $\ell$, going from $1$ to $500$.
-
-As in the previous study, the experimental bias becomes negligible when $K$ increases, confirming the approximation result~\eqref{eq:mean.error}.
-Besides, the experimental variance  is displayed in Fig.~\ref{fig:res.size.sine}, and compared with the associated theoretical variance~\eqref{eq:cov.error.2}.
-Each color corresponds to these experimental results obtained for a given value of $K$.
-This result validates the asymptotic behavior provided by~\eqref{eq:cov.error.2}, and we can show that the second-order moment $\gamma[\ell,\ell]$ is less dependent on $K$ when $K$ is large.
+Then, {\sf SigExt} is run on $500$ realizations of the discrete signal $\bx$ for $200$ different values of training dataset size $K$, logarithmically equispaced from $450$ to $2000$.
+For each of these values, we determine the experimental bias $\mu_{\xp}[N-1+\ell]$ and variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$.
+Fig.~\ref{fig:res.size.sine} shows the the experimental forecasting variance in function of the training dataset size for three different values the forecasting sample index: $\ell=1$, $\ell=10$, and $\ell=100$.
 
@@
 \includegraphics[width=.48\textwidth]{VarianceKSine.eps}
-\caption{Evolution of the experimental and theoretical forecasting variance in function of the forecasting sample index for different values of $K$.}
+\caption{Evolution of the experimental forecasting variance in function of the dataset size for three different values the forecasting sample index $\ell$.}
 \label{fig:res.size.sine}
@@
 
-\paragraph{Summary}
-Both previous experimental results combined with the theoretical asymptotic equation~\eqref{eq:cov.error.2} allow us to describe the influence of the noise variance and the size of the training dataset on the variance of the forecasting noise, empirically summarized as follows:
-\begin{equation}
-\gamma[N-1+\ell,N-1+\ell] \underset{K\to\infty}{\approx} \dfrac{\sigma^2}{K}g[\ell] .
-\end{equation} 
-where $g$ is a bounded positive function. The empirical result is coherent with the theoretical result provided by Theorem~\ref{th:error}.
+{\color{blue}
+{\color{blue}
+Comment Fig. + Comparison with Theorem 1
+}
 
-This study neglects the analysis of the influence of the parameter $M$, whose influence on the value of the experimental variance is numerically not significant as long as $M\ll 2K$.
+%\paragraph{Summary}
+%Both previous experimental results combined with the theoretical asymptotic equation~\eqref{eq:cov.error.2} allow us to describe the influence of the noise variance and the size of the training dataset on the variance of the forecasting noise, empirically summarized as follows:
+%\begin{equation}
+%\gamma[N-1+\ell,N-1+\ell] \underset{K\to\infty}{\approx} \dfrac{\sigma^2}{K}g[\ell] .
+%\end{equation} 
+%where $g$ is a bounded positive function. The empirical result is coherent with the theoretical result provided by Theorem~\ref{th:error}.
+}
+
+This study neglects the analysis of the influence of the parameter $M$, whose influence on the value of the experimental variance is numerically not significant as long as $M\ll 2K$ and $\ell<M$.
 The choice of this parameter is especially crucial when the deterministic component of the signal is no longer stationary.

### 2020-08-27 Scripts/EvalPerfSigExt/InfluenceNoiseSine.m

@@ -3,14 +3,14 @@
 % Email: adrien.meynard@duke.edu
 
 clear all; 
-close all; clc;
+close all; clc; warning off;
 addpath('../../Algorithm/');
 
 %% Parameters
 
 % forecasting parameters
 HOP = 1 ;
-L = 350 ;
+L = 100 ;
 extM = round( 1.5*L ) ; % dimension of embedding / signals length
 extK = round( 3*extM );  % number of points to estimate A / size of datasets
 
@@ -50,9 +50,9 @@ zK = x0(end-extM+1 : HOP: end).' ;
 
 %% Forecasting
 method.name = 'SigExt' ;
-nbXP = 3 ; % number of noise levels
-nbXPP = 3000 ; % number of XP / noise level
-Sigma = logspace(-3,-1,nbXP) ;
+nbXP = 200 ; % number of noise levels
+nbXPP = 500 ; % number of XP / noise level
+Sigma = logspace(-7,-1,nbXP) ;
 
 hopL = 10 ; % hop for forecasting
 lenForcast = length(1:hopL:L) ;
@@ -62,55 +62,19 @@ k = 1 ;
 for k = 1:nbXP
     sigman = Sigma(k) ;
     
-    % Theoretical foreacstion Matrix
-    S0 = (1/extK)*(X0*X0') + sigman^2*eye(extM) ;
-    r = zeros(1,extM) ; r(2) = 1 ;
-    S1 = (1/extK)*(Y0*X0') + sigman^2*toeplitz(zeros(extM,1),r) ;
-    A0 = S1 / S0 ; % ideal least square estimation
-    j = 1;
-    for l = 1:hopL:L
-        A0l{j} = A0^l ;
-        alpha0l(j,:) = A0l{j}(end,:) ;
-        j = j + 1 ;
-    end
-    a2 = sum(alpha0l.^2, 2) ;
-    
     for nb2 = 1:nbXPP
         noise = sigman*randn(N+L,1) ;
-        wK = noise((N-extM+1):N)/sigman ;
         x = x0.' + noise(1:N) ; % signal to be extended
         
-        % Evaluate h
-        X = [] ; Y = [] ;
-        for kk = 1: extK
-            X(:,kk) = x(end-extK-extM+kk: HOP: end-extK+kk-1) ;
-            Y(:,kk) = x(end-extK-extM+kk+1: HOP: end-extK+kk) ;
[…]

### 2020-08-27 Scripts/EvalPerfSigExt/ResultsInfluenceNoiseSine.m

@@ -1,30 +1,0 @@
-%% Display the results of InfluenceNoiseSine.m and ForecastAMFM_TBATS.r
-% Author: Adrien MEYNARD
-% Email: adrien.meynard@duke.edu
-
-clear all; close all;
-
-load('../../Results/PerfNoise.mat')
-L = size(VarXPm,2) ;
-nbXP = length (Sigma) ;
-
-%% Plot
-ll = 1:L ;
-
-for k = 1:nbXP
-    sigma = Sigma(k) ;
-    txtXP = ['Experimental \sigma = ',num2str(sigma,'%.0e')];
-    txtTH = ['Theoretical    \sigma = ',num2str(sigma,'%.0e')];
-    figure(1);
-    plot(ll,MeanXPm(k,:)/sigma,'linewidth',2,'DisplayName',txtXP); grid on; hold on;
-    figure(2);
-    fig = semilogy(ll,VarXPm(k,:),'+','linewidth',2,'MarkerSize',8,'DisplayName',txtXP); grid on; hold on;
-    semilogy(ll,VarTH(k,:),'-','color',fig.Color,'linewidth',2,'DisplayName',txtTH); hold on;
-end
-%legend('Experimental MSE[10]','Experimental MSE[100]','Experimemtal MSE[200]','Experimental MSE[500]','\sigma^2');
-figure(1); xlabel('Forecasting sample index ($\ell$)','interpreter', 'latex','fontsize',28); ylabel('Normalized experimental forecasting bias ($\mu_{xp}[N-1+\ell]/sigma$)','interpreter','latex','fontsize',28);
-legend show ; %ylim([-0.05 0.05]);
-figure(2); xlabel('Forecasting sample index ($\ell$)','interpreter', 'latex','fontsize',28); ylabel('Forecasting variance ($\gamma[N-1+\ell,N-1+\ell]$)','interpreter','latex','fontsize',28);
-lgd = legend('show') ;
-lgd.FontSize = 18 ;
-set(gca,'fontsize',24) ;

### 2020-08-27 Notes/results.tex
comments figs

@@
 For each of these values, we determine the experimental bias, denoted as $\mu_{\xp}[N-1+\ell]$, and experimental variance, denoted as $\gamma_{\xp}[N-1+\ell,N-1+\ell]$.
-Fig.~\ref{fig:res.noise.sine} shows the the experimental forecasting variance in function of the noise standard deviation for three different values the forecasting sample index: $\ell=1$, $\ell=10$, and $\ell=100$.
+Fig.~\ref{fig:res.noise.sine} shows the the experimental forecasting variance in function of the noise variance for three different values the forecasting sample index: $\ell=1$, $\ell=10$, and $\ell=100$.
 
@@
 \includegraphics[width=.48\textwidth]{VarianceNoiseSine.eps}
-\caption{Evolution of the experimental forecasting variance in function of the noise standard deviation for three different values the forecasting sample index $\ell$.}
+\caption{Evolution of the experimental forecasting variance in function of the noise variance for three different values the forecasting sample index $\ell$.}
 \label{fig:res.noise.sine}
@@
 
-{\color{blue}
-Comment Fig. + Comparison with Theorem 1
-}
+First, when the noise is weak, the variance of the forecasting estimator is dramatically high.
+For instance, in this specific example, the estimator $\tilde\bx[N-1+\ell]$ for $\ell=10$ is inaccurate as soon as the noise variance $\sigma^2$ is lower than $4\times 10^{-14}$ (see the red line in Fig.~\ref{fig:res.noise.sine}).
+This validates the result of Theorem~\ref{th:error}, highlighting the necessary presence of noise to ensure the numerical stability of the estimator.
+Furthermore, as soon as $\sigma$ is higher than this critical threshold, the estimator variance increases linearly with $\sigma^2$, as predicted by the bound~\eqref{eq:cov.error.2} provided in Theorem~\ref{th:error}.
+Note that the parameter $\ell$ has little influence on the quality of the estimator.
+This is due to the \textit{stationarity} of the studied signal, allowing long-range forecasts without deterioration of the estimator.
 
@@
 
-{\color{blue}
-{\color{blue}
-Comment Fig. + Comparison with Theorem 1
-}
+As anticipated by Theorem~\ref{th:error}, the variance of the estimator decreases linearly with $\frac1K$ as long as the dataset size is large enough.
+Indeed, in the log-log plot displayed in Fig.~\ref{fig:res.size.sine}, as soon as $K>700$, each of the three curves representing the forecasting variance tends towards an asymptote of slope $-1$.
 
@@
 %where $g$ is a bounded positive function. The empirical result is coherent with the theoretical result provided by Theorem~\ref{th:error}.
-}
 
-This study neglects the analysis of the influence of the parameter $M$, whose influence on the value of the experimental variance is numerically not significant as long as $M\ll 2K$ and $\ell<M$.
+This study overlooks the analysis of the influence of the parameter $M$, whose influence on the value of the experimental variance is numerically not significant as long as $M\ll 2K$ and $\ell<M$.
 The choice of this parameter is especially crucial when the deterministic component of the signal is no longer stationary.

### 2020-08-27 Scripts/EvalPerfSigExt/InfluenceParam.m

@@ -1,121 +1,0 @@
-clear all;
-close all; clc;
-addpath('../../Algorithm/');
-addpath(genpath('../../TimeFrequencyScaleRep'));
-addpath('../../Signals');
-
-%% load signal
-SigTyp = 'THO' ;
-
-switch SigTyp
-    case 'THO'
-        data = load('THO.mat');
-        xtot = data.THO;
-        fs = 100 ; % sampling frequency
-        
-        x = xtot(204e3:208e3);
-        mu = mean(x) ;
-        s = std(x) ;
-        x = (x - mu)/s ;
-
-        extSEC = 2 ; % the extension is of extSEC second
-        L = round(extSEC*fs) ;
-        xxTRUE = ( xtot((204e3-L):(208e3+L)) - mu ) / s ;
-
-    case 'PPG'
-        data = load('PPGsig');
-
-        xtot = data.PPG;
-        fs = data.fs; % sampling frequency
-
-        x = xtot(30e3:34e3);
-        mu = mean(x) ;
-        s = std(x) ;
-        x = (x - mu)/s ;
-        
-        extSEC = 2 ; % the extension is of extSEC second
-        L = round(extSEC*fs) ;
-        xxTRUE = ( xtot((30e3-L):(34e3+L)) - mu ) / s ;
-end
-
-N = length(x);
-t = linspace(0, (N-1)/fs, N);
-
-figure;
-plot(t,x,'linewidth',2);
-set(gca,'fontsize',16);
-xlabel('Time (s)'); ylabel('PPG signal'); grid on; axis tight;
-
-%% Influence of M (Fix K) 
-HOP = 1 ;
-Mh = round( linspace(2,5*L,100) );
-extK = 10*L ;
-
-method.name = 'lse' ;
-m = 1;
-for extM = Mh
-    xxLSE = SigExtension(x,fs,HOP,extK,extM,extSEC,method);
-%     xxLSEV = SigExtension(x,fs,HOP,extK,extM,extSEC,'lseV');
-%     xxDMD = SigExtension(x,fs,HOP,extK,extM,extSEC,'dmd');
-    forecastErrLSE_m(m) = sqrt( (1/(2*L)) * sum(abs(xxLSE - xxTRUE).^2) ) ;
[…]

### 2020-08-28 Notes/Supplementary material/SupMat.tex
sup mat

@@
 
-\setcounter{equation}{27}
+\setcounter{equation}{25}
 \setcounter{figure}{7}
@@
 \subsection{Preliminaries}
+
+\subsubsection{Notations}
 Recall the model~\eqref{I-eq:model.noise}. Based on the definition of matrices $\bX$ and $\bY$, we have:
@@
 \end{align*}
-Thus, 
-\begin{align*}
-\bA_0 :=\bS^{(1)}{\bS^{(0)}}^\inv\,,\quad\tilde\bA := (\bS^{(1)}+\bE^{(1)})(\bS^{(0)}+\bE^{(0)})^\inv\,.
-\end{align*}
-As a result, for $\ell\in \mathbb{N}$, 
+Thus, the matrix $\tilde\bA$, defined in equation~\eqref{I-eq:lse}, is expressed in the form:
+\begin{equation*}
+\tilde\bA = (\bS^{(1)}+\bE^{(1)})(\bS^{(0)}+\bE^{(0)})^\inv\,.
+\end{equation*}
+Let $\bA_0$ denote the deterministic matrix such that $\bA_0 \defeq \bS^{(1)}{\bS^{(0)}}^\inv$.
+We denote by $\balpha^{(\ell)}_0$ the last row of $\bA_0^\ell$.
+As a result, for $\ell\in \mathbb{N}^*$, the error vector $\bh^{(\ell)}$ defined by $\bh^{(\ell)}\defeq \balpha^{(\ell)} - \balpha^{(\ell)}_0$ satisfy the equation
 \begin{align}
 \nonumber
-\bh^{(\ell)} &= \balpha^{(\ell)} - \balpha^{(\ell)}_0 \\
-\nonumber
-&= \be_M^T\left(\tilde\bA^\ell-\bA_0^\ell\right) \\
+\bh^{(\ell)} &= \be_M^T\left(\tilde\bA^\ell-\bA_0^\ell\right) \\
 &= \be_M^T\left(\left((\bS^{(1)}+\bE^{(1)})(\bS^{(0)}+\bE^{(0)})^\inv\right)^\ell- \bA_0^\ell\right)\ .
@@
 
-%\subsection{Study of $\bh^{(\ell)}$}
 The randomness of $\bh^{(\ell)}$ completely comes from the error matrices.
 Besides, notice that the first $M-1$ rows in $\bE^{(1)}$ equal to the last $M-1$ rows of $\bE^{(0)}$.
-We gather all sources of randomness into an vector $\bg\in\RR^{M(M+1)}$, containing $M$ rows defined as
+We gather all sources of randomness into a vector $\bg\in\RR^{M(M+1)}$, defined as
 \begin{equation}
@@
 \end{equation}
-where "$\vec$" denotes the vectorization operator, concatenating the columns of a given matrix on top of one another.
+where "$\vec$" denotes the vectorization operator, that concatenates the columns of a given matrix on top of one another.
 Then, we can write $\bh^{(\ell)}$ as $\bh^{(\ell)}=f^{(\ell)}(\bg)$ where $f^{(\ell)}$ is a deterministic function such that:
@@
 \end{align*}
+
+In the following paragraphs, we provide some useful results to prove Theorem 1.
+
+\subsubsection{Expression of $\bA_0$}
+It follows from the signal model~\eqref{I-eq:sum.sine} that matrices $\bS^{(0)}$ and $\bS^{(1)}$ take the following form:
+\begin{align}
+\nonumber
+\bS^{(a)}[m,m'] &= \sigma^2\delta_{(m+a)m'}+\sum_{j,j'=1}^J\dfrac{\Omega_j\Omega_{j'}}{K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(N_0+m+a+k)+\varphi_j\right)\cos\left(2\pi \frac{f_{j'}}{\fs}(N_0+m'+k)+\varphi_{j'}\right) \\
+\nonumber
+&= \sigma^2\delta_{(m+a)m'}+\sum_{j=1}^J\dfrac{\Omega_j^2}{2K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)\\
+\nonumber
+&= \sigma^2\delta_{(m+a)m'}+\sum_{j=1}^J\Bigg( \dfrac{\Omega_j^2}2\cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \dfrac{\Omega_j^2}{2K}\underbrace{\sum_{k=0}^{K-1}\cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)}_{=0\ \mathrm{because}\ \frac{f_j}\fs=\frac{p'_j}K } \Bigg)\\
+&= \sigma^2\delta_{(m+a)m'}+\sum_{j=1}^J\dfrac{\Omega_j^2}2\cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right)\ .
+\label{eq:S1}
+\end{align}
+Thus, $\bS^{(0)}$ is a circulant matrix, and is therefore diagonalizable in the Fourier basis:
+\[
+\bS^{(0)} = \bU\bLambda^{(0)}\bU^*\ ,
+\]
+where $\bU[m,m']=\frac1{\sqrt{M}}e^{-2\ii\pi mm'/M}$ and $\bLambda^{(0)} = \mathrm{diag}(\lambda_0^{(0)},\dots,\lambda_{M-1}^{(0)})$ with
+\begin{align*}
+\lambda_m^{(0)} &= \sigma^2 + \sum_{j=1}^J\dfrac{\Omega_j^2}2\sum_{q=0}^{M-1} \cos\left(2\pi\frac{f_j}{\fs} q\right) e^{-2\ii\pi qm/M} \\
+&= \sigma^2 + \dfrac{M}{4}\sum_{j=1}^J\Omega_j^2(\delta_{m,p_j} + \delta_{m,M-p_j})\ .
+\end{align*}
+Therefore,
+\begin{align*}
+{\bS^{(0)}}^\inv  &= \bU{\bLambda^{(0)}}^\inv\bU^*\,,
+\end{align*}
+which leads to
+\begin{equation}
+{\bS^{(0)}}^\inv[m,m']  = \dfrac1{\sigma^2}\delta_{m,m'}-\dfrac{2}{M\sigma^2}\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j \dfrac{m-m'}{M}\right)\,.
+\label{eq:S0.inv}
+\end{equation}
+Consequently, combining equations~\eqref{eq:S1} and~\eqref{eq:S0.inv}, we have
+\begin{align}
+\nonumber
+\bA_0[m,m']  &= \sum_{q=0}^{M-1} {\bS^{(1)}}[m,q]{\bS^{(0)}}^\inv[q,m'] \\
+&= \delta_{m+1,m'} + \dfrac2M\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j\frac{m'}{M}\right)\delta_{m+1,M}\,.
+\label{eq:A0.sine}
+\end{align}
 
@@
 \end{align}
-where $C_\bz = 2\left(\sum_{j=1}^J \Omega_j\right)$. Besides, since odd-order moments of a zero-mean multivariate Gaussian random vector are zero, we have:
+where $C_\bz \defeq 2\left(\sum_{j=1}^J \Omega_j\right)$. Besides, since odd-order moments of a zero-mean multivariate Gaussian random vector are zero, we have:
 \begin{align}
@@
 &\hspace{30pt} - \delta_{m_{r'}+a_{r'},m'_{r'}}\bE\{\bg_1[r]\} \\
-& = 0\ ,
+& = 0\ .
 \label{eq:g1g2}
@@
 \subsubsection{Study of $f^{(\ell)}$}
-In this section, we provide bounds of the first and second-order derivatives of $f^{(\ell)}$ at the origin.
+In this section, we provide bounds on the first and second-order derivatives of $f^{(\ell)}$ at the origin.
 
@@
 \begin{equation}
-\nonumber
 \left.\dfrac{\partial f^{(\ell)}}{\partial \bg[r]}\right|_{\bg=\bzero} = \sum_{\lambda=0}^{\ell-1} \be_M^T\bA_0^\lambda\left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero}\bA_0^{\ell-1-\lambda}\ .
@@
 \end{align*}
-where $\bJ_{m_r,m'_r}\in\RR^{M\times M}$ is the matrix such that $\bJ_{m_r,m'_r}[m,m'] = \delta_{m,m_r}\delta_{m',m'_r}$.
+where $\bJ_{m_r,m'_r}\in\RR^{M\times M}$ is the matrix such that $\bJ_{m_r,m'_r}[m,m'] \defeq \delta_{m,m_r}\delta_{m',m'_r}$.
 Thus,
 \begin{align*}
-\left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=0} &= \left\{
+\left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero} &= \left\{
 \begin{array}{ll}
@@
 \end{align*}
-Let $\|\cdot\|_\infty$ denotes the uniform norm of a vector or a matrix, \ie $\|\bM\|_\infty = \max_{n,n'} \bM[n,m']$. Then,
+Let $\|\cdot\|_\infty$ denote the uniform norm of a vector or a matrix, \ie~$\|\bM\|_\infty = \max_{n,n'} \bM[n,n']$. Then,
 \begin{equation*}
-\left\| \left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=0}\right\|_\infty \leq \left(1+\|\bA_0\|_\infty\right)\|{\bS^{(0)}}^{-1}\|_\infty\ .
+\left\| \left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero}\right\|_\infty \leq \left(1+\|\bA_0\|_\infty\right)\left\|{\bS^{(0)}}^{-1}\right\|_\infty\ .
 \end{equation*}
@@
 \end{equation}
-and from~\eqref{eq:S0.inv}, we have that $\|{\bS^{(0)}}^{-1}\|_\infty\leq \dfrac1{\sigma^2}\left(1+\frac{2J}{M}\right)$. Then,
+and from~\eqref{eq:S0.inv}, we have that $\left\|{\bS^{(0)}}^{-1}\right\|_\infty\leq \dfrac1{\sigma^2}\left(1+\frac{2J}{M}\right)$. Then,
 \begin{equation}
-\left\| \left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=0}\right\|_\infty \leq \dfrac{C_{\bz,M}}{\sigma^2}\ ,
+\left\| \left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero}\right\|_\infty \leq \left(1+\max\left(1,\frac{2J}{M}\right)\right)\left(1+\frac{2J}{M}\right)\dfrac{1}{\sigma^2}\ .
 \label{eq:dA}
 \end{equation}
-where $C_{\bz,M}=\left(1+\max\left(1,\frac{2J}{M}\right)\right)\left(1+\frac{2J}{M}\right)$. Besides given expression~\eqref{eq:df}, for all $r\in\{0,\ldots,M(M+1)-1\}$, we have:
+Besides given expression~\eqref{eq:df}, for all $r\in\{0,\ldots,M(M+1)-1\}$, we have:
 \begin{align*}
-\left|\left.\dfrac{\partial f_m^{(\ell)}}{\partial \bg[r]}\right\vert_{\bg=\bzero}\right| &\leq d_{1,M,\ell} \|\bA_0\|_\infty^{\ell-1}\left\|\dfrac{\partial\tilde\bA}{\partial \bg[r]}\right\|_\infty\ .
+\left|\left.\dfrac{\partial f_m^{(\ell)}}{\partial \bg[r]}\right\vert_{\bg=\bzero}\right| &\leq 2M\|\bA_0^{\ell-1}\|_\infty \left\| \left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero}\right\|_\infty + M^2\sum_{\lambda=1}^{\ell-2} \|\bA_0^\lambda\|_\infty \left\| \left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero}\right\|_\infty  \|\bA_0^{\ell-1-\lambda}\|_\infty \\
+&\leq \left(2 + (\ell-2)M\right)M^{\ell-1} \|\bA_0\|_\infty^{\ell-1}\left\|\left.\dfrac{\partial\tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero}\right\|_\infty\ .
 \end{align*}
-where $d_{1,M,\ell}$ is only depending on $M$ and $\ell$.
 Therefore, given bounds~\eqref{eq:A0} and~\eqref{eq:dA}, we have:
@@
 \end{equation}
-where $d_{1,\bz,M,\ell}=d_{1,M,\ell}\left(1+\frac{2J}{M}\right)\left(\max\left(1,\frac{2J}{M}\right)\right)^{\ell}$.
+where $d_{1,\bz,M,\ell}\defeq \left(2 + (\ell-2)M\right)M^{\ell-1}\left(\max\left(1,\frac{2J}{M}\right)\right)^{\ell-1}\left(1+\max\left(1,\frac{2J}{M}\right)\right)\left(1+\frac{2J}{M}\right)$.
 
@@
 \begin{array}{ll}
-\|{\bS^{(0)}}^{-1}\|_\infty^2 &\mbox{if}\quad a_r=1\quad\mbox{or}\quad a_{r'}=1\,, \\
-2\left(1+\|\bA_0\|_\infty\right)\|{\bS^{(0)}}^{-1}\|_\infty^2 &\mbox{else} \\
+\left\|{\bS^{(0)}}^{-1}\right\|_\infty^2 &\mbox{if}\quad a_r=1\quad\mbox{or}\quad a_{r'}=1\,, \\
+2\left(1+\|\bA_0\|_\infty\right)\left\|{\bS^{(0)}}^{-1}\right\|_\infty^2 &\mbox{else} \\
 \end{array}
@@
 where $d_{2,M,\ell}$ and $d'_{2,M,\ell}$ are only depending on $M$ and $\ell$.
-Besides, given results~\eqref{eq:dA} and~\eqref{eq:d2A}, we have:
+Besides, given results~\eqref{eq:A0}, ~\eqref{eq:dA} and~\eqref{eq:d2A}, we have:
 \begin{equation}
@@
 \begin{align*}
-d_{2,\bz,M,\ell} &= d_{2,M,\ell}C_{\bz,M}^2\max\left(1,\frac{2J}{M}\right)^{\ell-2} + d'_{2,M,\ell}\max\left(1,\frac{2J}{M}\right)^{\ell-1}2\left(1+\max\left(1,\frac{2J}{M}\right)\right)\left(1+\frac{2J}{M}\right)^2\ .
+d_{2,\bz,M,\ell} &= d_{2,M,\ell}C_{\bz,M}^2\max\left(1,\frac{2J}{M}\right)^{\ell-2} + 2d'_{2,M,\ell}\max\left(1,\frac{2J}{M}\right)^{\ell-1}\left(1+\max\left(1,\frac{2J}{M}\right)\right)\left(1+\frac{2J}{M}\right)^2\ .
 \end{align*}
@@
 Outside the measurement interval $I$, denote by $\ell$ the index such that $n=N-1+\ell$.
-Then, given that $\bh^{(\ell)} = \balpha^{(\ell)} - \balpha_0^{(\ell)}$, we have
+Then, given that $\bh^{(\ell)} = \balpha^{(\ell)} - \balpha_0^{(\ell)}$, we deduce from expression~\eqref{I-eq:bias0} that
 \begin{align}
@@
 \label{eq:eps1.def}
-\bepsilon_1[\ell]&=  \balpha_0^{(\ell)}\bz_{K}  - \bz[N-1+\ell] \\
+\bepsilon_1[\ell]&=  \balpha_0^{(\ell)}\bz_{K}  - \bz[N-1+\ell]\,, \\
 \label{eq:eps2.def}
-\bepsilon_2[\ell] &=  \EE\{\bh^{(\ell)}\}\bz_{K}  \\
+\bepsilon_2[\ell] &=  \EE\{\bh^{(\ell)}\}\bz_{K}\,,  \\
 \label{eq:eps3.def}
@@
 
-Let us first determine an upper bound on $\left|\bepsilon_1[n]\right|$. We have
-\begin{align}
-\nonumber
-\bS^{(a)}[m,m'] &= \sigma^2\delta_{(m+a)m'}+\sum_{j,j'=1}^J\dfrac{\Omega_j\Omega_{j'}}{K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(N_0+m+a+k)+\varphi_j\right)\cos\left(2\pi \frac{f_{j'}}{\fs}(N_0+m'+k)+\varphi_{j'}\right) \\
-\nonumber
-&= \sigma^2\delta_{(m+a)m'}+\sum_{j=1}^J\dfrac{\Omega_j^2}{2K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)\\
-\nonumber
-&= \sigma^2\delta_{(m+a)m'}+\sum_{j=1}^J\Bigg( \dfrac{\Omega_j^2}2\cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \dfrac{\Omega_j^2}{2K}\underbrace{\sum_{k=0}^{K-1}\cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)}_{=0\ \mathrm{because}\ \frac{f_j}\fs=\frac{p'_j}K } \Bigg)\\
-&= \sigma^2\delta_{(m+a)m'}+\sum_{j=1}^J\dfrac{\Omega_j^2}2\cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right)\ .
-\label{eq:S1}
-\end{align}
-Thus, $\bS^{(0)}$ is a circulant matrix, and is therefore diagonalizable in the Fourier basis:
-\[
-\bS^{(0)} = \bU\bLambda^{(0)}\bU^*\ ,
-\]
-where $\bU[m,m']=\frac1{\sqrt{M}}e^{-2\ii\pi mm'/M}$ and $\bLambda^{(0)} = \mathrm{diag}(\lambda_0^{(0)},\dots,\lambda_{M-1}^{(0)})$ with
-\begin{align*}
-\lambda_m^{(0)} &= \sigma^2 + \sum_{j=1}^J\dfrac{\Omega_j^2}2\sum_{q=0}^{M-1} \cos\left(2\pi\frac{f_j}{\fs} q\right) e^{-2\ii\pi qm/M} \\
-&= \sigma^2 + \dfrac{M}{4}\sum_{j=1}^J\Omega_j^2(\delta_{m,p_j} + \delta_{m,M-p_j})\ .
-\end{align*}
-Therefore,
-\begin{align*}
-{\bS^{(0)}}^\inv  &= \bU{\bLambda^{(0)}}^\inv\bU^*\,,
-\end{align*}
-which leads to
-\begin{equation}
-{\bS^{(0)}}^\inv[m,m']  = \dfrac1{\sigma^2}\delta_{m,m'}-\dfrac{2}{M\sigma^2}\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j \dfrac{m-m'}{M}\right)\,.
-\label{eq:S0.inv}
-\end{equation}
-Consequently, combining equations~\eqref{eq:S1} and~\eqref{eq:S0.inv}, we have
-\begin{align}
-\nonumber
-\bA_0[m,m']  &= \sum_{q=0}^{M-1} {\bS^{(1)}}[m,q]{\bS^{(0)}}^\inv[q,m'] \\
-&= \delta_{m+1,m'} + \dfrac2M\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j\frac{m'}{M}\right)\delta_{m+1,M}\,.
-\label{eq:A0.sine}
-\end{align}
-Since $\balpha_0^{(1)}$ is the last row of $\bA_0$, we have
+Let us first determine an upper bound on $\left|\bepsilon_1[n]\right|$.  Since $\balpha_0^{(1)}$ is the last row of $\bA_0$, we deduce from the expression~\eqref{eq:A0.sine} of $\bA_0$ that
 \begin{equation}
@@
 \end{equation}
-Thus, given the bounds~\eqref{eq:gg.bound} on $\EE\{\bg[r]\bg[r']\}$ and~\eqref{eq:d2f.bound} on the second derivative of $f_m^{(\ell)}$, we have:
+Thus, given the bounds~\eqref{eq:gg.bound} on $\left|\EE\{\bg[r]\bg[r']\}\right|$ and~\eqref{eq:d2f.bound} on the second derivative of $f_m^{(\ell)}$, we have:
 \begin{align}
@@
 \begin{align*}
-c^{(2)}_1 &=\dfrac{M^3(M+1)^2}{2}C_\bz d_{2,\bz,M,\ell}\ ,\\
-c^{(2)}_2 &=\dfrac{M^3(M+1)^2}{4}C_\bz^3 d_{2,\bz,M,\ell} \ .
+c^{(2)}_1 &\defeq\dfrac{M^3(M+1)^2}{2}C_\bz d_{2,\bz,M,\ell}\ ,\\
+c^{(2)}_2 &\defeq\dfrac{M^3(M+1)^2}{4}C_\bz^3 d_{2,\bz,M,\ell} \ .
 \end{align*}
@@
 \end{equation}
-where $ c^{(3)} = M^2(M+1)d_{1,\bz,M,\ell}C_\bz$.
+where $ c^{(3)} \defeq M^2(M+1)d_{1,\bz,M,\ell}C_\bz$.
 
@@
 & \eta_6[n] &= \sigma^2\EE\left\{\left(\bh^{(\ell)}\bw_K\right)^2\right\}\,, \\
-\eta_7[n] &= 2\sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\ell)}\bw_K\} \ .\\ 
+\eta_7[n] &= 2\sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\ell)}\bw_K\} \ . 
 \end{align*}
@@
 
-First, since $\|\balpha_0^{(\ell}\|^2\leq M\|\balpha_0^{(\ell)}\|_\infty^2\leq M\|\bA_0^{\ell}\|_\infty^2$, we have:
+First, since $\|\balpha_0^{(\ell)}\|^2\leq M\|\balpha_0^{(\ell)}\|_\infty^2\leq M\|\bA_0^{\ell}\|_\infty^2$, we have:
 \begin{align}
@@
 \label{eq:sum5}
-\eta_5[n] &= o\left(\dfrac1K\right) \\
+\eta_5[n] &= o\left(\dfrac1K\right)\,, \\
 \label{eq:sum6}
@@
 \begin{align*}
-\left|\EE\{\bg_2^{(\ell)}[r]\bw_K[m]\bw_K[m']\}\right| &= \dfrac1K\sum_{k,=0}^{K-1} \EE\left\{\bw_{k}[m_{r}+a_{r}]\bw_{k}[m'_{r}]\bw_K[m]\bw_K[m']\right\} - \delta_{m_r+a_r,m'_r} \EE\left\{\bw_K[m]\bw_K[m']\right\} \\
-&\leq \dfrac4K\ .
+\left|\EE\{\bg_2^{(\ell)}[r]\bw_K[m]\bw_K[m']\}\right| &= \dfrac1K\left|\sum_{k=0}^{K-1} \EE\left\{\bw_{k}[m_{r}+a_{r}]\bw_{k}[m'_{r}]\bw_K[m]\bw_K[m']\right\} - \delta_{m_r+a_r,m'_r} \EE\left\{\bw_K[m]\bw_K[m']\right\}\right| \\
+&= \dfrac1K\left|\sum_{k=0}^{K-1} \delta_{k+m_{r}+a_{r},K+m} \delta_{k+m'_{r},K+m'} +  \delta_{k+m_{r}+a_{r},K+m'} \delta_{k+m'_{r},K+m}\right| \\
+&\leq \dfrac2K\ .
 \end{align*}
@@
 \nonumber
-\left| \eta_7[n] \right| &\leq 2\sigma^2\|\balpha_0^{(\ell)}\|_\infty M^3(M+1)\dfrac{d_{1,\bz,M,\ell}}{\sigma^2} \dfrac{4\sigma^2}K + o\left( \dfrac1K\right)\\
-&\leq 8 M^{\ell+2}(M+1)d_{1,\bz,M,\ell}\left(\max\left(1,\frac{2J}{M}\right)\right)^\ell\dfrac{\sigma^2}{K} + o\left( \dfrac1K\right)\ .
+\left| \eta_7[n] \right| &\leq 2\sigma^2\|\balpha_0^{(\ell)}\|_\infty M^3(M+1)\dfrac{d_{1,\bz,M,\ell}}{\sigma^2} \dfrac{2\sigma^2}K + o\left( \dfrac1K\right)\\
+&\leq 4 M^{\ell+2}(M+1)d_{1,\bz,M,\ell}\left(\max\left(1,\frac{2J}{M}\right)\right)^\ell\dfrac{\sigma^2}{K} + o\left( \dfrac1K\right)\ .
 \label{eq:sum7}
@@
 c_2^{(n)} & = \dfrac12C_\bz^2 M^2(M+1)^2d_{1,\bz,M,\ell}^2 + M^\ell(M+1)C_\bz^2 d_{1,\bz,M,\ell}\left(\max\left(1,\frac{2J}{M}\right)\right)^\ell\\
-c_3^{(n)} & =  8 M^{\ell+2}(M+1)d_{1,\bz,M,\ell}\left(\max\left(1,\frac{2J}{M}\right)\right)^\ell\ .
+c_3^{(n)} & =  4 M^{\ell+2}(M+1)d_{1,\bz,M,\ell}\left(\max\left(1,\frac{2J}{M}\right)\right)^\ell\ .
 \end{align*}
@@
 \begin{align*}
-\left|\gamma[n,n]\right| & \leq  \sqrt{\gamma[n,n]\gamma[n',n']} \\
+\left|\gamma[n,n']\right| & \leq  \sqrt{\gamma[n,n]\gamma[n',n']} \\
 & \leq  \sqrt{\gamma[n,n]\gamma[n',n']} \\
-& \leq   \sqrt{c_0^{(n)}c_0^{(n')}}\sigma^2\sqrt{1 + \dfrac{1}{K\sigma^2} \left( \dfrac1{c_0^{(n)}}\left( \dfrac{c_1^{(n)}}{\sigma^2} + c_2^{(n)} + c_3^{(n)}\sigma^2\right) + \dfrac1{c_0^{(n')}}\left( \dfrac{c_1^{(n')}}{\sigma^2} + c_2^{(n')} + c_3^{(n')}\sigma^2\right)\right) + o\left(\dfrac1K\right)} \\
+& \leq   \sqrt{c_0^{(n)}c_0^{(n')}}\sigma^2\sqrt{1 + \dfrac{1}{K\sigma^2} \left( \dfrac1{c_0^{(n)}}\left( \dfrac{c_1^{(n)}}{\sigma^2} + c_2^{(n)} + c_3^{(n)}\sigma^2\right) + \dfrac1{c_0^{(n')}}\left( \dfrac{c_1^{(n')}}{\sigma^2} + c_2^{(n')} + c_3^{(n')}\sigma^2\right)\right) + o\left(\dfrac1K\right)}\ .
+\end{align*}
+A first-order Taylor expansion of this bound as $K\to\infty$ gives
+\begin{align*}
 & \leq   \sqrt{c_0^{(n)}c_0^{(n')}}\sigma^2 + \dfrac{1}{2K} \left( \sqrt{\dfrac{c_0^{(n')}}{c_0^{(n)}}}\left( \dfrac{c_1^{(n)}}{\sigma^2} + c_2^{(n)} + c_3^{(n)}\sigma^2\right) + \sqrt{\dfrac{c_0^{(n)}}{c_0^{(n')}}}\left( \dfrac{c_1^{(n')}}{\sigma^2} + c_2^{(n')} + c_3^{(n')}\sigma^2\right)\right) + o\left(\dfrac1K\right) \\
@@
 \begin{align*}
-\beta_1[n,n'] &= \sigma\EE\{\bw[n']\bh^{(\ell)}\}\bz_K \\
-\beta_2[n,n'] &= \sigma^2\EE\{\bw[n']\balpha_0^{(\ell)}\bw_K\} \\
+\beta_1[n,n'] &= \sigma\EE\{\bw[n']\bh^{(\ell)}\}\bz_K\,, \\
+\beta_2[n,n'] &= \sigma^2\EE\{\bw[n']\balpha_0^{(\ell)}\bw_K\}\,, \\
 \beta_3[n,n'] &= \sigma^2\EE\{\bw[n']\bh^{(\ell)}\bw_K\}\ .
@@
 \nonumber
-\beta_2[n,n'] &\leq \sigma^2\left\|\balpha_0^{(\ell)}\right\|_\infty\sum_{m=0}^{M-1}\left|\EE\{\bw[n']\bw_K[m]\}\right| \\
+\left|\beta_2[n,n']\right| &\leq \sigma^2\left\|\balpha_0^{(\ell)}\right\|_\infty\sum_{m=0}^{M-1}\left|\EE\{\bw[n']\bw_K[m]\}\right| \\
 \nonumber
@@
 \nonumber
-\beta_3[n,n'] &\leq \sigma^2\sum_{m=0}^{M-1}\left|\EE\{\bw[n']\bh^{(\ell)}[m]\bw_K[m]\}\right|\\
+\left|\beta_3[n,n']\right| &\leq \sigma^2\sum_{m=0}^{M-1}\left|\EE\{\bw[n']\bh^{(\ell)}[m]\bw_K[m]\}\right|\\
 \nonumber

### 2020-09-02 Notes/BoundEffRed.tex
code
edit paper

@@
 
-\title{An Efficient Forecasting Approach for the Real-Time Reduction of Boundary Effects\\ in Time-Frequency Representations}
+\title{An Efficient Forecasting Approach\\ to Reduce Boundary Effects in Real-Time Time-Frequency Analysis}
 
@@
 In the first class, judicious choice of analysis window whose support does not interfere with the boundary points can minimize the occurrence of aberrant patterns near the boundaries of the TF plane, for instance, \cite{Chui92wavelets,Depczynski99fast}.
-Due to the specific relationship of chosen analysis windows and TF analysis tool, these techniques do not make it possible to reduce the boundary effects of any TF representations.
+Due to the specific relationship of chosen analysis windows and TF analysis tool, these techniques do not make it possible to reduce the boundary effects of all TF representations.
 %
@@
 In the physically relevant dynamical models, the oscillation or trend, are usually modeled as the mean of a stochastic process, while in the stochastic models, the oscillation or trend are modeled by the covariance structure of the stochastic process.
-It is also possible to consider polynomial regression \cite{fan1996local} or modeling the mean by splines \cite{hall2005theory}, kernel functions \cite{chang2010training}, and wavelets \cite{marron1998exact}, or nonparametric regression \cite{fan1996local}, to estimate the mean function and hence forecasting the signal.
-Neural network models, such as the long short-term memory model \cite{vlachas2018data}, is another approach.
+It is also possible to consider polynomial regression~\cite{fan1996local} or modeling the mean by splines~\cite{hall2005theory}, kernel functions~\cite{chang2010training}, and wavelets~\cite{marron1998exact}, or nonparametric regression~\cite{fan1996local}, to estimate the mean function before forecasting the signal.
+Neural network models, such as the long short-term memory model~\cite{vlachas2018data}, is another approach.
 %
@@
 We claim and prove that forecasting oscillatory signals based on the simple dynamical model combined with the simple least square approach is sufficient for reducing the boundary effects for the TF analysis, or other kernel-based analysis.
+To our knowledge, such theoretical analysis does not exist in the literature.
 See the bottom right of Fig.~\ref{fig:ex.intro} for a snapshot of the result.

### 2020-09-02 Notes/algorithm.tex
edit paper

@@
 
-\paragraph{Dynamical Model and Forecasting} 
-Establishing a dynamical model consists in determining the relation linking $\bY$ to $\bX$, that is finding a function $f$ so that
-\[
+\paragraph{Dynamical Model} 
+Establishing a dynamical model consists of determining the relation linking $\bY$ to $\bX$, that is finding a function $f$ so that
+\begin{equation}
 \bY = f(\bX)\ .
-\]
+\label{eq:generic.model}
+\end{equation}
 In a general framework, forecasting means estimating the function $f$ from the observed values taken by the signal, in order to predict its future values.
@@
 \end{equation}
+From the Takens' embedding perspective~\cite{Takens81detecting}, $\bX$ can be viewed as a lag map.
+Under the manifold assumption of the intrinsic phase space that hosts the dynamics, $f$ describes the dynamics supported on the manifold space, and $\bA$ approximates the dynamics on the manifold.
 
+\paragraph{Forecasting}
 The forecasting method consists in estimating the unknown matrix $\bA$.

### 2020-09-02 Scripts/EvalPerfSigExt/InfluenceNoiseSine.m

@@ -39,15 +39,6 @@ end
 xx0 = xx00 + xx01 ; % extended signal
 x0 = xx0(1:N) ; % restriction to the measurement interval
 
-%% Ideal Forecasting vector
-
-X0 = [] ; Y0 = [] ;
-for kk = 1: extK
-    X0(:,kk) = x0(end-extK-extM+kk: HOP: end-extK+kk-1) ;
-    Y0(:,kk) = x0(end-extK-extM+kk+1: HOP: end-extK+kk) ;
-end
-zK = x0(end-extM+1 : HOP: end).' ;
-
 %% Forecasting
 method.name = 'SigExt' ;
 nbXP = 200 ; % number of noise levels

### 2020-09-02 Scripts/EvalPerfSigExt/ResultsInfluencesSine.m

@@ -13,7 +13,7 @@ loglog(Sigma.^2,abs(BiasXPm(:,1)),Sigma.^2,abs(BiasXPm(:,10)),Sigma.^2,abs(BiasX
 axis tight; ylim([1e-10 1e3]); grid on;
 xlabel('Noise Variance $\sigma^2$','interpreter','latex') ; ylabel('Experimental Bias $\mu_{\mathrm{xp}}$','interpreter','latex') ;
 legend({'$\ell=1$','$\ell=10$','$\ell=100$'},'interpreter','latex') ;
-set(gca,'fontsize',18) ;
+set(gca,'fontsize',24) ;
 
 figure;
 % subplot(2,1,2) ;
@@ -21,7 +21,7 @@ loglog(Sigma.^2,VarXPm(:,1),Sigma.^2,VarXPm(:,10),Sigma.^2,VarXPm(:,100),'linewi
 axis tight; ylim([1e-15 1e3]); grid on;
 xlabel('Noise Variance $\sigma^2$','interpreter','latex') ; ylabel('Experimental Variance $\gamma_{\mathrm{xp}}$','interpreter','latex') ;
 legend({'$\ell=1$','$\ell=10$','$\ell=100$'},'interpreter','latex') ;
-set(gca,'fontsize',18) ;
+set(gca,'fontsize',24) ;
 
 %% Influence of the dataset size
 load('../../Results/PerfSizeDataset.mat')
@@ -32,7 +32,7 @@ loglog(KK,abs(BiasXPm(:,1)),KK,abs(BiasXPm(:,10)),KK,abs(BiasXPm(:,100)),'linewi
 axis tight; ylim([1e-8 1e0]); grid on;
 xlabel('Dataset Size $K$','interpreter','latex') ; ylabel('Experimental Bias $\mu_{\mathrm{xp}}$','interpreter','latex') ;
 legend({'$\ell=1$','$\ell=10$','$\ell=100$'},'interpreter','latex') ; xticks([300 500 1000 2000]);
-set(gca,'fontsize',20) ;
+set(gca,'fontsize',24) ;
 
 figure;
 % subplot(2,1,2) ;
@@ -40,4 +40,4 @@ loglog(KK,VarXPm(:,1),KK,VarXPm(:,10),KK,VarXPm(:,100),'linewidth',2);
 axis tight; grid on;
[…]

### 2020-09-03 Notes/BoundEffRed.tex
update paper

@@
 A preliminary step should then be added to the algorithm to detect signal activity, and disable the forecasting step when possible.
-A more fundamental improvement would be to perform the forecasting step directly in the TF plane, to specifically predict the evolution of each instantaneous frequency.
-This leads to seeing a TF representation as a multivariate time series, with each frequency band containing a univariate time series.
+More fundamentally, one can also consider accelerating the computational time by optimizing the forecasting step, \ie~reduce $t_{\mathrm{forecast}}$.
+Indeed, each new forecast requires the inversion of the matrix $\bX\bX^T$ of size $M\times M$.
+However, the matrix $\bX$ used to forecast $\bx_{N+H}$ at the current iteration differs from the one used at the previous iteration to forecast $\bx_{N}$ by only $H$ columns.
+It then seems natural to take inspiration from the work of Strobach~\cite{Strobach97square}, generalized by Badeau \etal~\cite{Badeau04sliding}, which proposes a fast algorithm for the singular value decomposition of successive data matrices taking the same form as $\bX$.
+Their algorithm is limited to the case where $H=1$.
+Developing an extension of this algorithm to variations of $H>1$ columns would then allow to efficiently update the matrix $\left(\bX\bX^T\right)^{-1}$.
 These avenues will be explored in our future work.

### 2020-09-03 Notes/results.tex
update paper

@@
 This validates the result of Theorem~\ref{th:error}, highlighting the necessary presence of noise to ensure the numerical stability of the estimator.
-Furthermore, as soon as $\sigma$ is higher than this critical threshold, the estimator variance increases linearly with $\sigma^2$, as predicted by the bound~\eqref{eq:cov.error.2} provided in Theorem~\ref{th:error}.
+Furthermore, as soon as $\sigma^2$ is higher than this critical threshold, the estimator variance increases linearly with $\sigma^2$, as predicted by the bound~\eqref{eq:cov.error.2} provided in Theorem~\ref{th:error}.
 Note that the parameter $\ell$ has little influence on the quality of the estimator.
@@
 As anticipated by Theorem~\ref{th:error}, the variance of the estimator decreases linearly with $\frac1K$ as long as the dataset size is large enough.
-Indeed, in the log-log plot displayed in Fig.~\ref{fig:res.size.sine}, as soon as $K>700$, each of the three curves representing the forecasting variance tends towards an asymptote of slope $-1$.
-
-%\paragraph{Summary}
-%Both previous experimental results combined with the theoretical asymptotic equation~\eqref{eq:cov.error.2} allow us to describe the influence of the noise variance and the size of the training dataset on the variance of the forecasting noise, empirically summarized as follows:
-%\begin{equation}
-%\gamma[N-1+\ell,N-1+\ell] \underset{K\to\infty}{\approx} \dfrac{\sigma^2}{K}g[\ell] .
-%\end{equation} 
-%where $g$ is a bounded positive function. The empirical result is coherent with the theoretical result provided by Theorem~\ref{th:error}.
+Indeed, in the log-log plot displayed in Fig.~\ref{fig:res.size.sine}, as soon as $K>700$, each of the three curves representing the forecasting variance in function of $K$ tends towards an asymptote of slope $-1$.
 

### 2020-09-03 Notes/BoundEffRed.tex
update paper

@@
 A preliminary step should then be added to the algorithm to detect signal activity, and disable the forecasting step when possible.
-More fundamentally, one can also consider accelerating the computational time by optimizing the forecasting step, \ie~reduce $t_{\mathrm{forecast}}$.
+More fundamentally, one can also consider accelerating the computational time by optimizing the forecasting step to improve the real-time performance of {\sf BoundEffRed}.
 Indeed, each new forecast requires the inversion of the matrix $\bX\bX^T$ of size $M\times M$.

### 2020-09-08 Notes/Supplementary material/SupMat.tex
update paper
proof

@@
 
+\newtheorem{lemma}{Lemma}
+
 \input{../mydefs}
@@
 
-\setcounter{equation}{25}
+\setcounter{equation}{26}
 \setcounter{figure}{7}
@@
 \subsubsection{Notations}
+Considering $\bz\in\RR^N$ the deterministic signal defined in~\eqref{I-eq:sum.sine}, we denote by $\bz_k\in\RR^M$ the subsignal such that
+\[
+\bz_k[m] = \bz[N-K-M+k+m]\,,\quad \forall m\in\{0,\ldots,M-1\}\,,\ k\in\{0,\ldots,K\}\ .
+\]
+Define $\bZ\in\RR^{M \times K}$ and $\bZ'\in\RR^{M \times K}$, the matrices such that
+\begin{align*}
+\bZ &\defeq \begin{pmatrix}
+\bz_0 & \cdots & \bz_{K-1}
+\end{pmatrix}\\ 
+\bZ' &\defeq \begin{pmatrix}
+\bz_1 & \cdots & \bz_{K}
+\end{pmatrix}\,.
+\end{align*}
+Let $\bD\in\RR^{M\times M}$ be the matrix defined by $\bD[m,m']=\delta_{m+1,m'}$.
+
 Recall the model~\eqref{I-eq:model.noise}. Based on the definition of matrices $\bX$ and $\bY$, we have:
 \begin{align}
+\label{eq:XX.T}
 \dfrac1K\bX\bX^T &= \underbrace{\dfrac1K\bZ\bZ^T+\sigma^2\bI}_{\defeq S^{(0)}} + \bE^{(0)} \\
+\label{eq:YX.T}
 \dfrac1K\bY\bX^T &= \underbrace{\dfrac1K\bZ'\bZ^T +\sigma^2\bD}_{\defeq S^{(1)}}+ \bE^{(1)} \ ,
@@
 \end{equation*}
-Let $\bA_0$ denote the deterministic matrix such that $\bA_0 \defeq \bS^{(1)}{\bS^{(0)}}^\inv$.
+Define $\bA_0$ the deterministic matrix such that
+\begin{equation}
+\bA_0 \defeq \bS^{(1)}{\bS^{(0)}}^\inv\,.
+\label{eq:A0}
+\end{equation}
 We denote by $\balpha^{(\ell)}_0$ the last row of $\bA_0^\ell$.
@@
 Then, we can write $\bh^{(\ell)}$ as $\bh^{(\ell)}=f^{(\ell)}(\bg)$ where $f^{(\ell)}$ is a deterministic function such that:
-\begin{align*}
+\begin{align}
+\nonumber
 f^{(\ell)} : \RR^{M(M+1)} &\to \RR^{M} \\
  \bg &\mapsto \bh^{(\ell)}\ .
-\end{align*}
+\label{eq:f.ell}
+\end{align}
 
-In the following paragraphs, we provide some useful results to prove Theorem 1.
+In the following paragraph, we provide some useful lemmas to prove Theorem 1.
 
-\subsubsection{Expression of $\bA_0$}
+\subsubsection{Lemmas}
+
+\begin{lemma}[Expressions of $\bA_0$ and ${\bS^{(0)}}^{-1}$]
+Let $\bS^{(0)}$ be the $M\times M$ matrices defined in~\eqref{eq:XX.T}.
+Let $\bA_0$ the $M\times M$ matrices defined in~\eqref{eq:A0}.
+Assume the deterministic signal $\bz$ takes the form~\eqref{I-eq:sum.sine}, and the observed noisy signal takes the form~\eqref{I-eq:model.noise}.
+Then, the matrix $\bS^{(0)}$ is given by
+\begin{equation}
+{\bS^{(0)}}^\inv[m,m']  = \dfrac1{\sigma^2}\delta_{m,m'}-\dfrac{2}{M\sigma^2}\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j \dfrac{m-m'}{M}\right)\,,
+\label{eq:S0.inv}
+\end{equation}
+and the matrix $\bA_0$ is given by
+\begin{align}
+\bA_0[m,m']  &= \delta_{m+1,m'} + \dfrac2M\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j\frac{m'}{M}\right)\delta_{m+1,M}\,.
+\label{eq:A0.sine}
+\end{align}
+Let $\|\cdot\|_{\max}$ denote the maximum norm of a matrix, \ie~$\|\bM\|_{\max} = \max_{n,n'} \left|\bM[n,n']\right|$. Then,
+\begin{align}
+\label{eq:S0.inv.bound}
+\left\|{\bS^{(0)}}^\inv\right\|_{\max} &\leq \dfrac1{\sigma^2}\left(1+\frac{2J}{M}\right)\,, \\
+\left\|\bA_0\right\|_{\max} &\leq \max\left(1,\frac{2J}{M}\right)\ .
+\label{eq:A0.bound}
+\end{align}
+\end{lemma}
+
+\begin{proof}
 It follows from the signal model~\eqref{I-eq:sum.sine} that matrices $\bS^{(0)}$ and $\bS^{(1)}$ take the following form:
@@
 which leads to
-\begin{equation}
+\begin{equation*}
 {\bS^{(0)}}^\inv[m,m']  = \dfrac1{\sigma^2}\delta_{m,m'}-\dfrac{2}{M\sigma^2}\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j \dfrac{m-m'}{M}\right)\,.
-\label{eq:S0.inv}
-\end{equation}
-Consequently, combining equations~\eqref{eq:S1} and~\eqref{eq:S0.inv}, we have
-\begin{align}
-\nonumber
+\end{equation*}
+Directly, we have:
+\begin{align*}
+\left|{\bS^{(0)}}^\inv[m,m']\right|  &= \dfrac1{\sigma^2}\left|\delta_{m,m'}-\dfrac{2}{M}\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j \dfrac{m-m'}{M}\right)\right| \\
+&\leq \dfrac1{\sigma^2}\left( 1+\dfrac{2}{M}\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}} \right) \leq \dfrac1{\sigma^2}\left( 1+\dfrac{2J}{M} \right)\ .
+\end{align*}
+Thus, 
+\[
+\left\|{\bS^{(0)}}^\inv[m,m']\right\|_{\max}\leq\dfrac1{\sigma^2}\left( 1+\dfrac{2J}{M} \right)\ .
+\]
+Furthermore, combining equations~\eqref{eq:S1} and~\eqref{eq:S0.inv}, we have
+\begin{align*}
 \bA_0[m,m']  &= \sum_{q=0}^{M-1} {\bS^{(1)}}[m,q]{\bS^{(0)}}^\inv[q,m'] \\
 &= \delta_{m+1,m'} + \dfrac2M\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j\frac{m'}{M}\right)\delta_{m+1,M}\,.
-\label{eq:A0.sine}
-\end{align}
+\end{align*}
+Directly, we have:
+\begin{align*}
+\left|\bA_0[m,m']\right| &\leq \left\{
+\begin{array}{ll}
+1 & \mbox{if}\quad m<M-1\,, \\
+\dfrac{2}{M}\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}} & \mbox{if}\quad m=M-1\ . \\
+\end{array}
+\right.
+\end{align*}
+Thus, 
+\[
+\left\|\bA_0\right\|_{\max}\leq \max\left(1,\frac{2J}{M}\right)\ .
+\]
+\end{proof}
 
-\subsubsection{Study of $\bg$}
-In the following, for all $r\in\{0,\ldots,M(M+1)-1\}$ we note $\bg[r] = \sigma\bg_1[r] + \sigma^2\bg_2[r]$, where 
+\begin{lemma}[Moments of $\bg$]
+Let $\bg\in\RR^{M(M+1)}$ be the random vector defined in~\eqref{eq:vec}.
+Assume the deterministic signal $\bz$ takes the form~\eqref{I-eq:sum.sine}, and the observed noisy signal takes the form~\eqref{I-eq:model.noise}.
+Then, the second-order moments of $\bg$ are bounded as follows:
+\begin{equation}
+\left|\bE\{\bg[r]\bg[r']\}\right| \leq \dfrac1K \left( C_\bz^2\sigma^2 + 2 \sigma^4\right) + o\left(\dfrac1K\right)\,,\quad\forall (r,r')\in\{0,\ldots,M(M+1)-1\}^2\ ,
+\label{eq:gg.bound}
+\end{equation}
+where $C_\bz \defeq 2\left(\sum_{j=1}^J \Omega_j\right)$. Besides, higher-order moments of $\bg$ behave as $o\left( \dfrac1K \right)$.
+\end{lemma}
+
+\begin{proof}
+In the following, for all $r\in\{0,\ldots,M(M+1)-1\}$, we denote $\bg[r] = \sigma\bg_1[r] + \sigma^2\bg_2[r]$, where 
 \begin{align*}
-\bg_1[r] &= \bE_1^{(a_r)}[m_r,m'_r] \\
-\bg_2[r] &= \bE_2^{(a_r)}[m_r,m'_r] \ ,
+\bg_1[r] &= \bE_1^{(a_r)}[m_r,m'_r] = \dfrac1K\sum_{k=0}^{K-1} \bz_k[m_r+a_r]\bw_k[m'_r] + \bw_k[m_r+a_r]\bz_k[m'_r] \\
+\bg_2[r] &= \bE_2^{(a_r)}[m_r,m'_r] = \dfrac1K\sum_{k=0}^{K-1} \bw_k[m_r+a_r]\bw_k[m'_r] - \delta_{m_r+a_r,\,m'_r} \ ,
 \end{align*}
 and $m_r,m'_r, a_r$ are the corresponding coordinates of the matrices associated with $r$ through the vectorization operation~\eqref{eq:vec}.
-That is why, order-two moments of this random vector is split as follows:
+Thus, order-two moments of this random vector is split as follows:
 \begin{equation}
@@
 \end{equation}
-Then,
+By definition of the signal $\bz$ (see equation~\eqref{I-eq:sum.sine}), we have $|\bz[n]|\leq\sum_{j=1}^J\Omega_j$, for all $n\in\NN$. Thus, by a direct bound, we have
 \begin{align}
@@
 \nonumber
-& \leq \dfrac1{K^2}\left(\sum_{j=1}^J \Omega_j\right)^2 \sum_{k,k'=0}^{K-1} \EE\left\{\bw_k[m'_r]\bw_{k'}[m'_{r'}]\right\} + \EE\left\{\bw_k[m_r+a_r]\bw_{k'}[m'_{r'}]\right\} + \EE\left\{\bw_k[m'_r]\bw_{k'}[m_{r'}+a_{r'}]\right\}  \\
+& \leq \dfrac1{K^2}\left(\sum_{j=1}^J \Omega_j\right)^2 \sum_{k,k'=0}^{K-1} \left|\EE\left\{\bw_k[m'_r]\bw_{k'}[m'_{r'}]\right\}\right| + \left|\EE\left\{\bw_k[m_r+a_r]\bw_{k'}[m'_{r'}]\right\}\right|  \\
+&\hspace{110pt}  + \left|\EE\left\{\bw_k[m'_r]\bw_{k'}[m_{r'}+a_{r'}]\right\}\right| + \left|\EE\left\{\bw_k[m_r+a_r]\bw_{k'}[m_{r'}+a_{r'}]\right\}\right|
+\label{eq:sum.g1g1}
+\end{align}
+Besides, since $\bw$ is a white noise,
+\begin{align*}
+\dfrac1{K^2}\sum_{k,k'=0}^{K-1} \left|\EE\left\{\bw_k[m'_r]\bw_{k'}[m'_{r'}]\right\}\right| & = \dfrac1{K^2}\sum_{k,k'=0}^{K-1} \delta_{k+m'_r,\,k'+m'_{r'}}  \\
+&= \dfrac1{K^2}\left(K-\left|m'_r-m'_{r'}\right|\right) \ .
+\end{align*}
+Moreover, $0\leq\left|m'_r-m'_{r'}\right|\leq M-1$. Thus,
+\begin{align*}
+\dfrac1{K} - \dfrac{M-1}{K^2}\leq \dfrac1{K^2}\sum_{k,k'=0}^{K-1} \left|\EE\left\{\bw_k[m'_r]\bw_{k'}[m'_{r'}]\right\}\right| \leq \dfrac1{K}\ .
+\end{align*}
+Therefore,
+\begin{align*}
+\dfrac1{K^2}\sum_{k,k'=0}^{K-1} \left|\EE\left\{\bw_k[m'_r]\bw_{k'}[m'_{r'}]\right\}\right| = \dfrac1{K} + o\left(\dfrac1K\right)\ .
+\end{align*}
+Similar calculations lead to the same results for the other three terms making up the sum~\eqref{eq:sum.g1g1}.  Therefore, we have:
+\begin{align}
 \nonumber
-&\hspace{110pt}  + \EE\left\{\bw_k[m_r+a_r]\bw_{k'}[m_{r'}+a_{r'}]\right\} \\
-\nonumber
-&\leq \dfrac1{K^2}\left(\sum_{j=1}^J \Omega_j\right)^2 \sum_{k=0}^{K-1} 4 \\
-&\leq \dfrac{C_\bz^2}{K}\ ,
+\left|\bE\{\bg_1[r]\bg_1[r']\}\right| &\leq \left(\sum_{j=1}^J \Omega_j\right)^2 \left(\dfrac4{K} + o\left(\dfrac1{K}\right) \right) \\
+&\leq \dfrac{C_\bz^2}{K} + o\left(\dfrac1{K}\right)\ .
 \label{eq:g1g1}
 \end{align}
-where $C_\bz \defeq 2\left(\sum_{j=1}^J \Omega_j\right)$. Besides, since odd-order moments of a zero-mean multivariate Gaussian random vector are zero, we have:
+Besides, since odd-order moments of a zero-mean multivariate Gaussian random vector are zero, we have:
 \begin{align}
@@
 \end{equation}
-Besides,
+Besides, by a direct calculation, we have:
 \begin{align}
 \nonumber
-\left|\bE\{\bg_2[r]\bg_2[r']\}\right| &= \dfrac1{K^2}\bigg|\sum_{k,k'=0}^{K-1} \EE\left\{\bw_{k}[m_{r}+a_{r}]\bw_{k}[m'_{r}]\bw_{k'}[m_{r'}+a_{r'}]\bw_{k'}[m'_{r'}]\right\} - \delta_{m_{r}+a_{r},m'_{r}}\bE\{\bg_2[r']\} - \delta_{m_{r'}+a_{r'},m'_{r'}}\bE\{\bg_2[r]\}\\
+\bE\{\bg_2[r]\bg_2[r']\} &= \dfrac1{K^2}\sum_{k,k'=0}^{K-1} \EE\left\{\bw_{k}[m_{r}+a_{r}]\bw_{k}[m'_{r}]\bw_{k'}[m_{r'}+a_{r'}]\bw_{k'}[m'_{r'}]\right\} - \delta_{m_{r}+a_{r},m'_{r}}\dfrac1{K}\sum_{k'=0}^{K-1} \EE\left\{\bw_{k'}[m_{r'}+a_{r'}]\bw_{k'}[m'_{r'}]\right\} \\
 \nonumber
-&\hspace{30pt} + \delta_{m_{r}+a_{r},m'_{r}}\delta_{m_{r'}+a_{r'},m'_{r'}} \bigg| \\
+&\hspace{60pt} - \delta_{m_{r'}+a_{r'},m'_{r'}}\dfrac1{K}\sum_{k=0}^{K-1} \EE\left\{\bw_k[m'_r]\bw_{k'}[m'_{r'}]\right\} + \delta_{m_{r}+a_{r},m'_{r}}\delta_{m_{r'}+a_{r'},m'_{r'}} \\
+&= \dfrac1{K^2}\sum_{k,k'=0}^{K-1} \EE\left\{\bw_{k}[m_{r}+a_{r}]\bw_{k}[m'_{r}]\bw_{k'}[m_{r'}+a_{r'}]\bw_{k'}[m'_{r'}]\right\} - \delta_{m_{r}+a_{r},m'_{r}}\delta_{m_{r'}+a_{r'},m'_{r'}}
+\label{eq:sum.g2g2}
+\end{align}
+Moreover, using the results of the Isserlis' theorem~\cite{Isserlis16formula} to express fourth-order moments of a Gaussian random vector as a function of its second-order moments, we expand sum~\eqref{eq:sum.g2g2} as follows:
+\begin{align*}
+\bE\{\bg_2[r]\bg_2[r']\} &= \dfrac1{K^2} \sum_{k,k'=0}^{K-1} \EE\left\{\bw_{k}[m_{r}+a_{r}]\bw_{k}[m'_{r}]\right\}\EE\left\{\bw_{k'}[m_{r'}+a_{r'}]\bw_{k'}[m'_{r'}]\right\} - \delta_{m_{r}+a_{r},m'_{r}}\delta_{m_{r'}+a_{r'},m'_{r'}} \\
+&\hspace{30pt}+ \dfrac1{K^2}\sum_{k,k'=0}^{K-1}\EE\left\{\bw_{k}[m_{r}+a_{r}]\bw_{k'}[m_{r'}+a_{r'}]\right\}\EE\left\{\bw_{k}[m'_{r}]\bw_{k'}[m'_{r'}]\right\} \\
+&\hspace{30pt}+ \dfrac1{K^2}\sum_{k,k'=0}^{K-1} \EE\left\{\bw_{k}[m_{r}+a_{r}]\bw_{k'}[m'_{r'}]\right\}\EE\left\{\bw_{k}[m'_{r}]\bw_{k'}[m_{r'}+a_{r'}]\right\} \\
+& = \dfrac1{K^2}\sum_{k,k'=0}^{K-1} \delta_{k+m_r+a_r,k'+m_{r'}+a_{r'}}\delta_{k+m'r,k'+m'_{r'}} + \delta_{k+m_r+a_r,k'+m'_{r'}}\delta_{k+m'_r,k'+m_{r'}+a_{r'}} \\
+& = \dfrac1{K^2}\left(\delta_{m'_r-m'_{r'},m_r+a_r-m_{r'}-a_{r'}}(K-|m'_r-m'_{r'}|) +  \delta_{m_r+a_r-m'_{r'},m'_r-m_{r'}-a_{r'}}(K-|m_r+a_r-m'_{r'}|)\right) \\
+& = \dfrac1{K}\left(\delta_{m'_r-m'_{r'},m_r+a_r-m_{r'}-a_{r'}} + \delta_{m'_r+m'_{r'},m_r+a_r+m_{r'}+a_{r'}}\right) + o\left( \dfrac1K\right)\ .
+\end{align*}
+Therefore,
+\begin{equation}
+\left| \bE\{\bg_2[r]\bg_2[r']\} \right| \leq \dfrac2{K} + o\left( \dfrac1K\right)\ .
+\label{eq:g2g2}
+\end{equation}
+Thus, combining results~\eqref{eq:g1g1},~\eqref{eq:g1g2},~\eqref{eq:g2g1} and~\eqref{eq:g2g2} into expression~\eqref{eq:gg} gives the following bound:
+\begin{equation*}
+\left|\bE\{\bg[r]\bg[r']\}\right| \leq \dfrac1K \left( C_\bz^2\sigma^2 + 2 \sigma^4\right) + o\left(\dfrac1K\right)\ .
+\end{equation*}
+
+Concerning higher-order moments, let $T\geq 3$ denote the order of the moment defined by $\EE\{\prod_{\theta=1}^T\bg[r_\theta]\}$. Thus,
+\begin{align}
 \nonumber
-&= \dfrac1{K^2} \bigg| \sum_{k,k'=0}^{K-1} \EE\left\{\bw_{k}[m_{r}+a_{r}]\bw_{k}[m'_{r}]\right\}\EE\left\{\bw_{k'}[m_{r'}+a_{r'}]\bw_{k'}[m'_{r'}]\right\} - \delta_{m_{r}+a_{r},m'_{r}}\delta_{m_{r'}+a_{r'},m'_{r'}} \\
+\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}& = \EE\left\{\prod_{\theta=1}^T \left(\sigma\bg_1[r_\theta] + \sigma^2\bg_2[r_\theta] \right)\right\}\\
 \nonumber
-&\hspace{30pt}+ \dfrac1{K^2}\sum_{k,k'=0}^{K-1}\EE\left\{\bw_{k}[m_{r}+a_{r}]\bw_{k'}[m_{r'}+a_{r'}]\right\}\EE\left\{\bw_{k}[m'_{r}]\bw_{k'}[m'_{r'}]\right\} \\
-\nonumber
-&\hspace{30pt}+ \dfrac1{K^2}\sum_{k,k'=0}^{K-1} \EE\left\{\bw_{k}[m_{r}+a_{r}]\bw_{k'}[m'_{r'}]\right\}\EE\left\{\bw_{k}[m'_{r}]\bw_{k'}[m_{r'}+a_{r'}]\right\} \bigg| \\
-\nonumber
-& \leq \dfrac1{K^2} \sum_{k=0}^{K-1} 2 \\
-& \leq \dfrac2{K}\ .
-\label{eq:g2g2}
+&=\dfrac1{K^T}\EE\left\{\prod_{\theta=1}^T \left(\sum_{k=0}^{K-1}\rho_{\theta,k}(\bw[k+m'_{r_\theta}],\bw[k+m_{r_\theta}+a_{r_\theta}]) \right)\right\}\\ 
+&=\dfrac1{K^T}\sum_{k_1=0}^{K-1}\cdots\sum_{k_T=0}^{K-1}\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]) \right\} \,,
+\label{eq:multisum}
 \end{align}
-Thus, combining results~\eqref{eq:g1g1},~\eqref{eq:g1g2},~\eqref{eq:g2g1} and~\eqref{eq:g2g2} into expression~\eqref{eq:gg} gives the following bound:
+where
+\[
+\rho_{\theta,k}(u,v)=\sigma \bz_k[m_{r_\theta}+a_{r_\theta}]u + \sigma \bz_k[m'_{r_\theta}]v + \sigma^2uv - \delta_{m_{r_\theta}+a_{r_\theta},\,m'_{r_\theta}}\ .
+\]
+Thus,
+\begin{align*}
+\left|\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}\right| &\leq \dfrac{\nu_K}{K^T} C\ ,
+\end{align*}
+where
+\[
+C = \max_{(k_1,\ldots,k_T)} \left|\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]) \right\} \right| \,,
+\]
+and $\nu_K$ is the number of non-zero terms in the sum~\eqref{eq:multisum}.
+Note that $C$ is independent of $K$ (but depends on $\bz$ and $\sigma$).
+The behavior of the order $T$ moment in function of $K$ is therefore only determined by the ratio $\dfrac{\nu_K}{K^T}$.
+
+Let us bound $\nu_K$.
+Fix $k_1$.
+For each of the other sum indexes $k_\theta$ ($\theta\in\{2,\ldots,T\}$), there are four values that make $\rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}])$ not independent of $\rho_{1,k_1}(\bw[k_1+m'_{r_1}],\bw[k_1+m_{r_1}+a_{r_1}])$.
+Indeed, since $\bw$ is a white noise these quantities are independent except when $k_\theta$ is such that:
+\begin{align*}
+k_\theta &= k_1+m'_{r_1}-m'_{r_\theta} \\
+k_\theta &= k_1+m_{r_1}+a_{r_1} - m'_{r_\theta}\\
+k_\theta &= k_1+m'_{r_1} - m_{r_\theta}- a_{r_\theta} \\
+k_\theta &= k_1+m_{r_1}+a_{r_1} - m_{r_\theta}- a_{r_\theta}\ .
+\end{align*}
+Consequently, there exist at least $(K-4)^{T-1}$ combinations of $(k_2,\ldots,k_T)$ such that the correlation term given by $\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]) \right\}$ is zero.
+Because $T\geq 3$, we develop similar arguments on $k_2$ and $k_3$ to obtain the following maximum bound on the number of non-zero terms in the sum~\eqref{eq:multisum}:
+\[
+\nu_K\leq K^T - 3K(K-4)^{T-1} + 3K(K-4)(K-8)^{T-2}- K(K-4)(K-8)(K-12)^{T-3}\ .
+\]
+Thus,
+\begin{align*}
+\left|\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}\right| &\leq C \dfrac{K^T - 3K(K-4)^{T-1} + 3K(K-4)(K-8)^{T-2}- K(K-4)(K-8)(K-12)^{T-3}}{K^T} \\
+&\leq C \left( 1 - 3\left(1-\frac4K\right)^{T-1} + 3\left(1-\frac4K\right)\left(1-\frac8K\right)^{T-2}- \left(1-\frac4K\right)\left(1-\frac8K\right)\left(1-\frac{12}{K}\right)^{T-3} \right) \\
+&\leq C \left( 1 - 3 +\frac{12(T-1)}{K} + 3-\frac{12}K-\frac{24(T-2)}{K} -1 + \frac{4}K + \frac{8}K +\frac{12(T-3)}{K} + o\left(\dfrac1K\right)\right) \\
+&\leq o\left(\dfrac1K\right)\ . \\
+\end{align*}
+Therefore,
+\begin{equation*}
+\left|\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}\right| = o\left(\dfrac1K\right)\,,\quad\forall T\geq 3\ .
+\end{equation*}
+\end{proof}
+
+\begin{lemma}[Bounds on the derivatives of $f^{(\ell)}$ at the origin]
+Let $f^{(\ell)} : \RR^{M(M+1)} \to \RR^{M}$ denote the multivariate function defined in~\eqref{eq:f.ell}.
+Assume the deterministic signal $\bz$ takes the form~\eqref{I-eq:sum.sine}, and the observed noisy signal takes the form~\eqref{I-eq:model.noise}.
+Then, the first-order derivatives of $f^{(\ell)}$ at the origin are bounded as follows:
 \begin{equation}
-\left|\bE\{\bg[r]\bg[r']\}\right| \leq \dfrac1K \left( C_\bz^2\sigma^2 + 2 \sigma^4\right)\ .
-\label{eq:gg.bound}
+\left|\left.\dfrac{\partial f_m^{(\ell)}}{\partial \bg[r]}\right\vert_{\bg=\bzero}\right| \leq \dfrac{d_{1,\bz,M,\ell}}{\sigma^2}\,,\quad\forall m\in\{0,\ldots,M-1\}\,,\ r\in\{0,\ldots,M(M+1)-1\}\ ,
+\label{eq:df.bound}
 \end{equation}
-In the same way, we show that higher-order moments behave as $o\left( \dfrac1K \right)$.
+where
+\[
+d_{1,\bz,M,\ell}\defeq \left(2 + (\ell-2)M\right)M^{\ell-1}\left(\max\left(1,\frac{2J}{M}\right)\right)^{\ell-1}\left(1+\max\left(1,\frac{2J}{M}\right)\right)\left(1+\frac{2J}{M}\right)\ .
+\]
+Besides, the second-order derivatives of $f^{(\ell)}$ at the origin are bounded as follows:  
+\begin{equation}
+\left|\left.\dfrac{\partial^2 f_m^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero}\right| \leq \dfrac{d_{2,\bz,M,\ell}}{\sigma^4}\,,\quad\forall m\in\{0,\ldots,M-1\}\,,\ (r,r')\in\{0,\ldots,M(M+1)-1\}^2\ ,
+\label{eq:d2f.bound}
+\end{equation}
+where 
+\begin{align*}
+d_{2,\bz,M,\ell} &\defeq d_{2,M,\ell}C_{\bz,M}^2\max\left(1,\frac{2J}{M}\right)^{\ell-2} + 2d'_{2,M,\ell}\max\left(1,\frac{2J}{M}\right)^{\ell-1}\left(1+\max\left(1,\frac{2J}{M}\right)\right)\left(1+\frac{2J}{M}\right)^2\,,
+\end{align*}
+and $d_{2,M,\ell}$ and $d'_{2,M,\ell}$ are only depending on $M$ and $\ell$.
+\end{lemma}
 
-\subsubsection{Study of $f^{(\ell)}$}
-In this section, we provide bounds on the first and second-order derivatives of $f^{(\ell)}$ at the origin.
-
+\begin{proof}
 Concerning the first-order derivative, from~\eqref{eq:error.vec}, we have:
@@
 \end{align*}
-Let $\|\cdot\|_\infty$ denote the uniform norm of a vector or a matrix, \ie~$\|\bM\|_\infty = \max_{n,n'} \bM[n,n']$. Then,
+Then,
 \begin{equation*}
-\left\| \left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero}\right\|_\infty \leq \left(1+\|\bA_0\|_\infty\right)\left\|{\bS^{(0)}}^{-1}\right\|_\infty\ .
+\left\| \left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero}\right\|_{\max} \leq \left(1+\|\bA_0\|_{\max}\right)\left\|{\bS^{(0)}}^{-1}\right\|_{\max}\ .
 \end{equation*}
-Given equation~\eqref{eq:A0.sine}, we have 
+Given bounds~\eqref{eq:A0.bound} and~\eqref{eq:S0.inv.bound}, we have that:
 \begin{equation}
-\|\bA_0\|_\infty\leq \max\left(1,\frac{2J}{M}\right)\ ;
-\label{eq:A0}
-\end{equation}
-and from~\eqref{eq:S0.inv}, we have that $\left\|{\bS^{(0)}}^{-1}\right\|_\infty\leq \dfrac1{\sigma^2}\left(1+\frac{2J}{M}\right)$. Then,
-\begin{equation}
-\left\| \left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero}\right\|_\infty \leq \left(1+\max\left(1,\frac{2J}{M}\right)\right)\left(1+\frac{2J}{M}\right)\dfrac{1}{\sigma^2}\ .
+\left\| \left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero}\right\|_{\max} \leq \left(1+\max\left(1,\frac{2J}{M}\right)\right)\left(1+\frac{2J}{M}\right)\dfrac{1}{\sigma^2}\ .
 \label{eq:dA}
@@
 \begin{align*}
-\left|\left.\dfrac{\partial f_m^{(\ell)}}{\partial \bg[r]}\right\vert_{\bg=\bzero}\right| &\leq 2M\|\bA_0^{\ell-1}\|_\infty \left\| \left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero}\right\|_\infty + M^2\sum_{\lambda=1}^{\ell-2} \|\bA_0^\lambda\|_\infty \left\| \left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero}\right\|_\infty  \|\bA_0^{\ell-1-\lambda}\|_\infty \\
-&\leq \left(2 + (\ell-2)M\right)M^{\ell-1} \|\bA_0\|_\infty^{\ell-1}\left\|\left.\dfrac{\partial\tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero}\right\|_\infty\ .
+\left|\left.\dfrac{\partial f_m^{(\ell)}}{\partial \bg[r]}\right\vert_{\bg=\bzero}\right| &\leq 2M\|\bA_0^{\ell-1}\|_{\max} \left\| \left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero}\right\|_{\max} + M^2\sum_{\lambda=1}^{\ell-2} \|\bA_0^\lambda\|_{\max} \left\| \left.\dfrac{\partial \tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero}\right\|_{\max}  \|\bA_0^{\ell-1-\lambda}\|_{\max} \\
+&\leq \left(2 + (\ell-2)M\right)M^{\ell-1} \|\bA_0\|_{\max}^{\ell-1}\left\|\left.\dfrac{\partial\tilde\bA}{\partial \bg[r]}\right|_{\bg=\bzero}\right\|_{\max}\ .
 \end{align*}
-Therefore, given bounds~\eqref{eq:A0} and~\eqref{eq:dA}, we have:
-\begin{equation}
-\left|\left.\dfrac{\partial f_m^{(\ell)}}{\partial \bg[r]}\right\vert_{\bg=\bzero}\right| \leq \dfrac{d_{1,\bz,M,\ell}}{\sigma^2}\ ,
-\label{eq:df.bound}
-\end{equation}
-where $d_{1,\bz,M,\ell}\defeq \left(2 + (\ell-2)M\right)M^{\ell-1}\left(\max\left(1,\frac{2J}{M}\right)\right)^{\ell-1}\left(1+\max\left(1,\frac{2J}{M}\right)\right)\left(1+\frac{2J}{M}\right)$.
+Therefore, given bounds~\eqref{eq:A0.bound} and~\eqref{eq:dA}, we have:
+\begin{equation*}
+\left|\left.\dfrac{\partial f_m^{(\ell)}}{\partial \bg[r]}\right\vert_{\bg=\bzero}\right| \leq \dfrac{d_{1,\bz,M,\ell}}{\sigma^2}\ .
+\end{equation*}
 
@@
 \nonumber
-\left\| \left.\dfrac{\partial^2 \tilde\bA}{\partial \bg[r]\partial \bg[r']}\right|_{\bg=0}\right\|_\infty &\leq \left\{
+\left\| \left.\dfrac{\partial^2 \tilde\bA}{\partial \bg[r]\partial \bg[r']}\right|_{\bg=0}\right\|_{\max} &\leq \left\{
 \begin{array}{ll}
-\left\|{\bS^{(0)}}^{-1}\right\|_\infty^2 &\mbox{if}\quad a_r=1\quad\mbox{or}\quad a_{r'}=1\,, \\
-2\left(1+\|\bA_0\|_\infty\right)\left\|{\bS^{(0)}}^{-1}\right\|_\infty^2 &\mbox{else} \\
+\left\|{\bS^{(0)}}^{-1}\right\|_{\max}^2 &\mbox{if}\quad a_r=1\quad\mbox{or}\quad a_{r'}=1\,, \\
+2\left(1+\|\bA_0\|_{\max}\right)\left\|{\bS^{(0)}}^{-1}\right\|_{\max}^2 &\mbox{else} \\
 \end{array}
@@
 \begin{align*}
-\left|\left.\dfrac{\partial^2 f_m^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero}\right| &\leq d_{2,M,\ell} \|\bA_0\|_\infty^{\ell-2}\left\|\dfrac{\partial\tilde\bA}{\partial \bg[r]}\right\|_\infty\left\|\dfrac{\partial\tilde\bA}{\partial \bg[r']}\right\|_\infty + d'_{2,M,\ell}\|\bA_0\|_\infty^{\ell-1}\left\|\dfrac{\partial^2\tilde\bA}{\partial \bg[r]\bg[r']}\right\|_\infty \ ,
+\left|\left.\dfrac{\partial^2 f_m^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero}\right| &\leq d_{2,M,\ell} \|\bA_0\|_{\max}^{\ell-2}\left\|\dfrac{\partial\tilde\bA}{\partial \bg[r]}\right\|_{\max}\left\|\dfrac{\partial\tilde\bA}{\partial \bg[r']}\right\|_{\max} + d'_{2,M,\ell}\|\bA_0\|_{\max}^{\ell-1}\left\|\dfrac{\partial^2\tilde\bA}{\partial \bg[r]\bg[r']}\right\|_{\max} \ ,
 \end{align*}
@@
 Besides, given results~\eqref{eq:A0}, ~\eqref{eq:dA} and~\eqref{eq:d2A}, we have:
-\begin{equation}
-\left|\left.\dfrac{\partial^2 f_m^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero}\right| \leq \dfrac{d_{2,\bz,M,\ell}}{\sigma^4}\ ,
-\label{eq:d2f.bound}
-\end{equation}
-where 
-\begin{align*}
-d_{2,\bz,M,\ell} &= d_{2,M,\ell}C_{\bz,M}^2\max\left(1,\frac{2J}{M}\right)^{\ell-2} + 2d'_{2,M,\ell}\max\left(1,\frac{2J}{M}\right)^{\ell-1}\left(1+\max\left(1,\frac{2J}{M}\right)\right)\left(1+\frac{2J}{M}\right)^2\ .
-\end{align*}
+\begin{equation*}
+\left|\left.\dfrac{\partial^2 f_m^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero}\right| \leq \dfrac{d_{2,\bz,M,\ell}}{\sigma^4}\ .
+\end{equation*}
+\end{proof}
 
@@
 \nonumber
-\left|\bepsilon_2[\ell]\right| &\leq C_\bz\dfrac{M^3(M+1)^2}{4}\left\|\left.\dfrac{\partial^2 f^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero}\right\|_\infty \dfrac1K \left( C_\bz^2\sigma^2 + 2 \sigma^4\right) + o\left(\dfrac1K\right)  \\
+\left|\bepsilon_2[\ell]\right| &\leq C_\bz\dfrac{M^3(M+1)^2}{4}\left\|\left.\dfrac{\partial^2 f^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero}\right\|_{\max} \dfrac1K \left( C_\bz^2\sigma^2 + 2 \sigma^4\right) + o\left(\dfrac1K\right)  \\
 &\leq \dfrac1K \left(c^{(2)}_1 + \dfrac{c^{(2)}_2}{\sigma^2} \right) + o\left(\dfrac1K\right)\ ,
@@
 
-%\begin{align*}
-%\bh^{(\ell)}[m] &= f_m^{(\ell)}(\bg) \\
-%& = f_m^{(\ell)}(\bzero) + \sum_{r=0}^{M(M-1)}\left.\dfrac{\partial f_m^{(\ell)}}{\partial \bg[r]}\right\vert_{\bg=\bzero} \bg[r] + \dfrac12\sum_{r,r'=0}^{M(M-1)}\left.\dfrac{\partial^2 f_m^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero} \bg[r]\bg[r'] + o\left(\dfrac1K\right)\ .
-%\end{align*}
-
 \subsection{Expression of the Covariance $\bgamma$.}
@@
 
-First, since $\|\balpha_0^{(\ell)}\|^2\leq M\|\balpha_0^{(\ell)}\|_\infty^2\leq M\|\bA_0^{\ell}\|_\infty^2$, we have:
+First, since $\|\balpha_0^{(\ell)}\|^2\leq M\|\bA_0^{\ell}\|_{\max}^2$, we have:
 \begin{align}
 \nonumber
-\eta_1[n] &\leq \sigma^2M\left\|\bA_0^{\ell}\right\|_\infty^2 \\
+\eta_1[n] &\leq \sigma^2M\left\|\bA_0^{\ell}\right\|_{\max}^2 \\
 \nonumber
-&\leq\sigma^2M^{\ell}\left\|\bA_0\right\|^{2\ell}_\infty \\
+&\leq\sigma^2M^{\ell}\left\|\bA_0\right\|^{2\ell}_{\max} \\
 &\leq \sigma^2 M^{\ell}\max\left(1,\left(\frac{2J}{M}\right)^{2\ell}\right)\ .
@@
 \begin{align*}
-\left|\eta_4[n]\right| & \leq \sigma C_\bz\|\balpha_0^{(\ell)}\|_\infty \sum_{m,m'=0}^{M-1}\left|\EE\{\bh^{(\ell)}[m]\bw_K[m']\} \right|
+\left|\eta_4[n]\right| & \leq \sigma C_\bz\left\|\bA_0^{\ell}\right\|_{\max} \sum_{m,m'=0}^{M-1}\left|\EE\{\bh^{(\ell)}[m]\bw_K[m']\} \right|
 \end{align*}
@@
 \begin{align*}
-\left|\eta_7[n] \right| &\leq 2\sigma^2\|\balpha_0^{(\ell)}\|_\infty \sum_{m,m'=0}^{M-1} \left| \EE\{\bh^{(\ell)}[m]\bw_K[m]\bw_K[m']\} \right| \\
-&\leq 2\sigma^2\|\balpha_0^{(\ell)}\|_\infty \sum_{m,m'=0}^{M-1} \sum_{r=0}^{M(M+1)-1}\left|\left.\dfrac{\partial f_m^{(\ell)}}{\partial \bg[r]}\right\vert_{\bg=\bzero}\right| \left| \EE\{\bg^{(\ell)}[r]\bw_K[m]\bw_K[m']\} \right| + o\left( \dfrac1K\right)\ .
+\left|\eta_7[n] \right| &\leq 2\sigma^2\left\|\bA_0^{\ell}\right\|_{\max} \sum_{m,m'=0}^{M-1} \left| \EE\{\bh^{(\ell)}[m]\bw_K[m]\bw_K[m']\} \right| \\
+&\leq 2\sigma^2\left\|\bA_0^{\ell}\right\|_{\max} \sum_{m,m'=0}^{M-1} \sum_{r=0}^{M(M+1)-1}\left|\left.\dfrac{\partial f_m^{(\ell)}}{\partial \bg[r]}\right\vert_{\bg=\bzero}\right| \left| \EE\{\bg^{(\ell)}[r]\bw_K[m]\bw_K[m']\} \right| + o\left( \dfrac1K\right)\ .
 \end{align*}
@@
 \nonumber
-\left| \eta_7[n] \right| &\leq 2\sigma^2\|\balpha_0^{(\ell)}\|_\infty M^3(M+1)\dfrac{d_{1,\bz,M,\ell}}{\sigma^2} \dfrac{2\sigma^2}K + o\left( \dfrac1K\right)\\
+\left| \eta_7[n] \right| &\leq 2\sigma^2\left\|\bA_0^{\ell}\right\|_{\max} M^3(M+1)\dfrac{d_{1,\bz,M,\ell}}{\sigma^2} \dfrac{2\sigma^2}K + o\left( \dfrac1K\right)\\
 &\leq 4 M^{\ell+2}(M+1)d_{1,\bz,M,\ell}\left(\max\left(1,\frac{2J}{M}\right)\right)^\ell\dfrac{\sigma^2}{K} + o\left( \dfrac1K\right)\ .
@@
 \left|\gamma[n,n']\right| & \leq  \sqrt{\gamma[n,n]\gamma[n',n']} \\
-& \leq  \sqrt{\gamma[n,n]\gamma[n',n']} \\
 & \leq   \sqrt{c_0^{(n)}c_0^{(n')}}\sigma^2\sqrt{1 + \dfrac{1}{K\sigma^2} \left( \dfrac1{c_0^{(n)}}\left( \dfrac{c_1^{(n)}}{\sigma^2} + c_2^{(n)} + c_3^{(n)}\sigma^2\right) + \dfrac1{c_0^{(n')}}\left( \dfrac{c_1^{(n')}}{\sigma^2} + c_2^{(n')} + c_3^{(n')}\sigma^2\right)\right) + o\left(\dfrac1K\right)}\ .
@@
 \nonumber
-\left|\beta_2[n,n']\right| &\leq \sigma^2\left\|\balpha_0^{(\ell)}\right\|_\infty\sum_{m=0}^{M-1}\left|\EE\{\bw[n']\bw_K[m]\}\right| \\
+\left|\beta_2[n,n']\right| &\leq \sigma^2\left\|\bA_0^{\ell}\right\|_{\max}\sum_{m=0}^{M-1}\left|\EE\{\bw[n']\bw_K[m]\}\right| \\
 \nonumber
-&\leq \sigma^2\left\|\bA_0^{\ell}\right\|_\infty \\
+&\leq \sigma^2\left\|\bA_0^{\ell}\right\|_{\max} \\
 &\leq \sigma^2M^{\ell-1}\left(\max\left(1,\frac{2J}{M}\right)\right)^{\ell}\ .
@@
 
-%The multivariate version of the delta method is applicable to $\bh^{(\ell)}$ under the following conditions:
-%\begin{enumerate}[(i)]
-%\item\label{delta.1} $f^{(\ell)}$ is differentiable at the origin;
-%\item\label{delta.2} $\sqrt{K}\ \bg \xrightarrow[K\to\infty]{\cD} \cN(\bzero,\bGamma_0)$ where $\bGamma_0$ is a covariance matrix.
-%\end{enumerate}
-%When these conditions are satisfied, the delta method gives the asymptotic behavior of $\bh^{(\ell)}=f^{(\ell)}(\bg)$, that is
-%\begin{equation}
-%\sqrt{K}\ \bh^{(\ell)} \xrightarrow[K\to\infty]{\cD} \cN(\bzero,{\bF^{(\ell)}}^T\bGamma_0\bF^{(\ell)})\ ,
-%\label{eq:delta}
-%\end{equation}
-%where $\bF^{(\ell)}$ is the Jacobian matrix of $f^{(\ell)}$ at the origin, that is
-%\[
-%\bF^{(\ell)}[m,m'] = \left.\dfrac{\partial f^{(\ell)}_m}{\partial\bg[m']}\right\vert_{\bg=\bzero}\ .
-%\]
-%
-%Condition~\ref{delta.1} is satisfied. Indeed, the differentiation of $f^{(\ell)}$ can be checked as a composition of standard matrix derivation rules. {\color{red}Detail it}
-%
-%Concerning condition~\ref{delta.2}, by definition of $\bg$, we have:
-%\[
-%\bg = \sigma\bg_1 + \sigma^2\bg_2\ ,
-%\]
-%where $\bg_1$ and $\bg_2$ are defined by
-%\begin{align*}
-%\bg_1 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left( \tilde\bz_k\bw_k^T + \tilde\bw_k\bz_k^T \right)\quad\mbox{where}\quad \tilde\bz_k^T= \begin{pmatrix} \bz_k^T & \bz_{k+1}[M-1] \end{pmatrix}\,, \\
-%\bg_2 &= \dfrac1K\sum_{k=0}^{K-1}\vec\left(\tilde\bw_k\bw_k^T-\bI \right) \quad\mbox{where}\quad \tilde\bw_k^T= \begin{pmatrix} \bw_k^T & \bw_{k+1}[M-1] \end{pmatrix} \,.
-%\end{align*}
-%First, $\bg_1$ is intrinsically a Gaussian random vector since it is a linear combination of Gaussian random vectors.
-Second, using the central limit theorem under weak dependence, we can show that $\bg_2$ also converges towards a Gaussian random vector as $K\to\infty$.
-Hence, we have
-%\[
-%\sqrt{K}\ \bg \xrightarrow[K\to\infty]{\cD} \cN(\bzero,\bGamma_0)\ ,
-%\]
-%where $\bGamma_0$ is the limit covariance matrix so that $\bGamma_0=\lim\limits_{K\to\infty} K\,\EE\{\bg\bg^T\}$.
-Hence, conditions~\ref{delta.1} and~\ref{delta.1} are satisfied.
-The delta method can therefore be applied, and result~\eqref{eq:delta} is valid.
-
-%Therefore,
-%\begin{align*}
-%\bgamma[n,n] &\leq \sigma^2\left\|\balpha_0^{(\ell)}\right\|^2-\left(\bepsilon_2[n]+\bepsilon_3[n]\right)^2 + \EE\left\{\left(\bh^{(\ell)}\bz_K\right)^2\right\} + 2\sigma\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\ell)}\}\bz_K+ 2\sigma\EE\{\bh^{(\ell)}\bw_K\bh^{(\ell)}\}\bz_K \\
-%&+\sigma^2\EE\left\{\left(\bh^{(\ell)}\bw_K\right)^2\right\}  + 2\sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\ell)}\bw_K\}
-%\end{align*}
-
-%Take $n>N$, and denote $n=N-1+\ell$.  To derive the covariance, let us segregate the cases. 
-% 
-%\paragraph{When $n'\in I$}
-%From Lemma~1, we have that $h^{(\ell)}$ is asymptotically Gaussian when $K\to \infty$.
-Then, as a direct consequence of the Isserlis' theorem, odd-order moments are vanishing {\color{red}[there is a gap here.
-Note that when $K$ is finite, it is approximated by Gaussian but not Gaussian.
-The discrepancy should be made clear.]}.
-Then, inserting result~\eqref{eq:bias.ap} into equation~\eqref{I-eq:bias0}, we obtain:
-%\begin{equation}
-%\bgamma[n,n'] = \sigma\EE\{\bw[n']\bh^{(\ell)}\}\bz_K + \sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bw[n']\} + o(\sigma^2) 
-%\label{eq:gamma.nn}
-%\end{equation}
-%when $K\to \infty$.
-%In addition, we remark that
-%\[
-%\balpha_0^{(\ell)}\EE\{\bw_K\bw[n']\}=\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)}\ .
-%\]
-%Besides, according to result~\eqref{eq:cov.ap} the first term in~\eqref{eq:gamma.nn} vanishes when $K\to\infty$. We thus have
-%\begin{align*}
-%\bgamma[n,n'] &= \sigma^2\balpha_0^{(\ell)}[n'-(N-M)]\1_{(n'\geq N-M)}+ o(\sigma^2)
-%\end{align*}
-%when $K\to\infty$.
-%
-%
-%\paragraph{When $n'\geq N$}
-%Inserting equation~\eqref{eq:bias.ap} into result~\eqref{I-eq:cov01}, we obtain
-%\begin{align*}  
-%\bgamma[n,n'] = &\,\bz_K^T\EE\left\{{\balpha^{(\ell)}}^T\balpha^{(\lambda)}\right\}\bz_K + \sigma\EE\{\balpha^{(\ell)}\bw_K\balpha^{(\lambda)}\}\bz_K+ \sigma\EE\{\balpha^{(\lambda)}\bw_K\balpha^{(\ell)}\}\bz_K \\
-%& +\sigma^2\EE\{\balpha^{(\ell)}\bw_K\balpha^{(\lambda)}\bw_K\}-\bz[n]\bz[n']+o(\sigma^2) \\
-%= &\,\bz_K^T\EE\left\{{\bh^{(\ell)}}^T\bh^{(\lambda)}\right\}\bz_K + \balpha_0^{(\ell)}\sigma\EE\{\bw_K\bh^{(\lambda)}\}\bz_K + \sigma\EE\{\bh^{(\ell)}\bw_K\}\bz[n']  \\
-%&+ \sigma\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\}\bz_K+ \sigma\EE\{\bh^{(\lambda)}\bw_K\}\bz[n] + \sigma\balpha_0^{(\lambda)}\EE\{\bw_K\bh^{(\ell)}\}\bz_K \\
-%& + \sigma\EE\{\bh^{(\lambda)}\bw_K\bh^{(\ell)}\}\bz_K+\sigma^2\balpha_0^{(\ell)}\EE\{\bw_K\bh^{(\lambda)}\bw_K\}+\sigma^2\balpha_0^{(\lambda)}\EE\{\bw_K\bh^{(\ell)}\bw_K\} \\
-%&+\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\sigma^2\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} + o(\sigma^2)\ .
-%\end{align*}
-%From Lemma~1, we have that $\bh^{(\ell)}$ is asymptotically Gaussian.
-Then, as a direct consequence of the Isserlis' theorem, odd-order moments are vanishing.
-{\color{red}[The same gap here.]} {\color{blue}Considering this property and equation~\eqref{eq:cov.ap} gives
-%\begin{align*}
-%\bgamma[n,n'] = &\,\dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K + \sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\sigma^2\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} + o(\sigma^2)
-%\end{align*}
-%when $K\to\infty$. The Isserlis' theorem also gives the following result {\color{red}[the same gap here.]}
-%\begin{align*}
-%\EE\{\bh^{(\ell)}\bw_K\bh^{(\lambda)}\bw_K\} &= \sum_{m,m'=0}^{M-1}\EE\{\bh^{(\ell)}[m]\bw_K[m]\bh^{(\lambda)}[m']\bw_K[m']\} \\
-%&= \sum_{m,m'=0}^{M-1}\EE\{\bh^{(\ell)}[m]\bw_K[m]\}\EE\{\bh^{(\lambda)}[m']\bw_K[m']\} + \EE\{\bh^{(\ell)}[m]\bw_K[m']\}\EE\{\bh^{(\lambda)}[m']\bw_K[m]\} \\
-%&\hspace{35pt} + \EE\{\bh^{(\ell)}[m]\bh^{(\lambda)}[m']\}\underbrace{\EE\{\bw_K[m]\bw_K[m']\}}_{=\delta_{m,m'}} \\
-%&= \sum_{m=0}^{M-1} \EE\{\bh^{(\ell)}[m]\bh^{(\lambda)}[m]\} = \dfrac1K\Tr\left(\bGamma^{(\ell,\lambda)}\right)
-%\end{align*}
-%when $K\to \infty$.
-%We thus conclude that
-%\begin{align*}  
-%\bgamma[n,n'] &= \dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K+\sigma^2\left\langle\balpha^{(\ell)},\balpha^{(\lambda)}\right\rangle +\dfrac{\sigma^2}{K}\Tr\left(\bGamma^{(\ell,\lambda)}\right) + o(\sigma^2)
-%\end{align*}
-%when $K\to \infty$.}
+\bibliographystyle{IEEEtran}
+\bibliography{../AM041320}
 

### 2020-09-08 Notes/Supplementary material/SupMat.tex
proof

@@
 \subsubsection{Notations}
-Considering $\bz\in\RR^N$ the deterministic signal defined in~\eqref{I-eq:sum.sine}, we denote by $\bz_k\in\RR^M$ the subsignal such that
+Consider $\bz\in\RR^N$ the deterministic signal defined in~\eqref{I-eq:sum.sine}, and denote by $\bz_k\in\RR^M$ the subsignal such that
 \[
@@
 \end{align*}
-Let $\bD\in\RR^{M\times M}$ be the matrix defined by $\bD[m,m']=\delta_{m+1,m'}$.
-
-Recall the model~\eqref{I-eq:model.noise}. Based on the definition of matrices $\bX$ and $\bY$, we have:
+Let $\bD\in\RR^{M\times M}$ be the matrix defined by $\bD[m,m']=\delta_{m+1,m'}$. Recall the model~\eqref{I-eq:model.noise}. Based on the definition of matrices $\bX$ and $\bY$, we have:
 \begin{align}
 \label{eq:XX.T}
-\dfrac1K\bX\bX^T &= \underbrace{\dfrac1K\bZ\bZ^T+\sigma^2\bI}_{\defeq S^{(0)}} + \bE^{(0)} \\
+\dfrac1K\bX\bX^T &= \underbrace{\dfrac1K\bZ\bZ^T+\sigma^2\bI}_{\defeq \bS^{(0)}} + \bE^{(0)} \\
 \label{eq:YX.T}
-\dfrac1K\bY\bX^T &= \underbrace{\dfrac1K\bZ'\bZ^T +\sigma^2\bD}_{\defeq S^{(1)}}+ \bE^{(1)} \ ,
+\dfrac1K\bY\bX^T &= \underbrace{\dfrac1K\bZ'\bZ^T +\sigma^2\bD}_{\defeq \bS^{(1)}}+ \bE^{(1)} \ ,
 \end{align}
-where $\bE^{(a)} := \sigma\bE_1^{(a)} + \sigma^2\bE_2^{(a)}$,
+where $\bE^{(a)} \defeq \sigma\bE_1^{(a)} + \sigma^2\bE_2^{(a)}$,
 \[
@@
 \end{align*}
-Thus, the matrix $\tilde\bA$, defined in equation~\eqref{I-eq:lse}, is expressed in the form:
+Thus, the random matrix $\tilde\bA$, defined in equation~\eqref{I-eq:lse}, is expressed in function of the above-defined matrices as:
 \begin{equation*}
@@
 \begin{lemma}[Expressions of $\bA_0$ and ${\bS^{(0)}}^{-1}$]
-Let $\bS^{(0)}$ be the $M\times M$ matrices defined in~\eqref{eq:XX.T}.
-Let $\bA_0$ the $M\times M$ matrices defined in~\eqref{eq:A0}.
+Let $\bS^{(0)}$ be the $M\times M$ matrix defined in~\eqref{eq:XX.T}.
+Let $\bA_0$ the $M\times M$ matrix defined in~\eqref{eq:A0}.
 Assume the deterministic signal $\bz$ takes the form~\eqref{I-eq:sum.sine}, and the observed noisy signal takes the form~\eqref{I-eq:model.noise}.
-Then, the matrix $\bS^{(0)}$ is given by
+Then, the inverse of the matrix $\bS^{(0)}$ is given by
 \begin{equation}
@@
 \begin{proof}
-It follows from the signal model~\eqref{I-eq:sum.sine} that matrices $\bS^{(0)}$ and $\bS^{(1)}$ take the following form:
+It follows from the signal model~\eqref{I-eq:sum.sine} that the matrices $\bS^{(0)}$ and $\bS^{(1)}$ take the following form:
 \begin{align}
 \nonumber
-\bS^{(a)}[m,m'] &= \sigma^2\delta_{(m+a)m'}+\sum_{j,j'=1}^J\dfrac{\Omega_j\Omega_{j'}}{K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(N_0+m+a+k)+\varphi_j\right)\cos\left(2\pi \frac{f_{j'}}{\fs}(N_0+m'+k)+\varphi_{j'}\right) \\
+\bS^{(a)}[m,m'] &= \sigma^2\delta_{(m+a)\,,m'}+\sum_{j,j'=1}^J\dfrac{\Omega_j\Omega_{j'}}{K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(N_0+m+a+k)+\varphi_j\right)\cos\left(2\pi \frac{f_{j'}}{\fs}(N_0+m'+k)+\varphi_{j'}\right) \\
 \nonumber
-&= \sigma^2\delta_{(m+a)m'}+\sum_{j=1}^J\dfrac{\Omega_j^2}{2K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)\\
+&= \sigma^2\delta_{(m+a)\,,m'}+\sum_{j=1}^J\dfrac{\Omega_j^2}{2K}\sum_{k=0}^{K-1} \cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)\\
 \nonumber
-&= \sigma^2\delta_{(m+a)m'}+\sum_{j=1}^J\Bigg( \dfrac{\Omega_j^2}2\cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \dfrac{\Omega_j^2}{2K}\underbrace{\sum_{k=0}^{K-1}\cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)}_{=0\ \mathrm{because}\ \frac{f_j}\fs=\frac{p'_j}K } \Bigg)\\
-&= \sigma^2\delta_{(m+a)m'}+\sum_{j=1}^J\dfrac{\Omega_j^2}2\cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right)\ .
+&= \sigma^2\delta_{(m+a)\,,m'}+\sum_{j=1}^J\Bigg( \dfrac{\Omega_j^2}2\cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right) + \dfrac{\Omega_j^2}{2K}\underbrace{\sum_{k=0}^{K-1}\cos\left(2\pi \frac{f_j}{\fs}(2k+m+a+m'+2N_0)\right)}_{=0\ \mathrm{because}\ \frac{f_j}\fs=\frac{p'_j}K } \Bigg)\\
+&= \sigma^2\delta_{(m+a)\,,m'}+\sum_{j=1}^J\dfrac{\Omega_j^2}2\cos\left(2\pi \frac{f_j}{\fs}(m+a-m')\right)\ .
 \label{eq:S1}
@@
 which leads to
-\begin{equation*}
+\begin{equation}
 {\bS^{(0)}}^\inv[m,m']  = \dfrac1{\sigma^2}\delta_{m,m'}-\dfrac{2}{M\sigma^2}\sum_{j=1}^J\dfrac{1}{1+\frac{4\sigma^2}{M\Omega_j^2}}\cos\left(2\pi p_j \dfrac{m-m'}{M}\right)\,.
-\end{equation*}
+\label{eq:S0.inv.2}
+\end{equation}
 Directly, we have:
@@
 \]
-Furthermore, combining equations~\eqref{eq:S1} and~\eqref{eq:S0.inv}, we have
+Furthermore, combining equations~\eqref{eq:S1} and~\eqref{eq:S0.inv.2}, we have
 \begin{align*}
@@
 Assume the deterministic signal $\bz$ takes the form~\eqref{I-eq:sum.sine}, and the observed noisy signal takes the form~\eqref{I-eq:model.noise}.
-Then, the second-order moments of $\bg$ are bounded as follows:
+Then, as $K\to\infty$, the second-order moments of $\bg$ are bounded as follows:
 \begin{equation}
@@
 \begin{align*}
-\bg_1[r] &= \bE_1^{(a_r)}[m_r,m'_r] = \dfrac1K\sum_{k=0}^{K-1} \bz_k[m_r+a_r]\bw_k[m'_r] + \bw_k[m_r+a_r]\bz_k[m'_r] \\
-\bg_2[r] &= \bE_2^{(a_r)}[m_r,m'_r] = \dfrac1K\sum_{k=0}^{K-1} \bw_k[m_r+a_r]\bw_k[m'_r] - \delta_{m_r+a_r,\,m'_r} \ ,
+\bg_1[r] &= \bE_1^{(a_r)}[m_r,m'_r] = \dfrac1K\sum_{k=0}^{K-1} \bz_k[m_r+a_r]\bw_k[m'_r] + \bw_k[m_r+a_r]\bz_k[m'_r]\,, \\
+\bg_2[r] &= \bE_2^{(a_r)}[m_r,m'_r] = \dfrac1K\sum_{k=0}^{K-1} \bw_k[m_r+a_r]\bw_k[m'_r] - \delta_{m_r+a_r,\,m'_r} \,,
 \end{align*}
@@
 \nonumber
-&=\dfrac1{K^T}\EE\left\{\prod_{\theta=1}^T \left(\sum_{k=0}^{K-1}\rho_{\theta,k}(\bw[k+m'_{r_\theta}],\bw[k+m_{r_\theta}+a_{r_\theta}]) \right)\right\}\\ 
-&=\dfrac1{K^T}\sum_{k_1=0}^{K-1}\cdots\sum_{k_T=0}^{K-1}\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]) \right\} \,,
+&=\dfrac1{K^T}\EE\left\{\prod_{\theta=1}^T \left(\sum_{k=0}^{K-1}\rho_{\theta,k}\left(\bw[k+m'_{r_\theta}],\bw[k+m_{r_\theta}+a_{r_\theta}]\right) \right)\right\}\\ 
+&=\dfrac1{K^T}\sum_{k_1=0}^{K-1}\cdots\sum_{k_T=0}^{K-1}\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}\left(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]\right) \right\} \,,
 \label{eq:multisum}
@@
 Thus,
-\begin{align*}
-\left|\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}\right| &\leq \dfrac{\nu_K}{K^T} C\ ,
-\end{align*}
+\begin{equation*}
+\left|\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}\right| \leq \dfrac{\nu_K}{K^T} C_T\ ,
+\end{equation*}
 where
 \[
-C = \max_{(k_1,\ldots,k_T)} \left|\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]) \right\} \right| \,,
+C_T = \max_{(k_1,\ldots,k_T)} \left|\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]) \right\} \right| \,,
 \]
 and $\nu_K$ is the number of non-zero terms in the sum~\eqref{eq:multisum}.
-Note that $C$ is independent of $K$ (but depends on $\bz$ and $\sigma$).
+Note that $C_T$ is independent of $K$ (but depends on $\bz$ and $\sigma$).
 The behavior of the order $T$ moment in function of $K$ is therefore only determined by the ratio $\dfrac{\nu_K}{K^T}$.
@@
 Fix $k_1$.
-For each of the other sum indexes $k_\theta$ ($\theta\in\{2,\ldots,T\}$), there are four values that make $\rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}])$ not independent of $\rho_{1,k_1}(\bw[k_1+m'_{r_1}],\bw[k_1+m_{r_1}+a_{r_1}])$.
+For each of the other indexes of summation $k_\theta$ ($\theta\in\{2,\ldots,T\}$), there are four values that make $\rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}])$ not independent of $\rho_{1,k_1}(\bw[k_1+m'_{r_1}],\bw[k_1+m_{r_1}+a_{r_1}])$.
 Indeed, since $\bw$ is a white noise these quantities are independent except when $k_\theta$ is such that:
@@
 \end{align*}
-Consequently, there exist at least $(K-4)^{T-1}$ combinations of $(k_2,\ldots,k_T)$ such that the correlation term given by $\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]) \right\}$ is zero.
+Consequently, for each value of $k_1$ there exist at least $(K-4)^{T-1}$ combinations of $(k_2,\ldots,k_T)$ such that the correlation term given by $\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]) \right\}$ is zero.
 Because $T\geq 3$, we develop similar arguments on $k_2$ and $k_3$ to obtain the following maximum bound on the number of non-zero terms in the sum~\eqref{eq:multisum}:
@@
 \begin{align*}
-\left|\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}\right| &\leq C \dfrac{K^T - 3K(K-4)^{T-1} + 3K(K-4)(K-8)^{T-2}- K(K-4)(K-8)(K-12)^{T-3}}{K^T} \\
-&\leq C \left( 1 - 3\left(1-\frac4K\right)^{T-1} + 3\left(1-\frac4K\right)\left(1-\frac8K\right)^{T-2}- \left(1-\frac4K\right)\left(1-\frac8K\right)\left(1-\frac{12}{K}\right)^{T-3} \right) \\
-&\leq C \left( 1 - 3 +\frac{12(T-1)}{K} + 3-\frac{12}K-\frac{24(T-2)}{K} -1 + \frac{4}K + \frac{8}K +\frac{12(T-3)}{K} + o\left(\dfrac1K\right)\right) \\
-&\leq o\left(\dfrac1K\right)\ . \\
+\left|\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}\right| &\leq C_T \dfrac{K^T - 3K(K-4)^{T-1} + 3K(K-4)(K-8)^{T-2}- K(K-4)(K-8)(K-12)^{T-3}}{K^T} \\
+&\leq C_T \left( 1 - 3\left(1-\frac4K\right)^{T-1} + 3\left(1-\frac4K\right)\left(1-\frac8K\right)^{T-2}- \left(1-\frac4K\right)\left(1-\frac8K\right)\left(1-\frac{12}{K}\right)^{T-3} \right) \\
+&\leq C_T \left( 1 - 3 +\frac{12(T-1)}{K} + 3-\frac{12}K-\frac{24(T-2)}{K} -1 + \frac{4}K + \frac{8}K +\frac{12(T-3)}{K} + o\left(\dfrac1K\right)\right) \\
+&\leq C_T\ o\left(\dfrac1K\right)\ . \\
 \end{align*}
@@
 \begin{align*}
-d_{2,\bz,M,\ell} &\defeq d_{2,M,\ell}C_{\bz,M}^2\max\left(1,\frac{2J}{M}\right)^{\ell-2} + 2d'_{2,M,\ell}\max\left(1,\frac{2J}{M}\right)^{\ell-1}\left(1+\max\left(1,\frac{2J}{M}\right)\right)\left(1+\frac{2J}{M}\right)^2\,,
+d_{2,\bz,M,\ell} &\defeq \left(1+\frac{2J}{M}\right)^2 \left(\max\left(1,\frac{2J}{M}\right)\right)^{\ell-2}\left(1+\max\left(1,\frac{2J}{M}\right)\right)\left(d_{2,M,\ell} + (d_{2,M,\ell}+2d'_{2,M,\ell})\max\left(1,\frac{2J}{M}\right)\right)\,,
 \end{align*}

### 2020-09-09 Notes/Supplementary material/SupMat.tex
proof
typos

@@
 \begin{equation*}
-\tilde\bA = (\bS^{(1)}+\bE^{(1)})(\bS^{(0)}+\bE^{(0)})^\inv\,.
+\tilde\bA = \left(\bS^{(1)}+\bE^{(1)}\right)\left(\bS^{(0)}+\bE^{(0)}\right)^{-1}\,.
 \end{equation*}
@@
 \[
-\left\|{\bS^{(0)}}^\inv[m,m']\right\|_{\max}\leq\dfrac1{\sigma^2}\left( 1+\dfrac{2J}{M} \right)\ .
+\left\|{\bS^{(0)}}^\inv\right\|_{\max}\leq\dfrac1{\sigma^2}\left( 1+\dfrac{2J}{M} \right)\ .
 \]
@@
 
-\paragraph{If $n'\geq N$.}
+\paragraph{If $n'\geq N$}
 When $n\geq N$ and $n'\geq N$, applying the Cauchy-Schwarz inequality, we obtain the following bound on the covariance $\gamma[n,n]$:
@@
 
-\paragraph{If $n'\in I$.}
+\paragraph{If $n'\in I$}
 When $n\geq N$ and $n'\in I$, equation~\eqref{I-eq:cov00} shows us that:

### 2020-09-09 Notes/Supplementary material/SupMat.tex
typos

@@
 \bz_0 & \cdots & \bz_{K-1}
-\end{pmatrix}\\ 
+\end{pmatrix}\,,\\ 
 \bZ' &\defeq \begin{pmatrix}
@@
 Consequently, for each value of $k_1$ there exist at least $(K-4)^{T-1}$ combinations of $(k_2,\ldots,k_T)$ such that the correlation term given by $\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]) \right\}$ is zero.
-Because $T\geq 3$, we develop similar arguments on $k_2$ and $k_3$ to obtain the following maximum bound on the number of non-zero terms in the sum~\eqref{eq:multisum}:
+Because $T\geq 3$, we develop similar arguments on $k_2$ and $k_3$ to determine other cases where this correlation term vanishes.
+We subtract these cases to $K^T$, the total number of combinations of $(k_1,\ldots,k_T)$ to obtain the following maximum bound on the number of non-zero terms in the sum~\eqref{eq:multisum}:
 \[

### 2020-09-09 Notes/Supplementary material/SupMat.tex
proof

@@
 \[
-\left\|\bA_0\right\|_{\max}\leq \max\left(1,\frac{2J}{M}\right)\ .
+\left\|\bA_0\right\|_{\max}\leq \max\left(1,\frac{2J}{M}\right)\ .\qedhere
 \]
@@
 \]
-and $\nu_K$ is the number of non-zero terms in the sum~\eqref{eq:multisum}.
+and $\nu_K$ is the number of nonzero terms in the sum~\eqref{eq:multisum}.
 Note that $C_T$ is independent of $K$ (but depends on $\bz$ and $\sigma$).
@@
 \end{align*}
-Consequently, for each value of $k_1$ there exist at least $(K-4)^{T-1}$ combinations of $(k_2,\ldots,k_T)$ such that the correlation term given by $\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]) \right\}$ is zero.
+Consequently, for each value of $k_1$ there exist at least $(K-4)^{T-1}$ combinations of $(k_2,\ldots,k_T)$ where we have
+\begin{align*}
+\EE\left\{\prod_{\theta=1}^T \rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]) \right\} &= \EE\left\{ \rho_{1,k_1}(\bw[k_1+m'_{r_1}],\bw[k_1+m_{r_1}+a_{r_1}]) \right\} \\
+&\hspace{20pt}\times\EE\left\{\prod_{\theta=2}^T \rho_{\theta,k_\theta}(\bw[k_\theta+m'_{r_\theta}],\bw[k_\theta+m_{r_\theta}+a_{r_\theta}]) \right\}
+& = 0\,,
+\end{align*}
+because $\EE\left\{ \rho_{1,k_1}(\bw[k_1+m'_{r_1}],\bw[k_1+m_{r_1}+a_{r_1}]) \right\}=0$.
+Therefore, at least $K(K-4)^{T-1}$ of the sum~\eqref{eq:multisum} are zero.
 Because $T\geq 3$, we develop similar arguments on $k_2$ and $k_3$ to determine other cases where this correlation term vanishes.
-We subtract these cases to $K^T$, the total number of combinations of $(k_1,\ldots,k_T)$ to obtain the following maximum bound on the number of non-zero terms in the sum~\eqref{eq:multisum}:
+We subtract these cases to $K^T$, the total number of combinations of $(k_1,\ldots,k_T)$ to obtain the following maximum bound on the number of nonzero terms in the sum~\eqref{eq:multisum}:
 \[
@@
 \begin{equation*}
-\left|\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}\right| = o\left(\dfrac1K\right)\,,\quad\forall T\geq 3\ .
+\left|\EE\left\{\prod_{\theta=1}^T\bg[r_\theta]\right\}\right| = o\left(\dfrac1K\right)\,,\quad\forall T\geq 3\ .\qedhere
 \end{equation*}
@@
 \begin{equation*}
-\left|\left.\dfrac{\partial^2 f_m^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero}\right| \leq \dfrac{d_{2,\bz,M,\ell}}{\sigma^4}\ .
+\left|\left.\dfrac{\partial^2 f_m^{(\ell)}}{\partial \bg[r]\partial \bg[r']}\right\vert_{\bg=\bzero}\right| \leq \dfrac{d_{2,\bz,M,\ell}}{\sigma^4}\ .\qedhere
 \end{equation*}

### 2020-09-10 Notes/BoundEffRed.tex
affilations

@@
         Hau-Tieng~Wu
-\thanks{A. Meynard and H.-T. Wu are with the Department
+\thanks{A. Meynard is with the Department
 of Mathematics, Duke University, Durham,
-NC, 27708 USA e-mail: adrien.meynard@duke.edu}}
+NC, 27708 USA.
+
+H.-T.
+Wu is with the Department of Mathematics and Department of Statistical Science, Duke University, Durham, NC, 27708 USA; Mathematics Division, National Center for Theoretical Sciences, Taipei, Taiwan.
+
+A. Meynard is the corresponding author. e-mail: adrien.meynard@duke.edu
+}}
 

### 2020-09-10 Notes/Supplementary material/SupMat.tex
affilations

@@
         Hau-Tieng~Wu
-\thanks{A. Meynard and H.-T. Wu are with the Department
+\thanks{A. Meynard is with the Department
 of Mathematics, Duke University, Durham,
-NC, 27708 USA e-mail: adrien.meynard@duke.edu}}
+NC, 27708 USA.
+
+H.-T.
+Wu is with the Department of Mathematics and Department of Statistical Science, Duke University, Durham, NC, 27708 USA; Mathematics Division, National Center for Theoretical Sciences, Taipei, Taiwan.
+
+A. Meynard is the corresponding author. e-mail: adrien.meynard@duke.edu}}
 

### 2020-09-10 Notes/BoundEffRed.tex
typos

@@
 
-A. Meynard is the corresponding author. e-mail: adrien.meynard@duke.edu
+A. Meynard is the corresponding author (e-mail: adrien.meynard@duke.edu).
 }}
@@
 
-Usually, the measured signals exhibit nonstationary behavior, and the observed quantities might be interfered by transient phenomena that can vary rapidly and irregularly.
+Usually, the measured signals exhibit nonstationary behavior, and the observed quantities might be interfered with by transient phenomena that can vary rapidly and irregularly.
 In this paper, we focus on oscillatory time series.
-These signals might, for example, oscillate fast with large amplitude at one moment, and then oscillate slowly with small amplitude at the next moment.
+These signals might, for example, oscillate fast with large amplitudes at one moment, and then oscillate slowly with small amplitudes at the next moment.
 In order to adapt the analysis to nonstationarities, local spectral analysis is generally performed~\cite{Stoica05spectral,Matz97generalized}.
@@
 Attempts to minimize the boundary effects generally consists in softening the discontinuity on signal edges.
-Usually the approaches can be classified into two main classes, choosing a proper analysis window and extending/forecasting the signal.
+Usually, the approaches can be classified into two main classes, choosing a proper analysis window and extending/forecasting the signal.
 In the first class, judicious choice of analysis window whose support does not interfere with the boundary points can minimize the occurrence of aberrant patterns near the boundaries of the TF plane, for instance, \cite{Chui92wavelets,Depczynski99fast}.
@@
 %
-The above is a far from exhaustive list, and we refer readers with interest to \cite{hyndman2018forecasting} for a friendly monograph on the general forecasting topic. 
+The above list is far from exhaustive, and we refer readers with interest to \cite{hyndman2018forecasting} for a friendly monograph on the general forecasting topic. 
 %
@@
 The numerical results also confirmed the robustness to noise of {\sf BoundEffRed}, as well as its ability to be applied to many TF representations.
-An additional application to ECG is included into the supplementary material.
+An additional application to ECG is included in the Supplementary Materials.
 
@@
 In particular, we have noticed (see Fig.~\ref{fig:THO.failure}) that when the regular oscillations of the observed signal break, the forecasting step is no longer relevant, and only slows down the calculation of the TF representation.
-A preliminary step should then be added to the algorithm to detect signal activity, and disable the forecasting step when possible.
+A preliminary step should then be added to the algorithm to detect signal activity and disable the forecasting step when possible.
 More fundamentally, one can also consider accelerating the computational time by optimizing the forecasting step to improve the real-time performance of {\sf BoundEffRed}.
@@
 Their algorithm is limited to the case where $H=1$.
-Developing an extension of this algorithm to variations of $H>1$ columns would then allow to efficiently update the matrix $\left(\bX\bX^T\right)^{-1}$.
+Developing an extension of this algorithm to variations of $H>1$ columns would then allow us to efficiently update the matrix $\left(\bX\bX^T\right)^{-1}$.
 These avenues will be explored in our future work.

### 2020-09-10 Notes/algorithm.tex
proof
typos

@@
 where $\fs$ denotes the sampling frequency. 
-Let $M$ and $K$ be two integers such that $M<N$ and $K+M<N$. Then, for all $k\in\{0,\ldots,K-1\}$, we extract from $\bx\in\RR^N$ the subsignal $\bx_k\in\RR^M$ given by:
+Let $M$ and $K$ be two positive integers such that $M<K$ and $K+M\leq N$. Then, for all $k\in\{0,\ldots,K-1\}$, we extract from $\bx\in\RR^N$ the subsignal $\bx_k\in\RR^M$ given by:
 \begin{equation}
@@
 \begin{pmatrix}
-\bx[N-K+(k-1)-(M-1)] \\
+\bx[N-K-M+k] \\
 \vdots \\
-\bx[N-K+(k-1)]
+\bx[N-K+k-1]
 \end{pmatrix}\ .
@@
 In other words, we adopt a classical strategy in the study of dynamical systems, the {\em linearization of a nonlinear system}, when the system is sufficiently regular.
-Notice that this linearized dynamical model can be written equivalently in function of the subsignals $\bx_k$, as:
+Notice that this linearized dynamical model can be written equivalently according to the subsignals $\bx_k$, as:
 \begin{equation}

### 2020-09-11 Notes/BoundEffRed.tex
typo

@@
 In the first class, judicious choice of analysis window whose support does not interfere with the boundary points can minimize the occurrence of aberrant patterns near the boundaries of the TF plane, for instance, \cite{Chui92wavelets,Depczynski99fast}.
-Due to the specific relationship of chosen analysis windows and TF analysis tool, these techniques do not make it possible to reduce the boundary effects of all TF representations.
+Due to the specific relationship between the chosen analysis windows and the TF analysis tool, these techniques do not make it possible to reduce the boundary effects of all TF representations.
 %

### 2020-09-11 Notes/Supplementary material/SupMat.tex
typos
typo

@@
 
-\title{Supplementary material for ``An Efficient Forecasting Approach to Reduce Boundary Effects in Real-Time Time-Frequency Analysis''}
+\title{Supplementary Materials for ``An Efficient Forecasting Approach to Reduce Boundary Effects in Real-Time Time-Frequency Analysis''}
 
@@
 
-A. Meynard is the corresponding author. e-mail: adrien.meynard@duke.edu}}
+A. Meynard is the corresponding author (e-mail: adrien.meynard@duke.edu).}}
 

### 2020-09-11 Paper/results.tex
folder name

--- 2020-09-03 Notes/results.tex
@@
 \item The TBATS method~\cite{DeLivera11forecasting} is based on a classical decomposition of times series into a trend, seasonal and ARMA components, with a specific dynamic for the seasonal component.
-This model demands the estimation of numerous parameters and, by implication, may be slow.
+This model demands the estimation of numerous parameters and, by implication, may require a long computation time.
 \end{itemize}
@@
 For each of these values, we determine the experimental bias, denoted as $\mu_{\xp}[N-1+\ell]$, and experimental variance, denoted as $\gamma_{\xp}[N-1+\ell,N-1+\ell]$.
-Fig.~\ref{fig:res.noise.sine} shows the the experimental forecasting variance in function of the noise variance for three different values the forecasting sample index: $\ell=1$, $\ell=10$, and $\ell=100$.
+Fig.~\ref{fig:res.noise.sine} shows the experimental forecasting variance as a function of the noise variance for three different values the forecasting sample index: $\ell=1$, $\ell=10$, and $\ell=100$.
 
@@
 For each of these values, we determine the experimental bias $\mu_{\xp}[N-1+\ell]$ and variance $\gamma_{\xp}[N-1+\ell,N-1+\ell]$.
-Fig.~\ref{fig:res.size.sine} shows the the experimental forecasting variance in function of the training dataset size for three different values the forecasting sample index: $\ell=1$, $\ell=10$, and $\ell=100$.
+Fig.~\ref{fig:res.size.sine} shows the experimental forecasting variance as a function of the training dataset size for three different values the forecasting sample index: $\ell=1$, $\ell=10$, and $\ell=100$.
 
@@
 As anticipated by Theorem~\ref{th:error}, the variance of the estimator decreases linearly with $\frac1K$ as long as the dataset size is large enough.
-Indeed, in the log-log plot displayed in Fig.~\ref{fig:res.size.sine}, as soon as $K>700$, each of the three curves representing the forecasting variance in function of $K$ tends towards an asymptote of slope $-1$.
+Indeed, in the log-log plot displayed in Fig.~\ref{fig:res.size.sine}, as soon as $K>700$, each of the three curves representing the forecasting variance as a function of $K$ tends towards an asymptote of slope $-1$.
 
@@
 %
-Note that the MSE of {\sf SigExt} is in average higher than the MSE of EDMD or the symmetric extension, with a huge standard deviation.
+Note that the MSE of {\sf SigExt} is on average higher than the MSE of EDMD or the symmetric extension, with a huge standard deviation.
 This variability is caused by the presence of few segments contaminated by artifacts so that it is unpredictable via a too simple dynamical model like~\eqref{eq:dyn.model}.
@@
 To prevent the STFT from being sensitive to the boundaries, we set $L=700$.
-In this way, the evaluation of the spectral content of the signal near its boundary is not limited by a lack of information all along the window support.
+In this way, the evaluation of the spectral content of the signal near its boundary is not limited by a lack of information in the interval delimited by the window support.
 From now on, all results are given for $L$ equal to the half of the width of the window used in the TF transform.
@@
 A t-test is performed to compare the performance index of {\sf SigExt} with those of other methods.
-Under the null hypothesis that the means are equal, with a $5\%$ significance level, the t-tests show no statistical significant difference between {\sf SigExt} and EDMD or GPR, regardless of the representation considered.
+Under the null hypothesis that the means are equal, with a $5\%$ significance level, the t-tests show no statistically significant difference between {\sf SigExt} and EDMD or GPR, regardless of the representation considered.
 
@@
 For all the TF representations considered, the results show that {\sf SigExt} reduces boundary effects about as efficiently as the extensions given by EDMD or GPR.
-As for the respiratory signal, t-tests show no statistical significant difference between {\sf SigExt} and EDMD or GPR, regardless of the representation considered.
-For a visual inspection, the TF representation of SST resulting from the {\sf BoundEffRed} strategy is shown in the bottom-right panel of Fig.~\ref{fig:ex.intro}, where SST is applied to the portion of PPG displayed in Fig.~\ref{fig:ppg}.
+As for the respiratory signal, t-tests show no statistically significant difference between {\sf SigExt} and EDMD or GPR, regardless of the representation considered.
+For visual inspection, the TF representation of SST resulting from the {\sf BoundEffRed} strategy is shown in the bottom-right panel of Fig.~\ref{fig:ex.intro}, where SST is applied to the portion of PPG displayed in Fig.~\ref{fig:ppg}.
 It produces a significant improvement in the quality of the SST near boundaries.
@@
 It is thus an additional noise to the measurement noise actually contained in the signal.
-Fig.~\ref{fig:otd.noise} shows the averaged performance index of {\sf BoundEffRed} for different values of the Signal-to-Noise Ratio (SNR).
+Fig.~\ref{fig:otd.noise} shows the average performance index of {\sf BoundEffRed} for different values of the Signal-to-Noise Ratio (SNR).
 Whatever the representation considered, {\sf BoundEffRed} is relatively robust to noise, as long as the SNR remains above $10$~dB.

### 2020-09-11 Paper/theoryOLD.tex
folder name

\subsection{Signal Model}
We model the deterministic part of the observed signal as a multicomponent harmonic signal; that is, a sum of sine waves:
\begin{equation}
\bz[n]=\sum_{j=1}^J\Omega_j\cos\left(2\pi f_j \frac{n}{\fs} + \varphi_j\right)\ ,
\label{eq:sum.sine}
\end{equation}
where $J$ denotes the number of components, $\Omega_j>0$ the amplitude of the $j$-th component, $f_j$ its frequency, and $\varphi_j\in[0,2\pi)$ its initial phase.
%
For the sake of simplicity, we make an additional assumption on the frequencies of each component. We assume that for all $j\in\{1,\dots,J\}$:
\begin{equation}
\exists\, p_j,p_j'\in\NN^*: \quad f_j = \dfrac{p_j}{M}\fs = \dfrac{p'_j}{K}\fs\ .
\end{equation}

In addition, the observed signal is assumed to be corrupted by an additive Gaussian white noise. Therefore, the measured discrete signal $\bx$ is written as:
\begin{equation}
\bx = \bz + \sigma\bw\ ,
\label{eq:model.noise}
\end{equation}
where $\bz$ follows model~\eqref{eq:sum.sine}, $\bw$ is a Gaussian white noise, whose variance is normalized to one. Thus, $\sigma^2$ denotes the variance of the additive noise $\sigma\bw$.

\subsection{Forecasting Error}
On the forecasting interval, we decompose the estimated signal $\tilde\bx$ as follows:
\begin{equation}
\tilde \bx[n] = \bz[n] + \bepsilon[n]\ ,
\label{eq:forecasting.error}
\end{equation}
where $\bepsilon$ is the forecasting error. When $n\in I=\{0,\dots,N-1\}$, this error contains only the measurement noise, that is $\bepsilon[n]=\sigma\bw[n]$. Outside the interval $I$, the importance of the forecasting error $\bepsilon$ is also affected by the loss of information resulting from the linearization of the dynamical model we consider in~\eqref{eq:dyn.model}. To evaluate the actual behavior of the forward forecasting error $\bepsilon[n]$ when $n\geq N$, we determine its first two moments.
\begin{enumerate}
\item 
The mean, or estimation bias, is such that: 
\begin{align}
\bmu[n]&\defeq\EE\{\bepsilon[n]\} 
=\EE\{\tilde\bx[n]\} - \bz[n]\ .\label{eq:defn:bias0}
\end{align}
Given the forecasting strategy, we have $\bmu[n]=0$ when $n\in I$ and
\begin{equation}
\bmu[n]= \EE\{\balpha^{(n-N+1)}\}\bz_{K} + \sigma\EE\{\balpha^{(n-N+1)}\bw_{K}\} - \bz[n]
\label{eq:bias0}
\end{equation}
when $n\geq N$.
\item 
The covariance is given by:
\begin{align*}
\bgamma[n,n'] &\defeq \EE\{\left(\bepsilon[n]-\bmu[n]\right)\left(\bepsilon[n']-\bmu[n']\right)\}\\
&= \EE\{\tilde\bx[n]\tilde\bx[n']\}-\bz[n]\bz[n'] -\bmu[n]\bz[n'] \\
&\hspace{15pt} - \bmu[n']\bz[n] -\bmu[n]\bmu[n']\ .
\end{align*}
Thus by definition of the noise, we have $\bgamma[n,n'] =\sigma^2\delta_{n,n'}$ when $(n,n')\in I^2$. When $n\geq N$, let us denote $\ell=n-N+1$. Then, we have two cases.
\begin{enumerate}[label=(\roman*)]
\item If $n'\in I$:
\begin{align}
\nonumber
\bgamma[n,n'] &= \sigma\EE\{\bw[n']\balpha^{(\ell)}\}\bz_K + \sigma^2\EE\{\bw[n']\balpha^{(\ell)}\bw_K\} \\
&\hspace{15pt} -\bz[n]\bmu[n'] - \bmu[n]\bmu[n']\ .
\label{eq:cov00}
\end{align}
\item If $n'=N-1+\lambda\geq N$:
\begin{align}
\nonumber
\hspace{-6pt}\bgamma[n,n'] &= \bz_K^T\EE\left\{{\balpha^{(\ell)}}^T\balpha^{(\lambda)}\right\}\bz_K + \sigma\EE\{\balpha^{(\ell)}\bw_K\balpha^{(\lambda)}\}\bz_K \\
\nonumber
&\hspace{-3pt} + \sigma\EE\{\balpha^{(\lambda)}\bw_K\balpha^{(\ell)}\}\bz_K + \sigma^2\EE\{\balpha^{(\ell)}\bw_K\balpha^{(\lambda)}\bw_K\}  \\
&\hspace{-3pt} -\!\bz[n]\bz[n']-\!\bz[n]\bmu[n']-\!\bz[n]\bmu[n'] -\! \bmu[n]\bmu[n']\,.
\label{eq:cov01}
\end{align}
\end{enumerate}
Besides, we recall that $\bgamma[n,n']=\bgamma[n',n]$.
\end{enumerate}

Expressions~\eqref{eq:bias0},~\eqref{eq:cov00}, and~\eqref{eq:cov01} show that these quantities depend on the behavior of the forecasting random vector $\balpha^{(\ell)}$. In Lemma~\ref{lm:error}, we specify the asymptotic behavior of the forecasting vector $\balpha^{(\ell)}$ when the dataset size $K$ is great.
\begin{lemma}
\label{lm:error}
Let $\bx$ be a random vector defined by~\eqref{eq:model.noise}. Let $\balpha^{(\ell)}$ be the associated forecasting vector for the estimation of $\bx[N-1+\ell]$, given by~\eqref{eq:alpha.l} and obtained from the least square estimation~\eqref{eq:lse}. Let $\balpha_0^{(\ell)}$ be the last row of the matrix $\bA_0^\ell$, where $\bA_0$ is defined by:
\begin{equation}
\bA_0 = \left(\dfrac1K\bZ'\bZ^T+\sigma^2\bD\right)\left(\dfrac1K\bZ\bZ^T+\sigma^2\bI\right)^{\inv}\ ,
\label{eq:lse.0}
\end{equation}
where $\bZ=\begin{pmatrix} \bz_0 & \cdots & \bz_{K-1}\end{pmatrix}$, $\bZ'=\begin{pmatrix} \bz_1 & \cdots & \bz_{K}\end{pmatrix}$, $\bz_k$ is the $k$-th subsignal extracted from $\bz$ in the same way as $\bx_k$ is defined from $\bx$ in~\eqref{eq:xk}, and $\bD\in\RR^{M\times M}$ is the Toeplitz matrix such that $\bD[m,m'] = \delta_{m+1,m'}$.

Let $\bh^{(\ell)}$ be the error vector given by
\begin{equation*}
\bh^{(\ell)} = \balpha^{(\ell)} - \balpha_0^{(\ell)}\, .
\end{equation*}
Then, the random vector $\bh^{(\ell)}$ converges in law to a zero-mean Gaussian random vector when $K\to\infty$; that is,
\begin{equation}
\sqrt{K}\,\bh^{(\ell)}  \xrightarrow[K\to\infty]{\cD} \cN\left(\bzero,\bGamma^{(\ell,\ell)}\right)\ ,
\end{equation}
with $\bGamma^{(\ell,\ell)}:={\bF^{(\ell)}}^T\bGamma_0\bF^{(\ell)}$, where $\bGamma_0\in\RR^{M(M+1)\times M(M+1)}$ is a covariance matrix and $\bF^{(\ell)}\in\RR^{M(M+1)\times M}$ is a Jacobian matrix, and $\bGamma^{(\ell,\ell)}$ does not depend on $\sigma$. 
% such that:
%{\color{red}
%\begin{align*}
%\bGamma_0[mM+n,pM+q] &= 2\sigma^2\sum_{j=1}^J\Omega_j^2\cos\left(2\pi\frac{f_j}{\fs}(m-n)\right)\\
%& \hspace{26pt} \times\cos\left(2\pi\frac{f_j}{\fs}(p-q)\right) + o(\sigma^2) \\
%\bF^{(\ell)}[] &= \ . 
%\end{align*}
%}
\end{lemma}

\begin{proof}
See the Supplementary Material. The proof is based on the multivariate delta method (see paragraph 7.2 in~\cite{Alho05statistical}), which gives an asymptotic approximation of a random vector by a Gaussian random vector.
\end{proof}

Consequently, the covariance between $\sqrt{K}\bh^{(\ell)}$ and $\sqrt{K}\bh^{(\lambda)}$ remains bounded, \ie:
\begin{equation*}
K\,\EE\left\{{\bh^{(\ell)}}^T\bh^{(\lambda)}\right\} \xrightarrow[K\to\infty]{} \bGamma^{(\ell,\lambda)} = {\bF^{(\ell)}}^T\bGamma_0\bF^{(\lambda)}\ .
\end{equation*}

\begin{theorem}
\label{th:error}
Let $\bx\in\RR^N$ be a discrete-time random signal following model~\eqref{eq:model.noise}. Let $\tilde\bx$ denotes its forecasting, obtained using the extension Algorithm~\ref{alg:extension}. Let $n\geq N$ be a sample index. Then, the first-order moment of the forecasting error $\epsilon[n]$ in~\eqref{eq:forecasting.error} defined in \eqref{eq:defn:bias0} satisfies
\begin{equation}
\frac{1}{\sigma}\bmu[n] = o(1)
\label{eq:mean.error}
\end{equation}
when $K\to\infty$.
Its second-order moment $\bgamma[n,n']$ satisfies the following approximation equations:
\begin{enumerate}[label=(\roman*)]
\item if $n'\in I=\{0,\ldots,N-1\}$:
\begin{equation}
\bgamma[n,n']\to \sigma^2\balpha_0^{(n-N-1)}[n'-(N-M)]\1_{(n'\geq N-M)}
\label{eq:cov.error.1}
\end{equation}
when $K\to\infty$;
\item if $n'\geq N$:
\begin{equation}
\hspace{-15pt}\bgamma[n,n'] = \dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K + \dfrac{\sigma^2}{K}\Tr\left(\bGamma^{(\ell,\lambda)}\right) + \sigma^2\left\langle\balpha_0^{(\ell)}\!, \balpha_0^{(\lambda)}\right\rangle ,
\label{eq:cov.error.2}
\end{equation}
when $K\to\infty$, where $\ell\!=n-N+1$ and $\lambda\!=n'-N+1$.
\end{enumerate}
\end{theorem}

\begin{proof}
See the Supplementary Material. The proof is mainly based on the results provided by Lemma~\ref{lm:error}, combined with the Isserlis' theorem~\cite{Isserlis16formula}, which provides a formula for the computation of higher-order moments of Gaussian random variables.
\end{proof}

Ideally, the forecasting error would behave like the measurement noise $\sigma\bw$, \ie~a zero-mean noise whose variance is of the order of $\sigma^2$. Theorem~\ref{th:error} shows that the forecasting error is asymptotically unbiased. Concerning the covariance of the forecasting error, although equations~\eqref{eq:cov.error.1} and~\eqref{eq:cov.error.2} are not easily readable, one can evaluate the dependence of the variance in function of the tuning parameters, that are adjusted by the user. Let us focus on the forecasting error variance $\gamma[n,n]$ when $n\geq N$. First, as expected, the variance increases linearly with the noise variance $\sigma^2$. Second, it asymptotically depends linearly on the ratio $\frac{1}{K}$. This shows the need to use a sufficiently large dataset to obtain an accurate forecast. Third, the dependency on the subsignals lengths $M$ and the forecasting index $\ell=n-N+1$ is hidden in the expression of the covariance matrix $\bGamma^{(\ell,\ell)}$. We discuss this dependency in more detail in section~\ref{ssse:res.sine}.

\subsection{Adaptive Harmonic Model}\label{RemarkAHM}

One can extend the previous result to the case where the instantaneous frequencies and amplitudes of the components of the deterministic part of the observed signal are {\em slowly varying}. It is an \textit{AM-FM model} that, in its is continuous-time version, takes the following form
\begin{equation}
z(t) = \sum_{j=1}^J a_j(t)\cos(2\pi\phi_j(t))\,,
\end{equation}
where $a_j(t)$ and $\phi'_j(t)$ describe how large and fast the signal oscillate at time $t$. 
Clearly, \eqref{eq:sum.sine} is a special case satisfying the AM-FM model when the $a_j$ and $\phi'_j$ are both constants. 
%
In many practical signals, the amplitude and frequency do not change fast. It is thus reasonable to further restrict the regularity and variations of the instantaneous amplitudes and frequencies of these AM-FM functions. When the speeds of variation of the instantaneous amplitudes $a_j$ and frequencies $\phi'_j$ are small, the signal can be ``locally'' well approximated by a harmonic function in \eqref{eq:sum.sine}; that is, 
%
locally at time $t_0$, $z(t)$ can be well approximated by $z_0(t) := \sum_{j=1}^J a_j(t_0)\cos(2\pi(\phi_j(t_0)-t_0\phi'_j(t_0)+\phi'_j(t_0)t))$ by the Taylor expansion. An AM-FM function satisfying the slow variation of the instantaneous amplitudes $a_j$ and frequencies $\phi'_j$ is referred to the adaptive harmonic model (AHM) (see~\cite{Chen14nonparametric,Daubechies16conceft} for mathematical details).
It is thus clear that the forecasting error caused by the {\sf SigExt} algorithm additionally depends on the speed of variation of the instantaneous amplitudes $a_j$ and frequencies $\phi'_j$. 
%
For the forecasting purpose, it is thus clear that when $K$ is not too large, the signal can be well approximated by \eqref{eq:sum.sine}. Hence, Theorem \ref{th:error} can still be applied to approximate the error. However, how to determine the optimal $K$ based on the signal depends on the application and is out of the scope of this paper. It will be explored in future work for specific application problems.
%\end{remark}

\subsection{Performance of the Boundary Effects Reduction}
\label{sse:perf.BoundEffRed}
While we do not provide a generic proof of the boundary effect reduction on {\em any} TF analysis tools, we discuss a particular case of SST. Since SST is designed to analyze signals satisfying the AHM, as is discussed in Remark \ref{RemarkAHM}, Theorem~\ref{th:error} ensures that the forecasting error $\bepsilon$ defined in~\eqref{eq:forecasting.error} is controlled and bounded in terms of mean and covariance. Recall Theorem 3 in~\cite{Chen14nonparametric}, which states that when the additive noise is stationary and maybe modulated by a smooth function with a small fluctuation, the SST of the observed signal remains close to the SST of the clean signal, throughout the TF plane. 
%
We refer the reader to~\cite{Chen14nonparametric} for a precise quantification of the error made in the TF plane, which depends notably on the covariance of the additive noise and on the speed of variation of the amplitudes and instantaneous frequencies composing the signal. 
%
In our case, while the noise of the historical data is stationary, the forecasting error depends on the historical noise, and hence the overall ``noise'' is non-stationary. However, the dependence only appears in the extended part of the signal. We claim that the proof of Theorem 3 in~\cite{Chen14nonparametric} can be generalized to explain the robustness of SST to the noise, and a systematic proof will be reported in our future theoretical work. While we do not provide a theoretical justification, we verify experimentally that Algorithm~\ref{alg:boundary} is efficient for a large number of representations in the following section~\ref{se:results}. 
%
Therefore, in our case, this means that the boundary effect is strongly reduced since the impact of the forecasting error does SST is limited. An immediate application is that the instantaneous frequencies can now be well estimated continuously up to the edges of the TF plane, and a real-time acquisition of the instantaneous frequency and amplitude information is possible.

### 2021-01-06 Paper/algorithm.tex
scripts

--- 2020-09-10 Notes/algorithm.tex
@@
 Indeed, let $\tilde\bA$ denotes the estimate of $\bA$.
-We then obtain the forecasting of the signal at time $\frac{N-1+\ell}{\fs}$ by:
+{\color{red}We then obtain the forecasting of $\bx_{k+\ell}$ by recursively applying the linear relation~\eqref{eq:dyn.model}:
+\begin{equation*}
+\tilde\bx_{K+\ell} = \underbrace{\tilde\bA\tilde\bA\cdots\tilde\bA}_{\scriptstyle\ell\mbox{ \footnotesize {times}}}\bx_{K} = \tilde\bA^\ell\bx_{K}\ .
+\end{equation*}  
+Let $\be_M$ denote the unit vector of length $M$ given by $\be_M = \begin{pmatrix} 0 & \cdots & 0 & 1\end{pmatrix}^T$.
+Consequently, given definition~\eqref{eq:xk}, the forecasting of the signal at time $\frac{N-1+\ell}{\fs}$ is provided by:
 \begin{equation}
 \label{eq:prediction}
-\tilde \bx [N-1+\ell] = \balpha^{(\ell)}\bx_{K}\ ,
+\tilde \bx [N-1+\ell] = \be_M^T\tilde\bx_{K+\ell} = \balpha^{(\ell)}\bx_{K}\ ,
 \end{equation}  
-where $\balpha^{(\ell)}$ denotes the last row of $\tilde\bA^\ell$; that is to say,
-\begin{equation}
-\balpha^{(\ell)}=\be_M^T\tilde\bA^\ell\ ,
-\label{eq:alpha.l}
-\end{equation}
-where $\be_M$ is the unit vector of length $M$ given by $\be_M = \begin{pmatrix} 0 & \cdots & 0 & 1\end{pmatrix}^T$.
+where $\balpha^{(\ell)}$ denotes the last row of $\tilde\bA^\ell$, \ie~$\balpha^{(\ell)}=\be_M^T\tilde\bA^\ell$.
+}
 

### 2021-01-06 Paper/results.tex
scripts

@@
 \subsection{Extension Methods for Comparison}
+\label{sse:methods}
 To highlight the fact that the linear dynamical model is sufficient to catch most of the dynamical behavior of signals following the AHM, we compare the performance of Algorithm~\ref{alg:boundary} with a simple extension obtained by pointwise symmetrization~\cite{Kharitonenko02wavelet}.
@@
 \includegraphics[width=.48\textwidth]{VarianceNoiseSine.eps}
-\caption{Evolution of the experimental forecasting variance in function of the noise variance for three different values the forecasting sample index $\ell$.}
+\caption{Evolution of the experimental forecasting variance as a function of the noise variance for three different values the forecasting sample index $\ell$.}
 \label{fig:res.noise.sine}
@@
 \begin{align*}
-\phi_1[n] &= \frac{p_1}{M}\left( n + \frac{0.01}{2\pi}\cos\left(2\pi\frac{n}{N}\right) \right) \\[-1mm]
-\phi_2[n] & = p_2\frac{n}{M} + \frac{20}{2N\fs}n^2
+\phi_1[n] &= \frac{p_1}{P}\left( n + \frac{0.01}{2\pi}\cos\left(2\pi\frac{n}{N}\right) \right) \\[-1mm]
+\phi_2[n] & = p_2\frac{n}{P} + \frac{20}{2N\fs}n^2\ .
 \end{align*}
-Besides, the noise is chosen to be Gaussian, $\bw\sim\cN(\bzero,\bI)$, and we take $N=10^4$, $M=750$, $p_1=10$, $p_2=23$.
+Besides, the noise is chosen to be Gaussian, $\bw\sim\cN(\bzero,\bI)$, and we take $N=10^4$, $P=750$, $p_1=10$, $p_2=23$.
 
-The forecasting methods are applied to $1000$ realizations of the above-described signal.
-The mean and the standard deviation (SD) of the resulting MSEs are given in Table~\ref{tab:mse.sine}.
+{\color{red}
+\paragraph{Influence of the subsignals length $M$}
+To evaluate the forecasting quality, {\sf SigExt} is applied to $1000$ realizations of the above-described signal.
+Here, we forecast signal extensions of $0.1$~s, \ie~$L=700$.
+Besides, we set $K=2625$.
+Table~\ref{tab:mse.sine} shows the resulting experimental MSE according to four  different selected values of the subsignals length $M$.
+
+}
+
+\paragraph{Comparison of the extensions methods}
+The same set of experiments is run on the extension methods outlined in section~\ref{sse:methods}.
+For comparison, the mean and the standard deviation (SD) of the resulting MSEs are given in Table~\ref{tab:mse.sine}.
 They show that the naive extension we propose gives satisfying results, in particular in comparison with the point-symmetric extension.
@@
   \hline
-   \multirow{2}{40pt}{\centering Extension method} & \multirow{2}{70pt}{\centering MSE (mean $\pm$ SD)}  & \multirow{2}{80pt}{\centering Averaged computing time (seconds)} \\
+   \multirow{2}{60pt}{\centering Extension method} & \multirow{2}{50pt}{\centering MSE\\ (mean $\pm$ SD)}  & \multirow{2}{80pt}{\centering Averaged computing time (seconds)} \\
     &  & \\
    \hhline{|=#=|=|}
-   {\sf SigExt} & $1.028 \pm  0.404$ & $1.069$ \\
+   {\sf SigExt ($M=100$)} & $1.028 \pm  0.404$ & $1.069$ \\
+   \hline
+   {\sf SigExt ($M=750$)} & $1.028 \pm  0.404$ & $1.069$ \\
+   \hline
+   {\sf SigExt ($M=1500$)} & $1.028 \pm  0.404$ & $1.069$ \\
+   \hline
+   {\sf SigExt ($M=3000$)} & $1.028 \pm  0.404$ & $1.069$ \\
    \hline
@@
 \caption{PPG Signal: Performance of the Boundary-Free TF Representations According to the Extension Method}
-\begin{tabular}{|c||c|c|c|c|}
+\begin{tabular}{|c||c||c|c|c|c|}
   \hline
-   \multirow{2}{40pt}{\centering Extension method} & \multicolumn{4}{c|}{Performance index $D$ (mean $\pm$ SD)} \\
-   \cline{2-5}
-      & STFT & SST & ConceFT & RS\\
-   \hhline{|=#=|=|=|=|}
-   \multirow{2}{40pt}{\centering {\sf SigExt}} & $0.280$ & $0.309$ & $0.367$ & $0.534$ \\
-    & $\pm 0.107$ & $\pm 0.112$ & $\pm 0.183$ & $\pm 0.160$ \\
+   \multirow{2}{38pt}{\centering Extension method} & \multirow{2}{29pt}{\centering MSE} &\multicolumn{4}{c|}{Performance index $D$ (mean $\pm$ SD)} \\
+   \cline{3-6}
+      & & STFT & SST & ConceFT & RS\\
+   \hhline{|=#=#=|=|=|=|}
+   \multirow{2}{38pt}{\centering {\sf SigExt}} & $0.018$ & $0.280$ & $0.309$ & $0.367$ & $0.534$ \\
+    & $\pm 0.014$ & $\pm 0.107$ & $\pm 0.112$ & $\pm 0.183$ & $\pm 0.160$ \\
    \hline
-   \multirow{2}{40pt}{\centering Symmetric} & $1.168$ & $1.209$ & $1.310$ & $0.983$ \\
-   & $\pm 0.390$ & $\pm 0.340$ & $\pm 0.140$ & $\pm 0.304$ \\
+   \multirow{2}{38pt}{\centering Symmetric} & $0.037$ & $1.168$ & $1.209$ & $1.310$ & $0.983$ \\
+   & $\pm 0.007$ & $\pm 0.390$ & $\pm 0.340$ & $\pm 0.140$ & $\pm 0.304$ \\
    \hline
-   \multirow{2}{40pt}{\centering EDMD} & $0.289$ & $0.319$ & $0.375$ & $0.503$ \\
-   & $\pm 0.126$ & $\pm 0.134$ & $\pm 0.163$ & $\pm 0.163$ \\
+   \multirow{2}{38pt}{\centering EDMD} & $0.012$ & $0.289$ & $0.319$ & $0.375$ & $0.503$ \\
+   & $\pm 0.005$ & $\pm 0.126$ & $\pm 0.134$ & $\pm 0.163$ & $\pm 0.163$ \\
    \hline
-   \multirow{2}{40pt}{\centering GPR} & $0.276$ & $0.303$ & $0.361$ & $0.544$ \\
-   & $\pm 0.106$ & $\pm 0.110$ & $\pm 0.165$ & $\pm 0.157$ \\
+   \multirow{2}{38pt}{\centering GPR} & $0.018$ & $0.276$ & $0.303$ & $0.361$ & $0.544$ \\
+   & $\pm 0.013$ & $\pm 0.106$ & $\pm 0.110$ & $\pm 0.165$ & $\pm 0.157$ \\
    \hline

### 2021-01-06 Scripts/EvalPerfBoundEffRed/CompPerfAHM.m

@@ -46,15 +46,30 @@ nbXP = 1000 ;
 
 for ind = 1:nbXP
     noise = sigman*randn(N+2*L,1) ;
-    x = x0.' + noise((L+1):(N+L)) ; % signal to be extended
+    xx = x0.' + noise((L+1):(N+L)) ; % ground truth
+    
+    x = xx((L+1):(N+L)) ; % signal to be extended
 
+    % SigExt
     method.name = 'SigExt' ;
     tic;
     xxLSE = SigExtension(x,fs,HOP,extK,extM,extSEC,method).' ; % Forecasted signal via SigExt
     LSEtime(ind) = toc;
-    MeanLSE(ind,:) = xxLSE((N+L+1):end) - xx0((N+L+1):end) ;
-    VarLSE(ind,:) = ( xxLSE((N+L+1):end) - xx0((N+L+1):end) ).^2 ;
+    MeanLSE(ind,:) = xxLSE((N+L+1):end) - xx((N+L+1):end) ;
+    VarLSE(ind,:) = ( xxLSE((N+L+1):end) - xx((N+L+1):end) ).^2 ;
     
+
+    [~, SSTxx, ~] = sqSTFT_RT(xxLSE, basicTF.fmin, basicTF.fmax, basicTF.df, basicTF.hop, basicTF.win, LL, 1, 10, 0, 0) ;
+    [~, ~, conceFTxx, ~] = ConceFT_sqSTFT_RT(xxLSE, basicTF.fmin, basicTF.fmax, basicTF.df, basicTF.hop, basicTF.win, LL, 1, 10, 1, 0, 0) ;
+    [STFTxx, ~] = STFT_RT(xxLSE, basicTF.fmin, basicTF.fmax, basicTF.df, basicTF.hop, basicTF.win, LL, 1, 10) ;
+    [~, RSxx, ~, ~] = ConceFT_rsSTFT_RT(xxLSE, basicTF.fmin, basicTF.fmax, basicTF.df, basicTF.hop, basicTF.win, LL, 1, 10, 1) ;
+    
+    OTD.SST.LSE(ind) = slicedOT(SSTxx, SSTxx0) ;
+    OTD.conceFT.LSE(ind) = slicedOT(conceFTxx, conceFTxx0) ;
+    OTD.STFT.LSE(ind) = slicedOT(STFTxx, STFTxx0) ;
[…]

### 2021-01-06 Scripts/EvalPerfSigExt/InfluenceNoiseSine.m

@@ -1,4 +1,4 @@
-%% Evaluation of Gaussian white noise on SigExt's performance on a sine wave
+%% Evaluation on a sine wave of the Gaussian white noise influence on SigExt's performance
 % Author: Adrien MEYNARD
 % Email: adrien.meynard@duke.edu
 

### 2021-01-06 Scripts/EvalPerfSigExt/InfluenceSizeDatasetSine.m

@@ -1,4 +1,4 @@
-%% Evaluation the "training dataset" on SigExt's performance on a sine wave
+%% Evaluation on a sine wave of the "training dataset" influence on SigExt's performance
 % Author: Adrien MEYNARD
 % Email: adrien.meynard@duke.edu
 

### 2021-01-06 Scripts/EvalPerfSigExt/InfluenceSubsigLengthAHM.m

@@ -43,19 +43,23 @@ sigman = 1e-2 ;
 
 %% Forecasting
 method.name = 'SigExt' ;
-nbXP = 100 ;
+nbXP = 1000 ;
 
-extMval = round( logspace(log10(0.05),log10(3),50)*L ) ; % dimension of embedding / signals length
+extMval = [100 750 1500 3000]; % dimension of embedding / signals length
 
 for indXP = 1:nbXP
     noise = sigman*randn(N+L,1) ;
-    x = x0.' + noise(1:N) ; % signal to be extended
+    xx = x0.' + noise ;
+    
+    x = xx(1:N) ; % signal to be extended
 
     indM = 1 ;
     for extM = extMval
-        xxSigExt = forecasting(x,L,HOP,extK,extM,method).' ; % Forecasted signal via SigExt
-        MeanSigExt(indXP,indM) = mean( xxSigExt - xx0((N+1):end) ) ;
-        VarSigExt(indXP,indM) = mean( ( xxSigExt - xx0((N+1):end) ).^2 );
+        tic;
+        xxSigExt = forecasting(x,L,HOP,extK,extM,method).' ; % Forecasted signal via SigExt  
+        CPUtimeXP(indXP,indM) = toc;
+        MeanSigExt(indXP,indM) = mean( xxSigExt - xx((N+1):end) ) ;
+        VarSigExt(indXP,indM) = mean( ( xxSigExt - xx((N+1):end) ).^2 );
         
         indM = indM+1 ;
     end

### 2021-01-06 Scripts/EvalPerfSigExt/ResultsInfluenceSubSigAHM.m

@@ -2,19 +2,39 @@
 % Author: Adrien MEYNARD
 % Email: adrien.meynard@duke.edu
 
+clear all; close all; clc;
+
 load('../../Results/PerfSubsigLengthAHM') ;
 
 BiasXP = abs(mean(MeanSigExt)) ;
-VarianceXP = mean(VarSigExt) ;
+MSEXP = mean(VarSigExt) ;
+CPUtimeXP = mean(CPUtimeXP) ;
 
-figure;
-plot(extMval,BiasXP,'linewidth',2) ;
-xlabel('Subsignals length $M$','interpreter','latex') ;
-ylabel('Experimental Bias $\gamma_{\mathrm{xp}}$','interpreter','latex') ;
-set(gca,'fontsize',24) ; grid on ;
+% figure;
+% plot(extMval,BiasXP,'linewidth',2) ; axis tight ;
+% xlabel('Subsignals length $M$','interpreter','latex') ;
+% ylabel('Experimental Bias $\mu_{\mathrm{xp}}$','interpreter','latex') ;
+% set(gca,'fontsize',24) ; grid on ;
+% 
+% figure;
+% plot(extMval,MSEXP,'linewidth',2) ; axis tight ;
+% xlabel('Subsignals length $M$','interpreter','latex') ;
+% ylabel('Experimental MSE','interpreter','latex') ;
+% set(gca,'fontsize',24) ; grid on ;
 
-figure;
-plot(extMval,VarianceXP,'linewidth',2) ;
-xlabel('Subsignals length $M$','interpreter','latex') ;
-ylabel('Experimental Variance $\gamma_{\mathrm{xp}}$','interpreter','latex') ;
-set(gca,'fontsize',24) ; grid on ;
+
+fprintf('============= Forecasting Performance =============\n')
+fprintf(' _________________________________________________\n')
+fprintf('|    Extension   |  Computing  |        MSE      |\n')
+fprintf('|     Method     | time (sec.) |   Mean  |   SD  |\n')
+fprintf('|----------------|-------------|---------|-------|\n')
+fprintf('|SigExt (M=%i)|     %.3f   |  %.3f  | %.3f |\n', extM, CPUtimeXP(indM) , mean(MSEXP(:,indM)), std(MSEXP(:,indM)) )
+
+
+indM = 1 ;
+for extM = exMval
[…]

### 2021-01-06 Paper/results.tex
new script

@@
 To evaluate the forecasting quality, {\sf SigExt} is applied to $1000$ realizations of the above-described signal.
-Here, we forecast signal extensions of $0.1$~s, \ie~$L=700$.
-Besides, we set $K=2625$.
-Table~\ref{tab:mse.sine} shows the resulting experimental MSE according to four  different selected values of the subsignals length $M$.
-
+We forecast signal extensions of $0.1$~s, \ie~$L=700$.
+Table~\ref{tab:mse.sine} shows the resulting mean and standard deviation (SD) of the experimental MSE, according to four different values of the subsignals length $M$.
+Note that choosing a too small value for $M$ does not provide enough information to forecast the signal satisfactorily.
+Furthermore, a value of $M$ that is too large makes the algorithm insensitive to local nonstationarities produced by frequency variations.
+The user must then make a compromise on the choice of $M$ in order to optimize the performance of the algorithm.
+A choice of $M$ such that the subsignals contain about a dozen oscillations is generally appropriate.
 }
@@
 The same set of experiments is run on the extension methods outlined in section~\ref{sse:methods}.
-For comparison, the mean and the standard deviation (SD) of the resulting MSEs are given in Table~\ref{tab:mse.sine}.
-They show that the naive extension we propose gives satisfying results, in particular in comparison with the point-symmetric extension.
+The results, displayed in Table~\ref{tab:mse.sine}, show that the naive extension we propose gives satisfying results, in particular in comparison with the point-symmetric extension.
 Besides, even though the other more sophisticated methods, like GPR, give MSE values that have a slightly smaller standard deviation, these methods are substantially limited by the computing time they require, which prevents them from being used to exploit real-time data.
 Thus, {\sf SigExt} is the extension method that optimizes the trade-off between forecasting quality and computing time.
+{\color{red} Besides, the last column of Table~\ref{tab:THO} gives the performance index $D$ of the boundary-free STFT associated with each extension method.
+This illustrates the ability of our algorithm to reduce boundary effects to a level comparable to what is possible with the GPR or TBATS extensions.
+The performance of {\sf BoundEffRed} on other TF representations is more thoroughly analyzed on real-life signals in the following.}
 
@@
 \centering
-\caption{AHM Signal: Performance of the Extension Methods}
-\begin{tabular}{|c||c|c|}
+\caption{AHM Signal: Performance of the Extension Methods and the Associated Boundary-Free STFT}
+\begin{tabular}{|c||c|c||c|}
   \hline
-   \multirow{2}{60pt}{\centering Extension method} & \multirow{2}{50pt}{\centering MSE\\ (mean $\pm$ SD)}  & \multirow{2}{80pt}{\centering Averaged computing time (seconds)} \\
-    &  & \\
-   \hhline{|=#=|=|}
-   {\sf SigExt ($M=100$)} & $1.028 \pm  0.404$ & $1.069$ \\
+   \multirow{2}{60pt}{\centering Extension method}  & \multirow{2}{30pt}{\centering CPU time (s)} & \multirow{2}{50pt}{\centering MSE\\ (mean $\pm$ SD)} & \multirow{2}{50pt}{\centering Index $D$ of the STFT}\\
+    &  & & \\
+   \hhline{|=#=|=#=|}
+   {\sf SigExt ($M=100$)} & $0.008$ & $1.133 \pm  0.077$ & \\
    \hline
-   {\sf SigExt ($M=750$)} & $1.028 \pm  0.404$ & $1.069$ \\
+   {\sf SigExt ($M=750$)} & $0.122$ & $0.479 \pm  0.166$ & \\
    \hline
-   {\sf SigExt ($M=1500$)} & $1.028 \pm  0.404$ & $1.069$ \\
+   {\sf SigExt ($M=1500$)} & $0.634$ & $0.907 \pm  0.377$ & \\
    \hline
-   {\sf SigExt ($M=3000$)} & $1.028 \pm  0.404$ & $1.069$ \\
+   {\sf SigExt ($M=3000$)} & $2.881$ & $5.365 \pm  3.919$ & \\
    \hline
-   Symmetric & $4.368 \pm  0.001$ & $0.001$ \\
+   Symmetric &  {\color{red}$<0.001$} & {\color{red}$4.368 \pm  0.001$} &\\
    \hline
-   EDMD & $1.156  \pm  0.003$ &$11.10$\\
+   EDMD & {\color{red}$11.10$} & {\color{red}$1.156  \pm  0.003$} & \\
    \hline
-   GPR & $1.025  \pm  0.403$ &$244.7$ \\
+   GPR & {\color{red}$244.7$} & {\color{red}$1.025  \pm  0.403$} &\\
    \hline
-   TBATS & $1.501 \pm  1.354$ & $633.6$ \\
+   TBATS  & {\color{red}$633.6$} & {\color{red}$1.501 \pm  1.354$} & \\
    \hline

### 2021-01-06 Scripts/EvalPerfSigExt/ResultsAHM.m

@@ -1,0 +1,48 @@
+%% Display the results of InfluenceSubsigLengthAHM.m and OtherMethAHM.m
+% Author: Adrien MEYNARD
+% Email: adrien.meynard@duke.edu
+
+clear all; close all; clc;
+
+
+%% Display the results of InfluenceSubsigLengthAHM.m
+
+load('../../Results/PerfSubsigLengthAHM') ;
+
+BiasXP = abs(mean(MeanSigExt)) ;
+MSEXP = VarSigExt ;
+CPUtimeSigExt = mean(CPUtimeXP) ;
+
+
+fprintf('============= Forecasting Performance =============\n')
+fprintf(' ________________________________________________\n')
+fprintf('|    Extension    |  Computing  |        MSE      |\n')
+fprintf('|     Method      | time (sec.) |   Mean  |   SD  |\n')
+fprintf('|-----------------|-------------|---------|-------|\n')
+
+indM = 1 ;
+for extM = extMval
+    fprintf('| SigExt (M=%4i) |     %.3f   |  %.3f  | %.3f |\n', extM, CPUtimeSigExt(indM) , mean(MSEXP(:,indM)), std(MSEXP(:,indM)) ) ;
+    indM = indM + 1 ;
+end
+indM = indM +1 ;
+
+%% Display the results of OtherMethAHM.m
+
+load('../../Results/PerfOtherMethodsAHM') ;
+
+CPUtimeSYM = mean(CPUtimeXP.SYM) ;
+MSESYM = MSE.SYM ;
+
+CPUtimeEDMD = mean(CPUtimeXP.EDMD) ;
+MSEEDMD = MSE.EDMD ;
+
+CPUtimeGPR = mean(CPUtimeXP.GPR) ;
+MSEGPR = MSE.GPR ;
+
+
+fprintf('|  Symmetrization |     %.3f   |  %.3f  | %.3f |\n', CPUtimeSYM , mean(MSESYM), std(MSESYM) ) ;
+fprintf('|       EDMD      |     %.3f   |  %.3f  | %.3f |\n', CPUtimeEDMD , mean(MSEEDMD), std(MSEEDMD) ) ;
+fprintf('|       GPR       |     %.3f   |  %.3f  | %.3f |\n', CPUtimeGPR , mean(MSEGPR), std(MSEGPR) ) ;
+
+fprintf('|_________________|_____________|_________|_______|\n\n')

### 2021-01-06 Scripts/EvalPerfSigExt/ResultsInfluenceSubSigAHM.m

@@ -1,39 +1,0 @@
-%% Display the results of InfluenceSubsigLengthAHM.m
-% Author: Adrien MEYNARD
-% Email: adrien.meynard@duke.edu
-
-clear all; close all; clc;
-
-load('../../Results/PerfSubsigLengthAHM') ;
-
-BiasXP = abs(mean(MeanSigExt)) ;
-MSEXP = VarSigExt ;
-CPUtimeXP = mean(CPUtimeXP) ;
-
-% figure;
-% plot(extMval,BiasXP,'linewidth',2) ; axis tight ;
-% xlabel('Subsignals length $M$','interpreter','latex') ;
-% ylabel('Experimental Bias $\mu_{\mathrm{xp}}$','interpreter','latex') ;
-% set(gca,'fontsize',24) ; grid on ;
-% 
-% figure;
-% plot(extMval,MSEXP,'linewidth',2) ; axis tight ;
-% xlabel('Subsignals length $M$','interpreter','latex') ;
-% ylabel('Experimental MSE','interpreter','latex') ;
-% set(gca,'fontsize',24) ; grid on ;
-
-
-fprintf('============= Forecasting Performance =============\n')
-fprintf(' ________________________________________________\n')
-fprintf('|    Extension    |  Computing  |        MSE      |\n')
-fprintf('|     Method      | time (sec.) |   Mean  |   SD  |\n')
-fprintf('|-----------------|-------------|---------|-------|\n')
-
-indM = 1 ;
-for extM = extMval
-    fprintf('| SigExt (M=%4i) |     %.3f   |  %.3f  | %.3f |\n', extM, CPUtimeXP(indM) , mean(MSEXP(:,indM)), std(MSEXP(:,indM)) ) ;
-    indM = indM + 1 ;
-end
-indM = indM +1 ;
-
-fprintf('|_________________|_____________|_________|_______|\n\n')

### 2021-01-10 Paper/BoundEffRed.tex
review

--- 2020-09-11 Notes/BoundEffRed.tex
@@
 There exist extension schemes based on physically relevant dynamical models, such as the Extended Dynamic Mode Decomposition~\cite{Williams15data} (EDMD) and the Gaussian process regression~\cite{Rasmussen06gaussian,Roberts13Gaussian} (GPR).
+% 
+{\color{red}
+In speech processing, dynamic mode predictors have also been proposed~\cite{Vargas11speech} to forecast signals falling into the so-called \textit{source--filter} model.
+This gradient-based technique rely on the shadowing approach proposed in~\cite{Grebogi90shadowing}.
+} 
+%
 There also exist extension schemes based on stochastic models, such as the Trigonometric, Box-Cox transformation, ARMA errors, Trend and Seasonal components (TBATS) algorithm~\cite{DeLivera11forecasting}, the dynamic linear models \cite{west2006bayesian}.
@@
 The numerical results also confirmed the robustness to noise of {\sf BoundEffRed}, as well as its ability to be applied to many TF representations.
-An additional application to ECG is included in the Supplementary Materials.
+An additional applications to ECG is included in the Supplementary Materials.
 

### 2021-01-10 Paper/algorithm.tex
review

@@
 \end{equation}
+%
+{\color{blue}
 From the Takens' embedding perspective~\cite{Takens81detecting}, $\bX$ can be viewed as a lag map.
 Under the manifold assumption of the intrinsic phase space that hosts the dynamics, $f$ describes the dynamics supported on the manifold space, and $\bA$ approximates the dynamics on the manifold.
+}
 
 \paragraph{Forecasting}
-The forecasting method consists in estimating the unknown matrix $\bA$.
-Indeed, let $\tilde\bA$ denotes the estimate of $\bA$.
-{\color{red}We then obtain the forecasting of $\bx_{k+\ell}$ by recursively applying the linear relation~\eqref{eq:dyn.model}:
+The forecasting method consists in estimating the unknown matrix $\bA$. Indeed, let $\tilde\bA$ denotes the estimate of $\bA$. 
+%
+{\color{red}
+We then obtain the forecasting of $\bx_{k+\ell}$ by recursively applying the linear relation~\eqref{eq:dyn.model}:
 \begin{equation*}
@@
 Let $\be_M$ denote the unit vector of length $M$ given by $\be_M = \begin{pmatrix} 0 & \cdots & 0 & 1\end{pmatrix}^T$.
-Consequently, given definition~\eqref{eq:xk}, the forecasting of the signal at time $\frac{N-1+\ell}{\fs}$ is provided by:
+Consequently, given definition~\eqref{eq:xk}, the forecasting of the signal at time $\frac{N-1+\ell}{\fs}$ is provided by
 \begin{equation}

### 2021-01-10 Paper/results.tex
review

@@
 Indeed, in the log-log plot displayed in Fig.~\ref{fig:res.size.sine}, as soon as $K>700$, each of the three curves representing the forecasting variance as a function of $K$ tends towards an asymptote of slope $-1$.
+%
+{\color{red}
+In practice, the asymptotic behavior of {\sf SigExt}, illustrated here on a synthetic signal, can be generalized to the study of real-life signals, as long as the duration of the recorded signal largely exceeds that of the desired extension, which allows the user to set $K\gg L$.
+}
 
@@
 Thus, {\sf SigExt} is the extension method that optimizes the trade-off between forecasting quality and computing time.
-{\color{red} Besides, the last column of Table~\ref{tab:THO} gives the performance index $D$ of the boundary-free STFT associated with each extension method.
+%
+{\color{red}
+Besides, the last column of Table~\ref{tab:mse.sine} gives the performance index $D$ of the boundary-free STFT associated with each extension method.
 This illustrates the ability of our algorithm to reduce boundary effects to a level comparable to what is possible with the GPR or TBATS extensions.
-The performance of {\sf BoundEffRed} on other TF representations is more thoroughly analyzed on real-life signals in the following.}
+The performance of {\sf BoundEffRed} on other TF representations is more thoroughly analyzed on real-life signals in the following.
+}
 
@@
    \hhline{|=#=|=#=|}
-   {\sf SigExt ($M=100$)} & $0.008$ & $1.133 \pm  0.077$ & \\
+   {\sf SigExt ($M=100$)} & $0.008$ & $1.133 \pm  0.077$ & {\color{blue}$x \pm  x$}\\
    \hline
-   {\sf SigExt ($M=750$)} & $0.122$ & $0.479 \pm  0.166$ & \\
+   {\sf SigExt ($M=750$)} & $0.122$ & $0.479 \pm  0.166$ & {\color{blue}$x \pm  x$}\\
    \hline
-   {\sf SigExt ($M=1500$)} & $0.634$ & $0.907 \pm  0.377$ & \\
+   {\sf SigExt ($M=1500$)} & $0.634$ & $0.907 \pm  0.377$ & {\color{blue}$x \pm  x$}\\
    \hline
-   {\sf SigExt ($M=3000$)} & $2.881$ & $5.365 \pm  3.919$ & \\
+   {\sf SigExt ($M=3000$)} & $2.881$ & $5.365 \pm  3.919$ & {\color{blue}$x \pm  x$}\\
    \hline
-   Symmetric &  {\color{red}$<0.001$} & {\color{red}$4.368 \pm  0.001$} &\\
+   Symmetric &  {\color{blue}$<0.001$} & {\color{blue}$x \pm  x$} & {\color{blue}$x \pm  x$}\\
    \hline
-   EDMD & {\color{red}$11.10$} & {\color{red}$1.156  \pm  0.003$} & \\
+   EDMD & {\color{blue}$x$} & {\color{blue}$x  \pm  x$} & {\color{blue}$x \pm  x$}\\
    \hline
-   GPR & {\color{red}$244.7$} & {\color{red}$1.025  \pm  0.403$} &\\
+   GPR & {\color{blue}$x$} & {\color{blue}$x  \pm  x$} & {\color{blue}$x \pm  x$}\\
    \hline
-   TBATS  & {\color{red}$633.6$} & {\color{red}$1.501 \pm  1.354$} & \\
+   TBATS  & {\color{blue}$x$} & {\color{blue}$x \pm  x$} & {\color{blue}$x \pm  x$}\\
    \hline
@@
 \subsection{Real Physiological Signals}
+\label{sse:physio.sig}
 

### 2021-01-10 Paper/theory.tex
review

--- 2020-09-11 Paper/theoryOLD.tex
@@
 \subsection{Signal Model}
-We model the deterministic part of the observed signal as a multicomponent harmonic signal; that is, a sum of sine waves:
+\label{sse:model.sine}
+We model the meaningful part of the observed signal as a deterministic multicomponent harmonic signal; that is, a sum of sine waves:
 \begin{equation}
@@
 \end{equation}
-where $\bz$ follows model~\eqref{eq:sum.sine}, $\bw$ is a Gaussian white noise, whose variance is normalized to one. Thus, $\sigma^2$ denotes the variance of the additive noise $\sigma\bw$.
+where $\bz$ follows model~\eqref{eq:sum.sine}, $\bw$ is a Gaussian white noise, whose variance is normalized to one, and $\sigma>0$.
+Clearly, $\sigma^2$ denotes the variance of the additive noise $\sigma\bw$.
 
@@
 \item If $n'\in I$:
-\begin{align}
-\nonumber
-\bgamma[n,n'] &= \sigma\EE\{\bw[n']\balpha^{(\ell)}\}\bz_K + \sigma^2\EE\{\bw[n']\balpha^{(\ell)}\bw_K\} \\
-&\hspace{15pt} -\bz[n]\bmu[n'] - \bmu[n]\bmu[n']\ .
+\begin{equation}
+\bgamma[n,n'] = \sigma\EE\{\bw[n']\balpha^{(\ell)}\}\bz_K + \sigma^2\EE\{\bw[n']\balpha^{(\ell)}\bw_K\}\ .
 \label{eq:cov00}
-\end{align}
+\end{equation}
 \item If $n'=N-1+\lambda\geq N$:
@@
 
-Expressions~\eqref{eq:bias0},~\eqref{eq:cov00}, and~\eqref{eq:cov01} show that these quantities depend on the behavior of the forecasting random vector $\balpha^{(\ell)}$.
-In Lemma~\ref{lm:error}, we specify the asymptotic behavior of the forecasting vector $\balpha^{(\ell)}$ when the dataset size $K$ is great.
-\begin{lemma}
-\label{lm:error}
-Let $\bx$ be a random vector defined by~\eqref{eq:model.noise}.
-Let $\balpha^{(\ell)}$ be the associated forecasting vector for the estimation of $\bx[N-1+\ell]$, given by~\eqref{eq:alpha.l} and obtained from the least square estimation~\eqref{eq:lse}.
-Let $\balpha_0^{(\ell)}$ be the last row of the matrix $\bA_0^\ell$, where $\bA_0$ is defined by:
-\begin{equation}
-\bA_0 = \left(\dfrac1K\bZ'\bZ^T+\sigma^2\bD\right)\left(\dfrac1K\bZ\bZ^T+\sigma^2\bI\right)^{\inv}\ ,
-\label{eq:lse.0}
-\end{equation}
-where $\bZ=\begin{pmatrix} \bz_0 & \cdots & \bz_{K-1}\end{pmatrix}$, $\bZ'=\begin{pmatrix} \bz_1 & \cdots & \bz_{K}\end{pmatrix}$, $\bz_k$ is the $k$-th subsignal extracted from $\bz$ in the same way as $\bx_k$ is defined from $\bx$ in~\eqref{eq:xk}, and $\bD\in\RR^{M\times M}$ is the Toeplitz matrix such that $\bD[m,m'] = \delta_{m+1,m'}$.
-
-Let $\bh^{(\ell)}$ be the error vector given by
-\begin{equation*}
-\bh^{(\ell)} = \balpha^{(\ell)} - \balpha_0^{(\ell)}\, .
-\end{equation*}
-Then, the random vector $\bh^{(\ell)}$ converges in law to a zero-mean Gaussian random vector when $K\to\infty$; that is,
-\begin{equation}
-\sqrt{K}\,\bh^{(\ell)}  \xrightarrow[K\to\infty]{\cD} \cN\left(\bzero,\bGamma^{(\ell,\ell)}\right)\ ,
-\end{equation}
-with $\bGamma^{(\ell,\ell)}:={\bF^{(\ell)}}^T\bGamma_0\bF^{(\ell)}$, where $\bGamma_0\in\RR^{M(M+1)\times M(M+1)}$ is a covariance matrix and $\bF^{(\ell)}\in\RR^{M(M+1)\times M}$ is a Jacobian matrix, and $\bGamma^{(\ell,\ell)}$ does not depend on $\sigma$.
-% such that:
-%{\color{red}
-%\begin{align*}
-%\bGamma_0[mM+n,pM+q] &= 2\sigma^2\sum_{j=1}^J\Omega_j^2\cos\left(2\pi\frac{f_j}{\fs}(m-n)\right)\\
-%& \hspace{26pt} \times\cos\left(2\pi\frac{f_j}{\fs}(p-q)\right) + o(\sigma^2) \\
-%\bF^{(\ell)}[] &= \ . 
-%\end{align*}
-%}
-\end{lemma}
-
-\begin{proof}
-See the Supplementary Material.
-The proof is based on the multivariate delta method (see paragraph 7.2 in~\cite{Alho05statistical}), which gives an asymptotic approximation of a random vector by a Gaussian random vector.
-\end{proof}
-
-Consequently, the covariance between $\sqrt{K}\bh^{(\ell)}$ and $\sqrt{K}\bh^{(\lambda)}$ remains bounded, \ie:
-\begin{equation*}
-K\,\EE\left\{{\bh^{(\ell)}}^T\bh^{(\lambda)}\right\} \xrightarrow[K\to\infty]{} \bGamma^{(\ell,\lambda)} = {\bF^{(\ell)}}^T\bGamma_0\bF^{(\lambda)}\ .
-\end{equation*}
-
+In Theorem~\ref{th:error}, we specify the asymptotic behavior of the forecasting bias and covariance when the dataset size $K$ is great.
 \begin{theorem}
@@
 \begin{equation}
-\frac{1}{\sigma}\bmu[n] = o(1)
+\left|\bmu[n]\right| \leq a^{(n)}_0\sigma^2 + \dfrac1K \left( \dfrac{a^{(n)}_1}{\sigma^2} + a^{(n)}_2 \right) + o\left( \dfrac1K \right)
 \label{eq:mean.error}
 \end{equation}
-when $K\to\infty$.
+as $K\to\infty$, where $a^{(n)}_0$, $a^{(n)}_1$, and $a^{(n)}_2$ are positive quantities, independent of $K$ and $\sigma$.
 Its second-order moment $\bgamma[n,n']$ satisfies the following approximation equations:
@@
 \begin{equation}
-\bgamma[n,n']\to \sigma^2\balpha_0^{(n-N-1)}[n'-(N-M)]\1_{(n'\geq N-M)}
+\left|\bgamma[n,n']\right|\leq b^{(n,n')}_0\sigma^2 + \!\dfrac1K\! \left( b^{(n,n')}_1\! + b^{(n,n')}_2\sigma^2 \! \right) + o\left( \dfrac1K \right)
 \label{eq:cov.error.1}
 \end{equation}
-when $K\to\infty$;
+as $K\to\infty$, where $b^{(n,n')}_0$, $b^{(n,n')}_1$, and $b^{(n,n')}_2$ are positive quantities, independent of $K$ and $\sigma$;
 \item if $n'\geq N$:
-\begin{equation}
-\hspace{-15pt}\bgamma[n,n'] = \dfrac1K\bz_K^T\bGamma^{(\ell,\lambda)}\bz_K + \dfrac{\sigma^2}{K}\Tr\left(\bGamma^{(\ell,\lambda)}\right) + \sigma^2\left\langle\balpha_0^{(\ell)}\!, \balpha_0^{(\lambda)}\right\rangle ,
+\begin{align}
+\nonumber
+\left|\bgamma[n,n']\right| &\leq c^{(n,n')}_0\sigma^2 \\
+&\hspace{10pt} + \!\dfrac1K\! \left(\! \dfrac{c^{(n,n')}_1}{\sigma^2} + c^{(n,n')}_2 + c^{(n,n')}_3\sigma^2 \right) + o\!\left( \!\dfrac1K \!\right)
 \label{eq:cov.error.2}
-\end{equation}
-when $K\to\infty$, where $\ell\!=n-N+1$ and $\lambda\!=n'-N+1$.
+\end{align}
+as $K\to\infty$, where $c^{(n,n')}_0$, $c^{(n,n')}_1$, $c^{(n,n')}_2$ and $c^{(n,n')}_3$ are positive quantities, independent of $K$ and $\sigma$.
 \end{enumerate}
@@
 \begin{proof}
-See the Supplementary Material.
-The proof is mainly based on the results provided by Lemma~\ref{lm:error}, combined with the Isserlis' theorem~\cite{Isserlis16formula}, which provides a formula for the computation of higher-order moments of Gaussian random variables.
+See the Supplementary Materials.
+The proof is mainly based on Taylor expansions of the forecasting error.
+Nonoptimal values of $a^{(n)}_0$, $a^{(n)}_1$,\dots, $b^{(n,n')}_0$,\dots, $c^{(n,n')}_3$ are also provided in the proof.
 \end{proof}
+%The Isserlis' theorem~\cite{Isserlis16formula}, which provides a formula for the computation of higher-order moments of Gaussian random variables.
 
-Ideally, the forecasting error would behave like the measurement noise $\sigma\bw$, \ie~a zero-mean noise whose variance is of the order of $\sigma^2$.
-Theorem~\ref{th:error} shows that the forecasting error is asymptotically unbiased.
-Concerning the covariance of the forecasting error, although equations~\eqref{eq:cov.error.1} and~\eqref{eq:cov.error.2} are not easily readable, one can evaluate the dependence of the variance in function of the tuning parameters, that are adjusted by the user.
-Let us focus on the forecasting error variance $\gamma[n,n]$ when $n\geq N$.
-First, as expected, the variance increases linearly with the noise variance $\sigma^2$.
-Second, it asymptotically depends linearly on the ratio $\frac{1}{K}$.
+The forecasting bias and covariance asymptotically depend linearly on the ratio $\frac{1}{K}$.
 This shows the need to use a sufficiently large dataset to obtain an accurate forecast.
-Third, the dependency on the subsignals lengths $M$ and the forecasting index $\ell=n-N+1$ is hidden in the expression of the covariance matrix $\bGamma^{(\ell,\ell)}$.
-We discuss this dependency in more detail in section~\ref{ssse:res.sine}.
+Ideally when $K\to\infty$, the forecasting error would behave like the measurement noise $\sigma\bw$, \ie~a zero-mean white noise whose variance is of the order of $\sigma^2$.
+However, Theorem~\ref{th:error} shows that the estimator may remain asymptotically biased (the first term in the bound of equation~\eqref{eq:mean.error} being independent of $K$).
+Concerning the covariance of the forecasting error, the expected asymptotic behavior is verified.
+Indeed, when $K \to\infty$, the variance of the forecasting estimator increases at most linearly with the noise variance $\sigma^2$, since is bounded by $b_0^{(n,n')}\sigma^2$.
 
-\subsection{Adaptive Harmonic Model}\label{RemarkAHM}
+Assume now that $K$ is sufficiently great and fixed.
+The noise influences the quality of the forecasting estimator in two ways.
+On the one hand, when the noise variance $\sigma^2$ is great, the bias and the variance of the estimator potentially increase linearly with $\sigma^2$.
+This reflects the classical behavior of an estimator of a signal degraded by additive noise.
+On the other hand, due to terms in $1/\sigma^2$, when the noise variance $\sigma^2$ is small, the bias and the variance of the forecasting estimator are no longer controlled via equations~\eqref{eq:mean.error} and~\eqref{eq:cov.error.2}.
+This illustrates the necessary presence of noise to obtain an accurate prediction.
+Indeed, if $\sigma$ is too small, the matrix $\bX\bX^T$ is ill-conditioned and its inversion in~\eqref{eq:lse} is sensitive.
+The forecasting is then of poor quality.
+Noise therefore appears as a regularization term to ensure the invertibility of $\bX\bX^T$.
+This seemingly counterintuitive fact indicates the challenge we encounter to do forecasting with the dynamic model~\eqref{eq:generic.model}, the degeneracy.
+Note that from the lag map perspective, $\bX$ and $\bY$ consist of points located on a low dimensional manifold.
+Thus, locally the ranks of $\bX$ and $\bY$ are low, which leads to the degeneracy.
+However, when noise exists, it naturally fills in deficient ranks, and stabilizes the algorithm.
+
+The dependencies of the forecasting quality on the subsignals lengths $M$ and the forecasting index $\ell=n-N+1$ are hidden in the expression of the parameters $a^{(n)}_0$, $a^{(n)}_1$,\dots, $b^{(n,n')}_0$,\dots, $c^{(n,n')}_3$.
+The numerical results, discussed in section~\ref{ssse:res.sine}, allow us to better evaluate these dependencies.
+
+\subsection{Adaptive Harmonic Model}
+\label{RemarkAHM}
 
@@
 \end{equation}
-where $a_j(t)$ and $\phi'_j(t)$ describe how large and fast the signal oscillate at time $t$. 
+where $a_j(t)$ and $\phi'_j(t)$ describe how large and fast the signal oscillates at time $t$. 
 Clearly, \eqref{eq:sum.sine} is a special case satisfying the AM-FM model when the $a_j$ and $\phi'_j$ are both constants. 
@@
 %
-For the forecasting purpose, it is thus clear that when $K$ is not too large, the signal can be well approximated by \eqref{eq:sum.sine}.
-Hence, Theorem \ref{th:error} can still be applied to approximate the error.
+For the forecasting purpose, it is thus clear that when $K$ is not too large, the signal can be well approximated by~\eqref{eq:sum.sine}.
+Hence, Theorem~\ref{th:error} can still be applied to approximate the error.
 However, how to determine the optimal $K$ based on the signal depends on the application and is out of the scope of this paper.
@@
 
+{\color{red}
+The models of sections~\ref{sse:model.sine} and~\ref{RemarkAHM} consider the meaningful part of the signal to be a deterministic component.
+These models are purposely adapted to signals showing local line spectra.
+The physiological signals we are interested in, such as respiratory or cardiac signals, have this characteristic (see section~\ref{sse:physio.sig}).
+Signals with wider local spectra, such as electroencephalography signals, do not fall into this category, and are more faithfully modeled as random signals.
+Thus, the theoretical justifications proposed above are no longer applicable to guarantee the forecasting quality of {\sf SigExt}.
+}
+
 \subsection{Performance of the Boundary Effects Reduction}
@@
 While we do not provide a generic proof of the boundary effect reduction on {\em any} TF analysis tools, we discuss a particular case of SST.
-Since SST is designed to analyze signals satisfying the AHM, as is discussed in Remark \ref{RemarkAHM}, Theorem~\ref{th:error} ensures that the forecasting error $\bepsilon$ defined in~\eqref{eq:forecasting.error} is controlled and bounded in terms of mean and covariance.
-Recall Theorem 3 in~\cite{Chen14nonparametric}, which states that when the additive noise is stationary and maybe modulated by a smooth function with a small fluctuation, the SST of the observed signal remains close to the SST of the clean signal, throughout the TF plane.
+Since SST is designed to analyze signals satisfying the AHM, as is discussed in section~\ref{RemarkAHM}, Theorem~\ref{th:error} ensures that the forecasting error $\bepsilon$ defined in~\eqref{eq:forecasting.error} is controlled and bounded in terms of mean and covariance.
+Recall Theorem 3 in~\cite{Chen14nonparametric}, which states that when the additive noise is stationary and may be modulated by a smooth function with a small fluctuation, the SST of the observed signal remains close to the SST of the clean signal, throughout the TF plane.
 %
@@
 %
-In our case, while the noise of the historical data is stationary, the forecasting error depends on the historical noise, and hence the overall ``noise'' is non-stationary.
+In our case, while the noise of the historical data is stationary, the forecasting error depends on the historical noise, and hence the overall ``noise'' is nonstationary.
 However, the dependence only appears in the extended part of the signal.
 We claim that the proof of Theorem 3 in~\cite{Chen14nonparametric} can be generalized to explain the robustness of SST to the noise, and a systematic proof will be reported in our future theoretical work.
-While we do not provide a theoretical justification, we verify experimentally that Algorithm~\ref{alg:boundary} is efficient for a large number of representations in the following section~\ref{se:results}.
+We verify experimentally that Algorithm~\ref{alg:boundary} is efficient for a large number of representations in the following section~\ref{se:results}.
 %
 Therefore, in our case, this means that the boundary effect is strongly reduced since the impact of the forecasting error does SST is limited.
-An immediate application is that the instantaneous frequencies can now be well estimated continuously up to the edges of the TF plane, and a real-time acquisition of the instantaneous frequency and amplitude information is possible.
+An immediate application is that the instantaneous frequencies can now be well estimated continuously up to the edges of the TF plane, and real-time acquisition of the instantaneous frequency and amplitude information is possible.

### 2021-01-10 Paper/response/response.tex
typos

\documentclass[11pt,DIV=16]{scrartcl}

\usepackage[shortlabels]{enumitem}
\usepackage{xcolor}
\begin{document}

\section*{Reviewer 1}
Thanks for your careful evaluation and your comments and suggestions. Please find below answers to your remarks.
\begin{enumerate}[1)]
\item
\textit{page 3, equation (7). The authors should provide the definition of the matrix $\tilde{A}^\ell$.}\\
\textbf{The definition is now provided by the unnumbered equation between equation (5) and (6).}
\item
\textit{page 3, second column. Based on the definitions of $\mathbf{X}$ and $\mathbf{Y}$ given in the first column as well as equations (4) and (5), it is unclear how equation (6) could allow to forecast the signal at time $\frac{N-1+\ell}{f_s}$. Please give more details.}\\
\textbf{The paragraph has been rephrased, in order to make the forecasting procedure clearer. In particular, we detail how we obtain the estimation of $\mathbf{x}_{k+\ell}$ by recursively applying the linear relation (4). The estimation of the signal at time $\frac{N-1+\ell}{f_s}$ is then directly deduced from the last element of this vector.}
\item
\textit{page 7, second column. The authors wrote that the choice of the $M$ parameter is especially crucial when the the deterministic component of the signal is no longer stationary, like in the AHM model.
However, when they present an application of the proposed method to a signal satisfying the AHM model, neither they discuss how they select the $M$ value, nor they provide the analysis of the dependency of the performance of the proposed algorithm on the M value. Please include in section IV-C2 the evolution of the performance of the proposed method as a function of M and an explanation on how to tune its value in applications.}\\
\textbf{We have added the new paragraph IV-C2a where we discuss the influence of the parameter $M$ on the forecasting performance of an adaptive harmonic signal. In particular we mention that that choosing a too small value for $M$ does not provide enough information to forecast the signal satisfactorily, while a value of $M$ that is too large makes the algorithm insensitive to local nonstationarities produced by frequency variations. The user must then make a compromise on the choice of $M$ in order to optimize the performance of the algorithm. A choice of $M$ such that the subsignals contain about a
dozen oscillations is generally appropriate.}
\item
\textit{page 8, Table I. The authors present the performance of the extension method when applied to a toy AHM model. But they do not provide an insight of the impact, in this specific case, on the associated "boundary-free" TFR. Please include in this table the "performance index $D$" estimation proposed in section "Evaluation Metrics".}\\
\textbf{Table III now includes a new column containing the performance index $D$ of the boundary-free STFT associated with each of the extension method that we tested. For the sake of clarity, the performance of the boundary-free versions of the other TF representations (SST, RS and ConceFT) are not provided in this section. Their study is referred to the analysis of the real physiological signals (section IV-D).}
\item
\textit{page 10, Table III. Please provide the performance results associated with the first metric you proposed to use, the MSE. }\\
\textbf{This results have been added to Table III.}
\item
\textit{All real life signals proposed in the work are nonstationary signals which contain one single nonstationary component plus noise. In the toy example of a  AHM signal, section IV-C2, you presented the case of a signal containing two nonstationary components.
Please add the example of a real life signal containing two or more simple nonstationary components plus noise.}\\
\textbf{\color{blue} Idea?}
\end{enumerate}

\section*{Reviewer 2}
Thanks for your careful evaluation of the paper and your comments, which we tried to take into account as much as we could. You will find below answers to your remarks, suggestions and questions.
\begin{enumerate}[1)]
\item
\textit{The approach is similar to the shadowing approach suggested by C. Gregobi in the Nonlinear Dynamics community, (see his Physics Review Letters in 1990 Shadowing of physical trajectories in chaotic dynamics) and subsequent papers on shadowing in multiple domains e.g. by Vargas et al in IEEE Transactions on Audio, Speech and Language in 2011.}\\
\textbf{We have mentioned this approach in the introduction of our article. We precise that dynamic mode predictors have been proposed in speech processing by Vargas and McLaughlin, to forecast signals  falling into the so-called source--filter model. This gradient-based technique rely on the shadowing approach proposed in Grebogi et al. Nevertheless,  since the source--filter is not adapted to model physiological signals on which we focus, we do not implement this method in our article.}
\item
\textit{The signal is assumed to be deterministic with additive noise, in my experience oscillatory signals of interest in many applications are a combination of deterministic and stochastic components. How robust is the scheme?}\\
\textbf{A new paragraph has been added to precise the limitations of the theoretical justifications on the performance of \textsf{SigExt}. Indeed, the models of sections III-A and III-C consider the meaningful part of the signal to be a deterministic component. These models are purposely adapted to signals showing local line spectra. The physiological signals we are interested in, such as respiratory or cardiac signals, have this characteristic (see section IV-D). Signals with wider local spectra, such as electroencephalography signals, do not fall into this category, and are more faithfully modeled as random signals. Thus, the theoretical justifications proposed above are no longer applicable to guarantee the forecasting quality of \textsf{SigExt}. A more robust approach to the study of signals following this model is envisaged in our future work.}
\item
\textit{The reference to Takens is slightly confusing could this be explained more clearly. You are not referring to nonlinear dynamic signals so I am not sure it is relevant.} \\
\textbf{\color{blue} Hau-Tieng?}
\item
\textit{The paper discusses asymptotic behaviour of the forecasting algorithm but the time series lengths in practice are quite short so what impact does that have on performance.} \\
\textbf{In practice, the asymptotic behavior of \textsf{SigExt}, illustrated on a synthetic signal in the paper, can be generalized to the study of real-life signals, as long as the duration of the recorded signal largely exceeds that of the desired extension, which allows the user to set $K \gg L$. In other words, to ensure the relevance of the extension produced by \textsf{SigExt}, the user should be able to use the previously recorded signal for at least three times the duration of the desired extension, approximately. For example, in section IV-D2, the analysis of the PPG signal requires 7-second-long extensions to effectively reduce boundary effects in the TF domain. About twenty seconds of recording are thus necessary to ensure the fidelity of our algorithm.}
\end{enumerate}

\end{document}

### 2021-01-13 Paper/results.tex
paper

@@
 A choice of $M$ such that the subsignals contain about a dozen oscillations is generally appropriate.
-}
 
@@
 The same set of experiments is run on the extension methods outlined in section~\ref{sse:methods}.
-The results, displayed in Table~\ref{tab:mse.sine}, show that the naive extension we propose gives satisfying results, in particular in comparison with the point-symmetric extension.
-Besides, even though the other more sophisticated methods, like GPR, give MSE values that have a slightly smaller standard deviation, these methods are substantially limited by the computing time they require, which prevents them from being used to exploit real-time data.
+The results, displayed in Table~\ref{tab:mse.sine}, show that the naive extension we propose gives satisfying results.
+Indeed, when the subsignal length $M$ is optimally chosen, {\sf SigExt} performance is of the same order as that of ore sophisticated methods, like GPR or TBATS.
+Besides, even though these methods could be slightly more robust to perturbations, they are substantially limited by the computing time they require, which prevents them from being used to exploit real-time data.
 Thus, {\sf SigExt} is the extension method that optimizes the trade-off between forecasting quality and computing time.
-%
-{\color{red}
 Besides, the last column of Table~\ref{tab:mse.sine} gives the performance index $D$ of the boundary-free STFT associated with each extension method.
@@
    \hline
-   Symmetric &  {\color{blue}$<0.001$} & {\color{blue}$x \pm  x$} & {\color{blue}$x \pm  x$}\\
+   Symmetric &  $<0.001$ & $5.452 \pm  0.002$ & {\color{blue}$x \pm  x$}\\
    \hline
-   EDMD & {\color{blue}$x$} & {\color{blue}$x  \pm  x$} & {\color{blue}$x \pm  x$}\\
+   EDMD & $1.587$ & $0.983  \pm  0.004$ & {\color{blue}$x \pm  x$}\\
    \hline
-   GPR & {\color{blue}$x$} & {\color{blue}$x  \pm  x$} & {\color{blue}$x \pm  x$}\\
+   GPR & $515.787$ & $0.479  \pm  0.166$ & {\color{blue}$x \pm  x$}\\
    \hline

### 2021-01-23 Scripts/EvalPerfBoundEffRed/ResultsCompPerfAHM.m
paper

@@ -16,32 +16,39 @@ MSE.GPR = VarianceXP.GPR ;
 % MSE.TBATS = dataTBATS(:,1).^2 + dataTBATS(:,2); %VarianceXP.TBATS ;
 
 fprintf('===================== Forecasting Performance ===================\n')
-fprintf(' _________________________________________________________________\n')
-fprintf('|    Extension   |  Computing  |                 MSE              |\n')
-fprintf('|     Method     | time (sec.) |   Mean  |   SD  |t-test to SigExt|\n')
-fprintf('|----------------|-------------|---------|-------|----------------|\n')
+fprintf(' _______________________________________________________________\n')
+fprintf('|    Extension   |  Computing  |              MSE               |\n')
+fprintf('|     Method     | time (sec.) |  Mean |   SD  |t-test to SigExt|\n')
+fprintf('|----------------|-------------|-------|-------|----------------|\n')
 
 indM = 1 ;
 for extM = extMval
-    fprintf('|SigExt (M=%4i) |   %7.3f   |  %5.3f  | %5.3f |                |\n', extM, CPUtimeXP.LSE(indM) , mean(MSE.LSE(:,indM)), std(MSE.LSE(:,indM)) ) ;
+    fprintf('|SigExt (M=%4i) |   %7.3f   | %5.3f | %5.3f |                |\n', extM, CPUtimeXP.LSE(indM) , mean(MSE.LSE(:,indM)), std(MSE.LSE(:,indM)) ) ;
     indM = indM + 1 ;
 end
 indM = indM +1 ;
 
-fprintf('| Symmetrization |   %7.3f   |  %5.3f  | %5.3f |      %u       |\n', CPUtimeXP.SYM , mean(MSE.SYM), std(MSE.SYM), ttest(MSE.SYM,MSE.LSE(:,2)))
-fprintf('|      EDMD      |   %7.3f   |  %5.3f  | %5.3f |      %u       |\n', CPUtimeXP.EDMD , mean(MSE.EDMD), std(MSE.EDMD), ttest(MSE.EDMD,MSE.LSE(:,2)))
[…]

### 2021-01-23 Scripts/EvalPerfSigExt/ResultsAHM.m

@@ -22,7 +22,7 @@ fprintf('|-----------------|-------------|---------|-------|\n')
 
 indM = 1 ;
 for extM = extMval
-    fprintf('| SigExt (M=%4i) |     %.3f   |  %.3f  | %.3f |\n', extM, CPUtimeSigExt(indM) , mean(MSEXP(:,indM)), std(MSEXP(:,indM)) ) ;
+    fprintf('| SigExt (M=%4i) |   %7.3f   |  %5.3f  | %5.3f |\n', extM, CPUtimeSigExt(indM) , mean(MSEXP(:,indM)), std(MSEXP(:,indM)) ) ;
     indM = indM + 1 ;
 end
 indM = indM +1 ;
@@ -41,8 +41,8 @@ CPUtimeGPR = mean(CPUtimeXP.GPR) ;
 MSEGPR = MSE.GPR ;
 
 
-fprintf('|  Symmetrization |     %.3f   |  %.3f  | %.3f |\n', CPUtimeSYM , mean(MSESYM), std(MSESYM) ) ;
-fprintf('|       EDMD      |     %.3f   |  %.3f  | %.3f |\n', CPUtimeEDMD , mean(MSEEDMD), std(MSEEDMD) ) ;
-fprintf('|       GPR       |   %.3f   |  %.3f  | %.3f |\n', CPUtimeGPR , mean(MSEGPR), std(MSEGPR) ) ;
+fprintf('|  Symmetrization |   %7.3f   |  %5.3f  | %5.3f |\n', CPUtimeSYM , mean(MSESYM), std(MSESYM) ) ;
+fprintf('|       EDMD      |   %7.3f   |  %5.3f  | %5.3f |\n', CPUtimeEDMD , mean(MSEEDMD), std(MSEEDMD) ) ;
+fprintf('|       GPR       |   %7.3f   |  %5.3f  | %5.3f |\n', CPUtimeGPR , mean(MSEGPR), std(MSEGPR) ) ;
 
 fprintf('|_________________|_____________|_________|_______|\n\n')

### 2021-01-24 Paper/Supplementary material/SupMat.tex
review
paper

--- 2020-09-11 Notes/Supplementary material/SupMat.tex
@@
 
-\setcounter{equation}{26}
-\setcounter{figure}{7}
+\setcounter{equation}{25}
+\setcounter{figure}{8}
 \setcounter{table}{3}
@@
 
+{\color{red}
+\section{Application to a Multicomponent Cardiac Signal}
+We consider a cardiac signal, namely an arterial blood pressure recording, sampled at $1000$~Hz.
+In addition to the blood pressure measurement, this signal contains a slow varying component, which is the respiratory component.
+Fig.~\ref{fig:abp} displays the ordinary and boundary-free SSTs of a 150-second-long extract of this signal, obtained via {\sf BoundEffRed}.
+This TF analysis brings out both components---the cardiac component, the most energetic, is located around the fundamental frequency $1$~Hz and its multiples, while the respiratory component, less visible, is located around $0.2$~Hz.
+The performance index of this representation takes the value $D=0.XXX$.
+Note that, here, boundary effects are only reduced on the right-side boundary of the TF domain.
+This means that {\sf BoundEffRed} has reduced the right-side boundary effects by about $XX\%$ with respect to the ordinary SST.
+This shows the ability of our algorithm to work on signals containing several nonstationary components.
+
+\begin{figure}
+\centering
+%\includegraphics[width=.48\textwidth]{ABPsst.eps}
+%\includegraphics[width=.48\textwidth]{ABPberSST.eps}
+\caption{Ordinary SST (left) and boundary-free SST (right) of an arterial blood pressure recording}
+\label{fig:abp}
+\end{figure} 
+}
+
 \bibliographystyle{IEEEtran}

### 2021-01-24 Paper/results.tex
paper

@@
    \hhline{|=#=|=#=|}
-   {\sf SigExt ($M=100$)} & $0.008$ & $1.133 \pm  0.077$ & {\color{blue}$x \pm  x$}\\
+   {\sf SigExt ($M=100$)} & $0.008$ & $1.133 \pm  0.077$ & $0.0091 \pm  0.0019$\\
    \hline
-   {\sf SigExt ($M=750$)} & $0.122$ & $0.479 \pm  0.166$ & {\color{blue}$x \pm  x$}\\
+   {\sf SigExt ($M=750$)} & $0.122$ & $0.479 \pm  0.166$ & $0.0056 \pm  0.0019$\\
    \hline
-   {\sf SigExt ($M=1500$)} & $0.634$ & $0.907 \pm  0.377$ & {\color{blue}$x \pm  x$}\\
+   {\sf SigExt ($M\!=\!1500$)} & $0.634$ & $0.907 \pm  0.377$ & $0.0065 \pm  0.0023$\\
    \hline
-   {\sf SigExt ($M=3000$)} & $2.881$ & $5.365 \pm  3.919$ & {\color{blue}$x \pm  x$}\\
+   {\sf SigExt ($M\!=\!3000$)} & $2.881$ & $5.365 \pm  3.919$ & $0.0117 \pm  0.0038$\\
    \hline
-   Symmetric &  $<0.001$ & $5.452 \pm  0.002$ & {\color{blue}$x \pm  x$}\\
+   Symmetric &  $<0.001$ & $5.452 \pm  0.002$ & $1.8209 \pm  0.0025$\\
    \hline
-   EDMD & $1.587$ & $0.983  \pm  0.004$ & {\color{blue}$x \pm  x$}\\
+   EDMD & $1.587$ & $0.983  \pm  0.004$ & $0.1145 \pm  0.0018$\\
    \hline
-   GPR & $515.787$ & $0.479  \pm  0.166$ & {\color{blue}$x \pm  x$}\\
+   GPR & $515.787$ & $0.479  \pm  0.166$ & $0.0119 \pm  0.0037$\\
    \hline
@@
 
+{\color{red}
+Then, we consider a second respiratory signal that contains not only the respiratory cycle, but also the cardiogenic artifact which reflects the hemodynamic component.
+Fig.~\ref{fig:resp.2} displays the boundary-free SST of a 180-second-long extract of this signal, sample at $125$~Hz, obtained via {\sf BoundEffRed}.
+This TF analysis brings out both components---the respiratory component, the most energetic, is located around the fundamental frequency $0.25$~Hz and its multiples, while the cardiac component, less visible, is located around $1.7$~Hz.
+The performance index of this representation takes the value $D=0.314$.
+Note that, here, boundary effects are only reduced on the right-side boundary of the TF domain.
+This means that {\sf BoundEffRed} has reduced the right-side boundary effects by about $69\%$ with respect to the ordinary SST.
+This shows the ability of our algorithm to work on signals containing several nonstationary components.
+
+\begin{figure}
+\centering
+\includegraphics[width=.48\textwidth]{Resp2comp.eps}
+\caption{Boundary-free SST of a two-component respiratory signal.}
+\label{fig:resp.2}
+\end{figure} 
+}
+
 \subsubsection{Photoplethysmogram}

### 2021-01-24 Scripts/BoundEffRedExamples/ABPexample.m

@@ -1,0 +1,201 @@
+clear all; close all; clc;
+addpath(genpath('../../TimeFrequencyScaleRep'));
+addpath('../../Algorithm/');
+
+%% load Arterial Blood Pressure signal
+addpath ../../Signals/
+data = load('ABP');
+
+xtot = data.abp';
+
+
+HOP0 = 8 ;
+xtot = xtot(1:HOP0:end) ;
+fs = 1000/HOP0 ; % sampling frequency
+
+n0 = 1 ;
+N = ceil(3*length(xtot)/4) ;
+nf = n0 + N-1 ;
+
+x = xtot(n0:nf);
+mu = mean(x) ;
+s = std(x) ;
+x = (x - mu)/s ;
+
+%% Forecasting
+
+HOP = 1 ;
+extSEC = 48 ; % the extension is of extSEC second
+L = round( extSEC*fs ) ;
+extM = round( 1.25*L ) ; % dimension of embedding / signals length
+extK = round( 1.5*extM );  % number of points to estimate A / size of datasets
+
+method.name = 'SigExt' ;
+tic; 
+xxLSE = forecasting(x,L,HOP,extK,extM,method);
+xxLSE = [x; xxLSE] ;
+LSEtime = toc;
+
+% method.name = 'edmd' ;
+% method.param = 100 ;
+% tic; 
+% xxEDMD = forecasting(x,L,HOP,extK,extM,method); 
+% xxEDMD = [x; xxEDMD] ;
+% EDMDtime = toc;
+% 
+% method.name = 'gpr' ;
+% tic; 
+% xxGPR = forecasting(x,L,HOP,extK,extM,method); 
+% xxGPR = [x; xxGPR] ;
+% GPRtime = toc;
+% 
+% fprintf(' __________________________________________\n')
+% fprintf('| Extension Method | Computing time (sec.) |\n')
+% fprintf('|------------------------------------------|\n')
+% fprintf('|     SigExt       |         %.2f          |\n', LSEtime)
+% fprintf('|  EDMD extension  |         %.2f          |\n', EDMDtime)
+% fprintf('|  GPR extension   |       %.2f          |\n', GPRtime)
[…]

### 2021-01-24 Scripts/BoundEffRedExamples/AWFexample.m

@@ -1,0 +1,197 @@
+clear all; close all; clc;
+addpath(genpath('../../TimeFrequencyScaleRep'));
+addpath('../../Algorithm/');
+
+%% load AWF
+addpath ../../Signals/
+data = load('AWF');
+
+xtot = data.AWF ;
+fs = data.fs ; % sampling frequency
+
+HOP0 = 5 ;
+xtot = xtot(1:HOP0:end) ;
+fs = fs/HOP0 ;
+
+n0 = 1 ;
+N = ceil(4*length(xtot)/5) ; %10e3 ;
+nf = n0 + N-1 ;
+
+x = xtot(n0:nf);
+mu = mean(x) ;
+s = std(x) ;
+x = (x - mu)/s ;
+
+%% Forecasting
+
+HOP = 1 ;
+extSEC = 57 ; % the extension is of extSEC second
+L = round( extSEC*fs ) ;
+extM = round( 1.5*L ) ; % dimension of embedding / signals length
+extK = round( 1.8*extM );  % number of points to estimate A / size of datasets
+
+method.name = 'SigExt' ;
+tic; 
+xxLSE = forecasting(x,L,HOP,extK,extM,method);
+xxLSE = [x; xxLSE] ;
+LSEtime = toc;
+
+% method.name = 'edmd' ;
+% method.param = 100 ;
+% tic; 
+% xxEDMD = forecasting(x,L,HOP,extK,extM,method); 
+% xxEDMD = [x; xxEDMD] ;
+% EDMDtime = toc;
+% 
+% method.name = 'gpr' ;
+% tic; 
+% xxGPR = forecasting(x,L,HOP,extK,extM,method); 
+% xxGPR = [x; xxGPR] ;
+% GPRtime = toc;
+% 
+% fprintf(' __________________________________________\n')
+% fprintf('| Extension Method | Computing time (sec.) |\n')
+% fprintf('|------------------------------------------|\n')
+% fprintf('|     SigExt       |         %.2f          |\n', LSEtime)
+% fprintf('|  EDMD extension  |         %.2f          |\n', EDMDtime)
+% fprintf('|  GPR extension   |       %.2f          |\n', GPRtime)
+% fprintf('|__________________|_______________________|\n')
[…]

### 2021-01-24 Paper/Supplementary material/SupMat.tex
figs

@@
 This TF analysis brings out both components---the cardiac component, the most energetic, is located around the fundamental frequency $1$~Hz and its multiples, while the respiratory component, less visible, is located around $0.2$~Hz.
-The performance index of this representation takes the value $D=0.XXX$.
+The performance index of this representation takes the value $D=0.850$.
 Note that, here, boundary effects are only reduced on the right-side boundary of the TF domain.
-This means that {\sf BoundEffRed} has reduced the right-side boundary effects by about $XX\%$ with respect to the ordinary SST.
+This means that {\sf BoundEffRed} has reduced the right-side boundary effects by about $15\%$ with respect to the ordinary SST.
 This shows the ability of our algorithm to work on signals containing several nonstationary components.
@@
 \centering
-%\includegraphics[width=.48\textwidth]{ABPsst.eps}
-%\includegraphics[width=.48\textwidth]{ABPberSST.eps}
+\includegraphics[width=.48\textwidth]{ABPsst.eps}
+\includegraphics[width=.48\textwidth]{ABPberSST.eps}
 \caption{Ordinary SST (left) and boundary-free SST (right) of an arterial blood pressure recording}

### 2021-01-24 Scripts/BoundEffRedExamples/ABPexample.m

@@ -25,7 +25,7 @@ x = (x - mu)/s ;
 %% Forecasting
 
 HOP = 1 ;
-extSEC = 48 ; % the extension is of extSEC second
+extSEC = 30 ; % the extension is of extSEC second
 L = round( extSEC*fs ) ;
 extM = round( 1.25*L ) ; % dimension of embedding / signals length
 extK = round( 1.5*extM );  % number of points to estimate A / size of datasets

### 2021-01-25 Scripts/EvalPerfSigExt/ResultsAHM.m
paper and results

@@ -22,7 +22,7 @@ fprintf('|-----------------|-------------|---------|-------|\n')
 
 indM = 1 ;
 for extM = extMval
-    fprintf('| SigExt (M=%4i) |   %7.3f   |  %5.3f  | %5.3f |\n', extM, CPUtimeSigExt(indM) , mean(MSEXP(:,indM)), std(MSEXP(:,indM)) ) ;
+    fprintf('| SigExt (M=%4i) |  %8.3f   |  %5.3f  | %5.3f |\n', extM, CPUtimeSigExt(indM) , mean(MSEXP(:,indM)), std(MSEXP(:,indM)) ) ;
     indM = indM + 1 ;
 end
 indM = indM +1 ;
@@ -40,9 +40,14 @@ MSEEDMD = MSE.EDMD ;
 CPUtimeGPR = mean(CPUtimeXP.GPR) ;
 MSEGPR = MSE.GPR ;
 
+dataTBATS = table2array( readtable('../../Results/PerfAHM_TBATS.csv','Range','B:D','TreatAsEmpty','NA') ) ;
+MSETBATS = dataTBATS(:,2) ;
+CPUtimeTBATS = dataTBATS(~isnan(dataTBATS(:,3)),3) ;
 
-fprintf('|  Symmetrization |   %7.3f   |  %5.3f  | %5.3f |\n', CPUtimeSYM , mean(MSESYM), std(MSESYM) ) ;
-fprintf('|       EDMD      |   %7.3f   |  %5.3f  | %5.3f |\n', CPUtimeEDMD , mean(MSEEDMD), std(MSEEDMD) ) ;
-fprintf('|       GPR       |   %7.3f   |  %5.3f  | %5.3f |\n', CPUtimeGPR , mean(MSEGPR), std(MSEGPR) ) ;
 
+
+fprintf('|  Symmetrization |  %8.3f   |  %5.3f  | %5.3f |\n', CPUtimeSYM , mean(MSESYM), std(MSESYM) ) ;
+fprintf('|       EDMD      |  %8.3f   |  %5.3f  | %5.3f |\n', CPUtimeEDMD , mean(MSEEDMD), std(MSEEDMD) ) ;
+fprintf('|       GPR       |  %8.3f   |  %5.3f  | %5.3f |\n', CPUtimeGPR , mean(MSEGPR), std(MSEGPR) ) ;
[…]

### 2021-01-25 Paper/response/response.tex
response

@@
 Please add the example of a real life signal containing two or more simple nonstationary components plus noise.}\\
-\textbf{\color{blue} Idea?}
+\textbf{Section IV-D1 now includes the application of \textsf{BoundEffRed} to a respiratory signal that contains not only the respiratory cycle, but also a
+cardiogenic artifact. We then show the ability of our algorithm to work on
+signals containing several nonstationary components.
+Another application to an atrial blood pressure signal is provided in section III of the Supplementary Materials.
+This signal also contains both simultaneously a cardiac component and a respiratory component.}
 \end{enumerate}

### 2021-01-25 Scripts/EvalPerfSigExt/ForecastAMFM_TBATS.r

@@ -42,7 +42,7 @@ extSEC = 0.1 # the extension is of extSEC second
 L = round( extSEC*fs )
 tt = seq(from = 0, to = 1+L/fs, by = 1/fs)
 
-x0 = xx0[1:N] # restriction to the measurement interval
+
 xextR = xx0[(N+1):(N+L)] # Part to be estimated
 
 sigman = 1e-2
@@ -58,7 +58,8 @@ VarTBATS = matrix(0,nrow=nbXP,ncol=L)
 start_time = Sys.time()
 for (ind in 1:nbXP){
   noise = sigman*rnorm(N+L)
-  x = x0 + noise[1:N] # signal to be extended
+  xx = xx0 + noise # signal to be extended
+  x = xx[1:N] # restriction to the measurement interval
   
   xxTBATS = TBATSextension(x,ss,L,xextR)
   
@@ -66,12 +67,12 @@ for (ind in 1:nbXP){
   VarTBATS[ind,] = ( xxTBATS[(N+1):(N+L)] - xx0[(N+1):(N+L)] )^2
   
   FileName = paste('../../Results/extTBATS_',ind,'.csv',sep="")
-  write.csv(xxTBATS,FileName)
+  write.csv(c(xxTBATS,xx),FileName)
 }
 end_time = Sys.time()
 
 BiasXP_TBATS = rowMeans(MeanTBATS)
-VarianceXP_TBATS = colMeans(VarTBATS)
+VarianceXP_TBATS = rowMeans(VarTBATS)
 CPUtimeXP_TBATS = (1/nbXP)*(end_time - start_time)
 
 length(CPUtimeXP_TBATS) = length(BiasXP_TBATS)

### 2021-01-25 Paper/BoundEffRed.tex
paper

--- 2020-09-11 Notes/BoundEffRed.tex
@@
 There exist extension schemes based on physically relevant dynamical models, such as the Extended Dynamic Mode Decomposition~\cite{Williams15data} (EDMD) and the Gaussian process regression~\cite{Rasmussen06gaussian,Roberts13Gaussian} (GPR).
+% 
+In speech processing, dynamic mode predictors have also been proposed~\cite{Vargas11speech} to forecast signals falling into the so-called \textit{source--filter} model.
+This gradient-based technique rely on the shadowing approach proposed in~\cite{Grebogi90shadowing}.
+%
 There also exist extension schemes based on stochastic models, such as the Trigonometric, Box-Cox transformation, ARMA errors, Trend and Seasonal components (TBATS) algorithm~\cite{DeLivera11forecasting}, the dynamic linear models \cite{west2006bayesian}.
@@
 The numerical results also confirmed the robustness to noise of {\sf BoundEffRed}, as well as its ability to be applied to many TF representations.
-An additional application to ECG is included in the Supplementary Materials.
+Additional applications to ECG and atrial blood pressure signals are included in the Supplementary Materials.
 

### 2021-01-25 Paper/results.tex
paper and results
paper

@@
 %
-{\color{red}
 In practice, the asymptotic behavior of {\sf SigExt}, illustrated here on a synthetic signal, can be generalized to the study of real-life signals, as long as the duration of the recorded signal largely exceeds that of the desired extension, which allows the user to set $K\gg L$.
-}
 
@@
 
-{\color{red}
 \paragraph{Influence of the subsignals length $M$}
@@
 The performance of {\sf BoundEffRed} on other TF representations is more thoroughly analyzed on real-life signals in the following.
-}
 
@@
    \hline
-   TBATS  & {\color{blue}$x$} & {\color{blue}$x \pm  x$} & {\color{blue}$x \pm  x$}\\
+   TBATS  & $1776.778$ & $0.686 \pm  0.467$ & $0.3847 \pm  0.1397$\\
    \hline
@@
 
-{\color{red}
 Then, we consider a second respiratory signal that contains not only the respiratory cycle, but also the cardiogenic artifact which reflects the hemodynamic component.
@@
 \end{figure}
-}
 

### 2021-01-26 Paper/Supplementary material/SupMat.tex
paper

@@
 
-{\color{red}
 \section{Application to a Multicomponent Cardiac Signal}
@@
 In addition to the blood pressure measurement, this signal contains a slow varying component, which is the respiratory component.
-Fig.~\ref{fig:abp} displays the ordinary and boundary-free SSTs of a 150-second-long extract of this signal, obtained via {\sf BoundEffRed}.
+Fig.~\ref{fig:abp} displays the ordinary and boundary-free SSTs of a 150-second-long extract of this signal.
 This TF analysis brings out both components---the cardiac component, the most energetic, is located around the fundamental frequency $1$~Hz and its multiples, while the respiratory component, less visible, is located around $0.2$~Hz.
@@
 \centering
-\includegraphics[width=.48\textwidth]{ABPsst.eps}
-\includegraphics[width=.48\textwidth]{ABPberSST.eps}
-\caption{Ordinary SST (left) and boundary-free SST (right) of an arterial blood pressure recording}
+\includegraphics[width=\textwidth]{ABPsst.eps}
+\caption{Ordinary SST (left) and boundary-free SST (right) of an arterial blood pressure recording. The red dashed rectangle highlights the
+area where boundary effects may appear.}
 \label{fig:abp}
 \end{figure}
-}
 

### 2021-01-26 Paper/algorithm.tex
paper

@@
 %
-{\color{blue}
-From the Takens' embedding perspective~\cite{Takens81detecting}, $\bX$ can be viewed as a lag map.
+%{\color{blue}
+%From the Takens' embedding perspective~\cite{Takens81detecting}, $\bX$ can be viewed as a lag map.
 Under the manifold assumption of the intrinsic phase space that hosts the dynamics, $f$ describes the dynamics supported on the manifold space, and $\bA$ approximates the dynamics on the manifold.
-}
+%}
 
@@
 %
-{\color{red}
 We then obtain the forecasting of $\bx_{k+\ell}$ by recursively applying the linear relation~\eqref{eq:dyn.model}:
@@
 where $\balpha^{(\ell)}$ denotes the last row of $\tilde\bA^\ell$, \ie~$\balpha^{(\ell)}=\be_M^T\tilde\bA^\ell$.
-}
 

### 2021-01-26 Paper/results.tex
paper

@@
 Table~\ref{tab:mse.sine} shows the resulting mean and standard deviation (SD) of the experimental MSE, according to four different values of the subsignals length $M$.
-Note that choosing a too small value for $M$ does not provide enough information to forecast the signal satisfactorily.
+Note that choosing a too small $M$ does not provide enough information to forecast the signal satisfactorily.
 Furthermore, a value of $M$ that is too large makes the algorithm insensitive to local nonstationarities produced by frequency variations.
@@
 Then, we consider a second respiratory signal that contains not only the respiratory cycle, but also the cardiogenic artifact which reflects the hemodynamic component.
-Fig.~\ref{fig:resp.2} displays the boundary-free SST of a 180-second-long extract of this signal, sample at $125$~Hz, obtained via {\sf BoundEffRed}.
+Fig.~\ref{fig:resp.2} displays the ordinary SST and the boundary-free SST of a 180-second-long extract of this signal, sample at $125$~Hz.
 This TF analysis brings out both components---the respiratory component, the most energetic, is located around the fundamental frequency $0.25$~Hz and its multiples, while the cardiac component, less visible, is located around $1.7$~Hz.
@@
 \includegraphics[width=.48\textwidth]{Resp2comp.eps}
-\caption{Boundary-free SST of a two-component respiratory signal.}
+\caption{Ordinary SST (left) and boundary-free SST (right) of a two-component respiratory signal. The red dashed rectangle highlights the area where boundary effects may appear.}
 \label{fig:resp.2}

### 2021-01-26 Paper/theory.tex
paper

@@
 %
-For the forecasting purpose, it is thus clear that when $K$ is not too large, the signal can be well approximated by~\eqref{eq:sum.sine}.
+For the forecasting purpose, it is thus clear that when the speed of variation of $a_j$ and $\phi'_j$ are slow $K$ can be chosen large while the signal can be well approximated by~\eqref{eq:sum.sine}.
 Hence, Theorem~\ref{th:error} can still be applied to approximate the error.
@@
 
-{\color{red}
 The models of sections~\ref{sse:model.sine} and~\ref{RemarkAHM} consider the meaningful part of the signal to be a deterministic component.
@@
 Thus, the theoretical justifications proposed above are no longer applicable to guarantee the forecasting quality of {\sf SigExt}.
-}
 

### 2021-01-26 Scripts/BoundEffRedExamples/ABPexample.m

@@ -36,56 +36,19 @@ xxLSE = forecasting(x,L,HOP,extK,extM,method);
 xxLSE = [x; xxLSE] ;
 LSEtime = toc;
 
-% method.name = 'edmd' ;
-% method.param = 100 ;
-% tic; 
-% xxEDMD = forecasting(x,L,HOP,extK,extM,method); 
-% xxEDMD = [x; xxEDMD] ;
-% EDMDtime = toc;
-% 
-% method.name = 'gpr' ;
-% tic; 
-% xxGPR = forecasting(x,L,HOP,extK,extM,method); 
-% xxGPR = [x; xxGPR] ;
-% GPRtime = toc;
-% 
-% fprintf(' __________________________________________\n')
-% fprintf('| Extension Method | Computing time (sec.) |\n')
-% fprintf('|------------------------------------------|\n')
-% fprintf('|     SigExt       |         %.2f          |\n', LSEtime)
-% fprintf('|  EDMD extension  |         %.2f          |\n', EDMDtime)
-% fprintf('|  GPR extension   |       %.2f          |\n', GPRtime)
-fprintf('|__________________|_______________________|\n')
-
 %% Plot
 t = linspace(0, (N-1)/fs, N) ;
 tt = linspace(0, (N-1+L)/fs, N+L) ;
 
 xxTRUE = ( xtot(n0:(nf+L)) - mu ) / s ;
-xxZP = [ zeros(L,1); x; zeros(L,1) ] ; % zero padding
 
 figure;
 
-% subplot(3,1,1)
 plot(tt,xxLSE,tt,xxTRUE,'--',t,x,'linewidth',2); grid on;
 ylabel('Signals'); axis tight; xlim([t(end)-1.5*extSEC tt(end)]);
 xticks([]);
 legend({'{\sf SigExt} extension','Ground truth extension','Original signal'},'location','northwest','interpreter','latex');
 set(gca,'fontsize',18);
-
-
-% subplot(3,1,2)
-% plot(tt,xxEDMD,tt,xxTRUE,'--',t,x,'linewidth',2); grid on;
-% ylabel('Signals'); axis tight; xlim([t(end)-1.5*extSEC tt(end)]);
-% xticks([]);
[…]

### 2021-01-26 Scripts/BoundEffRedExamples/figPPG.m

@@ -62,14 +62,19 @@ figure;
 subplot(3,2,[1 2]);
 plot(t,x,'k','linewidth',2); axis tight ; xlim([t(end)-15 t(end)]) ;
 xlabel('Time (s)'); ylabel('Signal'); 
-set(gca,'fontsize',18); grid on;
+set(gca,'fontsize',22); grid on;
 
 subplot(3,2,[3 5]);
 imagesc(tsst,fsst,log1p(abs(sstS)/1e1)); axis xy; xlim([t(end)-15 t(end)])
 xlabel('Time (s)'); ylabel('Frequency (Hz)'); colormap(1-gray)
-set(gca,'fontsize',18);
+set(gca,'fontsize',22);
 
 subplot(3,2,[4 6]);
 imagesc(tsstEXT,fsst,log1p(abs(sstLSE)/1e1)); axis xy; xlim([t(end)-15 t(end)])
 xlabel('Time (s)'); ylabel('Frequency (Hz)'); colormap(1-gray)
-set(gca,'fontsize',18);
+set(gca,'fontsize',22);
+
+dim = [.42 0.11 .045 0.515] ;
+annotation('rectangle',dim,'Color','red','linewidth',2,'linestyle','--') ;
+dim = [.86 0.11 .045 0.515] ;
+annotation('rectangle',dim,'Color','red','linewidth',2,'linestyle','--') ;

### 2021-01-26 Paper/results.tex
paper

@@
 Then, we consider a second respiratory signal that contains not only the respiratory cycle, but also the cardiogenic artifact which reflects the hemodynamic component.
-Fig.~\ref{fig:resp.2} displays the ordinary SST and the boundary-free SST of a 180-second-long extract of this signal, sample at $125$~Hz.
+The top of Fig.~\ref{fig:resp.2} displays the ordinary SST and the boundary-free SST of a 180-second-long extract of this signal, sample at $125$~Hz.
 This TF analysis brings out both components---the respiratory component, the most energetic, is located around the fundamental frequency $0.25$~Hz and its multiples, while the cardiac component, less visible, is located around $1.7$~Hz.
-The performance index of this representation takes the value $D=0.314$.
 Note that, here, boundary effects are only reduced on the right-side boundary of the TF domain.
+The middle of Fig.~\ref{fig:resp.2} show a zoom near the right-boundary of the ordinary SST, while the bottom of Fig.~\ref{fig:resp.2} respectively show the respective zoom on the boundary-free SST.
+In this area, the boundary-free SST makes it possible to disentangle the components contained in the signal.
+Moreover, the performance index of this representation takes the value $D=0.314$.
 This means that {\sf BoundEffRed} has reduced the right-side boundary effects by about $69\%$ with respect to the ordinary SST.
@@
 \includegraphics[width=.48\textwidth]{Resp2comp.eps}
-\caption{Ordinary SST (left) and boundary-free SST (right) of a two-component respiratory signal. The red dashed rectangle highlights the area where boundary effects may appear.}
+\includegraphics[width=.48\textwidth]{Resp2compZOOM.eps}
+\caption{Ordinary SST (top-left) and boundary-free SST (top-right) of a two-component respiratory signal.
+Middle and bottom: zoom on the areas delimited by the red dashed rectangle; these images are rotated 90 degrees clockwise.}
 \label{fig:resp.2}

### 2021-01-27 Paper/response/response.tex
paper

@@
 \textbf{We have added the new paragraph IV-C2a where we discuss the influence of the parameter $M$ on the forecasting performance of an adaptive harmonic signal.
-In particular we mention that that choosing a too small value for $M$ does not provide enough information to forecast the signal satisfactorily, while a value of $M$ that is too large makes the algorithm insensitive to local nonstationarities produced by frequency variations.
+In particular we mention that choosing a too small value for $M$ does not provide enough information to forecast the signal satisfactorily, while a value of $M$ that is too large makes the algorithm insensitive to local nonstationarities produced by frequency variations.
 The user must then make a compromise on the choice of $M$ in order to optimize the performance of the algorithm.
@@
 \textbf{Section IV-D1 now includes the application of \textsf{BoundEffRed} to a respiratory signal that contains not only the respiratory cycle, but also a
-cardiogenic artifact. We then show the ability of our algorithm to work on
+cardiac component, knowm as the cardiogenic artifact. We then show the ability of our algorithm to work on
 signals containing several nonstationary components.
@@
 \textit{The reference to Takens is slightly confusing could this be explained more clearly. You are not referring to nonlinear dynamic signals so I am not sure it is relevant.} \\
-\textbf{\color{blue} Hau-Tieng?}
+\textbf{Reviewer's comment is judicious.
+The paragraph comparing the linear dynamical model that we propose with the Takens' embedding wasn't appropriate.
+For the sake of clarity, this paragraph has been removed.}
 \item

### 2021-01-27 Paper/BoundEffRed.tex
paper
fig

@@
 The result of the SST of a PPG is shown in the lower-left corner of Fig.~\ref{fig:ex.intro} (see section~\ref{ssse:ppg} for a comprehensive description).
-The distortion resulting from boundary effects is clearly visible on the right side of this representation.
+The distortion resulting from boundary effects is clearly visible on the right side of this representation---this area is highlighted by the red dashed rectangle.
 Indeed, while in the major left part of the image, clear lines stand out, they become blurred as they approach the right boundary of the image.
@@
 \includegraphics[width=.48\textwidth]{PPGsig.eps}
-%\includegraphics[width=.24\textwidth]{zoomSSTintro.eps}
-%\includegraphics[width=.24\textwidth]{zoomSSTBoundEffRed.eps}
 \caption{A segment of PPG signal (top) and the right boundary of a TF representation determined by the SST without extension (bottom left), and the right boundary of a TF representation determined by the SST with the proposed boundary effect reduction algorithm by forecasting (bottom right).
-The window length for the SST is 12~seconds.}
+The window length for the SST is 12~seconds.
+A video of these real-time SSTs is available online at \url{https://github.com/AdMeynard/BoundaryEffectsReduction}}
 \label{fig:ex.intro}

### 2021-01-27 Paper/results.tex
paper

@@
 We forecast signal extensions of $0.1$~s, \ie~$L=700$.
-Table~\ref{tab:mse.sine} shows the resulting mean and standard deviation (SD) of the experimental MSE, according to four different values of the subsignals length $M$.
+Table~\ref{tab:mse.sine} shows the resulting mean and standard deviation (SD) of the experimental MSE, according to three different values of the subsignals length $M$.
 Note that choosing a too small $M$ does not provide enough information to forecast the signal satisfactorily.
-Furthermore, a value of $M$ that is too large makes the algorithm insensitive to local nonstationarities produced by frequency variations.
+Furthermore, a too large $M$ makes the algorithm insensitive to local nonstationarities produced by frequency variations.
 The user must then make a compromise on the choice of $M$ in order to optimize the performance of the algorithm.
@@
    \hline
-   {\sf SigExt ($M\!=\!3000$)} & $2.881$ & $5.365 \pm  3.919$ & $0.0117 \pm  0.0038$\\
-   \hline
+%   {\sf SigExt ($M\!=\!3000$)} & $2.881$ & $5.365 \pm  3.919$ & $0.0117 \pm  0.0038$\\
+%   \hline
    Symmetric &  $<0.001$ & $5.452 \pm  0.002$ & $1.8209 \pm  0.0025$\\
@@
 
-Then, we consider a second respiratory signal that contains not only the respiratory cycle, but also the cardiogenic artifact which reflects the hemodynamic component.
+Then, we consider a second respiratory signal that contains not only the respiratory cycle, but also the cardiogenic artifact which reflects the hemodynamic component~\cite{Smith94recognition}.
 The top of Fig.~\ref{fig:resp.2} displays the ordinary SST and the boundary-free SST of a 180-second-long extract of this signal, sample at $125$~Hz.
@@
 \caption{Ordinary SST (top-left) and boundary-free SST (top-right) of a two-component respiratory signal.
-Middle and bottom: zoom on the areas delimited by the red dashed rectangle; these images are rotated 90 degrees clockwise.}
+Middle and bottom: zoom on the areas delimited by the red dashed rectangles; these images are rotated 90 degrees clockwise.}
 \label{fig:resp.2}

### 2021-01-28 Paper/response/response.tex
polish

@@
 \textbf{We have added the new paragraph IV-C2a where we discuss the influence of the parameter $M$ on the forecasting performance of an adaptive harmonic signal.
-In particular we mention that choosing a too small value for $M$ does not provide enough information to forecast the signal satisfactorily, while a value of $M$ that is too large makes the algorithm insensitive to local nonstationarities produced by frequency variations.
+In particular, we mention that choosing a too small $M$ does not provide enough information to forecast the signal satisfactorily, while a too large $M$ makes the algorithm insensitive to local nonstationarities produced by frequency variations.
 The user must then make a compromise on the choice of $M$ in order to optimize the performance of the algorithm.
-A choice of $M$ such that the subsignals contain about a
-dozen oscillations is generally appropriate.}
+A choice of $M$ such that the subsignals contain about a dozen oscillations is generally appropriate.}
 \item
@@
 Please add the example of a real life signal containing two or more simple nonstationary components plus noise.}\\
-\textbf{Section IV-D1 now includes the application of \textsf{BoundEffRed} to a respiratory signal that contains not only the respiratory cycle, but also a
-cardiac component, knowm as the cardiogenic artifact. We then show the ability of our algorithm to work on
-signals containing several nonstationary components.
+\textbf{Section IV-D2 now includes the application of \textsf{BoundEffRed} to a respiratory signal that contains not only the respiratory cycle, but also a
+cardiac component, knowm as the cardiogenic artifact.
+We then show the ability of our algorithm to work on signals containing several nonstationary components.
 Another application to an atrial blood pressure signal is provided in section III of the Supplementary Materials.
@@
 In other words, to ensure the relevance of the extension produced by \textsf{SigExt}, the user should be able to use the previously recorded signal for at least three times the duration of the desired extension, approximately.
-For example, in section IV-D2, the analysis of the PPG signal requires 7-second-long extensions to effectively reduce boundary effects in the TF domain.
+For example, in section IV-D3, the analysis of the PPG signal requires 7-second-long extensions to effectively reduce boundary effects in the TF domain.
 About twenty seconds of recording are thus necessary to ensure the fidelity of our algorithm.}

### 2021-01-28 Paper/BoundEffRed.tex
polish

@@
 The window length for the SST is 12~seconds.
-A video of these real-time SSTs is available online at \url{https://github.com/AdMeynard/BoundaryEffectsReduction}}
+Videos of these real-time SSTs are available online at \url{https://github.com/AdMeynard/BoundaryEffectsReduction}}
 \label{fig:ex.intro}
@@
 %
+{\color{red}
 In speech processing, dynamic mode predictors have also been proposed~\cite{Vargas11speech} to forecast signals falling into the so-called \textit{source--filter} model.
 This gradient-based technique rely on the shadowing approach proposed in~\cite{Grebogi90shadowing}.
+}
 %
@@
 To our knowledge, such theoretical analysis does not exist in the literature.
-See the bottom right of Fig.~\ref{fig:ex.intro} for a snapshot of the result.
+See the bottom of Fig.~\ref{fig:ex.intro}---in particular, the regions indicated by the dashed boxes---for a snapshot of the result.
 The main benefit of this simple approach is a numerically efficient solution with a theoretical guarantee for real-time analysis purposes.

### 2021-01-28 Paper/Supplementary material/SupMat.tex
polish

@@
 \setcounter{equation}{25}
-\setcounter{figure}{8}
+\setcounter{figure}{9}
 \setcounter{table}{3}
@@
 
+{\color{red}
 \section{Application to a Multicomponent Cardiac Signal}
-We consider a cardiac signal, namely an arterial blood pressure recording, sampled at $1000$~Hz.
+We consider a cardiac signal, namely an arterial blood pressure (ABP) recording, sampled at $1000$~Hz.
 In addition to the blood pressure measurement, this signal contains a slow varying component, which is the respiratory component.
-Fig.~\ref{fig:abp} displays the ordinary and boundary-free SSTs of a 150-second-long extract of this signal.
-This TF analysis brings out both components---the cardiac component, the most energetic, is located around the fundamental frequency $1$~Hz and its multiples, while the respiratory component, less visible, is located around $0.2$~Hz.
-The performance index of this representation takes the value $D=0.850$.
-Note that, here, boundary effects are only reduced on the right-side boundary of the TF domain.
-This means that {\sf BoundEffRed} has reduced the right-side boundary effects by about $15\%$ with respect to the ordinary SST.
+The top Fig.~\ref{fig:abp} displays an excerpt of the signal along with a 7-second-long extension obtained by the {\sf SigExt} forecasting.
+The forecasting breaks the waveform of the oscillations because of its inability to forecast the high-frequency harmonics contained in the signal.
+Nevertheless, as long as the low-frequency harmonics of the components contained in the signal are preserved, the forecasting is sufficient to reduce boundary effects in the low-frequency TF domain.
+
+The ordinary and boundary-free SSTs of this signal are displayed in the bottom of Fig.~\ref{fig:abp}.
+These TF analyses brings out both components---the cardiac component, the most energetic, indicated by the blue arrows, is located around the fundamental frequency $1$~Hz and its multiples, while the respiratory component, less visible, indicated by the green arrows, is located around $0.2$~Hz.
+Clearly, near the boundaries, {\sf BoundEffRed} helps improve the quality of the TF representation.
+Besides, the performance index of this representation takes the value $D=0.679$.
+This means that {\sf BoundEffRed} has reduced the right-side boundary effects by about $32\%$ with respect to the ordinary SST.
 This shows the ability of our algorithm to work on signals containing several nonstationary components.
@@
 \centering
+\includegraphics[width=\textwidth]{ABP.eps}
 \includegraphics[width=\textwidth]{ABPsst.eps}
-\caption{Ordinary SST (left) and boundary-free SST (right) of an arterial blood pressure recording. The red dashed rectangle highlights the
-area where boundary effects may appear.}
+\caption{Atrial blood pressure signal (top), and the associated
+{\sf SigExt} extension (blue).
+Ordinary SST (bottom left) and boundary-free SST (bottom right) of the arterial blood pressure signal.
+The window length for the SST is 18~seconds.
+The green arrows indicate the instantaneous frequency of the respiratory component.
+The blue arrows, on the other hand, indicate the instantaneous frequency of the blood pressure component.
+The red dashed boxes highlight the areas where boundary effects may appear.}
 \label{fig:abp}
 \end{figure} 
+}
 

### 2021-01-28 Paper/algorithm.tex
polish

@@
 \paragraph{Dynamical Model} 
-Establishing a dynamical model consists of determining the relation linking $\bY$ to $\bX$, that is finding a function $f$ so that
+Establishing a dynamical model means determining the relation linking $\bY$ to $\bX$; that is, finding a function $f$ so that
 \begin{equation}
@@
 
+{\color{red}
 \paragraph{Forecasting}
-The forecasting method consists in estimating the unknown matrix $\bA$. Indeed, let $\tilde\bA$ denotes the estimate of $\bA$. 
+Determining the forecasting amounts to estimating the unknown matrix $\bA$. Indeed, let $\tilde\bA$ denotes the estimate of $\bA$. 
 %
@@
 where $\balpha^{(\ell)}$ denotes the last row of $\tilde\bA^\ell$, \ie~$\balpha^{(\ell)}=\be_M^T\tilde\bA^\ell$.
+}
 

### 2021-01-28 Paper/response/coverletter.tex
polish

\documentclass[11pt,letterpaper,roman]{moderncv}

\usepackage[english]{babel}
\usepackage{ragged2e}
\usepackage{graphicx} 
\usepackage{xcolor}

\moderncvstyle{classic}
\moderncvcolor{blue} 
\nopagenumbers{}
\usepackage[utf8]{inputenc}
\usepackage[scale=0.75]{geometry}
\usepackage{footmisc}
%\usepackage[style=mla]{biblatex}

\name{Adrien Meynard\\}{Hau-Tieng Wu}
\title{Cover letter}
\address{Duke University}{Durham, NC}{USA}
\email{adrien.meynard@duke.edu}

%\bibliography{Sampta17}

\begin{document}
\recipient{Prof. Lorenzo Galleani}{IEEE Transactions on Signal Processing}
\date{January 29, 2021}
\opening{Dear Associate Editor,}
\closing{Yours faithfully,}

\makelettertitle
\justify

Please find the revised manuscript:
\begin{center}
\emph{An Efficient Forecasting Approach to Reduce Boundary Effects in Real-Time Time-Frequency Analysis}
\end{center}
which we are submitting for possible publication in \textit{IEEE Transactions on Signal Processing}.

Thank you for the first evaluation of our article. In this new version, we have taken into account the reviewers' comments and suggestions as much as we could. In particular, we would like to mention that we have addressed the points that you emphasized in the decision letter. Here are our responses.

\begin{enumerate}
\item
\textit{In section IV.C-2, discuss how the performances of the proposed method change with respect to $M$, and explain how to tune its value in applications (comment 3 by Reviewer 1).}\\
\textbf{We have added the new paragraph IV-C2a where we discuss the influence of the parameter $M$ on the forecasting performance of an adaptive harmonic signal. In particular, we mention that choosing a too small $M$ does not provide enough information to forecast the signal satisfactorily, while a too large $M$ makes the algorithm insensitive to local nonstationarities produced by frequency variations. The user must then make a compromise on the choice of $M$ in order to optimize the performance of the algorithm. A choice of $M$ such that the subsignals contain about a dozen oscillations is generally appropriate.}
\item
\textit{Discuss the shadowing approach proposed by C. Gregobi and compare it to your method (comment 1 by Reviewer 2).}\\
\textbf{We have mentioned this approach in the introduction of our article. We precise that dynamic mode predictors have been proposed in speech processing by Vargas and McLaughlin, to forecast signals  falling into the so-called source--filter model. This gradient-based technique rely on the shadowing approach proposed in Grebogi et al. Nevertheless,  since the source--filter is not adapted to model physiological signals on which we focus, we do not implement this method in our article.}
\item
\textit{Discuss the robustness of your method with respect to possible stochastic components in the signal (comment 2 by Reviewer 2).}\\
\textbf{A new paragraph has been added to precise the limitations of the theoretical justifications on the performance of \textsf{SigExt}. Indeed, the models of sections III-A and III-C consider the meaningful part of the signal to be a deterministic component. These models are purposely adapted to signals showing local line spectra. The physiological signals we are interested in, such as respiratory or cardiac signals, have this characteristic (see section IV-D). Signals with wider local spectra, such as electroencephalography signals, do not fall into this category, and are more faithfully modeled as random signals. Thus, the theoretical justifications proposed above are no longer applicable to guarantee the forecasting quality of \textsf{SigExt}. A more robust approach to the study of signals following this model is envisaged in our future work.}
\end{enumerate}

A document containing detailed responses to the reviewers' comments is enclosed to this letter.

\makeletterclosing

\end{document}

### 2021-01-28 Paper/results.tex
polish

@@
 %
+{\color{red}
 In practice, the asymptotic behavior of {\sf SigExt}, illustrated here on a synthetic signal, can be generalized to the study of real-life signals, as long as the duration of the recorded signal largely exceeds that of the desired extension, which allows the user to set $K\gg L$.
+}
 
@@
 
+{\color{red}
 \paragraph{Influence of the subsignals length $M$}
@@
 A choice of $M$ such that the subsignals contain about a dozen oscillations is generally appropriate.
+}
 
@@
 Thus, {\sf SigExt} is the extension method that optimizes the trade-off between forecasting quality and computing time.
+%
+{\color{red}
 Besides, the last column of Table~\ref{tab:mse.sine} gives the performance index $D$ of the boundary-free STFT associated with each extension method.
@@
 The performance of {\sf BoundEffRed} on other TF representations is more thoroughly analyzed on real-life signals in the following.
+}
 
@@
 
-Then, we consider a second respiratory signal that contains not only the respiratory cycle, but also the cardiogenic artifact which reflects the hemodynamic component~\cite{Smith94recognition}.
-The top of Fig.~\ref{fig:resp.2} displays the ordinary SST and the boundary-free SST of a 180-second-long extract of this signal, sample at $125$~Hz.
+{\color{red}
+\subsubsection{Two-Component Respiratory Signal}
+Then, we consider a second respiratory signal that contains not only the respiratory cycle, but also a cardiac component known as the cardiogenic artifact~\cite{Smith94recognition}.
+The top of Fig.~\ref{fig:resp.2} shows an excerpt of the measured signal, as well as a 10-second-long {\sf SigExt} extension.
+Below, the simultaneously recorded ECG signal emphasizes the coincidences between the cardiogenic artifact of the respiratory signal and the ECG.
+
+\begin{figure}
+\centering
+\includegraphics[width=.48\textwidth]{Resp2.eps}
+\caption{Two-component respiratory signal (top), and the associated {\sf SigExt} extension (blue); the simultaneously recorded ECG signal is below.}
+\label{fig:resp.2}
+\end{figure}
+
+The top of Fig.~\ref{fig:resp.2.sst} displays the ordinary SST and the boundary-free SST of this signal, sampled at $125$~Hz.
 This TF analysis brings out both components---the respiratory component, the most energetic, is located around the fundamental frequency $0.25$~Hz and its multiples, while the cardiac component, less visible, is located around $1.7$~Hz.
 Note that, here, boundary effects are only reduced on the right-side boundary of the TF domain.
-The middle of Fig.~\ref{fig:resp.2} show a zoom near the right-boundary of the ordinary SST, while the bottom of Fig.~\ref{fig:resp.2} respectively show the respective zoom on the boundary-free SST.
+The middle of Fig.~\ref{fig:resp.2.sst} show a zoom near the right-boundary of the ordinary SST, while the bottom of Fig.~\ref{fig:resp.2.sst} respectively show the respective zoom on the boundary-free SST.
 In this area, the boundary-free SST makes it possible to disentangle the components contained in the signal.
-Moreover, the performance index of this representation takes the value $D=0.314$.
-This means that {\sf BoundEffRed} has reduced the right-side boundary effects by about $69\%$ with respect to the ordinary SST.
+Moreover, the performance index of this representation takes the value $D=0.205$.
+This means that {\sf BoundEffRed} has reduced the right-side boundary effects by about $80\%$ with respect to the ordinary SST.
 This shows the ability of our algorithm to work on signals containing several nonstationary components.
@@
 \caption{Ordinary SST (top-left) and boundary-free SST (top-right) of a two-component respiratory signal.
-Middle and bottom: zoom on the areas delimited by the red dashed rectangles; these images are rotated 90 degrees clockwise.}
-\label{fig:resp.2}
+Middle and bottom: zoom on the areas delimited by the red dashed rectangles; these images are rotated 90 degrees clockwise.
+The window length for the SSTs is 35~seconds.}
+\label{fig:resp.2.sst}
 \end{figure}
+}
 

### 2021-01-28 Algorithm/BoundEffRed_RT.m

@@ -1,13 +1,14 @@
 function [TFRtot, ForecastTime, TFRtime] = BoundEffRed_RT(xtot,fs,forecastMethod,basicTF,VideoWriting,varargin)
 % BOUNDEFFRED_RT Real-time implementation of BoundEffRed
-% Usage:	[TFRtot, dt] = BoundEffRed_RT(xtot,fs,forecastMethod,basicTF,VideoWriting)
+% Usage:	[TFRtot, dt] = BoundEffRed_RT(xtot,fs,forecastMethod,basicTF,VideoWriting,VideoName)
 %
 % Input:
 %   xtot: signal to analyze
 %   fs: sampling frequency
 %   forecastMethod: forecasting method
 %   basicTF: time-frequency representation and associated parameters
-%   VideoWriting: When set to 1, the real-time TF represention image is recorded 
+%   VideoWriting: When set to 1, the real-time TF represention image is recorded
+%   VideoName: name of the video to be recorded when VideoWriting=1
 %
 % Output:
 %   TFRtot: Whole time-frequency representation

### 2021-01-28 Scripts/BoundEffRedExamples/AWFexample.m

@@ -4,17 +4,17 @@ addpath('../../Algorithm/');
 
 %% load AWF
 addpath ../../Signals/
-data = load('AWF');
+data = load('respII');
 
 xtot = data.AWF ;
-fs = data.fs ; % sampling frequency
+fs = data.fs.AWF ; % sampling frequency
 
 HOP0 = 5 ;
 xtot = xtot(1:HOP0:end) ;
 fs = fs/HOP0 ;
 
 n0 = 1 ;
-N = ceil(4*length(xtot)/5) ; %10e3 ;
+N = ceil(4*length(xtot)/5) ;
 nf = n0 + N-1 ;
 
 x = xtot(n0:nf);
@@ -25,10 +25,10 @@ x = (x - mu)/s ;
 %% Forecasting
 
 HOP = 1 ;
-extSEC = 57 ; % the extension is of extSEC second
+extSEC = 17.5 ; % the extension is of extSEC second
 L = round( extSEC*fs ) ;
 extM = round( 1.5*L ) ; % dimension of embedding / signals length
-extK = round( 1.8*extM );  % number of points to estimate A / size of datasets
+extK = round( 2.5*extM );  % number of points to estimate A / size of datasets
 
 method.name = 'SigExt' ;
 tic; 
@@ -36,21 +36,15 @@ xxLSE = forecasting(x,L,HOP,extK,extM,method);
 xxLSE = [x; xxLSE] ;
 LSEtime = toc;
 
-%% Plot
+xxTRUE = ( xtot(n0:(nf+L)) - mu ) / s ;
+
 t = linspace(0, (N-1)/fs, N) ;
 tt = linspace(0, (N-1+L)/fs, N+L) ;
 
-xxTRUE = ( xtot(n0:(nf+L)) - mu ) / s ;
-
-figure;
-plot(tt,xxLSE,tt,xxTRUE,'--',t,x,'linewidth',2); grid on;
-ylabel('Signals'); axis tight; xlim([t(end)-18 t(end)+10]); %xlim([t(end)-1.5*extSEC tt(end)]);
-legend({'{\sf SigExt} extension','Ground truth extension','Original signal'},'location','northwest','interpreter','latex');
-set(gca,'fontsize',18);
-
 %% SST and STFT
 basicTF.hop = 5 ;
-twin = 114 ;
[…]

### 2021-01-28 Scripts/EvalPerfBoundEffRed/CompPerfAHM_TBATS.m

@@ -1,4 +1,4 @@
-%% Evaluation of the Performance of SigExt on a AHM oscillatory signal
+%% Evaluation, on a AHM signal, of the performance of TBATS for the reduction of boundary effects
 % Author: Adrien MEYNARD
 % Email: adrien.meynard@duke.edu
 

### 2021-01-28 Scripts/EvalPerfBoundEffRed/EvalPPG.m

@@ -1,4 +1,4 @@
-%% Evaluate BoundEffRed on an PPG
+%% Evaluate BoundEffRed on a PPG signal
 % Author: Adrien MEYNARD
 % Email: adrien.meynard@duke.edu
 

### 2021-01-28 Scripts/EvalPerfBoundEffRed/EvalTHO.m

@@ -1,4 +1,4 @@
-%% Evaluate BoundEffRed on a THO
+%% Evaluate BoundEffRed on a respiratory signal
 % Author: Adrien MEYNARD
 % Email: adrien.meynard@duke.edu
 

### 2021-01-28 Scripts/EvalPerfSigExt/InfluenceSizeDatasetSine.m

@@ -1,4 +1,4 @@
-%% Evaluation on a sine wave of the "training dataset" influence on SigExt's performance
+%% Evaluation, on a sine wave, of the influence of the size of the "training dataset" on SigExt's performance
 % Author: Adrien MEYNARD
 % Email: adrien.meynard@duke.edu
 

### 2021-01-28 Scripts/EvalPerfSigExt/OtherMethAHM.m

@@ -1,4 +1,4 @@
-%% Evaluation on a AHM signal of the sunsignals length influence on SigExt's performance
+%% Evaluation, on a AHM signal, of the performance of alternative extension methods
 % Author: Adrien MEYNARD
 % Email: adrien.meynard@duke.edu
 

### 2021-01-29 Paper/results.tex
minor

@@
 Then, we consider a second respiratory signal that contains not only the respiratory cycle, but also a cardiac component known as the cardiogenic artifact~\cite{Smith94recognition}.
-The top of Fig.~\ref{fig:resp.2} shows an excerpt of the measured signal, as well as a 10-second-long {\sf SigExt} extension.
-Below, the simultaneously recorded ECG signal emphasizes the coincidences between the cardiogenic artifact of the respiratory signal and the ECG.
+The top of Fig.~\ref{fig:resp.2} shows an excerpt of the measured signal, along with a 10-second-long {\sf SigExt} extension.
+Below, the simultaneously recorded ECG signal emphasizes the coincidences between the cardiogenic artifact of the respiratory signal and the QRS complex of the ECG signal.
 

### 2021-01-31 Paper/response/response.tex
typo
new XPs

@@
 \item
-\textit{page 3, equation (7). The authors should provide the definition of the matrix $\tilde{A}^\ell$.}\\
+\textit{page 3, equation (7). The authors should provide the definition of the matrix $\tilde{\mathbf{A}}^\ell$.}\\
 \textbf{The definition is now provided by the unnumbered equation between equation (5) and (6).}
@@
 We then show the ability of our algorithm to work on signals containing several nonstationary components.
-Another application to an atrial blood pressure signal is provided in section III of the Supplementary Materials.
+Another application to a PPG signal is provided in section III of the Supplementary Materials.
 This signal also contains both simultaneously a cardiac component and a respiratory component.}

### 2021-01-31 Paper/Supplementary material/SupMat.tex
minor
new XPs

@@
 Consequently, the forecast produced by {\sf SigExt} is moderately satisfactory.
+Note that TBATS is the only one that seems to accurately capture the locations of QRS complexes after 37 seconds.
+We will explore this long-term prediction capability in the future work.
 
@@
 \section{Application to a Multicomponent Cardiac Signal}
-We consider a cardiac signal, namely an arterial blood pressure (ABP) recording, sampled at $1000$~Hz.
-In addition to the blood pressure measurement, this signal contains a slow varying component, which is the respiratory component.
-The top Fig.~\ref{fig:abp} displays an excerpt of the signal along with a 7-second-long extension obtained by the {\sf SigExt} forecasting.
-The forecasting breaks the waveform of the oscillations because of its inability to forecast the high-frequency harmonics contained in the signal.
+We consider a cardiac signal, namely a photoplethysmogram (PPG) recording, sampled at $300$~Hz.
+In addition to the cardiac cycle measurement, this signal contains a slow varying component, which is the respiratory component, known as respiration induced intensity variation (RIIV)~\cite{Nilsson13respiration}.
+The top of Fig.~\ref{fig:ppg2} displays an excerpt of the signal along with a 3-second-long extension obtained by the {\sf SigExt} forecasting.
+The lower part of Fig.~\ref{fig:ppg2} shows the respiratory signal recording the concentration of $\mathrm{CO}_2$.
+This signal was recorded simultaneously with the PPG signal, and visually highlights the low-frequency respiratory component contained in the PPG signal.
+Indeed, the intervals where the concentration of $\mathrm{CO}_2$ drops---highlighted by the bluish areas---coincide with the decreases in the PPG signal.
+Note also that the forecasting breaks the waveform of the oscillations because of its inability to forecast the high-frequency harmonics contained in the signal.
 Nevertheless, as long as the low-frequency harmonics of the components contained in the signal are preserved, the forecasting is sufficient to reduce boundary effects in the low-frequency TF domain.
 
-The ordinary and boundary-free SSTs of this signal are displayed in the bottom of Fig.~\ref{fig:abp}.
-These TF analyses brings out both components---the cardiac component, the most energetic, indicated by the blue arrows, is located around the fundamental frequency $1$~Hz and its multiples, while the respiratory component, less visible, indicated by the green arrows, is located around $0.2$~Hz.
+The ordinary and boundary-free SSTs of this signal are displayed in Fig.~\ref{fig:ppg2.sst}.
+These TF analyses brings out both components---the fundamental frequency of the cardiac component, the most energetic one, indicated by the blue arrows, is located around $1.3$~Hz, while the respiratory component, less visible, indicated by the green arrows, is located around $0.2$~Hz  and its multiples.
 Clearly, near the boundaries, {\sf BoundEffRed} helps improve the quality of the TF representation.
-Besides, the performance index of this representation takes the value $D=0.679$.
-This means that {\sf BoundEffRed} has reduced the right-side boundary effects by about $32\%$ with respect to the ordinary SST.
+Besides, the performance index of this representation takes the value $D=0.491$.
+This means that {\sf BoundEffRed} has reduced the right-side boundary effects by about $51\%$ with respect to the ordinary SST.
 This shows the ability of our algorithm to work on signals containing several nonstationary components.
@@
 \centering
-\includegraphics[width=\textwidth]{ABP.eps}
-\includegraphics[width=\textwidth]{ABPsst.eps}
-\caption{Atrial blood pressure signal (top), and the associated
+\includegraphics[width=\textwidth]{PPG2.eps}
+\caption{PPG signal (top, yellow), and the associated
 {\sf SigExt} extension (blue).
-Ordinary SST (bottom left) and boundary-free SST (bottom right) of the arterial blood pressure signal.
-The window length for the SST is 18~seconds.
+The simultaneously recorded concentration of $\mathrm{CO}_2$ is below.
+The bluish areas show the synchrony between the drops in $\mathrm{CO}_2$ concentration and the decreases in the PPG signal.}
+\label{fig:ppg2}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{PPG2sst.eps}
+\caption{Ordinary SST (left) and boundary-free SST (right) of the PPG signal.
+The window length for the SSTs is 18~seconds.
 The green arrows indicate the instantaneous frequency of the respiratory component.
@@
 The red dashed boxes highlight the areas where boundary effects may appear.}
-\label{fig:abp}
+\label{fig:ppg2.sst}
 \end{figure} 

### 2021-01-31 Paper/results.tex
new XPs

--- 2021-01-28 Paper/results.tex
@@
 Then, we consider a second respiratory signal that contains not only the respiratory cycle, but also a cardiac component known as the cardiogenic artifact~\cite{Smith94recognition}.
-The top of Fig.~\ref{fig:resp.2} shows an excerpt of the measured signal, as well as a 10-second-long {\sf SigExt} extension.
-Below, the simultaneously recorded ECG signal emphasizes the coincidences between the cardiogenic artifact of the respiratory signal and the ECG.
+The top of Fig.~\ref{fig:resp.2} shows an excerpt of the measured thoracic impedance-based respiratory signal, recorded during a bronchoscopic sedation procedure, along with the {\sf SigExt} extension.
+Below, the simultaneously recorded ECG signal is depicted.
+The magenta dotted lines emphasize the coincidences between the cardiogenic artifact of the respiratory signal and the QRS complexes of the ECG signal.
+Such cardiac information can be harvested and utilized when other first-line heart rate information resource is not available or broken~\cite{Lu19recycling}.
 
@@
 \includegraphics[width=.48\textwidth]{Resp2.eps}
-\caption{Two-component respiratory signal (top), and the associated {\sf SigExt} extension (blue); the simultaneously recorded ECG signal is below.}
+\caption{Two-component respiratory signal (top, yellow), and the associated {\sf SigExt} extension (blue); the simultaneously recorded ECG signal is below.}
 \label{fig:resp.2}
@@
 
-The top of Fig.~\ref{fig:resp.2.sst} displays the ordinary SST and the boundary-free SST of this signal, sampled at $125$~Hz.
-This TF analysis brings out both components---the respiratory component, the most energetic, is located around the fundamental frequency $0.25$~Hz and its multiples, while the cardiac component, less visible, is located around $1.7$~Hz.
-Note that, here, boundary effects are only reduced on the right-side boundary of the TF domain.
+The top of Fig.~\ref{fig:resp.2.sst} displays the ordinary SST and the boundary-free SST of this respiratory signal, sampled at $64$~Hz.
+This TF analysis brings out both components---the respiratory component, indicated by the green arrows, is located around the fundamental frequency $0.3$~Hz and its multiples, while the cardiac component, indicated by the blue arrows, is located around $1.5$~Hz.
+Note that, here, boundary effects are only reduced on the right-side boundary of the TF domain, delimited by the red dashed boxes.
 The middle of Fig.~\ref{fig:resp.2.sst} show a zoom near the right-boundary of the ordinary SST, while the bottom of Fig.~\ref{fig:resp.2.sst} respectively show the respective zoom on the boundary-free SST.
 In this area, the boundary-free SST makes it possible to disentangle the components contained in the signal.
-Moreover, the performance index of this representation takes the value $D=0.205$.
-This means that {\sf BoundEffRed} has reduced the right-side boundary effects by about $80\%$ with respect to the ordinary SST.
+Moreover, the performance index of this representation takes the value $D=0.705$.
+This means that {\sf BoundEffRed} has reduced the right-side boundary effects by about $30\%$ with respect to the ordinary SST.
 This shows the ability of our algorithm to work on signals containing several nonstationary components.
@@
 Middle and bottom: zoom on the areas delimited by the red dashed rectangles; these images are rotated 90 degrees clockwise.
-The window length for the SSTs is 35~seconds.}
+The window length for the SSTs is 20~seconds.
+The instantaneous frequency of the respiration is indicated by the green arrows, while the instantaneous frequency of the cardiogenic artifact is indicated by the blue arrows.}
 \label{fig:resp.2.sst}

### 2021-02-01 Paper/BoundEffRed.tex
new XPs
paper v2

@@
 The window length for the SST is 12~seconds.
-Videos of these real-time SSTs are available online at \url{https://github.com/AdMeynard/BoundaryEffectsReduction}}
+{\color{red}Videos of these real-time SSTs are available online at \url{https://github.com/AdMeynard/BoundaryEffectsReduction}.}}
 \label{fig:ex.intro}
@@
 The numerical results also confirmed the robustness to noise of {\sf BoundEffRed}, as well as its ability to be applied to many TF representations.
-Additional applications to ECG and atrial blood pressure signals are included in the Supplementary Materials.
+Additional applications to ECG and PPG are included in the Supplementary Materials.
 

### 2021-02-01 Paper/results.tex
paper v2

@@
 \includegraphics[width=.48\textwidth]{Resp2.eps}
-\caption{Two-component respiratory signal (top, yellow), and the associated {\sf SigExt} extension (blue); the simultaneously recorded ECG signal is below.}
+\caption{Two-component respiratory signal (top, yellow), and the associated {\sf SigExt} extension (blue); the simultaneously recorded ECG signal is below.
+The vertical magenta dotted lines indicate the cardiac cycles in the respiratory signal.}
 \label{fig:resp.2}
@@
 \label{ssse:ppg}
-We perform a study similar to the previous one on a $640$-second-long photoplethysmogram (PPG) signal extracted from the Physionet dataset~\cite{Pimentel17toward, Goldberger00physiobank}, sampled at $\fs=125$~Hz.
+We consider a $640$-second photoplethysmogram (PPG) signal extracted from the Physionet dataset~\cite{Pimentel17toward, Goldberger00physiobank}, sampled at $\fs=125$~Hz.
 A portion of this signal is displayed in Fig.~\ref{fig:ppg}.
-The estimated 5 seconds extensions of this segment obtained by {\sf SigExt}, EDMD, and GPR forecastings  are superimposed to the ground-truth extension in Fig.~\ref{fig:ppg}.
+The estimated 5-second extensions of this segment obtained by {\sf SigExt}, EDMD, and GPR forecastings  are superimposed to the ground-truth extension in Fig.~\ref{fig:ppg}.
 

### 2021-02-01 Scripts/BoundEffRedExamples/RespIIexample.m

@@ -59,22 +59,29 @@ set(gca,'fontsize',24);
 
 xECG = data.ECG ;
 xECG = (xECG - mean(xECG))/std(xECG) ;
+
+wo = 60/(512/2) ;
+bw = wo/5;
+[num,den] = iirnotch(wo,bw) ;
+yECG = filter(num,den,xECG) ;
+
 tECG = data.timeECG ;
 
-subplot('Position',[0.045 0.255 0.942 0.265]);
-plot(tECG-tm,xECG,'k','linewidth',2); grid on;
-xlabel('Time (s)'); ylabel('ECG signal'); axis tight; xlim([0 t(end)-tm+7.5]);
+subplot('Position',[0.045 0.255 0.942 0.225]);
+plot(tECG-tm,yECG,'k','linewidth',2); grid on;
+xlabel('Time (s)'); ylabel('ECG signal');
+xlim([0 t(end)-tm+7.5]); ylim([-1.5 6]);
 set(gca,'fontsize',24);
 
-annotation('line',[0.142 0.142],[0.672 0.505],'Color','m','LineWidth',3,'LineStyle',':');
-annotation('line',[0.173 0.173],[0.584 0.502],'Color','m','LineWidth',3,'LineStyle',':');
-annotation('line',[0.205 0.205],[0.582 0.497],'Color','m','LineWidth',3,'LineStyle',':');
-annotation('line',[0.301 0.301],[0.661 0.506],'Color','m','LineWidth',3,'LineStyle',':');
-annotation('line',[0.333 0.333],[0.595 0.504],'Color','m','LineWidth',3,'LineStyle',':');
-annotation('line',[0.456 0.456],[0.595 0.477],'Color','m','LineWidth',3,'LineStyle',':');
-annotation('line',[0.487 0.487],[0.595 0.513],'Color','m','LineWidth',3,'LineStyle',':');
-annotation('line',[0.586 0.586],[0.623 0.504],'Color','m','LineWidth',3,'LineStyle',':');
-annotation('line',[0.619 0.619],[0.603 0.516],'Color','m','LineWidth',3,'LineStyle',':');
+annotation('line',[0.142 0.142],[0.672 0.435],'Color','m','LineWidth',4,'LineStyle',':');
+annotation('line',[0.174 0.174],[0.584 0.436],'Color','m','LineWidth',4,'LineStyle',':');
[…]

### 2021-02-01 Paper/BoundEffRed.tex
typo

@@
 We derive the corresponding algorithm for boundary effects reduction.
-In section~\ref{se:theoretical}, we show that the dynamic model we consider is sufficient to extend signals takings the form of sums of sine waves.
-An evaluation of the theoretical performance of our algorithm on a class of signals, the sums of sine waves, is given in section~\ref{se:theoretical}.
+In section~\ref{se:theoretical}, we show that the dynamic model we consider is sufficient to extend signals taking the form of sums of sine waves.
+An evaluation of the theoretical performance of our algorithm on this class of signals is given in section~\ref{se:theoretical}.
 In section~\ref{se:results}, we compare our extension method with more sophisticated methods such as EDMD, GPR, or TBATS.

### 2021-02-14 Paper/BoundEffRed.tex
colors

@@
 The window length for the SST is 12~seconds.
-{\color{red}Videos of these real-time SSTs are available online at \url{https://github.com/AdMeynard/BoundaryEffectsReduction}.}}
+Videos of these real-time SSTs are available online at \url{https://github.com/AdMeynard/BoundaryEffectsReduction}.}
 \label{fig:ex.intro}
@@
 %
-{\color{red}
 In speech processing, dynamic mode predictors have also been proposed~\cite{Vargas11speech} to forecast signals falling into the so-called \textit{source--filter} model.
 This gradient-based technique rely on the shadowing approach proposed in~\cite{Grebogi90shadowing}.
-}
 %

### 2021-02-14 Paper/results.tex
colors

@@
 %
-{\color{red}
 Besides, the last column of Table~\ref{tab:mse.sine} gives the performance index $D$ of the boundary-free STFT associated with each extension method.
@@
 
-{\color{red}
 \subsubsection{Two-Component Respiratory Signal}

### 2021-02-22 Paper/BoundEffRed.tex
final version

@@
 A. Meynard is the corresponding author (e-mail: adrien.meynard@duke.edu).
+
+This paper has supplementary downloadable material available at http://ieeexplore.ieee.org., provided by the author.
+The material includes a technical report that contains proofs of mathematical statements
+and additional illustrations. This material is 1.2~MB in size.
 }}
@@
 \]
-Then, the local analysis of $x$ around the instant $\tau\in I$ rely on the evaluation of the following inner product:
+Then, the local analysis of $x$ around the instant $\tau\in I$ relies on the evaluation of the following inner product:
 \begin{equation}
@@
 This phenomenon is usually understood as the \emph{boundary effect}.
-The result of the SST of a PPG is shown in the lower-left corner of Fig.~\ref{fig:ex.intro} (see section~\ref{ssse:ppg} for a comprehensive description).
+The result of the SST of a PPG is shown in the lower-left corner of Fig.~\ref{fig:ex.intro} (see Section~\ref{ssse:ppg} for a comprehensive description).
 The distortion resulting from boundary effects is clearly visible on the right side of this representation---this area is highlighted by the red dashed rectangle.
@@
 In speech processing, dynamic mode predictors have also been proposed~\cite{Vargas11speech} to forecast signals falling into the so-called \textit{source--filter} model.
-This gradient-based technique rely on the shadowing approach proposed in~\cite{Grebogi90shadowing}.
+This gradient-based technique relies on the shadowing approach proposed in~\cite{Grebogi90shadowing}.
 %
@@
 We derive the corresponding algorithm for boundary effects reduction.
-In section~\ref{se:theoretical}, we show that the dynamic model we consider is sufficient to extend signals taking the form of sums of sine waves.
-An evaluation of the theoretical performance of our algorithm on this class of signals is given in section~\ref{se:theoretical}.
-In section~\ref{se:results}, we compare our extension method with more sophisticated methods such as EDMD, GPR, or TBATS.
+In Section~\ref{se:theoretical}, we show that the dynamic model we consider is sufficient to extend signals taking the form of sums of sine waves.
+An evaluation of the theoretical performance of our algorithm on this class of signals is given in Section~\ref{se:theoretical}.
+In Section~\ref{se:results}, we compare our extension method with more sophisticated methods such as EDMD, GPR, or TBATS.
 We show that our algorithm gives fast results of reasonable quality.

### 2021-02-22 Paper/algorithm.tex
colors
final version

@@
 For instance, the dynamic mode decomposition~\cite{Schmid10dynamic,Williams15data} or other more complicated models \cite{Roberts13Gaussian,DeLivera11forecasting,west2006bayesian,vlachas2018data}, allow this by setting additional constraints on the behavior of $f$.
-We will see, in section~\ref{se:theoretical}, that considering such a complex dynamic model is not necessary for the study of the oscillatory signals of interest to us.
+We will see, in Section~\ref{se:theoretical}, that considering such a complex dynamic model is not necessary for the study of the oscillatory signals of interest to us.
 That is why we consider here a naive dynamical model, assuming that we have the following relation:
@@
 
-{\color{red}
 \paragraph{Forecasting}
@@
 \end{equation}  
-where $\balpha^{(\ell)}$ denotes the last row of $\tilde\bA^\ell$, \ie~$\balpha^{(\ell)}=\be_M^T\tilde\bA^\ell$.
-}
+where $\balpha^{(\ell)}$ denotes the last row of $\tilde\bA^\ell$, \ie, $\balpha^{(\ell)}=\be_M^T\tilde\bA^\ell$.
 
@@
 \]
-Therefore, solving the problem~\eqref{eq:ls.pb}, \ie~$\nabla \cL(\tilde\bA)=\bzero$, gives the following estimate $\tilde\bA$ of the dynamical model matrix $\bA$:
+Therefore, solving the problem~\eqref{eq:ls.pb}, \ie, $\nabla \cL(\tilde\bA)=\bzero$, gives the following estimate $\tilde\bA$ of the dynamical model matrix $\bA$:
 \begin{align}

### 2021-02-22 Paper/results.tex
final version

@@
 \end{itemize}
-In subsections~\ref{ssse:res.sine} and~\ref{ssse:res.ahm}, we study the influence of these parameters. A comparison with the theoretical results of section~\ref{se:theoretical} is also available.
+In Sections~\ref{ssse:res.sine} and~\ref{ssse:res.ahm}, we study the influence of these parameters. A comparison with the theoretical results of Section~\ref{se:theoretical} is also available.
 
@@
 To evaluate the forecasting quality, {\sf SigExt} is applied to $1000$ realizations of the above-described signal.
-We forecast signal extensions of $0.1$~s, \ie~$L=700$.
+We forecast signal extensions of $0.1$~s, \ie, $L=700$.
 Table~\ref{tab:mse.sine} shows the resulting mean and standard deviation (SD) of the experimental MSE, according to three different values of the subsignals length $M$.
-Note that choosing a too small $M$ does not provide enough information to forecast the signal satisfactorily.
-Furthermore, a too large $M$ makes the algorithm insensitive to local nonstationarities produced by frequency variations.
+Note that choosing too small an $M$ does not provide enough information to forecast the signal satisfactorily.
+Furthermore, too large an $M$ makes the algorithm insensitive to local nonstationarities produced by frequency variations.
 The user must then make a compromise on the choice of $M$ in order to optimize the performance of the algorithm.
@@
 \paragraph{Comparison of the extensions methods}
-The same set of experiments is run on the extension methods outlined in section~\ref{sse:methods}.
+The same set of experiments is run on the extension methods outlined in Section~\ref{sse:methods}.
 The results, displayed in Table~\ref{tab:mse.sine}, show that the naive extension we propose gives satisfying results.
@@
 Besides, the last column of Table~\ref{tab:mse.sine} gives the performance index $D$ of the boundary-free STFT associated with each extension method.
-This illustrates the ability of our algorithm to reduce boundary effects to a level comparable to what is possible with the GPR or TBATS extensions.
+This illustrates the ability of our algorithm to reduce boundary effects to a level comparable with what is possible with the GPR or TBATS extensions.
 The performance of {\sf BoundEffRed} on other TF representations is more thoroughly analyzed on real-life signals in the following.
@@
 
-From that long signal, we build a dataset of $379$ non\-overlapping signals of 60 seconds, \ie~$N=6000$.
-For each segment, we apply the forecasting methods introduced in section~\ref{ssse:res.ahm}, including the {\sf SigExt} method detailed in Algorithm~\ref{alg:extension}.
+From that long signal, we build a dataset of $379$ non\-overlapping signals of 60 seconds, \ie, $N=6000$.
+For each segment, we apply the forecasting methods introduced in Section~\ref{ssse:res.ahm}, including the {\sf SigExt} method detailed in Algorithm~\ref{alg:extension}.
 We forecast $7$-second-long extensions on each segment of the signal, corresponding to $L =700$.
 Thus, in order to catch the dynamical behavior, the size of the training signal $M$ is chosen so that $M=\lfloor 1.5L\rfloor$.
-As a result of section~\ref{ssse:res.sine}, we take: $K=\lfloor2.5M\rfloor$.
+As a result of Section~\ref{ssse:res.sine}, we take: $K=\lfloor2.5M\rfloor$.
 The extensions obtained on one of these subsignals are shown in Fig.~\ref{fig:tho}.
@@
 \subsection{Real-Time Implementation}
-We simulate the real-time implementation of the SST from {\sf BoundEffRed} applied to the PPG described in section~\ref{ssse:ppg}, subsampled by a factor of $2$.
+We simulate the real-time implementation of the SST from {\sf BoundEffRed} applied to the PPG described in Section~\ref{ssse:ppg}, subsampled by a factor of $2$.
 The test is performed on a 2-Core Intel Core i5 CPU running at $1.7$~GHz and $7.7$~GB of RAM.
@@
 For this signal, sampled at $\fs=65.5$~Hz, the suitable window is $8$ seconds long.
-Therefore, the extension must be over $4$ seconds, \ie~$L=250$.
+Therefore, the extension must be over $4$ seconds, \ie, $L=250$.
 The forecasting step in each iteration of {\sf BoundEffRed} then takes no more than $t_{\mathrm{forecast}}=46$~ms.

### 2021-02-22 Paper/theory.tex
polish
final version

@@
 \end{equation}
-where $J$ denotes the number of components, $\Omega_j>0$ the amplitude of the $j$-th component, $f_j$ its frequency, and $\varphi_j\in[0,2\pi)$ its initial phase.
+where $J$ denotes the number of components, $\Omega_j>0$ the amplitude of the $j$th component, $f_j$ its frequency, and $\varphi_j\in[0,2\pi)$ its initial phase.
 %
@@
 This shows the need to use a sufficiently large dataset to obtain an accurate forecast.
-Ideally when $K\to\infty$, the forecasting error would behave like the measurement noise $\sigma\bw$, \ie~a zero-mean white noise whose variance is of the order of $\sigma^2$.
+Ideally when $K\to\infty$, the forecasting error would behave like the measurement noise $\sigma\bw$, \ie, a zero-mean white noise whose variance is of the order of $\sigma^2$.
 However, Theorem~\ref{th:error} shows that the estimator may remain asymptotically biased (the first term in the bound of equation~\eqref{eq:mean.error} being independent of $K$).
@@
 The dependencies of the forecasting quality on the subsignals lengths $M$ and the forecasting index $\ell=n-N+1$ are hidden in the expression of the parameters $a^{(n)}_0$, $a^{(n)}_1$,\dots, $b^{(n,n')}_0$,\dots, $c^{(n,n')}_3$.
-The numerical results, discussed in section~\ref{ssse:res.sine}, allow us to better evaluate these dependencies.
+The numerical results, discussed in Section~\ref{ssse:res.sine}, allow us to better evaluate these dependencies.
 
@@
 
-The models of sections~\ref{sse:model.sine} and~\ref{RemarkAHM} consider the meaningful part of the signal to be a deterministic component.
+The models described in Sections~\ref{sse:model.sine} and~\ref{RemarkAHM} consider the meaningful part of the signal to be a deterministic component.
 These models are purposely adapted to signals showing local line spectra.
-The physiological signals we are interested in, such as respiratory or cardiac signals, have this characteristic (see section~\ref{sse:physio.sig}).
+The physiological signals we are interested in, such as respiratory or cardiac signals, have this characteristic (see Section~\ref{sse:physio.sig}).
 Signals with wider local spectra, such as electroencephalography signals, do not fall into this category, and are more faithfully modeled as random signals.
@@
 While we do not provide a generic proof of the boundary effect reduction on {\em any} TF analysis tools, we discuss a particular case of SST.
-Since SST is designed to analyze signals satisfying the AHM, as is discussed in section~\ref{RemarkAHM}, Theorem~\ref{th:error} ensures that the forecasting error $\bepsilon$ defined in~\eqref{eq:forecasting.error} is controlled and bounded in terms of mean and covariance.
+Since SST is designed to analyze signals satisfying the AHM, as is discussed in Section~\ref{RemarkAHM}, Theorem~\ref{th:error} ensures that the forecasting error $\bepsilon$ defined in~\eqref{eq:forecasting.error} is controlled and bounded in terms of mean and covariance.
 Recall Theorem 3 in~\cite{Chen14nonparametric}, which states that when the additive noise is stationary and may be modulated by a smooth function with a small fluctuation, the SST of the observed signal remains close to the SST of the clean signal, throughout the TF plane.
@@
 We claim that the proof of Theorem 3 in~\cite{Chen14nonparametric} can be generalized to explain the robustness of SST to the noise, and a systematic proof will be reported in our future theoretical work.
-We verify experimentally that Algorithm~\ref{alg:boundary} is efficient for a large number of representations in the following section~\ref{se:results}.
+We verify experimentally that Algorithm~\ref{alg:boundary} is efficient for a large number of representations in the following Section~\ref{se:results}.
 %
